### 2022-10-19 talk/talk.qmd
convert md to qmd

---
format: 
    revealjs:
        scrollable: true
        width: 1440
        height: 1080
        toc: true
        slide-number: true
        theme: [default, custom.scss]

title: "Mainstreaming Scientific Racism"
author: "Dan Hicks & Emilio Lobato"
institute: "UC Merced"
---

# Race science

## Some definitions

scientific racism
 ~ (purportedly) justifying racial inequality and colonialism by appealing to the epistemic authority of science

race science - wide sense
 ~ scientific research that is amenable to scientific racism

race science - narrow sense
 ~ scientific research that itself engages in scientific racism

- Ex: genotype clustering studies [@RosenbergGeneticStructureHuman2002] or polygenic scores for educational attainment [@HardenGeneticAssociationsMathematics2019] 
can be wide-sense race science without being narrow-sense race science [@CarlsonQuantifyingContextualizingImpact2020; WillsAreClustersRaces2017]

- This talk will focus on narrow-sense race science

## Race science in the 20th century: Pre-*Brown* {.smaller}

:::: columns 
::: {.column width="30%"}
![[Charles Davenport (1866-1944)](https://en.wikipedia.org/wiki/Charles_Davenport), circa 1929. Source: <https://en.wikipedia.org/wiki/File:Charles_Benedict_Davenport.jpg>](img/Charles_Benedict_Davenport.jpg){width="100%"}
:::

::: {.column width="30%"}
![[Franz Boas (1858-1942)](https://en.wikipedia.org/wiki/Franz_Boas), circa 1915. Source: <https://en.wikipedia.org/wiki/File:FranzBoas.jpg>](img/FranzBoas.jpg)
:::

::: {.column width="30%"}
![[Margaret Mead (1901-1978)](https://en.wikipedia.org/wiki/Margaret_Mead), 1948. Source: <https://en.wikipedia.org/wiki/File:Margaret_Mead_(1901-1978).jpg>](img/Margaret_Mead_(1901-1978).jpg)
:::
::::

- 1910: Davenport founds Eugenics Records Office at Cold Spring Harbor
- 1926: NRC "Committee on the Negro" includes both Davenport and Boas: "neither was influential enough to veto his rival's participation" [@BarkanRetreatScientificRacism1992 113]
- 1930s: *Coming of Age in Samoa* *[cite]* popularizes cultural anthropology, "free[ing] anthropology from the shackles of biology" [@BarkanRetreatScientificRacism1992 134]
- 1940: Carnegie Institute defunds Eugenics Records Office
- 1950: "For all practical social purposes 'race' is not so much a biological phenomenon as a social myth." *[https://unesdoc.unesco.org/ark:/48223/pf0000128291]*

## Pioneer Fund

*[todo]*

## *Brown v Board* and *Mankind Quarterly*

:::: columns
::: {.column width="30%"}
![[Reginald Ruggles Gates (1882-1962)](https://en.wikipedia.org/wiki/Reginald_Ruggles_Gates), 1921. Source: <https://en.wikipedia.org/wiki/File:Reginald_Ruggles_Gates_Portrait.jpg>](img/Reginald_Ruggles_Gates_Portrait.jpg)
:::

::: {.column width="30%"}
![[Henry Garrett (1894-1973)](https://en.wikipedia.org/wiki/Henry_Garrett_(psychologist)). Source: <https://www.psychometricsociety.org/post/past-present-and-incoming-presidents>, citing American Psychologist (1946), 1(9), 371](img/henry_e_garrett.png)
:::

::: {.column width="30%"}
![[George Robert Gayre of Gayre and Nigg (1907-1996)](https://en.wikipedia.org/wiki/Robert_Gayre). Source: <https://www.patrickcomerford.com/2020/07/robert-gair-racist-from-rathmines-who.html>, original uncredited](img/Robert Gayre profile.jpg)
:::
::::

> Ultimately, anti-'racial' science played a defining role in the Supreme Court’s dismissal of the 'separate-but-equal' principle in education .... [@SchafferScientificRacismAgain2007 261]

> [E]veryone agrees that races or ethnic groups exist and that they are in fact the raw material by means of which human evolution has taken place. (Gates, *MQ* 1:1 [1960], 11)

# Race science as agnotology and echo chamber

## The Tobacco Strategy

Five core strategies used by industry to manufacture doubt and create the appearance of legitimate scientific controversy [after @FernandezPintoKnowBetterNot2017, esp. 59]

1. Emphasize scientific uncertainty
2. Support friendly research
3. Recruit distinguished scientists as spokesperson
4. [Construct echo chambers]{.red}
	- "creation and distribution of non-peer-reviewed journals and pamphlets"
	- "funding and organizing of symposiums" 
	- ["In this way, the tobacco industry circumvented peer review standards in publication"]{.red}
5. Attack unfavorable research

\

[Did Pioneer and *Mankind Quarterly* function in this way for race science?]{.blue}

# Data and methods

## Pioneer-funded researchers

::::: columns
:::: column
::: {.smaller}
| | |
|:-----|:-----|
| [Thomas J. Bouchard, Jr.](https://en.wikipedia.org/wiki/Thomas_J._Bouchard_Jr.) | psychology |
| Brunetto Chiarelli | | anthropology? |
| [Hans Eysenck](https://en.wikipedia.org/wiki/Hans_Eysenck) | psychology |
| [Robert Gordon](https://en.wikipedia.org/wiki/Robert_A._Gordon) | sociology |
| [Linda Gottfredson](https://en.wikipedia.org/wiki/Linda_Gottfredson) | psychology |
| [Garrett Hardin](https://en.wikipedia.org/wiki/Garrett_Hardin) | ecology |
| [Joseph M. Horn](https://en.wikipedia.org/wiki/Joseph_M._Horn) | psychology |
| [Lloyd Humphreys](https://en.wikipedia.org/wiki/Lloyd_Humphreys) | psychology |
| [Arthur Jensen](https://en.wikipedia.org/wiki/Arthur_Jensen) | psychology |
| [Michael Levin](https://en.wikipedia.org/wiki/Michael_Levin) | philosophy |
| [Richard Lynn](https://en.wikipedia.org/wiki/Richard_Lynn) | psychology |
| [R. Travis Osborne](https://en.wikipedia.org/wiki/R._Travis_Osborne) | psychology |
| [J. Phillippe Rushton](https://en.wikipedia.org/wiki/J._Phillippe_Rushton) | psychology |
| [Audrey M. Shuey](https://en.wikipedia.org/wiki/Audrey_M._Shuey) | psychology |
| [Philip A. Vernon](https://en.wikipedia.org/wiki/Philip_A._Vernon) | psychology |
| [Daniel Vining, Jr.](https://en.wikipedia.org/wiki/Daniel_Vining_Jr.) | demography |
:::
::::

:::: column
- @MillerPioneerFundBankrolling1994
- [List of Grantees from PF website](https://web.archive.org/web/20130103005545/http://www.pioneerfund.org/Grantees.html)

\

- WoS author search results for 14 authors (Sept/Oct 2021)
- Identify journals that published more than 5/14
::::
:::::

## Corpus

|                                |       |
|:-------------------------------|------:|
| Behavior Genetics              |  2268 |
| Behavioral & Brain Sciences    |   898 |
| Intelligence                   |  1237 |
| **Mankind Quarterly**          |  1821 |
| Person. & Individ. Differences |  7274 |
| Psychological Reports          | 21398 |

- 34k journal articles
- Published 1960-2010
- 3 journals published by American Psychological Association not available
- PID is missing many relevant book reviews and essays
	- eg, [Rushton's review of Lynn's book on Pioneer](https://www.sciencedirect.com/science/article/pii/S0191886901001337) *[confirm not included]*

<!-- 
\

- Extract article full text from PDF or HTML
- Extract noun phrases using spaCy NLP
 -->

## Topic modeling

![Diachronic evolution of topic categories in *Philosophy of Science*, 1934–2015. @MalaterreWhatThisThing2019 fig. 2.](img/MalaterreWhatThisThing2019.jpeg){width="80%"}

- Statistical method for clustering documents and words simultaneously
- Documents are composed of topics; topics are composed of words
- Custom R package [`tmfast`](https://github.com/dhicks/tmfast)

## Gould's critique of factor analysis

> As a tool for simplification, [PCA/factor analysis] has proved its great value in many disciplines. But many factorists have gone beyond simplification, and tried to define factors as causal entities. This error of reification has plagued the technique since its inception .... [F]actors, by themselves, are neither things nor causes; they are mathematical abstractions. [@GouldMismeasureManRevised1996 284-5]

- `tmfast` uses methods from PCA and factor analysis [@RoheVintageFactorAnalysis2020]
	- (partial) PCA
	- varimax rotation

- Need to avoid hasty reification of topics

## Topic modeling as phenomena construction

Topics as phenomena [@BogenSavingPhenomena1988; @WoodwardDataPhenomenaRestatement2009]

- simplified patterns extracted from word occurrence data
- may or may not be stable:\
  will we see the same topics in a different corpus? [@MalaterreEarlyDaysContemporary2022]
- constructed with minimal role for substantive theory
- **explananda, not explanans**

These words tend to be used together; **but why?**\
These test items tend to be answered in the same way; **but why?**

### 2022-10-19 talk/talk.qmd
incorporate UMAP vis into talk

@@
 
-# Race science
+# Race science in the 20th century
 
+<!--
 ## Some definitions
@@
 - This talk will focus on narrow-sense race science
+-->
 
-## Race science in the 20th century: Pre-*Brown* {.smaller}
+## Before *Brown* {.smaller}
 
@@
 ![[Charles Davenport (1866-1944)](https://en.wikipedia.org/wiki/Charles_Davenport), circa 1929.
-Source: <https://en.wikipedia.org/wiki/File:Charles_Benedict_Davenport.jpg>](img/Charles_Benedict_Davenport.jpg){width="100%"}
+[Source](https://en.wikipedia.org/wiki/File:Charles_Benedict_Davenport.jpg)](img/Charles_Benedict_Davenport.jpg){width="100%"}
 :::
@@
 ::: {.column width="30%"}
-![[Franz Boas (1858-1942)](https://en.wikipedia.org/wiki/Franz_Boas), circa 1915. Source: <https://en.wikipedia.org/wiki/File:FranzBoas.jpg>](img/FranzBoas.jpg)
+![[Franz Boas (1858-1942)](https://en.wikipedia.org/wiki/Franz_Boas), circa 1915. [Source](https://en.wikipedia.org/wiki/File:FranzBoas.jpg)](img/FranzBoas.jpg)
 :::
@@
 ::: {.column width="30%"}
-![[Margaret Mead (1901-1978)](https://en.wikipedia.org/wiki/Margaret_Mead), 1948. Source: <https://en.wikipedia.org/wiki/File:Margaret_Mead_(1901-1978).jpg>](img/Margaret_Mead_(1901-1978).jpg)
+![[Margaret Mead (1901-1978)](https://en.wikipedia.org/wiki/Margaret_Mead), 1948. [Source](https://en.wikipedia.org/wiki/File:Margaret_Mead_(1901-1978).jpg)](img/Margaret_Mead_(1901-1978).jpg)
 :::
@@
 ![[Reginald Ruggles Gates (1882-1962)](https://en.wikipedia.org/wiki/Reginald_Ruggles_Gates), 1921.
-Source: <https://en.wikipedia.org/wiki/File:Reginald_Ruggles_Gates_Portrait.jpg>](img/Reginald_Ruggles_Gates_Portrait.jpg)
+[Source](https://en.wikipedia.org/wiki/File:Reginald_Ruggles_Gates_Portrait.jpg)](img/Reginald_Ruggles_Gates_Portrait.jpg)
 :::
@@
 ![[Henry Garrett (1894-1973)](https://en.wikipedia.org/wiki/Henry_Garrett_(psychologist)).
-Source: <https://www.psychometricsociety.org/post/past-present-and-incoming-presidents>, citing American Psychologist (1946), 1(9), 371](img/henry_e_garrett.png)
+[Source](https://www.psychometricsociety.org/post/past-present-and-incoming-presidents), citing American Psychologist (1946), 1(9), 371](img/henry_e_garrett.png){width="100%"}
 :::
@@
 ![[George Robert Gayre of Gayre and Nigg (1907-1996)](https://en.wikipedia.org/wiki/Robert_Gayre).
-Source: <https://www.patrickcomerford.com/2020/07/robert-gair-racist-from-rathmines-who.html>, original uncredited](img/Robert Gayre profile.jpg)
+[Source](https://www.patrickcomerford.com/2020/07/robert-gair-racist-from-rathmines-who.html), original uncredited](img/Robert Gayre profile.jpg)
 :::
@@
 
-> [E]veryone agrees that races or ethnic groups exist and that they are in fact the raw material by means of which human evolution has taken place. (Gates, *MQ* 1:1 [1960], 11)
+## *MQ* was founded as race science
 
-# Race science as agnotology and echo chamber
+> [[E]veryone agrees that races or ethnic groups exist and that they are in fact the raw material by means of which human evolution has taken place]{.red} ....
+
+> A scientific approach to racial questions, [free from sentiment, propaganda or bias]{.blue}, may lead to a more reasonable attitude to the problems which agitate considerable sections of mankind.
+This need for a calmer approach to problems of race on the part of [civilised peoples as well as more primitive tribes]{.red} makes it desirable that a journal should exist in which the problems of racial origin and racial relationships can be quietly discussed, [without rancour or bigotry and with the primary aim of elucidating facts]{.blue}.
+(Gates, *MQ* 1:1 [1960], 11)
+
+# Race science agntology
 
@@
 # Data and methods
+
+## Did PF and MQ function as a race science echo chamber?
+
+- Empirical, text-mining approach
+- Identify Pioneer-funded researchers
+- Identify journals that published their work
+- Analyze all articles published in those journals, 1960-2010
 
@@
 - Published 1960-2010
+
+\
 - 3 journals published by American Psychological Association not available
@@
 
+<!-- 
 ## Gould's critique of factor analysis
@@
 These test items tend to be answered in the same way; **but why?**
+ -->
+
+# Mainstreaming race science
+
+## Umap interactive
+
+```{r}
+#| out-width: 1440px
+#| out-height: 1080px
+#| fig-cap: "UMAP visualization of Hellinger similarity, medium vocabulary, k = 30"
+library(readr)
+library(here)
+library(plotly)
+
+vis_dir = here('out')
+
+umap_vis = read_rds(here(vis_dir, '07_md_30.Rds'))
+umap_vis
+```

### 2022-10-19 scripts/06_hellinger_betas.R

@@ -11,26 +11,13 @@ data_dir = here('data')
 tm_dir = here(data_dir, '04_tm')
 out_dir = here('out')
 
+source(here('R', 'compare_betas.R'))
+
 model_files = list.files(tm_dir, 'tmfast') |> 
     set_names('lg', 'md', 'sm')
 
 # model = here(tm_dir, '04_md_tmfast.Rds') |>
 #     read_rds()
-
-compare_betas = function(beta1, beta2, vocab) {
-    fill = function(beta) {
-        beta |> 
-            complete(token = vocab, topic, fill = list(beta = 0)) |> 
-            build_matrix(topic, token, beta) %>%
-            .[, vocab]
-    }
-    beta1 = fill(beta1)
-    beta2 = fill(beta2)
-    
-    assert_that(all(colnames(beta1) == colnames(beta2)))
-    
-    hellinger(beta1, beta2)
-}
 
 # compare_betas(betas$`5`, betas$`10`)
 

### 2022-10-19 scripts/07_umap_interactive.R

@@ -71,7 +71,11 @@ umap_plot = weight_df |>
 if (interactive()) umap_plot
 
 p = ggplotly(umap_plot, 
-             tooltip = c('id', 'so', 'yr', 'ti', 'tp', 'tr')) |> 
+             tooltip = c('id', 'so', 'yr', 'ti', 'tp', 'tr'), 
+             width = 1*1440, 
+             height = .8*1080) |> 
+    layout(legend = list(font = list(size = 24)), 
+           hoverlabel = list(font = list(size = 24))) |> 
     style(visible = FALSE, traces = (1:k)+5) |> 
     layout(updatemenus = list(
         list(

### 2022-10-31 talk/talk.qmd
working on slides

@@
         scrollable: true
-        width: 1440
+        width: 1.1*1440
         height: 1080
         toc: true
+        toc-depth: 1
+        toc-title: Overview
         slide-number: true
@@
 author: "Dan Hicks & Emilio Lobato"
-institute: "UC Merced"
+institute: "University of California, Merced"
+bibliography: ../paper/race_science.yaml
 ---
@@
 # Race science in the 20th century
+
+```{r setup}
+library(readr)
+library(here)
+library(plotly)
+library(gt)
+
+vis_dir = here('out')
+```
 
@@
 :::: columns 
-::: {.column width="30%"}
+::: {.column width="33%"}
 ![[Charles Davenport (1866-1944)](https://en.wikipedia.org/wiki/Charles_Davenport), circa 1929.
-[Source](https://en.wikipedia.org/wiki/File:Charles_Benedict_Davenport.jpg)](img/Charles_Benedict_Davenport.jpg){width="100%"}
+[Source](https://en.wikipedia.org/wiki/File:Charles_Benedict_Davenport.jpg)](img/Charles_Benedict_Davenport.jpg){width="70%"}
 :::
 
-::: {.column width="30%"}
-![[Franz Boas (1858-1942)](https://en.wikipedia.org/wiki/Franz_Boas), circa 1915. [Source](https://en.wikipedia.org/wiki/File:FranzBoas.jpg)](img/FranzBoas.jpg)
+::: {.column width="33%"}
+![[Franz Boas (1858-1942)](https://en.wikipedia.org/wiki/Franz_Boas), circa 1915. [Source](https://en.wikipedia.org/wiki/File:FranzBoas.jpg)](img/FranzBoas.jpg){width="70%"}
 :::
 
-::: {.column width="30%"}
-![[Margaret Mead (1901-1978)](https://en.wikipedia.org/wiki/Margaret_Mead), 1948. [Source](https://en.wikipedia.org/wiki/File:Margaret_Mead_(1901-1978).jpg)](img/Margaret_Mead_(1901-1978).jpg)
+::: {.column width="33%"}
+![[Margaret Mead (1901-1978)](https://en.wikipedia.org/wiki/Margaret_Mead), 1948.
+[Source](https://en.wikipedia.org/wiki/File:Margaret_Mead_(1901-1978).jpg)](img/Margaret_Mead_(1901-1978).jpg){width="70%"}
 :::
@@
 - 1926: NRC "Committee on the Negro" includes both Davenport and Boas: "neither was influential enough to veto his rival's participation" [@BarkanRetreatScientificRacism1992 113]
-- 1930s: *Coming of Age in Samoa* *[cite]* popularizes cultural anthropology, "free[ing] anthropology from the shackles of biology" [@BarkanRetreatScientificRacism1992 134]
+- 1930s: *Coming of Age in Samoa* [@MeadComingAgeSamoa1928] popularizes cultural anthropology, "free[ing] anthropology from the shackles of biology" [@BarkanRetreatScientificRacism1992 134]
 - 1940: Carnegie Institute defunds Eugenics Records Office
-- 1950: "For all practical social purposes 'race' is not so much a biological phenomenon as a social myth." *[https://unesdoc.unesco.org/ark:/48223/pf0000128291]*
+- 1950: "For all practical social purposes 'race' is not so much a biological phenomenon as a social myth" [@BeagleholeRaceQuestion1950; though see @BrattainRaceRacismAntiracism2007]
 
-## Pioneer Fund
+## The decline of eugenics and the Pioneer Fund {.smaller}
 
-*[todo]*
+:::: columns
+::: {.column width="30%"}
+![[Wickliffe Draper (1891-1972)](https://en.wikipedia.org/wiki/Wickliffe_Preston_Draper), textile heir, eugenicist, segregationist, 1918.
+[Source](https://en.wikipedia.org/wiki/File:Wickliffe_Draper.jpg)](img/Draper.jpg)
+:::
+
+::: {.column width="30%"}
+![[Harry Laughlin (1880-1943)](https://en.wikipedia.org/wiki/Harry_Laughlin), 
+superintendent of Eugenics Records Office, anti-immigrationist, c. 1929. [Source](https://en.wikipedia.org/wiki/File:Harry_Laughlin.jpg)](img/Harry_Laughlin.jpg){width="100%"}
+:::
+
+::: {.column width="30%"}
+![[Frederick Osborn (1889-1981)](https://en.wikipedia.org/wiki/Frederick_Osborn), railroad heir, eugenicist, shaped postwar pivot to "positive eugenics," undated.
+[Source](https://www.meehanreports.com/eugen-PT-pt1.htm)](img/osbornphoto.jpg){width="100%"}
+:::
+::::
+
+Founded in 1937 to "support academic research and the \'[sic]dissemination of information, into the 'problem of heredity and eugenics' and 'the problems of race betterment'" [@MehlerFoundationFascismNew1989 21, quoting Laughlin]
+<!-- 
+- Presidents
+	- Harry Laughlin, 1937-1941
+	- Frederick Osborn, 1941–1958
+	- Harry Weyher, Jr., 1958-2002
+	- J. Phillippe Rushton, 2002-12
+	- Richard Lynn, 2012-??
+ -->
 
@@
 
-> Ultimately, anti-'racial' science played a defining role in the Supreme Court’s dismissal of the 'separate-but-equal' principle in education .... [@SchafferScientificRacismAgain2007 261]
+> Ultimately, anti-'racial' science played a defining role in the Supreme Court’s dismissal of the 'separate-but-equal' principle in education ....
+[@SchafferScientificRacismAgain2007 261; see also @JacksonScienceSegregationRace2005]
 
@@
 
-> A scientific approach to racial questions, [free from sentiment, propaganda or bias]{.blue}, may lead to a more reasonable attitude to the problems which agitate considerable sections of mankind.
+> [A scientific approach to racial questions, free from sentiment, propaganda or bias]{.blue}, may lead to a more reasonable attitude to the problems which agitate considerable sections of mankind.
 This need for a calmer approach to problems of race on the part of [civilised peoples as well as more primitive tribes]{.red} makes it desirable that a journal should exist in which the problems of racial origin and racial relationships can be quietly discussed, [without rancour or bigotry and with the primary aim of elucidating facts]{.blue}.
@@
 
-# Race science agntology
+# Race science agnotology
 
-## The Tobacco Strategy
+## The Tobacco Strategy {.smallish}
 
@@
 
-[Did Pioneer and *Mankind Quarterly* function in this way for race science?]{.blue}
+[Did Pioneer and *Mankind Quarterly* function in this way for race science?]{.red}
 
-# Data and methods
-
-## Did PF and MQ function as a race science echo chamber?
+<!-- 
+## PF and MQ as a race science echo chamber?
 
@@
 - Analyze all articles published in those journals, 1960-2010
+
+\
+ -->
 
@@
 | [Thomas J. Bouchard, Jr.](https://en.wikipedia.org/wiki/Thomas_J._Bouchard_Jr.) | psychology |
-| Brunetto Chiarelli | | anthropology? |
+| Brunetto Chiarelli | anthropology? |
 | [Hans Eysenck](https://en.wikipedia.org/wiki/Hans_Eysenck) | psychology |
@@
 
-## Corpus
+## Pioneer-funded researchers
+
+![Journals publishing 6 or more Pioneer-funded researchers, WOS](img/wos_results.png)
+
+## Corpus {.smallish}
 
@@
 | **Mankind Quarterly**          |  1821 |
-| Person. & Individ. Differences |  7274 |
+| Personality & Individual<br> Differences |  7274 |
 | Psychological Reports          | 21398 |
@@
 
-\
 - 3 journals published by American Psychological Association not available
+- In B&BS, focal papers + commentaries all bundled together
 - PID is missing many relevant book reviews and essays
-	- eg, [Rushton's review of Lynn's book on Pioneer](https://www.sciencedirect.com/science/article/pii/S0191886901001337) *[confirm not included]*
+	- eg, [Rushton's review of Lynn's book on Pioneer](https://www.sciencedirect.com/science/article/pii/S0191886901001337)
 
@@
 
-![Diachronic evolution of topic categories in *Philosophy of Science*, 1934–2015. @MalaterreWhatThisThing2019 fig. 2.](img/MalaterreWhatThisThing2019.jpeg){width="80%"}
+:::: columns
+::: {.column width="60%"}
+![Diachronic evolution of topic categories in *Philosophy of Science*, 1934–2015. @MalaterreWhatThisThing2019 fig. 2.](img/MalaterreWhatThisThing2019.jpeg){width="100%"}
+:::
 
+::: {.column width="40%"}
 - Statistical method for clustering documents and words simultaneously
-- Documents are composed of topics; topics are composed of words
+- [Can we identify a race science topic?]{.red}
+- [Does this topic start in MQ, spread to mainstream journals?]{.red}
+:::
+::::
+
+::: {.smaller}
 - Custom R package [`tmfast`](https://github.com/dhicks/tmfast)
+- 18 different models (3 vocabulary sizes $\times$ $k$ = 5, 10, 20, ..., 50)
+:::
 
-<!-- 
-## Gould's critique of factor analysis
+## UMAP interactive
+
+```{r}
+#| out-width: 1.1*1440px
+#| out-height: 1080px
+#| fig-cap: "UMAP visualization of Hellinger similarity, medium vocabulary, k = 30"
+
+umap_vis = read_rds(here(vis_dir, '07_md_30.Rds'))
+umap_vis
+```
+
+## 3 focal topics {.smaller}
+
+```{r}
+beta_gt = read_rds(here(vis_dir, '??_beta_table.Rds'))
+beta_gt
+```
+
+---
+
+*[focal topics by journal over time]*
+
+# Conclusion
+
+- Race & intelligence research did not appear to originate in MQ and spread to other journals
+- Instead, Intelligence and PID were the primary "echo chambers" used to legitimize this psychological form of race science
+- "The agnotology is coming from inside the house"
+
+## Different topics or different communities?  {visibility="uncounted"}
+
+![Authors by topic](../out/_authors.png)
+
+## Gould's critique of factor analysis {visibility="uncounted"}
 
@@
 
-## Topic modeling as phenomena construction
+## Topic modeling as phenomena construction {visibility="uncounted"}
 
@@
 These test items tend to be answered in the same way; **but why?**
- -->
 
-# Mainstreaming race science
-
-## Umap interactive
-
-```{r}
-#| out-width: 1440px
-#| out-height: 1080px
-#| fig-cap: "UMAP visualization of Hellinger similarity, medium vocabulary, k = 30"
-library(readr)
-library(here)
-library(plotly)
-
-vis_dir = here('out')
-
-umap_vis = read_rds(here(vis_dir, '07_md_30.Rds'))
-umap_vis
-```
+## References {visibility="uncounted"}

### 2022-10-31 R/author_data.R

@@ -1,7 +1,5 @@
-# library(tidyverse)
-# library(here)
-
-author_data = function(data_dir = here::here('data')) {
+author_data = function(data_dir = here::here('data'), 
+                       canon_file = 'scratch_authors_canonicalized.csv') {
     mainstream_df = arrow::open_dataset(here(data_dir, 
                                       '00_authors_mainstr')) |> 
         select(article_id = doi, given, family) |> 
@@ -9,7 +7,7 @@ author_data = function(data_dir = here::here('data')) {
                author = str_to_title(author)) |> 
         collect()
     
-    #' Flip Lastname, Firstname names
+    # Flip Lastname, Firstname names
     flip = function(string) {
         split = str_match(string, '([^,]+)(, )?(.+)?')
         if_else(is.na(split[,3]) | split[,4] == 'Jr.', 
@@ -26,11 +24,35 @@ author_data = function(data_dir = here::here('data')) {
         mutate(author = flip(authors)) |> 
         select(-authors)
     
+    ## Canonicalized names
+    canonicalized_df = read_csv(here(data_dir, canon_file)) |> 
+        filter(!is.na(canonical))
+    
     comb_df = bind_rows(mq_df, 
-                        mainstream_df)
+                        mainstream_df) |> 
+        select(-given, -family) |> 
+        filter(author != 'Na Na') |> 
+        left_join(canonicalized_df, by = 'author') |> 
+        mutate(author = if_else(!is.na(canonical), 
+                                canonical, 
+                                author)) |> 
+        select(-canonical)
     return(comb_df)
 }
 
-# author_data() |> 
-#     count(author) |> 
-#     arrange(desc(n))
+
+mq = tibble(author = c('Arthur R. Jensen', 
+                       'J. Philippe Rushton', 
+                       'Michael Levin', 
+                       'Robert M. Gordon', 
+                       'Linda S. Gottfredson', 
+                       'Daniel R. Vining, Jr.', 
+                       'Richard Lynn', 
+                       'R. Travis Osborne', 
+                       'Hans J. Eysenck', 
+                       'Thomas J. Bouchard, Jr.', 
+                       'Joseph M. Horn', 
+                       'Philip A. Vernon', 
+                       'Brunetto Chiarelli',
+                       'Lloyd G. Humphreys', 
+                       'Audrey M. Shuey'))

### 2022-11-02 talk/talk.qmd
working on slides, incld Netlify upload

@@
         height: 1080
-        toc: true
+        toc: false
         toc-depth: 1
@@
 author: "Dan Hicks & Emilio Lobato"
-institute: "University of California, Merced"
+institute: "University of California, Merced<br><https://mainstream-race-science.netlify.app>"
 bibliography: ../paper/race_science.yaml
@@
 
-# Race science in the 20th century
+# Introduction
+
+## The return(?) of race science
+
+:::: columns
+::: column
+![](img/Saini.png){width="70%"}\
+@SainiSuperiorReturnRace2019
+:::
+::: column
+> [Many people assumed that the eugenicists had all but disappeared with the Nazi regime, and that race science was pretty much finished at the same time.]{.blue .fragment}
+
+> [Intellectual racism has always existed, and indeed for a chunk of history, it thrived.
+I believe it is still the toxic little seed at the heart of academia.
+However dead you might think it is, it needs only a little water, and now it’s raining.]{.red .fragment}
+:::
+::::
+
+## Institutions of race science
+
+Pioneer Fund
+: Founded 1937, during the decline of (institutionalized) eugenics
+
+*Mankind Quarterly*
+: Founded 1960 by marginalized academics as part of the backlash to *Brown v Board*
+
+. . .
+
+\
+
+The Tobacco Strategy
+: Used by regulated industries to create the appearance of legitimate scientific controversy and thereby block or delay regulation [@OreskesMerchantsDoubtHow2011; @FernandezPintoKnowBetterNot2017]
+
+## Echo chambers
+
+- "creation and distribution of non-peer reviewed journals and pamphlets"
+- "funding and organizing of symposiums" for industry-funded research
+- ["In this way, the tobacco industry circumvented peer review standards in publication"]{.red} [@FernandezPintoKnowBetterNot2017]
+
+. . .
+
+\
+
+- "a bounded, enclosed media space that has the potential to both magnify the messages delivered within it and [insulate them from rebuttal]{.red}" [@JamiesonEchoChamberRush2008 76]
+
+. . .
+
+\
+
+[**Did Pioneer and *Mankind Quarterly* nurture race science by protecting it from mainstream peer review?**]{}
+
+<!-- 
+## PF and MQ as a race science echo chamber?
+
+- Empirical, text-mining approach
+- Identify Pioneer-funded researchers
+- Identify journals that published their work
+- Analyze all articles published in those journals, 1960-2010
+
+\
+ -->
+
+<!-- 
+# Race science, 1910-1960
 
@@
 
-## Before *Brown* {.smaller}
+<!--
+## Rise and fall of eugenics {.smaller}
 
@@
 ![[Charles Davenport (1866-1944)](https://en.wikipedia.org/wiki/Charles_Davenport), circa 1929.
-[Source](https://en.wikipedia.org/wiki/File:Charles_Benedict_Davenport.jpg)](img/Charles_Benedict_Davenport.jpg){width="70%"}
+[Source](https://en.wikipedia.org/wiki/File:Charles_Benedict_Davenport.jpg)](img/Charles_Benedict_Davenport.jpg){width="80%"}
 :::
@@
 ::: {.column width="33%"}
-![[Franz Boas (1858-1942)](https://en.wikipedia.org/wiki/Franz_Boas), circa 1915. [Source](https://en.wikipedia.org/wiki/File:FranzBoas.jpg)](img/FranzBoas.jpg){width="70%"}
+![[Franz Boas (1858-1942)](https://en.wikipedia.org/wiki/Franz_Boas), circa 1915. [Source](https://en.wikipedia.org/wiki/File:FranzBoas.jpg)](img/FranzBoas.jpg){width="80%"}
 :::
@@
 ![[Margaret Mead (1901-1978)](https://en.wikipedia.org/wiki/Margaret_Mead), 1948.
-[Source](https://en.wikipedia.org/wiki/File:Margaret_Mead_(1901-1978).jpg)](img/Margaret_Mead_(1901-1978).jpg){width="70%"}
+[Source](https://en.wikipedia.org/wiki/File:Margaret_Mead_(1901-1978).jpg)](img/Margaret_Mead_(1901-1978).jpg){width="80%"}
 :::
@@
 - 1940: Carnegie Institute defunds Eugenics Records Office
-- 1950: "For all practical social purposes 'race' is not so much a biological phenomenon as a social myth" [@BeagleholeRaceQuestion1950; though see @BrattainRaceRacismAntiracism2007]
+<!-- - 1950: "For all practical social purposes 'race' is not so much a biological phenomenon as a social myth" [@BeagleholeRaceQuestion1950; though see @BrattainRaceRacismAntiracism2007] -->
 
+<!--
 ## The decline of eugenics and the Pioneer Fund {.smaller}
@@
 	- Harry Laughlin, 1937-1941
-	- Frederick Osborn, 1941–1958
+	- Frederick Osborn, 1941-1958
 	- Harry Weyher, Jr., 1958-2002
@@
 
-## *Brown v Board* and *Mankind Quarterly*
+<!--
+## *Brown v Board* (1954) and *Mankind Quarterly*
 
 :::: columns
-::: {.column width="30%"}
+::: {.column width="32%"}
+![[Henry Garrett (1894-1973)](https://en.wikipedia.org/wiki/Henry_Garrett_(psychologist)).
+[Source](https://www.psychometricsociety.org/post/past-present-and-incoming-presidents), citing American Psychologist (1946), 1(9), 371](img/henry_e_garrett.png){width="100%"}
+:::
+
+::: {.column width="32%"}
 ![[Reginald Ruggles Gates (1882-1962)](https://en.wikipedia.org/wiki/Reginald_Ruggles_Gates), 1921.
@@
 
-::: {.column width="30%"}
-![[Henry Garrett (1894-1973)](https://en.wikipedia.org/wiki/Henry_Garrett_(psychologist)).
-[Source](https://www.psychometricsociety.org/post/past-present-and-incoming-presidents), citing American Psychologist (1946), 1(9), 371](img/henry_e_garrett.png){width="100%"}
-:::
-
-::: {.column width="30%"}
+::: {.column width="32%"}
 ![[George Robert Gayre of Gayre and Nigg (1907-1996)](https://en.wikipedia.org/wiki/Robert_Gayre).
@@
 
-# Race science agnotology
+-->
 
-## The Tobacco Strategy {.smallish}
-
-Five core strategies used by industry to manufacture doubt and create the appearance of legitimate scientific controversy [after @FernandezPintoKnowBetterNot2017, esp. 59]
-
-1. Emphasize scientific uncertainty
-2. Support friendly research
-3. Recruit distinguished scientists as spokesperson
-4. [Construct echo chambers]{.red}
-	- "creation and distribution of non-peer-reviewed journals and pamphlets"
-	- "funding and organizing of symposiums" 
-	- ["In this way, the tobacco industry circumvented peer review standards in publication"]{.red}
-5. Attack unfavorable research
-
-\
-
-[Did Pioneer and *Mankind Quarterly* function in this way for race science?]{.red}
-
-<!-- 
-## PF and MQ as a race science echo chamber?
-
-- Empirical, text-mining approach
-- Identify Pioneer-funded researchers
-- Identify journals that published their work
-- Analyze all articles published in those journals, 1960-2010
-
-\
- -->
+# Mainstreaming race science
 
@@
 
-![Journals publishing 6 or more Pioneer-funded researchers, WOS](img/wos_results.png)
+![Journals publishing 6 or more Pioneer-funded researchers, WOS author search results](img/wos_results.png)
 
@@
 ::: {.column width="60%"}
-![Diachronic evolution of topic categories in *Philosophy of Science*, 1934–2015. @MalaterreWhatThisThing2019 fig. 2.](img/MalaterreWhatThisThing2019.jpeg){width="100%"}
+![Diachronic evolution of topic categories in *Philosophy of Science*, 1934-2015. @MalaterreWhatThisThing2019 fig. 2.](img/MalaterreWhatThisThing2019.jpeg){width="100%"}
 :::
@@
 
-- Race & intelligence research did not appear to originate in MQ and spread to other journals
-- Instead, Intelligence and PID were the primary "echo chambers" used to legitimize this psychological form of race science
-- "The agnotology is coming from inside the house"
+- Race & intelligence research did not appear to originate in *MQ* and spread to other journals
+- Instead, *Intelligence* and *PID* were the primary venues used to legitimize this psychological form of race science
+- "The race science is coming from inside the house!"
+
+# Thanks! {visibility="uncounted"}
+
+<https://mainstream-race-science.netlify.app>
+
+# Bonus content {visibility="uncounted"}
 
@@
 
-![Authors by topic](../out/_authors.png)
+![Authors by topic](img/_authors.png)
 

### 2022-11-02 scripts/08_topics_over_time.R
Create 08_topics_over_time.R
created 08_topics_over_time

@@ -1,0 +1,61 @@
+library(tidyverse)
+library(tidytext)
+library(tmfast)
+library(here)
+library(arrow)
+# remotes::install_github("mtennekes/cols4all")
+library(cols4all)
+library(patchwork)
+library(glue)
+
+data_dir = here('data')
+tm_dir = here(data_dir, '04_tm')
+out_dir = here('out')
+
+meta_ar = open_dataset(here(data_dir, '01_metadata'))
+
+phrases_ar = open_dataset(here(data_dir, '00_phrases')) |> 
+    select(-year) |> 
+    mutate(n = log1p(n))
+
+topics_over_time = phrases_ar |>
+    group_by(article_id) |>
+    summarize(total_phrases = sum(n)) |>
+    inner_join(meta_ar, by = 'article_id') |>
+   # filter(container.title != 'Psychological Reports') |> 
+    collect()
+
+sm_tmf = readRDS(here(tm_dir, '04_sm_tmfast.Rds'))
+md_tmf = readRDS(here(tm_dir, '04_md_tmfast.Rds'))
+
+create_dataframes = function(model, k = c(5, seq(10, 50, by = 10)), topics = NULL) {
+    
+    full_topic_list = list()
+    
+    for (i in k){
+        out_prefix = glue('df_{i}')
+        dataframe = tidy(model, k = i, matrix = 'gamma') |>
+            filter(gamma > .05) |>
+            left_join(topics_over_time, by = c('document' = 'article_id')) |> 
+            group_by(topic, container.title, year) |> 
+            summarise(gamma = median(gamma))
+        
+        if (is.null(topics)) {
+            #what to do if topics is null
+            full_topic_list[out_prefix] = list(dataframe)
+            
+        } else {
+            #what to do if topics is not null
+            #this is where filtering happens
+            dataframe = dataframe |> 
+                filter(topic %in% topics)
+            full_topic_list[out_prefix] = list(dataframe)
+            
+        }
+        
+    }
+    return(bind_rows(full_topic_list, .id = 'k'))
+}
+
+test = create_dataframes(sm_tmf, topics = 'V05')
+med_test = create_dataframes(md_tmf, topics = c('V07','V12'))

### 2022-11-02 talk/talk.qmd
full draft of talk

@@
 
-# Introduction
+# Introduction {visibility="uncounted"}
 
@@
 
-# Mainstreaming race science
+# Mainstreaming race science {visibility="uncounted"}
 
@@
 
----
+## Focal topics, by journal, over time
 
-*[focal topics by journal over time]*
+![](img/08_presentation.png)
 
-# Conclusion
+# Conclusions {visibility="hidden"}
+## Conclusions
 
-- Race & intelligence research did not appear to originate in *MQ* and spread to other journals
-- Instead, *Intelligence* and *PID* were the primary venues used to legitimize this psychological form of race science
+- Most Pioneer-funded researchers did not publish heavily in *MQ*
+- Race & intelligence research did not appear to originate in *MQ*
+
+\
+
+- Instead, *Psychological Reports* and later *Intelligence* and *PID* were the primary venues for this psychological form of race science
 - "The race science is coming from inside the house!"

### 2022-11-02 scripts/08_topics_over_time.R

@@ -1,10 +1,12 @@
 library(tidyverse)
 library(tidytext)
+theme_set(theme_bw())
+
 library(tmfast)
 library(here)
 library(arrow)
 # remotes::install_github("mtennekes/cols4all")
-library(cols4all)
+# library(cols4all)
 library(patchwork)
 library(glue)
 
@@ -18,27 +20,32 @@ phrases_ar = open_dataset(here(data_dir, '00_phrases')) |>
     select(-year) |> 
     mutate(n = log1p(n))
 
-topics_over_time = phrases_ar |>
+articles_df = phrases_ar |>
     group_by(article_id) |>
     summarize(total_phrases = sum(n)) |>
     inner_join(meta_ar, by = 'article_id') |>
-   # filter(container.title != 'Psychological Reports') |> 
+    # filter(container.title != 'Psychological Reports') |> 
     collect()
 
 sm_tmf = readRDS(here(tm_dir, '04_sm_tmfast.Rds'))
 md_tmf = readRDS(here(tm_dir, '04_md_tmfast.Rds'))
 
-create_dataframes = function(model, k = c(5, seq(10, 50, by = 10)), topics = NULL) {
-    
+create_dataframes = function(model, 
+                             k = c(5, seq(10, 50, by = 10)), 
+                             topics = NULL, 
+                             articles = articles_df, 
+                             agg_fn = median) {
     full_topic_list = list()
     
     for (i in k){
-        out_prefix = glue('df_{i}')
+        out_prefix = as.character(i)
+        ## Median value of gamma by year-topic-journal
         dataframe = tidy(model, k = i, matrix = 'gamma') |>
             filter(gamma > .05) |>
-            left_join(topics_over_time, by = c('document' = 'article_id')) |> 
+            left_join(articles, by = c('document' = 'article_id')) |> 
             group_by(topic, container.title, year) |> 
-            summarise(gamma = median(gamma))
+            summarise(gamma = agg_fn(gamma), 
+                      .groups = 'drop')
         
         if (is.null(topics)) {
             #what to do if topics is null
@@ -50,12 +57,48 @@ create_dataframes = function(model, k = c(5, seq(10, 50, by = 10)), topics = NUL
             dataframe = dataframe |> 
                 filter(topic %in% topics)
             full_topic_list[out_prefix] = list(dataframe)
-            
         }
-        
     }
+    
     return(bind_rows(full_topic_list, .id = 'k'))
 }
 
[…]

### 2022-11-04 talk/talk.qmd
tweak talk

@@
 Pioneer Fund
-: Founded 1937, during the decline of (institutionalized) eugenics
+: Founded 1937, during the decline of (institutionalized) eugenics\
+  [@MehlerFoundationFascismNew1989; @TuckerFundingScientificRacism2002]
 
 *Mankind Quarterly*
-: Founded 1960 by marginalized academics as part of the backlash to *Brown v Board*
+: Founded 1960 by marginalized academics as part of the backlash to *Brown v Board* [@MehlerFoundationFascismNew1989; @SchafferScientificRacismAgain2007; @JacksonScienceSegregationRace2005 ch. 7]
 
@@
 The Tobacco Strategy
-: Used by regulated industries to create the appearance of legitimate scientific controversy and thereby block or delay regulation [@OreskesMerchantsDoubtHow2011; @FernandezPintoKnowBetterNot2017]
+: Used by regulated industries to create the appearance of legitimate scientific controversy and thereby block or delay regulation\
+  [@OreskesMerchantsDoubtHow2011; @FernandezPintoKnowBetterNot2017]
 
@@
 
-## Corpus {.smallish}
+## Corpus {}
 
@@
 
+::: {.aside .smaller}
 - 3 journals published by American Psychological Association not available
 - In B&BS, focal papers + commentaries all bundled together
-- PID is missing many relevant book reviews and essays
-	- eg, [Rushton's review of Lynn's book on Pioneer](https://www.sciencedirect.com/science/article/pii/S0191886901001337)
-
-<!-- 
-\
-
-- Extract article full text from PDF or HTML
-- Extract noun phrases using spaCy NLP
- -->
+- PID is missing many relevant book reviews and essays, eg, [Rushton's review of Lynn's book on Pioneer](https://www.sciencedirect.com/science/article/pii/S0191886901001337)
+:::
 
@@
 
-::: {.smaller}
+::: aside
 - Custom R package [`tmfast`](https://github.com/dhicks/tmfast)
@@
 # Conclusions {visibility="hidden"}
-## Conclusions
+## Conclusions {.smallish}
 
 - Most Pioneer-funded researchers did not publish heavily in *MQ*
-- Race & intelligence research did not appear to originate in *MQ*
+- *MQ* was the primary venue for one of two race science topics
+- The other race science topic did not originate in *MQ*
+
+. . .
 
@@
 
-- Instead, *Psychological Reports* and later *Intelligence* and *PID* were the primary venues for this psychological form of race science
+- *MQ* was the primary venue for one form of race science
+	- Anthropological? 
+- While mainstream journals were the primary venues for a distinct form of race science
+	- *Psychological Reports*, later *Intelligence* and *PID*
+	- Psychological? 
 - "The race science is coming from inside the house!"

### 2022-11-05 talk/talk.qmd
tweak thresholds and counting in topics over time

@@
 
-. . .
+::: {.incremental}
+- "a bounded, enclosed media space that has the potential to both magnify the messages delivered within it and [insulate them from rebuttal]{.red}" [@JamiesonEchoChamberRush2008 76]
+:::
 
-\
-
-- "a bounded, enclosed media space that has the potential to both magnify the messages delivered within it and [insulate them from rebuttal]{.red}" [@JamiesonEchoChamberRush2008 76]
-
-. . .
-
-\
-
-[**Did Pioneer and *Mankind Quarterly* nurture race science by protecting it from mainstream peer review?**]{}
+[**Did Pioneer and *Mankind Quarterly* nurture race science by protecting it from mainstream peer review?**]{.fragment}\
+[**Our analysis indicates *MQ* did not play this role**]{.fragment .red}
 
@@
 
-## Corpus {}
+## Text-mining corpus
 
@@
 - Statistical method for clustering documents and words simultaneously
-- [Can we identify a race science topic?]{.red}
-- [Does this topic start in MQ, spread to mainstream journals?]{.red}
+
+\
+
+- [Can we identify a race science topic?]{.blue}
+- [Does this topic start in MQ, spread to mainstream journals?]{.blue}
 :::
@@
 - *MQ* was the primary venue for one form of race science
-	- Anthropological? 
+	- "Anthropology-flavored"? 
 - While mainstream journals were the primary venues for a distinct form of race science
 	- *Psychological Reports*, later *Intelligence* and *PID*
-	- Psychological? 
+	- "Psychology-flavored"? 
 - "The race science is coming from inside the house!"

### 2022-11-05 scripts/08_topics_over_time.R

@@ -30,18 +30,26 @@ articles_df = phrases_ar |>
 sm_tmf = readRDS(here(tm_dir, '04_sm_tmfast.Rds'))
 md_tmf = readRDS(here(tm_dir, '04_md_tmfast.Rds'))
 
+md_tmf |> 
+    tidy(k = 30, matrix = 'gamma') |> 
+    filter(topic %in% c('V05', 'V07', 'V19')) |> 
+    ggplot(aes(gamma, color = topic)) +
+    stat_ecdf() +
+    geom_hline(yintercept = .9)
+
 create_dataframes = function(model, 
                              k = c(5, seq(10, 50, by = 10)), 
                              topics = NULL, 
                              articles = articles_df, 
-                             agg_fn = median) {
+                             gamma_treshhold = .05,
+                             agg_fn = mean) {
     full_topic_list = list()
     
     for (i in k){
         out_prefix = as.character(i)
         ## Median value of gamma by year-topic-journal
         dataframe = tidy(model, k = i, matrix = 'gamma') |>
-            filter(gamma > .05) |>
+            filter(gamma > gamma_treshhold) |>
             left_join(articles, by = c('document' = 'article_id')) |> 
             group_by(topic, container.title, year) |> 
             summarise(gamma = agg_fn(gamma), 
@@ -76,15 +84,16 @@ ggplot(topics_time_md, aes(year, gamma,
 
 topics_time_md_count = create_dataframes(md_tmf, 
                                          topics = c('V07', 'V19', 'V05'), 
-                                         # agg_fn = \(x)(n())
-                                         agg_fn = sum)
+                                         agg_fn = \(x)(n()), 
+                                         # agg_fn = sum,
+                                         gamma_treshhold = .07)
 topics_time_md_count |> 
     filter(k == 30) |> 
     group_by(k, topic, container.title) |> 
     mutate(gamma_sm = slider::slide_index_dbl(gamma, year, 
                                               mean, 
-                                              .before = 4, 
-                                              .after = 4)) |> 
+                                              .before = 2, 
+                                              .after = 2)) |> 
     ungroup() |> 
     ggplot(aes(year,  
                color = container.title, 
@@ -92,13 +101,14 @@ topics_time_md_count |>
[…]

### 2022-11-08 talk/talk.qmd
race science def'n slide

@@
 # Introduction {visibility="uncounted"}
+
+```{r setup}
+library(readr)
+library(here)
+library(plotly)
+library(gt)
+
+vis_dir = here('out')
+```
 
@@
 <!-- 
-## PF and MQ as a race science echo chamber?
-
-- Empirical, text-mining approach
-- Identify Pioneer-funded researchers
-- Identify journals that published their work
-- Analyze all articles published in those journals, 1960-2010
-
-\
- -->
-
-<!-- 
 # Race science, 1910-1960
-
-```{r setup}
-library(readr)
-library(here)
-library(plotly)
-library(gt)
-
-vis_dir = here('out')
-```
-
-<!--
-## Some definitions
-
-scientific racism
- ~ (purportedly) justifying racial inequality and colonialism by appealing to the epistemic authority of science
-
-race science - wide sense
- ~ scientific research that is amenable to scientific racism
-
-race science - narrow sense
- ~ scientific research that itself engages in scientific racism
-
-- Ex: genotype clustering studies [@RosenbergGeneticStructureHuman2002] or polygenic scores for educational attainment [@HardenGeneticAssociationsMathematics2019] 
-can be wide-sense race science without being narrow-sense race science [@CarlsonQuantifyingContextualizingImpact2020; WillsAreClustersRaces2017]
-
-- This talk will focus on narrow-sense race science
--->
 
@@
 
+## Some definitions {visibility="uncounted"}
+
+scientific racism
+ ~ (purportedly) justifying racial inequality and colonialism by appealing to the epistemic authority of science\
+   [compare @LewontinBourgeoisIdeologyOrigins1982]
+
+race science - wide sense
+ ~ scientific research that is amenable to scientific racism
+
+race science - narrow sense
+ ~ scientific research that itself engages in scientific racism
+
+\
+
+Ex: genotype clustering studies [@RosenbergGeneticStructureHuman2002] or polygenic scores for educational attainment [@HardenGeneticAssociationsMathematics2019] 
+can be wide-sense race science without being narrow-sense race science [@WillsAreClustersRaces2017; @CarlsonQuantifyingContextualizingImpact2020]
+
 ## Different topics or different communities?  {visibility="uncounted"}

### 2022-11-08 R/author_data.R
update author overlap diagram

@@ -26,7 +26,8 @@ author_data = function(data_dir = here::here('data'),
     
     ## Canonicalized names
     canonicalized_df = read_csv(here(data_dir, canon_file)) |> 
-        filter(!is.na(canonical))
+        filter(!is.na(canonical)) |> 
+        select(-...3)
     
     comb_df = bind_rows(mq_df, 
                         mainstream_df) |> 

### 2023-03-20 paper/paper.Rmd
rticles template for paper

---
title: "Mainstreaming Race Science"
authors:
  - name: "Daniel J. Hicks"
    affiliation: 1
    corresponding_author: true
  - name: "Emilio Lobato"
    affiliation: 1
address:
  - code: 1
    address: UC Merced
corresponding_author: "To whom correspondence should be addressed; E-mail: dhicks4@ucmerced.edu."
abstract: |
  Enter the text of your abstract here.
acknowledgements: |
  Please include your acknowledgments here, set in a single paragraph.
bibliography: scibib.bib
output: 
  rticles::science_article:
      move_figures: false
---

<!-- 
Science submission instructions

General information: Research Articles
<https://www.science.org/content/page/science-information-authors>
- 5 printed pages in the journal
- 2000 to 3000 words of main text, in addition to an abstract, 3 to 5 display items (figures or tables) with brief legends, about 50 main-text references, and a structured acknowledgments section
- main text should be divided into sections with brief subheadings
- Materials and Methods should be included in supplementary materials and should be followed by additional data and figures needed to support the paper’s conclusions
- All data must be available in either the main text or the supplementary materials, or must be deposited at a publicly accessible repository and cited in the paper.

Initial submission
<https://www.science.org/content/page/instructions-preparing-initial-manuscript>
- Title: 96 characters
- Authors
- Affiliations:
- One Sentence Summary: 125 characters
- Abstract: Structured as BACKGROUND (1 sentence), OBJECTIVES/METHODS, RESULTS, CONCLUSION (1 sentence); 125 words or less
- Main Text:  The introduction should provide sufficient background information to make the article intelligible to readers in other disciplines and sufficient context so that the significance of the experimental findings is clear. Technical terms should be defined.
- References and Notes: There should be a single reference list that combines references cited in the main paper as well as those cited in the supplementary materials, and each reference should appear only once in the list. 
- Acknowledgements: Funding; Authors contributions (see link); Competing interests;  Data and materials availability
List of Supplementary Materials
- The width of figures, when printed, will usually be 5.7 cm (2.24 inches or 1 column), 12.1 cm (4.76 inches or 2 columns), or 18.4 cm (7.24 inches or 3 columns).

-->

# Introduction

This is a basic implementation of the Science template.

# Uniform Distributions

Uniform distributions text with Figure.

```{r once, echo=FALSE, fig.cap='test once', fig.align='center', out.width='50%'}
plot(runif(10))
```

And post uniform distributions text...

\begin{table}[ht]
\begin{center}
\begin{tabular}{ll}
once & 1 \\
once & 2\\
\end{tabular}
\end{center}
\end{table}

With a table.

# Poisson Distributions

Poisson distributions text...

```{r twice, echo=FALSE, fig.cap='test twice'}
plot(rpois(10,1))
```

And post Poisson distributions text...

\begin{table}[ht]
\begin{center}
\begin{tabular}{ll}
twice & 1 \\
twice & 2\\
\end{tabular}
\end{center}
\end{table}

With a table for something.

# Formatting Citations

\begin{enumerate}
\item Please cite your references and notes in text {\it only\/} using
  the standard \LaTeX\ \verb+\cite+ command, not another command
  driven by outside macros.
\item Please separate multiple citations within a single \verb+\cite+
  command using commas only; there should be {\it no space\/}
  between reference keynames.  That is, if you are citing two
  papers whose bibliography keys are \texttt{keyname1} and
  \texttt{keyname2}, the in-text cite should read
  \verb+\cite{keyname1,keyname2}+, {\it not\/}
  \verb+\cite{keyname1, keyname2}+.
\end{enumerate}

\noindent Failure to follow these guidelines could lead
to the omission of the references in an accepted paper when the source
file is translated to Word via HTML.

# Handling Math, Tables, and Figures

Following are a few things to keep in mind in coding equations,
tables, and figures for submission to {\it Science}.

\paragraph*{In-line math.}  The utility that we use for converting
from \LaTeX\ to HTML handles in-line math relatively well.  It is best
to avoid using built-up fractions in in-line equations, and going for
the more boring ``slash'' presentation whenever possible --- that is,
for \verb+$a/b$+ (which comes out as $a/b$) rather than
\verb+$\frac{a}{b}$+ (which compiles as $\frac{a}{b}$).  Likewise,
HTML isn't tooled to handle certain overaccented special characters
in-line; for $\hat{\alpha}$ (coded \verb+$\hat{\alpha}$+), for
example, the HTML translation code will return [\^{}$(\alpha)$].
Don't drive yourself crazy --- but if it's possible to avoid such
constructs, please do so.  Please do not code arrays or matrices as
in-line math; display them instead.  And please keep your coding as
\TeX-y as possible --- avoid using specialized math macro packages
like \texttt{amstex.sty}.

\paragraph*{Displayed math.} Our HTML converter sets up \TeX\
displayed equations using nested HTML tables.  That works well for an
HTML presentation, but Word chokes when it comes across a nested
table in an HTML file.  We surmount that problem by simply cutting the
displayed equations out of the HTML before it's imported into Word,
and then replacing them in the Word document using either images or
equations generated by a Word equation editor.  Strictly speaking,
this procedure doesn't bear on how you should prepare your manuscript
--- although, for reasons best consigned to a note \cite{nattex}, we'd
prefer that you use native \TeX\ commands within displayed-math
environments, rather than \LaTeX\ sub-environments.

\paragraph*{Tables.}  The HTML converter that we use seems to handle
reasonably well simple tables generated using the \LaTeX\
\texttt{\{tabular\}} environment.  For very complicated tables, you
may want to consider generating them in a word processing program and
including them as a separate file.

\paragraph*{Figures.}  Figure callouts within the text should not be
in the form of \LaTeX\ references, but should simply be typed in ---
that is, \verb+(Fig. 1)+ rather than \verb+\ref{fig1}+.  For the
figures themselves, treatment can differ depending on whether the
manuscript is an initial submission or a final revision for acceptance
and publication.  For an initial submission and review copy, you can
use the \LaTeX\ \verb+{figure}+ environment and the
\verb+\includegraphics+ command to include your PostScript figures at
the end of the compiled PostScript file.  For the final revision,
however, the \verb+{figure}+ environment should {\it not\/} be used;
instead, the figure captions themselves should be typed in as regular
text at the end of the source file (an example is included here), and
the figures should be uploaded separately according to the Art
Department's instructions.

# What to Send In

What you should send to {\it Science\/} will depend on the stage your manuscript is in:

\begin{itemize}
\item {\bf Important:} If you're sending in the initial submission of
  your manuscript (that is, the copy for evaluation and peer review),
  please send in {\it only\/} a PostScript or PDF version of the
  compiled file (including figures).  Please do not send in the \TeX\ 
  source, \texttt{.sty}, \texttt{.bbl}, or other associated files with
  your initial submission.  (For more information, please see the
  instructions at our Web submission site,
  http://www.submit2science.org/ .)
\item When the time comes for you to send in your revised final
  manuscript (i.e., after peer review), we require that you include
  all source files and generated files in your upload.  Thus, if the
  name of your main source document is \texttt{ltxfile.tex}, you
  need to include:
\begin{itemize}
\item \texttt{ltxfile.tex}.
\item \texttt{ltxfile.aux}, the auxilliary file generated by the
  compilation.
\item A PostScript file (compiled using \texttt{dvips} or some other
  driver) of the \texttt{.dvi} file generated from
  \texttt{ltxfile.tex}, or a PDF file distilled from that
  PostScript.  You do not need to include the actual \texttt{.dvi}
  file in your upload.
\item From B{\small{IB}}\TeX\ users, your bibliography (\texttt{.bib})
  file, {\it and\/} the generated file \texttt{ltxfile.bbl} created
  when you run B{\small{IB}}\TeX.
\item Any additional \texttt{.sty} and \texttt{.bst} files called by
  the source code (though, for reasons noted earlier, we {\it
    strongly\/} discourage the use of such files beyond those
  mentioned in this document).
\end{itemize}
\end{itemize}

# Appendix
\appendix
Appendix figures and tables will also need to be moved. A basic example is provided here so this feature is also tested.

First we have a figure.

```{r thrice, echo=FALSE, fig.cap='test thrice'}
plot(rnorm(10))
```

And next we have a table

\begin{table}[ht]
\begin{center}
\begin{tabular}{ll}
thrice & 1 \\
thrice & 2\\
\end{tabular}
\end{center}
\end{table}

### 2023-04-20 R/args.R
complete makefile (through 09)

@@ -1,0 +1,8 @@
+args = commandArgs(trailingOnly = TRUE)
+if ('--force' %in% args) {
+    force = TRUE
+} else {
+    force = FALSE
+}
+# message(glue::glue('force is {force}'))
+cli::cli_alert_warning('force is {force}')

### 2023-04-20 R/author_data.R

@@ -1,5 +1,5 @@
 author_data = function(data_dir = here::here('data'), 
-                       canon_file = 'scratch_authors_canonicalized.csv') {
+                       canon_file = '00_authors_canonicalized.csv') {
     mainstream_df = arrow::open_dataset(here(data_dir, 
                                       '00_authors_mainstr')) |> 
         select(article_id = doi, given, family) |> 

### 2023-04-20 R/compare_betas.R

@@ -1,0 +1,18 @@
+compare_betas = function(beta1, beta2 = NULL, vocab) {
+    fill = function(beta) {
+        beta |> 
+            complete(token = vocab, topic, fill = list(beta = 0)) |> 
+            build_matrix(topic, token, beta) %>%
+            .[, vocab]
+    }
+    beta1 = fill(beta1)
+    if (is.null(beta2)) {
+        beta2 = beta1
+    } else {
+        beta2 = fill(beta2)
+    }
+    
+    assertthat::assert_that(all(colnames(beta1) == colnames(beta2)))
+    
+    hellinger(beta1, beta2)
+}

### 2023-04-20 R/phrases.R

@@ -1,0 +1,6 @@
+phrases = function() {
+    open_dataset(here(data_dir, '00_phrases'))  |> 
+        select(-year) |> 
+        filter(!is.na(phrase))
+}
+

### 2023-04-20 scripts/01_combine_datasets.R

@@ -1,3 +1,4 @@
+renv::load(here::here())
 library(dplyr)
 library(arrow)
 
@@ -18,7 +19,8 @@ write_csv = FALSE  ## Write phrases out to a single CSV?
 # phrases_comb = open_dataset(c(phrases_mq, phrases_ms))
 # write_dataset(phrases_comb, here(data_dir, '01_phrases'))
 
-phrases_comb = open_dataset(here(data_dir, '00_phrases'))
+source(here('R', 'phrases.R'))
+phrases_comb = phrases()
 
 if (write_csv) {
     phrases_comb |> 

### 2023-04-20 scripts/02_describe_corpus.R

@@ -1,4 +1,5 @@
-library(tidyverse)
+renv::load(here::here())
+ library(tidyverse)
 theme_set(theme_minimal())
 library(arrow)
 
@@ -7,8 +8,8 @@ library(here)
 data_dir = here('data')
 
 ## Load metadata ----
-meta_ar = open_dataset(here(data_dir, '01_metadata')) |> 
-    collect()
+meta_ar = open_dataset(here(data_dir, '01_metadata'))
+    # collect()
 
 nrow(meta_ar)
 
@@ -20,9 +21,10 @@ nrow(meta_ar)
 # phrases_ar = open_dataset(here(data_dir, '01_phrases'))
 ## Approach 3:  Manually combined parquet folders
 ## This seems to be the most stable option
-#phrases_ar = open_dataset(here(data_dir, '00_phrases'))
+source(here('R', 'phrases.R'))
+phrases_ar = phrases()
 
-phrases_ar = read_csv(here(data_dir, '01_phrases.csv'))
+# phrases_ar = read_csv(here(data_dir, '01_phrases.csv'))
 
 ## Descriptive plots ----
 meta_ar |>
@@ -79,25 +81,33 @@ ggplot(combined,
     scale_y_log10()
 
 #created a joined df of everything rather than just total phrases by journal-year
-full = inner_join(meta_ar, phrases_ar, 
-                  by = c('article_id', 'year')) |> 
+full = inner_join(phrases_ar, meta_ar, 
+                  by = c('article_id'), 
+                  multiple = 'all') |> 
     collect()
 
 #identifying missing values
-which(is.na(full$article_id)) #none
-which(is.na(full$title)) #lots 437,092
-which(is.na(full$volume)) #lots 927,591
-which(is.na(full$issue)) #none
-which(is.na(full$container.title)) #none
-which(is.na(full$year)) #none
-which(is.na(full$phrase)) #lots 18,144
-which(is.na(full$n)) #none
+n_distinct(full$article_id)
+sum(is.na(full$article_id)) #none
+sum(is.na(full$title)) #lots 441,751
+sum(is.na(full$volume)) #lots 928,591
+sum(is.na(full$issue)) #1,375
+sum(is.na(full$container.title)) #none
+sum(is.na(full$n)) #none
 
-#check which article_id exist only in meta_ar or phrases_ar
-meta_ar[!meta_ar$article_id %in% phrases_ar$article_id,] #two article_id unique to meta_ar; pdfs for these articles are corrupt
[…]

### 2023-04-20 scripts/03_vocab.R

@@ -1,4 +1,5 @@
 ## This script uses length-proportional conditional entropy to select the analysis vocabulary
+renv::load(here::here())
 library(tidyverse)
 theme_set(theme_bw())
 library(tmfast)
@@ -18,8 +19,8 @@ vocab_ratio = c(1/5, 1, 5)
 ## Load data ----
 message('Loading data')
 meta_ar = open_dataset(here(data_dir, '01_metadata'))
-phrases_ar = open_dataset(here(data_dir, '00_phrases')) |> 
-    select(-year)
+source(here('R', 'phrases.R'))
+phrases_ar = phrases()
 
 ## 6.3M distinct phrases
 ## ~16 sec

### 2023-04-20 scripts/04_topic_models.R

@@ -1,5 +1,5 @@
 ## Fit topic models using tmfast
-## TODO: abstract process of loading vocab and fitting, to facilitate fitting multiple models (vocab size x k)
+renv::load(here::here())
 library(tidyverse)
 theme_set(theme_bw())
 # remotes::install_github('dhicks/tmfast')
@@ -13,6 +13,11 @@ library(glue)
 data_dir = here('data')
 
 k = c(5, seq(10, 50, by = 10))
+## Models are not re-fit automatically.  To refit: 
+## 1. Pass --force through the command line 
+source(here('R', 'args.R'))
+## 2. Set `force` manually here
+force = TRUE
 
 ## Load data ----
 meta_ar = open_dataset(here(data_dir, '01_metadata'))
@@ -22,8 +27,8 @@ article_ids = meta_ar |>
     filter(container.title != 'Psychological Reports') |> 
     pull(article_id)
 
-phrases_ar = open_dataset(here(data_dir, '00_phrases')) |> 
-    select(-year) |> 
+source(here('R', 'phrases.R'))
+phrases_ar = phrases() |> 
     mutate(n = log1p(n))
 
 vocabs = list.files(here(data_dir), '03-vocab') %>%
@@ -32,7 +37,11 @@ vocabs = list.files(here(data_dir), '03-vocab') %>%
     map(read_rds)
 
 ## Loop over vocabs ----
-process_vocab = function(this_vocab, vocab_name, force = FALSE) {
+process_vocab = function(this_vocab, 
+                         vocab_name, 
+                         force = FALSE, 
+                         verbose = FALSE,
+                         ...) {
     message(glue('Vocabulary {vocab_name} ----------------------------------------------'))
     file_prefix = str_c('04_', vocab_name)
     make_path = function(...) {
@@ -47,10 +56,12 @@ process_vocab = function(this_vocab, vocab_name, force = FALSE) {
     
     ## What's dropped with this vocabulary? 
     dropped_file = make_path('dropped.csv')
-    if (file.exists(dropped_file) || force) {
+    if (file.exists(dropped_file) && !force) {
         dropped_df = read_csv(dropped_file, show_col_types = FALSE)
     } else {
-        dropped_df = anti_join(meta_ar, phrases_df, by = 'article_id') |> 
+        dropped_df = anti_join(meta_ar, 
+                               phrases_df, 
+                               by = 'article_id') |> 
             count(container.title) |> 
             collect()
         write_csv(dropped_df, make_path('dropped.csv'))
[…]

### 2023-04-20 scripts/05_all_gamma.R

@@ -1,9 +1,11 @@
+renv::load(here::here())
 library(tidyverse)
 theme_set(theme_minimal())
 library(tmfast)
 library(magrittr)
 library(furrr)
 plan(multisession, workers = availableCores() - 2)
+warning('This script uses multisession parallel processing.')
 
 library(arrow)
 library(here)
@@ -21,6 +23,7 @@ model_files = list.files(tm_dir, 'tmfast') %>%
     set_names(c('lg', 'md', 'sm'))
 
 all_gammas = function(path, name) {
+    message(glue('Processing topic model {name}'))
     out_prefix = glue('05_{name}')
     model = read_rds(path)
     
@@ -47,6 +50,7 @@ all_gammas = function(path, name) {
            width = 4, height = 3, scale = 1.5, bg = 'white')
     
     ## Renormalization
+    message('Calculating renormalization exponents')
     get_power = function(this_k, model) {
         ee = expected_entropy(peak_alpha(this_k, 1, peak = .8, scale = 1))
         model |> 
@@ -54,16 +58,19 @@ all_gammas = function(path, name) {
             target_power(document, gamma, ee)
     }
     all_k = model$n
-    exponents = future_map_dbl(all_k, get_power, model)
+    exponents = future_map_dbl(all_k, get_power, model, 
+                               .progress = TRUE)
     tibble(k = all_k, 
            exponent = exponents) |> 
         write_rds(here(data_dir, glue(out_prefix, '_exponents.Rds')))
     
+    message('Applying renormalization exponents')
     gamma = tidy_all(model, matrix = 'gamma') |> 
         group_split(k) |> 
         map2_dfr(exponents, ~ renorm(.x, document, gamma, .y))
 
     ## Combine w/ metadata to order by year
+    message('Joining metadata with topic-doc distributions')
     comb_df = right_join(meta_df, gamma, 
                          by = c('article_id' = 'document'), 
                          multiple = 'all') |> 
@@ -72,6 +79,7 @@ all_gammas = function(path, name) {
                                         .fun = min))
     
     ## topic-doc entropies
+    message('Calculating topic-doc entropies')
     comb_df |> 
         group_by(k, container.title, article_id) |> 
         summarize(H = sum(-gamma * log2(gamma + 1e-15))) |> 
@@ -89,6 +97,7 @@ all_gammas = function(path, name) {
            width = 6, height = 3, scale = 1.5, bg = 'white')
     
[…]

### 2023-04-20 scripts/06_hellinger_betas.R

@@ -1,3 +1,4 @@
+renv::load(here::here())
 library(tidyverse)
 theme_set(theme_minimal())
 library(tmfast)
@@ -8,6 +9,9 @@ library(assertthat)
 library(glue)
 library(furrr)
 plan(multisession, workers = availableCores() - 2)
+
+library(patchwork)
+library(tidytext)
 
 data_dir = here('data')
 tm_dir = here(data_dir, '04_tm')
@@ -20,6 +24,7 @@ model_files = list.files(tm_dir, 'tmfast') %>%
     set_names('lg', 'md', 'sm')
 
 hellinger_betas = function(model_file, name) {
+    message(glue('{name} vocabulary'))
     out_prefix = glue('06_{name}')
     model = read_rds(model_file)
     
@@ -31,18 +36,73 @@ hellinger_betas = function(model_file, name) {
             target_power(topic, beta, ee)
     }
     all_k = model$n
-    exponents = future_map_dbl(all_k, get_power, model)
+    exponents = future_map_dbl(all_k, get_power, model, 
+                               .progress = TRUE)
     # return(exponents)
     
     betas = tidy_all(model) |> 
         group_split(k) |> 
         map2(exponents, ~ renorm(.x, topic, beta, .y))
     
+    ## Silge plots
+    betas |> 
+        list_rbind() |> 
+        group_by(k, topic) |> 
+        top_n(10, beta) |> 
+        ungroup() |> 
+        arrange(k, topic, desc(beta)) |> 
+        mutate(token = reorder_within(token, 
+                                      beta,
+                                      list(k, topic))) |> 
+        group_by(k) |> 
+        group_map(~ {ggplot(.x, 
+                            aes(token, beta)) +
+                geom_point() +
+                geom_linerange(aes(ymax = beta), ymin = 0) +
+                scale_x_reordered() +
+                scale_y_continuous(
+                    breaks = scales::pretty_breaks(n = 3)) +
+                coord_flip(ylim = c(0, NA)) +
+                facet_wrap(vars(topic),
+                           scales = 'free',
+                           ncol = 5) +
+                theme(plot.background = 
+                          element_rect(colour = "black", 
+                                       fill = NA, 
+                                       linewidth = 1))}, 
+                .keep = TRUE) |> 
+        wrap_plots(design = 
+                       'AB
+               #B
+               CD
+               CD
+               CD
+               CD
+               #D
+               #D
+               EF
+               EF
+               EF
[…]

### 2023-04-20 scripts/07_umap_interactive.R

@@ -1,22 +1,29 @@
+renv::load(here::here())
 library(tidyverse)
 theme_set(theme_minimal())
 library(tmfast)
 library(plotly)
 library(writexl)
+library(htmlwidgets)
 
 library(arrow)
 library(here)
 library(glue)
 
-vocab = 'sm'
-k = 30
-force = FALSE  ## Force re-calculating UMAP projection
-outliers = .01 ## Used to calculate window for visualization
+outliers = .01    ## Used to calculate window for visualization
+seed = 2023-04-06 ## RNG seed used by UMAP
+
+## UMAP embeddings are not re-fit automatically.  To refit: 
+## 1. Pass --force through the command line 
+source(here('R', 'args.R'))
+## 2. Set `force` manually here
+# force = TRUE
 
 ## Load data ----
 out_dir = here('out')
 data_dir = here('data')
 tm_dir = here('data', '04_tm')
+prefix = '07'
 
 source(here('R', 'author_data.R'))
 meta_df = open_dataset(here('data', '01_metadata')) |> 
@@ -25,137 +32,177 @@ meta_df = open_dataset(here('data', '01_metadata')) |>
     collect() |>
     nest(authors = author)
 
-phrases_ar = open_dataset(here(data_dir, '01_phrases.csv'), 
-                          format = 'csv')
-model = here(tm_dir, glue('04_{vocab}_tmfast.Rds')) |> 
-    read_rds()
-# umap = here(tm_dir, glue('04_{vocab}_{k}_umap.Rds')) |> 
-#     read_rds()
+source(here('R', 'phrases.R'))
+phrases_ar = phrases()
 
-exponent = here(data_dir, glue('05_{vocab}_exponents.Rds')) |> 
-    read_rds() |> 
-    pull(exponent, name = k) |> 
-    magrittr::extract(as.character(k))
-
-
-gamma = tidy(model, k, matrix = 'gamma', exponent = exponent) |> 
-    filter(document %in% meta_df$article_id)
-beta  = tidy(model, k, matrix = 'beta')
-
-umap_file = here(data_dir, glue('07_{vocab}_{k}_umap.Rds'))
-if (!file.exists(umap_file) || force) {
-    message(glue('Calculating UMAP projection for {vocab}-{k}'))
-    umap_df = hellinger(gamma, prob1 = gamma) |> 
-        umap(verbose = TRUE)
-    write_rds(umap_df, umap_file)
-} else {
[…]

### 2023-04-20 scripts/08_topics_over_time.R

@@ -1,3 +1,4 @@
+renv::load(here::here())
 library(tidyverse)
 library(tidytext)
 theme_set(theme_bw())
@@ -16,8 +17,8 @@ out_dir = here('out')
 
 meta_ar = open_dataset(here(data_dir, '01_metadata'))
 
-phrases_ar = open_dataset(here(data_dir, '00_phrases')) |> 
-    select(-year) |> 
+source(here('R', 'phrases.R'))
+phrases_ar = phrases() |> 
     mutate(n = log1p(n))
 
 articles_df = phrases_ar |>

### 2023-04-20 scripts/09_umap_static.R

@@ -1,0 +1,141 @@
+renv::load(here::here())
+library(tidyverse)
+theme_set(theme_minimal())
+library(geomtextpath)
+library(tmfast)
+
+library(arrow)
+library(here)
+library(glue)
+
+out_dir = here('out')
+data_dir = here('data')
+tm_dir = here('data', '04_tm')
+
+prefix = '09'
+
+outliers = .01 ## Used to calculate window for visualization
+
+
+source(here('R', 'author_data.R'))
+meta_df = open_dataset(here('data', '01_metadata')) |> 
+    left_join(author_data(), by = 'article_id') |> 
+    filter(container.title != 'Psychological Reports') |> 
+    collect() |>
+    nest(authors = author)
+
+## 2d KDE, wrapped in a pipe-friendly way
+kde2d_pipe = function(dataf, var1, var2, ...) {
+    MASS::kde2d(dataf[[var1]], 
+                dataf[[var2]], 
+                ...)
+}
+
+## Calculate 2d KDE and return a tidy dataframe
+calc_densities = function(dataf, var1, var2, group_var, group_values, ...) {
+    var1 = ensym(var1)
+    var2 = ensym(var2)
+    group_var = ensym(group_var)
+    dens = dataf |> 
+        filter(!is.na({{ var1 }}), 
+               !is.na({{ var2 }}), 
+               {{ group_var }} %in% group_values) |> 
+        mutate({{ group_var }} := fct_drop({{ group_var }})) %>%
+        split(.[[rlang::as_string(group_var)]]) |> 
+        map(kde2d_pipe, 
+            rlang::as_string(var1), 
+            rlang::as_string(var2), 
+            ...)
+    
+    tidy_density = function(d) {
+        d$z |> 
+            as_tibble(.name_repair = 'unique') |> 
+            mutate(!!var1 := d$x) |> 
+            pivot_longer(-{{ var1 }}, 
+                         names_to = rlang::as_string(var2), 
+                         values_to = 'z') |> 
+            mutate(!!var2 := rep_along({{ var2 }}, d$y))
+    }
+    dens |> 
+        map(tidy_density) |> 
+        list_rbind(names_to = rlang::as_string(group_var))
+}
+
+## Static UMAP visualization with highlighted topics
+visualize = function(vocab, k, highlight_topics, write = TRUE) {
+    model = here(tm_dir, glue('04_{vocab}_tmfast.Rds')) |> 
[…]

### 2023-04-20 paper/dataset.qmd
paper files

# Data and Methods

## Corpus Assembly

Corpus assembly is summarized in figure @fig-corpus.  The corpus was assembled in two parts, one for *Mankind Quarterly* and the other for mainstream journals.

```{mermaid}
%%| fig-cap: Dataset construction and analysis overview
%%| fig-label: fig-corpus
graph TD

subgraph Mankind Quarterly Corpus
MQ("Mankind Quarterly<br>
    All articles, 1960-2004")
unz(Unz Review) --> MQ
pq(ProQuest) --> MQ
end

subgraph Pioneer Fund Corpus
au("Pioneer-funded authors<br>16") --> wos("Web of Science<br>
                                         author records")
wos --> journals
journals("Journals publishing<br> 
          6+ PF authors")
journals --> apa("APA journals<br>
                  not available")
journals --> generalist("Science & Nature<br>
                         excluded")
journals --> PF("Pioneer Fund Corpus<br>
                 5 journals<br>
                 All articles, 1960-2010")
end

subgraph Publishable Data and Code

MQ & PF --> corpus["Combined corpus<br>
                     (34k documents)"]
corpus --> pos("Noun phrase extraction")
pos --> vocab("Vocabulary selection<br>
                   (large, medium, small)")
vocab --> tm("Topic modeling<br>
               (k = 5, 10, 20, 30, 40, 50)")
tm --> space("Discursive space<br>analysis")
end
```

### Mankind Quarterly

An electronic archive of *Mankind Quarterly*, from the first issue through 2004, is available for free on the open web from *The Unz Review*, which has been characterized as a white nationalist website by the Southern Poverty Law Center *[<https://www.splcenter.org/hatewatch/2021/01/19/meet-white-nationalist-organizer-who-spewed-hate-against-lawmakers>]*.  Web scraping packages and techniques *[which?]* were used to retrieve every PDF available in this archive.  These PDFs were processed using the open-source OCR (optical character recognition) software `tesseract` and the R API `rtesseract` *[cites]*.  This part of corpus assembly was conducted in Fall 2021.

### Mainstream journals

To identify suitable mainstream journals for inclusion in the corpus, we first identified academic researchers who had been funded by the Pioneer Fund.  Pioneer is an American non-profit organization founded in 1937 to fund eugenics research.  After the *Brown v. Board of Education* ruling in 1954, Pioneer funded various segregationist efforts across the United States.  *[Shockley and Jensen]*.  We reviewed a critical profile of Pioneer *[cite]* as well as an archived page from the organization's own web site *[cite]*, which together listed 16 researchers who had received Pioneer funds.

We then used the author search tool in Clarivate's Web of Science platform, retrieving publication lists for 14 researchers.  These searches were conducted between 2021-09-24 and 2021-10-05 by DJH.  After parsing these results, we counted how many of the 14 researchers had published in each journal.  13 journals had published 6 or more of the 14 Pioneer-funded researchers.  We excluded *Mankind Quarterly* (already covered) as well as *Science* and *Nature* (as too general) from further consideration in this side of the corpus.  3 journals published by the American Psychological Association (*American Psychologist*, *Contemporary Psychology*, and *Journal of Educational Psychology*) had to be excluded due to confusion over could give us permission to use the archives for a text mining project, with both APA and ProQuest asserting that we needed to get permission from the other.  *European Journal of Personality*, published by SAGE, also had to be excluded because our institutional access only went back to 1999.

The remaining 5 journals are all published by large academic publishers — Elsevier, Springer, or *[Cambridge UP]* — and each item in the entire run of each journal has been assigned a DOI (digital object identifier) for archival purposes.  We used the Crossref API and `rcrossref` R interface to this API to retrieve metadata for each item published from 1960-2010 in each of these 6 journals.  These metadata included individual license information — confirming that the text of each item could be used for text mining projects — and a URL to an electronic version of the item.  These URLs were used to retrieve an HTML or PDF version of each item, except for *Personality and Individual Differences*.  This journal has published a relatively large number of non-article documents, such as book reviews and commentaries, which are not available at the URL included in the Crossref metadata.  (This is unfortunate, as it was not difficult to find highly relevant documents that we could not automatically retrieve and therefore excluded from our corpus *[review]*.)  Instead we used Elsevier's ScienceDirect API to search and retrieve all items from *Personality and Individual Differences* that are available.  PDFs were run through OCR and text extraction as necessary, as with the *Mankind Quarterly* documents.

All together, *[corpus summary, total and by journal; table]*

## Data preparation

After document retrieval and text extraction, we pre-processed the text using the `spaCy` NLP (natural language processing) Python library and the R API `spacyr` *[version, cites]*.  Specifically, we applied regular expressions to remove header/footer copyright notices and hyphenation, used spaCy to annotate and extract noun phrases (eg, "the intelligence test items"), and then cleaned and standardized these phrases (eg, removing the/an/a, coverting all text to lowercase, and replacing all whitespace with underscores: "intelligence_test_items").  We then counted the occurrence of each noun phrase in the document.  The aggregated "document-term matrix" was stored in Parquet format *[cite]* for performance reasons, and written and read using the `arrow` package for R *[cite]*.  *[availability]*

NLP-extracted noun phrases offer a number of advantages over the more traditional unigram ("single-word") tokens.  First, noun phrase extraction removes many standard stopwords (articles, common verbs) without relying on an a priori stopword list.  Phrases can be more informative than single terms, for example, distinguishing "intelligence test" from "hypothesis test."  While simple n-gram extraction will include phrases that are not especially meaningful.  Consider the sentence "Since our first analyses of feeding patterns in rats, we had been using a criterion of 40 minutes (Le Magnen & Tallon 1963; 1966)" *[10.1017__^__s0140525x0000042x]*.  Bigrams such as "since our" and "criterion 40" will likely be discarded in vocabulary selection, but add to the computational cost of vocabulary selection.  Noun phrase extraction is therefore more efficient.

After noun phrase extraction, the corpus comprised *[tokens]* of *[distinct noun phrases]*.

### Vocabulary selection

We took an information-theoretic approach to vocabulary selection.  Consider a game in which I draw a document from the corpus, then a single token from that document.  I tell you the term, and you have to guess which document I picked.  Intuitively, highly informative terms (in this project, noun phrase types) are distinctive, allowing you to dramatically narrow down the list of potential documents.  This "informativeness" of a term can be quantified as the KL (Kullback-Leibler) divergence from a "baseline" distribution of documents to the distribution conditional on the term.  Because the most informative terms tend to be typos and OCR errors — these are unique to a single document — we multiply the KL divergence by the logarithm of the total number of occurrences of the term across the entire corpus.  We refer to the resulting measure as $log(n) \Delta H$ or simply `ndH`.

*[Hicks 2021?]* used the `ndH` approach in a topic modeling analysis, where the baseline distribution was the uniform distribution across documents.  In the current project, we found that this approach heavily favored recurrent noun phrases in the longest documents.  Many of the documents published in *Psychological Reports* are extremely short, 1-2 reports of a single study; while many of the documents published in *Brain and Behavioral Sciences* are book-length collections that include a long review article and sometimes dozens of commentaries.  More generic noun phrases in the latter can occur orders of magnitude more often than highly distinctive phrases in the former, and so the $log(n)$ factor overwhelms the $\Delta H$ factor.

To address this, we used a different baseline distribution of documents.  Namely, we used a distribution in which document probability is proportional to length.  This makes phrases from short documents much more surprising, and hence dramatically increases their informativeness.  This was more effective at identifying useful phrases from across the corpus.

A common rule of thumb in topic modeling is that the vocabulary should have about 10 times as many distinct terms as the number of documents in the corpus *[cite]*.  However, we had some concerns with computational demands here:  the resulting document-term matrix would have roughly $33,000 \times 330,000$ or 10.9 billion entries; with 10% density this would require on the order of 4 GB of memory; while most of the analysis was to be conducted on the authors' laptops.  We therefore chose to work with three smaller vocabularies, $5 \times$, $1 \times$ and $\frac{1/5} \times$ the number of documents.  We refer to these as the "large," "medium," and "small" vocabulary, respectively, and they include *[counts]* distinct phrases.  We compare findings across vocabularies as a robustness check.

## Topic modeling

To fit topic models, we followed the approach proposed by *[cite]*, which uses varimax-rotated partial principal components instead of the variational inference methods used by standard topic model packages *[eg, stm]*.  This novel approach was implemented in the R package `tmfast`, available separately at *[url]*.  A simulation study of `tmfast` found that it was significantly faster and slightly less accurate at reconstructing known word-topic and topic-document distributions, compared to `stm`.

Topic models were fit for all three vocabularies (large, medium, and small) with $k = 5, 10, 20, 30, 40, 50$ (number of topics), resulting in a total of 18 models.

Following the approach of *[Hicks 2021?]*, we did not attempt to identify a unique best model for further analysis.  While we focus on the medium vocabulary, $k=30$ model as the "median" among the 18 models fit, we compare and contrast findings from this model with those from the other models.

### Topic model interpretation

After fitting topic models, our first research question was whether we could identify distinctive "race science" topics.  We addressed this question in two ways.  Following common practice in topic modeling, we extracted lists of the top $n$ (highest-probability) terms (noun phrases) from each topic in each model, with $n$ generally ranging between 5 and 15 as we explored the models.  *[tables + Silge plots]*

We also adapted the "discursive space" analysis proposed by *[Hicks 2021?]*.  This analysis combines the Hellinger distance between topic-document distributions — capturing how "far apart" two documents are — with dimensional reduction techniques used for data visualization.  This produces a two-dimensional visualization of individual documents, arranged so that documents with similar topic distributions are visually clustered and visually distinct from documents with different topic distributions.

Specifically, for each fitted topic model, we extracted topic-document distributions for each document and calculated pairwise Hellinger distances.  We then processed these distances using the UMAP algorithm for dimension reduction, as implemented in the `umap` package *[details; cite]*.  To reduce computational requirements for these steps, we first dropped documents published in the journal *[Psychological Reports]* (*[x documents removed, y remaining]*; note that the full corpus was used to fit the topic models, *Psychological Reports* was removed just for this visualization).  Using the `plotly` library, we construct an interactive visualization:  each point represents a single document, and points can be colored based on either their journal or highest-probability topic.  Tooltips are displayed when we hover over a point, including document metadata (title, journal, authors, year) as well as "key phrases."

To calculate key phrases for a given document $d$, we first identify the highest-probability topic $t$ for that document and the word-topic (noun phrase-topic) distribution $\beta_{w,t}$ for that topic.  We also retrieve the document-term matrix, containing the count $n_{w,d}$ of occurrences for each word (noun phrase), and calculate a weighting $n_{w,d} \beta_{w,t}$ for each word.  The key phrases are the 5 most highly-weighted terms for the given document.

### 2023-06-09 paper/paper.Rmd
building paper infrastructure

@@
   - code: 1
-    address: UC Merced
+    address: "University of California, Merced"
 corresponding_author: "To whom correspondence should be addressed; E-mail: dhicks4@ucmerced.edu."
@@
   Please include your acknowledgments here, set in a single paragraph.
-bibliography: scibib.bib
+bibliography: "references.bib"
 output: 
+  # https://github.com/rstudio/rticles/pull/486
   rticles::science_article:
+      keep_tex: true
       move_figures: false
+      draft: false
+      number_sections: true
 ---
@@
 
-This is a basic implementation of the Science template.
+foo \cite{JacksonScienceSegregationRace2005}
 
-# Uniform Distributions
-
-Uniform distributions text with Figure.
-
-```{r once, echo=FALSE, fig.cap='test once', fig.align='center', out.width='50%'}
-plot(runif(10))
-```
-
-And post uniform distributions text...
-
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{ll}
-once & 1 \\
-once & 2\\
-\end{tabular}
-\end{center}
-\end{table}
-
-With a table.
-
-# Poisson Distributions
-
-Poisson distributions text...
-
-```{r twice, echo=FALSE, fig.cap='test twice'}
-plot(rpois(10,1))
-```
-
-And post Poisson distributions text...
-
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{ll}
-twice & 1 \\
-twice & 2\\
-\end{tabular}
-\end{center}
-\end{table}
-
-With a table for something.
-
-# Formatting Citations
-
-\begin{enumerate}
-\item Please cite your references and notes in text {\it only\/} using
-  the standard \LaTeX\ \verb+\cite+ command, not another command
-  driven by outside macros.
-\item Please separate multiple citations within a single \verb+\cite+
-  command using commas only; there should be {\it no space\/}
-  between reference keynames.  That is, if you are citing two
-  papers whose bibliography keys are \texttt{keyname1} and
-  \texttt{keyname2}, the in-text cite should read
-  \verb+\cite{keyname1,keyname2}+, {\it not\/}
-  \verb+\cite{keyname1, keyname2}+.
-\end{enumerate}
-
-\noindent Failure to follow these guidelines could lead
-to the omission of the references in an accepted paper when the source
-file is translated to Word via HTML.
-
-# Handling Math, Tables, and Figures
-
-Following are a few things to keep in mind in coding equations,
-tables, and figures for submission to {\it Science}.
-
-\paragraph*{In-line math.}  The utility that we use for converting
-from \LaTeX\ to HTML handles in-line math relatively well.  It is best
-to avoid using built-up fractions in in-line equations, and going for
-the more boring ``slash'' presentation whenever possible --- that is,
-for \verb+$a/b$+ (which comes out as $a/b$) rather than
-\verb+$\frac{a}{b}$+ (which compiles as $\frac{a}{b}$).  Likewise,
-HTML isn't tooled to handle certain overaccented special characters
-in-line; for $\hat{\alpha}$ (coded \verb+$\hat{\alpha}$+), for
-example, the HTML translation code will return [\^{}$(\alpha)$].
-Don't drive yourself crazy --- but if it's possible to avoid such
-constructs, please do so.  Please do not code arrays or matrices as
-in-line math; display them instead.  And please keep your coding as
-\TeX-y as possible --- avoid using specialized math macro packages
-like \texttt{amstex.sty}.
-
-\paragraph*{Displayed math.} Our HTML converter sets up \TeX\
-displayed equations using nested HTML tables.  That works well for an
-HTML presentation, but Word chokes when it comes across a nested
-table in an HTML file.  We surmount that problem by simply cutting the
-displayed equations out of the HTML before it's imported into Word,
-and then replacing them in the Word document using either images or
-equations generated by a Word equation editor.  Strictly speaking,
-this procedure doesn't bear on how you should prepare your manuscript
---- although, for reasons best consigned to a note \cite{nattex}, we'd
-prefer that you use native \TeX\ commands within displayed-math
-environments, rather than \LaTeX\ sub-environments.
-
-\paragraph*{Tables.}  The HTML converter that we use seems to handle
-reasonably well simple tables generated using the \LaTeX\
-\texttt{\{tabular\}} environment.  For very complicated tables, you
-may want to consider generating them in a word processing program and
-including them as a separate file.
-
-\paragraph*{Figures.}  Figure callouts within the text should not be
-in the form of \LaTeX\ references, but should simply be typed in ---
-that is, \verb+(Fig. 1)+ rather than \verb+\ref{fig1}+.  For the
-figures themselves, treatment can differ depending on whether the
-manuscript is an initial submission or a final revision for acceptance
-and publication.  For an initial submission and review copy, you can
-use the \LaTeX\ \verb+{figure}+ environment and the
-\verb+\includegraphics+ command to include your PostScript figures at
-the end of the compiled PostScript file.  For the final revision,
-however, the \verb+{figure}+ environment should {\it not\/} be used;
-instead, the figure captions themselves should be typed in as regular
-text at the end of the source file (an example is included here), and
-the figures should be uploaded separately according to the Art
-Department's instructions.
-
-# What to Send In
-
-What you should send to {\it Science\/} will depend on the stage your manuscript is in:
-
-\begin{itemize}
-\item {\bf Important:} If you're sending in the initial submission of
-  your manuscript (that is, the copy for evaluation and peer review),
-  please send in {\it only\/} a PostScript or PDF version of the
-  compiled file (including figures).  Please do not send in the \TeX\ 
-  source, \texttt{.sty}, \texttt{.bbl}, or other associated files with
-  your initial submission.  (For more information, please see the
-  instructions at our Web submission site,
-  http://www.submit2science.org/ .)
-\item When the time comes for you to send in your revised final
-  manuscript (i.e., after peer review), we require that you include
-  all source files and generated files in your upload.  Thus, if the
-  name of your main source document is \texttt{ltxfile.tex}, you
-  need to include:
-\begin{itemize}
-\item \texttt{ltxfile.tex}.
-\item \texttt{ltxfile.aux}, the auxilliary file generated by the
-  compilation.
-\item A PostScript file (compiled using \texttt{dvips} or some other
-  driver) of the \texttt{.dvi} file generated from
-  \texttt{ltxfile.tex}, or a PDF file distilled from that
-  PostScript.  You do not need to include the actual \texttt{.dvi}
-  file in your upload.
-\item From B{\small{IB}}\TeX\ users, your bibliography (\texttt{.bib})
-  file, {\it and\/} the generated file \texttt{ltxfile.bbl} created
-  when you run B{\small{IB}}\TeX.
-\item Any additional \texttt{.sty} and \texttt{.bst} files called by
-  the source code (though, for reasons noted earlier, we {\it
-    strongly\/} discourage the use of such files beyond those
-  mentioned in this document).
-\end{itemize}
-\end{itemize}
-
-# Appendix
 \appendix
+# Materials and methods
 Appendix figures and tables will also need to be moved. A basic example is provided here so this feature is also tested.

### 2023-06-09 paper/paper.tex
building paper infrastructure

% Use only LaTeX2e, calling the article.cls class and 12-point type.
\documentclass[12pt]{article}

% Users of the {thebibliography} environment or BibTeX should use the
% scicite.sty package, downloadable from *Science* at
% www.sciencemag.org/about/authors/prep/TeX_help/ .
% This package should properly format in-text
% reference calls and reference-list numbers.

% Science template required packages
\usepackage{scicite}

\usepackage{times}

% add-ons for rticles
% based on arxiv.sty
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[hidelinks]{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}
\usepackage{graphicx}

% the basics
\usepackage{amsmath}
\usepackage{calc}
\usepackage{tabularx}

% for figure adjustment
\usepackage{placeins}
\usepackage{flafter}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Pandoc citation processing

% The following parameters seem to provide a reasonable page setup.

\topmargin 0.0cm
\oddsidemargin 0.2cm
\textwidth 16cm
\textheight 21cm
\footskip 1.0cm

%The next command sets up an environment for the abstract to your paper.
\usepackage{setspace}

\newenvironment{sciabstract}{%
\begin{quote} \singlespacing}
{\end{quote}}

% If your reference list includes text notes as well as references,
% include the following line; otherwise, comment it out.

\renewcommand\refname{References and Notes}

% The following lines set up an environment for the last note in the
% reference list, which commonly includes acknowledgments of funding,
% help, etc.  It's intended for users of BibTeX or the {thebibliography}
% environment.  Users who are hand-coding their references at the end
% using a list environment such as {enumerate} can simply add another
% item at the end, and it will be numbered automatically.

\newcounter{lastnote}
\newenvironment{scilastnote}{%
\setcounter{lastnote}{\value{enumiv}}%
\addtocounter{lastnote}{+1}%
\begin{list}%
{\arabic{lastnote}.}
{\setlength{\leftmargin}{.22in}}
{\setlength{\labelsep}{.5em}}}
{\end{list}}

% Include your paper's title here

\title{\bf Mainstreaming Race Science}

% Place the author information here.  Please hand-code the contact
% information and notecalls; do *not* use \footnote commands.  Let the
% author contact information appear immediately below the author names
% as shown.  We would also prefer that you don't change the type-size
% settings shown here.

\author{
Daniel J. Hicks,\textsuperscript{1}\textsuperscript{*}
and Emilio Lobato\textsuperscript{1}
\\
\\
\normalsize{\textsuperscript{1}University of California, Merced}\\
\\
\textsuperscript{*}To whom correspondence should be addressed; E-mail:
\href{mailto:dhicks4@ucmerced.edu}{\nolinkurl{dhicks4@ucmerced.edu}}.
}

% Include the date command, but leave its argument blank.

\date{}

%%%%%%%%%%%%%%%%% END OF PREAMBLE %%%%%%%%%%%%%%%%

\begin{document}
% Double-space the manuscript.

\baselineskip24pt

% Make the title.

\maketitle

% Place your abstract within the special {sciabstract} environment.

\begin{sciabstract}
Enter the text of your abstract here.
\end{sciabstract}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

foo \cite{JacksonScienceSegregationRace2005}
\bibliography{references.bib}
\bibliographystyle{Science}

% Following is a new environment, {scilastnote}, that's defined in the
% preamble and that allows authors to add a reference at the end of the
% list that's not signaled in the text; such references are used in
% *Science* for acknowledgments of funding, help, etc.

\begin{scilastnote}
\item Please include your acknowledgments here, set in a single
paragraph.
\end{scilastnote}

\renewcommand{\thetable}{\arabic{table}}
\renewcommand{\thefigure}{\arabic{figure}}
\setcounter{table}{0}
\setcounter{figure}{0}

%%%begfigs---

%%%endfigs---

%%%begtabs---

%%%endtabs---

% Needs a bit of work on the \ref{} TODO  before uncommenting
%\renewcommand{\thetable}{A\arabic{table}}
%\renewcommand{\thefigure}{A\arabic{figure}}
%\setcounter{table}{0}
%\setcounter{figure}{0}

%%%begappxfigs---

%%%endappxfigs---

%%%begappxtabs---

%%%endappxtabs---

\end{document}

### 2023-06-09 paper/supplement_paper.tex
building paper infrastructure

--- 2023-06-09 paper/paper.tex
@@
 \author{
-Daniel J. Hicks,\textsuperscript{1}\textsuperscript{*}
-and Emilio Lobato\textsuperscript{1}
-\\
-\\
-\normalsize{\textsuperscript{1}University of California, Merced}\\
-\\
-\textsuperscript{*}To whom correspondence should be addressed; E-mail:
+Daniel J. Hicks,\textsuperscript{*}
+and Emilio Lobato
+\\\textsuperscript{*}To whom correspondence should be addressed; E-mail:
 \href{mailto:dhicks4@ucmerced.edu}{\nolinkurl{dhicks4@ucmerced.edu}}.
@@
 \begin{document}
-% Double-space the manuscript.
+\maketitle
+\section*{Supplementary Text}
+\renewcommand{\thesection}{S\arabic{section}}
+\hypertarget{materials-and-methods}{%
+\section{Materials and methods}\label{materials-and-methods}}
 
-\baselineskip24pt
+Appendix figures and tables will also need to be moved. A basic example
+is provided here so this feature is also tested.
 
-% Make the title.
+First we have a figure.
 
-\maketitle
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/thrice-1.pdf}
+\caption{test thrice}
+\end{figure}
 
-% Place your abstract within the special {sciabstract} environment.
+And next we have a table
 
-\begin{sciabstract}
-Enter the text of your abstract here.
-\end{sciabstract}
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-foo \cite{JacksonScienceSegregationRace2005}
-\bibliography{references.bib}
-\bibliographystyle{Science}
-
-% Following is a new environment, {scilastnote}, that's defined in the
-% preamble and that allows authors to add a reference at the end of the
-% list that's not signaled in the text; such references are used in
-% *Science* for acknowledgments of funding, help, etc.
-
-\begin{scilastnote}
-\item Please include your acknowledgments here, set in a single
-paragraph.
-\end{scilastnote}
-
-\renewcommand{\thetable}{\arabic{table}}
-\renewcommand{\thefigure}{\arabic{figure}}
-\setcounter{table}{0}
-\setcounter{figure}{0}
-
-%%%begfigs---
-
-%%%endfigs---
-
-%%%begtabs---
-
-%%%endtabs---
-
-% Needs a bit of work on the \ref{} TODO  before uncommenting
-%\renewcommand{\thetable}{A\arabic{table}}
-%\renewcommand{\thefigure}{A\arabic{figure}}
-%\setcounter{table}{0}
-%\setcounter{figure}{0}
-
-%%%begappxfigs---
-
-%%%endappxfigs---
-
-%%%begappxtabs---
-
-%%%endappxtabs---
+\begin{table}[ht]
+\begin{center}
+\begin{tabular}{ll}
+thrice & 1 \\
+thrice & 2\\
+\end{tabular}
+\end{center}
+\end{table}
 

### 2023-06-09 scripts/scratch_interrater_reliability.R

@@ -1,13 +1,34 @@
+library(tidyverse)
 library(readxl)
 library(irr)
 library(here)
 
 data_dir = here('data')
 
-data = read_xlsx(here(data_dir, '07_md_30_V19_coded.xlsx'))
+data_el = read_xlsx(here(data_dir, '07_md_30_V19 -EL codes.xlsx')) |> 
+    filter(!is.na(race_discourse)) |> 
+    rename(race_discourse_el = race_discourse, 
+           notes_el = notes) |> 
+    mutate(race_discourse_el = race_discourse_el == 1L)
+data_djh = read_xlsx(here(data_dir, '07_md_30_v19-DJH.xlsx')) |> 
+    filter(!is.na(race_discourse)) |> 
+    rename(race_discourse_djh = race_discourse) |> 
+    mutate(race_discourse_djh = race_discourse_djh == 'y')
 
-kappa2(data[,c(9,10)], "unweighted")
+data = inner_join(data_el, data_djh)
 
-data2 = subset(data, DJH == 0 & EL == 0)
+## documents coded
+nrow(data)
+## quick crosstab
+## "false positive" rate is 15%; 25% non-consensus
+count(data, race_discourse_djh, race_discourse_el) |> 
+    mutate(share = n / sum(n))
 
-write.csv(data2, here(data_dir, "07_md_30_V19_false_positives.csv"), row.names = FALSE)
+
+kappa2(data[,c('race_discourse_djh','race_discourse_el')], 
+       "unweighted")
+
+data2 = subset(data, !race_discourse_djh & !race_discourse_el)
+
+write.csv(data2, here(data_dir, "07_md_30_V19_false_positives.csv"), 
+          row.names = FALSE)

### 2023-06-09 paper/paper.Rmd
readme for paper

@@
   # https://github.com/rstudio/rticles/pull/486
+  # skeleton: <https://github.com/christopherkenny/rticles/tree/master/inst/rmarkdown/templates/science>
   rticles::science_article:
@@
       draft: false
-      number_sections: true
+      number_sections: false
 ---

### 2023-06-12 paper/paper.Rmd
working on main text

@@
     address: "University of California, Merced"
-corresponding_author: "To whom correspondence should be addressed; E-mail: dhicks4@ucmerced.edu."
+corresponding_author: "Corresponding author: Daniel J. Hicks, dhicks4@ucmerced.edu."
 abstract: |
-  Enter the text of your abstract here.
+  *[structured]*
 acknowledgements: |
-  Please include your acknowledgments here, set in a single paragraph.
+  *[strutured]* Thanks to Emily Merchant and John Jackson for some initial discussions that helped clarify the scope of the project and identify some key resources related to the Pioneer Fund and *Mankind Quarterly*.
+Thanks to Anthony Sainez for help retrieving the *Mankind Quarterly* articles.
+Thanks to Derek Devnich and James Dooley for their work attempting to secure the APA-published articles.
+EL's work on this project was supported by UC Merced.
+DH's work on this project was not supported by any specific funding.
 bibliography: "references.bib"
@@
 
-foo \cite{JacksonScienceSegregationRace2005}
+*[hey you should read this paper]*
+
+*[incorporate def'ns]*
+scientific racism 
+ ~ purporting to justify racial inequality and colonialism by appealing to the epistemic authority of science
+
+race science
+ ~ (pseudo-)scientific research that can be utilized for scientific racism
+
+race science discourse
+ ~ treating race science as a legitimate area of scientific research; this includes methodological critiques of race science and empirical tests that falsify race science hypotheses
+
+# Race science 1910-1960
+
+## The rise and fall of eugenics
+
+- scientific racism is thoroughly entangled with eugenics, but still conceptually distinct [@SlobodianUnequalMindHow2023]
+    - sterilization laws focused on categories of criminality and mental ability, rather than race per se
+    - contemporary defenders of eugenics at least sometimes claim to oppose racism 
+    - claims of innate racial differences in intelligence can be made without dysgenic prophecies of a rapidly-expanding biological underclass
+
+- 1910: Davenport founds Eugenics Records Office at Cold Spring Harbor
+- 1926: NRC "Committee on the Negro" includes both Davenport and Boas: "neither was influential enough to veto his rival's participation" [@BarkanRetreatScientificRacism1992 113]
+- 1930s: *Coming of Age in Samoa* [@MeadComingAgeSamoa1928] popularizes cultural anthropology, "free[ing] anthropology from the shackles of biology" [@BarkanRetreatScientificRacism1992 134]
+- 1937: Pioneer Fund is founded to "support academic research and the \'[sic]dissemination of information, into the 'problem of heredity and eugenics' and 'the problems of race betterment'" [@MehlerFoundationFascismNew1989 21, quoting Laughlin]
+- 1940: Carnegie Institute defunds Eugenics Records Office
+- 1950: "For all practical social purposes 'race' is not so much a biological phenomenon as a social myth" [@BeagleholeRaceQuestion1950; though see @BrattainRaceRacismAntiracism2007]
+
+## *Brown v Board* and *Mankind Quarterly*
+
+Histories of scientific racism in the second half of the twentieth century have often emphasized the parascholarly journal *Mankind Quarterly* (MQ) \cite{
+MehlerFoundationFascismNew1989,
+WinstonScienceServiceFar1998,
+SchafferScientificRacismAgain2007,
+SainiSuperiorReturnRace2019,
+WinstonScientificRacismNorth2020,
+AdamsMisAppropriationBiological2021,
+SainiDraperMillionsPhilanthropic2022
+}, with or without its connections to PF.
+MQ was founded in 1960 by biologist R.
+Ruggles Gates (1882-1962), psychologist Henry Garrett (1894-1973), and non-academic anthropologist G.
+Robert Gayre (self-styled as "Gayre of Gayre and Nigg"; 1907-1996).
+Gates' professional status was closely tied to eugenics and the explicit scientific racism of the 1920s, and he had been thoroughly marginalized in the wake of World War II \cite{WinstonScienceServiceFar1998}.
+In contrast, Garrett had been president of the American Psychological Association in 1946 and chair of Psychology at Columbia from 1941 to 1955 *[cite]*.
+
+In the landmark case *Brown v Board of Education of Topeka* (1954), the US Supreme Court had banned *de jure* educational segregation.
+The Court's decision relied on expert testimony from psychologists and education researchers; but the segregationists had also put forward their own experts, including Henry Garrett \cite{WinstonScienceServiceFar1998, JacksonScienceSegregationRace2005, SchafferScientificRacismAgain2007}.
+Shortly after *Brown*, Garrett left Columbia, taking a visiting position at the University of Virginia, and becoming increasingly involved in segregationist, eugenicist, and even the neofascist Northern League.
+
+In the context of *Brown*, MQ was created as what contemporary scholars call an "echo chamber" \cite{FernandezPintoKnowBetterNot2017}, a favorable venue for race scientists to publish their views, on the grounds that an "equalitarian dogma" created a censorious "taboo" against their research in mainstream publications \cite{TuckerFundingScientificRacism2002, JacksonMythicalTabooRace2020}.
+Historians, philosophers, and sociologists of science have shown that echo chambers have been a key part of the "tobacco strategy," used by numerous regulated industries — most infamously tobacco, but also fossil fuels and chemical manufacturing, among others — to "manufacture doubt" and delay regulation to protect public and environmental health *[cites]*.
+
+In light of the origins of MQ, the significant attention paid to the journal by historians of scientific racism, and contemporary research on the "tobacco strategy," it is surprising that MQ did not play a significant role in nurturing race science.
+
+# *Mankind Quarterly* and Pioneer-funded researchers
+
+We identified 16 researchers who had received funding from Pioneer; 14 of these researchers had profiles in the Web of Science (WoS) author search *[table + WoS pub coucnt]*, allowing us identify 13 WoS-indexed journals that had published 6 or more of these authors.
+
+| | | |
+|:-----|-----:|:----|
+| Thomas J. Bouchard, Jr. | psychology | |
+| Brunetto Chiarelli | anthropology? | |
+| Hans Eysenck | psychology | |
+| Robert Gordon | sociology | |
+| Linda Gottfredson | psychology | |
+| Garrett Hardin | ecology | |
+| Joseph M. Horn | psychology | |
+| Lloyd Humphreys | psychology | |
+| Arthur Jensen | psychology | |
+| Michael Levin | philosophy | |
+| Richard Lynn | psychology | |
+| R. Travis Osborne | psychology | |
+| J. Phillippe Rushton | psychology | |
+| Audrey M. Shuey | psychology | |
+| Philip A. Vernon | psychology | |
+| Daniel Vining, Jr. | demography | |
+
+Table: Pioneer-funded researchers.  Either identified in *[cite]* or named on Pioneer's website, along with identified discipline and WoS author search result counts.
+
+![Journals publishing 6 or more Pioneer-funded researchers, WoS author search results *[date coverage]*](img/wos_results.png){width="4.76in" height="2.6in"}
+
+Figure *[fig]* shows that, while MQ is on this list of "Pioneer-publishing" journals, a number of mainstream journals are more prominent: *Personality and Individual Differences* (PID), *Intelligence* (I), *Behavior Genetics* (BG), and *Psychological Reports* (PR).
+In addition, only psychologist Richard Lynn appears to have published heavily in MQ.
+Lynn has been an assistant editor of MQ since 1979 *[check this: https://rationalwiki.org/wiki/Mankind_Quarterly]* and president of PF since the death of psychologist J.
+Phillippe Rushton in 2012 *[cite]*.
+By contrast, a number of Pioneer-funded researchers have published in PID, I, and to a lesser degree BG: Bouchard, Eysenck, Jensen, Rushton, Vernon, and also Lynn.
+
+PID was founded in 1980, with Eysenck as editor-in-chief and an editorial board including Jensen and Lynn.
+In the inaugural editorial, Eysenck identified "studies of the genetic determinants of individual differences in the areas of personality and intelligence" as one of the journal's eight major areas of interest.
+Eysenck remained editor-in-chief until his death in 1997.
+In 2005 the editorial board still included Jensen and Lynn.
+PID was first published by Pergamon Press, a mainstream academic press, and today is published by Elsevier.
+According to the Scimago Journal Rankings, based on Scopus citation data, in 2022 PID had an H-index of 193, #29 in Psychology, and a citation index of 5.27, #24 in Psychology.
+
+*[similar para on I]*
+
+*[segue to topic model analysis]*
+
+# Topic model analysis
+
+*[max ~50 refs]*
 
@@
 # Materials and methods
-Appendix figures and tables will also need to be moved. A basic example is provided here so this feature is also tested.
 
-First we have a figure.
+*[probably some introduction]*
 
-```{r thrice, echo=FALSE, fig.cap='test thrice'}
-plot(rnorm(10))
-```
+## Corpus assembly
 
-And next we have a table
+Corpus assembly is summarized in figure *[xref]*.  The corpus was assembled in two parts, one for *Mankind Quarterly* and the other for mainstream journals.
 
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{ll}
-thrice & 1 \\
-thrice & 2\\
-\end{tabular}
-\end{center}
-\end{table}
+*[figure]*
+
+### Mankind Quarterly
+
+An electronic archive of *Mankind Quarterly*, from the first issue in *[1969]* through 2004, is available for free on the open web from *The Unz Review*, which has been characterized as a white nationalist website by the Southern Poverty Law Center *[<https://www.splcenter.org/hatewatch/2021/01/19/meet-white-nationalist-organizer-who-spewed-hate-against-lawmakers>]*.
+Web scraping packages and techniques *[which?]* were used to retrieve every PDF available in this archive.
+These PDFs were processed using the open-source OCR (optical character recognition) software `tesseract` and the R API `rtesseract` *[cites]*.
+This stage of corpus assembly was conducted in Fall 2021.
+
+### Mainstream journals
+
+To identify suitable mainstream journals for inclusion in the corpus, we first identified academic researchers who had been funded by the Pioneer Fund.
+Pioneer is an American non-profit organization founded in 1937 to fund eugenics research *[cite]*.
+After the *Brown v.
+Board of Education* ruling in 1954, Pioneer funded various segregationist efforts across the United States.
+*[Shockley and Jensen]*.
+We reviewed a critical profile of Pioneer *[cite]* as well as an archived page from the organization's own web site *[cite]*, which together listed 16 researchers who had received Pioneer funds.
+
+We then used the author search tool in Clarivate's Web of Science platform *[url]*, retrieving publication lists for 14 researchers.
+These searches were conducted between 2021-09-24 and 2021-10-05 by DJH.
+After parsing these results, we counted how many of the 14 researchers had published in each journal.
+13 journals had published 6 or more of the 14 Pioneer-funded researchers.
+We excluded *Mankind Quarterly* (as already included) as well as *Science* and *Nature* (as too general) from further consideration in this side of the corpus.
+3 journals published by the American Psychological Association (APA; *American Psychologist*, *Contemporary Psychology*, and *Journal of Educational Psychology*) had to be excluded due to confusion over who could give us permission to use the archives for a text mining project, with both APA and ProQuest asserting that we needed to get permission from the other entity.
+*European Journal of Personality*, published by SAGE, also had to be excluded because our institutional access only went back to 1999.
+
+The remaining 5 journals are all published by major academic publishers — Elsevier, Springer, or Cambridge University Press — and each item in the entire run of each journal has been assigned a DOI (digital object identifier) for archival purposes.
+We used the Crossref API and `rcrossref` R interface to this API *[cite]* to retrieve metadata for each item published from 1960-2010 in each of these 6 journals.
+These metadata included item-level license information — confirming that the text of each item could be used for text mining projects — and a URL to an electronic version of the item.
+These URLs were used to retrieve an HTML or PDF version of each item, except for *Personality and Individual Differences*.
+This journal has published a relatively large number of non-article documents, such as book reviews and commentaries, which are not available at the URL included in the Crossref metadata.
+(This is unfortunate, as it was not difficult to find highly relevant documents that we could not automatically retrieve and therefore excluded from our corpus *[review]*.)  Instead we used Elsevier's ScienceDirect API *[url]* to search and retrieve all items from *Personality and Individual Differences* that are available.
+PDFs were run through OCR and text extraction as necessary, as with the *Mankind Quarterly* documents.
+
+All together, *[corpus summary, total and by journal; table]*
+
+## Data preparation
+
+After document retrieval and text extraction, we pre-processed the text using the `spaCy` NLP (natural language processing) Python library and the R API `spacyr` *[version, cites]*.
+Specifically, we applied regular expressions to remove header/footer copyright notices and hyphenation, used spaCy to annotate and extract noun phrases (eg, "the intelligence test items"), and then cleaned and standardized these phrases (eg, removing the/an/a, coverting all text to lowercase, and replacing all whitespace with underscores: "intelligence_test_items").
+We then counted the occurrence of each noun phrase in the document.
+The aggregated "document-term matrix" was stored in Parquet format *[cite]* for performance reasons, and written and read using the `arrow` package for R *[cite]*.
+*[availability]*
+
+NLP-extracted noun phrases offer a number of advantages over the more traditional unigram ("single-word") tokens.
+First, noun phrase extraction removes many standard stopwords (articles, common verbs) without relying on a fixed, a priori stopword list.
+Phrases can be more informative than single terms, for example, distinguishing "intelligence test" from "hypothesis test."  But simple n-gram extraction will include numerous phrases that are not especially meaningful.
+Consider the sentence "Since our first analyses of feeding patterns in rats, we had been using a criterion of 40 minutes (Le Magnen & Tallon 1963; 1966)" *[10.1017__^__s0140525x0000042x]*.
+Bigrams such as "since our" and "criterion 40" will likely be discarded in vocabulary selection, but significantly increase the computational cost of vocabulary selection.
+Noun phrase extraction is therefore more efficient.
+
+After noun phrase extraction, the corpus comprised *[tokens]* of *[distinct noun phrases]*.
+
+### Vocabulary selection
+
+We took an information-theoretic approach to vocabulary selection.
+Consider a game in which I draw a document from the corpus, then a single token from that document.
+I tell you the term, and you have to guess which document I picked.
+Intuitively, highly informative terms (in this project, noun phrase types) are distinctive, allowing you to dramatically narrow down the list of potential documents.
+This "informativeness" of a term can be quantified as the KL (Kullback-Leibler) divergence from a "baseline" distribution of documents to the distribution conditional on the term.
+Because the most informative terms tend to be typos and OCR errors — these are unique to a single document — we multiply the KL divergence by the logarithm of the total number of occurrences of the term across the entire corpus.
+We refer to the resulting measure as $log(n) \Delta H$ or simply `ndH`.
+
+*[Hicks 2021?]* used the `ndH` approach in a topic modeling analysis, where the baseline distribution was the uniform distribution across documents.
+In the current project, we found that this approach heavily favored recurrent noun phrases in the longest documents.
+Many of the documents published in *Psychological Reports* are extremely short, 1-2 reports of a single study; while many of the documents published in *Brain and Behavioral Sciences* are book-length collections that include a long review article and sometimes dozens of commentaries.
+Very generic noun phrases that happen to appear in the latter can occur orders of magnitude more often than highly distinctive phrases in the former, and so the $log(n)$ factor overwhelms the $\Delta H$ factor.
+
+To address this, we used a different baseline distribution of documents, namely, one in which document probability is proportional to length.
+This makes phrases from short documents much more "surprising" (much less likely to occur according to the baseline), and hence substantially increases their informativeness.
+This was more effective at identifying useful phrases from across the corpus.
+
+A common rule of thumb in topic modeling is that the vocabulary should have about 10 times as many distinct terms as the number of documents in the corpus *[cite]*.
+However, we had some concerns with computational demands here:  the resulting document-term matrix would have roughly $10 \times 33,000^2$ or 10.9 billion entries; with 10% density this would require on the order of 4 GB of memory just for a single copy of the matrix; and most of the analysis was to be conducted on the authors' laptops.
+We therefore chose to work with three smaller vocabularies, $5 \times$, $1 \times$ and $\frac{1/5} \times$ the number of documents.
+We refer to these as the "large," "medium," and "small" vocabulary, respectively, and they include *[counts]* distinct phrases.
+We compare findings across vocabularies as a robustness check.
+
+## Topic modeling
+
+To fit topic models, we followed the approach proposed by *[cite]*, which uses varimax-rotated partial principal components instead of the variational inference methods used by standard topic model packages *[eg, stm]*.
+This novel approach was implemented in the R package `tmfast`, *[cite]*.
+A simulation study of `tmfast` found that it was significantly faster and only slightly less accurate at reconstructing known word-topic and topic-document distributions, compared to the standard topic modeling package `stm`.
+
+Topic models were fit for all three vocabularies (large, medium, and small) with $k = 5, 10, 20, 30, 40, 50$ (number of topics), resulting in a total of 18 models.
+
+Following the approach of *[Hicks 2021?]*, we did not attempt to identify a unique best model for further analysis.
+While we focus on the medium vocabulary, $k=30$ model as the "median" among the 18 models fit, we compare and contrast findings from this model with those from the other models.
+
+### Topic model interpretation
+
+After fitting topic models, our first research question was whether we could identify distinctive "race science" topics.
+We addressed this question in two ways.
+Following common practice in topic modeling, we extracted lists of the top $n$ (highest-probability) terms (noun phrases) from each topic in each model, with $n$ generally ranging between 5 and 15 as we explored the models.
+*[tables + Silge plots]*
+
+We also adapted the "discursive space" analysis proposed by *[Hicks 2021?]*.
+This analysis combines the Hellinger distance between topic-document distributions — capturing how "far apart" two documents are — with dimensional reduction techniques used for data visualization.
+This produces a two-dimensional visualization of individual documents, arranged so that documents with similar topic distributions are visually clustered and visually distinct from documents with different topic distributions.
+*[availability]*
+
+Specifically, for each fitted topic model, we extracted topic-document distributions for each document and calculated pairwise Hellinger distances.
+We then processed these distances using the UMAP algorithm for dimension reduction, as implemented in the `umap` package *[details; cite]*.
+To reduce computational requirements for these steps, we first dropped documents published in the journal *[Psychological Reports]* (*[x documents removed, y remaining]*; note that the full corpus was used to fit the topic models, *Psychological Reports* was removed just for this visualization).
+Using the `plotly` library *[cite]*, we construct an interactive visualization:  each point represents a single document, and points can be colored based on either their journal or highest-probability topic.
+Tooltips are displayed when we hover over a point, including document metadata (title, journal, authors, year) as well as "key phrases."
+
+To calculate key phrases for a given document $d$, we first identified the highest-probability topic $t$ for that document and the word-topic (noun phrase-topic) distribution $\beta_{w,t}$ for that topic.
+We also retrieved the document-term matrix, containing the count $n_{w,d}$ of occurrences for each word (noun phrase), and calculated a weighting $n_{w,d} \beta_{w,t}$ for each word.
+The key phrases are the 5 most highly-weighted terms for the given document.
+
+### Topic quality assessment
+
+We next conducted a topic quality check of topic 19 from the medium vocabulary, $k=30$ model.
+Keyword and inspection of the UMAP interactives seemed to indicate that this topic was a "race science discourse" topic in this model.
+
+To confirm this initial impression, a spreadsheet of all articles with $\gamma > 0.25$ for this topic was extracted (excluding articles published in *Psychological Reports*), and the top 200 articles were reviewed manually by both authors.
+For this qualitative analysis, we defined *scientific racism* as (purporting) to justify racial inequality and colonialism by appealing to the epistemic authority of science; *race science* as (pseudo-)scientific research that can be utilized for scientific racism; and *race science discourse* as treating race science as a legitimate area of scientific research; this includes methodological critiques of race science and empirical tests that falsify race science hypotheses.
+Our qualitative review of the top-200 topic-19 articles focused on classifying them as *race science discourse*.
+
+For the first round of review, both authors worked independently,  dichotomously coding each article as race science discourse or not.
+We calculated the "false positive rate" — documents with a very high $\gamma$ in this topic that were not race science discourse — and interrater reliability using Cohen's $\kappa$.
+*[false positive rate]* 
+*[Initial interrater reliability was low, $\kappa = 0.4$, with 50 out of 200 articles classified discordantly.]*  To understand the disagreement, both authors examined the 50 discordant articles using an interactive grounded theory approach.
+After this process, some initial codings were changed and *[almost]* all of the remaining *[n]* discordant articles were coded as follows:
+
+hereditarianism and/or eugenics
+ ~ 29 papers.
+On the “eugenics” side, mostly examines or otherwise discusses the "dysgenic" claim that "low-intelligence" people have more children than "high-intelligence" people.
+The “hereditarianism” side is more varied, with papers on environmentalist interventions to raise IQ, genetics of IQ, and several other sub-topics.
+
+Flynn effect 
+ ~ 15 papers.  Supporting, challenging, or otherwise focusing on Flynn’s findings that IQ has increased over time.
+
+Lynn/national IQ
+ ~ 3 papers.  Supports, challenges, or focuses on international IQ data/comparisons made by Richard Lynn.
+
+We agreed that, for all three categories of papers, there was reasonable ambiguity about whether they should be classified as race science discourse.
+These papers either do not explicitly mention race, or do so with just a brief mention of the "race and intelligence controversy."  On the other hand, they either contribute to or legitimatize areas of research that are deeply entangled with race science, and might thereby contribute to public perceptions of a legitimate scientific debate about race and intelligence.
+
+Based on this qualitative review, we judged that the topic model approach was sufficiently sensitive (low "false positive" rate) in identifying race science discourse.

### 2023-06-12 paper/paper.tex
readme for paper
working on main text

@@
 
+% From pandoc table feature
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+
 % Pandoc citation processing
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+% for Pandoc 2.8 to 2.10.1
+\newenvironment{cslreferences}%
+  {}%
+  {\par}
+% For Pandoc 2.11+
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc} % for calculating minipage widths
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 \\
-\textsuperscript{*}To whom correspondence should be addressed; E-mail:
+\textsuperscript{*}Corresponding author: Daniel J. Hicks,
 \href{mailto:dhicks4@ucmerced.edu}{\nolinkurl{dhicks4@ucmerced.edu}}.
@@
 \hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section*{Introduction}\label{introduction}}
 
-foo \cite{JacksonScienceSegregationRace2005}
+\emph{{[}hey you should read this paper{]}}
+
+\emph{{[}incorporate def'ns{]}} scientific racism \textasciitilde{}
+purporting to justify racial inequality and colonialism by appealing to
+the epistemic authority of science
+
+\begin{description}
+\tightlist
+\item[race science]
+(pseudo-)scientific research that can be utilized for scientific racism
+\item[race science discourse]
+treating race science as a legitimate area of scientific research; this
+includes methodological critiques of race science and empirical tests
+that falsify race science hypotheses
+\end{description}
+
+\hypertarget{race-science-1910-1960}{%
+\section*{Race science 1910-1960}\label{race-science-1910-1960}}
+
+\hypertarget{the-rise-and-fall-of-eugenics}{%
+\subsection*{The rise and fall of
+eugenics}\label{the-rise-and-fall-of-eugenics}}
+
+\begin{itemize}
+\tightlist
+\item
+  scientific racism is thoroughly entangled with eugenics, but still
+  conceptually distinct (Slobodian 2023)
+
+  \begin{itemize}
+  \tightlist
+  \item
+    sterilization laws focused on categories of criminality and mental
+    ability, rather than race per se
+  \item
+    contemporary defenders of eugenics at least sometimes claim to
+    oppose racism
+  \item
+    claims of innate racial differences in intelligence can be made
+    without dysgenic prophecies of a rapidly-expanding biological
+    underclass
+  \end{itemize}
+\item
+  1910: Davenport founds Eugenics Records Office at Cold Spring Harbor
+\item
+  1926: NRC ``Committee on the Negro'' includes both Davenport and Boas:
+  ``neither was influential enough to veto his rival's participation''
+  (Barkan 1992, 113)
+\item
+  1930s: \emph{Coming of Age in Samoa} (Mead 1928) popularizes cultural
+  anthropology, ``free{[}ing{]} anthropology from the shackles of
+  biology'' (Barkan 1992, 134)
+\item
+  1937: Pioneer Fund is founded to ``support academic research and the
+  '{[}sic{]}dissemination of information, into the `problem of heredity
+  and eugenics' and `the problems of race betterment'\,'' (Mehler 1989,
+  21, quoting Laughlin)
+\item
+  1940: Carnegie Institute defunds Eugenics Records Office
+\item
+  1950: ``For all practical social purposes `race' is not so much a
+  biological phenomenon as a social myth'' (Beaglehole et al. 1950;
+  though see Brattain 2007)
+\end{itemize}
+
+\hypertarget{brown-v-board-and-mankind-quarterly}{%
+\subsection*{\texorpdfstring{\emph{Brown v Board} and \emph{Mankind
+Quarterly}}{Brown v Board and Mankind Quarterly}}\label{brown-v-board-and-mankind-quarterly}}
+
+Histories of scientific racism in the second half of the twentieth
+century have often emphasized the parascholarly journal \emph{Mankind
+Quarterly} (MQ) \cite{
+MehlerFoundationFascismNew1989,
+WinstonScienceServiceFar1998,
+SchafferScientificRacismAgain2007,
+SainiSuperiorReturnRace2019,
+WinstonScientificRacismNorth2020,
+AdamsMisAppropriationBiological2021,
+SainiDraperMillionsPhilanthropic2022
+}, with or without its connections to PF. MQ was founded in 1960 by
+biologist R. Ruggles Gates (1882-1962), psychologist Henry Garrett
+(1894-1973), and non-academic anthropologist G. Robert Gayre
+(self-styled as ``Gayre of Gayre and Nigg''; 1907-1996). Gates'
+professional status was closely tied to eugenics and the explicit
+scientific racism of the 1920s, and he had been thoroughly marginalized
+in the wake of World War II \cite{WinstonScienceServiceFar1998}. In
+contrast, Garrett had been president of the American Psychological
+Association in 1946 and chair of Psychology at Columbia from 1941 to
+1955 \emph{{[}cite{]}}.
+
+In the landmark case \emph{Brown v Board of Education of Topeka} (1954),
+the US Supreme Court had banned \emph{de jure} educational segregation.
+The Court's decision relied on expert testimony from psychologists and
+education researchers; but the segregationists had also put forward
+their own experts, including Henry Garrett
+\cite{WinstonScienceServiceFar1998, JacksonScienceSegregationRace2005, SchafferScientificRacismAgain2007}.
+Shortly after \emph{Brown}, Garrett left Columbia, taking a visiting
+position at the University of Virginia, and becoming increasingly
+involved in segregationist, eugenicist, and even the neofascist Northern
+League.
+
+In the context of \emph{Brown}, MQ was created as what contemporary
+scholars call an ``echo chamber''
+\cite{FernandezPintoKnowBetterNot2017}, a favorable venue for race
+scientists to publish their views, on the grounds that an ``equalitarian
+dogma'' created a censorious ``taboo'' against their research in
+mainstream publications
+\cite{TuckerFundingScientificRacism2002, JacksonMythicalTabooRace2020}.
+Historians, philosophers, and sociologists of science have shown that
+echo chambers have been a key part of the ``tobacco strategy,'' used by
+numerous regulated industries --- most infamously tobacco, but also
+fossil fuels and chemical manufacturing, among others --- to
+``manufacture doubt'' and delay regulation to protect public and
+environmental health \emph{{[}cites{]}}.
+
+In light of the origins of MQ, the significant attention paid to the
+journal by historians of scientific racism, and contemporary research on
+the ``tobacco strategy,'' it is surprising that MQ did not play a
+significant role in nurturing race science.
+
+\hypertarget{mankind-quarterly-and-pioneer-funded-researchers}{%
+\section*{\texorpdfstring{\emph{Mankind Quarterly} and Pioneer-funded
+researchers}{Mankind Quarterly and Pioneer-funded researchers}}\label{mankind-quarterly-and-pioneer-funded-researchers}}
+
+We identified 16 researchers who had received funding from Pioneer; 14
+of these researchers had profiles in the Web of Science (WoS) author
+search \emph{{[}table + WoS pub coucnt{]}}, allowing us identify 13
+WoS-indexed journals that had published 6 or more of these authors.
+
+\begin{longtable}[]{@{}lrl@{}}
+\caption{Pioneer-funded researchers. Either identified in
+\emph{{[}cite{]}} or named on Pioneer's website, along with identified
+discipline and WoS author search result counts.}\tabularnewline
+\toprule
+\endhead
+Thomas J. Bouchard, Jr. & psychology & \\
+Brunetto Chiarelli & anthropology? & \\
+Hans Eysenck & psychology & \\
+Robert Gordon & sociology & \\
+Linda Gottfredson & psychology & \\
+Garrett Hardin & ecology & \\
+Joseph M. Horn & psychology & \\
+Lloyd Humphreys & psychology & \\
+Arthur Jensen & psychology & \\
+Michael Levin & philosophy & \\
+Richard Lynn & psychology & \\
+R. Travis Osborne & psychology & \\
+J. Phillippe Rushton & psychology & \\
+Audrey M. Shuey & psychology & \\
+Philip A. Vernon & psychology & \\
+Daniel Vining, Jr. & demography & \\
+\bottomrule
+\end{longtable}
+
+\begin{figure}
+\centering
+\includegraphics[width=4.76in,height=2.6in]{img/wos_results.png}
+\caption{Journals publishing 6 or more Pioneer-funded researchers, WoS
+author search results \emph{{[}date coverage{]}}}
+\end{figure}
+
+Figure \emph{{[}fig{]}} shows that, while MQ is on this list of
+``Pioneer-publishing'' journals, a number of mainstream journals are
+more prominent: \emph{Personality and Individual Differences} (PID),
+\emph{Intelligence} (I), \emph{Behavior Genetics} (BG), and
+\emph{Psychological Reports} (PR). In addition, only psychologist
+Richard Lynn appears to have published heavily in MQ. Lynn has been an
+assistant editor of MQ since 1979 \emph{{[}check this:
+\url{https://rationalwiki.org/wiki/Mankind_Quarterly}{]}} and president
+of PF since the death of psychologist J. Phillippe Rushton in 2012
+\emph{{[}cite{]}}. By contrast, a number of Pioneer-funded researchers
+have published in PID, I, and to a lesser degree BG: Eysenck, Jensen,
+Rushton, Vernon, and also Lynn.
+
+PID was founded in 1980, with Eysenck as editor-in-chief and an
+editorial board including Jensen and Lynn. In the inaugural editorial,
+Eysenck identified ``studies of the genetic determinants of individual
+differences in the areas of personality and intelligence'' as one of the
+journal's eight major areas of interest. Eysenck remained
+editor-in-chief until his death in 1997. In 2005 the editorial board
+still included Jensen and Lynn. PID was first published by Pergamon
+Press, a mainstream academic press, and today is published by Elsevier.
+According to the Scimago Journal Rankings, based on Scopus citation
+data, in 2022 PID had an H-index of 193, \#29 in Psychology, and a
+citation index of 5.27, \#24 in Psychology.
+
+\emph{{[}similar para on I{]}}
+
+\emph{{[}segue to topic model analysis{]}}
+
+\hypertarget{topic-model-analysis}{%
+\section*{Topic model analysis}\label{topic-model-analysis}}
+
+\emph{{[}max \textasciitilde50 refs{]}}
 \bibliography{references.bib}
@@
 \begin{scilastnote}
-\item Please include your acknowledgments here, set in a single
-paragraph.
+\item \emph{{[}strutured{]}} Thanks to Emily Merchant and John Jackson
+for some initial discussions that helped clarify the scope of the
+project and identify some key resources related to the Pioneer Fund and
+\emph{Mankind Quarterly}. Thanks to Anthony Sainez for help retrieving
+the \emph{Mankind Quarterly} articles. Thanks to Derek Devnich and James
+Dooley for their work attempting to secure the APA-published articles.
+EL's work on this project was supported by UC Merced. DH's work on this
+project was not supported by any specific funding.
 \end{scilastnote}

### 2023-06-12 paper/supplement_paper.tex
readme for paper
working on main text

@@
 and Emilio Lobato
-\\\textsuperscript{*}To whom correspondence should be addressed; E-mail:
+\\\textsuperscript{*}Corresponding author: Daniel J. Hicks,
 \href{mailto:dhicks4@ucmerced.edu}{\nolinkurl{dhicks4@ucmerced.edu}}.
@@
 \hypertarget{materials-and-methods}{%
-\section{Materials and methods}\label{materials-and-methods}}
+\section*{Materials and methods}\label{materials-and-methods}}
 
-Appendix figures and tables will also need to be moved. A basic example
-is provided here so this feature is also tested.
+\emph{{[}probably some introduction{]}}
 
-First we have a figure.
+\hypertarget{corpus-assembly}{%
+\subsection*{Corpus assembly}\label{corpus-assembly}}
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/thrice-1.pdf}
-\caption{test thrice}
-\end{figure}
+Corpus assembly is summarized in figure \emph{{[}xref{]}}. The corpus
+was assembled in two parts, one for \emph{Mankind Quarterly} and the
+other for mainstream journals.
 
-And next we have a table
+\emph{{[}figure{]}}
 
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{ll}
-thrice & 1 \\
-thrice & 2\\
-\end{tabular}
-\end{center}
-\end{table}
+\hypertarget{mankind-quarterly}{%
+\subsubsection*{Mankind Quarterly}\label{mankind-quarterly}}
+
+An electronic archive of \emph{Mankind Quarterly}, from the first issue
+in \emph{{[}1969{]}} through 2004, is available for free on the open web
+from \emph{The Unz Review}, which has been characterized as a white
+nationalist website by the Southern Poverty Law Center
+\emph{{[}\url{https://www.splcenter.org/hatewatch/2021/01/19/meet-white-nationalist-organizer-who-spewed-hate-against-lawmakers}{]}}.
+Web scraping packages and techniques \emph{{[}which?{]}} were used to
+retrieve every PDF available in this archive. These PDFs were processed
+using the open-source OCR (optical character recognition) software
+\texttt{tesseract} and the R API \texttt{rtesseract} \emph{{[}cites{]}}.
+This stage of corpus assembly was conducted in Fall 2021.
+
+\hypertarget{mainstream-journals}{%
+\subsubsection*{Mainstream journals}\label{mainstream-journals}}
+
+To identify suitable mainstream journals for inclusion in the corpus, we
+first identified academic researchers who had been funded by the Pioneer
+Fund. Pioneer is an American non-profit organization founded in 1937 to
+fund eugenics research \emph{{[}cite{]}}. After the \emph{Brown v. Board
+of Education} ruling in 1954, Pioneer funded various segregationist
+efforts across the United States. \emph{{[}Shockley and Jensen{]}}. We
+reviewed a critical profile of Pioneer \emph{{[}cite{]}} as well as an
+archived page from the organization's own web site \emph{{[}cite{]}},
+which together listed 16 researchers who had received Pioneer funds.
+
+We then used the author search tool in Clarivate's Web of Science
+platform \emph{{[}url{]}}, retrieving publication lists for 14
+researchers. These searches were conducted between 2021-09-24 and
+2021-10-05 by DJH. After parsing these results, we counted how many of
+the 14 researchers had published in each journal. 13 journals had
+published 6 or more of the 14 Pioneer-funded researchers. We excluded
+\emph{Mankind Quarterly} (as already included) as well as \emph{Science}
+and \emph{Nature} (as too general) from further consideration in this
+side of the corpus. 3 journals published by the American Psychological
+Association (APA; \emph{American Psychologist}, \emph{Contemporary
+Psychology}, and \emph{Journal of Educational Psychology}) had to be
+excluded due to confusion over who could give us permission to use the
+archives for a text mining project, with both APA and ProQuest asserting
+that we needed to get permission from the other entity. \emph{European
+Journal of Personality}, published by SAGE, also had to be excluded
+because our institutional access only went back to 1999.
+
+The remaining 5 journals are all published by major academic publishers
+--- Elsevier, Springer, or Cambridge University Press --- and each item
+in the entire run of each journal has been assigned a DOI (digital
+object identifier) for archival purposes. We used the Crossref API and
+\texttt{rcrossref} R interface to this API \emph{{[}cite{]}} to retrieve
+metadata for each item published from 1960-2010 in each of these 6
+journals. These metadata included item-level license information ---
+confirming that the text of each item could be used for text mining
+projects --- and a URL to an electronic version of the item. These URLs
+were used to retrieve an HTML or PDF version of each item, except for
+\emph{Personality and Individual Differences}. This journal has
+published a relatively large number of non-article documents, such as
+book reviews and commentaries, which are not available at the URL
+included in the Crossref metadata. (This is unfortunate, as it was not
+difficult to find highly relevant documents that we could not
+automatically retrieve and therefore excluded from our corpus
+\emph{{[}review{]}}.) Instead we used Elsevier's ScienceDirect API
+\emph{{[}url{]}} to search and retrieve all items from \emph{Personality
+and Individual Differences} that are available. PDFs were run through
+OCR and text extraction as necessary, as with the \emph{Mankind
+Quarterly} documents.
+
+All together, \emph{{[}corpus summary, total and by journal; table{]}}
+
+\hypertarget{data-preparation}{%
+\subsection*{Data preparation}\label{data-preparation}}
+
+After document retrieval and text extraction, we pre-processed the text
+using the \texttt{spaCy} NLP (natural language processing) Python
+library and the R API \texttt{spacyr} \emph{{[}version, cites{]}}.
+Specifically, we applied regular expressions to remove header/footer
+copyright notices and hyphenation, used spaCy to annotate and extract
+noun phrases (eg, ``the intelligence test items''), and then cleaned and
+standardized these phrases (eg, removing the/an/a, coverting all text to
+lowercase, and replacing all whitespace with underscores:
+``intelligence\_test\_items''). We then counted the occurrence of each
+noun phrase in the document. The aggregated ``document-term matrix'' was
+stored in Parquet format \emph{{[}cite{]}} for performance reasons, and
+written and read using the \texttt{arrow} package for R
+\emph{{[}cite{]}}. \emph{{[}availability{]}}
+
+NLP-extracted noun phrases offer a number of advantages over the more
+traditional unigram (``single-word'') tokens. First, noun phrase
+extraction removes many standard stopwords (articles, common verbs)
+without relying on a fixed, a priori stopword list. Phrases can be more
+informative than single terms, for example, distinguishing
+``intelligence test'' from ``hypothesis test.'' But simple n-gram
+extraction will include numerous phrases that are not especially
+meaningful. Consider the sentence ``Since our first analyses of feeding
+patterns in rats, we had been using a criterion of 40 minutes (Le Magnen
+\& Tallon 1963; 1966)'' *{[}10.1017\_\_\^{}\_\_s0140525x0000042x{]}*.
+Bigrams such as ``since our'' and ``criterion 40'' will likely be
+discarded in vocabulary selection, but significantly increase the
+computational cost of vocabulary selection. Noun phrase extraction is
+therefore more efficient.
+
+After noun phrase extraction, the corpus comprised \emph{{[}tokens{]}}
+of \emph{{[}distinct noun phrases{]}}.
+
+\hypertarget{vocabulary-selection}{%
+\subsubsection*{Vocabulary selection}\label{vocabulary-selection}}
+
+We took an information-theoretic approach to vocabulary selection.
+Consider a game in which I draw a document from the corpus, then a
+single token from that document. I tell you the term, and you have to
+guess which document I picked. Intuitively, highly informative terms (in
+this project, noun phrase types) are distinctive, allowing you to
+dramatically narrow down the list of potential documents. This
+``informativeness'' of a term can be quantified as the KL
+(Kullback-Leibler) divergence from a ``baseline'' distribution of
+documents to the distribution conditional on the term. Because the most
+informative terms tend to be typos and OCR errors --- these are unique
+to a single document --- we multiply the KL divergence by the logarithm
+of the total number of occurrences of the term across the entire corpus.
+We refer to the resulting measure as \(log(n) \Delta H\) or simply
+\texttt{ndH}.
+
+\emph{{[}Hicks 2021?{]}} used the \texttt{ndH} approach in a topic
+modeling analysis, where the baseline distribution was the uniform
+distribution across documents. In the current project, we found that
+this approach heavily favored recurrent noun phrases in the longest
+documents. Many of the documents published in \emph{Psychological
+Reports} are extremely short, 1-2 reports of a single study; while many
+of the documents published in \emph{Brain and Behavioral Sciences} are
+book-length collections that include a long review article and sometimes
+dozens of commentaries. Very generic noun phrases that happen to appear
+in the latter can occur orders of magnitude more often than highly
+distinctive phrases in the former, and so the \(log(n)\) factor
+overwhelms the \(\Delta H\) factor.
+
+To address this, we used a different baseline distribution of documents,
+namely, one in which document probability is proportional to length.
+This makes phrases from short documents much more ``surprising'' (much
+less likely to occur according to the baseline), and hence substantially
+increases their informativeness. This was more effective at identifying
+useful phrases from across the corpus.
+
+A common rule of thumb in topic modeling is that the vocabulary should
+have about 10 times as many distinct terms as the number of documents in
+the corpus \emph{{[}cite{]}}. However, we had some concerns with
+computational demands here: the resulting document-term matrix would
+have roughly \(10 \times 33,000^2\) or 10.9 billion entries; with 10\%
+density this would require on the order of 4 GB of memory just for a
+single copy of the matrix; and most of the analysis was to be conducted
+on the authors' laptops. We therefore chose to work with three smaller
+vocabularies, \(5 \times\), \(1 \times\) and \(\frac{1/5} \times\) the
+number of documents. We refer to these as the ``large,'' ``medium,'' and
+``small'' vocabulary, respectively, and they include \emph{{[}counts{]}}
+distinct phrases. We compare findings across vocabularies as a
+robustness check.
+
+\hypertarget{topic-modeling}{%
+\subsection*{Topic modeling}\label{topic-modeling}}
+
+To fit topic models, we followed the approach proposed by
+\emph{{[}cite{]}}, which uses varimax-rotated partial principal
+components instead of the variational inference methods used by standard
+topic model packages \emph{{[}eg, stm{]}}. This novel approach was
+implemented in the R package \texttt{tmfast}, \emph{{[}cite{]}}. A
+simulation study of \texttt{tmfast} found that it was significantly
+faster and only slightly less accurate at reconstructing known
+word-topic and topic-document distributions, compared to the standard
+topic modeling package \texttt{stm}.
+
+Topic models were fit for all three vocabularies (large, medium, and
+small) with \(k = 5, 10, 20, 30, 40, 50\) (number of topics), resulting
+in a total of 18 models.
+
+Following the approach of \emph{{[}Hicks 2021?{]}}, we did not attempt
+to identify a unique best model for further analysis. While we focus on
+the medium vocabulary, \(k=30\) model as the ``median'' among the 18
+models fit, we compare and contrast findings from this model with those
+from the other models.
+
+\hypertarget{topic-model-interpretation}{%
+\subsubsection*{Topic model
+interpretation}\label{topic-model-interpretation}}
+
+After fitting topic models, our first research question was whether we
+could identify distinctive ``race science'' topics. We addressed this
+question in two ways. Following common practice in topic modeling, we
+extracted lists of the top \(n\) (highest-probability) terms (noun
+phrases) from each topic in each model, with \(n\) generally ranging
+between 5 and 15 as we explored the models. \emph{{[}tables + Silge
+plots{]}}
+
+We also adapted the ``discursive space'' analysis proposed by
+\emph{{[}Hicks 2021?{]}}. This analysis combines the Hellinger distance
+between topic-document distributions --- capturing how ``far apart'' two
+documents are --- with dimensional reduction techniques used for data
+visualization. This produces a two-dimensional visualization of
+individual documents, arranged so that documents with similar topic
+distributions are visually clustered and visually distinct from
+documents with different topic distributions. \emph{{[}availability{]}}
+
+Specifically, for each fitted topic model, we extracted topic-document
+distributions for each document and calculated pairwise Hellinger
+distances. We then processed these distances using the UMAP algorithm
+for dimension reduction, as implemented in the \texttt{umap} package
+\emph{{[}details; cite{]}}. To reduce computational requirements for
+these steps, we first dropped documents published in the journal
+\emph{{[}Psychological Reports{]}} (\emph{{[}x documents removed, y
+remaining{]}}; note that the full corpus was used to fit the topic
+models, \emph{Psychological Reports} was removed just for this
+visualization). Using the \texttt{plotly} library \emph{{[}cite{]}}, we
+construct an interactive visualization: each point represents a single
+document, and points can be colored based on either their journal or
+highest-probability topic. Tooltips are displayed when we hover over a
+point, including document metadata (title, journal, authors, year) as
+well as ``key phrases.''
+
+To calculate key phrases for a given document \(d\), we first identified
+the highest-probability topic \(t\) for that document and the word-topic
+(noun phrase-topic) distribution \(\beta_{w,t}\) for that topic. We also
+retrieved the document-term matrix, containing the count \(n_{w,d}\) of
+occurrences for each word (noun phrase), and calculated a weighting
+\(n_{w,d} \beta_{w,t}\) for each word. The key phrases are the 5 most
+highly-weighted terms for the given document.
+
+\hypertarget{topic-quality-assessment}{%
+\subsubsection*{Topic quality
+assessment}\label{topic-quality-assessment}}
+
+We next conducted a topic quality check of topic 19 from the medium
+vocabulary, \(k=30\) model. Keyword and inspection of the UMAP
+interactives seemed to indicate that this topic was a ``race science
+discourse'' topic in this model.
+
+To confirm this initial impression, a spreadsheet of all articles with
+\(\gamma > 0.25\) for this topic was extracted (excluding articles
+published in \emph{Psychological Reports}), and the top 200 articles
+were reviewed manually by both authors. For this qualitative analysis,
+we defined \emph{scientific racism} as (purporting) to justify racial
+inequality and colonialism by appealing to the epistemic authority of
+science; \emph{race science} as (pseudo-)scientific research that can be
+utilized for scientific racism; and \emph{race science discourse} as
+treating race science as a legitimate area of scientific research; this
+includes methodological critiques of race science and empirical tests
+that falsify race science hypotheses. Our qualitative review of the
+top-200 topic-19 articles focused on classifying them as \emph{race
+science discourse}.
+
+For the first round of review, both authors worked independently,
+dichotomously coding each article as race science discourse or not. We
+calculated the ``false positive rate'' --- documents with a very high
+\(\gamma\) in this topic that were not race science discourse --- and
+interrater reliability using Cohen's \(\kappa\).\\
+\emph{{[}false positive rate{]}} \emph{{[}Initial interrater reliability
+was low, \(\kappa = 0.4\), with 50 out of 200 articles classified
+discordantly.{]}} To understand the disagreement, both authors examined
+the 50 discordant articles using an interactive grounded theory
+approach. After this process, some initial codings were changed and
+\emph{{[}almost{]}} all of the remaining \emph{{[}n{]}} discordant
+articles were coded as follows:
+
+\begin{description}
+\tightlist
+\item[hereditarianism and/or eugenics]
+29 papers. On the ``eugenics'' side, mostly examines or otherwise
+discusses the ``dysgenic'' claim that ``low-intelligence'' people have
+more children than ``high-intelligence'' people. The ``hereditarianism''
+side is more varied, with papers on environmentalist interventions to
+raise IQ, genetics of IQ, and several other sub-topics.
+\item[Flynn effect]
+15 papers. Supporting, challenging, or otherwise focusing on Flynn's
+findings that IQ has increased over time.
+\item[Lynn/national IQ]
+3 papers. Supports, challenges, or focuses on international IQ
+data/comparisons made by Richard Lynn.
+\end{description}
+
+We agreed that, for all three categories of papers, there was reasonable
+ambiguity about whether they should be classified as race science
+discourse. These papers either do not explicitly mention race, or do so
+with just a brief mention of the ``race and intelligence controversy.''
+On the other hand, they either contribute to or legitimatize areas of
+research that are deeply entangled with race science, and might thereby
+contribute to public perceptions of a legitimate scientific debate about
+race and intelligence.
+
+Based on this qualitative review, we judged that the topic model
+approach was sufficiently sensitive (low ``false positive'' rate) in
+identifying race science discourse.
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-BarkanRetreatScientificRacism1992}{}}%
+Barkan, Elazar. 1992. \emph{The {Retreat} of {Scientific Racism}:
+{Changing Concepts} of {Race} in {Britain} and the {United States
+Between} the {World Wars}}. {Cambridge University Press}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-BeagleholeRaceQuestion1950}{}}%
+Beaglehole, Ernest, Juan Comas, L. A. Costa Pinto, Franklin Frazier,
+Morris Ginsberg, Humayun Kabir, Claude Levi-Strauss, and Ashley Montagu.
+1950. {``The {Race Question}.''} \emph{UNESCO and Its Programme} 1 (3).
+
+\leavevmode\vadjust pre{\hypertarget{ref-BrattainRaceRacismAntiracism2007}{}}%
+Brattain, M. 2007. {``Race, {Racism}, and {Antiracism}: {UNESCO} and the
+{Politics} of {Presenting Science} to the {Postwar Public}.''} \emph{The
+American Historical Review} 112 (5): 1386--1413.
+\url{https://doi.org/10.1086/ahr.112.5.1386}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-MeadComingAgeSamoa1928}{}}%
+Mead, Margaret. 1928. \emph{Coming of {Age} in {Samoa}}. {New York}:
+{William Morrow and Company}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-MehlerFoundationFascismNew1989}{}}%
+Mehler, Barry. 1989. {``Foundation for Fascism: {The} New Eugenics
+Movement in the {United States}.''} \emph{Patterns of Prejudice} 23 (4):
+17--25. \url{https://doi.org/10.1080/0031322X.1989.9970026}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-SlobodianUnequalMindHow2023}{}}%
+Slobodian, Quinn. 2023. {``The {Unequal Mind}: {How Charles Murray} and
+{Neoliberal Think Tanks Revived IQ}.''} \emph{Capitalism: A Journal of
+History and Economics} 4 (1): 73--108.
+
+\end{CSLReferences}
 

### 2023-06-22 paper/paper.Rmd
paper stuff

@@
 
+*[for xref: https://bookdown.org/yihui/bookdown/a-single-document.html]*
+
 We identified 16 researchers who had received funding from Pioneer; 14 of these researchers had profiles in the Web of Science (WoS) author search *[table + WoS pub coucnt]*, allowing us identify 13 WoS-indexed journals that had published 6 or more of these authors.
@@
 
-## Data preparation
+## Text preparation
 
@@
 
-### Vocabulary selection
+### Data and code availability
+
+Document metadata retrieved from Crossref does not appear to be covered by any copyright or other intellectual property restrictions.
+However, due to copyright restrictions, we are unable to make document fulltext or Web of Science search results publicly available.
+Code for the corpus assembly and text preparation steps described above is available by request.
+
+The public analysis repository, <https://github.com/dhicks/race-science> *[DOI]*, includes document metadata and documentwise counts of NLP-extracted noun phrases ("document-term matrices"), along with code to reproduce the analysis of the following sections.
+*[repo readme: requires R 4.1.2; rig (<https://github.com/r-lib/rig/>) for managing multiple R installations]*
+
+## Vocabulary selection
 

### 2023-06-22 R/phrases.R
working through implications of removing whitespace phrases

@@ -1,6 +1,8 @@
 phrases = function() {
     open_dataset(here(data_dir, '00_phrases'))  |> 
         select(-year) |> 
-        filter(!is.na(phrase))
+        filter(!is.na(phrase), 
+               phrase != '', 
+               phrase != '_')
 }
 

### 2023-06-22 scripts/04_topic_models.R

@@ -12,12 +12,12 @@ library(glue)
 
 data_dir = here('data')
 
-k = c(5, seq(10, 50, by = 10))
+k = c(5, seq(10, 70, by = 10))
 ## Models are not re-fit automatically.  To refit: 
 ## 1. Pass --force through the command line 
 source(here('R', 'args.R'))
 ## 2. Set `force` manually here
-force = TRUE
+# force = TRUE
 
 ## Load data ----
 meta_ar = open_dataset(here(data_dir, '01_metadata'))

### 2023-06-22 scripts/05_all_gamma.R

@@ -100,7 +100,7 @@ all_gammas = function(path, name) {
     message('Building topic-doc tile visualizations')
     ggplot(comb_df, aes(topic, article_id, fill = log1p(gamma))) +
         geom_raster() +
-        geom_vline(xintercept = c(10, 20, 30, 40) + .5, 
+        geom_vline(xintercept = c(10, 20, 30, 40, 50, 60) + .5, 
                    color = 'black') +
         facet_grid(rows = vars(container.title), 
                    cols = vars(k), 
@@ -108,7 +108,7 @@ all_gammas = function(path, name) {
                    labeller = label_wrap_gen(width = 15), 
                    switch = 'y') +
         scale_x_discrete(breaks = c('V10', 'V20', 'V30',
-                                    'V40', 'V50')) +
+                                    'V40', 'V50', 'V60', 'V70')) +
         scale_y_discrete(breaks = NULL, 
                          name = '') +
         scale_fill_viridis_c(option = 'viridis', direction = 1) +
@@ -117,7 +117,7 @@ all_gammas = function(path, name) {
               legend.position = 'bottom', 
               plot.margin = margin(r = 10))
     ggsave(here(out_dir, glue(out_prefix, '_gamma.png')), 
-           width = 6, height = 6.5, scale = 1.5, bg = 'white')
+           width = 8, height = 6.5, scale = 1.5, bg = 'white')
 }
 
 iwalk(model_files, all_gammas)

### 2023-06-22 scripts/06_hellinger_betas.R

@@ -29,6 +29,7 @@ hellinger_betas = function(model_file, name) {
     model = read_rds(model_file)
     
     ## Renormalization
+    message('Calculating renormalization exponents')
     get_power = function(this_k, model) {
         ee = expected_entropy(.1, this_k)
         model |> 
@@ -40,6 +41,7 @@ hellinger_betas = function(model_file, name) {
                                .progress = TRUE)
     # return(exponents)
     
+    message('Extracting betas for Silge plots')
     betas = tidy_all(model) |> 
         group_split(k) |> 
         map2(exponents, ~ renorm(.x, topic, beta, .y))
@@ -72,31 +74,33 @@ hellinger_betas = function(model_file, name) {
                                        linewidth = 1))}, 
                 .keep = TRUE) |> 
         wrap_plots(design = 
-                       'AB
-               #B
-               CD
-               CD
-               CD
-               CD
-               #D
-               #D
-               EF
-               EF
-               EF
-               EF
-               EF
-               EF
-               EF
-               EF
-               #F
-               #F') +
+              'ABCD
+               #BCD
+               ##CD
+               ##CD
+               ###D
+               EFGH
+               EFGH
+               EFGH
+               EFGH
+               EFGH
+               EFGH
+               EFGH
+               EFGH
+               #FGH
+               #FGH
+               ##GH
+               ##GH
+               ###H
+               ###H') +
         labs(caption = glue('{name} vocabulary'))
     ggsave(here(out_dir, glue('06-{name}-silge.pdf')), 
-           width = 10, height = (10+30+50)/5, 
+           width = 4*5, height = (30+70)/5, 
            limitsize = FALSE,
            scale = 3)
     
     ## Pairwise Hellinger distance/similarity
+    message('Hellinger similarities')
     vocab = model$cols
     
     compared = map(1:(length(betas) - 1), 
@@ -126,6 +130,7 @@ hellinger_betas = function(model_file, name) {
                comparison = fct_inorder(comparison))
     
     ## Heat map
+    message('Constructing Hellinger visualizations')
     ggplot(comp_df, aes(topic_b, topic_a, fill = similarity)) +
         geom_raster() +
         facet_wrap(vars(comparison), 
@@ -142,24 +147,26 @@ hellinger_betas = function(model_file, name) {
               legend.position = c(.9, .2),
               legend.justification = c(1, 0))
[…]

### 2023-06-22 scripts/07_umap_interactive.R

@@ -201,8 +201,8 @@ make_umap = function(vocab,
 
 # debugonce(process_k)
 ## Generate a single spreadsheet for topic quality check
-# make_umap('md', 30, spreadsheet = 'V19', visualization = FALSE)
+# make_umap('md', 40, spreadsheet = 'V24', visualization = FALSE)
 ## Generate visualizations for all models of interest
-make_umap('md', c(20, 30, 40), seed = seed, force = force)
-make_umap('sm', c(20, 30, 40), seed = seed, force = force)
-make_umap('lg', c(20, 30, 40), seed = seed, force = force)
+make_umap('md', c(20, 30, 40, 50, 60, 70), seed = seed, force = force)
+make_umap('sm', c(20, 30, 40, 50, 60, 70), seed = seed, force = force)
+make_umap('lg', c(20, 30, 40, 50, 60, 70), seed = seed, force = force)

### 2023-06-22 scripts/08_topics_over_time.R

@@ -36,7 +36,7 @@ md_exp = read_rds(here(data_dir, '05_md_exponents.Rds')) |>
 
 md_tmf |> 
     tidy(k = 30, matrix = 'gamma', exponent = md_exp['30']) |> 
-    filter(topic %in% c('V05', 'V07', 'V19')) |> 
+    filter(topic %in% c('V05', 'V07', 'V22')) |> 
     ggplot(aes(gamma, color = topic)) +
     stat_ecdf() +
     geom_hline(yintercept = .9)
@@ -48,6 +48,7 @@ create_dataframes = function(model,
                              articles = articles_df, 
                              gamma_threshold = .05,
                              agg_fn = mean) {
+    ## Create a dataframe of aggregated gamma scores by year
     full_topic_list = list()
     
     for (i in k){
@@ -81,7 +82,7 @@ create_dataframes = function(model,
 
 ## Mean value of gamma, no threshold
 create_dataframes(md_tmf, md_exp,
-                  topics = c('V07', 'V19', 'V05'), 
+                  topics = c('V07', 'V22', 'V05', 'V24'), 
                   gamma_threshold = 0) |> 
     ggplot(aes(year, gamma, 
                color = container.title, 
@@ -93,7 +94,7 @@ create_dataframes(md_tmf, md_exp,
 
 ## Mean value of gamma, 1% threshold
 create_dataframes(md_tmf, md_exp,
-                  topics = c('V07', 'V19', 'V05'), 
+                  topics = c('V07', 'V22', 'V05', 'V24'), 
                   gamma_threshold = .1) |> 
     ggplot(aes(year, gamma, 
                color = container.title, 
@@ -105,7 +106,7 @@ create_dataframes(md_tmf, md_exp,
 
 ## Mean value of gamma, 25% threshold
 create_dataframes(md_tmf, md_exp,
-                  topics = c('V07', 'V19', 'V05'), 
+                  topics = c('V07', 'V22', 'V05'), 
                   gamma_threshold = .25) |> 
     ggplot(aes(year, gamma, 
                color = container.title, 
@@ -115,43 +116,63 @@ create_dataframes(md_tmf, md_exp,
                cols = vars(topic), 
                scales = 'free_y')
 
-count_plot = function(tmf, vocab, exp, this_k, topics, threshold) {
+count_plot = function(tmf, vocab, exp, this_k, topics, threshold, 
+                      plot = TRUE,
+                      position = 'identity', 
[…]

### 2023-06-22 scripts/09_umap_static.R

@@ -62,7 +62,9 @@ calc_densities = function(dataf, var1, var2, group_var, group_values, ...) {
 }
 
 ## Static UMAP visualization with highlighted topics
-visualize = function(vocab, k, highlight_topics, write = TRUE) {
+visualize = function(vocab, k, highlight_topics, 
+                     breaks_threshold = .15,
+                     write = TRUE) {
     model = here(tm_dir, glue('04_{vocab}_tmfast.Rds')) |> 
         read_rds()
     
@@ -102,7 +104,7 @@ visualize = function(vocab, k, highlight_topics, write = TRUE) {
         geom_point(alpha = .01) +
         geom_textcontour(aes(label = topic, z = z), linewidth = 1,
                          upright = TRUE, straight = TRUE,
-                         breaks = .15,
+                         breaks = breaks_threshold,
                          data = densities) +
         scale_color_viridis_d(guide = 'none') +
         lims(x = window$x, 
@@ -119,18 +121,21 @@ visualize = function(vocab, k, highlight_topics, write = TRUE) {
 }
 
 # debugonce(visualize)
-visualize('md', 30, c('V05', 'V07', 'V19'))
+visualize('md', 40, c('V05', 'V07', 'V22', 'V24'), 
+          breaks_threshold = .1,
+          write = FALSE)
 
 topics_of_interest = tribble(
     ~ vocab, ~ k, ~ highlight_topics,
-    'sm', 20, c('V04', 'V12', 'V20'),
-    'sm', 30, c('V04', 'V12', 'V20'),
-    'sm', 40, c('V04', 'V12', 'V20', 'V34'),
-    'md', 30, c('V05', 'V07', 'V19'),
-    'md', 40, c('V05', 'V07', 'V22', 'V24', 'V36'),
-    'lg', 20, c('V05', 'V06'),
-    'lg', 30, c('V06', 'V24', 'V27'),
-    'lg', 40, c('V06', 'V24', 'V26')
+    'sm', 30, c('V04', 'V23'),
+    'sm', 40, c('V04', 'V28', 'V36'),
+    'sm', 50, c('V04', 'V36', 'V47'),
+    'md', 30, c('V05', 'V07', 'V22'),
+    'md', 40, c('V05', 'V07', 'V22', 'V24', 'V35'),
+    'md', 50, c('V05', 'V07', 'V22', 'V24', 'V35')
[…]

### 2023-06-22 scripts/08_topics_over_time.R
updated 08 to produce big grid

@@ -6,11 +6,11 @@ theme_set(theme_bw())
 library(tmfast)
 library(here)
 library(arrow)
-# remotes::install_github("mtennekes/cols4all")
-# library(cols4all)
 library(patchwork)
 library(glue)
+library(ggh4x)
 
+## Load data ----
 data_dir = here('data')
 tm_dir = here(data_dir, '04_tm')
 out_dir = here('out')
@@ -28,27 +28,28 @@ articles_df = phrases_ar |>
     # filter(container.title != 'Psychological Reports') |> 
     collect()
 
-# sm_tmf = readRDS(here(tm_dir, '04_sm_tmfast.Rds'))
-md_tmf = readRDS(here(tm_dir, '04_md_tmfast.Rds'))
+model_files = list.files(tm_dir, '*_tmfast.Rds') %>%
+    here(tm_dir, .) |> 
+    set_names('lg', 'md', 'sm')
+exp_files = list.files(data_dir, '*_exponents.Rds') %>%
+    here(data_dir, .) |> 
+    set_names('lg', 'md', 'sm')
+## Helper to read exponents as a named num
+read_exp = function(file) {
+    file |> 
+        read_rds() |> 
+        pull(exponent, name = k)
+}
 
-md_exp = read_rds(here(data_dir, '05_md_exponents.Rds')) |> 
-    pull(exponent, name = k)
 
-md_tmf |> 
-    tidy(k = 30, matrix = 'gamma', exponent = md_exp['30']) |> 
-    filter(topic %in% c('V05', 'V07', 'V22')) |> 
-    ggplot(aes(gamma, color = topic)) +
-    stat_ecdf() +
-    geom_hline(yintercept = .9)
-
+## Create a dataframe of aggregated gamma scores by year ----
 create_dataframes = function(model, 
                              exponents,
-                             k = c(5, seq(10, 50, by = 10)), 
+                             k = c(5, seq(10, 70, by = 10)), 
                              topics = NULL, 
                              articles = articles_df, 
                              gamma_threshold = .05,
                              agg_fn = mean) {
-    ## Create a dataframe of aggregated gamma scores by year
     full_topic_list = list()
     
     for (i in k){
@@ -80,55 +81,57 @@ create_dataframes = function(model,
     return(bind_rows(full_topic_list, .id = 'k'))
 }
 
-## Mean value of gamma, no threshold
-create_dataframes(md_tmf, md_exp,
[…]

### 2023-06-22 scripts/09_umap_static.R

@@ -121,7 +121,7 @@ visualize = function(vocab, k, highlight_topics,
 }
 
 # debugonce(visualize)
-visualize('md', 40, c('V05', 'V07', 'V22', 'V24'), 
+visualize('md', 40, c('V07', 'V24', 'V05', 'V22', 'V35', 'V38'), 
           breaks_threshold = .1,
           write = FALSE)
 
@@ -129,13 +129,13 @@ topics_of_interest = tribble(
     ~ vocab, ~ k, ~ highlight_topics,
     'sm', 30, c('V04', 'V23'),
     'sm', 40, c('V04', 'V28', 'V36'),
-    'sm', 50, c('V04', 'V36', 'V47'),
+    'sm', 50, c('V04', 'V47', 'V36', 'V25', 'V46'),
     'md', 30, c('V05', 'V07', 'V22'),
-    'md', 40, c('V05', 'V07', 'V22', 'V24', 'V35'),
-    'md', 50, c('V05', 'V07', 'V22', 'V24', 'V35')
-    # 'lg', 20, c('V05', 'V06'),
-    # 'lg', 30, c('V06', 'V24', 'V27'),
-    # 'lg', 40, c('V06', 'V24', 'V26')
+    'md', 40, c('V07', 'V24', 'V05', 'V22', 'V35', 'V38'),
+    'md', 50, c('V07', 'V24', 'V05', 'V22', 'V35', 'V45'),
+    'lg', 30, c('V27', 'V06'),
+    'lg', 40, c('V27', 'V02', 'V06', 'V26', 'V32', 'V37'),
+    'lg', 50, c('V27', 'V06', 'V32', 'V34', 'V35', 'V49')
 )
 
 pwalk(topics_of_interest, 

### 2023-06-29 scripts/07_umap_interactive.R
finishing manual coding

@@ -35,6 +35,14 @@ meta_df = open_dataset(here('data', '01_metadata')) |>
 source(here('R', 'phrases.R'))
 phrases_ar = phrases()
 
+## Build article URLs; used by spreadsheet
+make_url = function(article_id) {
+    start = str_trunc(article_id, 4, side = 'right', ellipsis = '')
+    if_else(start == '10.1',
+            str_c('https://dx.doi.org/', article_id), 
+            str_c('https://www.unz.com/print/MankindQuarterly/Contents/?Period=', start))
+}
+
 ## Workhorse function, generates visualization for a 
 ## single model (vocab x k)
 process_k = function(k, 
@@ -82,10 +90,12 @@ process_k = function(k,
             mutate(authors = map_chr(authors, 
                                      ~ {. |> 
                                              pull(author) |> 
-                                             str_c(collapse = '; ')})) |> 
+                                             str_c(collapse = '; ')}), 
+                   url = make_url(article_id)) |> 
             select(topic, gamma, article_id, 
                    container.title, year, 
-                   title, phrases, authors) |> 
+                   authors, 
+                   title, phrases, url) |> 
             group_split(topic) |> 
             walk2(spreadsheet, 
                   ~ write_xlsx(.x, 

### 2023-06-29 scripts/10_interrater_reliability.R

@@ -1,0 +1,64 @@
+library(tidyverse)
+library(readxl)
+library(irr)
+library(here)
+
+data_dir = here('data')
+out_dir = here('out')
+
+data_el = read_xlsx(here(data_dir, '07_md_40_V24 -EL.xlsx')) |> 
+    filter(!is.na(race_science_discourse)) |> 
+    rename(race_discourse_el = race_science_discourse, 
+           notes_el = notes) |> 
+    mutate(race_discourse_el = race_discourse_el == 'y')
+data_djh = read_xlsx(here(data_dir, '07_md_40_V24-DJH.xlsx')) |> 
+    filter(!is.na(race_science_discourse)) |> 
+    rename(race_discourse_djh = race_science_discourse) |> 
+    mutate(race_discourse_djh = race_discourse_djh == 'y')
+
+# anti_join(data_djh, data_el, by = 'article_id') |> 
+#     pull(article_id)
+
+data = inner_join(data_el, data_djh, 
+                  by = c('topic', 'gamma', 'article_id', 
+                         'container.title', 'year', 'authors', 'title', 
+                         'phrases', 'url'))
+
+## documents coded
+nrow(data)
+
+## quick crosstabs
+count(data_djh, race_discourse_djh) |> 
+    mutate(share = n / sum(n))
+count(data_el, race_discourse_el) |> 
+    mutate(share = n / sum(n))
+data |> 
+    mutate(idx = row_number()) |> 
+    select(idx, starts_with('race_discourse')) |> 
+    pivot_longer(starts_with('race_discourse'), 
+                 names_to = 'rater', 
+                 values_to = 'race_discourse') |> 
+    ggplot(aes(idx, rater, fill = race_discourse)) +
+    geom_tile()
+## "false positive" rate is 8.3%; 5% non-consensus
+count(data, race_discourse_djh, race_discourse_el) |> 
+    mutate(share = n / sum(n))
+
+## 95% agreement
+agree(data[,c('race_discourse_djh','race_discourse_el')])
+## kappa 0.74
+kappa2(data[,c('race_discourse_djh','race_discourse_el')], 
+       "unweighted")
+## Bennett, Alpert and Goldstein's S
+## 0.90
+(2*agree(data[,c('race_discourse_djh','race_discourse_el')])$value/100 - 1) / (2-1)
+
+
+data |> 
+    filter(race_discourse_djh != race_discourse_el) |> #view()
[…]

### 2023-06-29 paper/paper.Rmd
working on paper

@@
 | Thomas J. Bouchard, Jr. | psychology | |
-| Brunetto Chiarelli | anthropology? | |
+| Brunetto Chiarelli | anthropology | |
 | Hans Eysenck | psychology | |
@@
 
-![Journals publishing 6 or more Pioneer-funded researchers, WoS author search results *[date coverage]*](img/wos_results.png){width="4.76in" height="2.6in"}
+![Journals publishing 6 or more Pioneer-funded researchers, Web of Science author search results *[date coverage]*](img/wos_results.png){width="4.76in" height="2.6in"}
 
-Figure *[fig]* shows that, while MQ is on this list of "Pioneer-publishing" journals, a number of mainstream journals are more prominent: *Personality and Individual Differences* (PID), *Intelligence* (I), *Behavior Genetics* (BG), and *Psychological Reports* (PR).
+Figure *[fig]* shows that, while MQ is among the "Pioneer-publishing" journals, a number of mainstream journals are more prominent: *Personality and Individual Differences* (PID), *Intelligence* (I), *Behavior Genetics* (BG), and *Psychological Reports* (PR).
 In addition, only psychologist Richard Lynn appears to have published heavily in MQ.
@@
 
-PID was founded in 1980, with Eysenck as editor-in-chief and an editorial board including Jensen and Lynn.
+*Personality and Individual Differences* (PID) was founded in 1980, with Eysenck as editor-in-chief and an editorial board including Jensen and Lynn.
 In the inaugural editorial, Eysenck identified "studies of the genetic determinants of individual differences in the areas of personality and intelligence" as one of the journal's eight major areas of interest.
@@
 PID was first published by Pergamon Press, a mainstream academic press, and today is published by Elsevier.
-According to the Scimago Journal Rankings, based on Scopus citation data, in 2022 PID had an H-index of 193, #29 in Psychology, and a citation index of 5.27, #24 in Psychology.
+According to the Scimago Journal Rankings, based on Scopus citation data, in 2022 PID had an H-index of 193, #29 in Psychology (Miscellaneous), and a 2-year citation index (impact factor) of 5.27, #24 in Psychology (Miscellaneous).
 
-*[similar para on I]*
-
-*[segue to topic model analysis]*
+*Intelligence* (I) was founded in 1977, with psychologist Douglas Detterman as editor-in-chief from the founding until 2016.
+Lloyd Humphreys was on the editorial board starting from 1977; by 1990 he had been joined by Jensen and Philip Vernon.
+Richard Lynn joined the editorial board sometime between 1998 and 2002.
+(Archive copies of the I editorial board page are not available from the journal's website from 1999 through 2001.)  I has been criticized for including Lynn and Gerhard Meisenberg on its editorial board until July-August 2018 \cite{SainiSuperiorReturnRace2019}.
+Meisenberg is also connected with the Pioneer Fund *[director?]*, and was editor-in-chief of MQ in 2015-18 *[check these]*.
+I is published by Elsevier.
+According to the Scimago Journal Rankings, based on Scopus citation data, in 2022 I had an H-index of 103, #32 in Experimental and Cognitive Psychology, and a 2-year citation index (impact factor) of 3.05, #31 in Experimental and Cognitive Psychology.
 
 # Topic model analysis
+
+The fact that Pioneer-funded researchers published heavily in two mainstream psychology journals does not tell us anything about the content of their publications or the findings of their research.
+We assembled a full-text corpus of articles published in MQ, PID, I, BG, PR, and Behavior and Brain Sciences (BBS) between 1960-2010 and used topic modeling to identify topics discussing race, intelligence, and both.
+*[tables: Ns for each article, topic model parameters; figs: articles by pub over time, topic heatmaps, Silge plots, race-intelligence-both grids]*
+
+We focus here on three topics identified in one of the 24 fitted topic models *[composite fig: topics over time + Silge plots]*.
+In this model, topic 07 is strongly associated with MQ:  MQ published dozens of articles in this topic each year, and no other journal ever published more than a handful, with most other journals publishing none in a given year.
+The Silge plot contains racial terms (`races`, `whites` and potentially `europe`, `africa`, and `india`) as well as `book`, likely reflecting the fact that MQ published a number of book reviews while the psychology journals did not.
+
+*[switch 05 with 22; or add together 05, 22, 35, 38?]*
+Topic 05 is strongly associated with I and PID in the same way.
+The Silge plot contains the names of two Pioneer-funded researchers, `jensen` and `vernon`, but no racial terms.
+(Three other topics only associated with mainstream intelligence journals and terms were also identified by this topic model.
+Topic 22 )  The Silge plot for topic 24, by contrast, suggests a topic focused squarely on race and intelligence, with multiple racial terms (`whites`, `blacks`, `racial_differences`, `races`) and the names of three prominent Pioneer-funded researchers, `jensen`, `lynn`, and `rushton`.
+Qualitative review of the top 121 papers in this topic (those with $gamma > 0.97$) confirmed that this topic has high precision for race science discourse, with *[x% (n/121)]* classified as race science discourse by the authors.
+
+In a given year, most papers in topic 24 are published in mainstream journals, not MQ.
+Jensen's "How Much Can We Boost IQ and Scholastic Achievement?" \cite{JensenHowMuchCan1969} is published in 1969 in Harvard Educational Review (not included in this study), and the time series suggests an increase in articles in topic 24 in both PS and MQ (the only two journals in our corpus that were active at the time).
+MQ shows another sharp increase in the late 1970s, around the time Lynn was brought on as assistant editor, *[confirming the claim that Lynn shifted MQ's focus to intelligence research]*.
+PID publishes multiple papers in topic 24 almost immediately after it is founded, with I showing a more gradual increase between the mid-1970s and mid-1990s.
+
+Notably, BG published very few articles in this topic.
+Outside of the field itself, behavior genetics is strongly associated with race science by both other academics and the general public *[cites/ev]*.
+Panofsky \cite{PanofskyMisbehavingScienceControversy2014} argues that, prior to Jensen's 1969 paper, behavior genetics emphasized the study of non-human animals and intentionally avoided associations with eugenics and public controversy more generally.
+In response to Jensen, critics such as Lewontin offered broad critiques of behavior genetics as such, and behavior geneticists in turn adopted a radical conception of academic freedom (without any sense of responsibility for the social implications of academic research) and a siege or wartime mentality, as demonstrated by Sandra Scarr's 1986 presidential address to the Behavior Genetics Association (BGA) \cite{ScarrThreeCheersBehavior1987}.
+Scarr's address coincided with the period between 1970-1990 when BG published articles in topic 24.
+
+However, Glayde Whitney's 1995 BGA presidential address was published in MQ rather than BG.
+According to Whitney, BG's editor refused to publish the address and the BGA Executive Committee (except for Whitney himself) voted to "issue an official statement of denouncement" of the address; though we were unable to find any such a statement, and Panofsky reports that "nothing official was done to Whitney."  In his address, Whitney criticized "the Marxist-Lysenkoist denial of genetics"; quoted Peter Brimelow — a white nationalist and later founder of the racist and xenophobic website VDARE — defining "racist" as "anyone who is winning an argument with a liberal"; and proposed as a "reasonable scientific hypothesis" that differences in murder rates between countries and cities was caused by racial genetic differences in intelligence, empathy, aggression, and impulsivity.
+
+The topic model analysis and the case of Whitney's address suggest that, by the 1990s, behavior genetics had begun to distance itself from race science.
+*[https://link.springer.com/search?use-oscar-shared-search=true&sortBy=newestFirst&dateTo=2010&sortOrder=newestFirst&facet-journal-id=10519&content-type=Article&query=race+intelligence&date=custom&dateFrom=&date-facet-mode=between&facet-start-year=1972&previous-start-year=1972&facet-end-year=2010&previous-end-year=2023]*
+
+A disciplinary history of behavior genetics published in 2010 \cite{McGueEndBehavioralGenetics2010} acknowledges the historical association with eugenics, but does not mention Jensen and other race science controversies since World War II.
+Instead postwar behavior genetics is portrayed as an endeavor to counterbalance the hegemony of the "Blank Slate," the supposedly widespread view that genetic differences play no role in explaining individual differences \cite{BatesonCorpseWearisomeDebate2002}.
+
+These findings for *Behavior Genetics* complicate at least the popular association of behavior genetics with race science, if not Panofsky's more nuanced analysis.
+
+*[for discussion: the possibility for white supremacist appropriation, ie, race science done by scientists who are not themselves scientific racists: CarlsonQuantifyingContextualizingImpact2020; probably some reviews of Paige Harden; O'Connor and Weatherall]*
 

### 2023-06-29 scripts/scratch_interrater_reliability.R

@@ -1,34 +1,0 @@
-library(tidyverse)
-library(readxl)
-library(irr)
-library(here)
-
-data_dir = here('data')
-
-data_el = read_xlsx(here(data_dir, '07_md_30_V19 -EL codes.xlsx')) |> 
-    filter(!is.na(race_discourse)) |> 
-    rename(race_discourse_el = race_discourse, 
-           notes_el = notes) |> 
-    mutate(race_discourse_el = race_discourse_el == 1L)
-data_djh = read_xlsx(here(data_dir, '07_md_30_v19-DJH.xlsx')) |> 
-    filter(!is.na(race_discourse)) |> 
-    rename(race_discourse_djh = race_discourse) |> 
-    mutate(race_discourse_djh = race_discourse_djh == 'y')
-
-data = inner_join(data_el, data_djh)
-
-## documents coded
-nrow(data)
-## quick crosstab
-## "false positive" rate is 15%; 25% non-consensus
-count(data, race_discourse_djh, race_discourse_el) |> 
-    mutate(share = n / sum(n))
-
-
-kappa2(data[,c('race_discourse_djh','race_discourse_el')], 
-       "unweighted")
-
-data2 = subset(data, !race_discourse_djh & !race_discourse_el)
-
-write.csv(data2, here(data_dir, "07_md_30_V19_false_positives.csv"), 
-          row.names = FALSE)

### 2023-06-29 paper/paper.Rmd
working on paper

@@
 ## The rise and fall of eugenics
+
+*[write this]*
 
@@
 
+![Silge plots and smoothed time series for three focal topics.
+Top row: top 15 terms by $\beta$ (term-topic distribution) for each topic.
+Bottom row: count of articles associated with each topic, by journal and year.
+Article counts use a threshold approach, with $\gamma$ (topic-document distribution) greater than 0.5.
+Thin lines give annual values, thick lines give 5-year running averages.
+Because *Behavior Genetics* and *Behavioral and Brain Sciences* are not prominent in any panel, neither is given a direct label.](img/focal_topics.png){width="4.76in" height="3.57in"}
+
 We focus here on three topics identified in one of the 24 fitted topic models *[composite fig: topics over time + Silge plots]*.
 In this model, topic 07 is strongly associated with MQ:  MQ published dozens of articles in this topic each year, and no other journal ever published more than a handful, with most other journals publishing none in a given year.
-The Silge plot contains racial terms (`races`, `whites` and potentially `europe`, `africa`, and `india`) as well as `book`, likely reflecting the fact that MQ published a number of book reviews while the psychology journals did not.
+The Silge plot contains racial terms (`races`, `whites`, `negroes` and potentially `europe`, `africa`, `india`, and `japan`) as well as `book`, likely reflecting the fact that MQ published a number of book reviews while the psychology journals did not.
 
-*[switch 05 with 22; or add together 05, 22, 35, 38?]*
-Topic 05 is strongly associated with I and PID in the same way.
-The Silge plot contains the names of two Pioneer-funded researchers, `jensen` and `vernon`, but no racial terms.
-(Three other topics only associated with mainstream intelligence journals and terms were also identified by this topic model.
-Topic 22 )  The Silge plot for topic 24, by contrast, suggests a topic focused squarely on race and intelligence, with multiple racial terms (`whites`, `blacks`, `racial_differences`, `races`) and the names of three prominent Pioneer-funded researchers, `jensen`, `lynn`, and `rushton`.
-Qualitative review of the top 121 papers in this topic (those with $gamma > 0.97$) confirmed that this topic has high precision for race science discourse, with *[x% (n/121)]* classified as race science discourse by the authors.
+Topic 22 is strongly associated with I and PID in the same way.
+The Silge plot does contain `jensen`, as well as a reference to Raymond Cattell, who played a major role in the development of factor analysis and intelligence testing also advocated eugenics, fascism, and Nazi race science\cite{MehlerBeyondismRaymondCattell1997}.
+However, the other authors named in this topic — John Horn and Peter Bentler — do not appear to have contributed to race science discourse.
+Instead this topic appears to identify "mainstream" (non-race science) intelligence research, specifically on factor analysis and the debate over whether intelligence in unidimensional or multidimensional.
+(Three other topics only associated with mainstream intelligence journals and terms were also identified by this topic model.)  This topic indicates that the model is not simply lumping race science research together with other intelligence research.
 
-In a given year, most papers in topic 24 are published in mainstream journals, not MQ.
-Jensen's "How Much Can We Boost IQ and Scholastic Achievement?" \cite{JensenHowMuchCan1969} is published in 1969 in Harvard Educational Review (not included in this study), and the time series suggests an increase in articles in topic 24 in both PS and MQ (the only two journals in our corpus that were active at the time).
+The Silge plot for topic 24, by contrast, suggests a topic focused squarely on race and intelligence, with multiple racial terms (`whites`, `blacks`, `racial_differences`, `races`, `race_differences`) and the names of three prominent Pioneer-funded researchers, `jensen`, `lynn`, and `rushton`.
+To confirm this impression from the Silge plot, we independently coded the top 121 papers in this topic (those with $gamma > 0.97$) as contributing to race science discourse or not.
+105 (87%) were coded as race science discourse by both authors; 10 (8.3%) were coded as not race science discourse by both; and discordant codes were given to the remaining 6 (5.0%) (95% agreement, Cohen's $\kappa = 0.74$).
+
+In almost all years, most papers in topic 24 were published in mainstream journals rather than MQ.
+Jensen's "How Much Can We Boost IQ and Scholastic Achievement?" \cite{JensenHowMuchCan1969} was published in 1969 in Harvard Educational Review (not included in this study), and the time series indicates that, around this time, there was an increase in articles in topic 24 in both PS and MQ (the only two journals in our corpus that were active at the time).
 MQ shows another sharp increase in the late 1970s, around the time Lynn was brought on as assistant editor, *[confirming the claim that Lynn shifted MQ's focus to intelligence research]*.
-PID publishes multiple papers in topic 24 almost immediately after it is founded, with I showing a more gradual increase between the mid-1970s and mid-1990s.
+PID published multiple papers in topic 24 almost immediately after it was founded, with I showing a more gradual increase between the mid-1970s and mid-1990s.
 
 Notably, BG published very few articles in this topic.
-Outside of the field itself, behavior genetics is strongly associated with race science by both other academics and the general public *[cites/ev]*.
+Outside of the field itself, behavior genetics is strongly associated with race science by both other academics and the general public.
 Panofsky \cite{PanofskyMisbehavingScienceControversy2014} argues that, prior to Jensen's 1969 paper, behavior genetics emphasized the study of non-human animals and intentionally avoided associations with eugenics and public controversy more generally.
 In response to Jensen, critics such as Lewontin offered broad critiques of behavior genetics as such, and behavior geneticists in turn adopted a radical conception of academic freedom (without any sense of responsibility for the social implications of academic research) and a siege or wartime mentality, as demonstrated by Sandra Scarr's 1986 presidential address to the Behavior Genetics Association (BGA) \cite{ScarrThreeCheersBehavior1987}.
-Scarr's address coincided with the period between 1970-1990 when BG published articles in topic 24.
+Scarr's address coincided with the period between 1970-1990 when BG published articles in topic 24 (including Scarr's address itself).
 
@@
 
-The topic model analysis and the case of Whitney's address suggest that, by the 1990s, behavior genetics had begun to distance itself from race science.
+The topic model analysis and the case of Whitney's address suggest that, by the 1990s, behavior genetics had at least somewhat distanced itself from race science.
 *[https://link.springer.com/search?use-oscar-shared-search=true&sortBy=newestFirst&dateTo=2010&sortOrder=newestFirst&facet-journal-id=10519&content-type=Article&query=race+intelligence&date=custom&dateFrom=&date-facet-mode=between&facet-start-year=1972&previous-start-year=1972&facet-end-year=2010&previous-end-year=2023]*
@@
 These findings for *Behavior Genetics* complicate at least the popular association of behavior genetics with race science, if not Panofsky's more nuanced analysis.
+
+# Discussion
+
+Our findings indicate that *Mankind Quarterly* (MQ) was less important as an echo chamber for late 20th century race science than mainstream psychology journals, especially *Intelligence* (I) and *Personality and Individual Differences* (PID).
+Indeed, MQ published a very different kind of race science from the race-and-intelligence research published in I and PID, even after *[Lynn joined the MQ editorial board]*.
+
+We stress that we are not identifying any particular scientist as a racist based on the topic model results.
+Scientific research can be utilized to promote scientific racism — and thus count as race science — independently of the views and attitudes of the scientist who conducted that research.\cite{TaberyWhyStudyingGenetics2015, CarlsonQuantifyingContextualizingImpact2020, HennWhyDNANo2021}  But this does not mean that scientists are not responsible for the downstream social effects of their
+research, even when these effects involve deliberate misinterpretation.*[Douglas, Kitcher, Kourany, Havstad; Block and Dworkin?]*  Like other citizens, scientists are responsible for mitigating the reasonably foreseen harms that result from their actions.
+And we now have decades of empirical evidence that research on the genetics of intelligence — even research that does not itself touch on race — can and will be used to promote scientific racism.\cite{MeyerWrestlingSocialBehavioral2023}
+
+*[engagement dilemma: technical critiques can legitimize]*
+
+*[this probably goes too far even for a discussion section]*
+
+Proponents of a strong conception of academic freedom often point to chapter 2 of philosopher John Stuart Mill's *On Liberty* \cite{MillLiberty1862}, "Of the Liberty of Thought and Discussion."  However, in chapter 1 of the same book, Mill states the Harm Principle as the overall thesis of his book: "The only end for which people are entitled, individually or collectively, to interfere with the liberty of action of any of their number is self-protection.
+The only purpose for which power can be rightfully exercised over any member of a civilized community, against his will, is to prevent harm to others." Even for Mill, free speech and free inquiry are appropriately restrained when the ideas in question will harm others.
+
+More recently, philosopher Shannon Dea \cite{DeaEvolvingSocialPurpose2021} has argued that academic freedom is importantly distinct from freedom of speech.
+Whereas freedom of speech is universal, enjoyed by all residents of countries where speech is legally protected, academic freedom is more exclusive, enjoyed only by college and university faculty.
+(Though, as Dea points out, non-tenure-track faculty often do not have effective protections for academic freedom.)  Freedom of speech primarily serves to protect the interests of the individual speaker: that you will not be punished by the state for stating unpopular opinions.
+But academic freedom is "other-directed," primarily serving to benefit the common good of society as a whole by enabling the pursuit of truth.
+
+Advocates of research on genetics and intelligence — both anti-racist liberals and race scientists — claim this work will lead to social benefits.
+This reflects a recognition that the purpose of academic freedom is to promote the common good, rather than merely satisfy an individual's curiosity.
+But the promised benefits are either vague or do not actually require understanding the genetics of intelligence.\cite{KaplanGaltonQuincunxProbabilistic2021, CoopLotteryLuckLegacy2022}
+
+Using agent-based models, Weatherall et al.\cite{WeatherallHowBeatScience2018} show that a strategy of "selective sharing" — promoting favorable results from independent research — can be effective at misleading a target audience.
 
@@
 
+*[drop]*
 We also adapted the "discursive space" analysis proposed by *[Hicks 2021?]*.
@@
 ### Topic quality assessment
+*[rewrite]*
 

### 2023-06-29 paper/paper.tex
working on paper

@@
 \begin{sciabstract}
-Enter the text of your abstract here.
+\emph{{[}structured{]}}
 \end{sciabstract}
@@
 eugenics}\label{the-rise-and-fall-of-eugenics}}
+
+\emph{{[}write this{]}}
 
@@
 
+\emph{{[}for xref:
+\url{https://bookdown.org/yihui/bookdown/a-single-document.html}{]}}
+
 We identified 16 researchers who had received funding from Pioneer; 14
@@
 Thomas J. Bouchard, Jr. & psychology & \\
-Brunetto Chiarelli & anthropology? & \\
+Brunetto Chiarelli & anthropology & \\
 Hans Eysenck & psychology & \\
@@
 \includegraphics[width=4.76in,height=2.6in]{img/wos_results.png}
-\caption{Journals publishing 6 or more Pioneer-funded researchers, WoS
-author search results \emph{{[}date coverage{]}}}
+\caption{Journals publishing 6 or more Pioneer-funded researchers, Web
+of Science author search results \emph{{[}date coverage{]}}}
 \end{figure}
 
-Figure \emph{{[}fig{]}} shows that, while MQ is on this list of
+Figure \emph{{[}fig{]}} shows that, while MQ is among the
 ``Pioneer-publishing'' journals, a number of mainstream journals are
@@
 \emph{{[}cite{]}}. By contrast, a number of Pioneer-funded researchers
-have published in PID, I, and to a lesser degree BG: Eysenck, Jensen,
-Rushton, Vernon, and also Lynn.
+have published in PID, I, and to a lesser degree BG: Bouchard, Eysenck,
+Jensen, Rushton, Vernon, and also Lynn.
 
-PID was founded in 1980, with Eysenck as editor-in-chief and an
-editorial board including Jensen and Lynn. In the inaugural editorial,
-Eysenck identified ``studies of the genetic determinants of individual
-differences in the areas of personality and intelligence'' as one of the
-journal's eight major areas of interest. Eysenck remained
-editor-in-chief until his death in 1997. In 2005 the editorial board
-still included Jensen and Lynn. PID was first published by Pergamon
-Press, a mainstream academic press, and today is published by Elsevier.
+\emph{Personality and Individual Differences} (PID) was founded in 1980,
+with Eysenck as editor-in-chief and an editorial board including Jensen
+and Lynn. In the inaugural editorial, Eysenck identified ``studies of
+the genetic determinants of individual differences in the areas of
+personality and intelligence'' as one of the journal's eight major areas
+of interest. Eysenck remained editor-in-chief until his death in 1997.
+In 2005 the editorial board still included Jensen and Lynn. PID was
+first published by Pergamon Press, a mainstream academic press, and
+today is published by Elsevier. According to the Scimago Journal
+Rankings, based on Scopus citation data, in 2022 PID had an H-index of
+193, \#29 in Psychology (Miscellaneous), and a 2-year citation index
+(impact factor) of 5.27, \#24 in Psychology (Miscellaneous).
+
+\emph{Intelligence} (I) was founded in 1977, with psychologist Douglas
+Detterman as editor-in-chief from the founding until 2016. Lloyd
+Humphreys was on the editorial board starting from 1977; by 1990 he had
+been joined by Jensen and Philip Vernon. Richard Lynn joined the
+editorial board sometime between 1998 and 2002. (Archive copies of the I
+editorial board page are not available from the journal's website from
+1999 through 2001.) I has been criticized for including Lynn and Gerhard
+Meisenberg on its editorial board until July-August 2018
+\cite{SainiSuperiorReturnRace2019}. Meisenberg is also connected with
+the Pioneer Fund \emph{{[}director?{]}}, and was editor-in-chief of MQ
+in 2015-18 \emph{{[}check these{]}}. I is published by Elsevier.
 According to the Scimago Journal Rankings, based on Scopus citation
-data, in 2022 PID had an H-index of 193, \#29 in Psychology, and a
-citation index of 5.27, \#24 in Psychology.
-
-\emph{{[}similar para on I{]}}
-
-\emph{{[}segue to topic model analysis{]}}
+data, in 2022 I had an H-index of 103, \#32 in Experimental and
+Cognitive Psychology, and a 2-year citation index (impact factor) of
+3.05, \#31 in Experimental and Cognitive Psychology.
 
@@
 \section*{Topic model analysis}\label{topic-model-analysis}}
+
+The fact that Pioneer-funded researchers published heavily in two
+mainstream psychology journals does not tell us anything about the
+content of their publications or the findings of their research. We
+assembled a full-text corpus of articles published in MQ, PID, I, BG,
+PR, and Behavior and Brain Sciences (BBS) between 1960-2010 and used
+topic modeling to identify topics discussing race, intelligence, and
+both. \emph{{[}tables: Ns for each article, topic model parameters;
+figs: articles by pub over time, topic heatmaps, Silge plots,
+race-intelligence-both grids{]}}
+
+\begin{figure}
+\centering
+\includegraphics[width=4.76in,height=3.57in]{img/focal_topics.png}
+\caption{Silge plots and smoothed time series for three focal topics.
+Top row: top 15 terms by \(\beta\) (term-topic distribution) for each
+topic. Bottom row: count of articles associated with each topic, by
+journal and year. Article counts use a threshold approach, with
+\(\gamma\) (topic-document distribution) greater than 0.5. Thin lines
+give annual values, thick lines give 5-year running averages. Because
+\emph{Behavior Genetics} and \emph{Behavioral and Brain Sciences} are
+not prominent in any panel, neither is given a direct label.}
+\end{figure}
+
+We focus here on three topics identified in one of the 24 fitted topic
+models \emph{{[}composite fig: topics over time + Silge plots{]}}. In
+this model, topic 07 is strongly associated with MQ: MQ published dozens
+of articles in this topic each year, and no other journal ever published
+more than a handful, with most other journals publishing none in a given
+year. The Silge plot contains racial terms (\texttt{races},
+\texttt{whites}, \texttt{negroes} and potentially \texttt{europe},
+\texttt{africa}, \texttt{india}, and \texttt{japan}) as well as
+\texttt{book}, likely reflecting the fact that MQ published a number of
+book reviews while the psychology journals did not.
+
+Topic 22 is strongly associated with I and PID in the same way. The
+Silge plot does contain \texttt{jensen}, as well as a reference to
+Raymond Cattell, who played a major role in the development of factor
+analysis and intelligence testing also advocated eugenics, fascism, and
+Nazi race science\cite{MehlerBeyondismRaymondCattell1997}. However, the
+other authors named in this topic --- John Horn and Peter Bentler --- do
+not appear to have contributed to race science discourse. Instead this
+topic appears to identify ``mainstream'' (non-race science) intelligence
+research, specifically on factor analysis and the debate over whether
+intelligence in unidimensional or multidimensional. (Three other topics
+only associated with mainstream intelligence journals and terms were
+also identified by this topic model.) This topic indicates that the
+model is not simply lumping race science research together with other
+intelligence research.
+
+The Silge plot for topic 24, by contrast, suggests a topic focused
+squarely on race and intelligence, with multiple racial terms
+(\texttt{whites}, \texttt{blacks}, \texttt{racial\_differences},
+\texttt{races}, \texttt{race\_differences}) and the names of three
+prominent Pioneer-funded researchers, \texttt{jensen}, \texttt{lynn},
+and \texttt{rushton}. To confirm this impression from the Silge plot, we
+independently coded the top 121 papers in this topic (those with
+\(gamma > 0.97\)) as contributing to race science discourse or not. 105
+(87\%) were coded as race science discourse by both authors; 10 (8.3\%)
+were coded as not race science discourse by both; and discordant codes
+were given to the remaining 6 (5.0\%) (95\% agreement, Cohen's
+\(\kappa = 0.74\)).
+
+In almost all years, most papers in topic 24 were published in
+mainstream journals rather than MQ. Jensen's ``How Much Can We Boost IQ
+and Scholastic Achievement?'' \cite{JensenHowMuchCan1969} was published
+in 1969 in Harvard Educational Review (not included in this study), and
+the time series indicates that, around this time, there was an increase
+in articles in topic 24 in both PS and MQ (the only two journals in our
+corpus that were active at the time). MQ shows another sharp increase in
+the late 1970s, around the time Lynn was brought on as assistant editor,
+\emph{{[}confirming the claim that Lynn shifted MQ's focus to
+intelligence research{]}}. PID published multiple papers in topic 24
+almost immediately after it was founded, with I showing a more gradual
+increase between the mid-1970s and mid-1990s.
+
+Notably, BG published very few articles in this topic. Outside of the
+field itself, behavior genetics is strongly associated with race science
+by both other academics and the general public. Panofsky
+\cite{PanofskyMisbehavingScienceControversy2014} argues that, prior to
+Jensen's 1969 paper, behavior genetics emphasized the study of non-human
+animals and intentionally avoided associations with eugenics and public
+controversy more generally. In response to Jensen, critics such as
+Lewontin offered broad critiques of behavior genetics as such, and
+behavior geneticists in turn adopted a radical conception of academic
+freedom (without any sense of responsibility for the social implications
+of academic research) and a siege or wartime mentality, as demonstrated
+by Sandra Scarr's 1986 presidential address to the Behavior Genetics
+Association (BGA) \cite{ScarrThreeCheersBehavior1987}. Scarr's address
+coincided with the period between 1970-1990 when BG published articles
+in topic 24 (including Scarr's address itself).
+
+However, Glayde Whitney's 1995 BGA presidential address was published in
+MQ rather than BG. According to Whitney, BG's editor refused to publish
+the address and the BGA Executive Committee (except for Whitney himself)
+voted to ``issue an official statement of denouncement'' of the address;
+though we were unable to find any such a statement, and Panofsky reports
+that ``nothing official was done to Whitney.'' In his address, Whitney
+criticized ``the Marxist-Lysenkoist denial of genetics''; quoted Peter
+Brimelow --- a white nationalist and later founder of the racist and
+xenophobic website VDARE --- defining ``racist'' as ``anyone who is
+winning an argument with a liberal''; and proposed as a ``reasonable
+scientific hypothesis'' that differences in murder rates between
+countries and cities was caused by racial genetic differences in
+intelligence, empathy, aggression, and impulsivity.
+
+The topic model analysis and the case of Whitney's address suggest that,
+by the 1990s, behavior genetics had at least somewhat distanced itself
+from race science.
+\emph{{[}\url{https://link.springer.com/search?use-oscar-shared-search=true\&sortBy=newestFirst\&dateTo=2010\&sortOrder=newestFirst\&facet-journal-id=10519\&content-type=Article\&query=race+intelligence\&date=custom\&dateFrom=\&date-facet-mode=between\&facet-start-year=1972\&previous-start-year=1972\&facet-end-year=2010\&previous-end-year=2023}{]}}
+
+A disciplinary history of behavior genetics published in 2010
+\cite{McGueEndBehavioralGenetics2010} acknowledges the historical
+association with eugenics, but does not mention Jensen and other race
+science controversies since World War II. Instead postwar behavior
+genetics is portrayed as an endeavor to counterbalance the hegemony of
+the ``Blank Slate,'' the supposedly widespread view that genetic
+differences play no role in explaining individual differences
+\cite{BatesonCorpseWearisomeDebate2002}.
+
+These findings for \emph{Behavior Genetics} complicate at least the
+popular association of behavior genetics with race science, if not
+Panofsky's more nuanced analysis.
+
+\hypertarget{discussion}{%
+\section*{Discussion}\label{discussion}}
+
+Our findings indicate that \emph{Mankind Quarterly} (MQ) was less
+important as an echo chamber for late 20th century race science than
+mainstream psychology journals, especially \emph{Intelligence} (I) and
+\emph{Personality and Individual Differences} (PID). Indeed, MQ
+published a very different kind of race science from the
+race-and-intelligence research published in I and PID, even after
+\emph{{[}Lynn joined the MQ editorial board{]}}.
+
+We stress that we are not identifying any particular scientist as a
+racist based on the topic model results. Scientific research can be
+utilized to promote scientific racism --- and thus count as race science
+--- independently of the views and attitudes of the scientist who
+conducted that
+research.\cite{TaberyWhyStudyingGenetics2015, CarlsonQuantifyingContextualizingImpact2020, HennWhyDNANo2021}
+But this does not mean that scientists are not responsible for the
+downstream social effects of their research, even when these effects
+involve deliberate misinterpretation.\emph{{[}Douglas, Kitcher, Kourany,
+Havstad; Block and Dworkin?{]}} Like other citizens, scientists are
+responsible for mitigating the reasonably foreseen harms that result
+from their actions. And we now have decades of empirical evidence that
+research on the genetics of intelligence --- even research that does not
+itself touch on race --- can and will be used to promote scientific
+racism.\cite{MeyerWrestlingSocialBehavioral2023}
+
+\emph{{[}engagement dilemma: technical critiques can legitimize{]}}
+
+\emph{{[}this probably goes too far even for a discussion section{]}}
+
+Proponents of a strong conception of academic freedom often point to
+chapter 2 of philosopher John Stuart Mill's \emph{On Liberty}
+\cite{MillLiberty1862}, ``Of the Liberty of Thought and Discussion.''
+However, in chapter 1 of the same book, Mill states the Harm Principle
+as the overall thesis of his book: ``The only end for which people are
+entitled, individually or collectively, to interfere with the liberty of
+action of any of their number is self-protection. The only purpose for
+which power can be rightfully exercised over any member of a civilized
+community, against his will, is to prevent harm to others.'' Even for
+Mill, free speech and free inquiry are appropriately restrained when the
+ideas in question will harm others.
+
+More recently, philosopher Shannon Dea
+\cite{DeaEvolvingSocialPurpose2021} has argued that academic freedom is
+importantly distinct from freedom of speech. Whereas freedom of speech
+is universal, enjoyed by all residents of countries where speech is
+legally protected, academic freedom is more exclusive, enjoyed only by
+college and university faculty. (Though, as Dea points out,
+non-tenure-track faculty often do not have effective protections for
+academic freedom.) Freedom of speech primarily serves to protect the
+interests of the individual speaker: that you will not be punished by
+the state for stating unpopular opinions. But academic freedom is
+``other-directed,'' primarily serving to benefit the common good of
+society as a whole by enabling the pursuit of truth.
+
+Advocates of research on genetics and intelligence --- both anti-racist
+liberals and race scientists --- claim this work will lead to social
+benefits. This reflects a recognition that the purpose of academic
+freedom is to promote the common good, rather than merely satisfy an
+individual's curiosity. But the promised benefits are either vague or do
+not actually require understanding the genetics of
+intelligence.\cite{KaplanGaltonQuincunxProbabilistic2021, CoopLotteryLuckLegacy2022}
+
+Using agent-based models, Weatherall et
+al.\cite{WeatherallHowBeatScience2018} show that a strategy of
+``selective sharing'' --- promoting favorable results from independent
+research --- can be effective at misleading a target audience.
+
+\emph{{[}for discussion: the possibility for white supremacist
+appropriation, ie, race science done by scientists who are not
+themselves scientific racists:
+CarlsonQuantifyingContextualizingImpact2020; probably some reviews of
+Paige Harden; O'Connor and Weatherall{]}}
 

### 2023-06-29 paper/supplement_paper.tex
working on paper

@@
 
-\hypertarget{data-preparation}{%
-\subsection*{Data preparation}\label{data-preparation}}
+\hypertarget{text-preparation}{%
+\subsection*{Text preparation}\label{text-preparation}}
 
@@
 
+\hypertarget{data-and-code-availability}{%
+\subsubsection*{Data and code
+availability}\label{data-and-code-availability}}
+
+Document metadata retrieved from Crossref does not appear to be covered
+by any copyright or other intellectual property restrictions. However,
+due to copyright restrictions, we are unable to make document fulltext
+or Web of Science search results publicly available. Code for the corpus
+assembly and text preparation steps described above is available by
+request.
+
+The public analysis repository,
+\url{https://github.com/dhicks/race-science} \emph{{[}DOI{]}}, includes
+document metadata and documentwise counts of NLP-extracted noun phrases
+(``document-term matrices''), along with code to reproduce the analysis
+of the following sections. \emph{{[}repo readme: requires R 4.1.2; rig
+(\url{https://github.com/r-lib/rig/}) for managing multiple R
+installations{]}}
+
 \hypertarget{vocabulary-selection}{%
-\subsubsection*{Vocabulary selection}\label{vocabulary-selection}}
+\subsection*{Vocabulary selection}\label{vocabulary-selection}}
 
@@
 
-We also adapted the ``discursive space'' analysis proposed by
-\emph{{[}Hicks 2021?{]}}. This analysis combines the Hellinger distance
-between topic-document distributions --- capturing how ``far apart'' two
-documents are --- with dimensional reduction techniques used for data
-visualization. This produces a two-dimensional visualization of
-individual documents, arranged so that documents with similar topic
-distributions are visually clustered and visually distinct from
-documents with different topic distributions. \emph{{[}availability{]}}
+\emph{{[}drop{]}} We also adapted the ``discursive space'' analysis
+proposed by \emph{{[}Hicks 2021?{]}}. This analysis combines the
+Hellinger distance between topic-document distributions --- capturing
+how ``far apart'' two documents are --- with dimensional reduction
+techniques used for data visualization. This produces a two-dimensional
+visualization of individual documents, arranged so that documents with
+similar topic distributions are visually clustered and visually distinct
+from documents with different topic distributions.
+\emph{{[}availability{]}}
 
@@
 assessment}\label{topic-quality-assessment}}
+
+\emph{{[}rewrite{]}}
 

### 2023-06-29 scripts/08_topics_over_time.R

@@ -173,32 +173,54 @@ md_tmf = read_rds(model_files['md'])
 md_exp = read_exp(exp_files['md'])
 
 ## Similar trends for low, medium, and high threshold
-count_plot(md_tmf, 'medium', md_exp, 40, c('V07', 'V24', 'V05'), .25)
-plot_50 = count_plot(md_tmf, 'medium', md_exp, 40, c('V07', 'V24', 'V05'), .50)
+count_plot(md_tmf, 'medium', md_exp, 40, c('V07', 'V24', 'V22'), .25)
+plot_50 = count_plot(md_tmf, 'medium', md_exp, 40, c('V07', 'V24', 'V22'), .50)
 plot_50
-count_plot(md_tmf, 'medium', md_exp, 40, c('V07', 'V24', 'V05'), .80)
+count_plot(md_tmf, 'medium', md_exp, 40, c('V07', 'V24', 'V22'), .80)
 
 ## Direct labelled version of plot_50
 jr_labels = tribble(
     ~ topic, ~ container.title, ~ label, ~ date, ~ y,
-    'V05', 'Personality and Individual Differences', 'Person. & Indiv. Diff.', 1970, 10,
-    'V05', 'Intelligence', 'Intelligence', 1973, 7,
+    'V22', 'Personality and Individual Differences', 'Person. & Indiv. Diff.', 1985, 13,
+    'V22', 'Intelligence', 'Intelligence', 1985, 10,
     'V07', 'Mankind Quarterly', 'Mankind Quarterly', 1995, 30, 
     'V24', 'Psychological Reports', 'Psychological Reports', 1997, 8
 )
 
-plot_50 +
+time_gg = plot_50 +
     geom_label(aes(x = date, y = y, label = label, 
                    fill = container.title), 
                color = 'black', size = 3,
                data = jr_labels)
+time_gg
+
+# ggsave(here(out_dir, '08_focal_topics.png'), 
+#        plot = time_gg,
+#        height = 5, width = 9, bg = 'white')
+
+silge_gg = tidy(md_tmf, matrix = 'beta', k = 40) |> 
+    filter(topic %in% c('V07', 'V22', 'V24')) |> 
+    group_by(topic) |> 
+    top_n(15, beta) |> 
+    arrange(topic, desc(beta)) |> 
+    mutate(token = reorder_within(token, beta, topic)) |> 
+    ggplot(aes(token, beta)) +
+    geom_point() +
[…]

### 2023-07-05 scripts/07_umap_interactive.R
remove UMAP from workflow and add IRR script

@@ -1,3 +1,4 @@
+## Formerly, this script generated two-dimensional UMAP embeddings, for use in interactive and static visualizations.  These visualizations were dropped from the analysis pipeline.  Because calculating these embeddings is computationally intensive, the relevant lines (last 3 lines at the end of the script) are commented out.  This script is still used in the workflow to generate a quality check spreadsheet for the focal topic.  
 renv::load(here::here())
 library(tidyverse)
 theme_set(theme_minimal())
@@ -211,8 +212,8 @@ make_umap = function(vocab,
 
 # debugonce(process_k)
 ## Generate a single spreadsheet for topic quality check
-# make_umap('md', 40, spreadsheet = 'V24', visualization = FALSE)
+make_umap('md', 40, spreadsheet = 'V24', visualization = FALSE)
 ## Generate visualizations for all models of interest
-make_umap('md', c(20, 30, 40, 50, 60, 70), seed = seed, force = force)
-make_umap('sm', c(20, 30, 40, 50, 60, 70), seed = seed, force = force)
-make_umap('lg', c(20, 30, 40, 50, 60, 70), seed = seed, force = force)
+# make_umap('md', c(20, 30, 40, 50, 60, 70), seed = seed, force = force)
+# make_umap('sm', c(20, 30, 40, 50, 60, 70), seed = seed, force = force)
+# make_umap('lg', c(20, 30, 40, 50, 60, 70), seed = seed, force = force)

### 2023-07-05 scripts/08_topics_over_time.R

@@ -6,7 +6,8 @@ theme_set(theme_bw())
 library(tmfast)
 library(here)
 library(arrow)
-library(patchwork)
+# library(patchwork)
+library(cowplot)
 library(glue)
 library(ggh4x)
 
@@ -173,10 +174,13 @@ md_tmf = read_rds(model_files['md'])
 md_exp = read_exp(exp_files['md'])
 
 ## Similar trends for low, medium, and high threshold
-count_plot(md_tmf, 'medium', md_exp, 40, c('V07', 'V24', 'V22'), .25)
 plot_50 = count_plot(md_tmf, 'medium', md_exp, 40, c('V07', 'V24', 'V22'), .50)
-plot_50
-count_plot(md_tmf, 'medium', md_exp, 40, c('V07', 'V24', 'V22'), .80)
+if (interactive()) {
+    count_plot(md_tmf, 'medium', md_exp, 40, c('V07', 'V24', 'V22'), .25)
+    # plot_50 = count_plot(md_tmf, 'medium', md_exp, 40, c('V07', 'V24', 'V22'), .50)
+    plot_50
+    count_plot(md_tmf, 'medium', md_exp, 40, c('V07', 'V24', 'V22'), .80)
+}
 
 ## Direct labelled version of plot_50
 jr_labels = tribble(
@@ -192,7 +196,7 @@ time_gg = plot_50 +
                    fill = container.title), 
                color = 'black', size = 3,
                data = jr_labels)
-time_gg
+if (interactive()) time_gg
 
 # ggsave(here(out_dir, '08_focal_topics.png'), 
 #        plot = time_gg,
@@ -211,19 +215,23 @@ silge_gg = tidy(md_tmf, matrix = 'beta', k = 40) |>
     scale_x_reordered() + 
     labs(y = 'β') +
     facet_wrap(vars(topic), scales = 'free')
-silge_gg
+if (interactive()) silge_gg
 
 # silge_gg / time_gg +
 #     plot_layout()
 
-cowplot::plot_grid(silge_gg, time_gg, ncol = 1, align = 'v', 
+focal_topics_gg = plot_grid(silge_gg, time_gg, ncol = 1, align = 'v', 
                    rel_heights = c(.7, 1))
+if (interactive()) focal_topics_gg
 
 ggsave(here(out_dir, '08_focal_topics.png'), 
+       plot = focal_topics_gg,
        height = 3*3, width = 4*3, bg = 'white')
 
[…]

### 2023-07-05 scripts/09_umap_static.R

@@ -1,3 +1,4 @@
+## This script generates static visualizations using the UMAP embeddings from script 07.  Because these visualizations are no longer used in the analysis workflow, this script is not triggered by `make all`.  
 renv::load(here::here())
 library(tidyverse)
 theme_set(theme_minimal())

### 2023-07-05 scripts/10_interrater_reliability.R

@@ -1,3 +1,4 @@
+renv::load(here::here())
 library(tidyverse)
 library(readxl)
 library(irr)
@@ -6,7 +7,7 @@ library(here)
 data_dir = here('data')
 out_dir = here('out')
 
-data_el = read_xlsx(here(data_dir, '07_md_40_V24 -EL.xlsx')) |> 
+data_el = read_xlsx(here(data_dir, '07_md_40_V24-EL.xlsx')) |> 
     filter(!is.na(race_science_discourse)) |> 
     rename(race_discourse_el = race_science_discourse, 
            notes_el = notes) |> 
@@ -51,7 +52,7 @@ kappa2(data[,c('race_discourse_djh','race_discourse_el')],
        "unweighted")
 ## Bennett, Alpert and Goldstein's S
 ## 0.90
-(2*agree(data[,c('race_discourse_djh','race_discourse_el')])$value/100 - 1) / (2-1)
+(2 * agree(data[,c('race_discourse_djh','race_discourse_el')])$value/100 - 1) / (2-1)
 
 
 data |> 

### 2023-07-07 paper/paper.Rmd
paper working notes

@@
 *[incorporate def'ns]*
+
 scientific racism 
@@
  ~ treating race science as a legitimate area of scientific research; this includes methodological critiques of race science and empirical tests that falsify race science hypotheses
+
+- pre-20th century history of race concepts and race science
+- expansion and contraction of racialized terms over time
 
@@
 
-*[this probably goes too far even for a discussion section]*
+*[limitation: colorblind racism (Bonilla-Silva) creates coding ambiguities]*
+
+*[limitation: false negatives (race science papers that don't appear in V24) are possible if, eg, they use many terms related to twin studies of heritability and thereby appear in V03]*
+
+*[replace w/: WinstonWhyMainstreamResearch2020: race science as a community project of psychology(?).
+radical academic freedom is a tool used to defend race science PigliucciWhatAreWe2013, RosemanTroublesomeReflectionRacism2014, LarsenMoreProvocativeLess2020.
+an important part of how these mainstream journals function as race science echo chambers.]*
 
@@
 *[for discussion: the possibility for white supremacist appropriation, ie, race science done by scientists who are not themselves scientific racists: CarlsonQuantifyingContextualizingImpact2020; probably some reviews of Paige Harden; O'Connor and Weatherall]*
+
+*[what can science do? Posselt, /Equity in Science/]*
 

### 2023-07-10 paper/paper.Rmd
enable cross-references

@@
   # skeleton: <https://github.com/christopherkenny/rticles/tree/master/inst/rmarkdown/templates/science>
-  rticles::science_article:
+  # rticles::science_article:
+  #   keep_tex: true
+  #    move_figures: false
+  #    draft: false
+  #    number_sections: false
+  bookdown::pdf_book:
+    base_format: rticles::science_article
     keep_tex: true
@@
 
-*[for xref: https://bookdown.org/yihui/bookdown/a-single-document.html]*
-
 We identified 16 researchers who had received funding from Pioneer; 14 of these researchers had profiles in the Web of Science (WoS) author search *[table + WoS pub coucnt]*, allowing us identify 13 WoS-indexed journals that had published 6 or more of these authors.
+See Table \@ref(tab:researchers).
 
@@
 
-Table: Pioneer-funded researchers.  Either identified in *[cite]* or named on Pioneer's website, along with identified discipline and WoS author search result counts.
+Table: (\#tab:researchers) Pioneer-funded researchers.  Either identified in *[cite]* or named on Pioneer's website, along with identified discipline and WoS author search result counts.
 
-![Journals publishing 6 or more Pioneer-funded researchers, Web of Science author search results *[date coverage]*](img/wos_results.png){width="4.76in" height="2.6in"}
+![Journals publishing 6 or more Pioneer-funded researchers, Web of Science author search results *[date coverage]* (\#fig:wos)](img/wos_results.png){width="4.76in" height="2.6in"}
 
-Figure *[fig]* shows that, while MQ is among the "Pioneer-publishing" journals, a number of mainstream journals are more prominent: *Personality and Individual Differences* (PID), *Intelligence* (I), *Behavior Genetics* (BG), and *Psychological Reports* (PR).
+Figure \@ref(fig:wos) shows that, while MQ is among the "Pioneer-publishing" journals, a number of mainstream journals are more prominent: *Personality and Individual Differences* (PID), *Intelligence* (I), *Behavior Genetics* (BG), and *Psychological Reports* (PR).
 In addition, only psychologist Richard Lynn appears to have published heavily in MQ.
@@
 
-The topic model analysis and the case of Whitney's address suggest that, by the 1990s, behavior genetics had at least somewhat distanced itself from race science.
+The topic model analysis and the case of Whitney's address suggest that, by the 1990s, behavior genetics had at least somewhat distanced itself from race science, though perhaps without repudiating it.
 *[https://link.springer.com/search?use-oscar-shared-search=true&sortBy=newestFirst&dateTo=2010&sortOrder=newestFirst&facet-journal-id=10519&content-type=Article&query=race+intelligence&date=custom&dateFrom=&date-facet-mode=between&facet-start-year=1972&previous-start-year=1972&facet-end-year=2010&previous-end-year=2023]*
 
-A disciplinary history of behavior genetics published in 2010 \cite{McGueEndBehavioralGenetics2010} acknowledges the historical association with eugenics, but does not mention Jensen and other race science controversies since World War II.
+To check this interpretation, we ran an independent search for documents with the keywords "race" and "intelligence" published in BG from 1972 to 2020, using Springer's journal search website.
+This search returned 125 documents.
+33 were presentation abstracts from meetings of the Behavior Genetics Association; we did not examine these further.
+14 of the remaining documents were published since 2010 (after the scope of the primary study).
+Among these, one obituary and two papers celebrated the work of John Loehlin, a prominent race-and-intelligence researcher who had been Director of the American Eugenics Society from 1968-1972 *[http://link.springer.com/article/10.1007/s10519-020-10024-w; http://link.springer.com/article/10.1007/s10519-014-9687-1; http://link.springer.com/article/10.1007/s10519-014-9673-7]*.
+Another was a retrospective of the work of Lindon Eaves, a geneticist who made at least one notable contribution to debates on the Scarr-Rowe hypothesis (interactions between heritability, race, and class).
+None of these 14 documents reported any studies of racial differences.
+
+28 documents in this sample were published between 1990 and 2010.
+One was an obituary of Jerry Hirsch, a critic of hereditarianism in general and Jensen in particular *[http://link.springer.com/article/10.1007/s10519-008-9231-2]*, and another was an obituary of David Rowe *[http://link.springer.com/article/10.1023/A%3A1026122928868]*.
+Only 2 of these 28 documents reported race differences of any kind: one examining at interactions among race, sex, and heritability for adolescent BMI (Body Mass Index, used as a measure of overweight/obesity) *[http://link.springer.com/article/10.1023/A%3A1021619329904]*; and the other at interactions among race, family history of alcoholism, and visuospatial performance *[http://link.springer.com/article/10.1007/BF02197241]*.
+Finally, Philip Vernon — one of the Pioneer-funded scientists we identified earlier — published a critical review of a book that attempted to address hereditarianism, the Flynn effect, race-and-intelligence research, and some related issues.
+
+This supplemental review provides additional support for the interpretation that behavior genetics had distanced itself from race science by the 1990s.
+Notably, this review brought our attention to a disciplinary history of behavior genetics published in 2010 \cite{McGueEndBehavioralGenetics2010}, which acknowledges the historical association with eugenics but does not mention Jensen and other race science controversies since World War II.
 Instead postwar behavior genetics is portrayed as an endeavor to counterbalance the hegemony of the "Blank Slate," the supposedly widespread view that genetic differences play no role in explaining individual differences \cite{BatesonCorpseWearisomeDebate2002}.
-
-These findings for *Behavior Genetics* complicate at least the popular association of behavior genetics with race science, if not Panofsky's more nuanced analysis.
+And none of the three documents celebrating Loehlin noted his postwar involvement in the eugenics movement.
+None of the documents examined in this supplemental review explicitly repudiate race science.
+Behavior geneticists seem to typically adopt a posture of objectivity and technical rigor, downplaying the extent of the field's involvement in race science in the past.
 
@@
 
-*[limitation: colorblind racism (Bonilla-Silva) creates coding ambiguities]*
+*[limitation: colorblind racism (Bonilla-Silva; WinstonWhyMainstreamResearch2020) creates coding ambiguities]*
 

### 2023-07-10 paper/paper.tex
enable cross-references

@@
 \\
-\textsuperscript{*}Corresponding author: Daniel J. Hicks,
-\href{mailto:dhicks4@ucmerced.edu}{\nolinkurl{dhicks4@ucmerced.edu}}.
+\textsuperscript{*}Corresponding author: Daniel J. Hicks, \href{mailto:dhicks4@ucmerced.edu}{\nolinkurl{dhicks4@ucmerced.edu}}.
 }
@@
 
-\emph{{[}incorporate def'ns{]}} scientific racism \textasciitilde{}
-purporting to justify racial inequality and colonialism by appealing to
-the epistemic authority of science
+\emph{{[}incorporate def'ns{]}}
 
@@
 \tightlist
+\item[scientific racism]
+purporting to justify racial inequality and colonialism by appealing to the epistemic authority of science
 \item[race science]
@@
 \item[race science discourse]
-treating race science as a legitimate area of scientific research; this
-includes methodological critiques of race science and empirical tests
-that falsify race science hypotheses
+treating race science as a legitimate area of scientific research; this includes methodological critiques of race science and empirical tests that falsify race science hypotheses
 \end{description}
+
+\begin{itemize}
+\tightlist
+\item
+  pre-20th century history of race concepts and race science
+\item
+  expansion and contraction of racialized terms over time
+\end{itemize}
 
@@
 \hypertarget{the-rise-and-fall-of-eugenics}{%
-\subsection*{The rise and fall of
-eugenics}\label{the-rise-and-fall-of-eugenics}}
+\subsection*{The rise and fall of eugenics}\label{the-rise-and-fall-of-eugenics}}
 
@@
 \item
-  scientific racism is thoroughly entangled with eugenics, but still
-  conceptually distinct (Slobodian 2023)
+  scientific racism is thoroughly entangled with eugenics, but still conceptually distinct (Slobodian 2023)
 
@@
   \item
-    sterilization laws focused on categories of criminality and mental
-    ability, rather than race per se
+    sterilization laws focused on categories of criminality and mental ability, rather than race per se
   \item
-    contemporary defenders of eugenics at least sometimes claim to
-    oppose racism
+    contemporary defenders of eugenics at least sometimes claim to oppose racism
   \item
-    claims of innate racial differences in intelligence can be made
-    without dysgenic prophecies of a rapidly-expanding biological
-    underclass
+    claims of innate racial differences in intelligence can be made without dysgenic prophecies of a rapidly-expanding biological underclass
   \end{itemize}
@@
 \item
-  1926: NRC ``Committee on the Negro'' includes both Davenport and Boas:
-  ``neither was influential enough to veto his rival's participation''
-  (Barkan 1992, 113)
+  1926: NRC ``Committee on the Negro'' includes both Davenport and Boas: ``neither was influential enough to veto his rival's participation'' (Barkan 1992, 113)
 \item
-  1930s: \emph{Coming of Age in Samoa} (Mead 1928) popularizes cultural
-  anthropology, ``free{[}ing{]} anthropology from the shackles of
-  biology'' (Barkan 1992, 134)
+  1930s: \emph{Coming of Age in Samoa} (Mead 1928) popularizes cultural anthropology, ``free{[}ing{]} anthropology from the shackles of biology'' (Barkan 1992, 134)
 \item
-  1937: Pioneer Fund is founded to ``support academic research and the
-  '{[}sic{]}dissemination of information, into the `problem of heredity
-  and eugenics' and `the problems of race betterment'\,'' (Mehler 1989,
-  21, quoting Laughlin)
+  1937: Pioneer Fund is founded to ``support academic research and the '{[}sic{]}dissemination of information, into the `problem of heredity and eugenics' and `the problems of race betterment'\,'' (Mehler 1989, 21, quoting Laughlin)
 \item
@@
 \item
-  1950: ``For all practical social purposes `race' is not so much a
-  biological phenomenon as a social myth'' (Beaglehole et al. 1950;
-  though see Brattain 2007)
+  1950: ``For all practical social purposes `race' is not so much a biological phenomenon as a social myth'' (Beaglehole et al. 1950; though see Brattain 2007)
 \end{itemize}
@@
 \hypertarget{brown-v-board-and-mankind-quarterly}{%
-\subsection*{\texorpdfstring{\emph{Brown v Board} and \emph{Mankind
-Quarterly}}{Brown v Board and Mankind Quarterly}}\label{brown-v-board-and-mankind-quarterly}}
+\subsection*{\texorpdfstring{\emph{Brown v Board} and \emph{Mankind Quarterly}}{Brown v Board and Mankind Quarterly}}\label{brown-v-board-and-mankind-quarterly}}
 
-Histories of scientific racism in the second half of the twentieth
-century have often emphasized the parascholarly journal \emph{Mankind
-Quarterly} (MQ) \cite{
+Histories of scientific racism in the second half of the twentieth century have often emphasized the parascholarly journal \emph{Mankind Quarterly} (MQ) \cite{
 MehlerFoundationFascismNew1989,
@@
 SainiDraperMillionsPhilanthropic2022
-}, with or without its connections to PF. MQ was founded in 1960 by
-biologist R. Ruggles Gates (1882-1962), psychologist Henry Garrett
-(1894-1973), and non-academic anthropologist G. Robert Gayre
-(self-styled as ``Gayre of Gayre and Nigg''; 1907-1996). Gates'
-professional status was closely tied to eugenics and the explicit
-scientific racism of the 1920s, and he had been thoroughly marginalized
-in the wake of World War II \cite{WinstonScienceServiceFar1998}. In
-contrast, Garrett had been president of the American Psychological
-Association in 1946 and chair of Psychology at Columbia from 1941 to
-1955 \emph{{[}cite{]}}.
+}, with or without its connections to PF.
+MQ was founded in 1960 by biologist R.
+Ruggles Gates (1882-1962), psychologist Henry Garrett (1894-1973), and non-academic anthropologist G.
+Robert Gayre (self-styled as ``Gayre of Gayre and Nigg''; 1907-1996).
+Gates' professional status was closely tied to eugenics and the explicit scientific racism of the 1920s, and he had been thoroughly marginalized in the wake of World War II \cite{WinstonScienceServiceFar1998}.
+In contrast, Garrett had been president of the American Psychological Association in 1946 and chair of Psychology at Columbia from 1941 to 1955 \emph{{[}cite{]}}.
 
-In the landmark case \emph{Brown v Board of Education of Topeka} (1954),
-the US Supreme Court had banned \emph{de jure} educational segregation.
-The Court's decision relied on expert testimony from psychologists and
-education researchers; but the segregationists had also put forward
-their own experts, including Henry Garrett
-\cite{WinstonScienceServiceFar1998, JacksonScienceSegregationRace2005, SchafferScientificRacismAgain2007}.
-Shortly after \emph{Brown}, Garrett left Columbia, taking a visiting
-position at the University of Virginia, and becoming increasingly
-involved in segregationist, eugenicist, and even the neofascist Northern
-League.
+In the landmark case \emph{Brown v Board of Education of Topeka} (1954), the US Supreme Court had banned \emph{de jure} educational segregation.
+The Court's decision relied on expert testimony from psychologists and education researchers; but the segregationists had also put forward their own experts, including Henry Garrett \cite{WinstonScienceServiceFar1998, JacksonScienceSegregationRace2005, SchafferScientificRacismAgain2007}.
+Shortly after \emph{Brown}, Garrett left Columbia, taking a visiting position at the University of Virginia, and becoming increasingly involved in segregationist, eugenicist, and even the neofascist Northern League.
 
-In the context of \emph{Brown}, MQ was created as what contemporary
-scholars call an ``echo chamber''
-\cite{FernandezPintoKnowBetterNot2017}, a favorable venue for race
-scientists to publish their views, on the grounds that an ``equalitarian
-dogma'' created a censorious ``taboo'' against their research in
-mainstream publications
-\cite{TuckerFundingScientificRacism2002, JacksonMythicalTabooRace2020}.
-Historians, philosophers, and sociologists of science have shown that
-echo chambers have been a key part of the ``tobacco strategy,'' used by
-numerous regulated industries --- most infamously tobacco, but also
-fossil fuels and chemical manufacturing, among others --- to
-``manufacture doubt'' and delay regulation to protect public and
-environmental health \emph{{[}cites{]}}.
+In the context of \emph{Brown}, MQ was created as what contemporary scholars call an ``echo chamber'' \cite{FernandezPintoKnowBetterNot2017}, a favorable venue for race scientists to publish their views, on the grounds that an ``equalitarian dogma'' created a censorious ``taboo'' against their research in mainstream publications \cite{TuckerFundingScientificRacism2002, JacksonMythicalTabooRace2020}.
+Historians, philosophers, and sociologists of science have shown that echo chambers have been a key part of the ``tobacco strategy,'' used by numerous regulated industries --- most infamously tobacco, but also fossil fuels and chemical manufacturing, among others --- to ``manufacture doubt'' and delay regulation to protect public and environmental health \emph{{[}cites{]}}.
 
-In light of the origins of MQ, the significant attention paid to the
-journal by historians of scientific racism, and contemporary research on
-the ``tobacco strategy,'' it is surprising that MQ did not play a
-significant role in nurturing race science.
+In light of the origins of MQ, the significant attention paid to the journal by historians of scientific racism, and contemporary research on the ``tobacco strategy,'' it is surprising that MQ did not play a significant role in nurturing race science.
 
 \hypertarget{mankind-quarterly-and-pioneer-funded-researchers}{%
-\section*{\texorpdfstring{\emph{Mankind Quarterly} and Pioneer-funded
-researchers}{Mankind Quarterly and Pioneer-funded researchers}}\label{mankind-quarterly-and-pioneer-funded-researchers}}
+\section*{\texorpdfstring{\emph{Mankind Quarterly} and Pioneer-funded researchers}{Mankind Quarterly and Pioneer-funded researchers}}\label{mankind-quarterly-and-pioneer-funded-researchers}}
 
-\emph{{[}for xref:
-\url{https://bookdown.org/yihui/bookdown/a-single-document.html}{]}}
-
-We identified 16 researchers who had received funding from Pioneer; 14
-of these researchers had profiles in the Web of Science (WoS) author
-search \emph{{[}table + WoS pub coucnt{]}}, allowing us identify 13
-WoS-indexed journals that had published 6 or more of these authors.
+We identified 16 researchers who had received funding from Pioneer; 14 of these researchers had profiles in the Web of Science (WoS) author search \emph{{[}table + WoS pub coucnt{]}}, allowing us identify 13 WoS-indexed journals that had published 6 or more of these authors.
+See Table \ref{tab:researchers}.
 
 \begin{longtable}[]{@{}lrl@{}}
-\caption{Pioneer-funded researchers. Either identified in
-\emph{{[}cite{]}} or named on Pioneer's website, along with identified
-discipline and WoS author search result counts.}\tabularnewline
+\caption{\label{tab:researchers} Pioneer-funded researchers.
+Either identified in \emph{{[}cite{]}} or named on Pioneer's website, along with identified discipline and WoS author search result counts.}\tabularnewline
 \toprule
@@
 \includegraphics[width=4.76in,height=2.6in]{img/wos_results.png}
-\caption{Journals publishing 6 or more Pioneer-funded researchers, Web
-of Science author search results \emph{{[}date coverage{]}}}
+\caption{Journals publishing 6 or more Pioneer-funded researchers, Web of Science author search results \emph{{[}date coverage{]}} \label{fig:wos}}
 \end{figure}
 
-Figure \emph{{[}fig{]}} shows that, while MQ is among the
-``Pioneer-publishing'' journals, a number of mainstream journals are
-more prominent: \emph{Personality and Individual Differences} (PID),
-\emph{Intelligence} (I), \emph{Behavior Genetics} (BG), and
-\emph{Psychological Reports} (PR). In addition, only psychologist
-Richard Lynn appears to have published heavily in MQ. Lynn has been an
-assistant editor of MQ since 1979 \emph{{[}check this:
-\url{https://rationalwiki.org/wiki/Mankind_Quarterly}{]}} and president
-of PF since the death of psychologist J. Phillippe Rushton in 2012
-\emph{{[}cite{]}}. By contrast, a number of Pioneer-funded researchers
-have published in PID, I, and to a lesser degree BG: Bouchard, Eysenck,
-Jensen, Rushton, Vernon, and also Lynn.
+Figure \ref{fig:wos} shows that, while MQ is among the ``Pioneer-publishing'' journals, a number of mainstream journals are more prominent: \emph{Personality and Individual Differences} (PID), \emph{Intelligence} (I), \emph{Behavior Genetics} (BG), and \emph{Psychological Reports} (PR).
+In addition, only psychologist Richard Lynn appears to have published heavily in MQ.
+Lynn has been an assistant editor of MQ since 1979 \emph{{[}check this: \url{https://rationalwiki.org/wiki/Mankind_Quarterly}{]}} and president of PF since the death of psychologist J.
+Phillippe Rushton in 2012 \emph{{[}cite{]}}.
+By contrast, a number of Pioneer-funded researchers have published in PID, I, and to a lesser degree BG: Bouchard, Eysenck, Jensen, Rushton, Vernon, and also Lynn.
 
-\emph{Personality and Individual Differences} (PID) was founded in 1980,
-with Eysenck as editor-in-chief and an editorial board including Jensen
-and Lynn. In the inaugural editorial, Eysenck identified ``studies of
-the genetic determinants of individual differences in the areas of
-personality and intelligence'' as one of the journal's eight major areas
-of interest. Eysenck remained editor-in-chief until his death in 1997.
-In 2005 the editorial board still included Jensen and Lynn. PID was
-first published by Pergamon Press, a mainstream academic press, and
-today is published by Elsevier. According to the Scimago Journal
-Rankings, based on Scopus citation data, in 2022 PID had an H-index of
-193, \#29 in Psychology (Miscellaneous), and a 2-year citation index
-(impact factor) of 5.27, \#24 in Psychology (Miscellaneous).
+\emph{Personality and Individual Differences} (PID) was founded in 1980, with Eysenck as editor-in-chief and an editorial board including Jensen and Lynn.
+In the inaugural editorial, Eysenck identified ``studies of the genetic determinants of individual differences in the areas of personality and intelligence'' as one of the journal's eight major areas of interest.
+Eysenck remained editor-in-chief until his death in 1997.
+In 2005 the editorial board still included Jensen and Lynn.
+PID was first published by Pergamon Press, a mainstream academic press, and today is published by Elsevier.
+According to the Scimago Journal Rankings, based on Scopus citation data, in 2022 PID had an H-index of 193, \#29 in Psychology (Miscellaneous), and a 2-year citation index (impact factor) of 5.27, \#24 in Psychology (Miscellaneous).
 
-\emph{Intelligence} (I) was founded in 1977, with psychologist Douglas
-Detterman as editor-in-chief from the founding until 2016. Lloyd
-Humphreys was on the editorial board starting from 1977; by 1990 he had
-been joined by Jensen and Philip Vernon. Richard Lynn joined the
-editorial board sometime between 1998 and 2002. (Archive copies of the I
-editorial board page are not available from the journal's website from
-1999 through 2001.) I has been criticized for including Lynn and Gerhard
-Meisenberg on its editorial board until July-August 2018
-\cite{SainiSuperiorReturnRace2019}. Meisenberg is also connected with
-the Pioneer Fund \emph{{[}director?{]}}, and was editor-in-chief of MQ
-in 2015-18 \emph{{[}check these{]}}. I is published by Elsevier.
-According to the Scimago Journal Rankings, based on Scopus citation
-data, in 2022 I had an H-index of 103, \#32 in Experimental and
-Cognitive Psychology, and a 2-year citation index (impact factor) of
-3.05, \#31 in Experimental and Cognitive Psychology.
+\emph{Intelligence} (I) was founded in 1977, with psychologist Douglas Detterman as editor-in-chief from the founding until 2016.
+Lloyd Humphreys was on the editorial board starting from 1977; by 1990 he had been joined by Jensen and Philip Vernon.
+Richard Lynn joined the editorial board sometime between 1998 and 2002.
+(Archive copies of the I editorial board page are not available from the journal's website from 1999 through 2001.) I has been criticized for including Lynn and Gerhard Meisenberg on its editorial board until July-August 2018 \cite{SainiSuperiorReturnRace2019}.
+Meisenberg is also connected with the Pioneer Fund \emph{{[}director?{]}}, and was editor-in-chief of MQ in 2015-18 \emph{{[}check these{]}}.
+I is published by Elsevier.
+According to the Scimago Journal Rankings, based on Scopus citation data, in 2022 I had an H-index of 103, \#32 in Experimental and Cognitive Psychology, and a 2-year citation index (impact factor) of 3.05, \#31 in Experimental and Cognitive Psychology.
 
@@
 
-The fact that Pioneer-funded researchers published heavily in two
-mainstream psychology journals does not tell us anything about the
-content of their publications or the findings of their research. We
-assembled a full-text corpus of articles published in MQ, PID, I, BG,
-PR, and Behavior and Brain Sciences (BBS) between 1960-2010 and used
-topic modeling to identify topics discussing race, intelligence, and
-both. \emph{{[}tables: Ns for each article, topic model parameters;
-figs: articles by pub over time, topic heatmaps, Silge plots,
-race-intelligence-both grids{]}}
+The fact that Pioneer-funded researchers published heavily in two mainstream psychology journals does not tell us anything about the content of their publications or the findings of their research.
+We assembled a full-text corpus of articles published in MQ, PID, I, BG, PR, and Behavior and Brain Sciences (BBS) between 1960-2010 and used topic modeling to identify topics discussing race, intelligence, and both.
+\emph{{[}tables: Ns for each article, topic model parameters; figs: articles by pub over time, topic heatmaps, Silge plots, race-intelligence-both grids{]}}
 
@@
 \caption{Silge plots and smoothed time series for three focal topics.
-Top row: top 15 terms by \(\beta\) (term-topic distribution) for each
-topic. Bottom row: count of articles associated with each topic, by
-journal and year. Article counts use a threshold approach, with
-\(\gamma\) (topic-document distribution) greater than 0.5. Thin lines
-give annual values, thick lines give 5-year running averages. Because
-\emph{Behavior Genetics} and \emph{Behavioral and Brain Sciences} are
-not prominent in any panel, neither is given a direct label.}
+Top row: top 15 terms by \(\beta\) (term-topic distribution) for each topic.
+Bottom row: count of articles associated with each topic, by journal and year.
+Article counts use a threshold approach, with \(\gamma\) (topic-document distribution) greater than 0.5.
+Thin lines give annual values, thick lines give 5-year running averages.
+Because \emph{Behavior Genetics} and \emph{Behavioral and Brain Sciences} are not prominent in any panel, neither is given a direct label.}
 \end{figure}
 
-We focus here on three topics identified in one of the 24 fitted topic
-models \emph{{[}composite fig: topics over time + Silge plots{]}}. In
-this model, topic 07 is strongly associated with MQ: MQ published dozens
-of articles in this topic each year, and no other journal ever published
-more than a handful, with most other journals publishing none in a given
-year. The Silge plot contains racial terms (\texttt{races},
-\texttt{whites}, \texttt{negroes} and potentially \texttt{europe},
-\texttt{africa}, \texttt{india}, and \texttt{japan}) as well as
-\texttt{book}, likely reflecting the fact that MQ published a number of
-book reviews while the psychology journals did not.
+We focus here on three topics identified in one of the 24 fitted topic models \emph{{[}composite fig: topics over time + Silge plots{]}}.
+In this model, topic 07 is strongly associated with MQ: MQ published dozens of articles in this topic each year, and no other journal ever published more than a handful, with most other journals publishing none in a given year.
+The Silge plot contains racial terms (\texttt{races}, \texttt{whites}, \texttt{negroes} and potentially \texttt{europe}, \texttt{africa}, \texttt{india}, and \texttt{japan}) as well as \texttt{book}, likely reflecting the fact that MQ published a number of book reviews while the psychology journals did not.
 
-Topic 22 is strongly associated with I and PID in the same way. The
-Silge plot does contain \texttt{jensen}, as well as a reference to
-Raymond Cattell, who played a major role in the development of factor
-analysis and intelligence testing also advocated eugenics, fascism, and
-Nazi race science\cite{MehlerBeyondismRaymondCattell1997}. However, the
-other authors named in this topic --- John Horn and Peter Bentler --- do
-not appear to have contributed to race science discourse. Instead this
-topic appears to identify ``mainstream'' (non-race science) intelligence
-research, specifically on factor analysis and the debate over whether
-intelligence in unidimensional or multidimensional. (Three other topics
-only associated with mainstream intelligence journals and terms were
-also identified by this topic model.) This topic indicates that the
-model is not simply lumping race science research together with other
-intelligence research.
+Topic 22 is strongly associated with I and PID in the same way.
+The Silge plot does contain \texttt{jensen}, as well as a reference to Raymond Cattell, who played a major role in the development of factor analysis and intelligence testing also advocated eugenics, fascism, and Nazi race science\cite{MehlerBeyondismRaymondCattell1997}.
+However, the other authors named in this topic --- John Horn and Peter Bentler --- do not appear to have contributed to race science discourse.
+Instead this topic appears to identify ``mainstream'' (non-race science) intelligence research, specifically on factor analysis and the debate over whether intelligence in unidimensional or multidimensional.
+(Three other topics only associated with mainstream intelligence journals and terms were also identified by this topic model.) This topic indicates that the model is not simply lumping race science research together with other intelligence research.
 
-The Silge plot for topic 24, by contrast, suggests a topic focused
-squarely on race and intelligence, with multiple racial terms
-(\texttt{whites}, \texttt{blacks}, \texttt{racial\_differences},
-\texttt{races}, \texttt{race\_differences}) and the names of three
-prominent Pioneer-funded researchers, \texttt{jensen}, \texttt{lynn},
-and \texttt{rushton}. To confirm this impression from the Silge plot, we
-independently coded the top 121 papers in this topic (those with
-\(gamma > 0.97\)) as contributing to race science discourse or not. 105
-(87\%) were coded as race science discourse by both authors; 10 (8.3\%)
-were coded as not race science discourse by both; and discordant codes
-were given to the remaining 6 (5.0\%) (95\% agreement, Cohen's
-\(\kappa = 0.74\)).
+The Silge plot for topic 24, by contrast, suggests a topic focused squarely on race and intelligence, with multiple racial terms (\texttt{whites}, \texttt{blacks}, \texttt{racial\_differences}, \texttt{races}, \texttt{race\_differences}) and the names of three prominent Pioneer-funded researchers, \texttt{jensen}, \texttt{lynn}, and \texttt{rushton}.
+To confirm this impression from the Silge plot, we independently coded the top 121 papers in this topic (those with \(gamma > 0.97\)) as contributing to race science discourse or not.
+105 (87\%) were coded as race science discourse by both authors; 10 (8.3\%) were coded as not race science discourse by both; and discordant codes were given to the remaining 6 (5.0\%) (95\% agreement, Cohen's \(\kappa = 0.74\)).
 
-In almost all years, most papers in topic 24 were published in
-mainstream journals rather than MQ. Jensen's ``How Much Can We Boost IQ
-and Scholastic Achievement?'' \cite{JensenHowMuchCan1969} was published
-in 1969 in Harvard Educational Review (not included in this study), and
-the time series indicates that, around this time, there was an increase
-in articles in topic 24 in both PS and MQ (the only two journals in our
-corpus that were active at the time). MQ shows another sharp increase in
-the late 1970s, around the time Lynn was brought on as assistant editor,
-\emph{{[}confirming the claim that Lynn shifted MQ's focus to
-intelligence research{]}}. PID published multiple papers in topic 24
-almost immediately after it was founded, with I showing a more gradual
-increase between the mid-1970s and mid-1990s.
+In almost all years, most papers in topic 24 were published in mainstream journals rather than MQ.
+Jensen's ``How Much Can We Boost IQ and Scholastic Achievement?'' \cite{JensenHowMuchCan1969} was published in 1969 in Harvard Educational Review (not included in this study), and the time series indicates that, around this time, there was an increase in articles in topic 24 in both PS and MQ (the only two journals in our corpus that were active at the time).
+MQ shows another sharp increase in the late 1970s, around the time Lynn was brought on as assistant editor, \emph{{[}confirming the claim that Lynn shifted MQ's focus to intelligence research{]}}.
+PID published multiple papers in topic 24 almost immediately after it was founded, with I showing a more gradual increase between the mid-1970s and mid-1990s.
 
-Notably, BG published very few articles in this topic. Outside of the
-field itself, behavior genetics is strongly associated with race science
-by both other academics and the general public. Panofsky
-\cite{PanofskyMisbehavingScienceControversy2014} argues that, prior to
-Jensen's 1969 paper, behavior genetics emphasized the study of non-human
-animals and intentionally avoided associations with eugenics and public
-controversy more generally. In response to Jensen, critics such as
-Lewontin offered broad critiques of behavior genetics as such, and
-behavior geneticists in turn adopted a radical conception of academic
-freedom (without any sense of responsibility for the social implications
-of academic research) and a siege or wartime mentality, as demonstrated
-by Sandra Scarr's 1986 presidential address to the Behavior Genetics
-Association (BGA) \cite{ScarrThreeCheersBehavior1987}. Scarr's address
-coincided with the period between 1970-1990 when BG published articles
-in topic 24 (including Scarr's address itself).
+Notably, BG published very few articles in this topic.
+Outside of the field itself, behavior genetics is strongly associated with race science by both other academics and the general public.
+Panofsky \cite{PanofskyMisbehavingScienceControversy2014} argues that, prior to Jensen's 1969 paper, behavior genetics emphasized the study of non-human animals and intentionally avoided associations with eugenics and public controversy more generally.
+In response to Jensen, critics such as Lewontin offered broad critiques of behavior genetics as such, and behavior geneticists in turn adopted a radical conception of academic freedom (without any sense of responsibility for the social implications of academic research) and a siege or wartime mentality, as demonstrated by Sandra Scarr's 1986 presidential address to the Behavior Genetics Association (BGA) \cite{ScarrThreeCheersBehavior1987}.
+Scarr's address coincided with the period between 1970-1990 when BG published articles in topic 24 (including Scarr's address itself).
 
-However, Glayde Whitney's 1995 BGA presidential address was published in
-MQ rather than BG. According to Whitney, BG's editor refused to publish
-the address and the BGA Executive Committee (except for Whitney himself)
-voted to ``issue an official statement of denouncement'' of the address;
-though we were unable to find any such a statement, and Panofsky reports
-that ``nothing official was done to Whitney.'' In his address, Whitney
-criticized ``the Marxist-Lysenkoist denial of genetics''; quoted Peter
-Brimelow --- a white nationalist and later founder of the racist and
-xenophobic website VDARE --- defining ``racist'' as ``anyone who is
-winning an argument with a liberal''; and proposed as a ``reasonable
-scientific hypothesis'' that differences in murder rates between
-countries and cities was caused by racial genetic differences in
-intelligence, empathy, aggression, and impulsivity.
+However, Glayde Whitney's 1995 BGA presidential address was published in MQ rather than BG.
+According to Whitney, BG's editor refused to publish the address and the BGA Executive Committee (except for Whitney himself) voted to ``issue an official statement of denouncement'' of the address; though we were unable to find any such a statement, and Panofsky reports that ``nothing official was done to Whitney.'' In his address, Whitney criticized ``the Marxist-Lysenkoist denial of genetics''; quoted Peter Brimelow --- a white nationalist and later founder of the racist and xenophobic website VDARE --- defining ``racist'' as ``anyone who is winning an argument with a liberal''; and proposed as a ``reasonable scientific hypothesis'' that differences in murder rates between countries and cities was caused by racial genetic differences in intelligence, empathy, aggression, and impulsivity.
 
-The topic model analysis and the case of Whitney's address suggest that,
-by the 1990s, behavior genetics had at least somewhat distanced itself
-from race science.
+The topic model analysis and the case of Whitney's address suggest that, by the 1990s, behavior genetics had at least somewhat distanced itself from race science, though perhaps without repudiating it.
 \emph{{[}\url{https://link.springer.com/search?use-oscar-shared-search=true\&sortBy=newestFirst\&dateTo=2010\&sortOrder=newestFirst\&facet-journal-id=10519\&content-type=Article\&query=race+intelligence\&date=custom\&dateFrom=\&date-facet-mode=between\&facet-start-year=1972\&previous-start-year=1972\&facet-end-year=2010\&previous-end-year=2023}{]}}
 
-A disciplinary history of behavior genetics published in 2010
-\cite{McGueEndBehavioralGenetics2010} acknowledges the historical
-association with eugenics, but does not mention Jensen and other race
-science controversies since World War II. Instead postwar behavior
-genetics is portrayed as an endeavor to counterbalance the hegemony of
-the ``Blank Slate,'' the supposedly widespread view that genetic
-differences play no role in explaining individual differences
-\cite{BatesonCorpseWearisomeDebate2002}.
+To check this interpretation, we ran an independent search for documents with the keywords ``race'' and ``intelligence'' published in BG from 1972 to 2020, using Springer's journal search website.
+This search returned 125 documents.
+33 were presentation abstracts from meetings of the Behavior Genetics Association; we did not examine these further.
+14 of the remaining documents were published since 2010 (after the scope of the primary study).
+Among these, one obituary and two papers celebrated the work of John Loehlin, a prominent race-and-intelligence researcher who had been Director of the American Eugenics Society from 1968-1972 \emph{{[}\url{http://link.springer.com/article/10.1007/s10519-020-10024-w}; \url{http://link.springer.com/article/10.1007/s10519-014-9687-1}; \url{http://link.springer.com/article/10.1007/s10519-014-9673-7}{]}}.
+Another was a retrospective of the work of Lindon Eaves, a geneticist who made at least one notable contribution to debates on the Scarr-Rowe hypothesis (interactions between heritability, race, and class).
+None of these 14 documents reported any studies of racial differences.
 
-These findings for \emph{Behavior Genetics} complicate at least the
-popular association of behavior genetics with race science, if not
-Panofsky's more nuanced analysis.
+28 documents in this sample were published between 1990 and 2010.
+One was an obituary of Jerry Hirsch, a critic of hereditarianism in general and Jensen in particular \emph{{[}\url{http://link.springer.com/article/10.1007/s10519-008-9231-2}{]}}, and another was an obituary of David Rowe \emph{{[}\url{http://link.springer.com/article/10.1023/A\%3A1026122928868}{]}}.
+Only 2 of these 28 documents reported race differences of any kind: one examining at interactions among race, sex, and heritability for adolescent BMI (Body Mass Index, used as a measure of overweight/obesity) \emph{{[}\url{http://link.springer.com/article/10.1023/A\%3A1021619329904}{]}}; and the other at interactions among race, family history of alcoholism, and visuospatial performance \emph{{[}\url{http://link.springer.com/article/10.1007/BF02197241}{]}}.
+Finally, Philip Vernon --- one of the Pioneer-funded scientists we identified earlier --- published a critical review of a book that attempted to address hereditarianism, the Flynn effect, race-and-intelligence research, and some related issues.
+
+This supplemental review provides additional support for the interpretation that behavior genetics had distanced itself from race science by the 1990s.
+Notably, this review brought our attention to a disciplinary history of behavior genetics published in 2010 \cite{McGueEndBehavioralGenetics2010}, which acknowledges the historical association with eugenics but does not mention Jensen and other race science controversies since World War II.
+Instead postwar behavior genetics is portrayed as an endeavor to counterbalance the hegemony of the ``Blank Slate,'' the supposedly widespread view that genetic differences play no role in explaining individual differences \cite{BatesonCorpseWearisomeDebate2002}.
+And none of the three documents celebrating Loehlin noted his postwar involvement in the eugenics movement.
+None of the documents examined in this supplemental review explicitly repudiate race science.
+Behavior geneticists seem to typically adopt a posture of objectivity and technical rigor, downplaying the extent of the field's involvement in race science in the past.
 
@@
 
-Our findings indicate that \emph{Mankind Quarterly} (MQ) was less
-important as an echo chamber for late 20th century race science than
-mainstream psychology journals, especially \emph{Intelligence} (I) and
-\emph{Personality and Individual Differences} (PID). Indeed, MQ
-published a very different kind of race science from the
-race-and-intelligence research published in I and PID, even after
-\emph{{[}Lynn joined the MQ editorial board{]}}.
+Our findings indicate that \emph{Mankind Quarterly} (MQ) was less important as an echo chamber for late 20th century race science than mainstream psychology journals, especially \emph{Intelligence} (I) and \emph{Personality and Individual Differences} (PID).
+Indeed, MQ published a very different kind of race science from the race-and-intelligence research published in I and PID, even after \emph{{[}Lynn joined the MQ editorial board{]}}.
 
-We stress that we are not identifying any particular scientist as a
-racist based on the topic model results. Scientific research can be
-utilized to promote scientific racism --- and thus count as race science
---- independently of the views and attitudes of the scientist who
-conducted that
-research.\cite{TaberyWhyStudyingGenetics2015, CarlsonQuantifyingContextualizingImpact2020, HennWhyDNANo2021}
-But this does not mean that scientists are not responsible for the
-downstream social effects of their research, even when these effects
-involve deliberate misinterpretation.\emph{{[}Douglas, Kitcher, Kourany,
-Havstad; Block and Dworkin?{]}} Like other citizens, scientists are
-responsible for mitigating the reasonably foreseen harms that result
-from their actions. And we now have decades of empirical evidence that
-research on the genetics of intelligence --- even research that does not
-itself touch on race --- can and will be used to promote scientific
-racism.\cite{MeyerWrestlingSocialBehavioral2023}
+We stress that we are not identifying any particular scientist as a racist based on the topic model results.
+Scientific research can be utilized to promote scientific racism --- and thus count as race science --- independently of the views and attitudes of the scientist who conducted that research.\cite{TaberyWhyStudyingGenetics2015, CarlsonQuantifyingContextualizingImpact2020, HennWhyDNANo2021} But this does not mean that scientists are not responsible for the downstream social effects of their
+research, even when these effects involve deliberate misinterpretation.\emph{{[}Douglas, Kitcher, Kourany, Havstad; Block and Dworkin?{]}} Like other citizens, scientists are responsible for mitigating the reasonably foreseen harms that result from their actions.
+And we now have decades of empirical evidence that research on the genetics of intelligence --- even research that does not itself touch on race --- can and will be used to promote scientific racism.\cite{MeyerWrestlingSocialBehavioral2023}
 
@@
 
-\emph{{[}this probably goes too far even for a discussion section{]}}
+\emph{{[}limitation: colorblind racism (Bonilla-Silva; WinstonWhyMainstreamResearch2020) creates coding ambiguities{]}}
 
-Proponents of a strong conception of academic freedom often point to
-chapter 2 of philosopher John Stuart Mill's \emph{On Liberty}
-\cite{MillLiberty1862}, ``Of the Liberty of Thought and Discussion.''
-However, in chapter 1 of the same book, Mill states the Harm Principle
-as the overall thesis of his book: ``The only end for which people are
-entitled, individually or collectively, to interfere with the liberty of
-action of any of their number is self-protection. The only purpose for
-which power can be rightfully exercised over any member of a civilized
-community, against his will, is to prevent harm to others.'' Even for
-Mill, free speech and free inquiry are appropriately restrained when the
-ideas in question will harm others.
+\emph{{[}limitation: false negatives (race science papers that don't appear in V24) are possible if, eg, they use many terms related to twin studies of heritability and thereby appear in V03{]}}
 
-More recently, philosopher Shannon Dea
-\cite{DeaEvolvingSocialPurpose2021} has argued that academic freedom is
-importantly distinct from freedom of speech. Whereas freedom of speech
-is universal, enjoyed by all residents of countries where speech is
-legally protected, academic freedom is more exclusive, enjoyed only by
-college and university faculty. (Though, as Dea points out,
-non-tenure-track faculty often do not have effective protections for
-academic freedom.) Freedom of speech primarily serves to protect the
-interests of the individual speaker: that you will not be punished by
-the state for stating unpopular opinions. But academic freedom is
-``other-directed,'' primarily serving to benefit the common good of
-society as a whole by enabling the pursuit of truth.
+\emph{{[}replace w/: WinstonWhyMainstreamResearch2020: race science as a community project of psychology(?).
+radical academic freedom is a tool used to defend race science PigliucciWhatAreWe2013, RosemanTroublesomeReflectionRacism2014, LarsenMoreProvocativeLess2020.
+an important part of how these mainstream journals function as race science echo chambers.{]}}
 
-Advocates of research on genetics and intelligence --- both anti-racist
-liberals and race scientists --- claim this work will lead to social
-benefits. This reflects a recognition that the purpose of academic
-freedom is to promote the common good, rather than merely satisfy an
-individual's curiosity. But the promised benefits are either vague or do
-not actually require understanding the genetics of
-intelligence.\cite{KaplanGaltonQuincunxProbabilistic2021, CoopLotteryLuckLegacy2022}
+Proponents of a strong conception of academic freedom often point to chapter 2 of philosopher John Stuart Mill's \emph{On Liberty} \cite{MillLiberty1862}, ``Of the Liberty of Thought and Discussion.'' However, in chapter 1 of the same book, Mill states the Harm Principle as the overall thesis of his book: ``The only end for which people are entitled, individually or collectively, to interfere with the liberty of action of any of their number is self-protection.
+The only purpose for which power can be rightfully exercised over any member of a civilized community, against his will, is to prevent harm to others.'' Even for Mill, free speech and free inquiry are appropriately restrained when the ideas in question will harm others.
 
-Using agent-based models, Weatherall et
-al.\cite{WeatherallHowBeatScience2018} show that a strategy of
-``selective sharing'' --- promoting favorable results from independent
-research --- can be effective at misleading a target audience.
+More recently, philosopher Shannon Dea \cite{DeaEvolvingSocialPurpose2021} has argued that academic freedom is importantly distinct from freedom of speech.
+Whereas freedom of speech is universal, enjoyed by all residents of countries where speech is legally protected, academic freedom is more exclusive, enjoyed only by college and university faculty.
+(Though, as Dea points out, non-tenure-track faculty often do not have effective protections for academic freedom.) Freedom of speech primarily serves to protect the interests of the individual speaker: that you will not be punished by the state for stating unpopular opinions.
+But academic freedom is ``other-directed,'' primarily serving to benefit the common good of society as a whole by enabling the pursuit of truth.
 
-\emph{{[}for discussion: the possibility for white supremacist
-appropriation, ie, race science done by scientists who are not
-themselves scientific racists:
-CarlsonQuantifyingContextualizingImpact2020; probably some reviews of
-Paige Harden; O'Connor and Weatherall{]}}
+Advocates of research on genetics and intelligence --- both anti-racist liberals and race scientists --- claim this work will lead to social benefits.
+This reflects a recognition that the purpose of academic freedom is to promote the common good, rather than merely satisfy an individual's curiosity.
+But the promised benefits are either vague or do not actually require understanding the genetics of intelligence.\cite{KaplanGaltonQuincunxProbabilistic2021, CoopLotteryLuckLegacy2022}
+
+Using agent-based models, Weatherall et al.\cite{WeatherallHowBeatScience2018} show that a strategy of ``selective sharing'' --- promoting favorable results from independent research --- can be effective at misleading a target audience.
+
+\emph{{[}for discussion: the possibility for white supremacist appropriation, ie, race science done by scientists who are not themselves scientific racists: CarlsonQuantifyingContextualizingImpact2020; probably some reviews of Paige Harden; O'Connor and Weatherall{]}}
+
+\emph{{[}what can science do? Posselt, /Equity in Science/{]}}
 
@@
 \begin{scilastnote}
-\item \emph{{[}strutured{]}} Thanks to Emily Merchant and John Jackson
-for some initial discussions that helped clarify the scope of the
-project and identify some key resources related to the Pioneer Fund and
-\emph{Mankind Quarterly}. Thanks to Anthony Sainez for help retrieving
-the \emph{Mankind Quarterly} articles. Thanks to Derek Devnich and James
-Dooley for their work attempting to secure the APA-published articles.
-EL's work on this project was supported by UC Merced. DH's work on this
-project was not supported by any specific funding.
+\item \emph{{[}strutured{]}} Thanks to Emily Merchant and John Jackson for some initial discussions that helped clarify the scope of the project and identify some key resources related to the Pioneer Fund and \emph{Mankind Quarterly}.
+Thanks to Anthony Sainez for help retrieving the \emph{Mankind Quarterly} articles.
+Thanks to Derek Devnich and James Dooley for their work attempting to secure the APA-published articles.
+EL's work on this project was supported by UC Merced.
+DH's work on this project was not supported by any specific funding.
 \end{scilastnote}

### 2023-07-10 paper/supplement_paper.tex
enable cross-references

@@
 and Emilio Lobato
-\\\textsuperscript{*}Corresponding author: Daniel J. Hicks,
-\href{mailto:dhicks4@ucmerced.edu}{\nolinkurl{dhicks4@ucmerced.edu}}.
+\\\textsuperscript{*}Corresponding author: Daniel J. Hicks, \href{mailto:dhicks4@ucmerced.edu}{\nolinkurl{dhicks4@ucmerced.edu}}.
 }
@@
 
-Corpus assembly is summarized in figure \emph{{[}xref{]}}. The corpus
-was assembled in two parts, one for \emph{Mankind Quarterly} and the
-other for mainstream journals.
+Corpus assembly is summarized in figure \emph{{[}xref{]}}. The corpus was assembled in two parts, one for \emph{Mankind Quarterly} and the other for mainstream journals.
 
@@
 
-An electronic archive of \emph{Mankind Quarterly}, from the first issue
-in \emph{{[}1969{]}} through 2004, is available for free on the open web
-from \emph{The Unz Review}, which has been characterized as a white
-nationalist website by the Southern Poverty Law Center
-\emph{{[}\url{https://www.splcenter.org/hatewatch/2021/01/19/meet-white-nationalist-organizer-who-spewed-hate-against-lawmakers}{]}}.
-Web scraping packages and techniques \emph{{[}which?{]}} were used to
-retrieve every PDF available in this archive. These PDFs were processed
-using the open-source OCR (optical character recognition) software
-\texttt{tesseract} and the R API \texttt{rtesseract} \emph{{[}cites{]}}.
+An electronic archive of \emph{Mankind Quarterly}, from the first issue in \emph{{[}1969{]}} through 2004, is available for free on the open web from \emph{The Unz Review}, which has been characterized as a white nationalist website by the Southern Poverty Law Center \emph{{[}\url{https://www.splcenter.org/hatewatch/2021/01/19/meet-white-nationalist-organizer-who-spewed-hate-against-lawmakers}{]}}.
+Web scraping packages and techniques \emph{{[}which?{]}} were used to retrieve every PDF available in this archive.
+These PDFs were processed using the open-source OCR (optical character recognition) software \texttt{tesseract} and the R API \texttt{rtesseract} \emph{{[}cites{]}}.
 This stage of corpus assembly was conducted in Fall 2021.
@@
 
-To identify suitable mainstream journals for inclusion in the corpus, we
-first identified academic researchers who had been funded by the Pioneer
-Fund. Pioneer is an American non-profit organization founded in 1937 to
-fund eugenics research \emph{{[}cite{]}}. After the \emph{Brown v. Board
-of Education} ruling in 1954, Pioneer funded various segregationist
-efforts across the United States. \emph{{[}Shockley and Jensen{]}}. We
-reviewed a critical profile of Pioneer \emph{{[}cite{]}} as well as an
-archived page from the organization's own web site \emph{{[}cite{]}},
-which together listed 16 researchers who had received Pioneer funds.
+To identify suitable mainstream journals for inclusion in the corpus, we first identified academic researchers who had been funded by the Pioneer Fund.
+Pioneer is an American non-profit organization founded in 1937 to fund eugenics research \emph{{[}cite{]}}.
+After the \emph{Brown v.
+Board of Education} ruling in 1954, Pioneer funded various segregationist efforts across the United States.
+\emph{{[}Shockley and Jensen{]}}.
+We reviewed a critical profile of Pioneer \emph{{[}cite{]}} as well as an archived page from the organization's own web site \emph{{[}cite{]}}, which together listed 16 researchers who had received Pioneer funds.
 
-We then used the author search tool in Clarivate's Web of Science
-platform \emph{{[}url{]}}, retrieving publication lists for 14
-researchers. These searches were conducted between 2021-09-24 and
-2021-10-05 by DJH. After parsing these results, we counted how many of
-the 14 researchers had published in each journal. 13 journals had
-published 6 or more of the 14 Pioneer-funded researchers. We excluded
-\emph{Mankind Quarterly} (as already included) as well as \emph{Science}
-and \emph{Nature} (as too general) from further consideration in this
-side of the corpus. 3 journals published by the American Psychological
-Association (APA; \emph{American Psychologist}, \emph{Contemporary
-Psychology}, and \emph{Journal of Educational Psychology}) had to be
-excluded due to confusion over who could give us permission to use the
-archives for a text mining project, with both APA and ProQuest asserting
-that we needed to get permission from the other entity. \emph{European
-Journal of Personality}, published by SAGE, also had to be excluded
-because our institutional access only went back to 1999.
+We then used the author search tool in Clarivate's Web of Science platform \emph{{[}url{]}}, retrieving publication lists for 14 researchers.
+These searches were conducted between 2021-09-24 and 2021-10-05 by DJH.
+After parsing these results, we counted how many of the 14 researchers had published in each journal.
+13 journals had published 6 or more of the 14 Pioneer-funded researchers.
+We excluded \emph{Mankind Quarterly} (as already included) as well as \emph{Science} and \emph{Nature} (as too general) from further consideration in this side of the corpus.
+3 journals published by the American Psychological Association (APA; \emph{American Psychologist}, \emph{Contemporary Psychology}, and \emph{Journal of Educational Psychology}) had to be excluded due to confusion over who could give us permission to use the archives for a text mining project, with both APA and ProQuest asserting that we needed to get permission from the other entity.
+\emph{European Journal of Personality}, published by SAGE, also had to be excluded because our institutional access only went back to 1999.
 
-The remaining 5 journals are all published by major academic publishers
---- Elsevier, Springer, or Cambridge University Press --- and each item
-in the entire run of each journal has been assigned a DOI (digital
-object identifier) for archival purposes. We used the Crossref API and
-\texttt{rcrossref} R interface to this API \emph{{[}cite{]}} to retrieve
-metadata for each item published from 1960-2010 in each of these 6
-journals. These metadata included item-level license information ---
-confirming that the text of each item could be used for text mining
-projects --- and a URL to an electronic version of the item. These URLs
-were used to retrieve an HTML or PDF version of each item, except for
-\emph{Personality and Individual Differences}. This journal has
-published a relatively large number of non-article documents, such as
-book reviews and commentaries, which are not available at the URL
-included in the Crossref metadata. (This is unfortunate, as it was not
-difficult to find highly relevant documents that we could not
-automatically retrieve and therefore excluded from our corpus
-\emph{{[}review{]}}.) Instead we used Elsevier's ScienceDirect API
-\emph{{[}url{]}} to search and retrieve all items from \emph{Personality
-and Individual Differences} that are available. PDFs were run through
-OCR and text extraction as necessary, as with the \emph{Mankind
-Quarterly} documents.
+The remaining 5 journals are all published by major academic publishers --- Elsevier, Springer, or Cambridge University Press --- and each item in the entire run of each journal has been assigned a DOI (digital object identifier) for archival purposes.
+We used the Crossref API and \texttt{rcrossref} R interface to this API \emph{{[}cite{]}} to retrieve metadata for each item published from 1960-2010 in each of these 6 journals.
+These metadata included item-level license information --- confirming that the text of each item could be used for text mining projects --- and a URL to an electronic version of the item.
+These URLs were used to retrieve an HTML or PDF version of each item, except for \emph{Personality and Individual Differences}.
+This journal has published a relatively large number of non-article documents, such as book reviews and commentaries, which are not available at the URL included in the Crossref metadata.
+(This is unfortunate, as it was not difficult to find highly relevant documents that we could not automatically retrieve and therefore excluded from our corpus \emph{{[}review{]}}.) Instead we used Elsevier's ScienceDirect API \emph{{[}url{]}} to search and retrieve all items from \emph{Personality and Individual Differences} that are available.
+PDFs were run through OCR and text extraction as necessary, as with the \emph{Mankind Quarterly} documents.
 
@@
 
-After document retrieval and text extraction, we pre-processed the text
-using the \texttt{spaCy} NLP (natural language processing) Python
-library and the R API \texttt{spacyr} \emph{{[}version, cites{]}}.
-Specifically, we applied regular expressions to remove header/footer
-copyright notices and hyphenation, used spaCy to annotate and extract
-noun phrases (eg, ``the intelligence test items''), and then cleaned and
-standardized these phrases (eg, removing the/an/a, coverting all text to
-lowercase, and replacing all whitespace with underscores:
-``intelligence\_test\_items''). We then counted the occurrence of each
-noun phrase in the document. The aggregated ``document-term matrix'' was
-stored in Parquet format \emph{{[}cite{]}} for performance reasons, and
-written and read using the \texttt{arrow} package for R
-\emph{{[}cite{]}}. \emph{{[}availability{]}}
+After document retrieval and text extraction, we pre-processed the text using the \texttt{spaCy} NLP (natural language processing) Python library and the R API \texttt{spacyr} \emph{{[}version, cites{]}}.
+Specifically, we applied regular expressions to remove header/footer copyright notices and hyphenation, used spaCy to annotate and extract noun phrases (eg, ``the intelligence test items''), and then cleaned and standardized these phrases (eg, removing the/an/a, coverting all text to lowercase, and replacing all whitespace with underscores: ``intelligence\_test\_items'').
+We then counted the occurrence of each noun phrase in the document.
+The aggregated ``document-term matrix'' was stored in Parquet format \emph{{[}cite{]}} for performance reasons, and written and read using the \texttt{arrow} package for R \emph{{[}cite{]}}.
+\emph{{[}availability{]}}
 
-NLP-extracted noun phrases offer a number of advantages over the more
-traditional unigram (``single-word'') tokens. First, noun phrase
-extraction removes many standard stopwords (articles, common verbs)
-without relying on a fixed, a priori stopword list. Phrases can be more
-informative than single terms, for example, distinguishing
-``intelligence test'' from ``hypothesis test.'' But simple n-gram
-extraction will include numerous phrases that are not especially
-meaningful. Consider the sentence ``Since our first analyses of feeding
-patterns in rats, we had been using a criterion of 40 minutes (Le Magnen
-\& Tallon 1963; 1966)'' *{[}10.1017\_\_\^{}\_\_s0140525x0000042x{]}*.
-Bigrams such as ``since our'' and ``criterion 40'' will likely be
-discarded in vocabulary selection, but significantly increase the
-computational cost of vocabulary selection. Noun phrase extraction is
-therefore more efficient.
+NLP-extracted noun phrases offer a number of advantages over the more traditional unigram (``single-word'') tokens.
+First, noun phrase extraction removes many standard stopwords (articles, common verbs) without relying on a fixed, a priori stopword list.
+Phrases can be more informative than single terms, for example, distinguishing ``intelligence test'' from ``hypothesis test.'' But simple n-gram extraction will include numerous phrases that are not especially meaningful.
+Consider the sentence ``Since our first analyses of feeding patterns in rats, we had been using a criterion of 40 minutes (Le Magnen \& Tallon 1963; 1966)'' *{[}10.1017\_\_\^{}\_\_s0140525x0000042x{]}*.
+Bigrams such as ``since our'' and ``criterion 40'' will likely be discarded in vocabulary selection, but significantly increase the computational cost of vocabulary selection.
+Noun phrase extraction is therefore more efficient.
 
-After noun phrase extraction, the corpus comprised \emph{{[}tokens{]}}
-of \emph{{[}distinct noun phrases{]}}.
+After noun phrase extraction, the corpus comprised \emph{{[}tokens{]}} of \emph{{[}distinct noun phrases{]}}.
 
 \hypertarget{data-and-code-availability}{%
-\subsubsection*{Data and code
-availability}\label{data-and-code-availability}}
+\subsubsection*{Data and code availability}\label{data-and-code-availability}}
 
-Document metadata retrieved from Crossref does not appear to be covered
-by any copyright or other intellectual property restrictions. However,
-due to copyright restrictions, we are unable to make document fulltext
-or Web of Science search results publicly available. Code for the corpus
-assembly and text preparation steps described above is available by
-request.
+Document metadata retrieved from Crossref does not appear to be covered by any copyright or other intellectual property restrictions.
+However, due to copyright restrictions, we are unable to make document fulltext or Web of Science search results publicly available.
+Code for the corpus assembly and text preparation steps described above is available by request.
 
-The public analysis repository,
-\url{https://github.com/dhicks/race-science} \emph{{[}DOI{]}}, includes
-document metadata and documentwise counts of NLP-extracted noun phrases
-(``document-term matrices''), along with code to reproduce the analysis
-of the following sections. \emph{{[}repo readme: requires R 4.1.2; rig
-(\url{https://github.com/r-lib/rig/}) for managing multiple R
-installations{]}}
+The public analysis repository, \url{https://github.com/dhicks/race-science} \emph{{[}DOI{]}}, includes document metadata and documentwise counts of NLP-extracted noun phrases (``document-term matrices''), along with code to reproduce the analysis of the following sections.
+\emph{{[}repo readme: requires R 4.1.2; rig (\url{https://github.com/r-lib/rig/}) for managing multiple R installations{]}}
 
@@
 We took an information-theoretic approach to vocabulary selection.
-Consider a game in which I draw a document from the corpus, then a
-single token from that document. I tell you the term, and you have to
-guess which document I picked. Intuitively, highly informative terms (in
-this project, noun phrase types) are distinctive, allowing you to
-dramatically narrow down the list of potential documents. This
-``informativeness'' of a term can be quantified as the KL
-(Kullback-Leibler) divergence from a ``baseline'' distribution of
-documents to the distribution conditional on the term. Because the most
-informative terms tend to be typos and OCR errors --- these are unique
-to a single document --- we multiply the KL divergence by the logarithm
-of the total number of occurrences of the term across the entire corpus.
-We refer to the resulting measure as \(log(n) \Delta H\) or simply
-\texttt{ndH}.
+Consider a game in which I draw a document from the corpus, then a single token from that document.
+I tell you the term, and you have to guess which document I picked.
+Intuitively, highly informative terms (in this project, noun phrase types) are distinctive, allowing you to dramatically narrow down the list of potential documents.
+This ``informativeness'' of a term can be quantified as the KL (Kullback-Leibler) divergence from a ``baseline'' distribution of documents to the distribution conditional on the term.
+Because the most informative terms tend to be typos and OCR errors --- these are unique to a single document --- we multiply the KL divergence by the logarithm of the total number of occurrences of the term across the entire corpus.
+We refer to the resulting measure as \(log(n) \Delta H\) or simply \texttt{ndH}.
 
-\emph{{[}Hicks 2021?{]}} used the \texttt{ndH} approach in a topic
-modeling analysis, where the baseline distribution was the uniform
-distribution across documents. In the current project, we found that
-this approach heavily favored recurrent noun phrases in the longest
-documents. Many of the documents published in \emph{Psychological
-Reports} are extremely short, 1-2 reports of a single study; while many
-of the documents published in \emph{Brain and Behavioral Sciences} are
-book-length collections that include a long review article and sometimes
-dozens of commentaries. Very generic noun phrases that happen to appear
-in the latter can occur orders of magnitude more often than highly
-distinctive phrases in the former, and so the \(log(n)\) factor
-overwhelms the \(\Delta H\) factor.
+\emph{{[}Hicks 2021?{]}} used the \texttt{ndH} approach in a topic modeling analysis, where the baseline distribution was the uniform distribution across documents.
+In the current project, we found that this approach heavily favored recurrent noun phrases in the longest documents.
+Many of the documents published in \emph{Psychological Reports} are extremely short, 1-2 reports of a single study; while many of the documents published in \emph{Brain and Behavioral Sciences} are book-length collections that include a long review article and sometimes dozens of commentaries.
+Very generic noun phrases that happen to appear in the latter can occur orders of magnitude more often than highly distinctive phrases in the former, and so the \(log(n)\) factor overwhelms the \(\Delta H\) factor.
 
-To address this, we used a different baseline distribution of documents,
-namely, one in which document probability is proportional to length.
-This makes phrases from short documents much more ``surprising'' (much
-less likely to occur according to the baseline), and hence substantially
-increases their informativeness. This was more effective at identifying
-useful phrases from across the corpus.
+To address this, we used a different baseline distribution of documents, namely, one in which document probability is proportional to length.
+This makes phrases from short documents much more ``surprising'' (much less likely to occur according to the baseline), and hence substantially increases their informativeness.
+This was more effective at identifying useful phrases from across the corpus.
 
-A common rule of thumb in topic modeling is that the vocabulary should
-have about 10 times as many distinct terms as the number of documents in
-the corpus \emph{{[}cite{]}}. However, we had some concerns with
-computational demands here: the resulting document-term matrix would
-have roughly \(10 \times 33,000^2\) or 10.9 billion entries; with 10\%
-density this would require on the order of 4 GB of memory just for a
-single copy of the matrix; and most of the analysis was to be conducted
-on the authors' laptops. We therefore chose to work with three smaller
-vocabularies, \(5 \times\), \(1 \times\) and \(\frac{1/5} \times\) the
-number of documents. We refer to these as the ``large,'' ``medium,'' and
-``small'' vocabulary, respectively, and they include \emph{{[}counts{]}}
-distinct phrases. We compare findings across vocabularies as a
-robustness check.
+A common rule of thumb in topic modeling is that the vocabulary should have about 10 times as many distinct terms as the number of documents in the corpus \emph{{[}cite{]}}.
+However, we had some concerns with computational demands here: the resulting document-term matrix would have roughly \(10 \times 33,000^2\) or 10.9 billion entries; with 10\% density this would require on the order of 4 GB of memory just for a single copy of the matrix; and most of the analysis was to be conducted on the authors' laptops.
+We therefore chose to work with three smaller vocabularies, \(5 \times\), \(1 \times\) and \(\frac{1/5} \times\) the number of documents.
+We refer to these as the ``large,'' ``medium,'' and ``small'' vocabulary, respectively, and they include \emph{{[}counts{]}} distinct phrases.
+We compare findings across vocabularies as a robustness check.
 
@@
 
-To fit topic models, we followed the approach proposed by
-\emph{{[}cite{]}}, which uses varimax-rotated partial principal
-components instead of the variational inference methods used by standard
-topic model packages \emph{{[}eg, stm{]}}. This novel approach was
-implemented in the R package \texttt{tmfast}, \emph{{[}cite{]}}. A
-simulation study of \texttt{tmfast} found that it was significantly
-faster and only slightly less accurate at reconstructing known
-word-topic and topic-document distributions, compared to the standard
-topic modeling package \texttt{stm}.
+To fit topic models, we followed the approach proposed by \emph{{[}cite{]}}, which uses varimax-rotated partial principal components instead of the variational inference methods used by standard topic model packages \emph{{[}eg, stm{]}}.
+This novel approach was implemented in the R package \texttt{tmfast}, \emph{{[}cite{]}}.
+A simulation study of \texttt{tmfast} found that it was significantly faster and only slightly less accurate at reconstructing known word-topic and topic-document distributions, compared to the standard topic modeling package \texttt{stm}.
 
-Topic models were fit for all three vocabularies (large, medium, and
-small) with \(k = 5, 10, 20, 30, 40, 50\) (number of topics), resulting
-in a total of 18 models.
+Topic models were fit for all three vocabularies (large, medium, and small) with \(k = 5, 10, 20, 30, 40, 50\) (number of topics), resulting in a total of 18 models.
 
-Following the approach of \emph{{[}Hicks 2021?{]}}, we did not attempt
-to identify a unique best model for further analysis. While we focus on
-the medium vocabulary, \(k=30\) model as the ``median'' among the 18
-models fit, we compare and contrast findings from this model with those
-from the other models.
+Following the approach of \emph{{[}Hicks 2021?{]}}, we did not attempt to identify a unique best model for further analysis.
+While we focus on the medium vocabulary, \(k=30\) model as the ``median'' among the 18 models fit, we compare and contrast findings from this model with those from the other models.
 
 \hypertarget{topic-model-interpretation}{%
-\subsubsection*{Topic model
-interpretation}\label{topic-model-interpretation}}
+\subsubsection*{Topic model interpretation}\label{topic-model-interpretation}}
 
-After fitting topic models, our first research question was whether we
-could identify distinctive ``race science'' topics. We addressed this
-question in two ways. Following common practice in topic modeling, we
-extracted lists of the top \(n\) (highest-probability) terms (noun
-phrases) from each topic in each model, with \(n\) generally ranging
-between 5 and 15 as we explored the models. \emph{{[}tables + Silge
-plots{]}}
+After fitting topic models, our first research question was whether we could identify distinctive ``race science'' topics.
+We addressed this question in two ways.
+Following common practice in topic modeling, we extracted lists of the top \(n\) (highest-probability) terms (noun phrases) from each topic in each model, with \(n\) generally ranging between 5 and 15 as we explored the models.
+\emph{{[}tables + Silge plots{]}}
 
-\emph{{[}drop{]}} We also adapted the ``discursive space'' analysis
-proposed by \emph{{[}Hicks 2021?{]}}. This analysis combines the
-Hellinger distance between topic-document distributions --- capturing
-how ``far apart'' two documents are --- with dimensional reduction
-techniques used for data visualization. This produces a two-dimensional
-visualization of individual documents, arranged so that documents with
-similar topic distributions are visually clustered and visually distinct
-from documents with different topic distributions.
+\emph{{[}drop{]}}
+We also adapted the ``discursive space'' analysis proposed by \emph{{[}Hicks 2021?{]}}.
+This analysis combines the Hellinger distance between topic-document distributions --- capturing how ``far apart'' two documents are --- with dimensional reduction techniques used for data visualization.
+This produces a two-dimensional visualization of individual documents, arranged so that documents with similar topic distributions are visually clustered and visually distinct from documents with different topic distributions.
 \emph{{[}availability{]}}
 
-Specifically, for each fitted topic model, we extracted topic-document
-distributions for each document and calculated pairwise Hellinger
-distances. We then processed these distances using the UMAP algorithm
-for dimension reduction, as implemented in the \texttt{umap} package
-\emph{{[}details; cite{]}}. To reduce computational requirements for
-these steps, we first dropped documents published in the journal
-\emph{{[}Psychological Reports{]}} (\emph{{[}x documents removed, y
-remaining{]}}; note that the full corpus was used to fit the topic
-models, \emph{Psychological Reports} was removed just for this
-visualization). Using the \texttt{plotly} library \emph{{[}cite{]}}, we
-construct an interactive visualization: each point represents a single
-document, and points can be colored based on either their journal or
-highest-probability topic. Tooltips are displayed when we hover over a
-point, including document metadata (title, journal, authors, year) as
-well as ``key phrases.''
+Specifically, for each fitted topic model, we extracted topic-document distributions for each document and calculated pairwise Hellinger distances.
+We then processed these distances using the UMAP algorithm for dimension reduction, as implemented in the \texttt{umap} package \emph{{[}details; cite{]}}.
+To reduce computational requirements for these steps, we first dropped documents published in the journal \emph{{[}Psychological Reports{]}} (\emph{{[}x documents removed, y remaining{]}}; note that the full corpus was used to fit the topic models, \emph{Psychological Reports} was removed just for this visualization).
+Using the \texttt{plotly} library \emph{{[}cite{]}}, we construct an interactive visualization: each point represents a single document, and points can be colored based on either their journal or highest-probability topic.
+Tooltips are displayed when we hover over a point, including document metadata (title, journal, authors, year) as well as ``key phrases.''
 
-To calculate key phrases for a given document \(d\), we first identified
-the highest-probability topic \(t\) for that document and the word-topic
-(noun phrase-topic) distribution \(\beta_{w,t}\) for that topic. We also
-retrieved the document-term matrix, containing the count \(n_{w,d}\) of
-occurrences for each word (noun phrase), and calculated a weighting
-\(n_{w,d} \beta_{w,t}\) for each word. The key phrases are the 5 most
-highly-weighted terms for the given document.
+To calculate key phrases for a given document \(d\), we first identified the highest-probability topic \(t\) for that document and the word-topic (noun phrase-topic) distribution \(\beta_{w,t}\) for that topic.
+We also retrieved the document-term matrix, containing the count \(n_{w,d}\) of occurrences for each word (noun phrase), and calculated a weighting \(n_{w,d} \beta_{w,t}\) for each word.
+The key phrases are the 5 most highly-weighted terms for the given document.
 
 \hypertarget{topic-quality-assessment}{%
-\subsubsection*{Topic quality
-assessment}\label{topic-quality-assessment}}
+\subsubsection*{Topic quality assessment}\label{topic-quality-assessment}}
 
@@
 
-We next conducted a topic quality check of topic 19 from the medium
-vocabulary, \(k=30\) model. Keyword and inspection of the UMAP
-interactives seemed to indicate that this topic was a ``race science
-discourse'' topic in this model.
+We next conducted a topic quality check of topic 19 from the medium vocabulary, \(k=30\) model.
+Keyword and inspection of the UMAP interactives seemed to indicate that this topic was a ``race science discourse'' topic in this model.
 
-To confirm this initial impression, a spreadsheet of all articles with
-\(\gamma > 0.25\) for this topic was extracted (excluding articles
-published in \emph{Psychological Reports}), and the top 200 articles
-were reviewed manually by both authors. For this qualitative analysis,
-we defined \emph{scientific racism} as (purporting) to justify racial
-inequality and colonialism by appealing to the epistemic authority of
-science; \emph{race science} as (pseudo-)scientific research that can be
-utilized for scientific racism; and \emph{race science discourse} as
-treating race science as a legitimate area of scientific research; this
-includes methodological critiques of race science and empirical tests
-that falsify race science hypotheses. Our qualitative review of the
-top-200 topic-19 articles focused on classifying them as \emph{race
-science discourse}.
+To confirm this initial impression, a spreadsheet of all articles with \(\gamma > 0.25\) for this topic was extracted (excluding articles published in \emph{Psychological Reports}), and the top 200 articles were reviewed manually by both authors.
+For this qualitative analysis, we defined \emph{scientific racism} as (purporting) to justify racial inequality and colonialism by appealing to the epistemic authority of science; \emph{race science} as (pseudo-)scientific research that can be utilized for scientific racism; and \emph{race science discourse} as treating race science as a legitimate area of scientific research; this includes methodological critiques of race science and empirical tests that falsify race science hypotheses.
+Our qualitative review of the top-200 topic-19 articles focused on classifying them as \emph{race science discourse}.
 
-For the first round of review, both authors worked independently,
-dichotomously coding each article as race science discourse or not. We
-calculated the ``false positive rate'' --- documents with a very high
-\(\gamma\) in this topic that were not race science discourse --- and
-interrater reliability using Cohen's \(\kappa\).\\
-\emph{{[}false positive rate{]}} \emph{{[}Initial interrater reliability
-was low, \(\kappa = 0.4\), with 50 out of 200 articles classified
-discordantly.{]}} To understand the disagreement, both authors examined
-the 50 discordant articles using an interactive grounded theory
-approach. After this process, some initial codings were changed and
-\emph{{[}almost{]}} all of the remaining \emph{{[}n{]}} discordant
-articles were coded as follows:
+For the first round of review, both authors worked independently, dichotomously coding each article as race science discourse or not.
+We calculated the ``false positive rate'' --- documents with a very high \(\gamma\) in this topic that were not race science discourse --- and interrater reliability using Cohen's \(\kappa\).\\
+\emph{{[}false positive rate{]}}
+\emph{{[}Initial interrater reliability was low, \(\kappa = 0.4\), with 50 out of 200 articles classified discordantly.{]}} To understand the disagreement, both authors examined the 50 discordant articles using an interactive grounded theory approach.
+After this process, some initial codings were changed and \emph{{[}almost{]}} all of the remaining \emph{{[}n{]}} discordant articles were coded as follows:
 
@@
 \item[hereditarianism and/or eugenics]
-29 papers. On the ``eugenics'' side, mostly examines or otherwise
-discusses the ``dysgenic'' claim that ``low-intelligence'' people have
-more children than ``high-intelligence'' people. The ``hereditarianism''
-side is more varied, with papers on environmentalist interventions to
-raise IQ, genetics of IQ, and several other sub-topics.
+29 papers.
+On the ``eugenics'' side, mostly examines or otherwise discusses the ``dysgenic'' claim that ``low-intelligence'' people have more children than ``high-intelligence'' people.
+The ``hereditarianism'' side is more varied, with papers on environmentalist interventions to raise IQ, genetics of IQ, and several other sub-topics.
 \item[Flynn effect]
-15 papers. Supporting, challenging, or otherwise focusing on Flynn's
-findings that IQ has increased over time.
+15 papers. Supporting, challenging, or otherwise focusing on Flynn's findings that IQ has increased over time.
 \item[Lynn/national IQ]
-3 papers. Supports, challenges, or focuses on international IQ
-data/comparisons made by Richard Lynn.
+3 papers. Supports, challenges, or focuses on international IQ data/comparisons made by Richard Lynn.
 \end{description}
 
-We agreed that, for all three categories of papers, there was reasonable
-ambiguity about whether they should be classified as race science
-discourse. These papers either do not explicitly mention race, or do so
-with just a brief mention of the ``race and intelligence controversy.''
-On the other hand, they either contribute to or legitimatize areas of
-research that are deeply entangled with race science, and might thereby
-contribute to public perceptions of a legitimate scientific debate about
-race and intelligence.
+We agreed that, for all three categories of papers, there was reasonable ambiguity about whether they should be classified as race science discourse.
+These papers either do not explicitly mention race, or do so with just a brief mention of the ``race and intelligence controversy.'' On the other hand, they either contribute to or legitimatize areas of research that are deeply entangled with race science, and might thereby contribute to public perceptions of a legitimate scientific debate about race and intelligence.
 
-Based on this qualitative review, we judged that the topic model
-approach was sufficiently sensitive (low ``false positive'' rate) in
-identifying race science discourse.
+Based on this qualitative review, we judged that the topic model approach was sufficiently sensitive (low ``false positive'' rate) in identifying race science discourse.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-BarkanRetreatScientificRacism1992}{}}%
-Barkan, Elazar. 1992. \emph{The {Retreat} of {Scientific Racism}:
-{Changing Concepts} of {Race} in {Britain} and the {United States
-Between} the {World Wars}}. {Cambridge University Press}.
+Barkan, Elazar. 1992. \emph{The {Retreat} of {Scientific Racism}: {Changing Concepts} of {Race} in {Britain} and the {United States Between} the {World Wars}}. {Cambridge University Press}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-BeagleholeRaceQuestion1950}{}}%
-Beaglehole, Ernest, Juan Comas, L. A. Costa Pinto, Franklin Frazier,
-Morris Ginsberg, Humayun Kabir, Claude Levi-Strauss, and Ashley Montagu.
-1950. {``The {Race Question}.''} \emph{UNESCO and Its Programme} 1 (3).
+Beaglehole, Ernest, Juan Comas, L.
+A.
+Costa Pinto, Franklin Frazier, Morris Ginsberg, Humayun Kabir, Claude Levi-Strauss, and Ashley Montagu.
+1950.
+{``The {Race Question}.''} \emph{UNESCO and Its Programme} 1 (3).
 
 \leavevmode\vadjust pre{\hypertarget{ref-BrattainRaceRacismAntiracism2007}{}}%
-Brattain, M. 2007. {``Race, {Racism}, and {Antiracism}: {UNESCO} and the
-{Politics} of {Presenting Science} to the {Postwar Public}.''} \emph{The
-American Historical Review} 112 (5): 1386--1413.
+Brattain, M.
+2007.
+{``Race, {Racism}, and {Antiracism}: {UNESCO} and the {Politics} of {Presenting Science} to the {Postwar Public}.''} \emph{The American Historical Review} 112 (5): 1386--1413.
 \url{https://doi.org/10.1086/ahr.112.5.1386}.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-MeadComingAgeSamoa1928}{}}%
-Mead, Margaret. 1928. \emph{Coming of {Age} in {Samoa}}. {New York}:
-{William Morrow and Company}.
+Mead, Margaret. 1928. \emph{Coming of {Age} in {Samoa}}. {New York}: {William Morrow and Company}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-MehlerFoundationFascismNew1989}{}}%
-Mehler, Barry. 1989. {``Foundation for Fascism: {The} New Eugenics
-Movement in the {United States}.''} \emph{Patterns of Prejudice} 23 (4):
-17--25. \url{https://doi.org/10.1080/0031322X.1989.9970026}.
+Mehler, Barry. 1989. {``Foundation for Fascism: {The} New Eugenics Movement in the {United States}.''} \emph{Patterns of Prejudice} 23 (4): 17--25. \url{https://doi.org/10.1080/0031322X.1989.9970026}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-SlobodianUnequalMindHow2023}{}}%
-Slobodian, Quinn. 2023. {``The {Unequal Mind}: {How Charles Murray} and
-{Neoliberal Think Tanks Revived IQ}.''} \emph{Capitalism: A Journal of
-History and Economics} 4 (1): 73--108.
+Slobodian, Quinn. 2023. {``The {Unequal Mind}: {How Charles Murray} and {Neoliberal Think Tanks Revived IQ}.''} \emph{Capitalism: A Journal of History and Economics} 4 (1): 73--108.
 

### 2023-07-14 paper/paper.Rmd
mostly filling in limitations; need to ensure stay below ~3k words

@@
 
+*[Pioneer]*
+
+*[probably don't include this]*
 - pre-20th century history of race concepts and race science
@@
 
-# Race science 1910-1960
+<!-- # Race science 1910-1960 -->
 
+<!-- 
 ## The rise and fall of eugenics
@@
 - 1950: "For all practical social purposes 'race' is not so much a biological phenomenon as a social myth" [@BeagleholeRaceQuestion1950; though see @BrattainRaceRacismAntiracism2007]
+ -->
 
-## *Brown v Board* and *Mankind Quarterly*
+# *Brown v Board* and *Mankind Quarterly*
 
@@
 
-In the landmark case *Brown v Board of Education of Topeka* (1954), the US Supreme Court had banned *de jure* educational segregation.
+In the landmark case *Brown v Board of Education of Topeka* (1954), the US Supreme Court banned *de jure* educational segregation.
 The Court's decision relied on expert testimony from psychologists and education researchers; but the segregationists had also put forward their own experts, including Henry Garrett \cite{WinstonScienceServiceFar1998, JacksonScienceSegregationRace2005, SchafferScientificRacismAgain2007}.
-Shortly after *Brown*, Garrett left Columbia, taking a visiting position at the University of Virginia, and becoming increasingly involved in segregationist, eugenicist, and even the neofascist Northern League.
+Shortly after *Brown*, Garrett left Columbia and became involved in segregationist and eugenicist efforts, and even the neofascist Northern League.
 
-In the context of *Brown*, MQ was created as what contemporary scholars call an "echo chamber" \cite{FernandezPintoKnowBetterNot2017}, a favorable venue for race scientists to publish their views, on the grounds that an "equalitarian dogma" created a censorious "taboo" against their research in mainstream publications \cite{TuckerFundingScientificRacism2002, JacksonMythicalTabooRace2020}.
-Historians, philosophers, and sociologists of science have shown that echo chambers have been a key part of the "tobacco strategy," used by numerous regulated industries — most infamously tobacco, but also fossil fuels and chemical manufacturing, among others — to "manufacture doubt" and delay regulation to protect public and environmental health *[cites]*.
+As part of the segregationist reaction to *Brown*, MQ was created as what contemporary historians and philosophers of science call an "echo chamber" \cite{FernandezPintoKnowBetterNot2017}, a favorable venue for race scientists to publish their views, on the grounds that an "equalitarian dogma" created a censorious "taboo" against their research in mainstream publications \cite{TuckerFundingScientificRacism2002, JacksonMythicalTabooRace2020}.
+Scholars have shown that echo chambers are a key part of the "tobacco strategy," used by numerous regulated industries — most infamously tobacco, but also fossil fuels and chemical manufacturing, among others — to "manufacture doubt" and delay regulation *[cites]*.
 
-In light of the origins of MQ, the significant attention paid to the journal by historians of scientific racism, and contemporary research on the "tobacco strategy," it is surprising that MQ did not play a significant role in nurturing race science.
+In light of the origins of MQ, the significant attention paid to the journal by historians of scientific racism, and contemporary research on the "tobacco strategy," we hypothesized that bibliometric and text mining analyses would confirm that MQ functioned as an echo chamber for Pioneer-funded scientists.
+This hypothesis was not supported.
+Instead, mainstream psychology journals served as the echo chambers for race science.
 
@@
 
-We identified 16 researchers who had received funding from Pioneer; 14 of these researchers had profiles in the Web of Science (WoS) author search *[table + WoS pub coucnt]*, allowing us identify 13 WoS-indexed journals that had published 6 or more of these authors.
+We identified 16 researchers who had received funding from Pioneer; 14 of these researchers had profiles in the Web of Science (WoS) author search, allowing us identify 13 WoS-indexed journals that had published 6 or more of these authors.
 See Table \@ref(tab:researchers).
@@
 | | | |
-|:-----|-----:|:----|
-| Thomas J. Bouchard, Jr. | psychology | |
-| Brunetto Chiarelli | anthropology | |
-| Hans Eysenck | psychology | |
-| Robert Gordon | sociology | |
-| Linda Gottfredson | psychology | |
-| Garrett Hardin | ecology | |
-| Joseph M. Horn | psychology | |
-| Lloyd Humphreys | psychology | |
-| Arthur Jensen | psychology | |
-| Michael Levin | philosophy | |
-| Richard Lynn | psychology | |
-| R. Travis Osborne | psychology | |
-| J. Phillippe Rushton | psychology | |
-| Audrey M. Shuey | psychology | |
-| Philip A. Vernon | psychology | |
-| Daniel Vining, Jr. | demography | |
+|:-----|-----:|----:|
+| Thomas J. Bouchard, Jr. (1937-) | psychology | 184 |
+| Brunetto Chiarelli (?-?) | anthropology | 91 |
+| Hans Eysenck (1916-1997) | psychology | 661 |
+| Robert A. Gordon (1932-) | sociology | 0|
+| Linda Gottfredson (1947-) | psychology | 71 |
+| Garrett Hardin (1915-2003) | ecology | 75 |
+| Joseph M. Horn (1940-) | psychology | 68 |
+| Lloyd Humphreys (1913-2003) | psychology | 0 |
+| Arthur Jensen (1923-2003) | psychology | 235 |
+| Michael Levin (1943-) | philosophy | 96 |
+| Richard Lynn (1930-) | psychology | 288 |
+| R. Travis Osborne (1913-2013) | psychology | 59 |
+| J. Phillippe Rushton (1943-2012) | psychology | 277 |
+| Audrey M. Shuey (1900-1977) | psychology | 10 |
+| Philip A. Vernon (1950-) | psychology | 227 |
+| Daniel Vining, Jr. (1944-) | demography | 33 |
 
-Table: (\#tab:researchers) Pioneer-funded researchers.  Either identified in *[cite]* or named on Pioneer's website, along with identified discipline and WoS author search result counts.
+Table: (\#tab:researchers) Pioneer-funded researchers.
+Either identified in *[cite]* or named on Pioneer's website, along with birth and death dates from Wikipedia, identified discipline, and WoS author search result counts.
+Brunetto Chiarelli does not have a Wikipedia page.
 
@@
 
-# Topic model analysis
+# Topic model analysis identifies race science discourse
 
@@
 We assembled a full-text corpus of articles published in MQ, PID, I, BG, PR, and Behavior and Brain Sciences (BBS) between 1960-2010 and used topic modeling to identify topics discussing race, intelligence, and both.
-*[tables: Ns for each article, topic model parameters; figs: articles by pub over time, topic heatmaps, Silge plots, race-intelligence-both grids]*
+*[supplemental tables: Ns for each article, topic model parameters; supplemental figs: articles by pub over time, topic heatmaps, Silge plots, race-intelligence-both grids]*
 
-![Silge plots and smoothed time series for three focal topics.
+![Silge plots \cite{SilgeTopicModeling2017} and smoothed time series for three focal topics.
 Top row: top 15 terms by $\beta$ (term-topic distribution) for each topic.
@@
 Thin lines give annual values, thick lines give 5-year running averages.
-Because *Behavior Genetics* and *Behavioral and Brain Sciences* are not prominent in any panel, neither is given a direct label.](img/focal_topics.png){width="4.76in" height="3.57in"}
+Because *Behavior Genetics* and *Behavioral and Brain Sciences* are not prominent in any panel, neither is given a direct label.
+(\#fig:focal)](img/focal_topics.png){width="4.76in" height="3.57in"}
 
-We focus here on three topics identified in one of the 24 fitted topic models *[composite fig: topics over time + Silge plots]*.
+We focus here on three topics identified in one of the 24 fitted topic models; see Figure \@(fig:focal).
 In this model, topic 07 is strongly associated with MQ:  MQ published dozens of articles in this topic each year, and no other journal ever published more than a handful, with most other journals publishing none in a given year.
-The Silge plot contains racial terms (`races`, `whites`, `negroes` and potentially `europe`, `africa`, `india`, and `japan`) as well as `book`, likely reflecting the fact that MQ published a number of book reviews while the psychology journals did not.
+The Silge plot \cite{SilgeTopicModeling2017} contains racial terms (`races`, `whites`, `negroes` and potentially `europe`, `africa`, `india`, and `japan`) as well as `book`, likely reflecting the fact that MQ published a number of book reviews while the psychology journals did not.
 
@@
 
-The Silge plot for topic 24, by contrast, suggests a topic focused squarely on race and intelligence, with multiple racial terms (`whites`, `blacks`, `racial_differences`, `races`, `race_differences`) and the names of three prominent Pioneer-funded researchers, `jensen`, `lynn`, and `rushton`.
-To confirm this impression from the Silge plot, we independently coded the top 121 papers in this topic (those with $gamma > 0.97$) as contributing to race science discourse or not.
+The Silge plot for topic 24, by contrast, suggests a distinct race science discourse topic, with multiple racial terms (`whites`, `blacks`, `racial_differences`, `races`, `race_differences`) and the names of three prominent Pioneer-funded researchers, `jensen`, `lynn`, and `rushton`.
+To check this interpretation, we independently coded the top 121 papers in this topic (those with $gamma > 0.97$) as contributing to race science discourse or not.
 105 (87%) were coded as race science discourse by both authors; 10 (8.3%) were coded as not race science discourse by both; and discordant codes were given to the remaining 6 (5.0%) (95% agreement, Cohen's $\kappa = 0.74$).
@@
 PID published multiple papers in topic 24 almost immediately after it was founded, with I showing a more gradual increase between the mid-1970s and mid-1990s.
+
+# Race science and behavior genetics
 
@@
 
-The topic model analysis and the case of Whitney's address suggest that, by the 1990s, behavior genetics had at least somewhat distanced itself from race science, though perhaps without repudiating it.
-*[https://link.springer.com/search?use-oscar-shared-search=true&sortBy=newestFirst&dateTo=2010&sortOrder=newestFirst&facet-journal-id=10519&content-type=Article&query=race+intelligence&date=custom&dateFrom=&date-facet-mode=between&facet-start-year=1972&previous-start-year=1972&facet-end-year=2010&previous-end-year=2023]*
+The topic model analysis and the case of Whitney's address suggest that, by the 1990s, behavior genetics may have somewhat distanced itself from race science, though perhaps without directly repudiating it \cite{GillbornSoftlySoftlyGenetics2016, HennWhyDNANo2021}.
 
+*[move most of this to supplement?]*
 To check this interpretation, we ran an independent search for documents with the keywords "race" and "intelligence" published in BG from 1972 to 2020, using Springer's journal search website.
@@
 14 of the remaining documents were published since 2010 (after the scope of the primary study).
-Among these, one obituary and two papers celebrated the work of John Loehlin, a prominent race-and-intelligence researcher who had been Director of the American Eugenics Society from 1968-1972 *[http://link.springer.com/article/10.1007/s10519-020-10024-w; http://link.springer.com/article/10.1007/s10519-014-9687-1; http://link.springer.com/article/10.1007/s10519-014-9673-7]*.
+Among these, one obituary and two papers celebrated the work of John Loehlin, a prominent race-and-intelligence researcher who had been Director of the American Eugenics Society from 1968-1972 \cite{JohnLoehlin19262020, WaldmanIntroductionFestschriftJohn2014, PlominGenotypeEnvironmentCorrelationEra2014}.
 Another was a retrospective of the work of Lindon Eaves, a geneticist who made at least one notable contribution to debates on the Scarr-Rowe hypothesis (interactions between heritability, race, and class).
@@
 28 documents in this sample were published between 1990 and 2010.
-One was an obituary of Jerry Hirsch, a critic of hereditarianism in general and Jensen in particular *[http://link.springer.com/article/10.1007/s10519-008-9231-2]*, and another was an obituary of David Rowe *[http://link.springer.com/article/10.1023/A%3A1026122928868]*.
-Only 2 of these 28 documents reported race differences of any kind: one examining at interactions among race, sex, and heritability for adolescent BMI (Body Mass Index, used as a measure of overweight/obesity) *[http://link.springer.com/article/10.1023/A%3A1021619329904]*; and the other at interactions among race, family history of alcoholism, and visuospatial performance *[http://link.springer.com/article/10.1007/BF02197241]*.
+One was an obituary of Jerry Hirsch, a critic of hereditarianism in general and Jensen in particular \cite{RoubertouxJerryHirsch202008}, and another was an obituary of David Rowe \cite{RodgersObituaryDavidChristian2003}.
+Only 2 of these 28 documents reported race differences of any kind: one examining at interactions among race, sex, and heritability for adolescent BMI (Body Mass Index, used as a measure of overweight/obesity) \cite{JacobsonGeneticSharedEnvironmental1998}; and the other at interactions among race, family history of alcoholism, and visuospatial performance \cite{BermanReducedVisuospatialPerformance1995}.
 Finally, Philip Vernon — one of the Pioneer-funded scientists we identified earlier — published a critical review of a book that attempted to address hereditarianism, the Flynn effect, race-and-intelligence research, and some related issues.
@@
 None of the documents examined in this supplemental review explicitly repudiate race science.
-Behavior geneticists seem to typically adopt a posture of objectivity and technical rigor, downplaying the extent of the field's involvement in race science in the past.
+Similarly, another disciplinary history of behavior genetics, published by Plomin in 2023 \cite{PlominCelebratingCenturyResearch2023}, briefly mentions the controversies surrounding Jensen's 1969 paper and later *The Bell Curve*, writing that Jensen's work was controversial because he "touched on ethnic differences"; though ultimately Plomin reassures the reader that "by the 1990s, quantitative genetic research had convinced most scientists of the importance of genetics for behavioral traits, including intelligence."  In short, behavior geneticists seem to typically adopt a posture of objectivity and technical rigor, downplaying the extent of the field's involvement in race science in the past, and displacing the source of race science controversies to external, "political," and "irrational" actors.
 
@@
 
-We stress that we are not identifying any particular scientist as a racist based on the topic model results.
-Scientific research can be utilized to promote scientific racism — and thus count as race science — independently of the views and attitudes of the scientist who conducted that research.\cite{TaberyWhyStudyingGenetics2015, CarlsonQuantifyingContextualizingImpact2020, HennWhyDNANo2021}  But this does not mean that scientists are not responsible for the downstream social effects of their
+*[for discussion: the possibility for white supremacist appropriation, ie, race science done by scientists who are not themselves scientific racists: CarlsonQuantifyingContextualizingImpact2020; probably some reviews of Paige Harden; O'Connor and Weatherall]*
+
+We stress that we are not identifying any particular scientist as a racist, scientific racist, or race scientist based on the topic model results.
+Scientific research can be utilized to promote scientific racism — and thus count as race science — independently of the views and attitudes of the scientist who conducted that research.\cite{TaberyWhyStudyingGenetics2015, GillbornSoftlySoftlyGenetics2016, CarlsonQuantifyingContextualizingImpact2020, HennWhyDNANo2021}  But this does not mean that scientists are not responsible for the downstream social effects of their
 research, even when these effects involve deliberate misinterpretation.*[Douglas, Kitcher, Kourany, Havstad; Block and Dworkin?]*  Like other citizens, scientists are responsible for mitigating the reasonably foreseen harms that result from their actions.
-And we now have decades of empirical evidence that research on the genetics of intelligence — even research that does not itself touch on race — can and will be used to promote scientific racism.\cite{MeyerWrestlingSocialBehavioral2023}
+And we now have decades of empirical evidence that research on the genetics of intelligence — even research that does not itself touch on race — can and will be used to promote scientific racism \cite{MeyerWrestlingSocialBehavioral2023}.
+Specifically, while our findings suggest that behavior genetics has attempted to distance itself from scientific racism, behavior genetics research — and research on human genetics more generally — may be especially susceptible to appropriation by scientific racists \cite{CarlsonQuantifyingContextualizingImpact2020, OgbunugaforDNABasketballBirthday2022}.
+A colorblind attitude of "we don't talk about race anymore" *[cite Bonilla-Silva?]* is likely insufficient to prevent particular research findings being utilized as race science.
 
-*[engagement dilemma: technical critiques can legitimize]*
+Our distinction between race science and race science discourse highlights an important "engagement dilemma" for scientists challenging race science and scientific racism.
+While the "tobacco strategy" for manufacturing doubt involves targeting mainstream scientists and promoting fringe scientific views, its primary audiences are lawmakers and the general public.
+The aim of the strategy is not to change the opinion of mainstream scientists, but instead to cultivate the impression that there is legitimate scientific uncertainty about the issue, reasonable disagreement among qualified experts *[cites]*.
+On the one hand, refusing to engage with fringe scientists — or, often, pseudoscientists \cite{BhakthavatsalamVirtueEpistemologicalApproach2021} — means that false or misleading statements go unchallenged.
+On the other hand, detailed, technical criticism legitimizes the fringe views, creating exactly the impression of reasonable scientific disagreement that the tobacco strategy is intended to produce.
 
-*[limitation: colorblind racism (Bonilla-Silva; WinstonWhyMainstreamResearch2020) creates coding ambiguities]*
+In the case of race science, mainstream psychologists have often taken a non-engagement approach — whether due to a strong conception of academic freedom or a desire to avoid stoking controversy — with the result that race scientists have been able to present themselves as the representatives of "mainstream science on intelligence."  On the other hand, critics — most prominently Gould, Kamin, and Lewontin — have offered sophisticated technical critiques.
+But these tend to spiral off into esoteric debates that are inaccessible to the general public, as with "Lewontin's fallacy" \cite{RosemanLewontinDidNot2021}.
+Consequently these critiques are likely to be either ignored by the general public or taken as evidence of legitimate scientific debate.
+*[more?]*
 
-*[limitation: false negatives (race science papers that don't appear in V24) are possible if, eg, they use many terms related to twin studies of heritability and thereby appear in V03]*
+## Limitations
 
+A key limitation of the current study stems from the inability of text mining techniques to draw on a broader cultural context than what is represented in the corpus.
+Scholars of racism and US race relations have noted a shift in the way racist attitudes are understood and expressed, from *[something]* to "colorblind racism" *[Bonilla-Silva]*, defined as *[defn]*.
+Colorblind racism involves the use of racialized "codes" or "dogwhistles," language that is superficially non-racial but carries racial implications, such as (in the US context) "violent inner city criminals" or "welfare queen" \cite{SaulDogwhistlesPoliticalManipulation2018}.
+Alderfer notes that the racist implications of *The Bell Curve* only emerge gradually, with the opening of the book using the colorblind language of "group differences" \cite{AlderferScienceNonsciencePsychologists2003}.
+Panofsky et al.
+show how the term "human biodiversity," originally developed for anti-racist purposes by biological anthropologist Jonathan Marks, has been inverted by white supremacists and is now a dogwhistle for biological racial hierarchy \cite{PanofskyHowWhiteNationalists2021}.
+These cases suggest further that colorblindness can facilitate white supremacist appropriation of non-racial research — even contrary to the researchers' own intentions — as scientific racist readers re-interpret racial-neutral language as dogwhistles \cite{GillbornSoftlySoftlyGenetics2016, WillsAreClustersRaces2017}.
+
+Techniques such as topic modeling may be able to account for discursive shifts from explicit to coded racial language, but only if the corpus contains a critical mass of documents using both sets of terms together.
+This would seem to require either a relatively gradual transition, the inclusion of reflective commentary, such as scholarship identifying and analyzing the transition, or both.
+It is unlikely that these requirements are met by the corpus used in this study.
+So it is plausible that the topic models used here have some rate of "false negatives," documents in the corpus that engage in race science discourse but in a colorblind way that avoids the use of the explicit racial terms in the race-and-intelligence topic.
+
+The phenomenon of colorblind racism is closed related to ambiguities that cannot be resolved even by human coders familiar with the broader cultural context.
+Two relevant examples are the "Flynn effect," a secular increase in intelligence test scores noted by philosopher James Flynn; and studies of "national IQ," exemplified by the work of Richard Lynn.
+The Flynn effect implies that environmental factors can create group differences in IQ that are comparable to Black-White differences, and thus directly challenges claims that the racial differences are biological.
+Using methods widely regarded as ad hoc, cherry-picking, and generally unreliable, Lynn and collaborators have claimed that national IQ averages are correlated with national GDP.
+Both the Flynn effect and "national IQ" are strongly associated with race-and-intelligence discourse in the US cultural context.
+But scientific journal articles on these topics may not include any racial language at all.
+For example, a meta-analysis of the Flynn effect does not discuss racial differences or use terms such as "White" or "Black" outside the reference list \cite{TrahanFlynnEffectMetaanalysis2014}.
+Should such articles be coded as race science discourse?
+We expect that even readers who accept our definitions of race science and race science discourse will reasonably disagree on how to answer this question.
+Topic models, of course, are not even capable of representing this essential ambiguity.
+
+<!--
 *[replace w/: WinstonWhyMainstreamResearch2020: race science as a community project of psychology(?).
@@
 
-Using agent-based models, Weatherall et al.\cite{WeatherallHowBeatScience2018} show that a strategy of "selective sharing" — promoting favorable results from independent research — can be effective at misleading a target audience.
-
-*[for discussion: the possibility for white supremacist appropriation, ie, race science done by scientists who are not themselves scientific racists: CarlsonQuantifyingContextualizingImpact2020; probably some reviews of Paige Harden; O'Connor and Weatherall]*
+-->
 

### 2023-07-14 paper/paper.tex
mostly filling in limitations; need to ensure stay below ~3k words

@@
 % Pandoc citation processing
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-% for Pandoc 2.8 to 2.10.1
-\newenvironment{cslreferences}%
-  {}%
-  {\par}
-% For Pandoc 2.11+
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
-  % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
- }%
- {}
-\usepackage{calc} % for calculating minipage widths
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 
-\begin{itemize}
-\tightlist
-\item
-  pre-20th century history of race concepts and race science
-\item
-  expansion and contraction of racialized terms over time
-\end{itemize}
+\emph{{[}Pioneer{]}}
 
-\hypertarget{race-science-1910-1960}{%
-\section*{Race science 1910-1960}\label{race-science-1910-1960}}
-
-\hypertarget{the-rise-and-fall-of-eugenics}{%
-\subsection*{The rise and fall of eugenics}\label{the-rise-and-fall-of-eugenics}}
-
-\emph{{[}write this{]}}
-
-\begin{itemize}
-\tightlist
-\item
-  scientific racism is thoroughly entangled with eugenics, but still conceptually distinct (Slobodian 2023)
-
-  \begin{itemize}
-  \tightlist
-  \item
-    sterilization laws focused on categories of criminality and mental ability, rather than race per se
-  \item
-    contemporary defenders of eugenics at least sometimes claim to oppose racism
-  \item
-    claims of innate racial differences in intelligence can be made without dysgenic prophecies of a rapidly-expanding biological underclass
-  \end{itemize}
-\item
-  1910: Davenport founds Eugenics Records Office at Cold Spring Harbor
-\item
-  1926: NRC ``Committee on the Negro'' includes both Davenport and Boas: ``neither was influential enough to veto his rival's participation'' (Barkan 1992, 113)
-\item
-  1930s: \emph{Coming of Age in Samoa} (Mead 1928) popularizes cultural anthropology, ``free{[}ing{]} anthropology from the shackles of biology'' (Barkan 1992, 134)
-\item
-  1937: Pioneer Fund is founded to ``support academic research and the '{[}sic{]}dissemination of information, into the `problem of heredity and eugenics' and `the problems of race betterment'\,'' (Mehler 1989, 21, quoting Laughlin)
-\item
-  1940: Carnegie Institute defunds Eugenics Records Office
-\item
-  1950: ``For all practical social purposes `race' is not so much a biological phenomenon as a social myth'' (Beaglehole et al. 1950; though see Brattain 2007)
-\end{itemize}
+\emph{{[}probably don't include this{]}}
+- pre-20th century history of race concepts and race science
+- expansion and contraction of racialized terms over time
 
 \hypertarget{brown-v-board-and-mankind-quarterly}{%
-\subsection*{\texorpdfstring{\emph{Brown v Board} and \emph{Mankind Quarterly}}{Brown v Board and Mankind Quarterly}}\label{brown-v-board-and-mankind-quarterly}}
+\section*{\texorpdfstring{\emph{Brown v Board} and \emph{Mankind Quarterly}}{Brown v Board and Mankind Quarterly}}\label{brown-v-board-and-mankind-quarterly}}
 
@@
 
-In the landmark case \emph{Brown v Board of Education of Topeka} (1954), the US Supreme Court had banned \emph{de jure} educational segregation.
+In the landmark case \emph{Brown v Board of Education of Topeka} (1954), the US Supreme Court banned \emph{de jure} educational segregation.
 The Court's decision relied on expert testimony from psychologists and education researchers; but the segregationists had also put forward their own experts, including Henry Garrett \cite{WinstonScienceServiceFar1998, JacksonScienceSegregationRace2005, SchafferScientificRacismAgain2007}.
-Shortly after \emph{Brown}, Garrett left Columbia, taking a visiting position at the University of Virginia, and becoming increasingly involved in segregationist, eugenicist, and even the neofascist Northern League.
+Shortly after \emph{Brown}, Garrett left Columbia and became involved in segregationist and eugenicist efforts, and even the neofascist Northern League.
 
-In the context of \emph{Brown}, MQ was created as what contemporary scholars call an ``echo chamber'' \cite{FernandezPintoKnowBetterNot2017}, a favorable venue for race scientists to publish their views, on the grounds that an ``equalitarian dogma'' created a censorious ``taboo'' against their research in mainstream publications \cite{TuckerFundingScientificRacism2002, JacksonMythicalTabooRace2020}.
-Historians, philosophers, and sociologists of science have shown that echo chambers have been a key part of the ``tobacco strategy,'' used by numerous regulated industries --- most infamously tobacco, but also fossil fuels and chemical manufacturing, among others --- to ``manufacture doubt'' and delay regulation to protect public and environmental health \emph{{[}cites{]}}.
+In the context of \emph{Brown}, MQ was created as what contemporary historians and philosophers of science call an ``echo chamber'' \cite{FernandezPintoKnowBetterNot2017}, a favorable venue for race scientists to publish their views, on the grounds that an ``equalitarian dogma'' created a censorious ``taboo'' against their research in mainstream publications \cite{TuckerFundingScientificRacism2002, JacksonMythicalTabooRace2020}.
+Scholars have shown that echo chambers are a key part of the ``tobacco strategy,'' used by numerous regulated industries --- most infamously tobacco, but also fossil fuels and chemical manufacturing, among others --- to ``manufacture doubt'' and delay regulation \emph{{[}cites{]}}.
 
-In light of the origins of MQ, the significant attention paid to the journal by historians of scientific racism, and contemporary research on the ``tobacco strategy,'' it is surprising that MQ did not play a significant role in nurturing race science.
+In light of the origins of MQ, the significant attention paid to the journal by historians of scientific racism, and contemporary research on the ``tobacco strategy,'' we hypothesized that bibliometric and text mining analyses would confirm that MQ functioned as an echo chamber.
+This hypothesis was not supported.
+Instead, mainstream psychology journals served as the echo chambers for race science.
 
@@
 
-We identified 16 researchers who had received funding from Pioneer; 14 of these researchers had profiles in the Web of Science (WoS) author search \emph{{[}table + WoS pub coucnt{]}}, allowing us identify 13 WoS-indexed journals that had published 6 or more of these authors.
+We identified 16 researchers who had received funding from Pioneer; 14 of these researchers had profiles in the Web of Science (WoS) author search, allowing us identify 13 WoS-indexed journals that had published 6 or more of these authors.
 See Table \ref{tab:researchers}.
 
-\begin{longtable}[]{@{}lrl@{}}
+\begin{longtable}[]{@{}lrr@{}}
 \caption{\label{tab:researchers} Pioneer-funded researchers.
-Either identified in \emph{{[}cite{]}} or named on Pioneer's website, along with identified discipline and WoS author search result counts.}\tabularnewline
+Either identified in \emph{{[}cite{]}} or named on Pioneer's website, along with birth and death dates from Wikipedia, identified discipline, and WoS author search result counts.
+Brunetto Chiarelli does not have a Wikipedia page.}\tabularnewline
 \toprule
 \endhead
-Thomas J. Bouchard, Jr. & psychology & \\
-Brunetto Chiarelli & anthropology & \\
-Hans Eysenck & psychology & \\
-Robert Gordon & sociology & \\
-Linda Gottfredson & psychology & \\
-Garrett Hardin & ecology & \\
-Joseph M. Horn & psychology & \\
-Lloyd Humphreys & psychology & \\
-Arthur Jensen & psychology & \\
-Michael Levin & philosophy & \\
-Richard Lynn & psychology & \\
-R. Travis Osborne & psychology & \\
-J. Phillippe Rushton & psychology & \\
-Audrey M. Shuey & psychology & \\
-Philip A. Vernon & psychology & \\
-Daniel Vining, Jr. & demography & \\
+Thomas J. Bouchard, Jr.~(1937-) & psychology & 184 \\
+Brunetto Chiarelli (?-?) & anthropology & 91 \\
+Hans Eysenck (1916-1997) & psychology & 661 \\
+Robert A. Gordon (1932-) & sociology & 0 \\
+Linda Gottfredson (1947-) & psychology & 71 \\
+Garrett Hardin (1915-2003) & ecology & 75 \\
+Joseph M. Horn (1940-) & psychology & 68 \\
+Lloyd Humphreys (1913-2003) & psychology & 0 \\
+Arthur Jensen (1923-2003) & psychology & 235 \\
+Michael Levin (1943-) & philosophy & 96 \\
+Richard Lynn (1930-) & psychology & 288 \\
+R. Travis Osborne (1913-2013) & psychology & 59 \\
+J. Phillippe Rushton (1943-2012) & psychology & 277 \\
+Audrey M. Shuey (1900-1977) & psychology & 10 \\
+Philip A. Vernon (1950-) & psychology & 227 \\
+Daniel Vining, Jr.~(1944-) & demography & 33 \\
 \bottomrule
@@
 
-\hypertarget{topic-model-analysis}{%
-\section*{Topic model analysis}\label{topic-model-analysis}}
+\hypertarget{topic-model-analysis-identifies-race-science-discourse}{%
+\section*{Topic model analysis identifies race science discourse}\label{topic-model-analysis-identifies-race-science-discourse}}
 
@@
 We assembled a full-text corpus of articles published in MQ, PID, I, BG, PR, and Behavior and Brain Sciences (BBS) between 1960-2010 and used topic modeling to identify topics discussing race, intelligence, and both.
-\emph{{[}tables: Ns for each article, topic model parameters; figs: articles by pub over time, topic heatmaps, Silge plots, race-intelligence-both grids{]}}
+\emph{{[}supplemental tables: Ns for each article, topic model parameters; supplemental figs: articles by pub over time, topic heatmaps, Silge plots, race-intelligence-both grids{]}}
 
@@
 \includegraphics[width=4.76in,height=3.57in]{img/focal_topics.png}
-\caption{Silge plots and smoothed time series for three focal topics.
+\caption{Silge plots \cite{SilgeTopicModeling2017} and smoothed time series for three focal topics.
 Top row: top 15 terms by \(\beta\) (term-topic distribution) for each topic.
@@
 Thin lines give annual values, thick lines give 5-year running averages.
-Because \emph{Behavior Genetics} and \emph{Behavioral and Brain Sciences} are not prominent in any panel, neither is given a direct label.}
+Because \emph{Behavior Genetics} and \emph{Behavioral and Brain Sciences} are not prominent in any panel, neither is given a direct label.
+\label{fig:focal}}
 \end{figure}
 
-We focus here on three topics identified in one of the 24 fitted topic models \emph{{[}composite fig: topics over time + Silge plots{]}}.
+We focus here on three topics identified in one of the 24 fitted topic models; see Figure @(fig:focal).
 In this model, topic 07 is strongly associated with MQ: MQ published dozens of articles in this topic each year, and no other journal ever published more than a handful, with most other journals publishing none in a given year.
-The Silge plot contains racial terms (\texttt{races}, \texttt{whites}, \texttt{negroes} and potentially \texttt{europe}, \texttt{africa}, \texttt{india}, and \texttt{japan}) as well as \texttt{book}, likely reflecting the fact that MQ published a number of book reviews while the psychology journals did not.
+The Silge plot \cite{SilgeTopicModeling2017} contains racial terms (\texttt{races}, \texttt{whites}, \texttt{negroes} and potentially \texttt{europe}, \texttt{africa}, \texttt{india}, and \texttt{japan}) as well as \texttt{book}, likely reflecting the fact that MQ published a number of book reviews while the psychology journals did not.
 
@@
 
-The Silge plot for topic 24, by contrast, suggests a topic focused squarely on race and intelligence, with multiple racial terms (\texttt{whites}, \texttt{blacks}, \texttt{racial\_differences}, \texttt{races}, \texttt{race\_differences}) and the names of three prominent Pioneer-funded researchers, \texttt{jensen}, \texttt{lynn}, and \texttt{rushton}.
-To confirm this impression from the Silge plot, we independently coded the top 121 papers in this topic (those with \(gamma > 0.97\)) as contributing to race science discourse or not.
+The Silge plot for topic 24, by contrast, suggests a distinct race science discourse topic, with multiple racial terms (\texttt{whites}, \texttt{blacks}, \texttt{racial\_differences}, \texttt{races}, \texttt{race\_differences}) and the names of three prominent Pioneer-funded researchers, \texttt{jensen}, \texttt{lynn}, and \texttt{rushton}.
+To check this interpretation, we independently coded the top 121 papers in this topic (those with \(gamma > 0.97\)) as contributing to race science discourse or not.
 105 (87\%) were coded as race science discourse by both authors; 10 (8.3\%) were coded as not race science discourse by both; and discordant codes were given to the remaining 6 (5.0\%) (95\% agreement, Cohen's \(\kappa = 0.74\)).
@@
 PID published multiple papers in topic 24 almost immediately after it was founded, with I showing a more gradual increase between the mid-1970s and mid-1990s.
+
+\hypertarget{race-science-and-behavior-genetics}{%
+\section*{Race science and behavior genetics}\label{race-science-and-behavior-genetics}}
 
@@
 
-The topic model analysis and the case of Whitney's address suggest that, by the 1990s, behavior genetics had at least somewhat distanced itself from race science, though perhaps without repudiating it.
-\emph{{[}\url{https://link.springer.com/search?use-oscar-shared-search=true\&sortBy=newestFirst\&dateTo=2010\&sortOrder=newestFirst\&facet-journal-id=10519\&content-type=Article\&query=race+intelligence\&date=custom\&dateFrom=\&date-facet-mode=between\&facet-start-year=1972\&previous-start-year=1972\&facet-end-year=2010\&previous-end-year=2023}{]}}
+The topic model analysis and the case of Whitney's address suggest that, by the 1990s, behavior genetics may have somewhat distanced itself from race science, though perhaps without directly repudiating it \cite{GillbornSoftlySoftlyGenetics2016, HennWhyDNANo2021}.
 
+\emph{{[}move most of this to supplement?{]}}
 To check this interpretation, we ran an independent search for documents with the keywords ``race'' and ``intelligence'' published in BG from 1972 to 2020, using Springer's journal search website.
@@
 14 of the remaining documents were published since 2010 (after the scope of the primary study).
-Among these, one obituary and two papers celebrated the work of John Loehlin, a prominent race-and-intelligence researcher who had been Director of the American Eugenics Society from 1968-1972 \emph{{[}\url{http://link.springer.com/article/10.1007/s10519-020-10024-w}; \url{http://link.springer.com/article/10.1007/s10519-014-9687-1}; \url{http://link.springer.com/article/10.1007/s10519-014-9673-7}{]}}.
+Among these, one obituary and two papers celebrated the work of John Loehlin, a prominent race-and-intelligence researcher who had been Director of the American Eugenics Society from 1968-1972 \cite{JohnLoehlin19262020, WaldmanIntroductionFestschriftJohn2014, PlominGenotypeEnvironmentCorrelationEra2014}.
 Another was a retrospective of the work of Lindon Eaves, a geneticist who made at least one notable contribution to debates on the Scarr-Rowe hypothesis (interactions between heritability, race, and class).
@@
 28 documents in this sample were published between 1990 and 2010.
-One was an obituary of Jerry Hirsch, a critic of hereditarianism in general and Jensen in particular \emph{{[}\url{http://link.springer.com/article/10.1007/s10519-008-9231-2}{]}}, and another was an obituary of David Rowe \emph{{[}\url{http://link.springer.com/article/10.1023/A\%3A1026122928868}{]}}.
-Only 2 of these 28 documents reported race differences of any kind: one examining at interactions among race, sex, and heritability for adolescent BMI (Body Mass Index, used as a measure of overweight/obesity) \emph{{[}\url{http://link.springer.com/article/10.1023/A\%3A1021619329904}{]}}; and the other at interactions among race, family history of alcoholism, and visuospatial performance \emph{{[}\url{http://link.springer.com/article/10.1007/BF02197241}{]}}.
+One was an obituary of Jerry Hirsch, a critic of hereditarianism in general and Jensen in particular \cite{RoubertouxJerryHirsch202008}, and another was an obituary of David Rowe \cite{RodgersObituaryDavidChristian2003}.
+Only 2 of these 28 documents reported race differences of any kind: one examining at interactions among race, sex, and heritability for adolescent BMI (Body Mass Index, used as a measure of overweight/obesity) \cite{JacobsonGeneticSharedEnvironmental1998}; and the other at interactions among race, family history of alcoholism, and visuospatial performance \cite{BermanReducedVisuospatialPerformance1995}.
 Finally, Philip Vernon --- one of the Pioneer-funded scientists we identified earlier --- published a critical review of a book that attempted to address hereditarianism, the Flynn effect, race-and-intelligence research, and some related issues.
@@
 None of the documents examined in this supplemental review explicitly repudiate race science.
-Behavior geneticists seem to typically adopt a posture of objectivity and technical rigor, downplaying the extent of the field's involvement in race science in the past.
+Similarly, another disciplinary history of behavior genetics, published by Plomin in 2023 \cite{PlominCelebratingCenturyResearch2023}, briefly mentions the controversies surrounding Jensen's 1969 paper and later \emph{The Bell Curve}, writing that Jensen's work was controversial because he ``touched on ethnic differences''; though ultimately Plomin reassures the reader that ``by the 1990s, quantitative genetic research had convinced most scientists of the importance of genetics for behavioral traits, including intelligence.'' In short, behavior geneticists seem to typically adopt a posture of objectivity and technical rigor, downplaying the extent of the field's involvement in race science in the past, and displacing the source of race science controversies to external, ``political,'' and ``irrational'' actors.
 
@@
 
-We stress that we are not identifying any particular scientist as a racist based on the topic model results.
-Scientific research can be utilized to promote scientific racism --- and thus count as race science --- independently of the views and attitudes of the scientist who conducted that research.\cite{TaberyWhyStudyingGenetics2015, CarlsonQuantifyingContextualizingImpact2020, HennWhyDNANo2021} But this does not mean that scientists are not responsible for the downstream social effects of their
+\emph{{[}for discussion: the possibility for white supremacist appropriation, ie, race science done by scientists who are not themselves scientific racists: CarlsonQuantifyingContextualizingImpact2020; probably some reviews of Paige Harden; O'Connor and Weatherall{]}}
+
+We stress that we are not identifying any particular scientist as a racist, scientific racist, or race scientist based on the topic model results.
+Scientific research can be utilized to promote scientific racism --- and thus count as race science --- independently of the views and attitudes of the scientist who conducted that research.\cite{TaberyWhyStudyingGenetics2015, GillbornSoftlySoftlyGenetics2016, CarlsonQuantifyingContextualizingImpact2020, HennWhyDNANo2021} But this does not mean that scientists are not responsible for the downstream social effects of their
 research, even when these effects involve deliberate misinterpretation.\emph{{[}Douglas, Kitcher, Kourany, Havstad; Block and Dworkin?{]}} Like other citizens, scientists are responsible for mitigating the reasonably foreseen harms that result from their actions.
-And we now have decades of empirical evidence that research on the genetics of intelligence --- even research that does not itself touch on race --- can and will be used to promote scientific racism.\cite{MeyerWrestlingSocialBehavioral2023}
+And we now have decades of empirical evidence that research on the genetics of intelligence --- even research that does not itself touch on race --- can and will be used to promote scientific racism \cite{MeyerWrestlingSocialBehavioral2023}.
+Specifically, while our findings suggest that behavior genetics has attempted to distance itself from scientific racism, behavior genetics research --- and research on human genetics more generally --- may be especially susceptible to appropriation by scientific racists \cite{CarlsonQuantifyingContextualizingImpact2020, OgbunugaforDNABasketballBirthday2022}.
+A colorblind attitude of ``we don't talk about race anymore'' \emph{{[}cite Bonilla-Silva?{]}} is likely insufficient to prevent particular research findings being utilized as race science.
 
-\emph{{[}engagement dilemma: technical critiques can legitimize{]}}
+Our distinction between race science and race science discourse highlights an important ``engagement dilemma'' for scientists challenging race science and scientific racism.
+While the ``tobacco strategy'' for manufacturing doubt involves targeting mainstream scientists and promoting fringe scientific views, its primary audiences are lawmakers and the general public.
+The aim of the strategy is not to change the opinion of mainstream scientists, but instead to cultivate the impression that there is legitimate scientific uncertainty about the issue, reasonable disagreement among qualified experts \emph{{[}cites{]}}.
+On the one hand, refusing to engage with fringe scientists --- or, often, pseudoscientists \cite{BhakthavatsalamVirtueEpistemologicalApproach2021} --- means that false or misleading statements go unchallenged.
+On the other hand, detailed, technical criticism legitimizes the fringe views, creating exactly the impression of reasonable scientific disagreement that the tobacco strategy is intended to produce.
 
-\emph{{[}limitation: colorblind racism (Bonilla-Silva; WinstonWhyMainstreamResearch2020) creates coding ambiguities{]}}
+In the case of race science, mainstream psychologists have often taken a non-engagement approach --- whether due to a strong conception of academic freedom or a desire to avoid stoking controversy --- with the result that race scientists have been able to present themselves as the representatives of ``mainstream science on intelligence.'' On the other hand, critics --- most prominently Gould, Kamin, and Lewontin --- have offered sophisticated technical critiques.
+But these tend to spiral off into esoteric debates that are inaccessible to the general public, as with ``Lewontin's fallacy'' \cite{RosemanLewontinDidNot2021}.
+Consequently these critiques are likely to be either ignored by the general public or taken as evidence of legitimate scientific debate.\\
+\emph{{[}more?{]}}
 
-\emph{{[}limitation: false negatives (race science papers that don't appear in V24) are possible if, eg, they use many terms related to twin studies of heritability and thereby appear in V03{]}}
+\hypertarget{limitations}{%
+\subsection*{Limitations}\label{limitations}}
 
-\emph{{[}replace w/: WinstonWhyMainstreamResearch2020: race science as a community project of psychology(?).
-radical academic freedom is a tool used to defend race science PigliucciWhatAreWe2013, RosemanTroublesomeReflectionRacism2014, LarsenMoreProvocativeLess2020.
-an important part of how these mainstream journals function as race science echo chambers.{]}}
+A key limitation of the current study stems from the inability of text mining techniques to draw on a broader cultural context than what is represented in the corpus.
+Scholars of racism and US race relations have noted a shift in the way racist attitudes are understood and expressed, from \emph{{[}something{]}} to ``colorblind racism'' \emph{{[}Bonilla-Silva{]}}, defined as \emph{{[}defn{]}}.
+Colorblind racism involves the use of racialized ``codes'' or ``dogwhistles,'' language that is superficially non-racial but carries racial implications, such as (in the US context) ``violent inner city criminals'' or ``welfare queen'' \cite{SaulDogwhistlesPoliticalManipulation2018}.
+Alderfer notes that the racist implications of \emph{The Bell Curve} only emerge gradually, with the opening of the book using the colorblind language of ``group differences'' \cite{AlderferScienceNonsciencePsychologists2003}.
+Panofsky et al.~show how the term ``human biodiversity,'' originally developed for anti-racist purposes by biological anthropologist Jonathan Marks, has been inverted by white supremacists and is now a dogwhistle for biological racial hierarchy \cite{PanofskyHowWhiteNationalists2021}.
+These cases suggest further that colorblindness can facilitate white supremacist appropriation of non-racial research --- even contrary to the researchers' own intentions --- as scientific racist readers re-interpret racial-neutral language as dogwhistles \cite{GillbornSoftlySoftlyGenetics2016, WillsAreClustersRaces2017}.
 
-Proponents of a strong conception of academic freedom often point to chapter 2 of philosopher John Stuart Mill's \emph{On Liberty} \cite{MillLiberty1862}, ``Of the Liberty of Thought and Discussion.'' However, in chapter 1 of the same book, Mill states the Harm Principle as the overall thesis of his book: ``The only end for which people are entitled, individually or collectively, to interfere with the liberty of action of any of their number is self-protection.
-The only purpose for which power can be rightfully exercised over any member of a civilized community, against his will, is to prevent harm to others.'' Even for Mill, free speech and free inquiry are appropriately restrained when the ideas in question will harm others.
+Techniques such as topic modeling may be able to account for discursive shifts from explicit to coded racial language, but only if the corpus contains a critical mass of documents using both sets of terms together.
+This would seem to require either a relatively gradual transition, the inclusion of reflective commentary, such as scholarship identifying and analyzing the transition, or both.
+It is unlikely that these requirements are met by the corpus used in this study.
+So it is plausible that the topic models used here have some rate of ``false negatives,'' documents in the corpus that engage in race science discourse but in a colorblind way that avoids the use of the explicit racial terms in the race-and-intelligence topic.
 
-More recently, philosopher Shannon Dea \cite{DeaEvolvingSocialPurpose2021} has argued that academic freedom is importantly distinct from freedom of speech.
-Whereas freedom of speech is universal, enjoyed by all residents of countries where speech is legally protected, academic freedom is more exclusive, enjoyed only by college and university faculty.
-(Though, as Dea points out, non-tenure-track faculty often do not have effective protections for academic freedom.) Freedom of speech primarily serves to protect the interests of the individual speaker: that you will not be punished by the state for stating unpopular opinions.
-But academic freedom is ``other-directed,'' primarily serving to benefit the common good of society as a whole by enabling the pursuit of truth.
-
-Advocates of research on genetics and intelligence --- both anti-racist liberals and race scientists --- claim this work will lead to social benefits.
-This reflects a recognition that the purpose of academic freedom is to promote the common good, rather than merely satisfy an individual's curiosity.
-But the promised benefits are either vague or do not actually require understanding the genetics of intelligence.\cite{KaplanGaltonQuincunxProbabilistic2021, CoopLotteryLuckLegacy2022}
-
-Using agent-based models, Weatherall et al.\cite{WeatherallHowBeatScience2018} show that a strategy of ``selective sharing'' --- promoting favorable results from independent research --- can be effective at misleading a target audience.
-
-\emph{{[}for discussion: the possibility for white supremacist appropriation, ie, race science done by scientists who are not themselves scientific racists: CarlsonQuantifyingContextualizingImpact2020; probably some reviews of Paige Harden; O'Connor and Weatherall{]}}
+The phenomenon of colorblind racism is closed related to ambiguities that cannot be resolved even by human coders familiar with the broader cultural context.
+Two relevant examples are the ``Flynn effect,'' a secular increase in intelligence test scores noted by philosopher James Flynn; and studies of ``national IQ,'' exemplified by the work of Richard Lynn.
+The Flynn effect implies that environmental factors can create group differences in IQ that are comparable to Black-White differences, and thus directly challenges claims that the racial differences are biological.
+Using methods widely regarded as ad hoc, cherry-picking, and generally unreliable, Lynn and collaborators have claimed that national IQ averages are correlated with national GDP.
+Both the Flynn effect and ``national IQ'' are strongly associated with race-and-intelligence discourse in the US cultural context.
+But scientific journal articles on these topics may not include any racial language at all.
+For example, a meta-analysis of the Flynn effect does not discuss racial differences or use terms such as ``White'' or ``Black'' outside the reference list \cite{TrahanFlynnEffectMetaanalysis2014}.
+Should such articles be coded as race science discourse?
+We expect that even readers who accept our definitions of race science and race science discourse will reasonably disagree on how to answer this question.
+Topic models, of course, are not even capable of representing this essential ambiguity.
 

### 2023-07-14 paper/supplement_paper.tex
mostly filling in limitations; need to ensure stay below ~3k words

@@
 
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-BarkanRetreatScientificRacism1992}{}}%
-Barkan, Elazar. 1992. \emph{The {Retreat} of {Scientific Racism}: {Changing Concepts} of {Race} in {Britain} and the {United States Between} the {World Wars}}. {Cambridge University Press}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-BeagleholeRaceQuestion1950}{}}%
-Beaglehole, Ernest, Juan Comas, L.
-A.
-Costa Pinto, Franklin Frazier, Morris Ginsberg, Humayun Kabir, Claude Levi-Strauss, and Ashley Montagu.
-1950.
-{``The {Race Question}.''} \emph{UNESCO and Its Programme} 1 (3).
-
-\leavevmode\vadjust pre{\hypertarget{ref-BrattainRaceRacismAntiracism2007}{}}%
-Brattain, M.
-2007.
-{``Race, {Racism}, and {Antiracism}: {UNESCO} and the {Politics} of {Presenting Science} to the {Postwar Public}.''} \emph{The American Historical Review} 112 (5): 1386--1413.
-\url{https://doi.org/10.1086/ahr.112.5.1386}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-MeadComingAgeSamoa1928}{}}%
-Mead, Margaret. 1928. \emph{Coming of {Age} in {Samoa}}. {New York}: {William Morrow and Company}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-MehlerFoundationFascismNew1989}{}}%
-Mehler, Barry. 1989. {``Foundation for Fascism: {The} New Eugenics Movement in the {United States}.''} \emph{Patterns of Prejudice} 23 (4): 17--25. \url{https://doi.org/10.1080/0031322X.1989.9970026}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-SlobodianUnequalMindHow2023}{}}%
-Slobodian, Quinn. 2023. {``The {Unequal Mind}: {How Charles Murray} and {Neoliberal Think Tanks Revived IQ}.''} \emph{Capitalism: A Journal of History and Economics} 4 (1): 73--108.
-
-\end{CSLReferences}
-
 \end{document}

### 2023-07-15 paper/paper.Rmd
updated paper with notes re: academic freedom and other values
added a few scattered notes in the commented out section on academic freedom. other values in scientific research, from ethical regulations to proposed norms of scientific culture to a recent paper proposing considerations for conducting socially responsible science. maybe some of this will be useful.

@@
 
+[EL notes on academic freedom as simply one value in science:
+
+Other ethical values that guide scientific research have been institutionalized via documents such as The Belmont Report.
+Principle values described in The Belmont Report are intended as guides and regulations of scientific researchers' behavior so as not to infringe on the rights or liberties of human study volunteers.
+This report was written in response to revelations in the 1970s of the Tuskegee Syphilis Study, wherein researchers at the CDC and HHS lied to participants - all of whom were impoversiehd black men - about the nature of the study and whether or not they had tested positive for syphilis, coerced their participation with extravagant compensation, and refused to allow men infected with syphilis to be treated at all.
+The Belmont Report outlines values guiding scientific research that focus on respect for persons, beneficence, and justice:
+Respect for persons - informed consent, participants are individuals with agency. 
+Beneficience - very much like Hippocratic "do no harm" or, if harm is unavoidable, the potential harms are outweighed by the benefits of research. 
+Justice - no population should disproportionately bear the burdens of being investigated scientifically; do not sample merely from convenient populations; when advances are made due to public funds (e.g.
+taxpayer money) the products of such advances are not restricted only to those who can afford them.
+
+How does race science research square against these other values in scientific research?
+Are participants ever honestly informed that they are (or were, in the case of a debriefing form) being evaluated to see which demographics are superior/inferior to which other ones?
+Granted, the Belmont Report is US-specific, but comparable other documents, regulations, or laws exist around the world that promote similar values.
+e.g.
+The Declaration of Helsinki states that the pursuit of knowledge cannot be used to trample on the rights of research participants.
+
+Maybe brief discussion of Mertonian Norms?
+Seems that race science research violates the norms of Universalism, Disinterestedness, and Organized Skepticism.
+Granted, most of the scientific community in practice violates all of the norms pretty frequently, so maybe this isn't a strong point to include.
+
+Recently published paper by Zivony et al in PLOS Computational Biology - "Ten simple rules for socially responsible science" - The authors call out a variety of biases in research (extreme male brain hypothesis of autism; face recognition software trained only on white faces; race differences in test scores) as failing to appropriately limit inferences based on the research's design or failing to properly ground their research in social theory or relevant historical context.
+
+Random oOther thought: academic freedom doesn't mean being free from criticism of one's methods, intentions, or theoretical starting point.
+Unless someone is calling for an academic policy that forbids race science as an academic area of inquiry, talk of academic freedom is non sequitor.
+Pointing out flaws in research and concluding the research is therefore wasted time/resources is not a threat to academic freedom.
+No one has forbidden research on geocentrism, but we've recognized for a few centuries now that it's a waste of time to pursue that line of inquiry.
+
+]
+
 -->

### 2023-07-19 paper/dataset.qmd
mostly filling in supplement refs

@@
-# Data and Methods
-
-## Corpus Assembly
-
-Corpus assembly is summarized in figure @fig-corpus.  The corpus was assembled in two parts, one for *Mankind Quarterly* and the other for mainstream journals.
-
 ```{mermaid}
@@
 ```
-
-### Mankind Quarterly
-
-An electronic archive of *Mankind Quarterly*, from the first issue through 2004, is available for free on the open web from *The Unz Review*, which has been characterized as a white nationalist website by the Southern Poverty Law Center *[<https://www.splcenter.org/hatewatch/2021/01/19/meet-white-nationalist-organizer-who-spewed-hate-against-lawmakers>]*.
-Web scraping packages and techniques *[which?]* were used to retrieve every PDF available in this archive.
-These PDFs were processed using the open-source OCR (optical character recognition) software `tesseract` and the R API `rtesseract` *[cites]*.
-This part of corpus assembly was conducted in Fall 2021.
-
-### Mainstream journals
-
-To identify suitable mainstream journals for inclusion in the corpus, we first identified academic researchers who had been funded by the Pioneer Fund.
-Pioneer is an American non-profit organization founded in 1937 to fund eugenics research.
-After the *Brown v.
-Board of Education* ruling in 1954, Pioneer funded various segregationist efforts across the United States.
-*[Shockley and Jensen]*.
-We reviewed a critical profile of Pioneer *[cite]* as well as an archived page from the organization's own web site *[cite]*, which together listed 16 researchers who had received Pioneer funds.
-
-We then used the author search tool in Clarivate's Web of Science platform, retrieving publication lists for 14 researchers.
-These searches were conducted between 2021-09-24 and 2021-10-05 by DJH.
-After parsing these results, we counted how many of the 14 researchers had published in each journal.
-13 journals had published 6 or more of the 14 Pioneer-funded researchers.
-We excluded *Mankind Quarterly* (already covered) as well as *Science* and *Nature* (as too general) from further consideration in this side of the corpus.
-3 journals published by the American Psychological Association (*American Psychologist*, *Contemporary Psychology*, and *Journal of Educational Psychology*) had to be excluded due to confusion over could give us permission to use the archives for a text mining project, with both APA and ProQuest asserting that we needed to get permission from the other.
-*European Journal of Personality*, published by SAGE, also had to be excluded because our institutional access only went back to 1999.
-
-The remaining 5 journals are all published by large academic publishers — Elsevier, Springer, or *[Cambridge UP]* — and each item in the entire run of each journal has been assigned a DOI (digital object identifier) for archival purposes.
-We used the Crossref API and `rcrossref` R interface to this API to retrieve metadata for each item published from 1960-2010 in each of these 6 journals.
-These metadata included individual license information — confirming that the text of each item could be used for text mining projects — and a URL to an electronic version of the item.
-These URLs were used to retrieve an HTML or PDF version of each item, except for *Personality and Individual Differences*.
-This journal has published a relatively large number of non-article documents, such as book reviews and commentaries, which are not available at the URL included in the Crossref metadata.
-(This is unfortunate, as it was not difficult to find highly relevant documents that we could not automatically retrieve and therefore excluded from our corpus *[review]*.)  Instead we used Elsevier's ScienceDirect API to search and retrieve all items from *Personality and Individual Differences* that are available.
-PDFs were run through OCR and text extraction as necessary, as with the *Mankind Quarterly* documents.
-
-All together, *[corpus summary, total and by journal; table]*
-
-## Data preparation
-
-After document retrieval and text extraction, we pre-processed the text using the `spaCy` NLP (natural language processing) Python library and the R API `spacyr` *[version, cites]*.
-Specifically, we applied regular expressions to remove header/footer copyright notices and hyphenation, used spaCy to annotate and extract noun phrases (eg, "the intelligence test items"), and then cleaned and standardized these phrases (eg, removing the/an/a, coverting all text to lowercase, and replacing all whitespace with underscores: "intelligence_test_items").
-We then counted the occurrence of each noun phrase in the document.
-The aggregated "document-term matrix" was stored in Parquet format *[cite]* for performance reasons, and written and read using the `arrow` package for R *[cite]*.
-*[availability]*
-
-NLP-extracted noun phrases offer a number of advantages over the more traditional unigram ("single-word") tokens.
-First, noun phrase extraction removes many standard stopwords (articles, common verbs) without relying on an a priori stopword list.
-Phrases can be more informative than single terms, for example, distinguishing "intelligence test" from "hypothesis test."  While simple n-gram extraction will include phrases that are not especially meaningful.
-Consider the sentence "Since our first analyses of feeding patterns in rats, we had been using a criterion of 40 minutes (Le Magnen & Tallon 1963; 1966)" *[10.1017__^__s0140525x0000042x]*.
-Bigrams such as "since our" and "criterion 40" will likely be discarded in vocabulary selection, but add to the computational cost of vocabulary selection.
-Noun phrase extraction is therefore more efficient.
-
-After noun phrase extraction, the corpus comprised *[tokens]* of *[distinct noun phrases]*.
-
-### Vocabulary selection
-
-We took an information-theoretic approach to vocabulary selection.
-Consider a game in which I draw a document from the corpus, then a single token from that document.
-I tell you the term, and you have to guess which document I picked.
-Intuitively, highly informative terms (in this project, noun phrase types) are distinctive, allowing you to dramatically narrow down the list of potential documents.
-This "informativeness" of a term can be quantified as the KL (Kullback-Leibler) divergence from a "baseline" distribution of documents to the distribution conditional on the term.
-Because the most informative terms tend to be typos and OCR errors — these are unique to a single document — we multiply the KL divergence by the logarithm of the total number of occurrences of the term across the entire corpus.
-We refer to the resulting measure as $log(n) \Delta H$ or simply `ndH`.
-
-*[Hicks 2021?]* used the `ndH` approach in a topic modeling analysis, where the baseline distribution was the uniform distribution across documents.
-In the current project, we found that this approach heavily favored recurrent noun phrases in the longest documents.
-Many of the documents published in *Psychological Reports* are extremely short, 1-2 reports of a single study; while many of the documents published in *Brain and Behavioral Sciences* are book-length collections that include a long review article and sometimes dozens of commentaries.
-More generic noun phrases in the latter can occur orders of magnitude more often than highly distinctive phrases in the former, and so the $log(n)$ factor overwhelms the $\Delta H$ factor.
-
-To address this, we used a different baseline distribution of documents.
-Namely, we used a distribution in which document probability is proportional to length.
-This makes phrases from short documents much more surprising, and hence dramatically increases their informativeness.
-This was more effective at identifying useful phrases from across the corpus.
-
-A common rule of thumb in topic modeling is that the vocabulary should have about 10 times as many distinct terms as the number of documents in the corpus *[cite]*.
-However, we had some concerns with computational demands here:  the resulting document-term matrix would have roughly $33,000 \times 330,000$ or 10.9 billion entries; with 10% density this would require on the order of 4 GB of memory; while most of the analysis was to be conducted on the authors' laptops.
-We therefore chose to work with three smaller vocabularies, $5 \times$, $1 \times$ and $\frac{1/5} \times$ the number of documents.
-We refer to these as the "large," "medium," and "small" vocabulary, respectively, and they include *[counts]* distinct phrases.
-We compare findings across vocabularies as a robustness check.
-
-## Topic modeling
-
-To fit topic models, we followed the approach proposed by *[cite]*, which uses varimax-rotated partial principal components instead of the variational inference methods used by standard topic model packages *[eg, stm]*.
-This novel approach was implemented in the R package `tmfast`, available separately at *[url]*.
-A simulation study of `tmfast` found that it was significantly faster and slightly less accurate at reconstructing known word-topic and topic-document distributions, compared to `stm`.
-
-Topic models were fit for all three vocabularies (large, medium, and small) with $k = 5, 10, 20, 30, 40, 50$ (number of topics), resulting in a total of 18 models.
-
-Following the approach of *[Hicks 2021?]*, we did not attempt to identify a unique best model for further analysis.
-While we focus on the medium vocabulary, $k=30$ model as the "median" among the 18 models fit, we compare and contrast findings from this model with those from the other models.
-
-### Topic model interpretation
-
-After fitting topic models, our first research question was whether we could identify distinctive "race science" topics.
-We addressed this question in two ways.
-Following common practice in topic modeling, we extracted lists of the top $n$ (highest-probability) terms (noun phrases) from each topic in each model, with $n$ generally ranging between 5 and 15 as we explored the models.
-*[tables + Silge plots]*
-
-We also adapted the "discursive space" analysis proposed by *[Hicks 2021?]*.
-This analysis combines the Hellinger distance between topic-document distributions — capturing how "far apart" two documents are — with dimensional reduction techniques used for data visualization.
-This produces a two-dimensional visualization of individual documents, arranged so that documents with similar topic distributions are visually clustered and visually distinct from documents with different topic distributions.
-
-Specifically, for each fitted topic model, we extracted topic-document distributions for each document and calculated pairwise Hellinger distances.
-We then processed these distances using the UMAP algorithm for dimension reduction, as implemented in the `umap` package *[details; cite]*.
-To reduce computational requirements for these steps, we first dropped documents published in the journal *[Psychological Reports]* (*[x documents removed, y remaining]*; note that the full corpus was used to fit the topic models, *Psychological Reports* was removed just for this visualization).
-Using the `plotly` library, we construct an interactive visualization:  each point represents a single document, and points can be colored based on either their journal or highest-probability topic.
-Tooltips are displayed when we hover over a point, including document metadata (title, journal, authors, year) as well as "key phrases."
-
-To calculate key phrases for a given document $d$, we first identify the highest-probability topic $t$ for that document and the word-topic (noun phrase-topic) distribution $\beta_{w,t}$ for that topic.
-We also retrieve the document-term matrix, containing the count $n_{w,d}$ of occurrences for each word (noun phrase), and calculate a weighting $n_{w,d} \beta_{w,t}$ for each word.
-The key phrases are the 5 most highly-weighted terms for the given document.

### 2023-07-19 paper/paper.Rmd
mostly filling in supplement refs

--- 2023-07-14 paper/paper.Rmd
@@
   # skeleton: <https://github.com/christopherkenny/rticles/tree/master/inst/rmarkdown/templates/science>
-  # rticles::science_article:
-  #   keep_tex: true
-  #    move_figures: false
-  #    draft: false
-  #    number_sections: false
   bookdown::pdf_book:
@@
     move_figures: false
-    draft: false
+    draft: true
     number_sections: false
@@
 
-![Journals publishing 6 or more Pioneer-funded researchers, Web of Science author search results *[date coverage]* (\#fig:wos)](img/wos_results.png){width="4.76in" height="2.6in"}
+![Journals publishing 6 or more Pioneer-funded researchers, Web of Science author search results. (\#fig:wos)](img/wos_results.png){width="4.76in" height="2.6in"}
 
@@
 
-Our findings indicate that *Mankind Quarterly* (MQ) was less important as an echo chamber for late 20th century race science than mainstream psychology journals, especially *Intelligence* (I) and *Personality and Individual Differences* (PID).
-Indeed, MQ published a very different kind of race science from the race-and-intelligence research published in I and PID, even after *[Lynn joined the MQ editorial board]*.
-
-*[for discussion: the possibility for white supremacist appropriation, ie, race science done by scientists who are not themselves scientific racists: CarlsonQuantifyingContextualizingImpact2020; probably some reviews of Paige Harden; O'Connor and Weatherall]*
+Using bibliometric and text-mining methods, this study finds that Pioneer-funded researchers published heavily in certain mainstream psychology journals, rather than *Mankind Quarterly*; and that a distinct topic of race science discourse can be identified in these same journals across the period 1960-2010.
+These findings indicate that *Mankind Quarterly* (MQ) was less important as an echo chamber for late 20th century race science than mainstream psychology journals, especially *Intelligence* (I) and *Personality and Individual Differences* (PID).
+Indeed, MQ published a very different kind of race science from the race-and-intelligence research published in I and PID.
 
@@
 And we now have decades of empirical evidence that research on the genetics of intelligence — even research that does not itself touch on race — can and will be used to promote scientific racism \cite{MeyerWrestlingSocialBehavioral2023}.
-Specifically, while our findings suggest that behavior genetics has attempted to distance itself from scientific racism, behavior genetics research — and research on human genetics more generally — may be especially susceptible to appropriation by scientific racists \cite{CarlsonQuantifyingContextualizingImpact2020, OgbunugaforDNABasketballBirthday2022}.
-A colorblind attitude of "we don't talk about race anymore" *[cite Bonilla-Silva?]* is likely insufficient to prevent particular research findings being utilized as race science.
+Specifically, while our findings suggest that behavior genetics has attempted to distance itself from scientific racism, behavior genetics research — and research on human genetics more generally — may be especially susceptible to appropriation by scientific racists \cite{CarlsonQuantifyingContextualizingImpact2020, OgbunugaforDNABasketballBirthday2022, *[Harden reviews]*}.
 
@@
 \appendix
+
 # Materials and methods
-
-*[probably some introduction]*
 
@@
 
-Corpus assembly is summarized in figure *[xref]*.  The corpus was assembled in two parts, one for *Mankind Quarterly* and the other for mainstream journals.
+![Corpus assembly. (\#fig:corpus)](img/dataset.png){width="8in" height="6.1in"}
 
-*[figure]*
+Corpus assembly is summarized in figure \@ref(fig:corpus).  The corpus was assembled in two parts, one for *Mankind Quarterly* and the other for mainstream journals.
 
@@
 
-An electronic archive of *Mankind Quarterly*, from the first issue in *[1969]* through 2004, is available for free on the open web from *The Unz Review*, which has been characterized as a white nationalist website by the Southern Poverty Law Center *[<https://www.splcenter.org/hatewatch/2021/01/19/meet-white-nationalist-organizer-who-spewed-hate-against-lawmakers>]*.
-Web scraping packages and techniques *[which?]* were used to retrieve every PDF available in this archive.
-These PDFs were processed using the open-source OCR (optical character recognition) software `tesseract` and the R API `rtesseract` *[cites]*.
-This stage of corpus assembly was conducted in Fall 2021.
+An electronic archive of *Mankind Quarterly*, from the first issue in 1969 through 2004, is available for free on the open web at the white nationalist website *The Unz Review*.
+The Python package Beautiful Soup (<https://www.crummy.com/software/BeautifulSoup/>, version not recorded) was used to retrieve every PDF available in this archive.
+After identifying some gaps (missing issues) in the archive from *The Unz Review*, we manually retrieved PDFs for further articles using ProQuest.
+Text was extracted using the R package `pdftools` \cite[3.0.1]{OomsPdftoolsTextExtraction2023}.
+This stage of corpus assembly was conducted in Fall 2021 with the assistance of Anthony Sainez.
 
@@
 To identify suitable mainstream journals for inclusion in the corpus, we first identified academic researchers who had been funded by the Pioneer Fund.
-Pioneer is an American non-profit organization founded in 1937 to fund eugenics research *[cite]*.
+Pioneer is an American non-profit organization founded in 1937 to fund eugenics research and propaganda \cite{TuckerFundingScientificRacism2002}.
 After the *Brown v.
-Board of Education* ruling in 1954, Pioneer funded various segregationist efforts across the United States.
-*[Shockley and Jensen]*.
-We reviewed a critical profile of Pioneer *[cite]* as well as an archived page from the organization's own web site *[cite]*, which together listed 16 researchers who had received Pioneer funds.
+Board of Education* ruling in 1954, Pioneer funded various segregationist efforts across the United States, including lectures on eugenics by Stanford physicist William Shockley \cite{JacksonScienceSegregationRace2005}.
+We reviewed a critical profile of Pioneer \cite{MillerPioneerFundBankrolling1994} as well as an archived page from the organization's own web site (<https://web.archive.org/web/20130103005545/http://www.pioneerfund.org/Grantees.html>), which together listed 16 researchers who had received Pioneer funds.
 
-We then used the author search tool in Clarivate's Web of Science platform *[url]*, retrieving publication lists for 14 researchers.
+We then used the author search tool in Clarivate's Web of Science platform (<https://www.webofscience.com/wos/author/search>), retrieving publication lists for 14 researchers.
 These searches were conducted between 2021-09-24 and 2021-10-05 by DJH.
@@
 The remaining 5 journals are all published by major academic publishers — Elsevier, Springer, or Cambridge University Press — and each item in the entire run of each journal has been assigned a DOI (digital object identifier) for archival purposes.
-We used the Crossref API and `rcrossref` R interface to this API *[cite]* to retrieve metadata for each item published from 1960-2010 in each of these 6 journals.
+We used the Crossref API and `rcrossref` R interface to this API \cite[1.1.0.99]{ChamberlainRcrossrefClientVarious2019} to retrieve metadata for each item published from 1960-2010 in each of these journals.
 These metadata included item-level license information — confirming that the text of each item could be used for text mining projects — and a URL to an electronic version of the item.
-These URLs were used to retrieve an HTML or PDF version of each item, except for *Personality and Individual Differences*.
-This journal has published a relatively large number of non-article documents, such as book reviews and commentaries, which are not available at the URL included in the Crossref metadata.
-(This is unfortunate, as it was not difficult to find highly relevant documents that we could not automatically retrieve and therefore excluded from our corpus *[review]*.)  Instead we used Elsevier's ScienceDirect API *[url]* to search and retrieve all items from *Personality and Individual Differences* that are available.
-PDFs were run through OCR and text extraction as necessary, as with the *Mankind Quarterly* documents.
+These URLs were used to retrieve a XML or PDF version of each item, except for *Personality and Individual Differences*.
+This journal has published a relatively large number of non-article documents, such as book reviews and commentaries, that are not available at the URL included in the Crossref metadata.
+(This is unfortunate, as it was not difficult to find highly relevant documents that we could not automatically retrieve and therefore excluded from our corpus, such as an exchange between Rushton and Flynn on the Flynn effect \cite{FlynnEvidenceRushtonGenetic1998,RushtonSecularGainsIQ1998,FlynnReplyRushtonGang1998}.)  Instead we used Elsevier's ScienceDirect API (<https://dev.elsevier.com/>) to independently search and retrieve all available items from *Personality and Individual Differences*.
+This stage of corpus assembly was conducted between 2021-11 and 2022-05.
 
-All together, *[corpus summary, total and by journal; table]*
+*Behavioral and Brain Sciences* typically uses a target article + commentary format; in some cases there are dozens of commentaries for a single article.
+In the Crossref DOI metadata, each target article and individual commentary is given its own DOI, with no distinction between contribution types or metadata links between a commentary and the corresponding target article.
+But text is only available in aggregate PDFs that bundle together the target article and all of the commentaries.
+In a few cases this results in PDFs that are hundreds of pages long and might be linked from the Crossref metadata 30+ times.
+In addition, the retrieved PDFs are not perfectly identical, because Cambridge UP's servers add a timestamped watermark to each page.
+We attempted to contact Cambridge for assistance but did not receive a response.
+We ultimately used a series of ad hoc measures to mitigate text duplication.
+Approximately 100,000 documents of 1-2 pages long were excluded before PDF retrieval.
+After PDF retrieval and text extraction, the timestamped watermarks were removed using a regular expression and the text was hashed using SHA 256.
+Hashes were used to construct groups of duplicate documents, and a single document (whichever one happened to be first in the dataframe) was chosen from each hash group for inclusion in the corpus.
+It is plausible that some duplicates made it through this process: where the watermark overlapped with the text, the regular expression may have been unable to identify and remove the watermark (a false negative result), and this difference in the watermarks (not the text) would produce different hashes.
+
+After XML/PDF versions were retrieved, text was extracted using either the `xml2` package \cite{WickhamXml2ParseXML2023} or `pdftools`.
+Figure \@ref(fig:counts) and table \@ref(tab:counts) show the number of fulltext documents included in the corpus.
+
+![Count of documents in the corpus, by journal and year. (\#fig:counts)](img/02_count.png){width="4.76in" height="2.6in"}
+
+Table: (\#tab:counts) Document counts, by journal, and years included in the corpus.
+
+|journal                                |     n| start|  end|
+|:--------------------------------------|-----:|-----:|----:|
+|Behavior Genetics                      |  2268|  1970| 2010|
+|Intelligence                           |  1237|  1977| 2010|
+|Mankind Quarterly                      |  1821|  1960| 2004|
+|Personality and Individual Differences |  7274|  1980| 2010|
+|Psychological Reports                  | 21398|  1960| 2010|
+|Behavioral and Brain Sciences          |   898|  1978| 2010|
+
+All together, the corpus includes 34,896 documents from 6 journals.  See \@ref(fig:counts) and \@ref(tab:counts).
 
@@
 
-After document retrieval and text extraction, we pre-processed the text using the `spaCy` NLP (natural language processing) Python library and the R API `spacyr` *[version, cites]*.
+After document retrieval and text extraction, we pre-processed the text using the `spaCy` NLP (natural language processing) Python library \cite[1.9.0]{SpaCyIndustrialstrengthNatural2018} and the R API `spacyr` \cite[1.2.1]{BenoitSpacyrWrapperSpaCy2020}.
 Specifically, we applied regular expressions to remove header/footer copyright notices and hyphenation, used spaCy to annotate and extract noun phrases (eg, "the intelligence test items"), and then cleaned and standardized these phrases (eg, removing the/an/a, coverting all text to lowercase, and replacing all whitespace with underscores: "intelligence_test_items").
 We then counted the occurrence of each noun phrase in the document.
-The aggregated "document-term matrix" was stored in Parquet format *[cite]* for performance reasons, and written and read using the `arrow` package for R *[cite]*.
-*[availability]*
+The aggregated "document-term matrix" was stored in Parquet format \cite{VohraApacheParquet2016} for performance reasons, and written and read using the `arrow` package for R \cite[7.0.0]{RichardsonArrowIntegrationApache2023}.
 
@@
 First, noun phrase extraction removes many standard stopwords (articles, common verbs) without relying on a fixed, a priori stopword list.
-Phrases can be more informative than single terms, for example, distinguishing "intelligence test" from "hypothesis test."  But simple n-gram extraction will include numerous phrases that are not especially meaningful.
-Consider the sentence "Since our first analyses of feeding patterns in rats, we had been using a criterion of 40 minutes (Le Magnen & Tallon 1963; 1966)" *[10.1017__^__s0140525x0000042x]*.
-Bigrams such as "since our" and "criterion 40" will likely be discarded in vocabulary selection, but significantly increase the computational cost of vocabulary selection.
+Phrases can be more informative than single terms, for example, distinguishing "intelligence test" from "hypothesis test."  While simple n-gram extraction will include numerous phrases that are not especially meaningful.
+Consider the sentence "Since our first analyses of feeding patterns in rats, we had been using a criterion of 40 minutes (Le Magnen & Tallon 1963; 1966)."\cite{SclafaniCorrelationCausationStudy1981}  Bigrams such as "since our" and "criterion 40" will likely be discarded in vocabulary selection, but significantly increase the computational cost of vocabulary selection.
 Noun phrase extraction is therefore more efficient.
 
-After noun phrase extraction, the corpus comprised *[tokens]* of *[distinct noun phrases]*.
+After noun phrase extraction, the corpus comprised 43,162,055 total tokens of 6,320,783 distinct phrases.
 
@@
 
-We took an information-theoretic approach to vocabulary selection.
-Consider a game in which I draw a document from the corpus, then a single token from that document.
+We took an information-theoretic approach to vocabulary selection.\cite{HicksProductivityInterdisciplinaryImpacts2021}  Consider a game in which I draw a document from the corpus, then a single token from that document.
 I tell you the term, and you have to guess which document I picked.
@@
 
-*[Hicks 2021?]* used the `ndH` approach in a topic modeling analysis, where the baseline distribution was the uniform distribution across documents.
-In the current project, we found that this approach heavily favored recurrent noun phrases in the longest documents.
-Many of the documents published in *Psychological Reports* are extremely short, 1-2 reports of a single study; while many of the documents published in *Brain and Behavioral Sciences* are book-length collections that include a long review article and sometimes dozens of commentaries.
+In the current project, we found that this `ndH` approach heavily favored recurrent noun phrases in the longest documents.
+Many of the documents published in *Psychological Reports* are extremely short, 1-2 page brief notices of a single study; while many of the documents published in *Brain and Behavioral Sciences* are book-length collections that include a long review article and sometimes dozens of commentaries.
 Very generic noun phrases that happen to appear in the latter can occur orders of magnitude more often than highly distinctive phrases in the former, and so the $log(n)$ factor overwhelms the $\Delta H$ factor.
@@
 
-A common rule of thumb in topic modeling is that the vocabulary should have about 10 times as many distinct terms as the number of documents in the corpus *[cite]*.
+A common rule of thumb in topic modeling is that the vocabulary should have about 10 times as many distinct terms as the number of documents in the corpus.
 However, we had some concerns with computational demands here:  the resulting document-term matrix would have roughly $10 \times 33,000^2$ or 10.9 billion entries; with 10% density this would require on the order of 4 GB of memory just for a single copy of the matrix; and most of the analysis was to be conducted on the authors' laptops.

### 2023-07-19 paper/paper.tex
mostly filling in supplement refs

@@
 
-In the context of \emph{Brown}, MQ was created as what contemporary historians and philosophers of science call an ``echo chamber'' \cite{FernandezPintoKnowBetterNot2017}, a favorable venue for race scientists to publish their views, on the grounds that an ``equalitarian dogma'' created a censorious ``taboo'' against their research in mainstream publications \cite{TuckerFundingScientificRacism2002, JacksonMythicalTabooRace2020}.
+As part of the segregationist reaction to \emph{Brown}, MQ was created as what contemporary historians and philosophers of science call an ``echo chamber'' \cite{FernandezPintoKnowBetterNot2017}, a favorable venue for race scientists to publish their views, on the grounds that an ``equalitarian dogma'' created a censorious ``taboo'' against their research in mainstream publications \cite{TuckerFundingScientificRacism2002, JacksonMythicalTabooRace2020}.
 Scholars have shown that echo chambers are a key part of the ``tobacco strategy,'' used by numerous regulated industries --- most infamously tobacco, but also fossil fuels and chemical manufacturing, among others --- to ``manufacture doubt'' and delay regulation \emph{{[}cites{]}}.
 
-In light of the origins of MQ, the significant attention paid to the journal by historians of scientific racism, and contemporary research on the ``tobacco strategy,'' we hypothesized that bibliometric and text mining analyses would confirm that MQ functioned as an echo chamber.
+In light of the origins of MQ, the significant attention paid to the journal by historians of scientific racism, and contemporary research on the ``tobacco strategy,'' we hypothesized that bibliometric and text mining analyses would confirm that MQ functioned as an echo chamber for Pioneer-funded scientists.
 This hypothesis was not supported.
@@
 \includegraphics[width=4.76in,height=2.6in]{img/wos_results.png}
-\caption{Journals publishing 6 or more Pioneer-funded researchers, Web of Science author search results \emph{{[}date coverage{]}} \label{fig:wos}}
+\caption{Journals publishing 6 or more Pioneer-funded researchers, Web of Science author search results. \label{fig:wos}}
 \end{figure}
@@
 
-Our findings indicate that \emph{Mankind Quarterly} (MQ) was less important as an echo chamber for late 20th century race science than mainstream psychology journals, especially \emph{Intelligence} (I) and \emph{Personality and Individual Differences} (PID).
-Indeed, MQ published a very different kind of race science from the race-and-intelligence research published in I and PID, even after \emph{{[}Lynn joined the MQ editorial board{]}}.
-
-\emph{{[}for discussion: the possibility for white supremacist appropriation, ie, race science done by scientists who are not themselves scientific racists: CarlsonQuantifyingContextualizingImpact2020; probably some reviews of Paige Harden; O'Connor and Weatherall{]}}
+Using bibliometric and text-mining methods, this study finds that Pioneer-funded researchers published heavily in certain mainstream psychology journals, rather than \emph{Mankind Quarterly}; and that a distinct topic of race science discourse can be identified in these same journals across the period 1960-2010.
+These findings indicate that \emph{Mankind Quarterly} (MQ) was less important as an echo chamber for late 20th century race science than mainstream psychology journals, especially \emph{Intelligence} (I) and \emph{Personality and Individual Differences} (PID).
+Indeed, MQ published a very different kind of race science from the race-and-intelligence research published in I and PID.
 
@@
 And we now have decades of empirical evidence that research on the genetics of intelligence --- even research that does not itself touch on race --- can and will be used to promote scientific racism \cite{MeyerWrestlingSocialBehavioral2023}.
-Specifically, while our findings suggest that behavior genetics has attempted to distance itself from scientific racism, behavior genetics research --- and research on human genetics more generally --- may be especially susceptible to appropriation by scientific racists \cite{CarlsonQuantifyingContextualizingImpact2020, OgbunugaforDNABasketballBirthday2022}.
-A colorblind attitude of ``we don't talk about race anymore'' \emph{{[}cite Bonilla-Silva?{]}} is likely insufficient to prevent particular research findings being utilized as race science.
+Specifically, while our findings suggest that behavior genetics has attempted to distance itself from scientific racism, behavior genetics research --- and research on human genetics more generally --- may be especially susceptible to appropriation by scientific racists \cite{CarlsonQuantifyingContextualizingImpact2020, OgbunugaforDNABasketballBirthday2022, *[Harden reviews]*}.
 
@@
 \emph{{[}max \textasciitilde50 refs{]}}
+
+\appendix
+
+\hypertarget{materials-and-methods}{%
+\section*{Materials and methods}\label{materials-and-methods}}
+
+\hypertarget{corpus-assembly}{%
+\subsection*{Corpus assembly}\label{corpus-assembly}}
+
+\begin{figure}
+\centering
+\includegraphics[width=8in,height=6.1in]{img/dataset.png}
+\caption{Corpus assembly. \label{fig:corpus}}
+\end{figure}
+
+Corpus assembly is summarized in figure \ref{fig:corpus}. The corpus was assembled in two parts, one for \emph{Mankind Quarterly} and the other for mainstream journals.
+
+\hypertarget{mankind-quarterly}{%
+\subsubsection*{Mankind Quarterly}\label{mankind-quarterly}}
+
+An electronic archive of \emph{Mankind Quarterly}, from the first issue in 1969 through 2004, is available for free on the open web at the white nationalist website \emph{The Unz Review}.
+The Python package Beautiful Soup (\url{https://www.crummy.com/software/BeautifulSoup/}, version not recorded) was used to retrieve every PDF available in this archive.
+Text was extracted using the R package \texttt{pdftools} \cite[3.0.1]{OomsPdftoolsTextExtraction2023}.
+This stage of corpus assembly was conducted in Fall 2021 with the assistance of Anthony Sainez.
+
+\hypertarget{mainstream-journals}{%
+\subsubsection*{Mainstream journals}\label{mainstream-journals}}
+
+To identify suitable mainstream journals for inclusion in the corpus, we first identified academic researchers who had been funded by the Pioneer Fund.
+Pioneer is an American non-profit organization founded in 1937 to fund eugenics research and propaganda \cite{TuckerFundingScientificRacism2002}.
+After the \emph{Brown v.
+Board of Education} ruling in 1954, Pioneer funded various segregationist efforts across the United States, including lectures on eugenics by Stanford physicist William Shockley \cite{JacksonScienceSegregationRace2005}.
+We reviewed a critical profile of Pioneer \cite{MillerPioneerFundBankrolling1994} as well as an archived page from the organization's own web site \url{https://web.archive.org/web/20130103005545/http://www.pioneerfund.org/Grantees.html}, which together listed 16 researchers who had received Pioneer funds.
+
+We then used the author search tool in Clarivate's Web of Science platform \emph{{[}url{]}}, retrieving publication lists for 14 researchers.
+These searches were conducted between 2021-09-24 and 2021-10-05 by DJH.
+After parsing these results, we counted how many of the 14 researchers had published in each journal.
+13 journals had published 6 or more of the 14 Pioneer-funded researchers.
+We excluded \emph{Mankind Quarterly} (as already included) as well as \emph{Science} and \emph{Nature} (as too general) from further consideration in this side of the corpus.
+3 journals published by the American Psychological Association (APA; \emph{American Psychologist}, \emph{Contemporary Psychology}, and \emph{Journal of Educational Psychology}) had to be excluded due to confusion over who could give us permission to use the archives for a text mining project, with both APA and ProQuest asserting that we needed to get permission from the other entity.
+\emph{European Journal of Personality}, published by SAGE, also had to be excluded because our institutional access only went back to 1999.
+
+The remaining 5 journals are all published by major academic publishers --- Elsevier, Springer, or Cambridge University Press --- and each item in the entire run of each journal has been assigned a DOI (digital object identifier) for archival purposes.
+We used the Crossref API and \texttt{rcrossref} R interface to this API \emph{{[}cite{]}} to retrieve metadata for each item published from 1960-2010 in each of these 6 journals.
+These metadata included item-level license information --- confirming that the text of each item could be used for text mining projects --- and a URL to an electronic version of the item.
+These URLs were used to retrieve an HTML or PDF version of each item, except for \emph{Personality and Individual Differences}.
+This journal has published a relatively large number of non-article documents, such as book reviews and commentaries, which are not available at the URL included in the Crossref metadata.
+(This is unfortunate, as it was not difficult to find highly relevant documents that we could not automatically retrieve and therefore excluded from our corpus \emph{{[}review{]}}.) Instead we used Elsevier's ScienceDirect API \emph{{[}url{]}} to search and retrieve all items from \emph{Personality and Individual Differences} that are available.
+PDFs were run through OCR and text extraction as necessary, as with the \emph{Mankind Quarterly} documents.
+
+\emph{Behavioral and Brain Sciences} typically uses a target article + commentary format; in some cases there are dozens of commentaries for a single article.
+In the Crossref DOI metadata, each target article and individual commentary is given its own DOI, with no distinction between contribution types or metadata links between a commentary and the corresponding target article.
+But text is only available in aggregate PDFs that bundle together the target article and all of the commentaries.
+In a few cases this results in PDFs that are hundreds of pages long and might be linked from the Crossref metadata 30+ times.
+In addition, the retrieved PDFs are not perfectly identical, because Cambridge UP's servers add a timestamped watermark to each page.
+We attempted to contact Cambridge for assistance but did not receive a response.
+We ultimately used a series of ad hoc measures to mitigate text duplication.
+Approximately 100,000 documents of 1-2 pages long were excluded before PDF retrieval.
+After PDF retrieval and text extraction, the timestamped watermarks were removed using a regular expression and the text was hashed using SHA 256.
+Hashes were used to construct groups of duplicate documents, and a single document (whichever one happened to be first in the dataframe) was chosen from each hash group for inclusion in the corpus.
+It is plausible that some duplicates made it through this process: where the watermark overlapped with the text, the regular expression may have been unable to identify and remove the watermark (a false negative result), and this difference in the watermarks (not the text) would produce different hashes.
+
+\begin{figure}
+\centering
+\includegraphics[width=4.76in,height=2.6in]{img/02_count.png}
+\caption{Count of documents in the corpus, by journal and year. \label{fig:counts}}
+\end{figure}
+
+\begin{longtable}[]{@{}lrrr@{}}
+\caption{\label{tab:counts} Document counts, by journal, and years included in the corpus.}\tabularnewline
+\toprule
+journal & n & start & end \\
+\midrule
+\endfirsthead
+\toprule
+journal & n & start & end \\
+\midrule
+\endhead
+Behavior Genetics & 2268 & 1970 & 2010 \\
+Intelligence & 1237 & 1977 & 2010 \\
+Mankind Quarterly & 1821 & 1960 & 2004 \\
+Personality and Individual Differences & 7274 & 1980 & 2010 \\
+Psychological Reports & 21398 & 1960 & 2010 \\
+Behavioral and Brain Sciences & 898 & 1978 & 2010 \\
+\bottomrule
+\end{longtable}
+
+All together, the corpus includes 34,896 documents from 6 journals. See \ref{fig:counts} and \ref{tab:counts}.
+
+\hypertarget{text-preparation}{%
+\subsection*{Text preparation}\label{text-preparation}}
+
+After document retrieval and text extraction, we pre-processed the text using the \texttt{spaCy} NLP (natural language processing) Python library and the R API \texttt{spacyr} \emph{{[}version, cites{]}}.
+Specifically, we applied regular expressions to remove header/footer copyright notices and hyphenation, used spaCy to annotate and extract noun phrases (eg, ``the intelligence test items''), and then cleaned and standardized these phrases (eg, removing the/an/a, coverting all text to lowercase, and replacing all whitespace with underscores: ``intelligence\_test\_items'').
+We then counted the occurrence of each noun phrase in the document.
+The aggregated ``document-term matrix'' was stored in Parquet format \emph{{[}cite{]}} for performance reasons, and written and read using the \texttt{arrow} package for R \emph{{[}cite{]}}.
+\emph{{[}availability{]}}
+
+NLP-extracted noun phrases offer a number of advantages over the more traditional unigram (``single-word'') tokens.
+First, noun phrase extraction removes many standard stopwords (articles, common verbs) without relying on a fixed, a priori stopword list.
+Phrases can be more informative than single terms, for example, distinguishing ``intelligence test'' from ``hypothesis test.'' But simple n-gram extraction will include numerous phrases that are not especially meaningful.
+Consider the sentence ``Since our first analyses of feeding patterns in rats, we had been using a criterion of 40 minutes (Le Magnen \& Tallon 1963; 1966)'' *{[}10.1017\_\_\^{}\_\_s0140525x0000042x{]}*.
+Bigrams such as ``since our'' and ``criterion 40'' will likely be discarded in vocabulary selection, but significantly increase the computational cost of vocabulary selection.
+Noun phrase extraction is therefore more efficient.
+
+After noun phrase extraction, the corpus comprised \emph{{[}tokens{]}} of \emph{{[}distinct noun phrases{]}}.
+
+\hypertarget{data-and-code-availability}{%
+\subsubsection*{Data and code availability}\label{data-and-code-availability}}
+
+Document metadata retrieved from Crossref does not appear to be covered by any copyright or other intellectual property restrictions.
+However, due to copyright restrictions, we are unable to make document fulltext or Web of Science search results publicly available.
+Code for the corpus assembly and text preparation steps described above is available by request.
+
+The public analysis repository, \url{https://github.com/dhicks/race-science} \emph{{[}DOI{]}}, includes document metadata and documentwise counts of NLP-extracted noun phrases (``document-term matrices''), along with code to reproduce the analysis of the following sections.
+\emph{{[}repo readme: requires R 4.1.2; rig (\url{https://github.com/r-lib/rig/}) for managing multiple R installations{]}}
+
+\hypertarget{vocabulary-selection}{%
+\subsection*{Vocabulary selection}\label{vocabulary-selection}}
+
+We took an information-theoretic approach to vocabulary selection.
+Consider a game in which I draw a document from the corpus, then a single token from that document.
+I tell you the term, and you have to guess which document I picked.
+Intuitively, highly informative terms (in this project, noun phrase types) are distinctive, allowing you to dramatically narrow down the list of potential documents.
+This ``informativeness'' of a term can be quantified as the KL (Kullback-Leibler) divergence from a ``baseline'' distribution of documents to the distribution conditional on the term.
+Because the most informative terms tend to be typos and OCR errors --- these are unique to a single document --- we multiply the KL divergence by the logarithm of the total number of occurrences of the term across the entire corpus.
+We refer to the resulting measure as \(log(n) \Delta H\) or simply \texttt{ndH}.
+
+\emph{{[}Hicks 2021?{]}} used the \texttt{ndH} approach in a topic modeling analysis, where the baseline distribution was the uniform distribution across documents.
+In the current project, we found that this approach heavily favored recurrent noun phrases in the longest documents.
+Many of the documents published in \emph{Psychological Reports} are extremely short, 1-2 reports of a single study; while many of the documents published in \emph{Brain and Behavioral Sciences} are book-length collections that include a long review article and sometimes dozens of commentaries.
+Very generic noun phrases that happen to appear in the latter can occur orders of magnitude more often than highly distinctive phrases in the former, and so the \(log(n)\) factor overwhelms the \(\Delta H\) factor.
+
+To address this, we used a different baseline distribution of documents, namely, one in which document probability is proportional to length.
+This makes phrases from short documents much more ``surprising'' (much less likely to occur according to the baseline), and hence substantially increases their informativeness.
+This was more effective at identifying useful phrases from across the corpus.
+
+A common rule of thumb in topic modeling is that the vocabulary should have about 10 times as many distinct terms as the number of documents in the corpus \emph{{[}cite{]}}.
+However, we had some concerns with computational demands here: the resulting document-term matrix would have roughly \(10 \times 33,000^2\) or 10.9 billion entries; with 10\% density this would require on the order of 4 GB of memory just for a single copy of the matrix; and most of the analysis was to be conducted on the authors' laptops.
+We therefore chose to work with three smaller vocabularies, \(5 \times\), \(1 \times\) and \(\frac{1/5} \times\) the number of documents.
+We refer to these as the ``large,'' ``medium,'' and ``small'' vocabulary, respectively, and they include \emph{{[}counts{]}} distinct phrases.
+We compare findings across vocabularies as a robustness check.
+
+\hypertarget{topic-modeling}{%
+\subsection*{Topic modeling}\label{topic-modeling}}
+
+To fit topic models, we followed the approach proposed by \emph{{[}cite{]}}, which uses varimax-rotated partial principal components instead of the variational inference methods used by standard topic model packages \emph{{[}eg, stm{]}}.
+This novel approach was implemented in the R package \texttt{tmfast}, \emph{{[}cite{]}}.
+A simulation study of \texttt{tmfast} found that it was significantly faster and only slightly less accurate at reconstructing known word-topic and topic-document distributions, compared to the standard topic modeling package \texttt{stm}.
+
+Topic models were fit for all three vocabularies (large, medium, and small) with \(k = 5, 10, 20, 30, 40, 50\) (number of topics), resulting in a total of 18 models.
+
+Following the approach of \emph{{[}Hicks 2021?{]}}, we did not attempt to identify a unique best model for further analysis.
+While we focus on the medium vocabulary, \(k=30\) model as the ``median'' among the 18 models fit, we compare and contrast findings from this model with those from the other models.
+
+\hypertarget{topic-model-interpretation}{%
+\subsubsection*{Topic model interpretation}\label{topic-model-interpretation}}
+
+After fitting topic models, our first research question was whether we could identify distinctive ``race science'' topics.
+We addressed this question in two ways.
+Following common practice in topic modeling, we extracted lists of the top \(n\) (highest-probability) terms (noun phrases) from each topic in each model, with \(n\) generally ranging between 5 and 15 as we explored the models.
+\emph{{[}tables + Silge plots{]}}
+
+\emph{{[}drop{]}}
+We also adapted the ``discursive space'' analysis proposed by \emph{{[}Hicks 2021?{]}}.
+This analysis combines the Hellinger distance between topic-document distributions --- capturing how ``far apart'' two documents are --- with dimensional reduction techniques used for data visualization.
+This produces a two-dimensional visualization of individual documents, arranged so that documents with similar topic distributions are visually clustered and visually distinct from documents with different topic distributions.
+\emph{{[}availability{]}}
+
+Specifically, for each fitted topic model, we extracted topic-document distributions for each document and calculated pairwise Hellinger distances.
+We then processed these distances using the UMAP algorithm for dimension reduction, as implemented in the \texttt{umap} package \emph{{[}details; cite{]}}.
+To reduce computational requirements for these steps, we first dropped documents published in the journal \emph{{[}Psychological Reports{]}} (\emph{{[}x documents removed, y remaining{]}}; note that the full corpus was used to fit the topic models, \emph{Psychological Reports} was removed just for this visualization).
+Using the \texttt{plotly} library \emph{{[}cite{]}}, we construct an interactive visualization: each point represents a single document, and points can be colored based on either their journal or highest-probability topic.
+Tooltips are displayed when we hover over a point, including document metadata (title, journal, authors, year) as well as ``key phrases.''
+
+To calculate key phrases for a given document \(d\), we first identified the highest-probability topic \(t\) for that document and the word-topic (noun phrase-topic) distribution \(\beta_{w,t}\) for that topic.
+We also retrieved the document-term matrix, containing the count \(n_{w,d}\) of occurrences for each word (noun phrase), and calculated a weighting \(n_{w,d} \beta_{w,t}\) for each word.
+The key phrases are the 5 most highly-weighted terms for the given document.
+
+\hypertarget{topic-quality-assessment}{%
+\subsubsection*{Topic quality assessment}\label{topic-quality-assessment}}
+
+\emph{{[}rewrite{]}}
+
+We next conducted a topic quality check of topic 19 from the medium vocabulary, \(k=30\) model.
+Keyword and inspection of the UMAP interactives seemed to indicate that this topic was a ``race science discourse'' topic in this model.
+
+To confirm this initial impression, a spreadsheet of all articles with \(\gamma > 0.25\) for this topic was extracted (excluding articles published in \emph{Psychological Reports}), and the top 200 articles were reviewed manually by both authors.
+For this qualitative analysis, we defined \emph{scientific racism} as (purporting) to justify racial inequality and colonialism by appealing to the epistemic authority of science; \emph{race science} as (pseudo-)scientific research that can be utilized for scientific racism; and \emph{race science discourse} as treating race science as a legitimate area of scientific research; this includes methodological critiques of race science and empirical tests that falsify race science hypotheses.
+Our qualitative review of the top-200 topic-19 articles focused on classifying them as \emph{race science discourse}.
+
+For the first round of review, both authors worked independently, dichotomously coding each article as race science discourse or not.
+We calculated the ``false positive rate'' --- documents with a very high \(\gamma\) in this topic that were not race science discourse --- and interrater reliability using Cohen's \(\kappa\).\\
+\emph{{[}false positive rate{]}}
+\emph{{[}Initial interrater reliability was low, \(\kappa = 0.4\), with 50 out of 200 articles classified discordantly.{]}} To understand the disagreement, both authors examined the 50 discordant articles using an interactive grounded theory approach.
+After this process, some initial codings were changed and \emph{{[}almost{]}} all of the remaining \emph{{[}n{]}} discordant articles were coded as follows:
+
+\begin{description}
+\tightlist
+\item[hereditarianism and/or eugenics]
+29 papers.
+On the ``eugenics'' side, mostly examines or otherwise discusses the ``dysgenic'' claim that ``low-intelligence'' people have more children than ``high-intelligence'' people.
+The ``hereditarianism'' side is more varied, with papers on environmentalist interventions to raise IQ, genetics of IQ, and several other sub-topics.
+\item[Flynn effect]
+15 papers. Supporting, challenging, or otherwise focusing on Flynn's findings that IQ has increased over time.
+\item[Lynn/national IQ]
+3 papers. Supports, challenges, or focuses on international IQ data/comparisons made by Richard Lynn.
+\end{description}
+
+We agreed that, for all three categories of papers, there was reasonable ambiguity about whether they should be classified as race science discourse.
+These papers either do not explicitly mention race, or do so with just a brief mention of the ``race and intelligence controversy.'' On the other hand, they either contribute to or legitimatize areas of research that are deeply entangled with race science, and might thereby contribute to public perceptions of a legitimate scientific debate about race and intelligence.
+
+Based on this qualitative review, we judged that the topic model approach was sufficiently sensitive (low ``false positive'' rate) in identifying race science discourse.
+
 \bibliography{references.bib}

### 2023-07-19 paper/supplement_paper.tex
mostly filling in supplement refs

@@
 
-\emph{{[}probably some introduction{]}}
-
 \hypertarget{corpus-assembly}{%
@@
 
-Corpus assembly is summarized in figure \emph{{[}xref{]}}. The corpus was assembled in two parts, one for \emph{Mankind Quarterly} and the other for mainstream journals.
+\begin{figure}
+\centering
+\includegraphics[width=8in,height=6.1in]{img/dataset.png}
+\caption{Corpus assembly. (\#fig:corpus)}
+\end{figure}
 
-\emph{{[}figure{]}}
+Corpus assembly is summarized in figure @ref(fig:corpus). The corpus was assembled in two parts, one for \emph{Mankind Quarterly} and the other for mainstream journals.
 
@@
 
-An electronic archive of \emph{Mankind Quarterly}, from the first issue in \emph{{[}1969{]}} through 2004, is available for free on the open web from \emph{The Unz Review}, which has been characterized as a white nationalist website by the Southern Poverty Law Center \emph{{[}\url{https://www.splcenter.org/hatewatch/2021/01/19/meet-white-nationalist-organizer-who-spewed-hate-against-lawmakers}{]}}.
-Web scraping packages and techniques \emph{{[}which?{]}} were used to retrieve every PDF available in this archive.
-These PDFs were processed using the open-source OCR (optical character recognition) software \texttt{tesseract} and the R API \texttt{rtesseract} \emph{{[}cites{]}}.
-This stage of corpus assembly was conducted in Fall 2021.
+An electronic archive of \emph{Mankind Quarterly}, from the first issue in 1969 through 2004, is available for free on the open web at the white nationalist website \emph{The Unz Review}.
+The Python package Beautiful Soup (\url{https://www.crummy.com/software/BeautifulSoup/}, version not recorded) was used to retrieve every PDF available in this archive.
+Text was extracted using the R package \texttt{pdftools} \cite[3.0.1]{OomsPdftoolsTextExtraction2023}.
+This stage of corpus assembly was conducted in Fall 2021 with the assistance of Anthony Sainez.
 
@@
 To identify suitable mainstream journals for inclusion in the corpus, we first identified academic researchers who had been funded by the Pioneer Fund.
-Pioneer is an American non-profit organization founded in 1937 to fund eugenics research \emph{{[}cite{]}}.
+Pioneer is an American non-profit organization founded in 1937 to fund eugenics research and propaganda \cite{TuckerFundingScientificRacism2002}.
 After the \emph{Brown v.
-Board of Education} ruling in 1954, Pioneer funded various segregationist efforts across the United States.
-\emph{{[}Shockley and Jensen{]}}.
-We reviewed a critical profile of Pioneer \emph{{[}cite{]}} as well as an archived page from the organization's own web site \emph{{[}cite{]}}, which together listed 16 researchers who had received Pioneer funds.
+Board of Education} ruling in 1954, Pioneer funded various segregationist efforts across the United States, including lectures on eugenics by Stanford physicist William Shockley \cite{JacksonScienceSegregationRace2005}.
+We reviewed a critical profile of Pioneer \cite{MillerPioneerFundBankrolling1994} as well as an archived page from the organization's own web site \url{https://web.archive.org/web/20130103005545/http://www.pioneerfund.org/Grantees.html}, which together listed 16 researchers who had received Pioneer funds.
 
@@
 
-All together, \emph{{[}corpus summary, total and by journal; table{]}}
+\emph{Behavioral and Brain Sciences} typically uses a target article + commentary format; in some cases there are dozens of commentaries for a single article.
+In the Crossref DOI metadata, each target article and individual commentary is given its own DOI, with no distinction between contribution types or metadata links between a commentary and the corresponding target article.
+But text is only available in aggregate PDFs that bundle together the target article and all of the commentaries.
+In a few cases this results in PDFs that are hundreds of pages long and might be linked from the Crossref metadata 30+ times.
+In addition, the retrieved PDFs are not perfectly identical, because Cambridge UP's servers add a timestamped watermark to each page.
+We attempted to contact Cambridge for assistance but did not receive a response.
+We ultimately used a series of ad hoc measures to mitigate text duplication.
+Approximately 100,000 documents of 1-2 pages long were excluded before PDF retrieval.
+After PDF retrieval and text extraction, the timestamped watermarks were removed using a regular expression and the text was hashed using SHA 256.
+Hashes were used to construct groups of duplicate documents, and a single document (whichever one happened to be first in the dataframe) was chosen from each hash group for inclusion in the corpus.
+It is plausible that some duplicates made it through this process: where the watermark overlapped with the text, the regular expression may have been unable to identify and remove the watermark (a false negative result), and this difference in the watermarks (not the text) would produce different hashes.
+
+\begin{figure}
+\centering
+\includegraphics[width=4.76in,height=2.6in]{img/02_count.png}
+\caption{Count of documents in the corpus, by journal and year. (\#fig:counts)}
+\end{figure}
+
+\begin{longtable}[]{@{}lrrr@{}}
+\caption{(\#tab:counts) Document counts, by journal, and years included in the corpus.}\tabularnewline
+\toprule
+journal & n & start & end \\
+\midrule
+\endfirsthead
+\toprule
+journal & n & start & end \\
+\midrule
+\endhead
+Behavior Genetics & 2268 & 1970 & 2010 \\
+Intelligence & 1237 & 1977 & 2010 \\
+Mankind Quarterly & 1821 & 1960 & 2004 \\
+Personality and Individual Differences & 7274 & 1980 & 2010 \\
+Psychological Reports & 21398 & 1960 & 2010 \\
+Behavioral and Brain Sciences & 898 & 1978 & 2010 \\
+\bottomrule
+\end{longtable}
+
+All together, the corpus includes 34,896 documents from 6 journals. See @ref(fig:counts) and @ref(tab:counts).
 

### 2023-07-19 scripts/02_describe_corpus.R

@@ -1,15 +1,17 @@
 renv::load(here::here())
- library(tidyverse)
+library(tidyverse)
 theme_set(theme_minimal())
+library(ggrepel)
 library(arrow)
 
 library(here)
 
 data_dir = here('data')
+out_dir = here('out')
 
 ## Load metadata ----
 meta_ar = open_dataset(here(data_dir, '01_metadata'))
-    # collect()
+# collect()
 
 nrow(meta_ar)
 
@@ -27,12 +29,62 @@ phrases_ar = phrases()
 # phrases_ar = read_csv(here(data_dir, '01_phrases.csv'))
 
 ## Descriptive plots ----
-meta_ar |>
+## Counts by journal ----
+count_df = meta_ar |>
     count(container.title, year) |>
+    collect()
+
+label_df = count_df |> 
+    group_by(container.title) |> 
+    slice_max(n, n = 1) |> 
+    mutate(container.title = str_wrap(container.title, width = 25))
+
+ggplot(count_df, 
+       aes(year, n, color = container.title, group = container.title)) +
+    geom_line(size = 1.25, color = 'black') +
+    geom_line(size = 1) +
+    geom_label_repel(
+        data = label_df, 
+        mapping = aes(fill = container.title, 
+                      label = container.title), 
+        hjust = 0, nudge_y = 20,
+        xlim = c(1960, 2100), direction = 'x',
+        color = 'black', alpha = .9) +
+    scale_color_viridis_d(guide = 'none', 
+                          aesthetics = c('fill', 'color')) +
+    # scale_y_sqrt() +
+    coord_cartesian(clip = 'off') +
+    theme(plot.margin = margin(t = 10, r = 50, b = 10, l = 10))
+
+ggsave(here(out_dir, '02_count.png'), 
+       width = 4.76, height = 3/4 * 4.76, scale = 1.75,
+       bg = 'white')
+
+meta_ar |> 
+    group_by(container.title) |> 
+    summarize(n = n(), 
+          start = min(year), 
+          end = max(year)) |> 
+    rename(journal = container.title) |> 
     collect() |>
-    ggplot(aes(year, n, color = container.title, group = container.title)) +
-    geom_line() +
-    geom_point()
+    knitr::kable(caption = '(\\#tab:counts) Document counts, by journal, and years included in the corpus.') |> 
+    write_lines(here(out_dir, '02_tab_counts.md'))
+
+
+## Phrases ----
[…]

### 2023-07-20 paper/paper.Rmd
~finished with supplement text

@@
     draft: true
-    number_sections: false
+    number_sections: true
 ---
@@
 
-# Introduction
+# Introduction {.unnumbered}
 
@@
 
-# *Brown v Board* and *Mankind Quarterly*
+# *Brown v Board* and *Mankind Quarterly* {.unnumbered}
 
@@
 
-# *Mankind Quarterly* and Pioneer-funded researchers
+# *Mankind Quarterly* and Pioneer-funded researchers {.unnumbered}
 
@@
 
-# Topic model analysis identifies race science discourse
+# Topic model analysis identifies race science discourse {.unnumbered}
 
-The fact that Pioneer-funded researchers published heavily in two mainstream psychology journals does not tell us anything about the content of their publications or the findings of their research.
+The fact that Pioneer-funded researchers published heavily in two mainstream psychology journals does not tell us anything about the content of their publications or the claims of their research.
 We assembled a full-text corpus of articles published in MQ, PID, I, BG, PR, and Behavior and Brain Sciences (BBS) between 1960-2010 and used topic modeling to identify topics discussing race, intelligence, and both.
@@
 
-# Race science and behavior genetics
+# Race science and behavior genetics {.unnumbered}
 
@@
 
-# Discussion
+# Discussion {.unnumbered}
 
@@
 We stress that we are not identifying any particular scientist as a racist, scientific racist, or race scientist based on the topic model results.
-Scientific research can be utilized to promote scientific racism — and thus count as race science — independently of the views and attitudes of the scientist who conducted that research.\cite{TaberyWhyStudyingGenetics2015, GillbornSoftlySoftlyGenetics2016, CarlsonQuantifyingContextualizingImpact2020, HennWhyDNANo2021}  But this does not mean that scientists are not responsible for the downstream social effects of their
+Scientific research can be utilized to promote scientific racism — and thus count as race science — independently of the views and attitudes of the scientist who conducted that research \cite{TaberyWhyStudyingGenetics2015, GillbornSoftlySoftlyGenetics2016, CarlsonQuantifyingContextualizingImpact2020, HennWhyDNANo2021}.
+But this does not mean that scientists are not responsible for the downstream social effects of their
 research, even when these effects involve deliberate misinterpretation.*[Douglas, Kitcher, Kourany, Havstad; Block and Dworkin?]*  Like other citizens, scientists are responsible for mitigating the reasonably foreseen harms that result from their actions.
@@
 
-## Limitations
+## Limitations {.unnumbered}
 
@@
 This reflects a recognition that the purpose of academic freedom is to promote the common good, rather than merely satisfy an individual's curiosity.
-But the promised benefits are either vague or do not actually require understanding the genetics of intelligence.\cite{KaplanGaltonQuincunxProbabilistic2021, CoopLotteryLuckLegacy2022}
+But the promised benefits are either vague or do not actually require understanding the genetics of intelligence \cite{KaplanGaltonQuincunxProbabilistic2021, CoopLotteryLuckLegacy2022}.
 
@@
 
+\clearpage
 \appendix
+\renewcommand{\thesection}{S\arabic{section}}
+\renewcommand\thefigure{\thesection.\arabic{figure}}    
+\setcounter{figure}{0} 
+\renewcommand\thetable{\thesection.\arabic{table}}    
+\setcounter{table}{0}
+\maketitle
 
@@
 After identifying some gaps (missing issues) in the archive from *The Unz Review*, we manually retrieved PDFs for further articles using ProQuest.
-Text was extracted using the R package `pdftools` \cite[3.0.1]{OomsPdftoolsTextExtraction2023}.
+Text was extracted using the R package `pdftools` \cite[version 3.0.1]{OomsPdftoolsTextExtraction2023}.
 This stage of corpus assembly was conducted in Fall 2021 with the assistance of Anthony Sainez.
@@
 The remaining 5 journals are all published by major academic publishers — Elsevier, Springer, or Cambridge University Press — and each item in the entire run of each journal has been assigned a DOI (digital object identifier) for archival purposes.
-We used the Crossref API and `rcrossref` R interface to this API \cite[1.1.0.99]{ChamberlainRcrossrefClientVarious2019} to retrieve metadata for each item published from 1960-2010 in each of these journals.
+We used the Crossref API and `rcrossref` R interface to this API \cite[version 1.1.0.99]{ChamberlainRcrossrefClientVarious2019} to retrieve metadata for each item published from 1960-2010 in each of these journals.
 These metadata included item-level license information — confirming that the text of each item could be used for text mining projects — and a URL to an electronic version of the item.
@@
 
-All together, the corpus includes 34,896 documents from 6 journals.  See \@ref(fig:counts) and \@ref(tab:counts).
+All together, the corpus includes 34,896 documents from 6 journals.  See figure \@ref(fig:counts) and table \@ref(tab:counts).
 
@@
 
-After document retrieval and text extraction, we pre-processed the text using the `spaCy` NLP (natural language processing) Python library \cite[1.9.0]{SpaCyIndustrialstrengthNatural2018} and the R API `spacyr` \cite[1.2.1]{BenoitSpacyrWrapperSpaCy2020}.
+After document retrieval and text extraction, we pre-processed the text using the `spaCy` NLP (natural language processing) Python library \cite[version 1.9.0]{SpaCyIndustrialstrengthNatural2018} and the R API `spacyr` \cite[version 1.2.1]{BenoitSpacyrWrapperSpaCy2020}.
 Specifically, we applied regular expressions to remove header/footer copyright notices and hyphenation, used spaCy to annotate and extract noun phrases (eg, "the intelligence test items"), and then cleaned and standardized these phrases (eg, removing the/an/a, coverting all text to lowercase, and replacing all whitespace with underscores: "intelligence_test_items").
 We then counted the occurrence of each noun phrase in the document.
-The aggregated "document-term matrix" was stored in Parquet format \cite{VohraApacheParquet2016} for performance reasons, and written and read using the `arrow` package for R \cite[7.0.0]{RichardsonArrowIntegrationApache2023}.
+The aggregated "document-term matrix" was stored in Parquet format \cite{VohraApacheParquet2016} for performance reasons, and written and read using the `arrow` package for R \cite[version 7.0.0]{RichardsonArrowIntegrationApache2023}.
 
@@
 
-We took an information-theoretic approach to vocabulary selection.\cite{HicksProductivityInterdisciplinaryImpacts2021}  Consider a game in which I draw a document from the corpus, then a single token from that document.
+We took an information-theoretic approach to vocabulary selection\cite{HicksProductivityInterdisciplinaryImpacts2021}.
+Consider a game in which I draw a document from the corpus, then a single token from that document.
 I tell you the term, and you have to guess which document I picked.
@@
 We therefore chose to work with three smaller vocabularies, $5 \times$, $1 \times$ and $\frac{1/5} \times$ the number of documents.
-We refer to these as the "large," "medium," and "small" vocabulary, respectively, and they include *[counts]* distinct phrases.
+We refer to these as the "large" (174,480 distinct phrases), "medium" (34,896) and "small" (6,979) vocabulary, respectively.
 We compare findings across vocabularies as a robustness check.
@@
 
-To fit topic models, we followed the approach proposed by *[cite]*, which uses varimax-rotated partial principal components instead of the variational inference methods used by standard topic model packages *[eg, stm]*.
-This novel approach was implemented in the R package `tmfast`, *[cite]*.
-A simulation study of `tmfast` found that it was significantly faster and only slightly less accurate at reconstructing known word-topic and topic-document distributions, compared to the standard topic modeling package `stm`.
+To fit topic models, we followed the approach proposed by Rohe and Zeng \cite{RoheVintageFactorAnalysis2020}, which uses varimax-rotated partial principal components instead of the variational inference methods used by standard topic model packages such as `stm` \cite{RobertsStmPackageStructural2019}.
+This novel approach was implemented in the R package `tmfast` \cite{HicksTmfastFitsTopic2023}.
+A simulation study of `tmfast` found that it was significantly faster and only slightly less accurate at reconstructing known word-topic and topic-document distributions, compared to `stm`.
 
-Topic models were fit for all three vocabularies (large, medium, and small) with $k = 5, 10, 20, 30, 40, 50$ (number of topics), resulting in a total of 18 models.
+Topic models were fit for all three vocabularies (large, medium, and small) with $k = 5, 10, 20, ..., 70$ (number of topics), resulting in a total of $24 = 3 \times 8$ models.
 
-Following the approach of *[Hicks 2021?]*, we did not attempt to identify a unique best model for further analysis.
-While we focus on the medium vocabulary, $k=30$ model as the "median" among the 18 models fit, we compare and contrast findings from this model with those from the other models.
+Following the approach of Hicks \cite{HicksProductivityInterdisciplinaryImpacts2021}, we did not attempt to identify a unique best fitted model for further analysis.
+While we focus on the medium vocabulary, $k=40$ model as the "median" among the 24 models fit, we compare and contrast findings from this model with those from the other models.
 
-### Topic model interpretation
+### Topic model interpretation and quality assessment
 
 After fitting topic models, our first research question was whether we could identify distinctive "race science" topics.
-We addressed this question in two ways.
-Following common practice in topic modeling, we extracted lists of the top $n$ (highest-probability) terms (noun phrases) from each topic in each model, with $n$ generally ranging between 5 and 15 as we explored the models.
-*[tables + Silge plots]*
+Figures \@ref(fig:gamma-lg), \@ref(fig:gamma-md), \@ref(fig:gamma-sm) show the gamma (topic-document) distributions for each value of $k$ for the three vocabularies.
+In these figures, each panel corresponds to a single journal-$k$ combination, each row of cells is a single document in that journal, and each column of singles is a single topic in that particular model.
+Topics correspond within columns of panels, but not within rows; for example, topic 04 for $k=10$ does not necessarily correspond to topic 04 for $k=20$.
+Color intensity indicates the value of gamma for that particular topic-document combination.
 
-*[drop]*
-We also adapted the "discursive space" analysis proposed by *[Hicks 2021?]*.
-This analysis combines the Hellinger distance between topic-document distributions — capturing how "far apart" two documents are — with dimensional reduction techniques used for data visualization.
-This produces a two-dimensional visualization of individual documents, arranged so that documents with similar topic distributions are visually clustered and visually distinct from documents with different topic distributions.
-*[availability]*
+![Fitted topic models: Gamma (topic-document) distributions, large vocabulary. (\#fig:gamma-lg)](img/05_lg_gamma.png){width="6in" height="4.8in"}
 
-Specifically, for each fitted topic model, we extracted topic-document distributions for each document and calculated pairwise Hellinger distances.
-We then processed these distances using the UMAP algorithm for dimension reduction, as implemented in the `umap` package *[details; cite]*.
-To reduce computational requirements for these steps, we first dropped documents published in the journal *[Psychological Reports]* (*[x documents removed, y remaining]*; note that the full corpus was used to fit the topic models, *Psychological Reports* was removed just for this visualization).
-Using the `plotly` library *[cite]*, we construct an interactive visualization:  each point represents a single document, and points can be colored based on either their journal or highest-probability topic.
-Tooltips are displayed when we hover over a point, including document metadata (title, journal, authors, year) as well as "key phrases."
+![Fitted topic models: Gamma (topic-document) distributions, medium vocabulary. (\#fig:gamma-md)](img/05_md_gamma.png){width="6in" height="4.8in"}
 
-To calculate key phrases for a given document $d$, we first identified the highest-probability topic $t$ for that document and the word-topic (noun phrase-topic) distribution $\beta_{w,t}$ for that topic.
-We also retrieved the document-term matrix, containing the count $n_{w,d}$ of occurrences for each word (noun phrase), and calculated a weighting $n_{w,d} \beta_{w,t}$ for each word.
-The key phrases are the 5 most highly-weighted terms for the given document.
+![Fitted topic models: Gamma (topic-document) distributions, small vocabulary. (\#fig:gamma-sm)](img/05_sm_gamma.png){width="6in" height="4.8in"}
 
-### Topic quality assessment
-*[rewrite]*
+For $k=5, 10$ the distributions are quite noisy, difficult to interpret, and generally don't form very strong visual clusters.
+As $k$ increases from 20 through 40, coherent bands start to appear for journal-distinctive topics in MQ, BG, I and, to a lesser degree, PID.
+Note that the distinctive topics for MQ and I do not obviously coincide.
+Above about $k=50$, the models appear to identify more fine-grained topics and the coherent bands fade.
 
-We next conducted a topic quality check of topic 19 from the medium vocabulary, $k=30$ model.
-Keyword and inspection of the UMAP interactives seemed to indicate that this topic was a "race science discourse" topic in this model.
+We next constructed Silge plots \cite{SilgeTopicModeling2017}, showing the top 15 (highest-probability) terms (noun phrases) from each topic in the model.
+These are provided as supplemental PDFs *[filenames]*.
+Perusing these term lists, we focused on topics in each model that contained racial terms.
+Every fitted topic model included at least one racial topic.
+Importantly, in some cases these topics contained terms related to intelligence research, but in other cases they did not.
 
-To confirm this initial impression, a spreadsheet of all articles with $\gamma > 0.25$ for this topic was extracted (excluding articles published in *Psychological Reports*), and the top 200 articles were reviewed manually by both authors.
-For this qualitative analysis, we defined *scientific racism* as (purporting) to justify racial inequality and colonialism by appealing to the epistemic authority of science; *race science* as (pseudo-)scientific research that can be utilized for scientific racism; and *race science discourse* as treating race science as a legitimate area of scientific research; this includes methodological critiques of race science and empirical tests that falsify race science hypotheses.
-Our qualitative review of the top-200 topic-19 articles focused on classifying them as *race science discourse*.
+The gamma distributions and Silge plots suggested that, once $k$ was sufficiently large, the topic models were distinguishing between two types of race science articles.
+We therefore examined the prevalence of topics that used racial terms, intelligence-research terms, or both, by journal and across all 24 models.
+Supplemental PDFs *[filenames]* provide these visualizations.
+In these visualizations, each row of facets corresponds to a value of $k$ (excluding 5 as too noisy), and each column of facets corresponds to one topic.
+Topics are clustered based on whether the top 15 terms contain racial terms ("race"), terms related to intelligence research ("intelligence"), or both.
+Line plots show the count of articles with $\gamma > 0.50$ *[confirm]* for the given topic, by journal; thick lines are 5-year running averages and thin lines are raw counts.
+
+Across all three vocabularies, for sufficiently large values of $k$ the models identify two or more different racial topics, one that appears all but exclusively in MQ and at least one that appears primarily in mainstream journals.
+
+We next conducted a topic quality check of topic 24 from the medium vocabulary, $k=40$ model, which appeared to identify race science research on intelligence published in mainstream journals.
+A spreadsheet of all articles that had a maximum value of $\gamma$ for this topic was extracted (excluding articles published in *Psychological Reports*), and the top 122 articles ($\gamma > 0.97$) were reviewed manually by both authors.
+We coded each article as *race science discourse*, or not, using our definition of race science discourse as treated race science as a legitimate area of scientific reearch; this includes methodological critiques of race science and empirical tests that falsify race science hypotheses.
 
 For the first round of review, both authors worked independently,  dichotomously coding each article as race science discourse or not.
-We calculated the "false positive rate" — documents with a very high $\gamma$ in this topic that were not race science discourse — and interrater reliability using Cohen's $\kappa$.
-*[false positive rate]* 
-*[Initial interrater reliability was low, $\kappa = 0.4$, with 50 out of 200 articles classified discordantly.]*  To understand the disagreement, both authors examined the 50 discordant articles using an interactive grounded theory approach.
-After this process, some initial codings were changed and *[almost]* all of the remaining *[n]* discordant articles were coded as follows:
-
-hereditarianism and/or eugenics
- ~ 29 papers.
-On the “eugenics” side, mostly examines or otherwise discusses the "dysgenic" claim that "low-intelligence" people have more children than "high-intelligence" people.
-The “hereditarianism” side is more varied, with papers on environmentalist interventions to raise IQ, genetics of IQ, and several other sub-topics.
-
-Flynn effect 
- ~ 15 papers.  Supporting, challenging, or otherwise focusing on Flynn’s findings that IQ has increased over time.
-
-Lynn/national IQ
- ~ 3 papers.  Supports, challenges, or focuses on international IQ data/comparisons made by Richard Lynn.
-
-We agreed that, for all three categories of papers, there was reasonable ambiguity about whether they should be classified as race science discourse.
-These papers either do not explicitly mention race, or do so with just a brief mention of the "race and intelligence controversy."  On the other hand, they either contribute to or legitimatize areas of research that are deeply entangled with race science, and might thereby contribute to public perceptions of a legitimate scientific debate about race and intelligence.
+We calculated the "false positive rate" — documents with maximum $\gamma$ in this topic that were not race science discourse — and interrater reliability using Cohen's $\kappa$.
+The consensus false positive rate was $8.3%$, with $98%$ agreement across the two raters ($\kappa = 0.86)$.
+Coding spreadsheets are included in the supplemental materials *[filenames]*.
+For each of the 10 consensus false documents, the inclusion was readily explained, but in different ways for different documents: some included text from other documents, as when an article started in the middle of a page; others discussed perceptions of race relations or racial animus; one piece in BG discussed differences in the herding behavior of two cattle breeds, referred to as "white" and "black."  The 3 non-consensus documents reflected essential ambiguity in the operationalization of race science discourse.
+One is a table of contents from a 1980 issue of I; arguably this document should have been excluded as irrelevant front matter, but arguably it legitimizes the research published in the issue.
+The second case is a paper on racial differences in emotional intelligence was classified as race science discourse by one author, while the other author felt it was more focused on psychometrics \cite{GignacGroupDifferencesEI2010}.
+A finally, a 1990 article in MQ (<https://www.unz.com/print/MankindQuarterly-1990q3-00108/>) proposed to offer an economic explanation for crime.
+While it has a few pages connecting race/ethnicity to both crime and poverty, it ultimately focuses on unemployment and alcohol sales, rather than race, as the key predictors of crime.
 
 Based on this qualitative review, we judged that the topic model approach was sufficiently sensitive (low "false positive" rate) in identifying race science discourse.
+However, "false negatives" are still possible, as some documents that human readers would classify as race science discourse might have been associated with other topics.
+
+\clearpage

### 2023-07-20 paper/paper.tex
~finished with supplement text

@@
 \section*{Introduction}\label{introduction}}
+\addcontentsline{toc}{section}{Introduction}
 
@@
 \section*{\texorpdfstring{\emph{Brown v Board} and \emph{Mankind Quarterly}}{Brown v Board and Mankind Quarterly}}\label{brown-v-board-and-mankind-quarterly}}
+\addcontentsline{toc}{section}{\emph{Brown v Board} and \emph{Mankind Quarterly}}
 
@@
 \section*{\texorpdfstring{\emph{Mankind Quarterly} and Pioneer-funded researchers}{Mankind Quarterly and Pioneer-funded researchers}}\label{mankind-quarterly-and-pioneer-funded-researchers}}
+\addcontentsline{toc}{section}{\emph{Mankind Quarterly} and Pioneer-funded researchers}
 
@@
 \section*{Topic model analysis identifies race science discourse}\label{topic-model-analysis-identifies-race-science-discourse}}
+\addcontentsline{toc}{section}{Topic model analysis identifies race science discourse}
 
-The fact that Pioneer-funded researchers published heavily in two mainstream psychology journals does not tell us anything about the content of their publications or the findings of their research.
+The fact that Pioneer-funded researchers published heavily in two mainstream psychology journals does not tell us anything about the content of their publications or the claims of their research.
 We assembled a full-text corpus of articles published in MQ, PID, I, BG, PR, and Behavior and Brain Sciences (BBS) between 1960-2010 and used topic modeling to identify topics discussing race, intelligence, and both.
@@
 \section*{Race science and behavior genetics}\label{race-science-and-behavior-genetics}}
+\addcontentsline{toc}{section}{Race science and behavior genetics}
 
@@
 \section*{Discussion}\label{discussion}}
+\addcontentsline{toc}{section}{Discussion}
 
@@
 We stress that we are not identifying any particular scientist as a racist, scientific racist, or race scientist based on the topic model results.
-Scientific research can be utilized to promote scientific racism --- and thus count as race science --- independently of the views and attitudes of the scientist who conducted that research.\cite{TaberyWhyStudyingGenetics2015, GillbornSoftlySoftlyGenetics2016, CarlsonQuantifyingContextualizingImpact2020, HennWhyDNANo2021} But this does not mean that scientists are not responsible for the downstream social effects of their
+Scientific research can be utilized to promote scientific racism --- and thus count as race science --- independently of the views and attitudes of the scientist who conducted that research \cite{TaberyWhyStudyingGenetics2015, GillbornSoftlySoftlyGenetics2016, CarlsonQuantifyingContextualizingImpact2020, HennWhyDNANo2021}.
+But this does not mean that scientists are not responsible for the downstream social effects of their
 research, even when these effects involve deliberate misinterpretation.\emph{{[}Douglas, Kitcher, Kourany, Havstad; Block and Dworkin?{]}} Like other citizens, scientists are responsible for mitigating the reasonably foreseen harms that result from their actions.
@@
 \subsection*{Limitations}\label{limitations}}
+\addcontentsline{toc}{subsection}{Limitations}
 
@@
 
+\clearpage
 \appendix
+\renewcommand{\thesection}{S\arabic{section}}
+\renewcommand\thefigure{\thesection.\arabic{figure}}    
+\setcounter{figure}{0} 
+\renewcommand\thetable{\thesection.\arabic{table}}    
+\setcounter{table}{0}
+\maketitle
 
 \hypertarget{materials-and-methods}{%
-\section*{Materials and methods}\label{materials-and-methods}}
+\section{Materials and methods}\label{materials-and-methods}}
 
 \hypertarget{corpus-assembly}{%
-\subsection*{Corpus assembly}\label{corpus-assembly}}
+\subsection{Corpus assembly}\label{corpus-assembly}}
 
@@
 \hypertarget{mankind-quarterly}{%
-\subsubsection*{Mankind Quarterly}\label{mankind-quarterly}}
+\subsubsection{Mankind Quarterly}\label{mankind-quarterly}}
 
@@
 The Python package Beautiful Soup (\url{https://www.crummy.com/software/BeautifulSoup/}, version not recorded) was used to retrieve every PDF available in this archive.
-Text was extracted using the R package \texttt{pdftools} \cite[3.0.1]{OomsPdftoolsTextExtraction2023}.
+After identifying some gaps (missing issues) in the archive from \emph{The Unz Review}, we manually retrieved PDFs for further articles using ProQuest.
+Text was extracted using the R package \texttt{pdftools} \cite[version 3.0.1]{OomsPdftoolsTextExtraction2023}.
 This stage of corpus assembly was conducted in Fall 2021 with the assistance of Anthony Sainez.
@@
 \hypertarget{mainstream-journals}{%
-\subsubsection*{Mainstream journals}\label{mainstream-journals}}
+\subsubsection{Mainstream journals}\label{mainstream-journals}}
 
@@
 Board of Education} ruling in 1954, Pioneer funded various segregationist efforts across the United States, including lectures on eugenics by Stanford physicist William Shockley \cite{JacksonScienceSegregationRace2005}.
-We reviewed a critical profile of Pioneer \cite{MillerPioneerFundBankrolling1994} as well as an archived page from the organization's own web site \url{https://web.archive.org/web/20130103005545/http://www.pioneerfund.org/Grantees.html}, which together listed 16 researchers who had received Pioneer funds.
+We reviewed a critical profile of Pioneer \cite{MillerPioneerFundBankrolling1994} as well as an archived page from the organization's own web site (\url{https://web.archive.org/web/20130103005545/http://www.pioneerfund.org/Grantees.html}), which together listed 16 researchers who had received Pioneer funds.
 
-We then used the author search tool in Clarivate's Web of Science platform \emph{{[}url{]}}, retrieving publication lists for 14 researchers.
+We then used the author search tool in Clarivate's Web of Science platform (\url{https://www.webofscience.com/wos/author/search}), retrieving publication lists for 14 researchers.
 These searches were conducted between 2021-09-24 and 2021-10-05 by DJH.
@@
 The remaining 5 journals are all published by major academic publishers --- Elsevier, Springer, or Cambridge University Press --- and each item in the entire run of each journal has been assigned a DOI (digital object identifier) for archival purposes.
-We used the Crossref API and \texttt{rcrossref} R interface to this API \emph{{[}cite{]}} to retrieve metadata for each item published from 1960-2010 in each of these 6 journals.
+We used the Crossref API and \texttt{rcrossref} R interface to this API \cite[version 1.1.0.99]{ChamberlainRcrossrefClientVarious2019} to retrieve metadata for each item published from 1960-2010 in each of these journals.
 These metadata included item-level license information --- confirming that the text of each item could be used for text mining projects --- and a URL to an electronic version of the item.
-These URLs were used to retrieve an HTML or PDF version of each item, except for \emph{Personality and Individual Differences}.
-This journal has published a relatively large number of non-article documents, such as book reviews and commentaries, which are not available at the URL included in the Crossref metadata.
-(This is unfortunate, as it was not difficult to find highly relevant documents that we could not automatically retrieve and therefore excluded from our corpus \emph{{[}review{]}}.) Instead we used Elsevier's ScienceDirect API \emph{{[}url{]}} to search and retrieve all items from \emph{Personality and Individual Differences} that are available.
-PDFs were run through OCR and text extraction as necessary, as with the \emph{Mankind Quarterly} documents.
+These URLs were used to retrieve a XML or PDF version of each item, except for \emph{Personality and Individual Differences}.
+This journal has published a relatively large number of non-article documents, such as book reviews and commentaries, that are not available at the URL included in the Crossref metadata.
+(This is unfortunate, as it was not difficult to find highly relevant documents that we could not automatically retrieve and therefore excluded from our corpus, such as an exchange between Rushton and Flynn on the Flynn effect \cite{FlynnEvidenceRushtonGenetic1998,RushtonSecularGainsIQ1998,FlynnReplyRushtonGang1998}.) Instead we used Elsevier's ScienceDirect API (\url{https://dev.elsevier.com/}) to independently search and retrieve all available items from \emph{Personality and Individual Differences}.
+This stage of corpus assembly was conducted between 2021-11 and 2022-05.
 
@@
 It is plausible that some duplicates made it through this process: where the watermark overlapped with the text, the regular expression may have been unable to identify and remove the watermark (a false negative result), and this difference in the watermarks (not the text) would produce different hashes.
+
+After XML/PDF versions were retrieved, text was extracted using either the \texttt{xml2} package \cite{WickhamXml2ParseXML2023} or \texttt{pdftools}.
+Figure \ref{fig:counts} and table \ref{tab:counts} show the number of fulltext documents included in the corpus.
 
@@
 
-All together, the corpus includes 34,896 documents from 6 journals. See \ref{fig:counts} and \ref{tab:counts}.
+All together, the corpus includes 34,896 documents from 6 journals. See figure \ref{fig:counts} and table \ref{tab:counts}.
 
 \hypertarget{text-preparation}{%
-\subsection*{Text preparation}\label{text-preparation}}
+\subsection{Text preparation}\label{text-preparation}}
 
-After document retrieval and text extraction, we pre-processed the text using the \texttt{spaCy} NLP (natural language processing) Python library and the R API \texttt{spacyr} \emph{{[}version, cites{]}}.
+After document retrieval and text extraction, we pre-processed the text using the \texttt{spaCy} NLP (natural language processing) Python library \cite[version 1.9.0]{SpaCyIndustrialstrengthNatural2018} and the R API \texttt{spacyr} \cite[version 1.2.1]{BenoitSpacyrWrapperSpaCy2020}.
 Specifically, we applied regular expressions to remove header/footer copyright notices and hyphenation, used spaCy to annotate and extract noun phrases (eg, ``the intelligence test items''), and then cleaned and standardized these phrases (eg, removing the/an/a, coverting all text to lowercase, and replacing all whitespace with underscores: ``intelligence\_test\_items'').
 We then counted the occurrence of each noun phrase in the document.
-The aggregated ``document-term matrix'' was stored in Parquet format \emph{{[}cite{]}} for performance reasons, and written and read using the \texttt{arrow} package for R \emph{{[}cite{]}}.
-\emph{{[}availability{]}}
+The aggregated ``document-term matrix'' was stored in Parquet format \cite{VohraApacheParquet2016} for performance reasons, and written and read using the \texttt{arrow} package for R \cite[version 7.0.0]{RichardsonArrowIntegrationApache2023}.
 
@@
 First, noun phrase extraction removes many standard stopwords (articles, common verbs) without relying on a fixed, a priori stopword list.
-Phrases can be more informative than single terms, for example, distinguishing ``intelligence test'' from ``hypothesis test.'' But simple n-gram extraction will include numerous phrases that are not especially meaningful.
-Consider the sentence ``Since our first analyses of feeding patterns in rats, we had been using a criterion of 40 minutes (Le Magnen \& Tallon 1963; 1966)'' *{[}10.1017\_\_\^{}\_\_s0140525x0000042x{]}*.
-Bigrams such as ``since our'' and ``criterion 40'' will likely be discarded in vocabulary selection, but significantly increase the computational cost of vocabulary selection.
+Phrases can be more informative than single terms, for example, distinguishing ``intelligence test'' from ``hypothesis test.'' While simple n-gram extraction will include numerous phrases that are not especially meaningful.
+Consider the sentence ``Since our first analyses of feeding patterns in rats, we had been using a criterion of 40 minutes (Le Magnen \& Tallon 1963; 1966).''\cite{SclafaniCorrelationCausationStudy1981} Bigrams such as ``since our'' and ``criterion 40'' will likely be discarded in vocabulary selection, but significantly increase the computational cost of vocabulary selection.
 Noun phrase extraction is therefore more efficient.
 
-After noun phrase extraction, the corpus comprised \emph{{[}tokens{]}} of \emph{{[}distinct noun phrases{]}}.
+After noun phrase extraction, the corpus comprised 43,162,055 total tokens of 6,320,783 distinct phrases.
 
 \hypertarget{data-and-code-availability}{%
-\subsubsection*{Data and code availability}\label{data-and-code-availability}}
+\subsubsection{Data and code availability}\label{data-and-code-availability}}
 
@@
 \hypertarget{vocabulary-selection}{%
-\subsection*{Vocabulary selection}\label{vocabulary-selection}}
+\subsection{Vocabulary selection}\label{vocabulary-selection}}
 
-We took an information-theoretic approach to vocabulary selection.
+We took an information-theoretic approach to vocabulary selection\cite{HicksProductivityInterdisciplinaryImpacts2021}.
 Consider a game in which I draw a document from the corpus, then a single token from that document.
@@
 
-\emph{{[}Hicks 2021?{]}} used the \texttt{ndH} approach in a topic modeling analysis, where the baseline distribution was the uniform distribution across documents.
-In the current project, we found that this approach heavily favored recurrent noun phrases in the longest documents.
-Many of the documents published in \emph{Psychological Reports} are extremely short, 1-2 reports of a single study; while many of the documents published in \emph{Brain and Behavioral Sciences} are book-length collections that include a long review article and sometimes dozens of commentaries.
+In the current project, we found that this \texttt{ndH} approach heavily favored recurrent noun phrases in the longest documents.
+Many of the documents published in \emph{Psychological Reports} are extremely short, 1-2 page brief notices of a single study; while many of the documents published in \emph{Brain and Behavioral Sciences} are book-length collections that include a long review article and sometimes dozens of commentaries.
 Very generic noun phrases that happen to appear in the latter can occur orders of magnitude more often than highly distinctive phrases in the former, and so the \(log(n)\) factor overwhelms the \(\Delta H\) factor.
@@
 
-A common rule of thumb in topic modeling is that the vocabulary should have about 10 times as many distinct terms as the number of documents in the corpus \emph{{[}cite{]}}.
+A common rule of thumb in topic modeling is that the vocabulary should have about 10 times as many distinct terms as the number of documents in the corpus.
 However, we had some concerns with computational demands here: the resulting document-term matrix would have roughly \(10 \times 33,000^2\) or 10.9 billion entries; with 10\% density this would require on the order of 4 GB of memory just for a single copy of the matrix; and most of the analysis was to be conducted on the authors' laptops.
 We therefore chose to work with three smaller vocabularies, \(5 \times\), \(1 \times\) and \(\frac{1/5} \times\) the number of documents.
-We refer to these as the ``large,'' ``medium,'' and ``small'' vocabulary, respectively, and they include \emph{{[}counts{]}} distinct phrases.
+We refer to these as the ``large'' (174,480 distinct phrases), ``medium'' (34,896) and ``small'' (6,979) vocabulary, respectively.
 We compare findings across vocabularies as a robustness check.
@@
 \hypertarget{topic-modeling}{%
-\subsection*{Topic modeling}\label{topic-modeling}}
+\subsection{Topic modeling}\label{topic-modeling}}
 
-To fit topic models, we followed the approach proposed by \emph{{[}cite{]}}, which uses varimax-rotated partial principal components instead of the variational inference methods used by standard topic model packages \emph{{[}eg, stm{]}}.
-This novel approach was implemented in the R package \texttt{tmfast}, \emph{{[}cite{]}}.
-A simulation study of \texttt{tmfast} found that it was significantly faster and only slightly less accurate at reconstructing known word-topic and topic-document distributions, compared to the standard topic modeling package \texttt{stm}.
+To fit topic models, we followed the approach proposed by Rohe and Zeng \cite{RoheVintageFactorAnalysis2020}, which uses varimax-rotated partial principal components instead of the variational inference methods used by standard topic model packages such as \texttt{stm} \cite{RobertsStmPackageStructural2019}.
+This novel approach was implemented in the R package \texttt{tmfast} \cite{HicksTmfastFitsTopic2023}.
+A simulation study of \texttt{tmfast} found that it was significantly faster and only slightly less accurate at reconstructing known word-topic and topic-document distributions, compared to \texttt{stm}.
 
-Topic models were fit for all three vocabularies (large, medium, and small) with \(k = 5, 10, 20, 30, 40, 50\) (number of topics), resulting in a total of 18 models.
+Topic models were fit for all three vocabularies (large, medium, and small) with \(k = 5, 10, 20, ..., 70\) (number of topics), resulting in a total of \(24 = 3 \times 8\) models.
 
-Following the approach of \emph{{[}Hicks 2021?{]}}, we did not attempt to identify a unique best model for further analysis.
-While we focus on the medium vocabulary, \(k=30\) model as the ``median'' among the 18 models fit, we compare and contrast findings from this model with those from the other models.
+Following the approach of Hicks \cite{HicksProductivityInterdisciplinaryImpacts2021}, we did not attempt to identify a unique best fitted model for further analysis.
+While we focus on the medium vocabulary, \(k=40\) model as the ``median'' among the 24 models fit, we compare and contrast findings from this model with those from the other models.
 
-\hypertarget{topic-model-interpretation}{%
-\subsubsection*{Topic model interpretation}\label{topic-model-interpretation}}
+\hypertarget{topic-model-interpretation-and-quality-assessment}{%
+\subsubsection{Topic model interpretation and quality assessment}\label{topic-model-interpretation-and-quality-assessment}}
 
 After fitting topic models, our first research question was whether we could identify distinctive ``race science'' topics.
-We addressed this question in two ways.
-Following common practice in topic modeling, we extracted lists of the top \(n\) (highest-probability) terms (noun phrases) from each topic in each model, with \(n\) generally ranging between 5 and 15 as we explored the models.
-\emph{{[}tables + Silge plots{]}}
+Figures \ref{fig:gamma-lg}, \ref{fig:gamma-md}, \ref{fig:gamma-sm} show the gamma (topic-document) distributions for each value of \(k\) for the three vocabularies.
+In these figures, each panel corresponds to a single journal-\(k\) combination, each row of cells is a single document in that journal, and each column of singles is a single topic in that particular model.
+Topics correspond within columns of panels, but not within rows; for example, topic 04 for \(k=10\) does not necessarily correspond to topic 04 for \(k=20\).
+Color intensity indicates the value of gamma for that particular topic-document combination.
 
-\emph{{[}drop{]}}
-We also adapted the ``discursive space'' analysis proposed by \emph{{[}Hicks 2021?{]}}.
-This analysis combines the Hellinger distance between topic-document distributions --- capturing how ``far apart'' two documents are --- with dimensional reduction techniques used for data visualization.
-This produces a two-dimensional visualization of individual documents, arranged so that documents with similar topic distributions are visually clustered and visually distinct from documents with different topic distributions.
-\emph{{[}availability{]}}
+\begin{figure}
+\centering
+\includegraphics[width=6in,height=4.8in]{img/05_lg_gamma.png}
+\caption{Fitted topic models: Gamma (topic-document) distributions, large vocabulary. \label{fig:gamma-lg}}
+\end{figure}
 
-Specifically, for each fitted topic model, we extracted topic-document distributions for each document and calculated pairwise Hellinger distances.
-We then processed these distances using the UMAP algorithm for dimension reduction, as implemented in the \texttt{umap} package \emph{{[}details; cite{]}}.
-To reduce computational requirements for these steps, we first dropped documents published in the journal \emph{{[}Psychological Reports{]}} (\emph{{[}x documents removed, y remaining{]}}; note that the full corpus was used to fit the topic models, \emph{Psychological Reports} was removed just for this visualization).
-Using the \texttt{plotly} library \emph{{[}cite{]}}, we construct an interactive visualization: each point represents a single document, and points can be colored based on either their journal or highest-probability topic.
-Tooltips are displayed when we hover over a point, including document metadata (title, journal, authors, year) as well as ``key phrases.''
+\begin{figure}
+\centering
+\includegraphics[width=6in,height=4.8in]{img/05_md_gamma.png}
+\caption{Fitted topic models: Gamma (topic-document) distributions, medium vocabulary. \label{fig:gamma-md}}
+\end{figure}
 
-To calculate key phrases for a given document \(d\), we first identified the highest-probability topic \(t\) for that document and the word-topic (noun phrase-topic) distribution \(\beta_{w,t}\) for that topic.
-We also retrieved the document-term matrix, containing the count \(n_{w,d}\) of occurrences for each word (noun phrase), and calculated a weighting \(n_{w,d} \beta_{w,t}\) for each word.
-The key phrases are the 5 most highly-weighted terms for the given document.
+\begin{figure}
+\centering
+\includegraphics[width=6in,height=4.8in]{img/05_sm_gamma.png}
+\caption{Fitted topic models: Gamma (topic-document) distributions, small vocabulary. \label{fig:gamma-sm}}
+\end{figure}
 
-\hypertarget{topic-quality-assessment}{%
-\subsubsection*{Topic quality assessment}\label{topic-quality-assessment}}
+For \(k=5, 10\) the distributions are quite noisy, difficult to interpret, and generally don't form very strong visual clusters.
+As \(k\) increases from 20 through 40, coherent bands start to appear for journal-distinctive topics in MQ, BG, I and, to a lesser degree, PID.
+Note that the distinctive topics for MQ and I do not obviously coincide.
+Above about \(k=50\), the models appear to identify more fine-grained topics and the coherent bands fade.
 
-\emph{{[}rewrite{]}}
+We next constructed Silge plots \cite{SilgeTopicModeling2017}, showing the top 15 (highest-probability) terms (noun phrases) from each topic in the model.
+These are provided as supplemental PDFs \emph{{[}filenames{]}}.
+Perusing these term lists, we focused on topics in each model that contained racial terms.
+Every fitted topic model included at least one racial topic.
+Importantly, in some cases these topics contained terms related to intelligence research, but in other cases they did not.
 
-We next conducted a topic quality check of topic 19 from the medium vocabulary, \(k=30\) model.
-Keyword and inspection of the UMAP interactives seemed to indicate that this topic was a ``race science discourse'' topic in this model.
+The gamma distributions and Silge plots suggested that, once \(k\) was sufficiently large, the topic models were distinguishing between two types of race science articles.
+We therefore examined the prevalence of topics that used racial terms, intelligence-research terms, or both, by journal and across all 24 models.
+Supplemental PDFs \emph{{[}filenames{]}} provide these visualizations.
+In these visualizations, each row of facets corresponds to a value of \(k\) (excluding 5 as too noisy), and each column of facets corresponds to one topic.
+Topics are clustered based on whether the top 15 terms contain racial terms (``race''), terms related to intelligence research (``intelligence''), or both.
+Line plots show the count of articles with \(\gamma > 0.50\) \emph{{[}confirm{]}} for the given topic, by journal; thick lines are 5-year running averages and thin lines are raw counts.
 
-To confirm this initial impression, a spreadsheet of all articles with \(\gamma > 0.25\) for this topic was extracted (excluding articles published in \emph{Psychological Reports}), and the top 200 articles were reviewed manually by both authors.
-For this qualitative analysis, we defined \emph{scientific racism} as (purporting) to justify racial inequality and colonialism by appealing to the epistemic authority of science; \emph{race science} as (pseudo-)scientific research that can be utilized for scientific racism; and \emph{race science discourse} as treating race science as a legitimate area of scientific research; this includes methodological critiques of race science and empirical tests that falsify race science hypotheses.
-Our qualitative review of the top-200 topic-19 articles focused on classifying them as \emph{race science discourse}.
+Across all three vocabularies, for sufficiently large values of \(k\) the models identify two or more different racial topics, one that appears all but exclusively in MQ and at least one that appears primarily in mainstream journals.
+
+We next conducted a topic quality check of topic 24 from the medium vocabulary, \(k=40\) model, which appeared to identify race science research on intelligence published in mainstream journals.
+A spreadsheet of all articles that had a maximum value of \(\gamma\) for this topic was extracted (excluding articles published in \emph{Psychological Reports}), and the top 122 articles (\(\gamma > 0.97\)) were reviewed manually by both authors.
+We coded each article as \emph{race science discourse}, or not, using our definition of race science discourse as treated race science as a legitimate area of scientific reearch; this includes methodological critiques of race science and empirical tests that falsify race science hypotheses.
 
 For the first round of review, both authors worked independently, dichotomously coding each article as race science discourse or not.
-We calculated the ``false positive rate'' --- documents with a very high \(\gamma\) in this topic that were not race science discourse --- and interrater reliability using Cohen's \(\kappa\).\\
-\emph{{[}false positive rate{]}}
-\emph{{[}Initial interrater reliability was low, \(\kappa = 0.4\), with 50 out of 200 articles classified discordantly.{]}} To understand the disagreement, both authors examined the 50 discordant articles using an interactive grounded theory approach.
-After this process, some initial codings were changed and \emph{{[}almost{]}} all of the remaining \emph{{[}n{]}} discordant articles were coded as follows:
-
-\begin{description}
-\tightlist
-\item[hereditarianism and/or eugenics]
-29 papers.
-On the ``eugenics'' side, mostly examines or otherwise discusses the ``dysgenic'' claim that ``low-intelligence'' people have more children than ``high-intelligence'' people.
-The ``hereditarianism'' side is more varied, with papers on environmentalist interventions to raise IQ, genetics of IQ, and several other sub-topics.
-\item[Flynn effect]
-15 papers. Supporting, challenging, or otherwise focusing on Flynn's findings that IQ has increased over time.
-\item[Lynn/national IQ]
-3 papers. Supports, challenges, or focuses on international IQ data/comparisons made by Richard Lynn.
-\end{description}
-
-We agreed that, for all three categories of papers, there was reasonable ambiguity about whether they should be classified as race science discourse.
-These papers either do not explicitly mention race, or do so with just a brief mention of the ``race and intelligence controversy.'' On the other hand, they either contribute to or legitimatize areas of research that are deeply entangled with race science, and might thereby contribute to public perceptions of a legitimate scientific debate about race and intelligence.
+We calculated the ``false positive rate'' --- documents with maximum \(\gamma\) in this topic that were not race science discourse --- and interrater reliability using Cohen's \(\kappa\).
+The consensus false positive rate was \(8.3%
+\), with \(98%
+\) agreement across the two raters (\(\kappa = 0.86)\).
+Coding spreadsheets are included in the supplemental materials \emph{{[}filenames{]}}.
+For each of the 10 consensus false documents, the inclusion was readily explained, but in different ways for different documents: some included text from other documents, as when an article started in the middle of a page; others discussed perceptions of race relations or racial animus; one piece in BG discussed differences in the herding behavior of two cattle breeds, referred to as ``white'' and ``black.'' The 3 non-consensus documents reflected essential ambiguity in the operationalization of race science discourse.
+One is a table of contents from a 1980 issue of I; arguably this document should have been excluded as irrelevant front matter, but arguably it legitimizes the research published in the issue.
+The second case is a paper on racial differences in emotional intelligence was classified as race science discourse by one author, while the other author felt it was more focused on psychometrics \cite{GignacGroupDifferencesEI2010}.
+A finally, a 1990 article in MQ (\url{https://www.unz.com/print/MankindQuarterly-1990q3-00108/}) proposed to offer an economic explanation for crime.
+While it has a few pages connecting race/ethnicity to both crime and poverty, it ultimately focuses on unemployment and alcohol sales, rather than race, as the key predictors of crime.
 
 Based on this qualitative review, we judged that the topic model approach was sufficiently sensitive (low ``false positive'' rate) in identifying race science discourse.
+However, ``false negatives'' are still possible, as some documents that human readers would classify as race science discourse might have been associated with other topics.
+
+\clearpage
 

### 2023-07-20 scripts/10_interrater_reliability.R

@@ -41,17 +41,17 @@ data |>
                  values_to = 'race_discourse') |> 
     ggplot(aes(idx, rater, fill = race_discourse)) +
     geom_tile()
-## "false positive" rate is 8.3%; 5% non-consensus
+## "false positive" rate is 8.3%; 1.7% non-consensus
 count(data, race_discourse_djh, race_discourse_el) |> 
     mutate(share = n / sum(n))
 
-## 95% agreement
+## 98% agreement
 agree(data[,c('race_discourse_djh','race_discourse_el')])
-## kappa 0.74
+## kappa 0.86
 kappa2(data[,c('race_discourse_djh','race_discourse_el')], 
        "unweighted")
 ## Bennett, Alpert and Goldstein's S
-## 0.90
+## 0.95
 (2 * agree(data[,c('race_discourse_djh','race_discourse_el')])$value/100 - 1) / (2-1)
 
 

### 2023-07-21 paper/paper.Rmd
mostly there?

@@
 
-*[incorporate def'ns]*
+We propose a distinction between scientific racism, race science, and race science discourse.
+*Scientific racism* refers to the social practice of purporting to justify racial inequality and colonialism by appealing to the epistemic authority of science.
+*Race science* refers to scientific and/or pseudoscientific research, along with its research products (journal articles, etc.), that can be utilized for scientific racism.
+*Race science discourse* characterizes a broader category, that treats race science as a legitimate area of scientific research.
+Race science discourse is a strict superset of race science, because the former includes methodological and empirical critiques of race science.
+For example, the Flynn effect (the finding that IQ scores have increased over time) poses a serious empirical challenge to hereditarian claims of racial differences in intelligence.
+But technical debates over the extent and causes of the Flynn effect likely create the impression among the general public of legitimate scientific disagreement, and in this way legitimize race science.
 
-scientific racism 
- ~ purporting to justify racial inequality and colonialism by appealing to the epistemic authority of science
+The distinction between scientific racism and race science allows us to bracket the intentions and mindset of the researchers involved in any particular piece of research.
+For the purposes of the current study, it is not important whether, for instance, Arthur Jensen is or is not correctly labelled a racist.
+It is important that much of Jensen's highest-profile research was useful and indeed utilized to purportedly justify racial inequality.
 
-race science
- ~ (pseudo-)scientific research that can be utilized for scientific racism
+# Scientific racism, eugenics, and *Brown* {.unnumbered}
 
-race science discourse
- ~ treating race science as a legitimate area of scientific research; this includes methodological critiques of race science and empirical tests that falsify race science hypotheses
-
-*[Pioneer]*
-
-*[probably don't include this]*
-- pre-20th century history of race concepts and race science
-- expansion and contraction of racialized terms over time
-
-<!-- # Race science 1910-1960 -->
-
-<!-- 
-## The rise and fall of eugenics
-
-*[write this]*
-
-- scientific racism is thoroughly entangled with eugenics, but still conceptually distinct [@SlobodianUnequalMindHow2023]
-    - sterilization laws focused on categories of criminality and mental ability, rather than race per se
-    - contemporary defenders of eugenics at least sometimes claim to oppose racism 
-    - claims of innate racial differences in intelligence can be made without dysgenic prophecies of a rapidly-expanding biological underclass
-
-- 1910: Davenport founds Eugenics Records Office at Cold Spring Harbor
-- 1926: NRC "Committee on the Negro" includes both Davenport and Boas: "neither was influential enough to veto his rival's participation" [@BarkanRetreatScientificRacism1992 113]
-- 1930s: *Coming of Age in Samoa* [@MeadComingAgeSamoa1928] popularizes cultural anthropology, "free[ing] anthropology from the shackles of biology" [@BarkanRetreatScientificRacism1992 134]
-- 1937: Pioneer Fund is founded to "support academic research and the \'[sic]dissemination of information, into the 'problem of heredity and eugenics' and 'the problems of race betterment'" [@MehlerFoundationFascismNew1989 21, quoting Laughlin]
-- 1940: Carnegie Institute defunds Eugenics Records Office
-- 1950: "For all practical social purposes 'race' is not so much a biological phenomenon as a social myth" [@BeagleholeRaceQuestion1950; though see @BrattainRaceRacismAntiracism2007]
- -->
-
-# *Brown v Board* and *Mankind Quarterly* {.unnumbered}
-
-Histories of scientific racism in the second half of the twentieth century have often emphasized the parascholarly journal *Mankind Quarterly* (MQ) \cite{
+Histories of scientific racism in the twentieth century have often emphasized the Pioneer Fund and the parascholarly journal *Mankind Quarterly* (MQ) \cite{
 MehlerFoundationFascismNew1989,
+BarkanRetreatScientificRacism1992,
 WinstonScienceServiceFar1998,
+TuckerFundingScientificRacism2002,
 SchafferScientificRacismAgain2007,
@@
 SainiDraperMillionsPhilanthropic2022
-}, with or without its connections to PF.
+}.
+Pioneer was formed in 1937, during the waning years of the eugenics movement in North America \cite{BarkanRetreatScientificRacism1992}, with the aim of "support[ing] academic research and the \'[sic]dissemination of information, into the 'problem of heredity and eugenics' and 'the problems of race betterment'" [@MehlerFoundationFascismNew1989 21, quoting Laughlin].
 MQ was founded in 1960 by biologist R.
@@
 Robert Gayre (self-styled as "Gayre of Gayre and Nigg"; 1907-1996).
-Gates' professional status was closely tied to eugenics and the explicit scientific racism of the 1920s, and he had been thoroughly marginalized in the wake of World War II \cite{WinstonScienceServiceFar1998}.
-In contrast, Garrett had been president of the American Psychological Association in 1946 and chair of Psychology at Columbia from 1941 to 1955 *[cite]*.
+Gates' professional status had risen and fallen with eugenics and the explicit scientific racism of the 1920s, and by the end of World War II he was thoroughly marginalized \cite{WinstonScienceServiceFar1998}.
+In contrast, Garrett had been president of the American Psychological Association in 1946 and chair of Psychology at Columbia from 1941 to 1955 \cite{WinstonScienceServiceFar1998}.
 
@@
 The Court's decision relied on expert testimony from psychologists and education researchers; but the segregationists had also put forward their own experts, including Henry Garrett \cite{WinstonScienceServiceFar1998, JacksonScienceSegregationRace2005, SchafferScientificRacismAgain2007}.
-Shortly after *Brown*, Garrett left Columbia and became involved in segregationist and eugenicist efforts, and even the neofascist Northern League.
 
 As part of the segregationist reaction to *Brown*, MQ was created as what contemporary historians and philosophers of science call an "echo chamber" \cite{FernandezPintoKnowBetterNot2017}, a favorable venue for race scientists to publish their views, on the grounds that an "equalitarian dogma" created a censorious "taboo" against their research in mainstream publications \cite{TuckerFundingScientificRacism2002, JacksonMythicalTabooRace2020}.
-Scholars have shown that echo chambers are a key part of the "tobacco strategy," used by numerous regulated industries — most infamously tobacco, but also fossil fuels and chemical manufacturing, among others — to "manufacture doubt" and delay regulation *[cites]*.
+Echo chambers are a key part of the "tobacco strategy," used by numerous regulated industries — most infamously tobacco, but also fossil fuels and chemical manufacturing, among others — to "manufacture doubt" and delay regulation \cite{OreskesMerchantsDoubtHow2011, JasnyEmpiricalExaminationEcho2015, FernandezPintoKnowBetterNot2017, WeatherallHowBeatScience2018}.
 
-In light of the origins of MQ, the significant attention paid to the journal by historians of scientific racism, and contemporary research on the "tobacco strategy," we hypothesized that bibliometric and text mining analyses would confirm that MQ functioned as an echo chamber for Pioneer-funded scientists.
+In light of the origins of Pioneer and MQ, the significant attention paid to the journal by historians of scientific racism, and contemporary research on the "tobacco strategy," we hypothesized that bibliometric and text mining analyses would confirm that MQ functioned as an echo chamber for Pioneer-funded scientists.
 This hypothesis was not supported.
@@
 Table: (\#tab:researchers) Pioneer-funded researchers.
-Either identified in *[cite]* or named on Pioneer's website, along with birth and death dates from Wikipedia, identified discipline, and WoS author search result counts.
+Either identified in \cite{MillerPioneerFundBankrolling1994} or named on an archive copy of Pioneer's website, along with birth and death dates from Wikipedia, attributed discipline, and WoS author search result counts.
 Brunetto Chiarelli does not have a Wikipedia page.
@@
 In addition, only psychologist Richard Lynn appears to have published heavily in MQ.
-Lynn has been an assistant editor of MQ since 1979 *[check this: https://rationalwiki.org/wiki/Mankind_Quarterly]* and president of PF since the death of psychologist J.
-Phillippe Rushton in 2012 *[cite]*.
+Lynn became an assistant editor of MQ in 1979 (vol.
+XX, Nos.
+1 & 2) and is listed as editor-in-chief on MQ's current website as of 2023-07-21.
+He has also been president of Pioneer since the death of psychologist J.
+Phillippe Rushton in 2012 \cite{BeirichPioneerFundAssets2013}.
 By contrast, a number of Pioneer-funded researchers have published in PID, I, and to a lesser degree BG: Bouchard, Eysenck, Jensen, Rushton, Vernon, and also Lynn.
@@
 PID was first published by Pergamon Press, a mainstream academic press, and today is published by Elsevier.
-According to the Scimago Journal Rankings, based on Scopus citation data, in 2022 PID had an H-index of 193, #29 in Psychology (Miscellaneous), and a 2-year citation index (impact factor) of 5.27, #24 in Psychology (Miscellaneous).
 
@@
 Richard Lynn joined the editorial board sometime between 1998 and 2002.
-(Archive copies of the I editorial board page are not available from the journal's website from 1999 through 2001.)  I has been criticized for including Lynn and Gerhard Meisenberg on its editorial board until July-August 2018 \cite{SainiSuperiorReturnRace2019}.
-Meisenberg is also connected with the Pioneer Fund *[director?]*, and was editor-in-chief of MQ in 2015-18 *[check these]*.
+(Archive copies of the I editorial board page are not available from the journal's website from 1999 through 2001.)  I has been criticized for including Lynn and Gerhard Meisenberg — who was editor-in-chief of MQ in 2015-18 — on its editorial board until 2018 \cite{SainiSuperiorReturnRace2019}.
 I is published by Elsevier.
-According to the Scimago Journal Rankings, based on Scopus citation data, in 2022 I had an H-index of 103, #32 in Experimental and Cognitive Psychology, and a 2-year citation index (impact factor) of 3.05, #31 in Experimental and Cognitive Psychology.
 
@@
 We assembled a full-text corpus of articles published in MQ, PID, I, BG, PR, and Behavior and Brain Sciences (BBS) between 1960-2010 and used topic modeling to identify topics discussing race, intelligence, and both.
-*[supplemental tables: Ns for each article, topic model parameters; supplemental figs: articles by pub over time, topic heatmaps, Silge plots, race-intelligence-both grids]*
 
@@
 
-We focus here on three topics identified in one of the 24 fitted topic models; see Figure \@(fig:focal).
-In this model, topic 07 is strongly associated with MQ:  MQ published dozens of articles in this topic each year, and no other journal ever published more than a handful, with most other journals publishing none in a given year.
-The Silge plot \cite{SilgeTopicModeling2017} contains racial terms (`races`, `whites`, `negroes` and potentially `europe`, `africa`, `india`, and `japan`) as well as `book`, likely reflecting the fact that MQ published a number of book reviews while the psychology journals did not.
+Figure \@ref(fig:focal) and this discussion focuses on three topics identified in one of the 24 fitted topic models; see figures \@ref(fig:gamma-lg), \@ref(fig:gamma-md), \@ref(fig:gamma-sm) and supplemental visualizations *[filenames]* for all topics from all models.
+In the focal model, topic 07 is strongly associated with MQ:  MQ published dozens of articles in this topic each year, and no other journal ever published more than a handful.
+The Silge plot of the top 15 terms \cite{SilgeTopicModeling2017} contains racial terms (`races`, `whites`, `negroes` and potentially `europe`, `africa`, `india`, and `japan`) as well as `book`, likely reflecting the fact that MQ published a number of book reviews, while the psychology journals either did not or these were not available for the corpus.
 
 Topic 22 is strongly associated with I and PID in the same way.
-The Silge plot does contain `jensen`, as well as a reference to Raymond Cattell, who played a major role in the development of factor analysis and intelligence testing also advocated eugenics, fascism, and Nazi race science\cite{MehlerBeyondismRaymondCattell1997}.
-However, the other authors named in this topic — John Horn and Peter Bentler — do not appear to have contributed to race science discourse.
-Instead this topic appears to identify "mainstream" (non-race science) intelligence research, specifically on factor analysis and the debate over whether intelligence in unidimensional or multidimensional.
+The Silge plot does contain `jensen`, as well as a reference to Raymond Cattell, who played a major role in the development of factor analysis and intelligence testing but also advocated for eugenics, fascism, and Nazi race science \cite{MehlerBeyondismRaymondCattell1997}.
+However, the other authors named in this topic — John Horn and Peter Bentler — do not appear to have contributed to scientific racism.
+Instead this topic appears to identify "mainstream" (non-race science) intelligence research, especially factor analysis and the debate over whether intelligence in unidimensional or multidimensional.
 (Three other topics only associated with mainstream intelligence journals and terms were also identified by this topic model.)  This topic indicates that the model is not simply lumping race science research together with other intelligence research.
@@
 The Silge plot for topic 24, by contrast, suggests a distinct race science discourse topic, with multiple racial terms (`whites`, `blacks`, `racial_differences`, `races`, `race_differences`) and the names of three prominent Pioneer-funded researchers, `jensen`, `lynn`, and `rushton`.
-To check this interpretation, we independently coded the top 121 papers in this topic (those with $gamma > 0.97$) as contributing to race science discourse or not.
-105 (87%) were coded as race science discourse by both authors; 10 (8.3%) were coded as not race science discourse by both; and discordant codes were given to the remaining 6 (5.0%) (95% agreement, Cohen's $\kappa = 0.74$).
+Independent qualitative coding of the top 121 papers in this topic (those with $gamma > 0.97$) confirmed this interpretation of the topic, with 108 (89%) coded as race science discourse by both authors (Cohen's $\kappa = 0.86$).
 
@@
 Jensen's "How Much Can We Boost IQ and Scholastic Achievement?" \cite{JensenHowMuchCan1969} was published in 1969 in Harvard Educational Review (not included in this study), and the time series indicates that, around this time, there was an increase in articles in topic 24 in both PS and MQ (the only two journals in our corpus that were active at the time).
-MQ shows another sharp increase in the late 1970s, around the time Lynn was brought on as assistant editor, *[confirming the claim that Lynn shifted MQ's focus to intelligence research]*.
+MQ shows another sharp increase in the late 1970s, around the time Lynn was brought on as assistant editor, consistent with the conclusion of a qualitative analysis by Adams and Pilloud that "Psychology became the dominant discipline upon the editorship of Richard Lynn and Gerhard Meisenberg in MQ, where some issues were almost entirely represented by IQ test–based studies" \cite[110]{AdamsMisAppropriationBiological2021}.
 PID published multiple papers in topic 24 almost immediately after it was founded, with I showing a more gradual increase between the mid-1970s and mid-1990s.
@@
 
-Notably, BG published very few articles in this topic.
-Outside of the field itself, behavior genetics is strongly associated with race science by both other academics and the general public.
+Among both academics and the general public, the field of behavior genetics is strongly associated with race science, and specifically race-and-intelligence research.
+It's therefore notably that BG published very few articles in any race-and-intelligence topic identified by any of the 24 topic models.
 Panofsky \cite{PanofskyMisbehavingScienceControversy2014} argues that, prior to Jensen's 1969 paper, behavior genetics emphasized the study of non-human animals and intentionally avoided associations with eugenics and public controversy more generally.
@@
 
-*[move most of this to supplement?]*
-To check this interpretation, we ran an independent search for documents with the keywords "race" and "intelligence" published in BG from 1972 to 2020, using Springer's journal search website.
-This search returned 125 documents.
-33 were presentation abstracts from meetings of the Behavior Genetics Association; we did not examine these further.
-14 of the remaining documents were published since 2010 (after the scope of the primary study).
-Among these, one obituary and two papers celebrated the work of John Loehlin, a prominent race-and-intelligence researcher who had been Director of the American Eugenics Society from 1968-1972 \cite{JohnLoehlin19262020, WaldmanIntroductionFestschriftJohn2014, PlominGenotypeEnvironmentCorrelationEra2014}.
-Another was a retrospective of the work of Lindon Eaves, a geneticist who made at least one notable contribution to debates on the Scarr-Rowe hypothesis (interactions between heritability, race, and class).
-None of these 14 documents reported any studies of racial differences.
-
-28 documents in this sample were published between 1990 and 2010.
-One was an obituary of Jerry Hirsch, a critic of hereditarianism in general and Jensen in particular \cite{RoubertouxJerryHirsch202008}, and another was an obituary of David Rowe \cite{RodgersObituaryDavidChristian2003}.
-Only 2 of these 28 documents reported race differences of any kind: one examining at interactions among race, sex, and heritability for adolescent BMI (Body Mass Index, used as a measure of overweight/obesity) \cite{JacobsonGeneticSharedEnvironmental1998}; and the other at interactions among race, family history of alcoholism, and visuospatial performance \cite{BermanReducedVisuospatialPerformance1995}.
-Finally, Philip Vernon — one of the Pioneer-funded scientists we identified earlier — published a critical review of a book that attempted to address hereditarianism, the Flynn effect, race-and-intelligence research, and some related issues.
-
-This supplemental review provides additional support for the interpretation that behavior genetics had distanced itself from race science by the 1990s.
-Notably, this review brought our attention to a disciplinary history of behavior genetics published in 2010 \cite{McGueEndBehavioralGenetics2010}, which acknowledges the historical association with eugenics but does not mention Jensen and other race science controversies since World War II.
+This interpretation was given additional support by a supplemental review of BG using Springer's journal search website; see supplemental text \@ref(bg).
+This review also brought to our attention two disciplinary histories of behavior genetics, one published in McGue in 2010 \cite{McGueEndBehavioralGenetics2010} and the other by Plomin in 2023 \cite{PlominCelebratingCenturyResearch2023}.
+McGue acknowledges the historical association with eugenics but does not mention Jensen and other race science controversies since World War II.
 Instead postwar behavior genetics is portrayed as an endeavor to counterbalance the hegemony of the "Blank Slate," the supposedly widespread view that genetic differences play no role in explaining individual differences \cite{BatesonCorpseWearisomeDebate2002}.
-And none of the three documents celebrating Loehlin noted his postwar involvement in the eugenics movement.
-None of the documents examined in this supplemental review explicitly repudiate race science.
-Similarly, another disciplinary history of behavior genetics, published by Plomin in 2023 \cite{PlominCelebratingCenturyResearch2023}, briefly mentions the controversies surrounding Jensen's 1969 paper and later *The Bell Curve*, writing that Jensen's work was controversial because he "touched on ethnic differences"; though ultimately Plomin reassures the reader that "by the 1990s, quantitative genetic research had convinced most scientists of the importance of genetics for behavioral traits, including intelligence."  In short, behavior geneticists seem to typically adopt a posture of objectivity and technical rigor, downplaying the extent of the field's involvement in race science in the past, and displacing the source of race science controversies to external, "political," and "irrational" actors.
+Plomin briefly mentions the controversies surrounding Jensen's 1969 paper and later *The Bell Curve*, writing that Jensen's work was controversial because he "touched on ethnic differences"; though ultimately Plomin reassures the reader that "by the 1990s, quantitative genetic research had convinced most scientists of the importance of genetics for behavioral traits, including intelligence."  All together, behavior geneticists acknowledge but then downplay the extent of the field's involvement in race science in the past, portraying themselves as models of scientific objectivity and technical rigor, while displacing the source of race science controversies on to external, "politically-motivated," and "irrational" actors.
 
@@
 
-Using bibliometric and text-mining methods, this study finds that Pioneer-funded researchers published heavily in certain mainstream psychology journals, rather than *Mankind Quarterly*; and that a distinct topic of race science discourse can be identified in these same journals across the period 1960-2010.
+Using bibliometric and text-mining methods, this study finds that Pioneer-funded researchers published heavily in certain mainstream psychology journals, much more than in *Mankind Quarterly*; and that a distinct topic of race science discourse, centered on race and intelligence, can be identified in these same journals across the period 1960-2010.
 These findings indicate that *Mankind Quarterly* (MQ) was less important as an echo chamber for late 20th century race science than mainstream psychology journals, especially *Intelligence* (I) and *Personality and Individual Differences* (PID).
@@
 But this does not mean that scientists are not responsible for the downstream social effects of their
-research, even when these effects involve deliberate misinterpretation.*[Douglas, Kitcher, Kourany, Havstad; Block and Dworkin?]*  Like other citizens, scientists are responsible for mitigating the reasonably foreseen harms that result from their actions.
-And we now have decades of empirical evidence that research on the genetics of intelligence — even research that does not itself touch on race — can and will be used to promote scientific racism \cite{MeyerWrestlingSocialBehavioral2023}.
-Specifically, while our findings suggest that behavior genetics has attempted to distance itself from scientific racism, behavior genetics research — and research on human genetics more generally — may be especially susceptible to appropriation by scientific racists \cite{CarlsonQuantifyingContextualizingImpact2020, OgbunugaforDNABasketballBirthday2022, *[Harden reviews]*}.
+research, even when these effects involve deliberate misinterpretation \cite{DouglasSciencePolicyValuefree2009, KitcherArgumentFreeInquiry1997, KouranyShouldKnowledgeBe2016, KouranyShouldKnowledgeBe2016, BlockIQHeritabilityInequality1974a}.
+Like other citizens, scientists are responsible for mitigating the reasonably foreseen harms that result from their actions.
+And we now have decades of empirical evidence that research on the genetics of intelligence — even research that does not itself touch on race — can and will be used to promote scientific racism and indeed racial violence \cite{MeyerWrestlingSocialBehavioral2023}.
+Specifically, while our findings suggest that behavior genetics has attempted to distance itself from scientific racism, it has done little to explicitly repudiate this research, especially in contrast to statements by biological anthropologists \cite{FuentesAAPAStatementRace2019} and human geneticists \cite{JacksonFacingOurHistory2023}.
+Behavior genetics research — and research on human genetics more generally — may be especially susceptible to appropriation by scientific racists, which only strengthens the obligations these researchers have to prevent their research from causing harm \cite{CarlsonQuantifyingContextualizingImpact2020, OgbunugaforDNABasketballBirthday2022}.
 
@@
 *[more?]*
-
-## Limitations {.unnumbered}
-
-A key limitation of the current study stems from the inability of text mining techniques to draw on a broader cultural context than what is represented in the corpus.
-Scholars of racism and US race relations have noted a shift in the way racist attitudes are understood and expressed, from *[something]* to "colorblind racism" *[Bonilla-Silva]*, defined as *[defn]*.
-Colorblind racism involves the use of racialized "codes" or "dogwhistles," language that is superficially non-racial but carries racial implications, such as (in the US context) "violent inner city criminals" or "welfare queen" \cite{SaulDogwhistlesPoliticalManipulation2018}.
-Alderfer notes that the racist implications of *The Bell Curve* only emerge gradually, with the opening of the book using the colorblind language of "group differences" \cite{AlderferScienceNonsciencePsychologists2003}.
-Panofsky et al.
-show how the term "human biodiversity," originally developed for anti-racist purposes by biological anthropologist Jonathan Marks, has been inverted by white supremacists and is now a dogwhistle for biological racial hierarchy \cite{PanofskyHowWhiteNationalists2021}.
-These cases suggest further that colorblindness can facilitate white supremacist appropriation of non-racial research — even contrary to the researchers' own intentions — as scientific racist readers re-interpret racial-neutral language as dogwhistles \cite{GillbornSoftlySoftlyGenetics2016, WillsAreClustersRaces2017}.
-
-Techniques such as topic modeling may be able to account for discursive shifts from explicit to coded racial language, but only if the corpus contains a critical mass of documents using both sets of terms together.
-This would seem to require either a relatively gradual transition, the inclusion of reflective commentary, such as scholarship identifying and analyzing the transition, or both.
-It is unlikely that these requirements are met by the corpus used in this study.
-So it is plausible that the topic models used here have some rate of "false negatives," documents in the corpus that engage in race science discourse but in a colorblind way that avoids the use of the explicit racial terms in the race-and-intelligence topic.
-
-The phenomenon of colorblind racism is closed related to ambiguities that cannot be resolved even by human coders familiar with the broader cultural context.
-Two relevant examples are the "Flynn effect," a secular increase in intelligence test scores noted by philosopher James Flynn; and studies of "national IQ," exemplified by the work of Richard Lynn.
-The Flynn effect implies that environmental factors can create group differences in IQ that are comparable to Black-White differences, and thus directly challenges claims that the racial differences are biological.
-Using methods widely regarded as ad hoc, cherry-picking, and generally unreliable, Lynn and collaborators have claimed that national IQ averages are correlated with national GDP.
-Both the Flynn effect and "national IQ" are strongly associated with race-and-intelligence discourse in the US cultural context.
-But scientific journal articles on these topics may not include any racial language at all.
-For example, a meta-analysis of the Flynn effect does not discuss racial differences or use terms such as "White" or "Black" outside the reference list \cite{TrahanFlynnEffectMetaanalysis2014}.
-Should such articles be coded as race science discourse?
-We expect that even readers who accept our definitions of race science and race science discourse will reasonably disagree on how to answer this question.
-Topic models, of course, are not even capable of representing this essential ambiguity.
 
@@
 \clearpage
+\pagenumbering{arabic}
+\renewcommand*{\thepage}{\thesection.\arabic{page}}
 \appendix
@@
 \setcounter{table}{0}
-\maketitle
 
@@
 
-![Count of documents in the corpus, by journal and year. (\#fig:counts)](img/02_count.png){width="4.76in" height="2.6in"}
+![Count of documents in the corpus, by journal and year. (\#fig:counts)](img/02_count.png){width="4.76in" height="3.6in"}
 
@@
 Topics are clustered based on whether the top 15 terms contain racial terms ("race"), terms related to intelligence research ("intelligence"), or both.
-Line plots show the count of articles with $\gamma > 0.50$ *[confirm]* for the given topic, by journal; thick lines are 5-year running averages and thin lines are raw counts.
+Line plots show the count of articles with $\gamma > 0.50$ for the given topic, by journal; thick lines are 5-year running averages and thin lines are raw counts.
 
@@
 We next conducted a topic quality check of topic 24 from the medium vocabulary, $k=40$ model, which appeared to identify race science research on intelligence published in mainstream journals.
-A spreadsheet of all articles that had a maximum value of $\gamma$ for this topic was extracted (excluding articles published in *Psychological Reports*), and the top 122 articles ($\gamma > 0.97$) were reviewed manually by both authors.
+A spreadsheet of all articles that had a maximum value of $\gamma$ for this topic was extracted (excluding articles published in *Psychological Reports*), and the top 121 articles ($\gamma > 0.97$) were reviewed manually by both authors.
 We coded each article as *race science discourse*, or not, using our definition of race science discourse as treated race science as a legitimate area of scientific reearch; this includes methodological critiques of race science and empirical tests that falsify race science hypotheses.
@@
 We calculated the "false positive rate" — documents with maximum $\gamma$ in this topic that were not race science discourse — and interrater reliability using Cohen's $\kappa$.
-The consensus false positive rate was $8.3%$, with $98%$ agreement across the two raters ($\kappa = 0.86)$.
+The consensus true positive rate was 89.3$ (108/121), while the consensus false positive rate was 8.3% (10/121), with $98%$ agreement across the two raters ($\kappa = 0.86)$.
 Coding spreadsheets are included in the supplemental materials *[filenames]*.
@@
 
+## Supplementary review of *Behavior Genetics* {#bg}
+
+To check the finding that very few race-and-intelligence documents had been published in BG, we ran an independent search for documents with the keywords "race" and "intelligence" published in BG from 1972 to 2020 using Springer's journal search website.
+This search returned 125 documents.
+33 were presentation abstracts from meetings of the Behavior Genetics Association; we did not examine these further.
+14 of the remaining documents were published since 2010 (after the scope of the primary study).
+Among these, one obituary and two papers celebrated the work of John Loehlin, a prominent race-and-intelligence researcher who had been Director of the American Eugenics Society from 1968-1972 \cite{JohnLoehlin19262020, WaldmanIntroductionFestschriftJohn2014, PlominGenotypeEnvironmentCorrelationEra2014}.
+Another was a retrospective of the work of Lindon Eaves, a geneticist who made at least one notable contribution to debates on the Scarr-Rowe hypothesis (interactions between heritability, race, and class).
+None of these 14 documents reported any studies of racial differences.
+
+28 documents in this sample were published between 1990 and 2010.
+One was an obituary of Jerry Hirsch, a critic of hereditarianism in general and Jensen in particular \cite{RoubertouxJerryHirsch202008}, and another was an obituary of David Rowe \cite{RodgersObituaryDavidChristian2003}.
+Only 2 of these 28 documents reported race differences of any kind: one examining at interactions among race, sex, and heritability for adolescent BMI (Body Mass Index, used as a measure of overweight/obesity) \cite{JacobsonGeneticSharedEnvironmental1998}; and the other at interactions among race, family history of alcoholism, and visuospatial performance \cite{BermanReducedVisuospatialPerformance1995}.
+Finally, Philip Vernon — one of the Pioneer-funded scientists we identified earlier — published a critical review of a book that attempted to address hereditarianism, the Flynn effect, race-and-intelligence research, and some related issues.
+
+At the same time, we were unable to find any articles published in BG that acknowledged the involvement of the field with scientific racism with anything like the force of the statements that have been made by biological anthropologists \cite{FuentesAAPAStatementRace2019} and human geneticists \cite{JacksonFacingOurHistory2023}.
+Notably, the ASHG report includes discussions of behavior genetics race science, and is critical of the field and the organization for failing to publicly reject the race-and-intelligence claims made by Shockley and Jensen in particular.
+A search for "eugenics" found articles that presented the issue as part of the distant past, and a search for "racism" turned up studies of racist attitudes.
+A statement on the Behavior Genetics Association website, dated 2021, acknowledges that "The history of our field is inextricably linked with racism, including the misuse of behavior genetic research to support violent eugenic policies," while primarily focusing on BGA member demographics (<https://www.bga.org/content.aspx?page_id=22&club_id=971921&module_id=567723>).
+
+## Limitations
+
+A key limitation of the current study stems from the inability of text mining techniques to draw on a broader cultural context than what is represented in the corpus.
+Scholars of racism and US race relations have noted a shift in the way racist attitudes are understood and expressed, from *[something]* to "colorblind racism" *[Bonilla-Silva]*, defined as *[defn]*.
+Colorblind racism involves the use of racialized "codes" or "dogwhistles," language that is superficially non-racial but carries racial implications, such as (in the US context) "violent inner city criminals" or "welfare queen" \cite{SaulDogwhistlesPoliticalManipulation2018}.
+Alderfer notes that the racist implications of *The Bell Curve* only emerge gradually, with the opening of the book using the colorblind language of "group differences" \cite{AlderferScienceNonsciencePsychologists2003}.
+Panofsky et al.
+show how the term "human biodiversity," originally developed for anti-racist purposes by biological anthropologist Jonathan Marks, has been inverted by white supremacists and is now a dogwhistle for biological racial hierarchy \cite{PanofskyHowWhiteNationalists2021}.
+These cases suggest further that colorblindness can facilitate white supremacist appropriation of non-racial research — even contrary to the researchers' own intentions — as scientific racist readers re-interpret racial-neutral language as dogwhistles \cite{GillbornSoftlySoftlyGenetics2016, WillsAreClustersRaces2017}.
+
+Techniques such as topic modeling may be able to account for discursive shifts from explicit to coded racial language, but only if the corpus contains a critical mass of documents using both sets of terms together.
+This would seem to require either a relatively gradual transition, the inclusion of reflective commentary, such as scholarship identifying and analyzing the transition, or both.
+It is unlikely that these requirements are met by the corpus used in this study.
+So it is plausible that the topic models used here have some rate of "false negatives," documents in the corpus that engage in race science discourse but in a colorblind way that avoids the use of the explicit racial terms in the race-and-intelligence topic.
+
+The phenomenon of colorblind racism is closed related to ambiguities that cannot be resolved even by human coders familiar with the broader cultural context.
+Two relevant examples are the "Flynn effect," a secular increase in intelligence test scores noted by philosopher James Flynn; and studies of "national IQ," exemplified by the work of Richard Lynn.
+The Flynn effect implies that environmental factors can create group differences in IQ that are comparable to Black-White differences, and thus directly challenges claims that the racial differences are biological.
+Using methods widely regarded as ad hoc, cherry-picking, and generally unreliable, Lynn and collaborators have claimed that national IQ averages are correlated with national GDP.
+Both the Flynn effect and "national IQ" are strongly associated with race-and-intelligence discourse in the US cultural context.
+But scientific journal articles on these topics may not include any racial language at all.
+For example, a meta-analysis of the Flynn effect does not discuss racial differences or use terms such as "White" or "Black" outside the reference list \cite{TrahanFlynnEffectMetaanalysis2014}.
+Should such articles be coded as race science discourse?
+We expect that even readers who accept our definitions of race science and race science discourse will reasonably disagree on how to answer this question.
+Topic models, of course, are not even capable of representing this essential ambiguity.
+
 \clearpage

### 2023-07-21 paper/paper.tex
mostly there?

@@
 % Pandoc citation processing
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+% for Pandoc 2.8 to 2.10.1
+\newenvironment{cslreferences}%
+  {}%
+  {\par}
+% For Pandoc 2.11+
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc} % for calculating minipage widths
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 
-\emph{{[}incorporate def'ns{]}}
+We propose a distinction between scientific racism, race science, and race science discourse.
+\emph{Scientific racism} refers to the social practice of purporting to justify racial inequality and colonialism by appealing to the epistemic authority of science.
+\emph{Race science} refers to scientific and/or pseudoscientific research, along with its research products (journal articles, etc.), that can be utilized for scientific racism.
+\emph{Race science discourse} characterizes a broader category, that treats race science as a legitimate area of scientific research.
+Race science discourse is a strict superset of race science, because the former includes methodological and empirical critiques of race science.
+For example, the Flynn effect (the finding that IQ scores have increased over time) poses a serious empirical challenge to hereditarian claims of racial differences in intelligence.
+But technical debates over the extent and causes of the Flynn effect likely create the impression among the general public of legitimate scientific disagreement, and in this way legitimize race science.
 
-\begin{description}
-\tightlist
-\item[scientific racism]
-purporting to justify racial inequality and colonialism by appealing to the epistemic authority of science
-\item[race science]
-(pseudo-)scientific research that can be utilized for scientific racism
-\item[race science discourse]
-treating race science as a legitimate area of scientific research; this includes methodological critiques of race science and empirical tests that falsify race science hypotheses
-\end{description}
+The distinction between scientific racism and race science allows us to bracket the intentions and mindset of the researchers involved in any particular piece of research.
+For the purposes of the current study, it is not important whether, for instance, Arthur Jensen is or is not correctly labelled a racist.
+It is important that much of Jensen's highest-profile research was useful and indeed utilized to purportedly justify racial inequality.
 
-\emph{{[}Pioneer{]}}
+\hypertarget{scientific-racism-eugenics-and-brown}{%
+\section*{\texorpdfstring{Scientific racism, eugenics, and \emph{Brown}}{Scientific racism, eugenics, and Brown}}\label{scientific-racism-eugenics-and-brown}}
+\addcontentsline{toc}{section}{Scientific racism, eugenics, and \emph{Brown}}
 
-\emph{{[}probably don't include this{]}}
-- pre-20th century history of race concepts and race science
-- expansion and contraction of racialized terms over time
-
-\hypertarget{brown-v-board-and-mankind-quarterly}{%
-\section*{\texorpdfstring{\emph{Brown v Board} and \emph{Mankind Quarterly}}{Brown v Board and Mankind Quarterly}}\label{brown-v-board-and-mankind-quarterly}}
-\addcontentsline{toc}{section}{\emph{Brown v Board} and \emph{Mankind Quarterly}}
-
-Histories of scientific racism in the second half of the twentieth century have often emphasized the parascholarly journal \emph{Mankind Quarterly} (MQ) \cite{
+Histories of scientific racism in the twentieth century have often emphasized the Pioneer Fund and the parascholarly journal \emph{Mankind Quarterly} (MQ) \cite{
 MehlerFoundationFascismNew1989,
+BarkanRetreatScientificRacism1992,
 WinstonScienceServiceFar1998,
+TuckerFundingScientificRacism2002,
 SchafferScientificRacismAgain2007,
@@
 SainiDraperMillionsPhilanthropic2022
-}, with or without its connections to PF.
+}.
+Pioneer was formed in 1937, during the waning years of the eugenics movement in North America \cite{BarkanRetreatScientificRacism1992}, with the aim of ``support{[}ing{]} academic research and the '{[}sic{]}dissemination of information, into the `problem of heredity and eugenics' and `the problems of race betterment'\,'' (Mehler 1989, 21, quoting Laughlin).
 MQ was founded in 1960 by biologist R.
@@
 Robert Gayre (self-styled as ``Gayre of Gayre and Nigg''; 1907-1996).
-Gates' professional status was closely tied to eugenics and the explicit scientific racism of the 1920s, and he had been thoroughly marginalized in the wake of World War II \cite{WinstonScienceServiceFar1998}.
-In contrast, Garrett had been president of the American Psychological Association in 1946 and chair of Psychology at Columbia from 1941 to 1955 \emph{{[}cite{]}}.
+Gates' professional status had risen and fallen with eugenics and the explicit scientific racism of the 1920s, and by the end of World War II he was thoroughly marginalized \cite{WinstonScienceServiceFar1998}.
+In contrast, Garrett had been president of the American Psychological Association in 1946 and chair of Psychology at Columbia from 1941 to 1955 \cite{WinstonScienceServiceFar1998}.
 
@@
 The Court's decision relied on expert testimony from psychologists and education researchers; but the segregationists had also put forward their own experts, including Henry Garrett \cite{WinstonScienceServiceFar1998, JacksonScienceSegregationRace2005, SchafferScientificRacismAgain2007}.
-Shortly after \emph{Brown}, Garrett left Columbia and became involved in segregationist and eugenicist efforts, and even the neofascist Northern League.
 
 As part of the segregationist reaction to \emph{Brown}, MQ was created as what contemporary historians and philosophers of science call an ``echo chamber'' \cite{FernandezPintoKnowBetterNot2017}, a favorable venue for race scientists to publish their views, on the grounds that an ``equalitarian dogma'' created a censorious ``taboo'' against their research in mainstream publications \cite{TuckerFundingScientificRacism2002, JacksonMythicalTabooRace2020}.
-Scholars have shown that echo chambers are a key part of the ``tobacco strategy,'' used by numerous regulated industries --- most infamously tobacco, but also fossil fuels and chemical manufacturing, among others --- to ``manufacture doubt'' and delay regulation \emph{{[}cites{]}}.
+Echo chambers are a key part of the ``tobacco strategy,'' used by numerous regulated industries --- most infamously tobacco, but also fossil fuels and chemical manufacturing, among others --- to ``manufacture doubt'' and delay regulation \cite{OreskesMerchantsDoubtHow2011, JasnyEmpiricalExaminationEcho2015, FernandezPintoKnowBetterNot2017, WeatherallHowBeatScience2018}.
 
-In light of the origins of MQ, the significant attention paid to the journal by historians of scientific racism, and contemporary research on the ``tobacco strategy,'' we hypothesized that bibliometric and text mining analyses would confirm that MQ functioned as an echo chamber for Pioneer-funded scientists.
+In light of the origins of Pioneer and MQ, the significant attention paid to the journal by historians of scientific racism, and contemporary research on the ``tobacco strategy,'' we hypothesized that bibliometric and text mining analyses would confirm that MQ functioned as an echo chamber for Pioneer-funded scientists.
 This hypothesis was not supported.
@@
 \caption{\label{tab:researchers} Pioneer-funded researchers.
-Either identified in \emph{{[}cite{]}} or named on Pioneer's website, along with birth and death dates from Wikipedia, identified discipline, and WoS author search result counts.
+Either identified in \cite{MillerPioneerFundBankrolling1994} or named on an archive copy of Pioneer's website, along with birth and death dates from Wikipedia, attributed discipline, and WoS author search result counts.
 Brunetto Chiarelli does not have a Wikipedia page.}\tabularnewline
@@
 In addition, only psychologist Richard Lynn appears to have published heavily in MQ.
-Lynn has been an assistant editor of MQ since 1979 \emph{{[}check this: \url{https://rationalwiki.org/wiki/Mankind_Quarterly}{]}} and president of PF since the death of psychologist J.
-Phillippe Rushton in 2012 \emph{{[}cite{]}}.
+Lynn became an assistant editor of MQ in 1979 (vol.~XX, Nos.
+1 \& 2) and is listed as editor-in-chief on MQ's current website as of 2023-07-21.
+He has also been president of Pioneer since the death of psychologist J.
+Phillippe Rushton in 2012 \cite{BeirichPioneerFundAssets2013}.
 By contrast, a number of Pioneer-funded researchers have published in PID, I, and to a lesser degree BG: Bouchard, Eysenck, Jensen, Rushton, Vernon, and also Lynn.
@@
 PID was first published by Pergamon Press, a mainstream academic press, and today is published by Elsevier.
-According to the Scimago Journal Rankings, based on Scopus citation data, in 2022 PID had an H-index of 193, \#29 in Psychology (Miscellaneous), and a 2-year citation index (impact factor) of 5.27, \#24 in Psychology (Miscellaneous).
 
@@
 Richard Lynn joined the editorial board sometime between 1998 and 2002.
-(Archive copies of the I editorial board page are not available from the journal's website from 1999 through 2001.) I has been criticized for including Lynn and Gerhard Meisenberg on its editorial board until July-August 2018 \cite{SainiSuperiorReturnRace2019}.
-Meisenberg is also connected with the Pioneer Fund \emph{{[}director?{]}}, and was editor-in-chief of MQ in 2015-18 \emph{{[}check these{]}}.
+(Archive copies of the I editorial board page are not available from the journal's website from 1999 through 2001.) I has been criticized for including Lynn and Gerhard Meisenberg --- who was editor-in-chief of MQ in 2015-18 --- on its editorial board until 2018 \cite{SainiSuperiorReturnRace2019}.
 I is published by Elsevier.
-According to the Scimago Journal Rankings, based on Scopus citation data, in 2022 I had an H-index of 103, \#32 in Experimental and Cognitive Psychology, and a 2-year citation index (impact factor) of 3.05, \#31 in Experimental and Cognitive Psychology.
 
@@
 We assembled a full-text corpus of articles published in MQ, PID, I, BG, PR, and Behavior and Brain Sciences (BBS) between 1960-2010 and used topic modeling to identify topics discussing race, intelligence, and both.
-\emph{{[}supplemental tables: Ns for each article, topic model parameters; supplemental figs: articles by pub over time, topic heatmaps, Silge plots, race-intelligence-both grids{]}}
 
@@
 
-We focus here on three topics identified in one of the 24 fitted topic models; see Figure @(fig:focal).
-In this model, topic 07 is strongly associated with MQ: MQ published dozens of articles in this topic each year, and no other journal ever published more than a handful, with most other journals publishing none in a given year.
-The Silge plot \cite{SilgeTopicModeling2017} contains racial terms (\texttt{races}, \texttt{whites}, \texttt{negroes} and potentially \texttt{europe}, \texttt{africa}, \texttt{india}, and \texttt{japan}) as well as \texttt{book}, likely reflecting the fact that MQ published a number of book reviews while the psychology journals did not.
+Figure \ref{fig:focal} and this discussion focuses on three topics identified in one of the 24 fitted topic models; see figures \ref{fig:gamma-lg}, \ref{fig:gamma-md}, \ref{fig:gamma-sm} and supplemental visualizations \emph{{[}filenames{]}} for all topics from all models.
+In the focal model, topic 07 is strongly associated with MQ: MQ published dozens of articles in this topic each year, and no other journal ever published more than a handful.
+The Silge plot of the top 15 terms \cite{SilgeTopicModeling2017} contains racial terms (\texttt{races}, \texttt{whites}, \texttt{negroes} and potentially \texttt{europe}, \texttt{africa}, \texttt{india}, and \texttt{japan}) as well as \texttt{book}, likely reflecting the fact that MQ published a number of book reviews, while the psychology journals either did not or these were not available for the corpus.
 
 Topic 22 is strongly associated with I and PID in the same way.
-The Silge plot does contain \texttt{jensen}, as well as a reference to Raymond Cattell, who played a major role in the development of factor analysis and intelligence testing also advocated eugenics, fascism, and Nazi race science\cite{MehlerBeyondismRaymondCattell1997}.
-However, the other authors named in this topic --- John Horn and Peter Bentler --- do not appear to have contributed to race science discourse.
-Instead this topic appears to identify ``mainstream'' (non-race science) intelligence research, specifically on factor analysis and the debate over whether intelligence in unidimensional or multidimensional.
+The Silge plot does contain \texttt{jensen}, as well as a reference to Raymond Cattell, who played a major role in the development of factor analysis and intelligence testing but also advocated for eugenics, fascism, and Nazi race science \cite{MehlerBeyondismRaymondCattell1997}.
+However, the other authors named in this topic --- John Horn and Peter Bentler --- do not appear to have contributed to scientific racism.
+Instead this topic appears to identify ``mainstream'' (non-race science) intelligence research, especially factor analysis and the debate over whether intelligence in unidimensional or multidimensional.
 (Three other topics only associated with mainstream intelligence journals and terms were also identified by this topic model.) This topic indicates that the model is not simply lumping race science research together with other intelligence research.
@@
 The Silge plot for topic 24, by contrast, suggests a distinct race science discourse topic, with multiple racial terms (\texttt{whites}, \texttt{blacks}, \texttt{racial\_differences}, \texttt{races}, \texttt{race\_differences}) and the names of three prominent Pioneer-funded researchers, \texttt{jensen}, \texttt{lynn}, and \texttt{rushton}.
-To check this interpretation, we independently coded the top 121 papers in this topic (those with \(gamma > 0.97\)) as contributing to race science discourse or not.
-105 (87\%) were coded as race science discourse by both authors; 10 (8.3\%) were coded as not race science discourse by both; and discordant codes were given to the remaining 6 (5.0\%) (95\% agreement, Cohen's \(\kappa = 0.74\)).
+Independent qualitative coding of the top 121 papers in this topic (those with \(gamma > 0.97\)) confirmed this interpretation of the topic, with 108 (89\%) coded as race science discourse by both authors (Cohen's \(\kappa = 0.86\)).
 
@@
 Jensen's ``How Much Can We Boost IQ and Scholastic Achievement?'' \cite{JensenHowMuchCan1969} was published in 1969 in Harvard Educational Review (not included in this study), and the time series indicates that, around this time, there was an increase in articles in topic 24 in both PS and MQ (the only two journals in our corpus that were active at the time).
-MQ shows another sharp increase in the late 1970s, around the time Lynn was brought on as assistant editor, \emph{{[}confirming the claim that Lynn shifted MQ's focus to intelligence research{]}}.
+MQ shows another sharp increase in the late 1970s, around the time Lynn was brought on as assistant editor, consistent with the conclusion of a qualitative analysis by Adams and Pilloud that ``Psychology became the dominant discipline upon the editorship of Richard Lynn and Gerhard Meisenberg in MQ, where some issues were almost entirely represented by IQ test--based studies'' \cite[110]{AdamsMisAppropriationBiological2021}.
 PID published multiple papers in topic 24 almost immediately after it was founded, with I showing a more gradual increase between the mid-1970s and mid-1990s.
@@
 
-\emph{{[}move most of this to supplement?{]}}
-To check this interpretation, we ran an independent search for documents with the keywords ``race'' and ``intelligence'' published in BG from 1972 to 2020, using Springer's journal search website.
-This search returned 125 documents.
-33 were presentation abstracts from meetings of the Behavior Genetics Association; we did not examine these further.
-14 of the remaining documents were published since 2010 (after the scope of the primary study).
-Among these, one obituary and two papers celebrated the work of John Loehlin, a prominent race-and-intelligence researcher who had been Director of the American Eugenics Society from 1968-1972 \cite{JohnLoehlin19262020, WaldmanIntroductionFestschriftJohn2014, PlominGenotypeEnvironmentCorrelationEra2014}.
-Another was a retrospective of the work of Lindon Eaves, a geneticist who made at least one notable contribution to debates on the Scarr-Rowe hypothesis (interactions between heritability, race, and class).
-None of these 14 documents reported any studies of racial differences.
-
-28 documents in this sample were published between 1990 and 2010.
-One was an obituary of Jerry Hirsch, a critic of hereditarianism in general and Jensen in particular \cite{RoubertouxJerryHirsch202008}, and another was an obituary of David Rowe \cite{RodgersObituaryDavidChristian2003}.
-Only 2 of these 28 documents reported race differences of any kind: one examining at interactions among race, sex, and heritability for adolescent BMI (Body Mass Index, used as a measure of overweight/obesity) \cite{JacobsonGeneticSharedEnvironmental1998}; and the other at interactions among race, family history of alcoholism, and visuospatial performance \cite{BermanReducedVisuospatialPerformance1995}.
-Finally, Philip Vernon --- one of the Pioneer-funded scientists we identified earlier --- published a critical review of a book that attempted to address hereditarianism, the Flynn effect, race-and-intelligence research, and some related issues.
-
-This supplemental review provides additional support for the interpretation that behavior genetics had distanced itself from race science by the 1990s.
-Notably, this review brought our attention to a disciplinary history of behavior genetics published in 2010 \cite{McGueEndBehavioralGenetics2010}, which acknowledges the historical association with eugenics but does not mention Jensen and other race science controversies since World War II.
+This interpretation was given additional support by a supplemental review of BG using Springer's journal search website; see supplemental text \ref{bg}.
+This review also brought to our attention two disciplinary histories of behavior genetics, one published in McGue in 2010 \cite{McGueEndBehavioralGenetics2010} and the other by Plomin in 2023 \cite{PlominCelebratingCenturyResearch2023}.
+McGue acknowledges the historical association with eugenics but does not mention Jensen and other race science controversies since World War II.
 Instead postwar behavior genetics is portrayed as an endeavor to counterbalance the hegemony of the ``Blank Slate,'' the supposedly widespread view that genetic differences play no role in explaining individual differences \cite{BatesonCorpseWearisomeDebate2002}.
-And none of the three documents celebrating Loehlin noted his postwar involvement in the eugenics movement.
-None of the documents examined in this supplemental review explicitly repudiate race science.
-Similarly, another disciplinary history of behavior genetics, published by Plomin in 2023 \cite{PlominCelebratingCenturyResearch2023}, briefly mentions the controversies surrounding Jensen's 1969 paper and later \emph{The Bell Curve}, writing that Jensen's work was controversial because he ``touched on ethnic differences''; though ultimately Plomin reassures the reader that ``by the 1990s, quantitative genetic research had convinced most scientists of the importance of genetics for behavioral traits, including intelligence.'' In short, behavior geneticists seem to typically adopt a posture of objectivity and technical rigor, downplaying the extent of the field's involvement in race science in the past, and displacing the source of race science controversies to external, ``political,'' and ``irrational'' actors.
+Plomin briefly mentions the controversies surrounding Jensen's 1969 paper and later \emph{The Bell Curve}, writing that Jensen's work was controversial because he ``touched on ethnic differences''; though ultimately Plomin reassures the reader that ``by the 1990s, quantitative genetic research had convinced most scientists of the importance of genetics for behavioral traits, including intelligence.'' All together, behavior geneticists acknowledge but then downplay the extent of the field's involvement in race science in the past, portraying themselves as models of scientific objectivity and technical rigor, while displacing the source of race science controversies on to external, ``politically-motivated,'' and ``irrational'' actors.
 
@@
 
-Using bibliometric and text-mining methods, this study finds that Pioneer-funded researchers published heavily in certain mainstream psychology journals, rather than \emph{Mankind Quarterly}; and that a distinct topic of race science discourse can be identified in these same journals across the period 1960-2010.
+Using bibliometric and text-mining methods, this study finds that Pioneer-funded researchers published heavily in certain mainstream psychology journals, much more than in \emph{Mankind Quarterly}; and that a distinct topic of race science discourse, centered on race and intelligence, can be identified in these same journals across the period 1960-2010.
 These findings indicate that \emph{Mankind Quarterly} (MQ) was less important as an echo chamber for late 20th century race science than mainstream psychology journals, especially \emph{Intelligence} (I) and \emph{Personality and Individual Differences} (PID).
@@
 But this does not mean that scientists are not responsible for the downstream social effects of their
-research, even when these effects involve deliberate misinterpretation.\emph{{[}Douglas, Kitcher, Kourany, Havstad; Block and Dworkin?{]}} Like other citizens, scientists are responsible for mitigating the reasonably foreseen harms that result from their actions.
-And we now have decades of empirical evidence that research on the genetics of intelligence --- even research that does not itself touch on race --- can and will be used to promote scientific racism \cite{MeyerWrestlingSocialBehavioral2023}.
-Specifically, while our findings suggest that behavior genetics has attempted to distance itself from scientific racism, behavior genetics research --- and research on human genetics more generally --- may be especially susceptible to appropriation by scientific racists \cite{CarlsonQuantifyingContextualizingImpact2020, OgbunugaforDNABasketballBirthday2022, *[Harden reviews]*}.
+research, even when these effects involve deliberate misinterpretation \cite{DouglasSciencePolicyValuefree2009, KitcherArgumentFreeInquiry1997, KouranyShouldKnowledgeBe2016, KouranyShouldKnowledgeBe2016, BlockIQHeritabilityInequality1974a}.
+Like other citizens, scientists are responsible for mitigating the reasonably foreseen harms that result from their actions.
+And we now have decades of empirical evidence that research on the genetics of intelligence --- even research that does not itself touch on race --- can and will be used to promote scientific racism and indeed racial violence \cite{MeyerWrestlingSocialBehavioral2023}.
+Specifically, while our findings suggest that behavior genetics has attempted to distance itself from scientific racism, it has done little to explicitly repudiate this research, especially in contrast to statements by biological anthropologists \cite{FuentesAAPAStatementRace2019} and human geneticists \cite{JacksonFacingOurHistory2023}.
+Behavior genetics research --- and research on human genetics more generally --- may be especially susceptible to appropriation by scientific racists, which only strengthens the obligations these researchers have to prevent their research from causing harm \cite{CarlsonQuantifyingContextualizingImpact2020, OgbunugaforDNABasketballBirthday2022}.
 
@@
 
-\hypertarget{limitations}{%
-\subsection*{Limitations}\label{limitations}}
-\addcontentsline{toc}{subsection}{Limitations}
-
-A key limitation of the current study stems from the inability of text mining techniques to draw on a broader cultural context than what is represented in the corpus.
-Scholars of racism and US race relations have noted a shift in the way racist attitudes are understood and expressed, from \emph{{[}something{]}} to ``colorblind racism'' \emph{{[}Bonilla-Silva{]}}, defined as \emph{{[}defn{]}}.
-Colorblind racism involves the use of racialized ``codes'' or ``dogwhistles,'' language that is superficially non-racial but carries racial implications, such as (in the US context) ``violent inner city criminals'' or ``welfare queen'' \cite{SaulDogwhistlesPoliticalManipulation2018}.
-Alderfer notes that the racist implications of \emph{The Bell Curve} only emerge gradually, with the opening of the book using the colorblind language of ``group differences'' \cite{AlderferScienceNonsciencePsychologists2003}.
-Panofsky et al.~show how the term ``human biodiversity,'' originally developed for anti-racist purposes by biological anthropologist Jonathan Marks, has been inverted by white supremacists and is now a dogwhistle for biological racial hierarchy \cite{PanofskyHowWhiteNationalists2021}.
-These cases suggest further that colorblindness can facilitate white supremacist appropriation of non-racial research --- even contrary to the researchers' own intentions --- as scientific racist readers re-interpret racial-neutral language as dogwhistles \cite{GillbornSoftlySoftlyGenetics2016, WillsAreClustersRaces2017}.
-
-Techniques such as topic modeling may be able to account for discursive shifts from explicit to coded racial language, but only if the corpus contains a critical mass of documents using both sets of terms together.
-This would seem to require either a relatively gradual transition, the inclusion of reflective commentary, such as scholarship identifying and analyzing the transition, or both.
-It is unlikely that these requirements are met by the corpus used in this study.
-So it is plausible that the topic models used here have some rate of ``false negatives,'' documents in the corpus that engage in race science discourse but in a colorblind way that avoids the use of the explicit racial terms in the race-and-intelligence topic.
-
-The phenomenon of colorblind racism is closed related to ambiguities that cannot be resolved even by human coders familiar with the broader cultural context.
-Two relevant examples are the ``Flynn effect,'' a secular increase in intelligence test scores noted by philosopher James Flynn; and studies of ``national IQ,'' exemplified by the work of Richard Lynn.
-The Flynn effect implies that environmental factors can create group differences in IQ that are comparable to Black-White differences, and thus directly challenges claims that the racial differences are biological.
-Using methods widely regarded as ad hoc, cherry-picking, and generally unreliable, Lynn and collaborators have claimed that national IQ averages are correlated with national GDP.
-Both the Flynn effect and ``national IQ'' are strongly associated with race-and-intelligence discourse in the US cultural context.
-But scientific journal articles on these topics may not include any racial language at all.
-For example, a meta-analysis of the Flynn effect does not discuss racial differences or use terms such as ``White'' or ``Black'' outside the reference list \cite{TrahanFlynnEffectMetaanalysis2014}.
-Should such articles be coded as race science discourse?
-We expect that even readers who accept our definitions of race science and race science discourse will reasonably disagree on how to answer this question.
-Topic models, of course, are not even capable of representing this essential ambiguity.
-
 \emph{{[}what can science do? Posselt, /Equity in Science/{]}}
@@
 \centering
-\includegraphics[width=4.76in,height=2.6in]{img/02_count.png}
+\includegraphics[width=4.76in,height=3.6in]{img/02_count.png}
 \caption{Count of documents in the corpus, by journal and year. \label{fig:counts}}
@@
 Topics are clustered based on whether the top 15 terms contain racial terms (``race''), terms related to intelligence research (``intelligence''), or both.
-Line plots show the count of articles with \(\gamma > 0.50\) \emph{{[}confirm{]}} for the given topic, by journal; thick lines are 5-year running averages and thin lines are raw counts.
+Line plots show the count of articles with \(\gamma > 0.50\) for the given topic, by journal; thick lines are 5-year running averages and thin lines are raw counts.
 
@@
 We next conducted a topic quality check of topic 24 from the medium vocabulary, \(k=40\) model, which appeared to identify race science research on intelligence published in mainstream journals.
-A spreadsheet of all articles that had a maximum value of \(\gamma\) for this topic was extracted (excluding articles published in \emph{Psychological Reports}), and the top 122 articles (\(\gamma > 0.97\)) were reviewed manually by both authors.
+A spreadsheet of all articles that had a maximum value of \(\gamma\) for this topic was extracted (excluding articles published in \emph{Psychological Reports}), and the top 121 articles (\(\gamma > 0.97\)) were reviewed manually by both authors.
 We coded each article as \emph{race science discourse}, or not, using our definition of race science discourse as treated race science as a legitimate area of scientific reearch; this includes methodological critiques of race science and empirical tests that falsify race science hypotheses.
@@
 We calculated the ``false positive rate'' --- documents with maximum \(\gamma\) in this topic that were not race science discourse --- and interrater reliability using Cohen's \(\kappa\).
-The consensus false positive rate was \(8.3%
-\), with \(98%
+The consensus true positive rate was 89.3\$ (108/121), while the consensus false positive rate was 8.3\% (10/121), with \(98%
 \) agreement across the two raters (\(\kappa = 0.86)\).
@@
 
+\hypertarget{bg}{%
+\subsection{\texorpdfstring{Supplementary review of \emph{Behavior Genetics}}{Supplementary review of Behavior Genetics}}\label{bg}}
+
+To check the finding that very few race-and-intelligence documents had been published in BG, we ran an independent search for documents with the keywords ``race'' and ``intelligence'' published in BG from 1972 to 2020 using Springer's journal search website.
+This search returned 125 documents.
+33 were presentation abstracts from meetings of the Behavior Genetics Association; we did not examine these further.
+14 of the remaining documents were published since 2010 (after the scope of the primary study).
+Among these, one obituary and two papers celebrated the work of John Loehlin, a prominent race-and-intelligence researcher who had been Director of the American Eugenics Society from 1968-1972 \cite{JohnLoehlin19262020, WaldmanIntroductionFestschriftJohn2014, PlominGenotypeEnvironmentCorrelationEra2014}.
+Another was a retrospective of the work of Lindon Eaves, a geneticist who made at least one notable contribution to debates on the Scarr-Rowe hypothesis (interactions between heritability, race, and class).
+None of these 14 documents reported any studies of racial differences.
+
+28 documents in this sample were published between 1990 and 2010.
+One was an obituary of Jerry Hirsch, a critic of hereditarianism in general and Jensen in particular \cite{RoubertouxJerryHirsch202008}, and another was an obituary of David Rowe \cite{RodgersObituaryDavidChristian2003}.
+Only 2 of these 28 documents reported race differences of any kind: one examining at interactions among race, sex, and heritability for adolescent BMI (Body Mass Index, used as a measure of overweight/obesity) \cite{JacobsonGeneticSharedEnvironmental1998}; and the other at interactions among race, family history of alcoholism, and visuospatial performance \cite{BermanReducedVisuospatialPerformance1995}.
+Finally, Philip Vernon --- one of the Pioneer-funded scientists we identified earlier --- published a critical review of a book that attempted to address hereditarianism, the Flynn effect, race-and-intelligence research, and some related issues.
+
+At the same time, we were unable to find any articles published in BG that acknowledged the involvement of the field with scientific racism with anything like the force of the statements that have been made by biological anthropologists \cite{FuentesAAPAStatementRace2019} and human geneticists \cite{JacksonFacingOurHistory2023}.
+Notably, the ASHG report includes discussions of behavior genetics race science, and is critical of the field and the organization for failing to publicly reject the race-and-intelligence claims made by Shockley and Jensen in particular.
+A search for ``eugenics'' found articles that presented the issue as part of the distant past, and a search for ``racism'' turned up studies of racist attitudes.
+A statement on the Behavior Genetics Association website, dated 2021, acknowledges that ``The history of our field is inextricably linked with racism, including the misuse of behavior genetic research to support violent eugenic policies,'' while primarily focusing on BGA member demographics (\url{https://www.bga.org/content.aspx?page_id=22\&club_id=971921\&module_id=567723}).
+
+\hypertarget{limitations}{%
+\subsection{Limitations}\label{limitations}}
+
+A key limitation of the current study stems from the inability of text mining techniques to draw on a broader cultural context than what is represented in the corpus.
+Scholars of racism and US race relations have noted a shift in the way racist attitudes are understood and expressed, from \emph{{[}something{]}} to ``colorblind racism'' \emph{{[}Bonilla-Silva{]}}, defined as \emph{{[}defn{]}}.
+Colorblind racism involves the use of racialized ``codes'' or ``dogwhistles,'' language that is superficially non-racial but carries racial implications, such as (in the US context) ``violent inner city criminals'' or ``welfare queen'' \cite{SaulDogwhistlesPoliticalManipulation2018}.
+Alderfer notes that the racist implications of \emph{The Bell Curve} only emerge gradually, with the opening of the book using the colorblind language of ``group differences'' \cite{AlderferScienceNonsciencePsychologists2003}.
+Panofsky et al.~show how the term ``human biodiversity,'' originally developed for anti-racist purposes by biological anthropologist Jonathan Marks, has been inverted by white supremacists and is now a dogwhistle for biological racial hierarchy \cite{PanofskyHowWhiteNationalists2021}.
+These cases suggest further that colorblindness can facilitate white supremacist appropriation of non-racial research --- even contrary to the researchers' own intentions --- as scientific racist readers re-interpret racial-neutral language as dogwhistles \cite{GillbornSoftlySoftlyGenetics2016, WillsAreClustersRaces2017}.
+
+Techniques such as topic modeling may be able to account for discursive shifts from explicit to coded racial language, but only if the corpus contains a critical mass of documents using both sets of terms together.
+This would seem to require either a relatively gradual transition, the inclusion of reflective commentary, such as scholarship identifying and analyzing the transition, or both.
+It is unlikely that these requirements are met by the corpus used in this study.
+So it is plausible that the topic models used here have some rate of ``false negatives,'' documents in the corpus that engage in race science discourse but in a colorblind way that avoids the use of the explicit racial terms in the race-and-intelligence topic.
+
+The phenomenon of colorblind racism is closed related to ambiguities that cannot be resolved even by human coders familiar with the broader cultural context.
+Two relevant examples are the ``Flynn effect,'' a secular increase in intelligence test scores noted by philosopher James Flynn; and studies of ``national IQ,'' exemplified by the work of Richard Lynn.
+The Flynn effect implies that environmental factors can create group differences in IQ that are comparable to Black-White differences, and thus directly challenges claims that the racial differences are biological.
+Using methods widely regarded as ad hoc, cherry-picking, and generally unreliable, Lynn and collaborators have claimed that national IQ averages are correlated with national GDP.
+Both the Flynn effect and ``national IQ'' are strongly associated with race-and-intelligence discourse in the US cultural context.
+But scientific journal articles on these topics may not include any racial language at all.
+For example, a meta-analysis of the Flynn effect does not discuss racial differences or use terms such as ``White'' or ``Black'' outside the reference list \cite{TrahanFlynnEffectMetaanalysis2014}.
+Should such articles be coded as race science discourse?
+We expect that even readers who accept our definitions of race science and race science discourse will reasonably disagree on how to answer this question.
+Topic models, of course, are not even capable of representing this essential ambiguity.
+
 \clearpage
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-MehlerFoundationFascismNew1989}{}}%
+Mehler, Barry. 1989. {``Foundation for Fascism: {The} New Eugenics Movement in the {United States}.''} \emph{Patterns of Prejudice} 23 (4): 17--25. \url{https://doi.org/10.1080/0031322X.1989.9970026}.
+
+\end{CSLReferences}
 

### 2023-08-04 paper/paper.Rmd
paper edits, bib file additions
edited paper to revise section on RS and BG, reducing discussion about BG and describing psychology as the discipline where RS is being maintained as per our results

@@
 
-# Race science and behavior genetics {.unnumbered}
+# Race science, behavior genetics, and psychology {.unnumbered}
 
+It's notable that BG published very few articles in any race-and-intelligence topic identified by any of the 24 topic models.
 Among both academics and the general public, the field of behavior genetics is strongly associated with race science, and specifically race-and-intelligence research.
+Panofsky \cite{PanofskyMisbehavingScienceControversy2014} argues that, prior to Jensen's 1969 paper, behavior genetics emphasized the study of non-human animals and intentionally avoided associations with eugenics and public controversy more generally.
+In response to Jensen, critics such as Lewontin offered broad critiques of behavior genetics as such, and behavior geneticists in turn adopted a radical conception of academic freedom (without any sense of responsibility for the social implications of academic research) and a siege or wartime mentality, as demonstrated by Sandra Scarr's 1986 presidential address to the Behavior Genetics Association (BGA) \cite{ScarrThreeCheersBehavior1987}.
+Scarr's address coincided with the period between 1970-1990 when BG published articles in topic 24 (including Scarr's address itself).
+However,
+
+The topic model analysis suggests that by the 1990s, behavior genetics may have distanced itself somewhat from race science, albeit without directly repudiating it.
+This interpretation is further supported by two disciplinary histories of behavior genetics, one published in McGue in 2010 \cite{McGueEndBehavioralGenetics2010} and the other by Plomin in 2023 \cite{PlominCelebratingCenturyResearch2023}.
+In addition, the 1995 BGA presidential address by Glayde Whitney - in which he criticized the "Marixist-Lysenkoist denial of genetics" and proposed that differences in murder rates between countries and cities was caused by racial genetic differences in intelligence, empathy, aggression, and impulsivity - was published in MQ rather than BG.
+As a discipline, behavior genetics may have been prominent in forming an echo chamber for race science in the past, but has been less receptive to such ideas in more recent history.
+
+Instead, our results suggest that psychology has been the dominant discipline in which the race science echo chamber is being maintained.
+This may be the result of racial essentialization being maintained more easily in a discipline like psychology than in other disciplines that contribute to behavior genetics, like genetics and anthropology.
+The typical practitioners in psychology do not have the degree of formal education about, and therefore knowledge of, relevant underlying genetic and anthropological research that has led many professional organizations in genetics and anthropology to formally reject the claim of race as meaningful biological concept \cite{anthropology1996aapa, american2018ashg}, undermining the core assumption necessary for race science to exist.
+As described by Winston, however, "[s]ince the 1990s, scientific racism has evolved in ways that sidestep or blunt the charge of essentialist categorization" (p.
+427, \cite{WinstonWhyMainstreamResearch2020}), such as shifting the rhetoric away from race to concepts such as "national IQ" that are deployed to preserve an essentialist categorization of different human populations.
+Professional psychological organizations routinely publish statements rejecting racism or acknowledging psychology's complicity in systemic racism \cite{american2021apology, aps2021statement}, but stop short of rejecting race as a meaningful biological categorization of people, allowing for race science research to be smuggled into psychology and given the appearance of continued scientific legitimacy.
+
+<!-- Among both academics and the general public, the field of behavior genetics is strongly associated with race science, and specifically race-and-intelligence research.
 It's therefore notably that BG published very few articles in any race-and-intelligence topic identified by any of the 24 topic models.
@@
 Scarr's address coincided with the period between 1970-1990 when BG published articles in topic 24 (including Scarr's address itself).
+-->
 
-However, Glayde Whitney's 1995 BGA presidential address was published in MQ rather than BG.
+<!-- However, Glayde Whitney's 1995 BGA presidential address was published in MQ rather than BG.
 According to Whitney, BG's editor refused to publish the address and the BGA Executive Committee (except for Whitney himself) voted to "issue an official statement of denouncement" of the address; though we were unable to find any such a statement, and Panofsky reports that "nothing official was done to Whitney."  In his address, Whitney criticized "the Marxist-Lysenkoist denial of genetics"; quoted Peter Brimelow — a white nationalist and later founder of the racist and xenophobic website VDARE — defining "racist" as "anyone who is winning an argument with a liberal"; and proposed as a "reasonable scientific hypothesis" that differences in murder rates between countries and cities was caused by racial genetic differences in intelligence, empathy, aggression, and impulsivity.
+-->
 
-The topic model analysis and the case of Whitney's address suggest that, by the 1990s, behavior genetics may have somewhat distanced itself from race science, though perhaps without directly repudiating it \cite{GillbornSoftlySoftlyGenetics2016, HennWhyDNANo2021}.
+<!-- The topic model analysis and the case of Whitney's address suggest that, by the 1990s, behavior genetics may have somewhat distanced itself from race science, though perhaps without directly repudiating it \cite{GillbornSoftlySoftlyGenetics2016, HennWhyDNANo2021}.
+-->
 
-This interpretation was given additional support by a supplemental review of BG using Springer's journal search website; see supplemental text \@ref(bg).
+<!-- This interpretation was given additional support by a supplemental review of BG using Springer's journal search website; see supplemental text \@ref(bg).
 This review also brought to our attention two disciplinary histories of behavior genetics, one published in McGue in 2010 \cite{McGueEndBehavioralGenetics2010} and the other by Plomin in 2023 \cite{PlominCelebratingCenturyResearch2023}.
@@
 Plomin briefly mentions the controversies surrounding Jensen's 1969 paper and later *The Bell Curve*, writing that Jensen's work was controversial because he "touched on ethnic differences"; though ultimately Plomin reassures the reader that "by the 1990s, quantitative genetic research had convinced most scientists of the importance of genetics for behavioral traits, including intelligence."  All together, behavior geneticists acknowledge but then downplay the extent of the field's involvement in race science in the past, portraying themselves as models of scientific objectivity and technical rigor, while displacing the source of race science controversies on to external, "politically-motivated," and "irrational" actors.
+-->
 

### 2023-08-30 paper/paper.Rmd
citations working again?

@@
 Professional psychological organizations routinely publish statements rejecting racism or acknowledging psychology's complicity in systemic racism \cite{american2021apology, aps2021statement}, but stop short of rejecting race as a meaningful biological categorization of people, allowing for race science research to be smuggled into psychology and given the appearance of continued scientific legitimacy.
-
-<!-- Among both academics and the general public, the field of behavior genetics is strongly associated with race science, and specifically race-and-intelligence research.
-It's therefore notably that BG published very few articles in any race-and-intelligence topic identified by any of the 24 topic models.
-Panofsky \cite{PanofskyMisbehavingScienceControversy2014} argues that, prior to Jensen's 1969 paper, behavior genetics emphasized the study of non-human animals and intentionally avoided associations with eugenics and public controversy more generally.
-In response to Jensen, critics such as Lewontin offered broad critiques of behavior genetics as such, and behavior geneticists in turn adopted a radical conception of academic freedom (without any sense of responsibility for the social implications of academic research) and a siege or wartime mentality, as demonstrated by Sandra Scarr's 1986 presidential address to the Behavior Genetics Association (BGA) \cite{ScarrThreeCheersBehavior1987}.
-Scarr's address coincided with the period between 1970-1990 when BG published articles in topic 24 (including Scarr's address itself).
--->
-
-<!-- However, Glayde Whitney's 1995 BGA presidential address was published in MQ rather than BG.
-According to Whitney, BG's editor refused to publish the address and the BGA Executive Committee (except for Whitney himself) voted to "issue an official statement of denouncement" of the address; though we were unable to find any such a statement, and Panofsky reports that "nothing official was done to Whitney."  In his address, Whitney criticized "the Marxist-Lysenkoist denial of genetics"; quoted Peter Brimelow — a white nationalist and later founder of the racist and xenophobic website VDARE — defining "racist" as "anyone who is winning an argument with a liberal"; and proposed as a "reasonable scientific hypothesis" that differences in murder rates between countries and cities was caused by racial genetic differences in intelligence, empathy, aggression, and impulsivity.
--->
-
-<!-- The topic model analysis and the case of Whitney's address suggest that, by the 1990s, behavior genetics may have somewhat distanced itself from race science, though perhaps without directly repudiating it \cite{GillbornSoftlySoftlyGenetics2016, HennWhyDNANo2021}.
--->
-
-<!-- This interpretation was given additional support by a supplemental review of BG using Springer's journal search website; see supplemental text \@ref(bg).
-This review also brought to our attention two disciplinary histories of behavior genetics, one published in McGue in 2010 \cite{McGueEndBehavioralGenetics2010} and the other by Plomin in 2023 \cite{PlominCelebratingCenturyResearch2023}.
-McGue acknowledges the historical association with eugenics but does not mention Jensen and other race science controversies since World War II.
-Instead postwar behavior genetics is portrayed as an endeavor to counterbalance the hegemony of the "Blank Slate," the supposedly widespread view that genetic differences play no role in explaining individual differences \cite{BatesonCorpseWearisomeDebate2002}.
-Plomin briefly mentions the controversies surrounding Jensen's 1969 paper and later *The Bell Curve*, writing that Jensen's work was controversial because he "touched on ethnic differences"; though ultimately Plomin reassures the reader that "by the 1990s, quantitative genetic research had convinced most scientists of the importance of genetics for behavioral traits, including intelligence."  All together, behavior geneticists acknowledge but then downplay the extent of the field's involvement in race science in the past, portraying themselves as models of scientific objectivity and technical rigor, while displacing the source of race science controversies on to external, "politically-motivated," and "irrational" actors.
--->
 

### 2023-08-30 paper/paper.Rmd
dual bib files working correctly

@@
 DH's work on this project was not supported by any specific funding.
-bibliography: "references.bib"
+# bibliography: "references.bib"
 output: 
@@
 \clearpage
+
+\bibliography{references.bib,refs-manual.bib}

### 2023-08-30 paper/paper.Rmd
ready to read

@@
 }.
-Pioneer was formed in 1937, during the waning years of the eugenics movement in North America \cite{BarkanRetreatScientificRacism1992}, with the aim of "support[ing] academic research and the \'[sic]dissemination of information, into the 'problem of heredity and eugenics' and 'the problems of race betterment'" [@MehlerFoundationFascismNew1989 21, quoting Laughlin].
+Pioneer was formed in 1937, during the waning years of the eugenics movement in North America \cite{BarkanRetreatScientificRacism1992}, with the aim of "support[ing] academic research and the \'[sic]dissemination of information, into the 'problem of heredity and eugenics' and 'the problems of race betterment'" \cite[p.21, quoting Laughlin]{MehlerFoundationFascismNew1989}.
 MQ was founded in 1960 by biologist R.

### 2023-08-30 paper/paper.Rmd
reread and updated TODO

@@
 *Scientific racism* refers to the social practice of purporting to justify racial inequality and colonialism by appealing to the epistemic authority of science.
-*Race science* refers to scientific and/or pseudoscientific research, along with its research products (journal articles, etc.), that can be utilized for scientific racism.
-*Race science discourse* characterizes a broader category, that treats race science as a legitimate area of scientific research.
-Race science discourse is a strict superset of race science, because the former includes methodological and empirical critiques of race science.
+*Race science* refers to scientific and/or pseudoscientific research, including research products (journal articles, etc.), that can be utilized for scientific racism.
+*Race science discourse* refers to a broader category, any treatment of race science as a legitimate area of scientific research.
+This includes methodological and empirical critiques of race science.
 For example, the Flynn effect (the finding that IQ scores have increased over time) poses a serious empirical challenge to hereditarian claims of racial differences in intelligence.
-But technical debates over the extent and causes of the Flynn effect likely create the impression among the general public of legitimate scientific disagreement, and in this way legitimize race science.
+But technical debates over the extent and causes of the Flynn effect likely create the impression among the general public of ordinary scientific disagreement, and in this way legitimize race science.
+So empirical research supporting the Flynn effect is usually not race science, but can still be race science discourse.
 
@@
 For the purposes of the current study, it is not important whether, for instance, Arthur Jensen is or is not correctly labelled a racist.
-It is important that much of Jensen's highest-profile research was useful and indeed utilized to purportedly justify racial inequality.
+It is important that much of Jensen's research was useful and indeed utilized to purportedly justify racial inequality.
 
@@
 We identified 16 researchers who had received funding from Pioneer; 14 of these researchers had profiles in the Web of Science (WoS) author search, allowing us identify 13 WoS-indexed journals that had published 6 or more of these authors.
-See Table \@ref(tab:researchers).
+See Table \@ref(tab:researchers) and Figure \@ref(fig:wos).
 
@@
 
-Topic 22 is strongly associated with I and PID in the same way.
+Topic 22 is strongly associated with I and PID in the same way that 07 is associated with MQ.
 The Silge plot does contain `jensen`, as well as a reference to Raymond Cattell, who played a major role in the development of factor analysis and intelligence testing but also advocated for eugenics, fascism, and Nazi race science \cite{MehlerBeyondismRaymondCattell1997}.
@@
 The Silge plot for topic 24, by contrast, suggests a distinct race science discourse topic, with multiple racial terms (`whites`, `blacks`, `racial_differences`, `races`, `race_differences`) and the names of three prominent Pioneer-funded researchers, `jensen`, `lynn`, and `rushton`.
-Independent qualitative coding of the top 121 papers in this topic (those with $gamma > 0.97$) confirmed this interpretation of the topic, with 108 (89%) coded as race science discourse by both authors (Cohen's $\kappa = 0.86$).
+Independent qualitative coding of the top 121 papers in this topic (those with $\gamma > 0.97$) confirmed this interpretation of the topic, with 108 (89%) coded as race science discourse by both authors (Cohen's $\kappa = 0.86$).
 
@@
 Jensen's "How Much Can We Boost IQ and Scholastic Achievement?" \cite{JensenHowMuchCan1969} was published in 1969 in Harvard Educational Review (not included in this study), and the time series indicates that, around this time, there was an increase in articles in topic 24 in both PS and MQ (the only two journals in our corpus that were active at the time).
-MQ shows another sharp increase in the late 1970s, around the time Lynn was brought on as assistant editor, consistent with the conclusion of a qualitative analysis by Adams and Pilloud that "Psychology became the dominant discipline upon the editorship of Richard Lynn and Gerhard Meisenberg in MQ, where some issues were almost entirely represented by IQ test–based studies" \cite[110]{AdamsMisAppropriationBiological2021}.
+MQ shows another sharp increase in the late 1970s, around the time Lynn was brought on as assistant editor.
+This finding is consistent with the conclusion of a qualitative analysis that "Psychology became the dominant discipline upon the editorship of Richard Lynn and Gerhard Meisenberg in MQ, where some issues were almost entirely represented by IQ test–based studies" \cite[p.
+110]{AdamsMisAppropriationBiological2021}.
 PID published multiple papers in topic 24 almost immediately after it was founded, with I showing a more gradual increase between the mid-1970s and mid-1990s.
@@
 Scarr's address coincided with the period between 1970-1990 when BG published articles in topic 24 (including Scarr's address itself).
-However,
 
-The topic model analysis suggests that by the 1990s, behavior genetics may have distanced itself somewhat from race science, albeit without directly repudiating it.
+However, the topic model analysis suggests that by the 1990s, behavior genetics may have distanced itself somewhat from race science, albeit without directly repudiating it.
 This interpretation is further supported by two disciplinary histories of behavior genetics, one published in McGue in 2010 \cite{McGueEndBehavioralGenetics2010} and the other by Plomin in 2023 \cite{PlominCelebratingCenturyResearch2023}.
-In addition, the 1995 BGA presidential address by Glayde Whitney - in which he criticized the "Marixist-Lysenkoist denial of genetics" and proposed that differences in murder rates between countries and cities was caused by racial genetic differences in intelligence, empathy, aggression, and impulsivity - was published in MQ rather than BG.
-As a discipline, behavior genetics may have been prominent in forming an echo chamber for race science in the past, but has been less receptive to such ideas in more recent history.
+In addition, the 1995 BGA presidential address by Glayde Whitney --- in which he criticized the "Marixist-Lysenkoist denial of genetics" and proposed that differences in murder rates between countries and cities was caused by racial genetic differences in intelligence, empathy, aggression, and impulsivity --- was published in MQ rather than BG.
+As a discipline, behavior genetics may have been prominent in forming an echo chamber for race science in the past, but appears to have been less receptive to such ideas in more recent history.
 
-Instead, our results suggest that psychology has been the dominant discipline in which the race science echo chamber is being maintained.
+Instead, our results suggest that psychology has been the dominant discipline in which the race science echo chamber has been maintained.
 This may be the result of racial essentialization being maintained more easily in a discipline like psychology than in other disciplines that contribute to behavior genetics, like genetics and anthropology.
-The typical practitioners in psychology do not have the degree of formal education about, and therefore knowledge of, relevant underlying genetic and anthropological research that has led many professional organizations in genetics and anthropology to formally reject the claim of race as meaningful biological concept \cite{anthropology1996aapa, american2018ashg}, undermining the core assumption necessary for race science to exist.
-As described by Winston, however, "[s]ince the 1990s, scientific racism has evolved in ways that sidestep or blunt the charge of essentialist categorization" (p.
-427, \cite{WinstonWhyMainstreamResearch2020}), such as shifting the rhetoric away from race to concepts such as "national IQ" that are deployed to preserve an essentialist categorization of different human populations.
-Professional psychological organizations routinely publish statements rejecting racism or acknowledging psychology's complicity in systemic racism \cite{american2021apology, aps2021statement}, but stop short of rejecting race as a meaningful biological categorization of people, allowing for race science research to be smuggled into psychology and given the appearance of continued scientific legitimacy.
+*[cite?]* The typical practitioners in psychology do not have the degree of formal education about, and therefore knowledge of, relevant underlying genetic and anthropological research that has led many professional organizations in genetics and anthropology to formally reject the claim of race as meaningful biological concept \cite{anthropology1996aapa, american2018ashg}, undermining the core assumption of race science.
+As described by Winston, however, "[s]ince the 1990s, scientific racism has evolved in ways that sidestep or blunt the charge of essentialist categorization" \cite[p.
+427]{WinstonWhyMainstreamResearch2020}, such as shifting the rhetoric away from race to concepts such as "national IQ" that are deployed to preserve an essentialist categorization of different human populations.
+Professional psychological organizations routinely publish statements rejecting racism or acknowledging psychology's complicity in systemic racism \cite{american2021apology, aps2021statement}, but stop short of rejecting race as a meaningful biological categorization of people, allowing for race science research to be smuggled into psychology and giving the appearance of continued scientific legitimacy.
 
@@
 
-We stress that we are not identifying any particular scientist as a racist, scientific racist, or race scientist based on the topic model results.
+We stress that we are not identifying any particular scientist as a racist, scientific racist, or race scientist based on the topic model results alone.
 Scientific research can be utilized to promote scientific racism — and thus count as race science — independently of the views and attitudes of the scientist who conducted that research \cite{TaberyWhyStudyingGenetics2015, GillbornSoftlySoftlyGenetics2016, CarlsonQuantifyingContextualizingImpact2020, HennWhyDNANo2021}.
 But this does not mean that scientists are not responsible for the downstream social effects of their
-research, even when these effects involve deliberate misinterpretation \cite{DouglasSciencePolicyValuefree2009, KitcherArgumentFreeInquiry1997, KouranyShouldKnowledgeBe2016, KouranyShouldKnowledgeBe2016, BlockIQHeritabilityInequality1974a}.
+research, even when these effects involve deliberate misinterpretation \cite{DouglasSciencePolicyValuefree2009, KitcherArgumentFreeInquiry1997, KouranyShouldKnowledgeBe2016, BlockIQHeritabilityInequality1974a}.
 Like other citizens, scientists are responsible for mitigating the reasonably foreseen harms that result from their actions.
 And we now have decades of empirical evidence that research on the genetics of intelligence — even research that does not itself touch on race — can and will be used to promote scientific racism and indeed racial violence \cite{MeyerWrestlingSocialBehavioral2023}.
-Specifically, while our findings suggest that behavior genetics has attempted to distance itself from scientific racism, it has done little to explicitly repudiate this research, especially in contrast to statements by biological anthropologists \cite{FuentesAAPAStatementRace2019} and human geneticists \cite{JacksonFacingOurHistory2023}.
+For instance, while our findings suggest that behavior genetics has attempted to distance itself from scientific racism, it has done little to explicitly repudiate this research, especially in contrast to statements by biological anthropologists \cite{FuentesAAPAStatementRace2019} and human geneticists \cite{JacksonFacingOurHistory2023}.
 Behavior genetics research — and research on human genetics more generally — may be especially susceptible to appropriation by scientific racists, which only strengthens the obligations these researchers have to prevent their research from causing harm \cite{CarlsonQuantifyingContextualizingImpact2020, OgbunugaforDNABasketballBirthday2022}.
@@
 Our distinction between race science and race science discourse highlights an important "engagement dilemma" for scientists challenging race science and scientific racism.
-While the "tobacco strategy" for manufacturing doubt involves targeting mainstream scientists and promoting fringe scientific views, its primary audiences are lawmakers and the general public.
+While the "tobacco strategy" for manufacturing doubt targets mainstream scientists and promotes fringe scientific views, its primary audiences are lawmakers and the general public.
 The aim of the strategy is not to change the opinion of mainstream scientists, but instead to cultivate the impression that there is legitimate scientific uncertainty about the issue, reasonable disagreement among qualified experts *[cites]*.
@@
 
-*[what can science do? Posselt, /Equity in Science/]*
-
-*[max ~50 refs]*
-
 \clearpage
@@
 This journal has published a relatively large number of non-article documents, such as book reviews and commentaries, that are not available at the URL included in the Crossref metadata.
-(This is unfortunate, as it was not difficult to find highly relevant documents that we could not automatically retrieve and therefore excluded from our corpus, such as an exchange between Rushton and Flynn on the Flynn effect \cite{FlynnEvidenceRushtonGenetic1998,RushtonSecularGainsIQ1998,FlynnReplyRushtonGang1998}.)  Instead we used Elsevier's ScienceDirect API (<https://dev.elsevier.com/>) to independently search and retrieve all available items from *Personality and Individual Differences*.
+(This is unfortunate, as it was not difficult to find highly relevant documents that we could not automatically retrieve and therefore had to be excluded from our corpus.
+One set of examples is an exchange between Rushton and Flynn on the Flynn effect \cite{FlynnEvidenceRushtonGenetic1998,RushtonSecularGainsIQ1998,FlynnReplyRushtonGang1998}.)  Instead we used Elsevier's ScienceDirect API (<https://dev.elsevier.com/>) to independently search and retrieve all available items from *Personality and Individual Differences*.
 This stage of corpus assembly was conducted between 2021-11 and 2022-05.
@@
 But text is only available in aggregate PDFs that bundle together the target article and all of the commentaries.
-In a few cases this results in PDFs that are hundreds of pages long and might be linked from the Crossref metadata 30+ times.
-In addition, the retrieved PDFs are not perfectly identical, because Cambridge UP's servers add a timestamped watermark to each page.
+In somes cases this results in PDFs that are hundreds of pages long and might be linked from the Crossref metadata 30+ times.
+In addition, the retrieved PDFs are not perfectly identical, because Cambridge UP's servers add a timestamped watermark to each page when the PDF is requested.
 We attempted to contact Cambridge for assistance but did not receive a response.
 We ultimately used a series of ad hoc measures to mitigate text duplication.
-Approximately 100,000 documents of 1-2 pages long were excluded before PDF retrieval.
+Approximately 100,000 documents that were 1-2 pages long were excluded before PDF retrieval.
 After PDF retrieval and text extraction, the timestamped watermarks were removed using a regular expression and the text was hashed using SHA 256.
 Hashes were used to construct groups of duplicate documents, and a single document (whichever one happened to be first in the dataframe) was chosen from each hash group for inclusion in the corpus.
-It is plausible that some duplicates made it through this process: where the watermark overlapped with the text, the regular expression may have been unable to identify and remove the watermark (a false negative result), and this difference in the watermarks (not the text) would produce different hashes.
+It is plausible that some duplicates made it through this process: where the watermark overlapped with the text, the regular expression likely would have been unable to identify and remove the watermark (a false negative result), and this small difference in the watermarks (not the text) would produce different hashes.
 
@@
 Figure \@ref(fig:counts) and table \@ref(tab:counts) show the number of fulltext documents included in the corpus.
+All together, the corpus includes 34,896 documents from 6 journals.
 
@@
 
-All together, the corpus includes 34,896 documents from 6 journals.  See figure \@ref(fig:counts) and table \@ref(tab:counts).
-
 ## Text preparation
@@
 
-NLP-extracted noun phrases offer a number of advantages over the more traditional unigram ("single-word") tokens.
-First, noun phrase extraction removes many standard stopwords (articles, common verbs) without relying on a fixed, a priori stopword list.
+NLP-extracted noun phrases offer a number of advantages over the more traditional unigram ("single word") terms.
+First, noun phrase extraction removes many standard stopwords (articles, common verbs, numbers) without relying on a fixed, a priori stopword list.
 Phrases can be more informative than single terms, for example, distinguishing "intelligence test" from "hypothesis test."  While simple n-gram extraction will include numerous phrases that are not especially meaningful.
-Consider the sentence "Since our first analyses of feeding patterns in rats, we had been using a criterion of 40 minutes (Le Magnen & Tallon 1963; 1966)."\cite{SclafaniCorrelationCausationStudy1981}  Bigrams such as "since our" and "criterion 40" will likely be discarded in vocabulary selection, but significantly increase the computational cost of vocabulary selection.
+Consider the sentence "Since our first analyses of feeding patterns in rats, we had been using a criterion of 40 minutes (Le Magnen & Tallon 1963; 1966)"\cite{SclafaniCorrelationCausationStudy1981}.
+Bigrams such as "since our" and "criterion 40" will likely be discarded in vocabulary selection, but significantly increase the computational cost of vocabulary selection.
 Noun phrase extraction is therefore more efficient.
@@
 Intuitively, highly informative terms (in this project, noun phrase types) are distinctive, allowing you to dramatically narrow down the list of potential documents.
-This "informativeness" of a term can be quantified as the KL (Kullback-Leibler) divergence from a "baseline" distribution of documents to the distribution conditional on the term.
+This "informativeness" of a term can be quantified as the KL (Kullback-Leibler) divergence from a "baseline" uniform distribution across documents to the distribution conditional on the term, which we abbreviate as $\Delta H$.
 Because the most informative terms tend to be typos and OCR errors — these are unique to a single document — we multiply the KL divergence by the logarithm of the total number of occurrences of the term across the entire corpus.
-We refer to the resulting measure as $log(n) \Delta H$ or simply `ndH`.
+We refer to the resulting measure as $log(n) \Delta H$ or `ndH`.
 
@@
 Many of the documents published in *Psychological Reports* are extremely short, 1-2 page brief notices of a single study; while many of the documents published in *Brain and Behavioral Sciences* are book-length collections that include a long review article and sometimes dozens of commentaries.
-Very generic noun phrases that happen to appear in the latter can occur orders of magnitude more often than highly distinctive phrases in the former, and so the $log(n)$ factor overwhelms the $\Delta H$ factor.
+Very generic noun phrases that happen to appear in BBS can occur orders of magnitude more often than highly distinctive phrases in the PS, and so the $log(n)$ factor overwhelms the $\Delta H$ factor.
 
@@
 However, we had some concerns with computational demands here:  the resulting document-term matrix would have roughly $10 \times 33,000^2$ or 10.9 billion entries; with 10% density this would require on the order of 4 GB of memory just for a single copy of the matrix; and most of the analysis was to be conducted on the authors' laptops.
-We therefore chose to work with three smaller vocabularies, $5 \times$, $1 \times$ and $\frac{1/5} \times$ the number of documents.
+We therefore chose to work with three smaller vocabularies, $5 \times$, $1 \times$ and $\frac{1}{5} \times$ the number of documents.
 We refer to these as the "large" (174,480 distinct phrases), "medium" (34,896) and "small" (6,979) vocabulary, respectively.
@@
 This novel approach was implemented in the R package `tmfast` \cite{HicksTmfastFitsTopic2023}.
-A simulation study of `tmfast` found that it was significantly faster and only slightly less accurate at reconstructing known word-topic and topic-document distributions, compared to `stm`.
+A simulation study of `tmfast` found that it was significantly faster and only slightly less accurate at reconstructing known word-topic and topic-document distributions, compared to `stm` \cite{HicksTmfastFitsTopic2023}.
 
@@
 
-Following the approach of Hicks \cite{HicksProductivityInterdisciplinaryImpacts2021}, we did not attempt to identify a unique best fitted model for further analysis.
-While we focus on the medium vocabulary, $k=40$ model as the "median" among the 24 models fit, we compare and contrast findings from this model with those from the other models.
+Following the approach of \cite{HicksProductivityInterdisciplinaryImpacts2021}, we did not attempt to identify a unique best fitted model for further analysis.
+While the manuscript text focuses on the medium vocabulary, $k=40$ model as the "median" among the 24 fitted models, we compare and contrast findings from this model with those from the other models.
 
@@
 
-After fitting topic models, our first research question was whether we could identify distinctive "race science" topics.
+After fitting topic models, our first research question was whether we could identify distinctive "race science discourse" topics.
 Figures \@ref(fig:gamma-lg), \@ref(fig:gamma-md), \@ref(fig:gamma-sm) show the gamma (topic-document) distributions for each value of $k$ for the three vocabularies.
-In these figures, each panel corresponds to a single journal-$k$ combination, each row of cells is a single document in that journal, and each column of singles is a single topic in that particular model.
+In these figures, each panel corresponds to a single journal-$k$ combination, each row of cells is a single document in that journal, and each column of cells is a single topic in that particular model.
 Topics correspond within columns of panels, but not within rows; for example, topic 04 for $k=10$ does not necessarily correspond to topic 04 for $k=20$.
@@
 A spreadsheet of all articles that had a maximum value of $\gamma$ for this topic was extracted (excluding articles published in *Psychological Reports*), and the top 121 articles ($\gamma > 0.97$) were reviewed manually by both authors.
-We coded each article as *race science discourse*, or not, using our definition of race science discourse as treated race science as a legitimate area of scientific reearch; this includes methodological critiques of race science and empirical tests that falsify race science hypotheses.
+We coded each article as *race science discourse*, or not, using our definition of race science discourse as treating race science as a legitimate area of scientific reearch; this includes methodological critiques of race science and empirical tests that falsify race science hypotheses.
 
@@
 We calculated the "false positive rate" — documents with maximum $\gamma$ in this topic that were not race science discourse — and interrater reliability using Cohen's $\kappa$.
-The consensus true positive rate was 89.3$ (108/121), while the consensus false positive rate was 8.3% (10/121), with $98%$ agreement across the two raters ($\kappa = 0.86)$.
+The consensus true positive rate was 89.3% (108/121), while the consensus false positive rate was 8.3% (10/121), with 98% agreement across the two raters ($\kappa = 0.86)$.
 Coding spreadsheets are included in the supplemental materials *[filenames]*.
-For each of the 10 consensus false documents, the inclusion was readily explained, but in different ways for different documents: some included text from other documents, as when an article started in the middle of a page; others discussed perceptions of race relations or racial animus; one piece in BG discussed differences in the herding behavior of two cattle breeds, referred to as "white" and "black."  The 3 non-consensus documents reflected essential ambiguity in the operationalization of race science discourse.
+For each of the 10 consensus false positive documents, their inclusion in the topic was readily explained, but in different ways for different documents.
+Some included text from other documents, as when an article started in the middle of a page; others discussed perceptions of race relations or racial animus; one piece in BG discussed differences in the herding behavior of two cattle breeds, referred to as "white" and "black."
+
+The 3 non-consensus documents reflected essential ambiguity in the operationalization of race science discourse.
 One is a table of contents from a 1980 issue of I; arguably this document should have been excluded as irrelevant front matter, but arguably it legitimizes the research published in the issue.
-The second case is a paper on racial differences in emotional intelligence was classified as race science discourse by one author, while the other author felt it was more focused on psychometrics \cite{GignacGroupDifferencesEI2010}.
+The second case is a paper on racial differences in emotional intelligence that was classified as race science discourse by one author, while the other author felt it was more focused on psychometrics \cite{GignacGroupDifferencesEI2010}.
 A finally, a 1990 article in MQ (<https://www.unz.com/print/MankindQuarterly-1990q3-00108/>) proposed to offer an economic explanation for crime.
@@
 
-28 documents in this sample were published between 1990 and 2010.
+28 documents in this sample from BG were published between 1990 and 2010.
 One was an obituary of Jerry Hirsch, a critic of hereditarianism in general and Jensen in particular \cite{RoubertouxJerryHirsch202008}, and another was an obituary of David Rowe \cite{RodgersObituaryDavidChristian2003}.
-Only 2 of these 28 documents reported race differences of any kind: one examining at interactions among race, sex, and heritability for adolescent BMI (Body Mass Index, used as a measure of overweight/obesity) \cite{JacobsonGeneticSharedEnvironmental1998}; and the other at interactions among race, family history of alcoholism, and visuospatial performance \cite{BermanReducedVisuospatialPerformance1995}.
+Only 2 of these 28 documents reported race differences of any kind: one examining interactions among race, sex, and heritability for adolescent BMI (Body Mass Index, used as a measure of overweight/obesity) \cite{JacobsonGeneticSharedEnvironmental1998}; and the other interactions among race, family history of alcoholism, and visuospatial performance \cite{BermanReducedVisuospatialPerformance1995}.
 Finally, Philip Vernon — one of the Pioneer-funded scientists we identified earlier — published a critical review of a book that attempted to address hereditarianism, the Flynn effect, race-and-intelligence research, and some related issues.
@@
 At the same time, we were unable to find any articles published in BG that acknowledged the involvement of the field with scientific racism with anything like the force of the statements that have been made by biological anthropologists \cite{FuentesAAPAStatementRace2019} and human geneticists \cite{JacksonFacingOurHistory2023}.
-Notably, the ASHG report includes discussions of behavior genetics race science, and is critical of the field and the organization for failing to publicly reject the race-and-intelligence claims made by Shockley and Jensen in particular.
-A search for "eugenics" found articles that presented the issue as part of the distant past, and a search for "racism" turned up studies of racist attitudes.
-A statement on the Behavior Genetics Association website, dated 2021, acknowledges that "The history of our field is inextricably linked with racism, including the misuse of behavior genetic research to support violent eugenic policies," while primarily focusing on BGA member demographics (<https://www.bga.org/content.aspx?page_id=22&club_id=971921&module_id=567723>).
+Notably, the ASHG report includes discussions of behavior genetics race science, and is critical of the field and the organization for failing to publicly reject the race-and-intelligence claims made by Shockley and Jensen in particular \cite{JacksonFacingOurHistory2023}.
+A search for "eugenics" in BG found articles that presented the issue as part of the distant past, and a search for "racism" turned up studies of racist attitudes.
+A statement on the Behavior Genetics Association website, dated 2021, acknowledges that "The history of our field is inextricably linked with racism, including the misuse of behavior genetic research to support violent eugenic policies," but primarily focuses on BGA member demographics (<https://www.bga.org/content.aspx?page_id=22&club_id=971921&module_id=567723>).
 
@@
 Techniques such as topic modeling may be able to account for discursive shifts from explicit to coded racial language, but only if the corpus contains a critical mass of documents using both sets of terms together.
-This would seem to require either a relatively gradual transition, the inclusion of reflective commentary, such as scholarship identifying and analyzing the transition, or both.
+This would seem to require either a relatively gradual transition, the inclusion of reflective commentary --- such as scholarship identifying and analyzing the transition --- or both.
 It is unlikely that these requirements are met by the corpus used in this study.
@@
 
-The phenomenon of colorblind racism is closed related to ambiguities that cannot be resolved even by human coders familiar with the broader cultural context.
+The phenomenon of colorblind racism is closely related to ambiguities that cannot be resolved even by human coders familiar with the broader cultural context.
 Two relevant examples are the "Flynn effect," a secular increase in intelligence test scores noted by philosopher James Flynn; and studies of "national IQ," exemplified by the work of Richard Lynn.

### 2023-08-30 paper/paper.Rmd
update TODO and readme

@@
 acknowledgements: |
-  *[strutured]* Thanks to Emily Merchant and John Jackson for some initial discussions that helped clarify the scope of the project and identify some key resources related to the Pioneer Fund and *Mankind Quarterly*.
+  | Thanks to Emily Merchant and John Jackson for some initial discussions that helped clarify the scope of the project and identify some key resources related to the Pioneer Fund and *Mankind Quarterly*.
 Thanks to Anthony Sainez for help retrieving the *Mankind Quarterly* articles.
 Thanks to Derek Devnich and James Dooley for their work attempting to secure the APA-published articles.
-EL's work on this project was supported by UC Merced.
-DH's work on this project was not supported by any specific funding.
+  | **Funding:** EL's work on this project was partially supported by UC Merced.  DH's work on this project was not supported by any specific funding. 
+  | **Author contributions:** DJH (conceptualization, data curation, formal analysis, investigation, methodology, project administration, software, validation, visualization, writing-original draft, writing-review & editing); EL (conceptualization, methodology, software, validation, visualization, writing-original draft, writing-review & editing)
+  | **Competing interests:** The authors have no competing interests to declare. 
+  | **Data and materials availability:** Web of Science search results and article fulltext are not publicly available due to intellectual property restrictions.
+Preprocessed analysis data (document-term counts and article metadata) and analysis scripts are available at *[doi]*.
 # bibliography: "references.bib"
@@
 -->
+
+FUNDING: EL's work on this project was partially funded by UC Merced.  No other specific funding was used for this project.  AUTHOR CONTRIBUTIONS:
 
@@
 -->
+
+\clearpage
+
+# References and Notes {.unnumbered}
+
+\bibliography{references.bib,refs-manual.bib}
 
@@
 Topic models, of course, are not even capable of representing this essential ambiguity.
-
-\clearpage
-
-\bibliography{references.bib,refs-manual.bib}

### 2023-09-01 paper/paper.Rmd
abstract and intro

@@
 abstract: |
-  *[structured]*
+  | **Background:** While anthropologists and human geneticists have long rejected the biological race concept, both social movements and prominent researchers have purported to justify racial inequality and colonialism by appealing to the epistemic authority of science.
+  | **Methods:** Bibliometric and text mining methods were used to trace the appearance of race science discourse in both mainstream scientific journals and the parascholarly journal *Mankind Quarterly* from 1960 to 2010.
+  | **Results:** Our analysis found two distinguishable lines of race science discourse.
+One is strongly associated with *Mankind Quarterly*, while the other — emphasizing race and intelligence — is strongly associated with mainstream psychology journals, especially *Intelligence* and *Personality and Individual Differences*.
+  | **Conclusion:** Psychology, and especially intelligence research, has played a significant role in mainstreaming race science.
+
 acknowledgements: |
@@
 - One Sentence Summary: 125 characters
-- Abstract: Structured as BACKGROUND (1 sentence), OBJECTIVES/METHODS, RESULTS, CONCLUSION (1 sentence); 125 words or less
 - Main Text:  The introduction should provide sufficient background information to make the article intelligible to readers in other disciplines and sufficient context so that the significance of the experimental findings is clear.
@@
 - References and Notes: There should be a single reference list that combines references cited in the main paper as well as those cited in the supplementary materials, and each reference should appear only once in the list.
-- Acknowledgements: Funding; Authors contributions (see link); Competing interests;  Data and materials availability
 List of Supplementary Materials
@@
 
-FUNDING: EL's work on this project was partially funded by UC Merced.  No other specific funding was used for this project.  AUTHOR CONTRIBUTIONS:
+| 
+| **One sentence summary:** Bibliometric and text mining analysis indicate that race science discourse appeared in mainstream psychology journals throughout the second half of the twentieth century.
 
@@
 
-*[hey you should read this paper]*
+By the end of the Second World War, a combination of technical developments in anthropology and genetics, along with revelations of Nazi atrocities, had led most members of the scientific community to reject attempts to scientifically justify racial inequalities and colonialism \cite{BarkanRetreatScientificRacism1992}.
+But eugenic thinking and scientific racism persisted throughout the second half of the twentieth century, and recent work has found that white supremacists continue to attempt to use scientific research to justify racial violence \cite{DusterBackdoorEugenics2003, PanofskyHowWhiteNationalists2021}.
 
-We propose a distinction between scientific racism, race science, and race science discourse.
+The aim of the study reported here was to trace race science in the scholarly and parascholarly literature, examining the ways that fringe ideas can present themselves as legitimate scholarly inquiry.
+
+We first propose a distinction between scientific racism, race science, and race science discourse.
 *Scientific racism* refers to the social practice of purporting to justify racial inequality and colonialism by appealing to the epistemic authority of science.
@@
 For example, the Flynn effect (the finding that IQ scores have increased over time) poses a serious empirical challenge to hereditarian claims of racial differences in intelligence.
-But technical debates over the extent and causes of the Flynn effect likely create the impression among the general public of ordinary scientific disagreement, and in this way legitimize race science.
-So empirical research supporting the Flynn effect is usually not race science, but can still be race science discourse.
+But technical debates over the extent and causes of the Flynn effect likely create the impression among the general public of ordinary scientific disagreement.
+In this way technical critiques of race science based on the Flynn effect could potentially legitimize race science.
+So empirical research supporting the Flynn effect is usually not itself race science, but can still be race science discourse.
 
-The distinction between scientific racism and race science allows us to bracket the intentions and mindset of the researchers involved in any particular piece of research.
-For the purposes of the current study, it is not important whether, for instance, Arthur Jensen is or is not correctly labelled a racist.
-It is important that much of Jensen's research was useful and indeed utilized to purportedly justify racial inequality.
+These distinction allow us to bracket the intentions and mindset of the researchers involved in any particular piece of research or debate.
+For the purposes of the current study, it is not important whether, for instance, Arthur Jensen is or is not correctly labelled a scientific racist \cite{JacksonJr.ArthurJensenEvolutionary2022}.
+But it is important that much of Jensen's research was useful and indeed utilized to purportedly justify racial inequality.
 
@@
 Robert Gayre (self-styled as "Gayre of Gayre and Nigg"; 1907-1996).
-Gates' professional status had risen and fallen with eugenics and the explicit scientific racism of the 1920s, and by the end of World War II he was thoroughly marginalized \cite{WinstonScienceServiceFar1998}.
+Gates' professional status had risen and fallen with eugenics and the explicit scientific racism of the 1920s, and by the end of the Second World War he was thoroughly marginalized \cite{WinstonScienceServiceFar1998}.
 In contrast, Garrett had been president of the American Psychological Association in 1946 and chair of Psychology at Columbia from 1941 to 1955 \cite{WinstonScienceServiceFar1998}.
@@
 In the landmark case *Brown v Board of Education of Topeka* (1954), the US Supreme Court banned *de jure* educational segregation.
-The Court's decision relied on expert testimony from psychologists and education researchers; but the segregationists had also put forward their own experts, including Henry Garrett \cite{WinstonScienceServiceFar1998, JacksonScienceSegregationRace2005, SchafferScientificRacismAgain2007}.
+The Court's decision relied on expert testimony from psychologists and education researchers; but the segregationists also put forward their own experts, including Henry Garrett \cite{WinstonScienceServiceFar1998, JacksonScienceSegregationRace2005, SchafferScientificRacismAgain2007}.
 
-As part of the segregationist reaction to *Brown*, MQ was created as what contemporary historians and philosophers of science call an "echo chamber" \cite{FernandezPintoKnowBetterNot2017}, a favorable venue for race scientists to publish their views, on the grounds that an "equalitarian dogma" created a censorious "taboo" against their research in mainstream publications \cite{TuckerFundingScientificRacism2002, JacksonMythicalTabooRace2020}.
+As part of the segregationist reaction to *Brown*, MQ was created as what contemporary historians, philosophers, and sociologists of science call an "echo chamber" \cite{FernandezPintoKnowBetterNot2017}, a favorable venue for race scientists to publish their views, on the grounds that an "equalitarian dogma" created a censorious "taboo" against their research in mainstream publications \cite{TuckerFundingScientificRacism2002, JacksonMythicalTabooRace2020}.
 Echo chambers are a key part of the "tobacco strategy," used by numerous regulated industries — most infamously tobacco, but also fossil fuels and chemical manufacturing, among others — to "manufacture doubt" and delay regulation \cite{OreskesMerchantsDoubtHow2011, JasnyEmpiricalExaminationEcho2015, FernandezPintoKnowBetterNot2017, WeatherallHowBeatScience2018}.
@@
 
-Figure \@ref(fig:focal) and this discussion focuses on three topics identified in one of the 24 fitted topic models; see figures \@ref(fig:gamma-lg), \@ref(fig:gamma-md), \@ref(fig:gamma-sm) and supplemental visualizations *[filenames]* for all topics from all models.
+Figure \@ref(fig:focal) and this discussion focuses on three topics identified in one of the 24 fitted topic models; see figures \@ref(fig:gamma-lg), \@ref(fig:gamma-md), \@ref(fig:gamma-sm) and supplemental visualizations S2-S4 for all topics from all models and S5-S7 for selected topics from all models.
 In the focal model, topic 07 is strongly associated with MQ:  MQ published dozens of articles in this topic each year, and no other journal ever published more than a handful.
@@
 While the "tobacco strategy" for manufacturing doubt targets mainstream scientists and promotes fringe scientific views, its primary audiences are lawmakers and the general public.
-The aim of the strategy is not to change the opinion of mainstream scientists, but instead to cultivate the impression that there is legitimate scientific uncertainty about the issue, reasonable disagreement among qualified experts *[cites]*.
+The aim of the strategy is not to change the opinion of mainstream scientists, but instead to cultivate the impression that there is legitimate scientific uncertainty about the issue, reasonable disagreement among qualified experts \cite{FernandezPintoKnowBetterNot2017}.
 On the one hand, refusing to engage with fringe scientists — or, often, pseudoscientists \cite{BhakthavatsalamVirtueEpistemologicalApproach2021} — means that false or misleading statements go unchallenged.
@@
 Consequently these critiques are likely to be either ignored by the general public or taken as evidence of legitimate scientific debate.
-*[more?]*
 
@@
 
-\bibliography{references.bib,refs-manual.bib}
+\bibliographystyle{Science}
+\bibliography{references.bib, refs-manual.bib}
 
@@
 \setcounter{table}{0}
+
+\clearpage
+
+# Supplemental Materials {.unnumbered}
+
+| S1. Materials and methods
+| Figures S2-S7
+| Data S8, S9
+
+\clearpage
 
@@
 The public analysis repository, <https://github.com/dhicks/race-science> *[DOI]*, includes document metadata and documentwise counts of NLP-extracted noun phrases ("document-term matrices"), along with code to reproduce the analysis of the following sections.
-*[repo readme: requires R 4.1.2; rig (<https://github.com/r-lib/rig/>) for managing multiple R installations]*
 
@@
 We next constructed Silge plots \cite{SilgeTopicModeling2017}, showing the top 15 (highest-probability) terms (noun phrases) from each topic in the model.
-These are provided as supplemental PDFs *[filenames]*.
+These are provided as supplemental PDFs S2-S4.
 Perusing these term lists, we focused on topics in each model that contained racial terms.
@@
 We therefore examined the prevalence of topics that used racial terms, intelligence-research terms, or both, by journal and across all 24 models.
-Supplemental PDFs *[filenames]* provide these visualizations.
+Supplemental PDFs S5-7 provide these visualizations.
 In these visualizations, each row of facets corresponds to a value of $k$ (excluding 5 as too noisy), and each column of facets corresponds to one topic.
@@
 The consensus true positive rate was 89.3% (108/121), while the consensus false positive rate was 8.3% (10/121), with 98% agreement across the two raters ($\kappa = 0.86)$.
-Coding spreadsheets are included in the supplemental materials *[filenames]*.
+Coding spreadsheets are included in the supplemental materials S8-S9.
 For each of the 10 consensus false positive documents, their inclusion in the topic was readily explained, but in different ways for different documents.

### 2023-09-01 paper/paper.Rmd
move academic freedom notes to new project

@@
 Consequently these critiques are likely to be either ignored by the general public or taken as evidence of legitimate scientific debate.
-
-<!--
-*[replace w/: WinstonWhyMainstreamResearch2020: race science as a community project of psychology(?).
-radical academic freedom is a tool used to defend race science PigliucciWhatAreWe2013, RosemanTroublesomeReflectionRacism2014, LarsenMoreProvocativeLess2020.
-an important part of how these mainstream journals function as race science echo chambers.]*
-
-Proponents of a strong conception of academic freedom often point to chapter 2 of philosopher John Stuart Mill's *On Liberty* \cite{MillLiberty1862}, "Of the Liberty of Thought and Discussion."  However, in chapter 1 of the same book, Mill states the Harm Principle as the overall thesis of his book: "The only end for which people are entitled, individually or collectively, to interfere with the liberty of action of any of their number is self-protection.
-The only purpose for which power can be rightfully exercised over any member of a civilized community, against his will, is to prevent harm to others." Even for Mill, free speech and free inquiry are appropriately restrained when the ideas in question will harm others.
-
-More recently, philosopher Shannon Dea \cite{DeaEvolvingSocialPurpose2021} has argued that academic freedom is importantly distinct from freedom of speech.
-Whereas freedom of speech is universal, enjoyed by all residents of countries where speech is legally protected, academic freedom is more exclusive, enjoyed only by college and university faculty.
-(Though, as Dea points out, non-tenure-track faculty often do not have effective protections for academic freedom.)  Freedom of speech primarily serves to protect the interests of the individual speaker: that you will not be punished by the state for stating unpopular opinions.
-But academic freedom is "other-directed," primarily serving to benefit the common good of society as a whole by enabling the pursuit of truth.
-
-Advocates of research on genetics and intelligence — both anti-racist liberals and race scientists — claim this work will lead to social benefits.
-This reflects a recognition that the purpose of academic freedom is to promote the common good, rather than merely satisfy an individual's curiosity.
-But the promised benefits are either vague or do not actually require understanding the genetics of intelligence \cite{KaplanGaltonQuincunxProbabilistic2021, CoopLotteryLuckLegacy2022}.
-
-[EL notes on academic freedom as simply one value in science:
-
-Other ethical values that guide scientific research have been institutionalized via documents such as The Belmont Report.
-Principle values described in The Belmont Report are intended as guides and regulations of scientific researchers' behavior so as not to infringe on the rights or liberties of human study volunteers.
-This report was written in response to revelations in the 1970s of the Tuskegee Syphilis Study, wherein researchers at the CDC and HHS lied to participants - all of whom were impoversiehd black men - about the nature of the study and whether or not they had tested positive for syphilis, coerced their participation with extravagant compensation, and refused to allow men infected with syphilis to be treated at all.
-The Belmont Report outlines values guiding scientific research that focus on respect for persons, beneficence, and justice:
-Respect for persons - informed consent, participants are individuals with agency. 
-Beneficience - very much like Hippocratic "do no harm" or, if harm is unavoidable, the potential harms are outweighed by the benefits of research. 
-Justice - no population should disproportionately bear the burdens of being investigated scientifically; do not sample merely from convenient populations; when advances are made due to public funds (e.g.
-taxpayer money) the products of such advances are not restricted only to those who can afford them.
-
-How does race science research square against these other values in scientific research?
-Are participants ever honestly informed that they are (or were, in the case of a debriefing form) being evaluated to see which demographics are superior/inferior to which other ones?
-Granted, the Belmont Report is US-specific, but comparable other documents, regulations, or laws exist around the world that promote similar values.
-e.g.
-The Declaration of Helsinki states that the pursuit of knowledge cannot be used to trample on the rights of research participants.
-
-Maybe brief discussion of Mertonian Norms?
-Seems that race science research violates the norms of Universalism, Disinterestedness, and Organized Skepticism.
-Granted, most of the scientific community in practice violates all of the norms pretty frequently, so maybe this isn't a strong point to include.
-
-Recently published paper by Zivony et al in PLOS Computational Biology - "Ten simple rules for socially responsible science" - The authors call out a variety of biases in research (extreme male brain hypothesis of autism; face recognition software trained only on white faces; race differences in test scores) as failing to appropriately limit inferences based on the research's design or failing to properly ground their research in social theory or relevant historical context.
-
-Random oOther thought: academic freedom doesn't mean being free from criticism of one's methods, intentions, or theoretical starting point.
-Unless someone is calling for an academic policy that forbids race science as an academic area of inquiry, talk of academic freedom is non sequitor.
-Pointing out flaws in research and concluding the research is therefore wasted time/resources is not a threat to academic freedom.
-No one has forbidden research on geocentrism, but we've recognized for a few centuries now that it's a waste of time to pursue that line of inquiry.
-
-]
-
--->
 

### 2023-09-04 paper/paper.Rmd
updating manuscript
updated manuscript to describe color-blind racism in more detail. added bonilla-silva's book to the bib file.

@@
 A key limitation of the current study stems from the inability of text mining techniques to draw on a broader cultural context than what is represented in the corpus.
-Scholars of racism and US race relations have noted a shift in the way racist attitudes are understood and expressed, from *[something]* to "colorblind racism" *[Bonilla-Silva]*, defined as *[defn]*.
-Colorblind racism involves the use of racialized "codes" or "dogwhistles," language that is superficially non-racial but carries racial implications, such as (in the US context) "violent inner city criminals" or "welfare queen" \cite{SaulDogwhistlesPoliticalManipulation2018}.
-Alderfer notes that the racist implications of *The Bell Curve* only emerge gradually, with the opening of the book using the colorblind language of "group differences" \cite{AlderferScienceNonsciencePsychologists2003}.
+Scholars of racism and US race relations have noted a shift in the way racist attitudes are understood and expressed, from overt and direct to a subtle "color-blind" racism \cite{bonillasilva2022racism}.
+Color-blind racism represents a change in the primary mechanisms by which white privilege is maintained, using non-racialized language in furtherance of the racial status quo both institutionally and individually, and increasingly reliant on the invisibility of socio-cultural mechanisms to preserve inequalities.
+Color-blind racism involves the use of racialized "codes" or "dogwhistles," language that is superficially non-racial but carries racial implications, such as (in the US context) "violent inner city criminals" or "welfare queen" \cite{SaulDogwhistlesPoliticalManipulation2018}.
+Alderfer notes that the racist implications of *The Bell Curve* only emerge gradually, with the opening of the book using the color-blind language of "group differences" \cite{AlderferScienceNonsciencePsychologists2003}.
 Panofsky et al.
 show how the term "human biodiversity," originally developed for anti-racist purposes by biological anthropologist Jonathan Marks, has been inverted by white supremacists and is now a dogwhistle for biological racial hierarchy \cite{PanofskyHowWhiteNationalists2021}.
-These cases suggest further that colorblindness can facilitate white supremacist appropriation of non-racial research — even contrary to the researchers' own intentions — as scientific racist readers re-interpret racial-neutral language as dogwhistles \cite{GillbornSoftlySoftlyGenetics2016, WillsAreClustersRaces2017}.
+These cases suggest further that color-blindness can facilitate white supremacist appropriation of non-racial research — even contrary to the researchers' own intentions — as scientific racist readers re-interpret racial-neutral language as dogwhistles \cite{GillbornSoftlySoftlyGenetics2016, WillsAreClustersRaces2017}.
 
@@
 It is unlikely that these requirements are met by the corpus used in this study.
-So it is plausible that the topic models used here have some rate of "false negatives," documents in the corpus that engage in race science discourse but in a colorblind way that avoids the use of the explicit racial terms in the race-and-intelligence topic.
+So it is plausible that the topic models used here have some rate of "false negatives," documents in the corpus that engage in race science discourse but in a color-blind way that avoids the use of the explicit racial terms in the race-and-intelligence topic.
 
-The phenomenon of colorblind racism is closely related to ambiguities that cannot be resolved even by human coders familiar with the broader cultural context.
+The phenomenon of color-blind racism is closely related to ambiguities that cannot be resolved even by human coders familiar with the broader cultural context.
 Two relevant examples are the "Flynn effect," a secular increase in intelligence test scores noted by philosopher James Flynn; and studies of "national IQ," exemplified by the work of Richard Lynn.

### 2023-09-05 paper/paper.Rmd
adding in cites for other professional orgs rejecting race as biology
Added in additional citations to professional anthropology and genetics organizations rejecting "race" as biologically meaningful.

@@
 This may be the result of racial essentialization being maintained more easily in a discipline like psychology than in other disciplines that contribute to behavior genetics, like genetics and anthropology.
-*[cite?]* The typical practitioners in psychology do not have the degree of formal education about, and therefore knowledge of, relevant underlying genetic and anthropological research that has led many professional organizations in genetics and anthropology to formally reject the claim of race as meaningful biological concept \cite{anthropology1996aapa, american2018ashg}, undermining the core assumption of race science.
+*[cite?]* The typical practitioners in psychology do not have the degree of formal education about, and therefore knowledge of, relevant underlying genetic and anthropological research that has led many professional organizations in genetics and anthropology to formally reject the claim of race as a biologically meaningful concept \cite{anthropology1996aapa, american2018ashg, board1998aaa, fuentes2019aapa, nhgri2023, rotimi2022ashg, wynshawboris2020ashg}, undermining the core assumption of race science.
 As described by Winston, however, "[s]ince the 1990s, scientific racism has evolved in ways that sidestep or blunt the charge of essentialist categorization" \cite[p.

### 2023-09-06 paper/paper.Rmd
merging citations

--- 2023-09-04 paper/paper.Rmd
@@
 This may be the result of racial essentialization being maintained more easily in a discipline like psychology than in other disciplines that contribute to behavior genetics, like genetics and anthropology.
-*[cite?]* The typical practitioners in psychology do not have the degree of formal education about, and therefore knowledge of, relevant underlying genetic and anthropological research that has led many professional organizations in genetics and anthropology to formally reject the claim of race as meaningful biological concept \cite{anthropology1996aapa, american2018ashg}, undermining the core assumption of race science.
+The typical practitioners in psychology do not have the degree of formal education about, and therefore knowledge of, relevant underlying genetic and anthropological research that has led many professional organizations in genetics and anthropology to formally reject race as a biologically meaningful concept \cite{anthropology1996aapa, american2018ashg, board1998aaa, FuentesAAPAStatementRace2019, nhgri2023, rotimi2022ashg, wynshawboris2020ashg}, undermining the core assumption of race science.
 As described by Winston, however, "[s]ince the 1990s, scientific racism has evolved in ways that sidestep or blunt the charge of essentialist categorization" \cite[p.
@@
 At the same time, we were unable to find any articles published in BG that acknowledged the involvement of the field with scientific racism with anything like the force of the statements that have been made by biological anthropologists \cite{FuentesAAPAStatementRace2019} and human geneticists \cite{JacksonFacingOurHistory2023}.
-Notably, the ASHG report includes discussions of behavior genetics race science, and is critical of the field and the organization for failing to publicly reject the race-and-intelligence claims made by Shockley and Jensen in particular \cite{JacksonFacingOurHistory2023}.
+Notably, the ASHG report includes discussions of behavior genetics race science, and in particular is critical of the field and the organization for failing to publicly reject the race-and-intelligence claims made by Shockley and Jensen \cite{JacksonFacingOurHistory2023}.
 A search for "eugenics" in BG found articles that presented the issue as part of the distant past, and a search for "racism" turned up studies of racist attitudes.

### 2023-09-06 paper/paper.Rmd
bib file formatting

@@
     number_sections: true
+header-includes:
+  - \usepackage{lineno}
+  - \linenumbers
 ---

### 2023-09-06 paper/paper.Rmd
sentences starting with numbers

@@
 After parsing these results, we counted how many of the 14 researchers had published in each journal.
-13 journals had published 6 or more of the 14 Pioneer-funded researchers.
+Thirteen journals had published 6 or more of the 14 Pioneer-funded researchers.
 We excluded *Mankind Quarterly* (as already included) as well as *Science* and *Nature* (as too general) from further consideration in this side of the corpus.
-3 journals published by the American Psychological Association (APA; *American Psychologist*, *Contemporary Psychology*, and *Journal of Educational Psychology*) had to be excluded due to confusion over who could give us permission to use the archives for a text mining project, with both APA and ProQuest asserting that we needed to get permission from the other entity.
+Three journals published by the American Psychological Association (APA; *American Psychologist*, *Contemporary Psychology*, and *Journal of Educational Psychology*) had to be excluded due to confusion over who could give us permission to use the archives for a text mining project, with both APA and ProQuest asserting that we needed to get permission from the other entity.
 *European Journal of Personality*, published by SAGE, also had to be excluded because our institutional access only went back to 1999.
@@
 This search returned 125 documents.
-33 were presentation abstracts from meetings of the Behavior Genetics Association; we did not examine these further.
-14 of the remaining documents were published since 2010 (after the scope of the primary study).
+Thirty-three were presentation abstracts from meetings of the Behavior Genetics Association; we did not examine these further.
+Fourteen of the remaining documents had been published since 2010 (after the scope of the primary study).
 Among these, one obituary and two papers celebrated the work of John Loehlin, a prominent race-and-intelligence researcher who had been Director of the American Eugenics Society from 1968-1972 \cite{JohnLoehlin19262020, WaldmanIntroductionFestschriftJohn2014, PlominGenotypeEnvironmentCorrelationEra2014}.
@@
 
-28 documents in this sample from BG were published between 1990 and 2010.
+Twenty-eight documents in this sample from BG were published between 1990 and 2010.
 One was an obituary of Jerry Hirsch, a critic of hereditarianism in general and Jensen in particular \cite{RoubertouxJerryHirsch202008}, and another was an obituary of David Rowe \cite{RodgersObituaryDavidChristian2003}.

### 2023-09-26 paper/paper.Rmd
updating paper and refs-manual
I've updated the "Race science, behavior genetics, and psychology" section to include more text relevant to psychology's lack of forceful repudiation of race science relative to other disciplines, and updated refs-manual to include the new references for that text.

@@
 | Michael Levin (1943-) | philosophy | 96 |
-| Richard Lynn (1930-) | psychology | 288 |
+| Richard Lynn (1930-2023) | psychology | 288 |
 | R. Travis Osborne (1913-2013) | psychology | 59 |
@@
 In addition, the 1995 BGA presidential address by Glayde Whitney --- in which he criticized the "Marixist-Lysenkoist denial of genetics" and proposed that differences in murder rates between countries and cities was caused by racial genetic differences in intelligence, empathy, aggression, and impulsivity --- was published in MQ rather than BG.
-As a discipline, behavior genetics may have been prominent in forming an echo chamber for race science in the past, but appears to have been less receptive to such ideas in more recent history.
+As a subdiscipline of psychology, behavior genetics may have been prominent in promoting race science in the past, but appears to have been less receptive to such ideas in more recent history.
 
-Instead, our results suggest that psychology has been the dominant discipline in which the race science echo chamber has been maintained.
+Instead, our results suggest that journals focusing subdisciplines within psychology beyond behavior genetics have been the environments in which the mainstreaming of race science has taken place.
 This may be the result of racial essentialization being maintained more easily in a discipline like psychology than in other disciplines that contribute to behavior genetics, like genetics and anthropology.
-The typical practitioners in psychology do not have the degree of formal education about, and therefore knowledge of, relevant underlying genetic and anthropological research that has led many professional organizations in genetics and anthropology to formally reject race as a biologically meaningful concept \cite{anthropology1996aapa, american2018ashg, board1998aaa, FuentesAAPAStatementRace2019, nhgri2023, rotimi2022ashg, wynshawboris2020ashg}, undermining the core assumption of race science.
+The typical practitioners in psychology do not have the degree of formal education about, and therefore knowledge of, relevant underlying genetic and anthropological research that has led many professional organizations in genetics and anthropology to formally and routinely reject race as a biologically meaningful concept over the past several decades \cite{anthropology1996aapa, american2018ashg, board1998aaa, committee2020asa, FuentesAAPAStatementRace2019, nhgri2023, rotimi2022ashg, wynshawboris2020ashg}, undermining the core assumption of race science.
+
 As described by Winston, however, "[s]ince the 1990s, scientific racism has evolved in ways that sidestep or blunt the charge of essentialist categorization" \cite[p.
 427]{WinstonWhyMainstreamResearch2020}, such as shifting the rhetoric away from race to concepts such as "national IQ" that are deployed to preserve an essentialist categorization of different human populations.
-Professional psychological organizations routinely publish statements rejecting racism or acknowledging psychology's complicity in systemic racism \cite{american2021apology, aps2021statement}, but stop short of rejecting race as a meaningful biological categorization of people, allowing for race science research to be smuggled into psychology and giving the appearance of continued scientific legitimacy.
+As such, another way to view the role that psychology has played in maintaining the legitimacy of race science is by attending to top-down influences within the profession.
+Professional psychological organizations routinely publish statements rejecting racism or acknowledging psychology's historical complicity in systemic racism, but do not uniformly reject race as a biological concept.
+The American Psychological Association (APA), whose members are predominantly clinical psychologists, and the Federation of Associations in Behavioral & Brain Sciences (FABBS), an organization whose purpose is to provide policy recommendations, have openly rejected race as biologically meaningful \cite{american2021apology, baron2022fabbs} whereas research-focused psychological organizations like the Association for Psychological Science (APS) and Psychonomic Society (PS) stop short of rejecting race as a biologically meaningful when they denounce racism\cite{aps2021statement, board2020psychonomic}.
+The lack of a uniform professional rejection of the core claim of race science allows for race science research to be smuggled into mainstream psychological research.
+The hesitancy of professional organizations like APS or PS are potentially due to the prominence with which several notable race scientists have achieved in scientific gate keeping positions, such as on journal editorial boards.
+Additionally, psychology as a discipline routinely upholds several mythological narratives concerning race science and the history of psychology, such as allegations of a taboo about race and intelligence research\cite{jackson2021mythical}, claims that scientists engaging in or calling for any kind of anti-racism in the field is an ideological corruption of dispassionate and value-neutral science\cite{roberts_2022}, or arguments that holding race science to the kinds of evidentiary standards that other psychological research is held to is somehow a violation of academic freedom\cite{herbert2023academic}.
 

### 2023-10-04 paper/paper.Rmd
added a reference?

--- 2023-09-06 paper/paper.Rmd
@@
 By the end of the Second World War, a combination of technical developments in anthropology and genetics, along with revelations of Nazi atrocities, had led most members of the scientific community to reject attempts to scientifically justify racial inequalities and colonialism \cite{BarkanRetreatScientificRacism1992}.
-But eugenic thinking and scientific racism persisted throughout the second half of the twentieth century, and recent work has found that white supremacists continue to attempt to use scientific research to justify racial violence \cite{DusterBackdoorEugenics2003, PanofskyHowWhiteNationalists2021}.
+But eugenic thinking and scientific racism persisted throughout the second half of the twentieth century, and recent work has found that white supremacists continue to attempt to use scientific research to justify racial violence \cite{DusterBackdoorEugenics2003, PanofskyHowWhiteNationalists2021, PronczukRacistResearcherExposed2022}.
 

### 2023-10-04 paper/paper.Rmd
edits based on comments from K, J, M

--- 2023-10-04 paper/paper.Rmd
@@
 Thanks to Derek Devnich and James Dooley for their work attempting to secure the APA-published articles.
+Thanks to Kevin Bird, John Jackson, and Manuela Fernández Pinot for comments on a draft of this paper.
   | **Funding:** EL's work on this project was partially supported by UC Merced.  DH's work on this project was not supported by any specific funding. 
@@
     number_sections: true
+    latex_engine: 'xelatex'
 header-includes:
@@
 The Court's decision relied on expert testimony from psychologists and education researchers; but the segregationists also put forward their own experts, including Henry Garrett \cite{WinstonScienceServiceFar1998, JacksonScienceSegregationRace2005, SchafferScientificRacismAgain2007}.
+As part of this segregationist reaction, MQ was created to provide a favorable venue for race scientists to publish their views, on the grounds that an "equalitarian dogma" created a censorious "taboo" against their research in mainstream publications \cite{TuckerFundingScientificRacism2002, JacksonMythicalTabooRace2020}.
 
-As part of the segregationist reaction to *Brown*, MQ was created as what contemporary historians, philosophers, and sociologists of science call an "echo chamber" \cite{FernandezPintoKnowBetterNot2017}, a favorable venue for race scientists to publish their views, on the grounds that an "equalitarian dogma" created a censorious "taboo" against their research in mainstream publications \cite{TuckerFundingScientificRacism2002, JacksonMythicalTabooRace2020}.
-Echo chambers are a key part of the "tobacco strategy," used by numerous regulated industries — most infamously tobacco, but also fossil fuels and chemical manufacturing, among others — to "manufacture doubt" and delay regulation \cite{OreskesMerchantsDoubtHow2011, JasnyEmpiricalExaminationEcho2015, FernandezPintoKnowBetterNot2017, WeatherallHowBeatScience2018}.
-
-In light of the origins of Pioneer and MQ, the significant attention paid to the journal by historians of scientific racism, and contemporary research on the "tobacco strategy," we hypothesized that bibliometric and text mining analyses would confirm that MQ functioned as an echo chamber for Pioneer-funded scientists.
+In light of the scientific racist origins of Pioneer and MQ, and the significant attention MQ has received in recent historiography, we hypothesized that bibliometric and text mining analyses would show race science ideas originating in MQ, and from there being disseminated to mainstream publications.
 This hypothesis was not supported.
-Instead, mainstream psychology journals served as the echo chambers for race science.
+Instead, mainstream psychology journals provided a venue for a distinct form of race science, based on intelligence research and originating outside of MQ.
 
@@
 
-Figure \@ref(fig:wos) shows that, while MQ is among the "Pioneer-publishing" journals, a number of mainstream journals are more prominent: *Personality and Individual Differences* (PID), *Intelligence* (I), *Behavior Genetics* (BG), and *Psychological Reports* (PR).
+Figure \@ref(fig:wos) shows that, while MQ is among the "Pioneer-publishing" journals, a number of mainstream journals are more prominent: *Personality and Individual Differences* (PID), *Intelligence* (Int), *Behavior Genetics* (BG), and *Psychological Reports* (PR).
 In addition, only psychologist Richard Lynn appears to have published heavily in MQ.
@@
 Phillippe Rushton in 2012 \cite{BeirichPioneerFundAssets2013}.
-By contrast, a number of Pioneer-funded researchers have published in PID, I, and to a lesser degree BG: Bouchard, Eysenck, Jensen, Rushton, Vernon, and also Lynn.
+By contrast, a number of Pioneer-funded researchers have published in PID, Int, and to a lesser degree BG: Bouchard, Eysenck, Jensen, Rushton, Vernon, and also Lynn.
 
@@
 
-*Intelligence* (I) was founded in 1977, with psychologist Douglas Detterman as editor-in-chief from the founding until 2016.
+*Intelligence* (Int) was founded in 1977, with psychologist Douglas Detterman as editor-in-chief from the founding until 2016.
 Lloyd Humphreys was on the editorial board starting from 1977; by 1990 he had been joined by Jensen and Philip Vernon.
 Richard Lynn joined the editorial board sometime between 1998 and 2002.
-(Archive copies of the I editorial board page are not available from the journal's website from 1999 through 2001.)  I has been criticized for including Lynn and Gerhard Meisenberg — who was editor-in-chief of MQ in 2015-18 — on its editorial board until 2018 \cite{SainiSuperiorReturnRace2019}.
-I is published by Elsevier.
+(Archive copies of the Int editorial board page are not available from the journal's website from 1999 through 2001.)  Int has been criticized for including Lynn and Gerhard Meisenberg — who was editor-in-chief of MQ in 2015-18 — on its editorial board until 2018 \cite{SainiSuperiorReturnRace2019}.
+Int is published by Elsevier.
 
@@
 The fact that Pioneer-funded researchers published heavily in two mainstream psychology journals does not tell us anything about the content of their publications or the claims of their research.
-We assembled a full-text corpus of articles published in MQ, PID, I, BG, PR, and Behavior and Brain Sciences (BBS) between 1960-2010 and used topic modeling to identify topics discussing race, intelligence, and both.
+We assembled a full-text corpus of articles published in MQ, PID, Int, BG, PR, and Behavior and Brain Sciences (BBS) between 1960-2010 and used topic modeling to identify topics discussing race, intelligence, and both.
 
@@
 In the focal model, topic 07 is strongly associated with MQ:  MQ published dozens of articles in this topic each year, and no other journal ever published more than a handful.
-The Silge plot of the top 15 terms \cite{SilgeTopicModeling2017} contains racial terms (`races`, `whites`, `negroes` and potentially `europe`, `africa`, `india`, and `japan`) as well as `book`, likely reflecting the fact that MQ published a number of book reviews, while the psychology journals either did not or these were not available for the corpus.
+The "Silge plot," showing the top 15 terms in the topic \cite{SilgeTopicModeling2017}, contains racial terms (`races`, `whites`, `negroes` and potentially `europe`, `africa`, `india`, and `japan`) as well as `book`, likely reflecting the fact that MQ published a number of book reviews, while the psychology journals either did not or these were not available for the corpus.
 
-Topic 22 is strongly associated with I and PID in the same way that 07 is associated with MQ.
+Topic 22 is strongly associated with Int and PID in the same way that 07 is associated with MQ.
 The Silge plot does contain `jensen`, as well as a reference to Raymond Cattell, who played a major role in the development of factor analysis and intelligence testing but also advocated for eugenics, fascism, and Nazi race science \cite{MehlerBeyondismRaymondCattell1997}.
 However, the other authors named in this topic — John Horn and Peter Bentler — do not appear to have contributed to scientific racism.
-Instead this topic appears to identify "mainstream" (non-race science) intelligence research, especially factor analysis and the debate over whether intelligence in unidimensional or multidimensional.
+Instead this topic appears to identify "mainstream" (non-race science) intelligence research, especially factor analysis and the debate over whether intelligence is unidimensional or multidimensional.
 (Three other topics only associated with mainstream intelligence journals and terms were also identified by this topic model.)  This topic indicates that the model is not simply lumping race science research together with other intelligence research.
@@
 In almost all years, most papers in topic 24 were published in mainstream journals rather than MQ.
-Jensen's "How Much Can We Boost IQ and Scholastic Achievement?" \cite{JensenHowMuchCan1969} was published in 1969 in Harvard Educational Review (not included in this study), and the time series indicates that, around this time, there was an increase in articles in topic 24 in both PS and MQ (the only two journals in our corpus that were active at the time).
-MQ shows another sharp increase in the late 1970s, around the time Lynn was brought on as assistant editor.
-This finding is consistent with the conclusion of a qualitative analysis that "Psychology became the dominant discipline upon the editorship of Richard Lynn and Gerhard Meisenberg in MQ, where some issues were almost entirely represented by IQ test–based studies" \cite[p.
-110]{AdamsMisAppropriationBiological2021}.
-PID published multiple papers in topic 24 almost immediately after it was founded, with I showing a more gradual increase between the mid-1970s and mid-1990s.
+Jensen's "How Much Can We Boost IQ and Scholastic Achievement?" \cite{JensenHowMuchCan1969} was published in 1969 in *Harvard Educational Review* (not included in this study), and the time series indicates that, in the early 1970s, there was an increase in articles in topic 24 in both PS and MQ (the only two journals in our corpus that were active at the time).
+MQ shows another sharp increase in the late 1980s; using content analysis, Adams and Pilloud found that psychology was the dominant discipline in MQ in the period 1992-2018 \cite{AdamsMisAppropriationBiological2021}.
+PID published multiple papers in topic 24 almost immediately after it was founded, with Int showing a more gradual increase between the mid-1970s and mid-1990s.
 
@@
 Panofsky \cite{PanofskyMisbehavingScienceControversy2014} argues that, prior to Jensen's 1969 paper, behavior genetics emphasized the study of non-human animals and intentionally avoided associations with eugenics and public controversy more generally.
-In response to Jensen, critics such as Lewontin offered broad critiques of behavior genetics as such, and behavior geneticists in turn adopted a radical conception of academic freedom (without any sense of responsibility for the social implications of academic research) and a siege or wartime mentality, as demonstrated by Sandra Scarr's 1986 presidential address to the Behavior Genetics Association (BGA) \cite{ScarrThreeCheersBehavior1987}.
+In response to Jensen, critics such as Lewontin offered broad critiques of behavior genetics as such, and behavior geneticists in turn adopted a radical conception of academic freedom (without any sense of responsibility for the social implications of academic research) and a siege or wartime mentality, as illustrated by Sandra Scarr's 1986 presidential address to the Behavior Genetics Association (BGA) \cite{ScarrThreeCheersBehavior1987}.
 Scarr's address coincided with the period between 1970-1990 when BG published articles in topic 24 (including Scarr's address itself).
@@
 However, the topic model analysis suggests that by the 1990s, behavior genetics may have distanced itself somewhat from race science, albeit without directly repudiating it.
-This interpretation is further supported by two disciplinary histories of behavior genetics, one published in McGue in 2010 \cite{McGueEndBehavioralGenetics2010} and the other by Plomin in 2023 \cite{PlominCelebratingCenturyResearch2023}.
-In addition, the 1995 BGA presidential address by Glayde Whitney --- in which he criticized the "Marixist-Lysenkoist denial of genetics" and proposed that differences in murder rates between countries and cities was caused by racial genetic differences in intelligence, empathy, aggression, and impulsivity --- was published in MQ rather than BG.
-As a discipline, behavior genetics may have been prominent in forming an echo chamber for race science in the past, but appears to have been less receptive to such ideas in more recent history.
+This interpretation is further supported by two disciplinary histories of behavior genetics published in BG in 2010 and 2023 \cite{McGueEndBehavioralGenetics2010, PlominCelebratingCenturyResearch2023}.
+In addition, the 1995 BGA presidential address by Glayde Whitney --- in which he criticized the "Marxist-Lysenkoist denial of genetics" and proposed that differences in murder rates between countries and cities was caused by racial genetic differences in intelligence, empathy, aggression, and impulsivity --- was published in MQ rather than BG.
+As a subdiscipline of psychology, behavior genetics may have been prominent in promoting race science in the past, but appears to have been less receptive to such ideas in more recent history.
 
-Instead, our results suggest that psychology has been the dominant discipline in which the race science echo chamber has been maintained.
-This may be the result of racial essentialization being maintained more easily in a discipline like psychology than in other disciplines that contribute to behavior genetics, like genetics and anthropology.
-The typical practitioners in psychology do not have the degree of formal education about, and therefore knowledge of, relevant underlying genetic and anthropological research that has led many professional organizations in genetics and anthropology to formally reject race as a biologically meaningful concept \cite{anthropology1996aapa, american2018ashg, board1998aaa, FuentesAAPAStatementRace2019, nhgri2023, rotimi2022ashg, wynshawboris2020ashg}, undermining the core assumption of race science.
-As described by Winston, however, "[s]ince the 1990s, scientific racism has evolved in ways that sidestep or blunt the charge of essentialist categorization" \cite[p.
-427]{WinstonWhyMainstreamResearch2020}, such as shifting the rhetoric away from race to concepts such as "national IQ" that are deployed to preserve an essentialist categorization of different human populations.
-Professional psychological organizations routinely publish statements rejecting racism or acknowledging psychology's complicity in systemic racism \cite{american2021apology, aps2021statement}, but stop short of rejecting race as a meaningful biological categorization of people, allowing for race science research to be smuggled into psychology and giving the appearance of continued scientific legitimacy.
+Instead, our results suggest that other subdisciplines of psychology have provided the primary venue for mainstreaming race science.
+Over the past few decades, many professional organizations in genetics and anthropology have made formal statements rejecting race as a biologically meaningful concept \cite{anthropology1996aapa, american2018ashg, board1998aaa, committee2020asa, FuentesAAPAStatementRace2019, nhgri2023, rotimi2022ashg, wynshawboris2020ashg}, undermining the core assumption of race science.
+The American Psychological Association (APA) --- whose members are predominantly clinical psychologists --- and the Federation of Associations in Behavioral & Brain Sciences (FABBS) --- an organization whose purpose is to provide policy recommendations --- have openly rejected race as biologically meaningful \cite{american2021apology, baron2022fabbs}.
+But research-focused psychological organizations like the Association for Psychological Science (APS) and Psychonomic Society (PS) stop short of rejecting race as biologically meaningful when they denounce racism \cite{aps2021statement, board2020psychonomic}.
+Psychology as a discipline maintains space for several mythological race science narratives, such as allegations of a taboo against race and intelligence research \cite{jackson2021mythical}, claims that scientists engaging in or calling for any kind of anti-racism in the field is an ideological corruption of dispassionate and value-neutral science \cite{roberts_2022}, or arguments that holding race science to the same evidentiary standards as other psychological research is a violation of academic freedom \cite{herbert2023academic}.
+And historically several race scientists have secured important gatekeeping positions within the scientific community, such as on the editorial boards of Int and PID.
 
@@
 
-Using bibliometric and text-mining methods, this study finds that Pioneer-funded researchers published heavily in certain mainstream psychology journals, much more than in *Mankind Quarterly*; and that a distinct topic of race science discourse, centered on race and intelligence, can be identified in these same journals across the period 1960-2010.
-These findings indicate that *Mankind Quarterly* (MQ) was less important as an echo chamber for late 20th century race science than mainstream psychology journals, especially *Intelligence* (I) and *Personality and Individual Differences* (PID).
-Indeed, MQ published a very different kind of race science from the race-and-intelligence research published in I and PID.
+Using bibliometric and text-mining methods, this study finds that Pioneer-funded researchers published heavily in certain mainstream psychology journals, much more than in *Mankind Quarterly*; and that a distinct topic of race science discourse, centered on race and intelligence, can be identified in these same mainstream journals across the period 1960-2010.
+These findings indicate that *Mankind Quarterly* (MQ) was less important as venue for late 20th century race science than mainstream psychology journals, especially *Intelligence* (Int) and *Personality and Individual Differences* (PID).
+Indeed, until the 1990s, MQ published a very different kind of race science from the race-and-intelligence research published in Int and PID.
+During this same period of time, race scientists served in prominent and influential positions in behavior genetics and psychology, including as society presidents and members of journal editorial boards.
+In some cases, some race scientists simultaneously maintained active connections to both the mainstream scientific community and white supremacist organizations \cite{WinstonScientificRacismNorth2020, JacksonJr.ArthurJensenEvolutionary2022, SlobodianUnequalMindHow2023}.
+As far as we are aware, no organization for scientific research in psychology has apologized for the historical contributions of their fields to scientific racism.
 
-We stress that we are not identifying any particular scientist as a racist, scientific racist, or race scientist based on the topic model results alone.
-Scientific research can be utilized to promote scientific racism — and thus count as race science — independently of the views and attitudes of the scientist who conducted that research \cite{TaberyWhyStudyingGenetics2015, GillbornSoftlySoftlyGenetics2016, CarlsonQuantifyingContextualizingImpact2020, HennWhyDNANo2021}.
-But this does not mean that scientists are not responsible for the downstream social effects of their
-research, even when these effects involve deliberate misinterpretation \cite{DouglasSciencePolicyValuefree2009, KitcherArgumentFreeInquiry1997, KouranyShouldKnowledgeBe2016, BlockIQHeritabilityInequality1974a}.
-Like other citizens, scientists are responsible for mitigating the reasonably foreseen harms that result from their actions.
-And we now have decades of empirical evidence that research on the genetics of intelligence — even research that does not itself touch on race — can and will be used to promote scientific racism and indeed racial violence \cite{MeyerWrestlingSocialBehavioral2023}.
-For instance, while our findings suggest that behavior genetics has attempted to distance itself from scientific racism, it has done little to explicitly repudiate this research, especially in contrast to statements by biological anthropologists \cite{FuentesAAPAStatementRace2019} and human geneticists \cite{JacksonFacingOurHistory2023}.
-Behavior genetics research — and research on human genetics more generally — may be especially susceptible to appropriation by scientific racists, which only strengthens the obligations these researchers have to prevent their research from causing harm \cite{CarlsonQuantifyingContextualizingImpact2020, OgbunugaforDNABasketballBirthday2022}.
+We are not identifying any particular scientist as a racist, scientific racist, or race scientist based on the topic model results alone.
+Such claims require an analysis of documentary evidence that goes beyond the scope of the current study \cite{JacksonJr.ArthurJensenEvolutionary2022}.
+We do assume that our readers, like us, regard white supremacy and scientific racism as morally odious and beyond the scope of reasonable debate \cite{SchroederLimitsDemocratizingScience2022}.
+However, scientific research can be appropriated to promote scientific racism — and thus count as race science — even when this is contrary to the intentions of the original researchers themselves \cite{TaberyWhyStudyingGenetics2015, GillbornSoftlySoftlyGenetics2016, CarlsonQuantifyingContextualizingImpact2020, HennWhyDNANo2021}.
+Panofsky et al.
+show how the term "human biodiversity," originally developed for anti-racist purposes by biological anthropologist Jonathan Marks, has been inverted by white supremacists and is now a dogwhistle for biological racial hierarchy \cite{PanofskyHowWhiteNationalists2021}.
 
-Our distinction between race science and race science discourse highlights an important "engagement dilemma" for scientists challenging race science and scientific racism.
-While the "tobacco strategy" for manufacturing doubt targets mainstream scientists and promotes fringe scientific views, its primary audiences are lawmakers and the general public.
-The aim of the strategy is not to change the opinion of mainstream scientists, but instead to cultivate the impression that there is legitimate scientific uncertainty about the issue, reasonable disagreement among qualified experts \cite{FernandezPintoKnowBetterNot2017}.
-On the one hand, refusing to engage with fringe scientists — or, often, pseudoscientists \cite{BhakthavatsalamVirtueEpistemologicalApproach2021} — means that false or misleading statements go unchallenged.
-On the other hand, detailed, technical criticism legitimizes the fringe views, creating exactly the impression of reasonable scientific disagreement that the tobacco strategy is intended to produce.
+Research on the genetics of intelligence has been socially harmful, not just in the somewhat abstract sense of promoting racial stigma, but in the concrete sense of being used to rationalize mass shootings and other acts of racial violence \cite{PronczukRacistResearcherExposed2022, MeyerWrestlingSocialBehavioral2023}.
+While scientists do not fully control the downstream social effects of their research, like other citizens scientists are responsible for mitigating reasonably foreseeable harms that result from their actions \cite{BlockIQHeritabilityInequality1974a, DouglasSciencePolicyValuefree2009, KitcherArgumentFreeInquiry1997, KouranyShouldKnowledgeBe2016}.
+These actions include not only individual decisions to research this topic or that, but also collective decisions about who is awarded positions of power or influence.
+Given the way research on intelligence and behavior genetics has been used historically, these fields may be especially susceptible to appropriation by scientific racists.
+This only strengthens the obligations of researchers in these areas to prevent their research from causing harm \cite{CarlsonQuantifyingContextualizingImpact2020, OgbunugaforDNABasketballBirthday2022}.
 
-In the case of race science, mainstream psychologists have often taken a non-engagement approach — whether due to a strong conception of academic freedom or a desire to avoid stoking controversy — with the result that race scientists have been able to present themselves as the representatives of "mainstream science on intelligence."  On the other hand, critics — most prominently Gould, Kamin, and Lewontin — have offered sophisticated technical critiques.
-But these tend to spiral off into esoteric debates that are inaccessible to the general public, as with "Lewontin's fallacy" \cite{RosemanLewontinDidNot2021}.
-Consequently these critiques are likely to be either ignored by the general public or taken as evidence of legitimate scientific debate.
+How should psychologists exercise these responsibilities?
+We believe biological anthropology could provide an important model for psychology.
+Anthropologists not only denounce racism in general, but the specific ways that anthropology has (and, in some cases, continues to be) used to provide cover for scientific racism.
+Anthropologists produce both moral and technical critiques of race science, in venues and language that are appropriate for the general public.
+And they recognize the importance of reparations, taking active steps to repair relationships with communities that have been harmed by anthropology.
 
@@
 For $k=5, 10$ the distributions are quite noisy, difficult to interpret, and generally don't form very strong visual clusters.
-As $k$ increases from 20 through 40, coherent bands start to appear for journal-distinctive topics in MQ, BG, I and, to a lesser degree, PID.
-Note that the distinctive topics for MQ and I do not obviously coincide.
+As $k$ increases from 20 through 40, coherent bands start to appear for journal-distinctive topics in MQ, BG, Int and, to a lesser degree, PID.
+Note that the distinctive topics for MQ and Int do not obviously coincide.
 Above about $k=50$, the models appear to identify more fine-grained topics and the coherent bands fade.
@@
 The 3 non-consensus documents reflected essential ambiguity in the operationalization of race science discourse.
-One is a table of contents from a 1980 issue of I; arguably this document should have been excluded as irrelevant front matter, but arguably it legitimizes the research published in the issue.
+One is a table of contents from a 1980 issue of Int; arguably this document should have been excluded as irrelevant front matter, but arguably it legitimizes the research published in the issue.
 The second case is a paper on racial differences in emotional intelligence that was classified as race science discourse by one author, while the other author felt it was more focused on psychometrics \cite{GignacGroupDifferencesEI2010}.
@@
 A key limitation of the current study stems from the inability of text mining techniques to draw on a broader cultural context than what is represented in the corpus.
-Scholars of racism and US race relations have noted a shift in the way racist attitudes are understood and expressed, from overt and direct to a subtle "color-blind" racism \cite{bonillasilva2022racism}.
+Scholars of racism and US race relations have noted a shift in the way racist attitudes are understood and expressed, from overt and direct to a subtle "color-blind" racism \cite{Bonilla-SilvaRacismRacistsColorBlind2006}.
 Color-blind racism represents a change in the primary mechanisms by which white privilege is maintained, using non-racialized language in furtherance of the racial status quo both institutionally and individually, and increasingly reliant on the invisibility of socio-cultural mechanisms to preserve inequalities.
@@
 Alderfer notes that the racist implications of *The Bell Curve* only emerge gradually, with the opening of the book using the color-blind language of "group differences" \cite{AlderferScienceNonsciencePsychologists2003}.
-Panofsky et al.
-show how the term "human biodiversity," originally developed for anti-racist purposes by biological anthropologist Jonathan Marks, has been inverted by white supremacists and is now a dogwhistle for biological racial hierarchy \cite{PanofskyHowWhiteNationalists2021}.
 These cases suggest further that color-blindness can facilitate white supremacist appropriation of non-racial research — even contrary to the researchers' own intentions — as scientific racist readers re-interpret racial-neutral language as dogwhistles \cite{GillbornSoftlySoftlyGenetics2016, WillsAreClustersRaces2017}.

### 2023-10-20 paper/paper.Rmd
tweak final paragraph

@@
     corresponding_author: true
-  - name: "Emilio Lobato"
+  - name: "Emilio J. C. Lobato"
     affiliation: 1
@@
 Thanks to Derek Devnich and James Dooley for their work attempting to secure the APA-published articles.
-Thanks to Kevin Bird, John Jackson, and Manuela Fernández Pinot for comments on a draft of this paper.
-  | **Funding:** EL's work on this project was partially supported by UC Merced.  DH's work on this project was not supported by any specific funding. 
+Thanks to Kevin Bird, John Jackson, and Manuela Fernández Pinto for comments on a draft of this paper.
+  | **Funding:** EL's work on this project was partially supported by UC Merced.  DJH's work on this project was not supported by any specific funding. 
   | **Author contributions:** DJH (conceptualization, data curation, formal analysis, investigation, methodology, project administration, software, validation, visualization, writing-original draft, writing-review & editing); EL (conceptualization, methodology, software, validation, visualization, writing-original draft, writing-review & editing)
@@
 Preprocessed analysis data (document-term counts and article metadata) and analysis scripts are available at *[doi]*.
-# bibliography: "references.bib"
+# bibliography: references.bib, refs-manual.bib
 output: 
@@
 In almost all years, most papers in topic 24 were published in mainstream journals rather than MQ.
-Jensen's "How Much Can We Boost IQ and Scholastic Achievement?" \cite{JensenHowMuchCan1969} was published in 1969 in *Harvard Educational Review* (not included in this study), and the time series indicates that, in the early 1970s, there was an increase in articles in topic 24 in both PS and MQ (the only two journals in our corpus that were active at the time).
+Jensen's "How Much Can We Boost IQ and Scholastic Achievement?" \cite{JensenHowMuchCan1969} was published in 1969 in *Harvard Educational Review* (not included in this study), and the time series indicates that, in the early 1970s, there was an increase in articles in topic 24 in both PR and MQ (the only two journals in our corpus that were active at the time).
 MQ shows another sharp increase in the late 1980s; using content analysis, Adams and Pilloud found that psychology was the dominant discipline in MQ in the period 1992-2018 \cite{AdamsMisAppropriationBiological2021}.
@@
 
-However, the topic model analysis suggests that by the 1990s, behavior genetics may have distanced itself somewhat from race science, albeit without directly repudiating it.
-This interpretation is further supported by two disciplinary histories of behavior genetics published in BG in 2010 and 2023 \cite{McGueEndBehavioralGenetics2010, PlominCelebratingCenturyResearch2023}.
-In addition, the 1995 BGA presidential address by Glayde Whitney --- in which he criticized the "Marxist-Lysenkoist denial of genetics" and proposed that differences in murder rates between countries and cities was caused by racial genetic differences in intelligence, empathy, aggression, and impulsivity --- was published in MQ rather than BG.
+However, the topic model analysis suggests that, by the 1990s, behavior genetics may have distanced itself somewhat from race science, albeit without directly repudiating it.
+This interpretation was supported by a supplemental analysis that focused on BG specifically (SM section 1.5).
+In addition, the 1995 BGA presidential address by Glayde Whitney --- in which he criticized the "Marxist-Lysenkoist denial of genetics" and proposed that differences in murder rates between countries and cities were caused by racial genetic differences in intelligence, empathy, aggression, and impulsivity --- was published in MQ rather than BG.
 As a subdiscipline of psychology, behavior genetics may have been prominent in promoting race science in the past, but appears to have been less receptive to such ideas in more recent history.
@@
 In some cases, some race scientists simultaneously maintained active connections to both the mainstream scientific community and white supremacist organizations \cite{WinstonScientificRacismNorth2020, JacksonJr.ArthurJensenEvolutionary2022, SlobodianUnequalMindHow2023}.
-As far as we are aware, no organization for scientific research in psychology has apologized for the historical contributions of their fields to scientific racism.
+Except for the APA, we are not aware of other major organizations for scientific research in psychology that have issued apologies for the historical contributions of their field to scientific racism.
 
@@
 How should psychologists exercise these responsibilities?
-We believe biological anthropology could provide an important model for psychology.
-Anthropologists not only denounce racism in general, but the specific ways that anthropology has (and, in some cases, continues to be) used to provide cover for scientific racism.
-Anthropologists produce both moral and technical critiques of race science, in venues and language that are appropriate for the general public.
-And they recognize the importance of reparations, taking active steps to repair relationships with communities that have been harmed by anthropology.
+We believe restorative or reparative justice provides a useful model \cite{WenzelRetributiveRestorativeJustice2008, WormerRestorativeJustice2013}.
+Unlike retributive justice (used by almost all criminal justice systems today), the aim of restorative justice is to repair the damaged relationships between victims and perpetrators of injustice.
+Typically, restorative justice requires perpetrators to not only acknowledge their actions and the harms that these actions had on victims, but also to work with victims to identify concrete actions that perpetrators can take to redress or mitigate these harms.
+From this perspective, explicit apologies for contributions to scientific racism are essential as a first step \cite{FuentesAAPAStatementRace2019, american2021apology, JacksonFacingOurHistory2023}, but not fully sufficient on their own.
 
+<!-- 
 \clearpage
@@
 \bibliography{references.bib, refs-manual.bib}
+ -->
 
@@
 Many of the documents published in *Psychological Reports* are extremely short, 1-2 page brief notices of a single study; while many of the documents published in *Brain and Behavioral Sciences* are book-length collections that include a long review article and sometimes dozens of commentaries.
-Very generic noun phrases that happen to appear in BBS can occur orders of magnitude more often than highly distinctive phrases in the PS, and so the $log(n)$ factor overwhelms the $\Delta H$ factor.
+Very generic noun phrases that happen to appear in BBS can occur orders of magnitude more often than highly distinctive phrases in PS, and so the $log(n)$ factor overwhelms the $\Delta H$ factor.
 
@@
 This novel approach was implemented in the R package `tmfast` \cite{HicksTmfastFitsTopic2023}.
-A simulation study of `tmfast` found that it was significantly faster and only slightly less accurate at reconstructing known word-topic and topic-document distributions, compared to `stm` \cite{HicksTmfastFitsTopic2023}.
+We conducted a simulation study of `tmfast`, which found that it was significantly faster and only slightly less accurate at reconstructing known word-topic and topic-document distributions, compared to `stm` \cite{HicksTmfastFitsTopic2023}.
 
@@
 The second case is a paper on racial differences in emotional intelligence that was classified as race science discourse by one author, while the other author felt it was more focused on psychometrics \cite{GignacGroupDifferencesEI2010}.
-A finally, a 1990 article in MQ (<https://www.unz.com/print/MankindQuarterly-1990q3-00108/>) proposed to offer an economic explanation for crime.
+And a 1990 article in MQ (<https://www.unz.com/print/MankindQuarterly-1990q3-00108/>) proposed to offer an economic explanation for crime.
 While it has a few pages connecting race/ethnicity to both crime and poverty, it ultimately focuses on unemployment and alcohol sales, rather than race, as the key predictors of crime.
@@
 
-## Supplementary review of *Behavior Genetics* {#bg}
+## Supplemental analysis of *Behavior Genetics* {#bg}
 
@@
 Topic models, of course, are not even capable of representing this essential ambiguity.
+
+\clearpage
+
+# References and Notes {.unnumbered}
+
+\bibliographystyle{Science}
+\bibliography{references.bib, refs-manual.bib}

### 2023-10-20 paper/paper.Rmd
Emilio initials

@@
 Thanks to Kevin Bird, John Jackson, and Manuela Fernández Pinto for comments on a draft of this paper.
-  | **Funding:** EL's work on this project was partially supported by UC Merced.  DJH's work on this project was not supported by any specific funding. 
-  | **Author contributions:** DJH (conceptualization, data curation, formal analysis, investigation, methodology, project administration, software, validation, visualization, writing-original draft, writing-review & editing); EL (conceptualization, methodology, software, validation, visualization, writing-original draft, writing-review & editing)
+  | **Funding:** EJCL's work on this project was partially supported by UC Merced.  DJH's work on this project was not supported by any specific funding. 
+  | **Author contributions:** DJH (conceptualization, data curation, formal analysis, investigation, methodology, project administration, software, validation, visualization, writing-original draft, writing-review & editing); EJCL (conceptualization, methodology, software, validation, visualization, writing-original draft, writing-review & editing)
   | **Competing interests:** The authors have no competing interests to declare. 

### 2023-10-23 paper/paper.Rmd
just do acknowledgments manually

@@
 
-acknowledgements: |
-  | Thanks to Emily Merchant and John Jackson for some initial discussions that helped clarify the scope of the project and identify some key resources related to the Pioneer Fund and *Mankind Quarterly*.
+# acknowledgements: |
+#   | Thanks to Emily Merchant and John Jackson for some initial discussions that helped clarify the scope of the project and identify some key resources related to the Pioneer Fund and *Mankind Quarterly*.
 Thanks to Anthony Sainez for help retrieving the *Mankind Quarterly* articles.
@@
 Thanks to Kevin Bird, John Jackson, and Manuela Fernández Pinto for comments on a draft of this paper.
-  | **Funding:** EJCL's work on this project was partially supported by UC Merced.  DJH's work on this project was not supported by any specific funding. 
-  | **Author contributions:** DJH (conceptualization, data curation, formal analysis, investigation, methodology, project administration, software, validation, visualization, writing-original draft, writing-review & editing); EJCL (conceptualization, methodology, software, validation, visualization, writing-original draft, writing-review & editing)
-  | **Competing interests:** The authors have no competing interests to declare. 
-  | **Data and materials availability:** Web of Science search results and article fulltext are not publicly available due to intellectual property restrictions.
+#   | **Funding:** EJCL's work on this project was partially supported by UC Merced.  DJH's work on this project was not supported by any specific funding. 
+#   | **Author contributions:** DJH (conceptualization, data curation, formal analysis, investigation, methodology, project administration, software, validation, visualization, writing-original draft, writing-review & editing); EJCL (conceptualization, methodology, software, validation, visualization, writing-original draft, writing-review & editing)
+#   | **Competing interests:** The authors have no competing interests to declare. 
+#   | **Data and materials availability:** Web of Science search results and article fulltext are not publicly available due to intellectual property restrictions.
 Preprocessed analysis data (document-term counts and article metadata) and analysis scripts are available at *[doi]*.
@@
 
-<!-- 
 \clearpage
@@
 \bibliography{references.bib, refs-manual.bib}
- -->
+
+| 
+| Thanks to Emily Merchant and John Jackson for some initial discussions that helped clarify the scope of the project and identify some key resources related to the Pioneer Fund and *Mankind Quarterly*.
+Thanks to Anthony Sainez for help retrieving the *Mankind Quarterly* articles.
+Thanks to Derek Devnich and James Dooley for their work attempting to secure the APA-published articles.
+Thanks to Kevin Bird, John Jackson, and Manuela Fernández Pinto for comments on a draft of this paper.
+| **Funding:** EJCL's work on this project was partially supported by UC Merced.  DJH's work on this project was not supported by any specific funding. 
+| **Author contributions:** DJH (conceptualization, data curation, formal analysis, investigation, methodology, project administration, software, validation, visualization, writing-original draft, writing-review & editing); EJCL (conceptualization, methodology, software, validation, visualization, writing-original draft, writing-review & editing)
+| **Competing interests:** The authors have no competing interests to declare. 
+| **Data and materials availability:** Web of Science search results and article fulltext are not publicly available due to intellectual property restrictions.
+Preprocessed analysis data (document-term counts and article metadata) and analysis scripts are available at *[doi]*.
 
@@
 
+<!-- 
 \clearpage
@@
 \bibliography{references.bib, refs-manual.bib}
+
+ -->

### 2023-10-27 paper/paper.qmd
initial conversion to quarto journal template

--- 2023-10-20 paper/paper.Rmd
@@
   - name: "Daniel J. Hicks"
-    affiliation: 1
-    corresponding_author: true
+    email: "dhicks4@ucmerced.edu"
+    affiliations: 
+      - id: ucm
+        name: "University of California, Merced"
+        city: "Merced"
+        state: "CA, USA"
+    corresponding: true
   - name: "Emilio J. C. Lobato"
-    affiliation: 1
-address:
-  - code: 1
-    address: "University of California, Merced"
-corresponding_author: "Corresponding author: Daniel J. Hicks, dhicks4@ucmerced.edu."
+    affiliations:
+      - ref: ucm
 abstract: |
@@
 Preprocessed analysis data (document-term counts and article metadata) and analysis scripts are available at *[doi]*.
-# bibliography: references.bib, refs-manual.bib
-output: 
-  # https://github.com/rstudio/rticles/pull/486
-  # skeleton: <https://github.com/christopherkenny/rticles/tree/master/inst/rmarkdown/templates/science>
-  bookdown::pdf_book:
-    base_format: rticles::science_article
+
+bibliography: [references.bib, refs-manual.bib]
+
+format: 
+  elsevier-pdf:
+    journal: 
+      name: "journal"
     keep_tex: true
-    move_figures: false
-    draft: true
-    number_sections: true
-    latex_engine: 'xelatex'
-header-includes:
-  - \usepackage{lineno}
-  - \linenumbers
+    latex_engine: "xelatex"
+
+# header-includes:
+#  - \usepackage{lineno}
+#  - \linenumbers
 ---
-
-<!-- 
-Science submission instructions
-
-General information: Research Articles
-<https://www.science.org/content/page/science-information-authors>
-- 5 printed pages in the journal
-- 2000 to 3000 words of main text, in addition to an abstract, 3 to 5 display items (figures or tables) with brief legends, about 50 main-text references, and a structured acknowledgments section
-- main text should be divided into sections with brief subheadings
-- Materials and Methods should be included in supplementary materials and should be followed by additional data and figures needed to support the paper’s conclusions
-- All data must be available in either the main text or the supplementary materials, or must be deposited at a publicly accessible repository and cited in the paper.
-
-Initial submission
-<https://www.science.org/content/page/instructions-preparing-initial-manuscript>
-- Title: 96 characters
-- Authors
-- Affiliations:
-- One Sentence Summary: 125 characters
-- Main Text:  The introduction should provide sufficient background information to make the article intelligible to readers in other disciplines and sufficient context so that the significance of the experimental findings is clear.
-Technical terms should be defined.
-- References and Notes: There should be a single reference list that combines references cited in the main paper as well as those cited in the supplementary materials, and each reference should appear only once in the list.
-List of Supplementary Materials
-- The width of figures, when printed, will usually be 5.7 cm (2.24 inches or 1 column), 12.1 cm (4.76 inches or 2 columns), or 18.4 cm (7.24 inches or 3 columns).
-
--->
 

### 2023-10-27 paper/paper.qmd
fixed refs?

@@
 bibliography: [references.bib, refs-manual.bib]
+csl: nature.csl
 
@@
 
-By the end of the Second World War, a combination of technical developments in anthropology and genetics, along with revelations of Nazi atrocities, had led most members of the scientific community to reject attempts to scientifically justify racial inequalities and colonialism \cite{BarkanRetreatScientificRacism1992}.
-But eugenic thinking and scientific racism persisted throughout the second half of the twentieth century, and recent work has found that white supremacists continue to attempt to use scientific research to justify racial violence \cite{DusterBackdoorEugenics2003, PanofskyHowWhiteNationalists2021, PronczukRacistResearcherExposed2022}.
+By the end of the Second World War, a combination of technical developments in anthropology and genetics, along with revelations of Nazi atrocities, had led most members of the scientific community to reject attempts to scientifically justify racial inequalities and colonialism [@BarkanRetreatScientificRacism1992].
+But eugenic thinking and scientific racism persisted throughout the second half of the twentieth century, and recent work has found that white supremacists continue to attempt to use scientific research to justify racial violence [@DusterBackdoorEugenics2003; @PanofskyHowWhiteNationalists2021; @PronczukRacistResearcherExposed2022].
 
@@
 These distinction allow us to bracket the intentions and mindset of the researchers involved in any particular piece of research or debate.
-For the purposes of the current study, it is not important whether, for instance, Arthur Jensen is or is not correctly labelled a scientific racist \cite{JacksonJr.ArthurJensenEvolutionary2022}.
+For the purposes of the current study, it is not important whether, for instance, Arthur Jensen is or is not correctly labelled a scientific racist [@JacksonJr.ArthurJensenEvolutionary2022].
 But it is important that much of Jensen's research was useful and indeed utilized to purportedly justify racial inequality.
@@
 
-Histories of scientific racism in the twentieth century have often emphasized the Pioneer Fund and the parascholarly journal *Mankind Quarterly* (MQ) \cite{
-MehlerFoundationFascismNew1989,
-BarkanRetreatScientificRacism1992,
-WinstonScienceServiceFar1998,
-TuckerFundingScientificRacism2002,
-SchafferScientificRacismAgain2007,
-SainiSuperiorReturnRace2019,
-WinstonScientificRacismNorth2020,
-AdamsMisAppropriationBiological2021,
-SainiDraperMillionsPhilanthropic2022
-}.
-Pioneer was formed in 1937, during the waning years of the eugenics movement in North America \cite{BarkanRetreatScientificRacism1992}, with the aim of "support[ing] academic research and the \'[sic]dissemination of information, into the 'problem of heredity and eugenics' and 'the problems of race betterment'" \cite[p.21, quoting Laughlin]{MehlerFoundationFascismNew1989}.
+Histories of scientific racism in the twentieth century have often emphasized the Pioneer Fund and the parascholarly journal *Mankind Quarterly* (MQ) [
+@BarkanRetreatScientificRacism1992;
+@MehlerFoundationFascismNew1989;
+@WinstonScienceServiceFar1998;
+@TuckerFundingScientificRacism2002;
+@SchafferScientificRacismAgain2007;
+@SainiSuperiorReturnRace2019;
+@WinstonScientificRacismNorth2020;
+@AdamsMisAppropriationBiological2021;
+@SainiDraperMillionsPhilanthropic2022].
+Pioneer was formed in 1937, during the waning years of the eugenics movement in North America [@BarkanRetreatScientificRacism1992], with the aim of "support[ing] academic research and the \'[sic]dissemination of information, into the 'problem of heredity and eugenics' and 'the problems of race betterment'" [@MehlerFoundationFascismNew1989 p.21, quoting Laughlin].
 MQ was founded in 1960 by biologist R.
@@
 Robert Gayre (self-styled as "Gayre of Gayre and Nigg"; 1907-1996).
-Gates' professional status had risen and fallen with eugenics and the explicit scientific racism of the 1920s, and by the end of the Second World War he was thoroughly marginalized \cite{WinstonScienceServiceFar1998}.
-In contrast, Garrett had been president of the American Psychological Association in 1946 and chair of Psychology at Columbia from 1941 to 1955 \cite{WinstonScienceServiceFar1998}.
+Gates' professional status had risen and fallen with eugenics and the explicit scientific racism of the 1920s, and by the end of the Second World War he was thoroughly marginalized [@WinstonScienceServiceFar1998].
+In contrast, Garrett had been president of the American Psychological Association in 1946 and chair of Psychology at Columbia from 1941 to 1955 [@WinstonScienceServiceFar1998].
 
 In the landmark case *Brown v Board of Education of Topeka* (1954), the US Supreme Court banned *de jure* educational segregation.
-The Court's decision relied on expert testimony from psychologists and education researchers; but the segregationists also put forward their own experts, including Henry Garrett \cite{WinstonScienceServiceFar1998, JacksonScienceSegregationRace2005, SchafferScientificRacismAgain2007}.
-As part of this segregationist reaction, MQ was created to provide a favorable venue for race scientists to publish their views, on the grounds that an "equalitarian dogma" created a censorious "taboo" against their research in mainstream publications \cite{TuckerFundingScientificRacism2002, JacksonMythicalTabooRace2020}.
+The Court's decision relied on expert testimony from psychologists and education researchers; but the segregationists also put forward their own experts, including Henry Garrett [@WinstonScienceServiceFar1998; @JacksonScienceSegregationRace2005; @SchafferScientificRacismAgain2007].
+As part of this segregationist reaction, MQ was created to provide a favorable venue for race scientists to publish their views, on the grounds that an "equalitarian dogma" created a censorious "taboo" against their research in mainstream publications [@TuckerFundingScientificRacism2002; @JacksonMythicalTabooRace2020].
 
@@
 We identified 16 researchers who had received funding from Pioneer; 14 of these researchers had profiles in the Web of Science (WoS) author search, allowing us identify 13 WoS-indexed journals that had published 6 or more of these authors.
-See Table \@ref(tab:researchers) and Figure \@ref(fig:wos).
+See Table @tbl-researchers and Figure @fig:wos.
 
@@
 
-Table: (\#tab:researchers) Pioneer-funded researchers.
-Either identified in \cite{MillerPioneerFundBankrolling1994} or named on an archive copy of Pioneer's website, along with birth and death dates from Wikipedia, attributed discipline, and WoS author search result counts.
+Table: Pioneer-funded researchers.
+Either identified in [@MillerPioneerFundBankrolling1994] or named on an archive copy of Pioneer's website, along with birth and death dates from Wikipedia, attributed discipline, and WoS author search result counts.
 Brunetto Chiarelli does not have a Wikipedia page.
+{#tbl-researchers}
 
@@
 
-Figure \@ref(fig:wos) shows that, while MQ is among the "Pioneer-publishing" journals, a number of mainstream journals are more prominent: *Personality and Individual Differences* (PID), *Intelligence* (Int), *Behavior Genetics* (BG), and *Psychological Reports* (PR).
+Figure @fig:wos shows that, while MQ is among the "Pioneer-publishing" journals, a number of mainstream journals are more prominent: *Personality and Individual Differences* (PID), *Intelligence* (Int), *Behavior Genetics* (BG), and *Psychological Reports* (PR).
 In addition, only psychologist Richard Lynn appears to have published heavily in MQ.
@@
 He has also been president of Pioneer since the death of psychologist J.
-Phillippe Rushton in 2012 \cite{BeirichPioneerFundAssets2013}.
+Phillippe Rushton in 2012 [@BeirichPioneerFundAssets2013].
 By contrast, a number of Pioneer-funded researchers have published in PID, Int, and to a lesser degree BG: Bouchard, Eysenck, Jensen, Rushton, Vernon, and also Lynn.
@@
 Richard Lynn joined the editorial board sometime between 1998 and 2002.
-(Archive copies of the Int editorial board page are not available from the journal's website from 1999 through 2001.)  Int has been criticized for including Lynn and Gerhard Meisenberg — who was editor-in-chief of MQ in 2015-18 — on its editorial board until 2018 \cite{SainiSuperiorReturnRace2019}.
+(Archive copies of the Int editorial board page are not available from the journal's website from 1999 through 2001.)  Int has been criticized for including Lynn and Gerhard Meisenberg — who was editor-in-chief of MQ in 2015-18 — on its editorial board until 2018 [@SainiSuperiorReturnRace2019}.
 Int is published by Elsevier.
 
-# Topic model analysis identifies race science discourse {.unnumbered}
+# Topic model analysis identifies race science discourse {.unnumbered]
 
@@
 
-![Silge plots \cite{SilgeTopicModeling2017} and smoothed time series for three focal topics.
+![Silge plots [@SilgeTopicModeling2017] and smoothed time series for three focal topics.
 Top row: top 15 terms by $\beta$ (term-topic distribution) for each topic.
@@
 
-Figure \@ref(fig:focal) and this discussion focuses on three topics identified in one of the 24 fitted topic models; see figures \@ref(fig:gamma-lg), \@ref(fig:gamma-md), \@ref(fig:gamma-sm) and supplemental visualizations S2-S4 for all topics from all models and S5-S7 for selected topics from all models.
+Figure @fig:focal and this discussion focuses on three topics identified in one of the 24 fitted topic models; see figures @fig:gamma-lg, @fig:gamma-md, @fig:gamma-sm and supplemental visualizations S2-S4 for all topics from all models and S5-S7 for selected topics from all models.
 In the focal model, topic 07 is strongly associated with MQ:  MQ published dozens of articles in this topic each year, and no other journal ever published more than a handful.
-The "Silge plot," showing the top 15 terms in the topic \cite{SilgeTopicModeling2017}, contains racial terms (`races`, `whites`, `negroes` and potentially `europe`, `africa`, `india`, and `japan`) as well as `book`, likely reflecting the fact that MQ published a number of book reviews, while the psychology journals either did not or these were not available for the corpus.
+The "Silge plot," showing the top 15 terms in the topic [@SilgeTopicModeling2017], contains racial terms (`races`, `whites`, `negroes` and potentially `europe`, `africa`, `india`, and `japan`) as well as `book`, likely reflecting the fact that MQ published a number of book reviews, while the psychology journals either did not or these were not available for the corpus.
 
 Topic 22 is strongly associated with Int and PID in the same way that 07 is associated with MQ.
-The Silge plot does contain `jensen`, as well as a reference to Raymond Cattell, who played a major role in the development of factor analysis and intelligence testing but also advocated for eugenics, fascism, and Nazi race science \cite{MehlerBeyondismRaymondCattell1997}.
+The Silge plot does contain `jensen`, as well as a reference to Raymond Cattell, who played a major role in the development of factor analysis and intelligence testing but also advocated for eugenics, fascism, and Nazi race science [@MehlerBeyondismRaymondCattell1997].
 However, the other authors named in this topic — John Horn and Peter Bentler — do not appear to have contributed to scientific racism.
@@
 In almost all years, most papers in topic 24 were published in mainstream journals rather than MQ.
-Jensen's "How Much Can We Boost IQ and Scholastic Achievement?" \cite{JensenHowMuchCan1969} was published in 1969 in *Harvard Educational Review* (not included in this study), and the time series indicates that, in the early 1970s, there was an increase in articles in topic 24 in both PR and MQ (the only two journals in our corpus that were active at the time).
-MQ shows another sharp increase in the late 1980s; using content analysis, Adams and Pilloud found that psychology was the dominant discipline in MQ in the period 1992-2018 \cite{AdamsMisAppropriationBiological2021}.
+Jensen's "How Much Can We Boost IQ and Scholastic Achievement?" [@JensenHowMuchCan1969] was published in 1969 in *Harvard Educational Review* (not included in this study), and the time series indicates that, in the early 1970s, there was an increase in articles in topic 24 in both PR and MQ (the only two journals in our corpus that were active at the time).
+MQ shows another sharp increase in the late 1980s; using content analysis, Adams and Pilloud found that psychology was the dominant discipline in MQ in the period 1992-2018 [@AdamsMisAppropriationBiological2021].
 PID published multiple papers in topic 24 almost immediately after it was founded, with Int showing a more gradual increase between the mid-1970s and mid-1990s.
@@
 Among both academics and the general public, the field of behavior genetics is strongly associated with race science, and specifically race-and-intelligence research.
-Panofsky \cite{PanofskyMisbehavingScienceControversy2014} argues that, prior to Jensen's 1969 paper, behavior genetics emphasized the study of non-human animals and intentionally avoided associations with eugenics and public controversy more generally.
-In response to Jensen, critics such as Lewontin offered broad critiques of behavior genetics as such, and behavior geneticists in turn adopted a radical conception of academic freedom (without any sense of responsibility for the social implications of academic research) and a siege or wartime mentality, as illustrated by Sandra Scarr's 1986 presidential address to the Behavior Genetics Association (BGA) \cite{ScarrThreeCheersBehavior1987}.
+Panofsky [@PanofskyMisbehavingScienceControversy2014] argues that, prior to Jensen's 1969 paper, behavior genetics emphasized the study of non-human animals and intentionally avoided associations with eugenics and public controversy more generally.
+In response to Jensen, critics such as Lewontin offered broad critiques of behavior genetics as such, and behavior geneticists in turn adopted a radical conception of academic freedom (without any sense of responsibility for the social implications of academic research) and a siege or wartime mentality, as illustrated by Sandra Scarr's 1986 presidential address to the Behavior Genetics Association (BGA) [@ScarrThreeCheersBehavior1987].
 Scarr's address coincided with the period between 1970-1990 when BG published articles in topic 24 (including Scarr's address itself).
@@
 Instead, our results suggest that other subdisciplines of psychology have provided the primary venue for mainstreaming race science.
-Over the past few decades, many professional organizations in genetics and anthropology have made formal statements rejecting race as a biologically meaningful concept \cite{anthropology1996aapa, american2018ashg, board1998aaa, committee2020asa, FuentesAAPAStatementRace2019, nhgri2023, rotimi2022ashg, wynshawboris2020ashg}, undermining the core assumption of race science.
-The American Psychological Association (APA) --- whose members are predominantly clinical psychologists --- and the Federation of Associations in Behavioral & Brain Sciences (FABBS) --- an organization whose purpose is to provide policy recommendations --- have openly rejected race as biologically meaningful \cite{american2021apology, baron2022fabbs}.
-But research-focused psychological organizations like the Association for Psychological Science (APS) and Psychonomic Society (PS) stop short of rejecting race as biologically meaningful when they denounce racism \cite{aps2021statement, board2020psychonomic}.
-Psychology as a discipline maintains space for several mythological race science narratives, such as allegations of a taboo against race and intelligence research \cite{jackson2021mythical}, claims that scientists engaging in or calling for any kind of anti-racism in the field is an ideological corruption of dispassionate and value-neutral science \cite{roberts_2022}, or arguments that holding race science to the same evidentiary standards as other psychological research is a violation of academic freedom \cite{herbert2023academic}.
+Over the past few decades, many professional organizations in genetics and anthropology have made formal statements rejecting race as a biologically meaningful concept [@anthropology1996aapa; @@american2018ashg, board1998aaa, committee2020asa; FuentesAAPAStatementRace2019; nhgri2023; rotimi2022ashg; @wynshawboris2020ashg], undermining the core assumption of race science.
+The American Psychological Association (APA) --- whose members are predominantly clinical psychologists --- and the Federation of Associations in Behavioral & Brain Sciences (FABBS) --- an organization whose purpose is to provide policy recommendations --- have openly rejected race as biologically meaningful [@american2021apology; @baron2022fabbs].
+But research-focused psychological organizations like the Association for Psychological Science (APS) and Psychonomic Society (PS) stop short of rejecting race as biologically meaningful when they denounce racism [@aps2021statement; @board2020psychonomic].
+Psychology as a discipline maintains space for several mythological race science narratives, such as allegations of a taboo against race and intelligence research [@jackson2021mythical], claims that scientists engaging in or calling for any kind of anti-racism in the field is an ideological corruption of dispassionate and value-neutral science [@roberts_2022], or arguments that holding race science to the same evidentiary standards as other psychological research is a violation of academic freedom [@herbert2023academic].
 And historically several race scientists have secured important gatekeeping positions within the scientific community, such as on the editorial boards of Int and PID.
@@
 During this same period of time, race scientists served in prominent and influential positions in behavior genetics and psychology, including as society presidents and members of journal editorial boards.
-In some cases, some race scientists simultaneously maintained active connections to both the mainstream scientific community and white supremacist organizations \cite{WinstonScientificRacismNorth2020, JacksonJr.ArthurJensenEvolutionary2022, SlobodianUnequalMindHow2023}.
+In some cases, some race scientists simultaneously maintained active connections to both the mainstream scientific community and white supremacist organizations [@WinstonScientificRacismNorth2020; @JacksonJr.ArthurJensenEvolutionary2022; @SlobodianUnequalMindHow2023].
 Except for the APA, we are not aware of other major organizations for scientific research in psychology that have issued apologies for the historical contributions of their field to scientific racism.
@@
 We are not identifying any particular scientist as a racist, scientific racist, or race scientist based on the topic model results alone.
-Such claims require an analysis of documentary evidence that goes beyond the scope of the current study \cite{JacksonJr.ArthurJensenEvolutionary2022}.
-We do assume that our readers, like us, regard white supremacy and scientific racism as morally odious and beyond the scope of reasonable debate \cite{SchroederLimitsDemocratizingScience2022}.
-However, scientific research can be appropriated to promote scientific racism — and thus count as race science — even when this is contrary to the intentions of the original researchers themselves \cite{TaberyWhyStudyingGenetics2015, GillbornSoftlySoftlyGenetics2016, CarlsonQuantifyingContextualizingImpact2020, HennWhyDNANo2021}.
+Such claims require an analysis of documentary evidence that goes beyond the scope of the current study [@JacksonJr.ArthurJensenEvolutionary2022].
+We do assume that our readers, like us, regard white supremacy and scientific racism as morally odious and beyond the scope of reasonable debate [@SchroederLimitsDemocratizingScience2022].
+However, scientific research can be appropriated to promote scientific racism — and thus count as race science — even when this is contrary to the intentions of the original researchers themselves [@TaberyWhyStudyingGenetics2015; @GillbornSoftlySoftlyGenetics2016; CarlsonQuantifyingContextualizingImpact2020; @HennWhyDNANo2021].
 Panofsky et al.
-show how the term "human biodiversity," originally developed for anti-racist purposes by biological anthropologist Jonathan Marks, has been inverted by white supremacists and is now a dogwhistle for biological racial hierarchy \cite{PanofskyHowWhiteNationalists2021}.
+show how the term "human biodiversity," originally developed for anti-racist purposes by biological anthropologist Jonathan Marks, has been inverted by white supremacists and is now a dogwhistle for biological racial hierarchy [@PanofskyHowWhiteNationalists2021].
 
-Research on the genetics of intelligence has been socially harmful, not just in the somewhat abstract sense of promoting racial stigma, but in the concrete sense of being used to rationalize mass shootings and other acts of racial violence \cite{PronczukRacistResearcherExposed2022, MeyerWrestlingSocialBehavioral2023}.
-While scientists do not fully control the downstream social effects of their research, like other citizens scientists are responsible for mitigating reasonably foreseeable harms that result from their actions \cite{BlockIQHeritabilityInequality1974a, DouglasSciencePolicyValuefree2009, KitcherArgumentFreeInquiry1997, KouranyShouldKnowledgeBe2016}.
+Research on the genetics of intelligence has been socially harmful, not just in the somewhat abstract sense of promoting racial stigma, but in the concrete sense of being used to rationalize mass shootings and other acts of racial violence [@PronczukRacistResearcherExposed2022; @MeyerWrestlingSocialBehavioral2023].
+While scientists do not fully control the downstream social effects of their research, like other citizens scientists are responsible for mitigating reasonably foreseeable harms that result from their actions [@BlockIQHeritabilityInequality1974a; @DouglasSciencePolicyValuefree2009; KitcherArgumentFreeInquiry1997; @KouranyShouldKnowledgeBe2016].
 These actions include not only individual decisions to research this topic or that, but also collective decisions about who is awarded positions of power or influence.
 Given the way research on intelligence and behavior genetics has been used historically, these fields may be especially susceptible to appropriation by scientific racists.
-This only strengthens the obligations of researchers in these areas to prevent their research from causing harm \cite{CarlsonQuantifyingContextualizingImpact2020, OgbunugaforDNABasketballBirthday2022}.
+This only strengthens the obligations of researchers in these areas to prevent their research from causing harm [@CarlsonQuantifyingContextualizingImpact2020; @OgbunugaforDNABasketballBirthday2022].
 
 How should psychologists exercise these responsibilities?
-We believe restorative or reparative justice provides a useful model \cite{WenzelRetributiveRestorativeJustice2008, WormerRestorativeJustice2013}.
+We believe restorative or reparative justice provides a useful model [@WenzelRetributiveRestorativeJustice2008; @WormerRestorativeJustice2013].
 Unlike retributive justice (used by almost all criminal justice systems today), the aim of restorative justice is to repair the damaged relationships between victims and perpetrators of injustice.
 Typically, restorative justice requires perpetrators to not only acknowledge their actions and the harms that these actions had on victims, but also to work with victims to identify concrete actions that perpetrators can take to redress or mitigate these harms.
-From this perspective, explicit apologies for contributions to scientific racism are essential as a first step \cite{FuentesAAPAStatementRace2019, american2021apology, JacksonFacingOurHistory2023}, but not fully sufficient on their own.
+From this perspective, explicit apologies for contributions to scientific racism are essential as a first step [@FuentesAAPAStatementRace2019; @american2021apology; @JacksonFacingOurHistory2023], but not fully sufficient on their own.
 
@@
 
-Corpus assembly is summarized in figure \@ref(fig:corpus).  The corpus was assembled in two parts, one for *Mankind Quarterly* and the other for mainstream journals.
+Corpus assembly is summarized in figure @fig:corpus.  The corpus was assembled in two parts, one for *Mankind Quarterly* and the other for mainstream journals.
 
@@
 To identify suitable mainstream journals for inclusion in the corpus, we first identified academic researchers who had been funded by the Pioneer Fund.
-Pioneer is an American non-profit organization founded in 1937 to fund eugenics research and propaganda \cite{TuckerFundingScientificRacism2002}.
+Pioneer is an American non-profit organization founded in 1937 to fund eugenics research and propaganda [@TuckerFundingScientificRacism2002].
 After the *Brown v.
-Board of Education* ruling in 1954, Pioneer funded various segregationist efforts across the United States, including lectures on eugenics by Stanford physicist William Shockley \cite{JacksonScienceSegregationRace2005}.
-We reviewed a critical profile of Pioneer \cite{MillerPioneerFundBankrolling1994} as well as an archived page from the organization's own web site (<https://web.archive.org/web/20130103005545/http://www.pioneerfund.org/Grantees.html>), which together listed 16 researchers who had received Pioneer funds.
+Board of Education* ruling in 1954, Pioneer funded various segregationist efforts across the United States, including lectures on eugenics by Stanford physicist William Shockley [@JacksonScienceSegregationRace2005}.
+We reviewed a critical profile of Pioneer [MillerPioneerFundBankrolling1994] as well as an archived page from the organization's own web site (<https://web.archive.org/web/20130103005545/http://www.pioneerfund.org/Grantees.html>), which together listed 16 researchers who had received Pioneer funds.
 
@@
 The remaining 5 journals are all published by major academic publishers — Elsevier, Springer, or Cambridge University Press — and each item in the entire run of each journal has been assigned a DOI (digital object identifier) for archival purposes.
-We used the Crossref API and `rcrossref` R interface to this API \cite[version 1.1.0.99]{ChamberlainRcrossrefClientVarious2019} to retrieve metadata for each item published from 1960-2010 in each of these journals.
+We used the Crossref API and `rcrossref` R interface to this API \cite[version 1.1.0.99]{ChamberlainRcrossrefClientVarious2019] to retrieve metadata for each item published from 1960-2010 in each of these journals.
 These metadata included item-level license information — confirming that the text of each item could be used for text mining projects — and a URL to an electronic version of the item.
@@
 
-After XML/PDF versions were retrieved, text was extracted using either the `xml2` package \cite{WickhamXml2ParseXML2023} or `pdftools`.
-Figure \@ref(fig:counts) and table \@ref(tab:counts) show the number of fulltext documents included in the corpus.
+After XML/PDF versions were retrieved, text was extracted using either the `xml2` package [@WickhamXml2ParseXML2023] or `pdftools`.
+Figure @fig:counts and table @tbl-counts show the number of fulltext documents included in the corpus.
 All together, the corpus includes 34,896 documents from 6 journals.
@@
 
-Table: (\#tab:counts) Document counts, by journal, and years included in the corpus.
+Table: (\#tbl-counts) Document counts, by journal, and years included in the corpus.
 
@@
 We then counted the occurrence of each noun phrase in the document.
-The aggregated "document-term matrix" was stored in Parquet format \cite{VohraApacheParquet2016} for performance reasons, and written and read using the `arrow` package for R \cite[version 7.0.0]{RichardsonArrowIntegrationApache2023}.
+The aggregated "document-term matrix" was stored in Parquet format [@VohraApacheParquet2016] for performance reasons, and written and read using the `arrow` package for R \cite[version 7.0.0]{RichardsonArrowIntegrationApache2023}.
 
@@
 Phrases can be more informative than single terms, for example, distinguishing "intelligence test" from "hypothesis test."  While simple n-gram extraction will include numerous phrases that are not especially meaningful.
-Consider the sentence "Since our first analyses of feeding patterns in rats, we had been using a criterion of 40 minutes (Le Magnen & Tallon 1963; 1966)"\cite{SclafaniCorrelationCausationStudy1981}.
+Consider the sentence "Since our first analyses of feeding patterns in rats, we had been using a criterion of 40 minutes (Le Magnen & Tallon 1963; 1966)"[@SclafaniCorrelationCausationStudy1981].
 Bigrams such as "since our" and "criterion 40" will likely be discarded in vocabulary selection, but significantly increase the computational cost of vocabulary selection.
@@
 
-We took an information-theoretic approach to vocabulary selection\cite{HicksProductivityInterdisciplinaryImpacts2021}.
+We took an information-theoretic approach to vocabulary selection[@HicksProductivityInterdisciplinaryImpacts2021].
 Consider a game in which I draw a document from the corpus, then a single token from that document.
@@
 
-To fit topic models, we followed the approach proposed by Rohe and Zeng \cite{RoheVintageFactorAnalysis2020}, which uses varimax-rotated partial principal components instead of the variational inference methods used by standard topic model packages such as `stm` \cite{RobertsStmPackageStructural2019}.
-This novel approach was implemented in the R package `tmfast` \cite{HicksTmfastFitsTopic2023}.
-We conducted a simulation study of `tmfast`, which found that it was significantly faster and only slightly less accurate at reconstructing known word-topic and topic-document distributions, compared to `stm` \cite{HicksTmfastFitsTopic2023}.
+To fit topic models, we followed the approach proposed by Rohe and Zeng [@RoheVintageFactorAnalysis2020], which uses varimax-rotated partial principal components instead of the variational inference methods used by standard topic model packages such as `stm` [@RobertsStmPackageStructural2019}.
+This novel approach was implemented in the R package `tmfast` [HicksTmfastFitsTopic2023].
+We conducted a simulation study of `tmfast`, which found that it was significantly faster and only slightly less accurate at reconstructing known word-topic and topic-document distributions, compared to `stm` [@HicksTmfastFitsTopic2023].
 
@@
 
-Following the approach of \cite{HicksProductivityInterdisciplinaryImpacts2021}, we did not attempt to identify a unique best fitted model for further analysis.
+Following the approach of [@HicksProductivityInterdisciplinaryImpacts2021], we did not attempt to identify a unique best fitted model for further analysis.
 While the manuscript text focuses on the medium vocabulary, $k=40$ model as the "median" among the 24 fitted models, we compare and contrast findings from this model with those from the other models.
@@
 After fitting topic models, our first research question was whether we could identify distinctive "race science discourse" topics.
-Figures \@ref(fig:gamma-lg), \@ref(fig:gamma-md), \@ref(fig:gamma-sm) show the gamma (topic-document) distributions for each value of $k$ for the three vocabularies.
+Figures @fig:gamma-lg, @fig:gamma-md, @fig:gamma-sm show the gamma (topic-document) distributions for each value of $k$ for the three vocabularies.
 In these figures, each panel corresponds to a single journal-$k$ combination, each row of cells is a single document in that journal, and each column of cells is a single topic in that particular model.
@@
 
-![Fitted topic models: Gamma (topic-document) distributions, large vocabulary. (\#fig:gamma-lg)](img/05_lg_gamma.png){width="6in" height="4.8in"}
+![Fitted topic models: Gamma (topic-document) distributions, large vocabulary. (\#fig:gamma-lg)](img/05_lg_gamma.png){width="6in" height="4.8in"]
 
@@
 
-We next constructed Silge plots \cite{SilgeTopicModeling2017}, showing the top 15 (highest-probability) terms (noun phrases) from each topic in the model.
+We next constructed Silge plots [@SilgeTopicModeling2017], showing the top 15 (highest-probability) terms (noun phrases) from each topic in the model.
 These are provided as supplemental PDFs S2-S4.
@@
 One is a table of contents from a 1980 issue of Int; arguably this document should have been excluded as irrelevant front matter, but arguably it legitimizes the research published in the issue.
-The second case is a paper on racial differences in emotional intelligence that was classified as race science discourse by one author, while the other author felt it was more focused on psychometrics \cite{GignacGroupDifferencesEI2010}.
+The second case is a paper on racial differences in emotional intelligence that was classified as race science discourse by one author, while the other author felt it was more focused on psychometrics [@GignacGroupDifferencesEI2010].
 And a 1990 article in MQ (<https://www.unz.com/print/MankindQuarterly-1990q3-00108/>) proposed to offer an economic explanation for crime.
@@
 Fourteen of the remaining documents had been published since 2010 (after the scope of the primary study).
-Among these, one obituary and two papers celebrated the work of John Loehlin, a prominent race-and-intelligence researcher who had been Director of the American Eugenics Society from 1968-1972 \cite{JohnLoehlin19262020, WaldmanIntroductionFestschriftJohn2014, PlominGenotypeEnvironmentCorrelationEra2014}.
+Among these, one obituary and two papers celebrated the work of John Loehlin, a prominent race-and-intelligence researcher who had been Director of the American Eugenics Society from 1968-1972 [@JohnLoehlin19262020; @WaldmanIntroductionFestschriftJohn2014; @PlominGenotypeEnvironmentCorrelationEra2014].
 Another was a retrospective of the work of Lindon Eaves, a geneticist who made at least one notable contribution to debates on the Scarr-Rowe hypothesis (interactions between heritability, race, and class).
@@
 Twenty-eight documents in this sample from BG were published between 1990 and 2010.
-One was an obituary of Jerry Hirsch, a critic of hereditarianism in general and Jensen in particular \cite{RoubertouxJerryHirsch202008}, and another was an obituary of David Rowe \cite{RodgersObituaryDavidChristian2003}.
-Only 2 of these 28 documents reported race differences of any kind: one examining interactions among race, sex, and heritability for adolescent BMI (Body Mass Index, used as a measure of overweight/obesity) \cite{JacobsonGeneticSharedEnvironmental1998}; and the other interactions among race, family history of alcoholism, and visuospatial performance \cite{BermanReducedVisuospatialPerformance1995}.
+One was an obituary of Jerry Hirsch, a critic of hereditarianism in general and Jensen in particular [@RoubertouxJerryHirsch202008], and another was an obituary of David Rowe [@RodgersObituaryDavidChristian2003].
+Only 2 of these 28 documents reported race differences of any kind: one examining interactions among race, sex, and heritability for adolescent BMI (Body Mass Index, used as a measure of overweight/obesity) [@JacobsonGeneticSharedEnvironmental1998]; and the other interactions among race, family history of alcoholism, and visuospatial performance [@BermanReducedVisuospatialPerformance1995].
 Finally, Philip Vernon — one of the Pioneer-funded scientists we identified earlier — published a critical review of a book that attempted to address hereditarianism, the Flynn effect, race-and-intelligence research, and some related issues.
 
-At the same time, we were unable to find any articles published in BG that acknowledged the involvement of the field with scientific racism with anything like the force of the statements that have been made by biological anthropologists \cite{FuentesAAPAStatementRace2019} and human geneticists \cite{JacksonFacingOurHistory2023}.
-Notably, the ASHG report includes discussions of behavior genetics race science, and in particular is critical of the field and the organization for failing to publicly reject the race-and-intelligence claims made by Shockley and Jensen \cite{JacksonFacingOurHistory2023}.
+At the same time, we were unable to find any articles published in BG that acknowledged the involvement of the field with scientific racism with anything like the force of the statements that have been made by biological anthropologists [@FuentesAAPAStatementRace2019} and human geneticists \cite{JacksonFacingOurHistory2023].
+Notably, the ASHG report includes discussions of behavior genetics race science, and in particular is critical of the field and the organization for failing to publicly reject the race-and-intelligence claims made by Shockley and Jensen [@JacksonFacingOurHistory2023].
 A search for "eugenics" in BG found articles that presented the issue as part of the distant past, and a search for "racism" turned up studies of racist attitudes.
@@
 A key limitation of the current study stems from the inability of text mining techniques to draw on a broader cultural context than what is represented in the corpus.
-Scholars of racism and US race relations have noted a shift in the way racist attitudes are understood and expressed, from overt and direct to a subtle "color-blind" racism \cite{Bonilla-SilvaRacismRacistsColorBlind2006}.
+Scholars of racism and US race relations have noted a shift in the way racist attitudes are understood and expressed, from overt and direct to a subtle "color-blind" racism [@Bonilla-SilvaRacismRacistsColorBlind2006].
 Color-blind racism represents a change in the primary mechanisms by which white privilege is maintained, using non-racialized language in furtherance of the racial status quo both institutionally and individually, and increasingly reliant on the invisibility of socio-cultural mechanisms to preserve inequalities.
-Color-blind racism involves the use of racialized "codes" or "dogwhistles," language that is superficially non-racial but carries racial implications, such as (in the US context) "violent inner city criminals" or "welfare queen" \cite{SaulDogwhistlesPoliticalManipulation2018}.
-Alderfer notes that the racist implications of *The Bell Curve* only emerge gradually, with the opening of the book using the color-blind language of "group differences" \cite{AlderferScienceNonsciencePsychologists2003}.
-These cases suggest further that color-blindness can facilitate white supremacist appropriation of non-racial research — even contrary to the researchers' own intentions — as scientific racist readers re-interpret racial-neutral language as dogwhistles \cite{GillbornSoftlySoftlyGenetics2016, WillsAreClustersRaces2017}.
+Color-blind racism involves the use of racialized "codes" or "dogwhistles," language that is superficially non-racial but carries racial implications, such as (in the US context) "violent inner city criminals" or "welfare queen" [@SaulDogwhistlesPoliticalManipulation2018].
+Alderfer notes that the racist implications of *The Bell Curve* only emerge gradually, with the opening of the book using the color-blind language of "group differences" [@AlderferScienceNonsciencePsychologists2003].
+These cases suggest further that color-blindness can facilitate white supremacist appropriation of non-racial research — even contrary to the researchers' own intentions — as scientific racist readers re-interpret racial-neutral language as dogwhistles [@GillbornSoftlySoftlyGenetics2016; @WillsAreClustersRaces2017].
 
@@
 But scientific journal articles on these topics may not include any racial language at all.
-For example, a meta-analysis of the Flynn effect does not discuss racial differences or use terms such as "White" or "Black" outside the reference list \cite{TrahanFlynnEffectMetaanalysis2014}.
+For example, a meta-analysis of the Flynn effect does not discuss racial differences or use terms such as "White" or "Black" outside the reference list [@TrahanFlynnEffectMetaanalysis2014].
 Should such articles be coded as race science discourse?

### 2023-10-27 paper/paper.qmd
crossrefs

@@
 We identified 16 researchers who had received funding from Pioneer; 14 of these researchers had profiles in the Web of Science (WoS) author search, allowing us identify 13 WoS-indexed journals that had published 6 or more of these authors.
-See Table @tbl-researchers and Figure @fig:wos.
+See @tbl-researchers and @fig-wos.
 
@@
 
-![Journals publishing 6 or more Pioneer-funded researchers, Web of Science author search results. (\#fig:wos)](img/wos_results.png){width="4.76in" height="2.6in"}
+![Journals publishing 6 or more Pioneer-funded researchers, Web of Science author search results.](img/wos_results.png){#fig-wos width="4.76in" height="2.6in"}
 
-Figure @fig:wos shows that, while MQ is among the "Pioneer-publishing" journals, a number of mainstream journals are more prominent: *Personality and Individual Differences* (PID), *Intelligence* (Int), *Behavior Genetics* (BG), and *Psychological Reports* (PR).
+@fig-wos shows that, while MQ is among the "Pioneer-publishing" journals, a number of mainstream journals are more prominent: *Personality and Individual Differences* (PID), *Intelligence* (Int), *Behavior Genetics* (BG), and *Psychological Reports* (PR).
 In addition, only psychologist Richard Lynn appears to have published heavily in MQ.
@@
 Richard Lynn joined the editorial board sometime between 1998 and 2002.
-(Archive copies of the Int editorial board page are not available from the journal's website from 1999 through 2001.)  Int has been criticized for including Lynn and Gerhard Meisenberg — who was editor-in-chief of MQ in 2015-18 — on its editorial board until 2018 [@SainiSuperiorReturnRace2019}.
+(Archive copies of the Int editorial board page are not available from the journal's website from 1999 through 2001.)  Int has been criticized for including Lynn and Gerhard Meisenberg — who was editor-in-chief of MQ in 2015-18 — on its editorial board until 2018 [@SainiSuperiorReturnRace2019].
 Int is published by Elsevier.
 
-# Topic model analysis identifies race science discourse {.unnumbered]
+# Topic model analysis identifies race science discourse {.unnumbered}
 
@@
 Thin lines give annual values, thick lines give 5-year running averages.
-Because *Behavior Genetics* and *Behavioral and Brain Sciences* are not prominent in any panel, neither is given a direct label.
-(\#fig:focal)](img/focal_topics.png){width="4.76in" height="3.57in"}
+Because *Behavior Genetics* and *Behavioral and Brain Sciences* are not prominent in any panel, neither is given a direct label.](img/focal_topics.png){#fig-focal width="4.76in" height="3.57in"}
 
-Figure @fig:focal and this discussion focuses on three topics identified in one of the 24 fitted topic models; see figures @fig:gamma-lg, @fig:gamma-md, @fig:gamma-sm and supplemental visualizations S2-S4 for all topics from all models and S5-S7 for selected topics from all models.
+@fig-focal and this discussion focuses on three topics identified in one of the 24 fitted topic models; see figures @fig-gamma-lg, @fig-gamma-md, @fig-gamma-sm and supplemental visualizations S2-S4 for all topics from all models and S5-S7 for selected topics from all models.
 In the focal model, topic 07 is strongly associated with MQ:  MQ published dozens of articles in this topic each year, and no other journal ever published more than a handful.
@@
 Instead, our results suggest that other subdisciplines of psychology have provided the primary venue for mainstreaming race science.
-Over the past few decades, many professional organizations in genetics and anthropology have made formal statements rejecting race as a biologically meaningful concept [@anthropology1996aapa; @@american2018ashg, board1998aaa, committee2020asa; FuentesAAPAStatementRace2019; nhgri2023; rotimi2022ashg; @wynshawboris2020ashg], undermining the core assumption of race science.
+Over the past few decades, many professional organizations in genetics and anthropology have made formal statements rejecting race as a biologically meaningful concept [@anthropology1996aapa; @american2018ashg; @board1998aaa; @committee2020asa; @FuentesAAPAStatementRace2019; @nhgri2023; @rotimi2022ashg; @wynshawboris2020ashg], undermining the core assumption of race science.
 The American Psychological Association (APA) --- whose members are predominantly clinical psychologists --- and the Federation of Associations in Behavioral & Brain Sciences (FABBS) --- an organization whose purpose is to provide policy recommendations --- have openly rejected race as biologically meaningful [@american2021apology; @baron2022fabbs].
@@
 We do assume that our readers, like us, regard white supremacy and scientific racism as morally odious and beyond the scope of reasonable debate [@SchroederLimitsDemocratizingScience2022].
-However, scientific research can be appropriated to promote scientific racism — and thus count as race science — even when this is contrary to the intentions of the original researchers themselves [@TaberyWhyStudyingGenetics2015; @GillbornSoftlySoftlyGenetics2016; CarlsonQuantifyingContextualizingImpact2020; @HennWhyDNANo2021].
+However, scientific research can be appropriated to promote scientific racism — and thus count as race science — even when this is contrary to the intentions of the original researchers themselves [@TaberyWhyStudyingGenetics2015; @GillbornSoftlySoftlyGenetics2016; @CarlsonQuantifyingContextualizingImpact2020; @HennWhyDNANo2021].
 Panofsky et al.
@@
 Research on the genetics of intelligence has been socially harmful, not just in the somewhat abstract sense of promoting racial stigma, but in the concrete sense of being used to rationalize mass shootings and other acts of racial violence [@PronczukRacistResearcherExposed2022; @MeyerWrestlingSocialBehavioral2023].
-While scientists do not fully control the downstream social effects of their research, like other citizens scientists are responsible for mitigating reasonably foreseeable harms that result from their actions [@BlockIQHeritabilityInequality1974a; @DouglasSciencePolicyValuefree2009; KitcherArgumentFreeInquiry1997; @KouranyShouldKnowledgeBe2016].
+While scientists do not fully control the downstream social effects of their research, like other citizens scientists are responsible for mitigating reasonably foreseeable harms that result from their actions [@BlockIQHeritabilityInequality1974a; @DouglasSciencePolicyValuefree2009; @KitcherArgumentFreeInquiry1997; @KouranyShouldKnowledgeBe2016].
 These actions include not only individual decisions to research this topic or that, but also collective decisions about who is awarded positions of power or influence.
@@
 From this perspective, explicit apologies for contributions to scientific racism are essential as a first step [@FuentesAAPAStatementRace2019; @american2021apology; @JacksonFacingOurHistory2023], but not fully sufficient on their own.
-
-\clearpage
-
-# References and Notes {.unnumbered}
-
-\bibliographystyle{Science}
-\bibliography{references.bib, refs-manual.bib}
-
-| 
-| Thanks to Emily Merchant and John Jackson for some initial discussions that helped clarify the scope of the project and identify some key resources related to the Pioneer Fund and *Mankind Quarterly*.
-Thanks to Anthony Sainez for help retrieving the *Mankind Quarterly* articles.
-Thanks to Derek Devnich and James Dooley for their work attempting to secure the APA-published articles.
-Thanks to Kevin Bird, John Jackson, and Manuela Fernández Pinto for comments on a draft of this paper.
-| **Funding:** EJCL's work on this project was partially supported by UC Merced.  DJH's work on this project was not supported by any specific funding. 
-| **Author contributions:** DJH (conceptualization, data curation, formal analysis, investigation, methodology, project administration, software, validation, visualization, writing-original draft, writing-review & editing); EJCL (conceptualization, methodology, software, validation, visualization, writing-original draft, writing-review & editing)
-| **Competing interests:** The authors have no competing interests to declare. 
-| **Data and materials availability:** Web of Science search results and article fulltext are not publicly available due to intellectual property restrictions.
-Preprocessed analysis data (document-term counts and article metadata) and analysis scripts are available at *[doi]*.
 
@@
 
+| 
 | S1. Materials and methods
@@
 
-![Corpus assembly. (\#fig:corpus)](img/dataset.png){width="8in" height="6.1in"}
+![Corpus assembly.](img/dataset.png){#fig-corpus width="8in" height="6.1in"}
 
-Corpus assembly is summarized in figure @fig:corpus.  The corpus was assembled in two parts, one for *Mankind Quarterly* and the other for mainstream journals.
+Corpus assembly is summarized in @fig-corpus.  The corpus was assembled in two parts, one for *Mankind Quarterly* and the other for mainstream journals.
 
@@
 After XML/PDF versions were retrieved, text was extracted using either the `xml2` package [@WickhamXml2ParseXML2023] or `pdftools`.
-Figure @fig:counts and table @tbl-counts show the number of fulltext documents included in the corpus.
+@fig-counts and table @tbl-counts show the number of fulltext documents included in the corpus.
 All together, the corpus includes 34,896 documents from 6 journals.
 
-![Count of documents in the corpus, by journal and year. (\#fig:counts)](img/02_count.png){width="4.76in" height="3.6in"}
+![Count of documents in the corpus, by journal and year.](img/02_count.png){#fig-counts width="4.76in" height="3.6in"}
 
-Table: (\#tbl-counts) Document counts, by journal, and years included in the corpus.
+Table: Document counts, by journal, and years included in the corpus. {#tbl-counts}
 
@@
 
-To fit topic models, we followed the approach proposed by Rohe and Zeng [@RoheVintageFactorAnalysis2020], which uses varimax-rotated partial principal components instead of the variational inference methods used by standard topic model packages such as `stm` [@RobertsStmPackageStructural2019}.
-This novel approach was implemented in the R package `tmfast` [HicksTmfastFitsTopic2023].
+To fit topic models, we followed the approach proposed by Rohe and Zeng [@RoheVintageFactorAnalysis2020], which uses varimax-rotated partial principal components instead of the variational inference methods used by standard topic model packages such as `stm` [@RobertsStmPackageStructural2019].
+This novel approach was implemented in the R package `tmfast` [@HicksTmfastFitsTopic2023].
 We conducted a simulation study of `tmfast`, which found that it was significantly faster and only slightly less accurate at reconstructing known word-topic and topic-document distributions, compared to `stm` [@HicksTmfastFitsTopic2023].
@@
 After fitting topic models, our first research question was whether we could identify distinctive "race science discourse" topics.
-Figures @fig:gamma-lg, @fig:gamma-md, @fig:gamma-sm show the gamma (topic-document) distributions for each value of $k$ for the three vocabularies.
+[@fig-gamma-lg; @fig-gamma-md; and @fig-gamma-sm] show the gamma (topic-document) distributions for each value of $k$ for the three vocabularies.
 In these figures, each panel corresponds to a single journal-$k$ combination, each row of cells is a single document in that journal, and each column of cells is a single topic in that particular model.
@@
 
-![Fitted topic models: Gamma (topic-document) distributions, large vocabulary. (\#fig:gamma-lg)](img/05_lg_gamma.png){width="6in" height="4.8in"]
+![Fitted topic models: Gamma (topic-document) distributions, large vocabulary.](img/05_lg_gamma.png){#fig-gamma-lg width="6in" height="4.8in"}
 
-![Fitted topic models: Gamma (topic-document) distributions, medium vocabulary. (\#fig:gamma-md)](img/05_md_gamma.png){width="6in" height="4.8in"}
+![Fitted topic models: Gamma (topic-document) distributions, medium vocabulary.](img/05_md_gamma.png){#fig-gamma-md width="6in" height="4.8in"}
 
-![Fitted topic models: Gamma (topic-document) distributions, small vocabulary. (\#fig:gamma-sm)](img/05_sm_gamma.png){width="6in" height="4.8in"}
+![Fitted topic models: Gamma (topic-document) distributions, small vocabulary.](img/05_sm_gamma.png){#fig-gamma-sm width="6in" height="4.8in"}
 
@@
 
-<!-- 
-\clearpage
-
-# References and Notes {.unnumbered}
-
-\bibliographystyle{Science}
-\bibliography{references.bib, refs-manual.bib}
-
- -->
+# References

### 2023-10-27 paper/paper.qmd
small conversion errors

@@
     latex_engine: "xelatex"
+    number-sections: false
 
@@
 
-# Introduction {.unnumbered}
+<!-- # Introduction {.unnumbered} -->
 
@@
 
-# Scientific racism, eugenics, and *Brown* {.unnumbered}
+<!-- # Scientific racism, eugenics, and *Brown* {.unnumbered} -->
 
@@
 
-# *Mankind Quarterly* and Pioneer-funded researchers {.unnumbered}
+# Results
+
+## *Mankind Quarterly* and Pioneer-funded researchers {.unnumbered}
 
@@
 
-# Topic model analysis identifies race science discourse {.unnumbered}
+## Topic model analysis identifies race science discourse {.unnumbered}
 
@@
 
-# Race science, behavior genetics, and psychology {.unnumbered}
+## Race science, behavior genetics, and psychology {.unnumbered}
 
@@
 \clearpage
+ -->
 
@@
 
-![Corpus assembly.](img/dataset.png){#fig-corpus width="8in" height="6.1in"}
+![Corpus assembly.](img/dataset.png){#fig-corpus width="120%"}
 
@@
 After identifying some gaps (missing issues) in the archive from *The Unz Review*, we manually retrieved PDFs for further articles using ProQuest.
-Text was extracted using the R package `pdftools` \cite[version 3.0.1]{OomsPdftoolsTextExtraction2023}.
+Text was extracted using the R package `pdftools` [@OomsPdftoolsTextExtraction2023 version 3.0.1].
 This stage of corpus assembly was conducted in Fall 2021 with the assistance of Anthony Sainez.
@@
 After the *Brown v.
-Board of Education* ruling in 1954, Pioneer funded various segregationist efforts across the United States, including lectures on eugenics by Stanford physicist William Shockley [@JacksonScienceSegregationRace2005}.
-We reviewed a critical profile of Pioneer [MillerPioneerFundBankrolling1994] as well as an archived page from the organization's own web site (<https://web.archive.org/web/20130103005545/http://www.pioneerfund.org/Grantees.html>), which together listed 16 researchers who had received Pioneer funds.
+Board of Education* ruling in 1954, Pioneer funded various segregationist efforts across the United States, including lectures on eugenics by Stanford physicist William Shockley [@JacksonScienceSegregationRace2005].
+We reviewed a critical profile of Pioneer [@MillerPioneerFundBankrolling1994] as well as an archived page from the organization's own web site (<https://web.archive.org/web/20130103005545/http://www.pioneerfund.org/Grantees.html>), which together listed 16 researchers who had received Pioneer funds.
 
@@
 The remaining 5 journals are all published by major academic publishers — Elsevier, Springer, or Cambridge University Press — and each item in the entire run of each journal has been assigned a DOI (digital object identifier) for archival purposes.
-We used the Crossref API and `rcrossref` R interface to this API \cite[version 1.1.0.99]{ChamberlainRcrossrefClientVarious2019] to retrieve metadata for each item published from 1960-2010 in each of these journals.
+We used the Crossref API and `rcrossref` R interface to this API [@ChamberlainRcrossrefClientVarious2019 version 1.1.0.99] to retrieve metadata for each item published from 1960-2010 in each of these journals.
 These metadata included item-level license information — confirming that the text of each item could be used for text mining projects — and a URL to an electronic version of the item.
@@
 (This is unfortunate, as it was not difficult to find highly relevant documents that we could not automatically retrieve and therefore had to be excluded from our corpus.
-One set of examples is an exchange between Rushton and Flynn on the Flynn effect \cite{FlynnEvidenceRushtonGenetic1998,RushtonSecularGainsIQ1998,FlynnReplyRushtonGang1998}.)  Instead we used Elsevier's ScienceDirect API (<https://dev.elsevier.com/>) to independently search and retrieve all available items from *Personality and Individual Differences*.
+One set of examples is an exchange between Rushton and Flynn on the Flynn effect [@FlynnEvidenceRushtonGenetic1998; @RushtonSecularGainsIQ1998; @FlynnReplyRushtonGang1998].)  Instead we used Elsevier's ScienceDirect API (<https://dev.elsevier.com/>) to independently search and retrieve all available items from *Personality and Individual Differences*.
 This stage of corpus assembly was conducted between 2021-11 and 2022-05.
@@
 After XML/PDF versions were retrieved, text was extracted using either the `xml2` package [@WickhamXml2ParseXML2023] or `pdftools`.
-@fig-counts and table @tbl-counts show the number of fulltext documents included in the corpus.
+@fig-counts and @tbl-counts show the number of fulltext documents included in the corpus.
 All together, the corpus includes 34,896 documents from 6 journals.
@@
 
-After document retrieval and text extraction, we pre-processed the text using the `spaCy` NLP (natural language processing) Python library \cite[version 1.9.0]{SpaCyIndustrialstrengthNatural2018} and the R API `spacyr` \cite[version 1.2.1]{BenoitSpacyrWrapperSpaCy2020}.
+After document retrieval and text extraction, we pre-processed the text using the `spaCy` NLP (natural language processing) Python library [@SpaCyIndustrialstrengthNatural2018 version 1.9.0] and the R API `spacyr` [@BenoitSpacyrWrapperSpaCy2020 version 1.2.1].
 Specifically, we applied regular expressions to remove header/footer copyright notices and hyphenation, used spaCy to annotate and extract noun phrases (eg, "the intelligence test items"), and then cleaned and standardized these phrases (eg, removing the/an/a, coverting all text to lowercase, and replacing all whitespace with underscores: "intelligence_test_items").
 We then counted the occurrence of each noun phrase in the document.
-The aggregated "document-term matrix" was stored in Parquet format [@VohraApacheParquet2016] for performance reasons, and written and read using the `arrow` package for R \cite[version 7.0.0]{RichardsonArrowIntegrationApache2023}.
+The aggregated "document-term matrix" was stored in Parquet format [@VohraApacheParquet2016] for performance reasons, and written and read using the `arrow` package for R [@RichardsonArrowIntegrationApache2023 version 7.0.0].
 
@@
 
-At the same time, we were unable to find any articles published in BG that acknowledged the involvement of the field with scientific racism with anything like the force of the statements that have been made by biological anthropologists [@FuentesAAPAStatementRace2019} and human geneticists \cite{JacksonFacingOurHistory2023].
+At the same time, we were unable to find any articles published in BG that acknowledged the involvement of the field with scientific racism with anything like the force of the statements that have been made by biological anthropologists [@FuentesAAPAStatementRace2019] and human geneticists [@JacksonFacingOurHistory2023].
 Notably, the ASHG report includes discussions of behavior genetics race science, and in particular is critical of the field and the organization for failing to publicly reject the race-and-intelligence claims made by Shockley and Jensen [@JacksonFacingOurHistory2023].

### 2023-11-02 paper/paper.qmd
reformatting mostly finished

@@
 ---
-title: "Mainstreaming Race Science"
+title: "Identifying Race Science in Mainstream Psychology, 1960-2010"
 authors:
@@
         city: "Merced"
-        state: "CA, USA"
+        country: "CA, USA"
     corresponding: true
@@
 abstract: |
-  | **Background:** While anthropologists and human geneticists have long rejected the biological race concept, both social movements and prominent researchers have purported to justify racial inequality and colonialism by appealing to the epistemic authority of science.
-  | **Methods:** Bibliometric and text mining methods were used to trace the appearance of race science discourse in both mainstream scientific journals and the parascholarly journal *Mankind Quarterly* from 1960 to 2010.
-  | **Results:** Our analysis found two distinguishable lines of race science discourse.
+  | While anthropologists and human geneticists have long rejected the biological race concept, both social movements and prominent researchers have purported to justify racial inequality and colonialism by appealing to the epistemic authority of science.
+  Bibliometric and text mining methods were used to trace the appearance of race science discourse in both mainstream scientific journals and the parascholarly journal *Mankind Quarterly* from 1960 to 2010.
+  Our analysis found two distinguishable lines of race science discourse.
 One is strongly associated with *Mankind Quarterly*, while the other — emphasizing race and intelligence — is strongly associated with mainstream psychology journals, especially *Intelligence* and *Personality and Individual Differences*.
-  | **Conclusion:** Psychology, and especially intelligence research, has played a significant role in mainstreaming race science.
+  Psychology, and especially intelligence research, has played a significant role in mainstreaming race science.
 
@@
 bibliography: [references.bib, refs-manual.bib]
-csl: nature.csl
+cite-method: "citeproc"
+csl: "nature-no-superscript.csl"
 
@@
 
+crossref:
+  fig-prefix: ""
+
 # header-includes:
@@
 
-| 
-| **One sentence summary:** Bibliometric and text mining analysis indicate that race science discourse appeared in mainstream psychology journals throughout the second half of the twentieth century.
+<!-- NHB submission guidelines: <https://www.nature.com/nathumbehav/submission-guidelines> -->
 
@@
 
-We identified 16 researchers who had received funding from Pioneer; 14 of these researchers had profiles in the Web of Science (WoS) author search, allowing us identify 13 WoS-indexed journals that had published 6 or more of these authors.
-See @tbl-researchers and @fig-wos.
+We identified 16 researchers who had received funding from Pioneer; 14 of these researchers had profiles in the Web of Science (WoS) author search, allowing us to identify 13 WoS-indexed journals that had published 6 or more of these authors.
+See @tbl-researchers and figure @fig-wos.
 
@@
 
-@fig-wos shows that, while MQ is among the "Pioneer-publishing" journals, a number of mainstream journals are more prominent: *Personality and Individual Differences* (PID), *Intelligence* (Int), *Behavior Genetics* (BG), and *Psychological Reports* (PR).
+Figure @fig-wos shows that, while MQ is among the "Pioneer-publishing" journals, a number of mainstream journals are more prominent: *Personality and Individual Differences* (PID), *Intelligence* (Int), *Behavior Genetics* (BG), and *Psychological Reports* (PR).
 In addition, only psychologist Richard Lynn appears to have published heavily in MQ.
@@
 
-@fig-focal and this discussion focuses on three topics identified in one of the 24 fitted topic models; see figures @fig-gamma-lg, @fig-gamma-md, @fig-gamma-sm and supplemental visualizations S2-S4 for all topics from all models and S5-S7 for selected topics from all models.
+Figure @fig-focal and this discussion focuses on three topics identified in one of the 24 fitted topic models; see figures [@fig-gamma-lg; @fig-gamma-md; @fig-gamma-sm; @fig-silge-lg; @fig-silge-md; @fig-silge-sm] for all topics from all models and figures [@fig-grid-lg; @fig-grid-md; @fig-grid-sm] for selected topics from all models.
 In the focal model, topic 07 is strongly associated with MQ:  MQ published dozens of articles in this topic each year, and no other journal ever published more than a handful.
@@
 
-Corpus assembly is summarized in @fig-corpus.  The corpus was assembled in two parts, one for *Mankind Quarterly* and the other for mainstream journals.
+Corpus assembly is summarized in figure @fig-corpus.  The corpus was assembled in two parts, one for *Mankind Quarterly* and the other for mainstream journals.
 
@@
 After XML/PDF versions were retrieved, text was extracted using either the `xml2` package [@WickhamXml2ParseXML2023] or `pdftools`.
-@fig-counts and @tbl-counts show the number of fulltext documents included in the corpus.
+Figure @fig-counts and @tbl-counts show the number of fulltext documents included in the corpus.
 All together, the corpus includes 34,896 documents from 6 journals.
@@
 After fitting topic models, our first research question was whether we could identify distinctive "race science discourse" topics.
-[@fig-gamma-lg; @fig-gamma-md; and @fig-gamma-sm] show the gamma (topic-document) distributions for each value of $k$ for the three vocabularies.
+Figures [@fig-gamma-lg; @fig-gamma-md; and @fig-gamma-sm] show the gamma (topic-document) distributions for each value of $k$ for the three vocabularies.
 In these figures, each panel corresponds to a single journal-$k$ combination, each row of cells is a single document in that journal, and each column of cells is a single topic in that particular model.
@@
 Color intensity indicates the value of gamma for that particular topic-document combination.
-
-![Fitted topic models: Gamma (topic-document) distributions, large vocabulary.](img/05_lg_gamma.png){#fig-gamma-lg width="6in" height="4.8in"}
-
-![Fitted topic models: Gamma (topic-document) distributions, medium vocabulary.](img/05_md_gamma.png){#fig-gamma-md width="6in" height="4.8in"}
-
-![Fitted topic models: Gamma (topic-document) distributions, small vocabulary.](img/05_sm_gamma.png){#fig-gamma-sm width="6in" height="4.8in"}
 
@@
 
-We next constructed Silge plots [@SilgeTopicModeling2017], showing the top 15 (highest-probability) terms (noun phrases) from each topic in the model.
-These are provided as supplemental PDFs S2-S4.
+We next constructed Silge plots [@SilgeTopicModeling2017], showing the top 15 (highest-probability) terms (noun phrases) from each topic in the model; see figures [@fig-silge-lg; @fig-silge-md; @fig-silge-sm].
 Perusing these term lists, we focused on topics in each model that contained racial terms.
@@
 The gamma distributions and Silge plots suggested that, once $k$ was sufficiently large, the topic models were distinguishing between two types of race science articles.
-We therefore examined the prevalence of topics that used racial terms, intelligence-research terms, or both, by journal and across all 24 models.
-Supplemental PDFs S5-7 provide these visualizations.
+We therefore examined the prevalence of topics that used racial terms, intelligence-research terms, or both, by journal and across all 24 models; see figures [@fig-grid-lg; @fig-grid-md; @fig-grid-sm].
 In these visualizations, each row of facets corresponds to a value of $k$ (excluding 5 as too noisy), and each column of facets corresponds to one topic.
@@
 # References
+
+::: {#refs}
+:::
+
+\clearpage
+
+# Extended Data
+
+\renewcommand{\thefigure}{ED-\arabic{figure}}
+\setcounter{figure}{0}
+
+![Fitted topic models: Gamma (topic-document) distributions, large vocabulary.](img/05_lg_gamma.png){#fig-gamma-lg width="6in" height="4.8in"}
+
+![Fitted topic models: Gamma (topic-document) distributions, medium vocabulary.](img/05_md_gamma.png){#fig-gamma-md width="6in" height="4.8in"}
+
+![Fitted topic models: Gamma (topic-document) distributions, small vocabulary.](img/05_sm_gamma.png){#fig-gamma-sm width="6in" height="4.8in"}
+
+![Silge plots (top 15 terms) for each topic and model, large vocabulary.](supplements/S2-lg-silge.pdf){#fig-silge-lg}
+
+![Silge plots (top 15 terms) for each topic and model, medium vocabulary.](supplements/S3-md-silge.pdf){#fig-silge-md}
+
+![Silge plots (top 15 terms) for each topic and model, small vocabulary.](supplements/S4-sm-silge.pdf){#fig-silge-sm}
+
+![Topics with "race," "intelligence," or both occurring in the top 15 terms, large vocabulary.](supplements/S5_grid_lg.pdf){#fig-grid-lg}
+
+![Topics with "race," "intelligence," or both occurring in the top 15 terms, medium vocabulary.](supplements/S6_grid_md.pdf){#fig-grid-md}
+
+![Topics with "race," "intelligence," or both occurring in the top 15 terms, small vocabulary.](supplements/S7_grid_sm.pdf){#fig-grid-sm}

### 2023-11-08 paper/paper.qmd
ready for nhb?

@@
 ---
-title: "Identifying Race Science in Mainstream Psychology, 1960-2010"
+title: "Race Science in Mainstream Psychology, 1960-2010"
 authors:
@@
   | While anthropologists and human geneticists have long rejected the biological race concept, both social movements and prominent researchers have purported to justify racial inequality and colonialism by appealing to the epistemic authority of science.
-  Bibliometric and text mining methods were used to trace the appearance of race science discourse in both mainstream scientific journals and the parascholarly journal *Mankind Quarterly* from 1960 to 2010.
+  In this study, bibliometric and text mining methods were used to trace the appearance of race science discourse in both mainstream behavioral science journals and the parascholarly journal *Mankind Quarterly* from 1960 to 2010.
   Our analysis found two distinguishable lines of race science discourse.
-One is strongly associated with *Mankind Quarterly*, while the other — emphasizing race and intelligence — is strongly associated with mainstream psychology journals, especially *Intelligence* and *Personality and Individual Differences*.
+One is strongly associated with *Mankind Quarterly*, while the other — emphasizing race and intelligence — is strongly associated with mainstream journals, especially *Intelligence* and *Personality and Individual Differences*.
   Psychology, and especially intelligence research, has played a significant role in mainstreaming race science.
@@
     number-sections: false
+    link-citations: true
 
@@
 
-The aim of the study reported here was to trace race science in the scholarly and parascholarly literature, examining the ways that fringe ideas can present themselves as legitimate scholarly inquiry.
+The aim of the study reported here was to trace race science in the scholarly and parascholarly literature, examining the ways that fringe ideas can present themselves in mainstream venues as legitimate scholarly inquiry.
 
@@
 
-These distinction allow us to bracket the intentions and mindset of the researchers involved in any particular piece of research or debate.
-For the purposes of the current study, it is not important whether, for instance, Arthur Jensen is or is not correctly labelled a scientific racist [@JacksonJr.ArthurJensenEvolutionary2022].
-But it is important that much of Jensen's research was useful and indeed utilized to purportedly justify racial inequality.
+These distinction allow us to bracket the intentions and attitudes of the researchers involved in any particular piece of research or debate.
+For the purposes of the current study, it is not important whether, for instance, psychologist Arthur Jensen is or is not correctly labelled a scientific racist [@JacksonJr.ArthurJensenEvolutionary2022].
+But it is important that much of Jensen's research was useful and indeed utilized to purportedly justify racial inequality, and so unambiguously qualifies as race science.
 
@@
 1 & 2) and is listed as editor-in-chief on MQ's current website as of 2023-07-21.
-He has also been president of Pioneer since the death of psychologist J.
-Phillippe Rushton in 2012 [@BeirichPioneerFundAssets2013].
-By contrast, a number of Pioneer-funded researchers have published in PID, Int, and to a lesser degree BG: Bouchard, Eysenck, Jensen, Rushton, Vernon, and also Lynn.
+He was also president of Pioneer from the death of psychologist J.
+Phillippe Rushton in 2012 [@BeirichPioneerFundAssets2013] until his own death in 2023.
+By contrast, a number of Pioneer-funded researchers have published a significant number of papers in PID, Int, and to a lesser degree BG: Bouchard, Eysenck, Jensen, Rushton, Vernon, and also Lynn.
 
@@
 We assembled a full-text corpus of articles published in MQ, PID, Int, BG, PR, and Behavior and Brain Sciences (BBS) between 1960-2010 and used topic modeling to identify topics discussing race, intelligence, and both.
+(See the materials and methods for a discussion of some important limitations with this corpus: journals published by the American Psychological Association could not be included, BBS bundles focus articles and commentaries together in a way that confounds simple document counts, and an unknown number of relevant documents in PID could not be included.)
 
@@
 
-Figure @fig-focal and this discussion focuses on three topics identified in one of the 24 fitted topic models; see figures [@fig-gamma-lg; @fig-gamma-md; @fig-gamma-sm; @fig-silge-lg; @fig-silge-md; @fig-silge-sm] for all topics from all models and figures [@fig-grid-lg; @fig-grid-md; @fig-grid-sm] for selected topics from all models.
+Figure @fig-focal and this discussion focuses on three topics identified in one of the 24 fitted topic models; see figures [@fig-gamma-lg; @fig-gamma-md; @fig-gamma-sm; @fig-silge-lg; @fig-silge-md; @fig-silge-sm] for all topics from all models, and figures [@fig-grid-lg; @fig-grid-md; @fig-grid-sm] for selected topics from all models.
 In the focal model, topic 07 is strongly associated with MQ:  MQ published dozens of articles in this topic each year, and no other journal ever published more than a handful.
-The "Silge plot," showing the top 15 terms in the topic [@SilgeTopicModeling2017], contains racial terms (`races`, `whites`, `negroes` and potentially `europe`, `africa`, `india`, and `japan`) as well as `book`, likely reflecting the fact that MQ published a number of book reviews, while the psychology journals either did not or these were not available for the corpus.
+The "Silge plot," showing the top 15 terms in the topic [@SilgeTopicModeling2017], contains a number of racial terms (`races`, `whites`, `negroes` and potentially `europe`, `africa`, `india`, and `japan`); as well as `book`, likely reflecting the fact that MQ published a number of book reviews, while the mainstream journals either did not publish book reviews or these were not available for the corpus.
 
@@
 Instead this topic appears to identify "mainstream" (non-race science) intelligence research, especially factor analysis and the debate over whether intelligence is unidimensional or multidimensional.
-(Three other topics only associated with mainstream intelligence journals and terms were also identified by this topic model.)  This topic indicates that the model is not simply lumping race science research together with other intelligence research.
+(Three other topics only associated with mainstream intelligence journals and terms were also identified by this topic model.)  This topic indicates that the model is able to distinguish "mainstream" intelligence research from race science discourse.
 
 The Silge plot for topic 24, by contrast, suggests a distinct race science discourse topic, with multiple racial terms (`whites`, `blacks`, `racial_differences`, `races`, `race_differences`) and the names of three prominent Pioneer-funded researchers, `jensen`, `lynn`, and `rushton`.
-Independent qualitative coding of the top 121 papers in this topic (those with $\gamma > 0.97$) confirmed this interpretation of the topic, with 108 (89%) coded as race science discourse by both authors (Cohen's $\kappa = 0.86$).
+Independent qualitative coding of 121 papers in this topic (those with $\gamma > 0.97$ and published in journals other than PR) confirmed this interpretation, with 108 (89%) coded as race science discourse by both authors (Cohen's $\kappa = 0.86$).
 
@@
 
-It's notable that BG published very few articles in any race-and-intelligence topic identified by any of the 24 topic models.
+BG published very few articles in any race-and-intelligence topic identified by any of the 24 topic models.
 Among both academics and the general public, the field of behavior genetics is strongly associated with race science, and specifically race-and-intelligence research.
@@
 However, the topic model analysis suggests that, by the 1990s, behavior genetics may have distanced itself somewhat from race science, albeit without directly repudiating it.
-This interpretation was supported by a supplemental analysis that focused on BG specifically (SM section 1.5).
+This interpretation was supported by a supplemental analysis that focused on BG specifically (see methods and materials)<!-- [@sec-bg] -->.
 In addition, the 1995 BGA presidential address by Glayde Whitney --- in which he criticized the "Marxist-Lysenkoist denial of genetics" and proposed that differences in murder rates between countries and cities were caused by racial genetic differences in intelligence, empathy, aggression, and impulsivity --- was published in MQ rather than BG.
-As a subdiscipline of psychology, behavior genetics may have been prominent in promoting race science in the past, but appears to have been less receptive to such ideas in more recent history.
-
+Behavior genetics may have been prominent in promoting race science in the past, but appears to have been less receptive to such ideas in more recent history.
 Instead, our results suggest that other subdisciplines of psychology have provided the primary venue for mainstreaming race science.
-Over the past few decades, many professional organizations in genetics and anthropology have made formal statements rejecting race as a biologically meaningful concept [@anthropology1996aapa; @american2018ashg; @board1998aaa; @committee2020asa; @FuentesAAPAStatementRace2019; @nhgri2023; @rotimi2022ashg; @wynshawboris2020ashg], undermining the core assumption of race science.
-The American Psychological Association (APA) --- whose members are predominantly clinical psychologists --- and the Federation of Associations in Behavioral & Brain Sciences (FABBS) --- an organization whose purpose is to provide policy recommendations --- have openly rejected race as biologically meaningful [@american2021apology; @baron2022fabbs].
-But research-focused psychological organizations like the Association for Psychological Science (APS) and Psychonomic Society (PS) stop short of rejecting race as biologically meaningful when they denounce racism [@aps2021statement; @board2020psychonomic].
-Psychology as a discipline maintains space for several mythological race science narratives, such as allegations of a taboo against race and intelligence research [@jackson2021mythical], claims that scientists engaging in or calling for any kind of anti-racism in the field is an ideological corruption of dispassionate and value-neutral science [@roberts_2022], or arguments that holding race science to the same evidentiary standards as other psychological research is a violation of academic freedom [@herbert2023academic].
-And historically several race scientists have secured important gatekeeping positions within the scientific community, such as on the editorial boards of Int and PID.
 
@@
 In some cases, some race scientists simultaneously maintained active connections to both the mainstream scientific community and white supremacist organizations [@WinstonScientificRacismNorth2020; @JacksonJr.ArthurJensenEvolutionary2022; @SlobodianUnequalMindHow2023].
-Except for the APA, we are not aware of other major organizations for scientific research in psychology that have issued apologies for the historical contributions of their field to scientific racism.
 
@@
 
+Over the past few decades, many professional organizations in genetics and anthropology have made formal statements rejecting race as a biologically meaningful concept [@anthropology1996aapa; @american2018ashg; @board1998aaa; @committee2020asa; @FuentesAAPAStatementRace2019; @nhgri2023; @rotimi2022ashg; @wynshawboris2020ashg], undermining the core assumption of race science.
+The American Psychological Association (APA) --- whose members are predominantly clinical psychologists --- the Federation of Associations in Behavioral & Brain Sciences (FABBS) --- an organization whose purpose is to provide policy recommendations --- and the editors of *Nature Human Behavior* have openly rejected race as biologically meaningful [@american2021apology; @baron2022fabbs; @ScienceMustRespect2022].
+But research-focused psychological organizations like the Association for Psychological Science (APS) and Psychonomic Society (PS) stop short of rejecting race as biologically meaningful when they denounce racism [@aps2021statement; @board2020psychonomic].
+Psychology as a discipline maintains space for several mythological race science narratives, such as allegations of a taboo against race and intelligence research [@jackson2021mythical], claims that scientists engaging in or calling for any kind of anti-racism in the field is an ideological corruption of dispassionate and value-neutral science [@roberts_2022], or arguments that holding race science to the same evidentiary standards as other psychological research is a violation of academic freedom [@herbert2023academic].
+And historically several race scientists have secured important gatekeeping positions within the scientific community, such as on the editorial boards of *Intelligence* and *Personality and Individual Differences*.
+
 Research on the genetics of intelligence has been socially harmful, not just in the somewhat abstract sense of promoting racial stigma, but in the concrete sense of being used to rationalize mass shootings and other acts of racial violence [@PronczukRacistResearcherExposed2022; @MeyerWrestlingSocialBehavioral2023].
 While scientists do not fully control the downstream social effects of their research, like other citizens scientists are responsible for mitigating reasonably foreseeable harms that result from their actions [@BlockIQHeritabilityInequality1974a; @DouglasSciencePolicyValuefree2009; @KitcherArgumentFreeInquiry1997; @KouranyShouldKnowledgeBe2016].
-These actions include not only individual decisions to research this topic or that, but also collective decisions about who is awarded positions of power or influence.
+These actions include not only individual decisions to research this topic or that, but also collective decisions about the allocation of scarce resources such as journal space, research funding, and positions of power and influence.
 Given the way research on intelligence and behavior genetics has been used historically, these fields may be especially susceptible to appropriation by scientific racists.
@@
 We ultimately used a series of ad hoc measures to mitigate text duplication.
-Approximately 100,000 documents that were 1-2 pages long were excluded before PDF retrieval.
+Approximately 100,000 documents that were 1-2 pages long (per Crossref metadata) were excluded before PDF retrieval.
 After PDF retrieval and text extraction, the timestamped watermarks were removed using a regular expression and the text was hashed using SHA 256.
-Hashes were used to construct groups of duplicate documents, and a single document (whichever one happened to be first in the dataframe) was chosen from each hash group for inclusion in the corpus.
+Hashes were used to efficiently construct groups of identical documents, and a single document (whichever one happened to be first in the dataframe) was chosen from each hash group for inclusion in the corpus.
 It is plausible that some duplicates made it through this process: where the watermark overlapped with the text, the regular expression likely would have been unable to identify and remove the watermark (a false negative result), and this small difference in the watermarks (not the text) would produce different hashes.
@@
 
-## Supplemental analysis of *Behavior Genetics* {#bg}
+## Supplemental analysis of *Behavior Genetics* {#sec-bg}
 

### 2024-03-29 paper/paper.qmd
working on reviewer comments

@@
 title: "Race Science in Mainstream Psychology, 1960-2010"
-authors:
-  - name: "Daniel J. Hicks"
-    email: "dhicks4@ucmerced.edu"
-    affiliations: 
-      - id: ucm
-        name: "University of California, Merced"
-        city: "Merced"
-        country: "CA, USA"
-    corresponding: true
-  - name: "Emilio J. C. Lobato"
-    affiliations:
-      - ref: ucm
+# authors:
+#   - name: "Daniel J. Hicks"
+#     email: "dhicks4@ucmerced.edu"
+#     affiliations: 
+#       - id: ucm
+#         name: "University of California, Merced"
+#         city: "Merced"
+#         country: "CA, USA"
+#     corresponding: true
+#   - name: "Emilio J. C. Lobato"
+#     affiliations:
+#       - ref: ucm
 abstract: |
@@
 We then used the author search tool in Clarivate's Web of Science platform (<https://www.webofscience.com/wos/author/search>), retrieving publication lists for 14 researchers.
-These searches were conducted between 2021-09-24 and 2021-10-05 by DJH.
+These searches were conducted between 2021-09-24 and 2021-10-05 by *[AUTHOR 1]*.
 After parsing these results, we counted how many of the 14 researchers had published in each journal.
@@
 
-The public analysis repository, <https://github.com/dhicks/race-science> *[DOI]*, includes document metadata and documentwise counts of NLP-extracted noun phrases ("document-term matrices"), along with code to reproduce the analysis of the following sections.
+The public analysis repository, <!-- <https://github.com/dhicks/race-science> --> *[URL and DOI will go here]*, includes document metadata and documentwise counts of NLP-extracted noun phrases ("document-term matrices"), along with code to reproduce the analysis of the following sections.
 
@@
 To fit topic models, we followed the approach proposed by Rohe and Zeng [@RoheVintageFactorAnalysis2020], which uses varimax-rotated partial principal components instead of the variational inference methods used by standard topic model packages such as `stm` [@RobertsStmPackageStructural2019].
-This novel approach was implemented in the R package `tmfast` [@HicksTmfastFitsTopic2023].
-We conducted a simulation study of `tmfast`, which found that it was significantly faster and only slightly less accurate at reconstructing known word-topic and topic-document distributions, compared to `stm` [@HicksTmfastFitsTopic2023].
+This novel approach was implemented in the R package *[redacted for anonymous review]* <!-- `tmfast` [@HicksTmfastFitsTopic2023] -->.
+<!-- We conducted a simulation study of `tmfast`, which found that it was significantly faster and only slightly less accurate at reconstructing known word-topic and topic-document distributions, compared to `stm` [@HicksTmfastFitsTopic2023].
+-->
 

### 2024-03-29 scripts/11-proof-of-concept.qmd
working on reviewer comments

This section uses simulated data to provide a proof-of-concept for our use of topic models to detect the "mainstreaming" of a marginalized topic.

```{r setup}
#| message: false
#| warning: false
library(tidyverse)
theme_set(theme_bw())
library(tmfast)

library(glue)
library(tictoc)
library(memoise)

set.seed(2024-03-18)

## Simulation parameters
k = 20            ## total number of topics
N = 100           ## papers per journal
vocab = 1000      ## vocabulary size
paper_size = 500  ## length of each paper
```

We construct a simulated corpus as follows.  We work with five journals, labelled A through E.  Journal A primarily publishes the marginalized topic (correspond to race science).  Ten journals, C-L, are mainstream journals, and publish an even mix from `r k-1` other topics, but avoid the marginalized topic (weighted 0 in the Dirichlet parameterization below).  These journals represent the context or comparison group for identifying the race science topic; their exact number is not important here.

Journal B is the venue by which the marginalized topic is "mainstreamed."  This journal is represented in a time-indexed way, $B_t$, with $t$ ranging from 0 to 5.  For simplicity, in the text each time-slice of B is treated as a distinct journal, giving us 17 journals total (A; the ten C-L; plus 6 time-slices of B).

This gives us three sets of journals: A, B (0-5), and C-L.  In the LDA model, these journal sets correspond to $\mathbf{\alpha}$, the parameter for the "prior" Dirichlet distribution $\Theta(\mathbf{\alpha})$. The `tmfast` package provides a helper function `peak_alpha()` for constructing $\mathbf{\alpha}$.

```{r}
## Construct journal theta ----
## A: marginal topic journal
A_alpha = peak_alpha(k, 1, peak = .9)
A_alpha
## B: mainstream journal that picks up marginal topic over time
B_alpha = map(0:5, ~peak_alpha(k, 1, peak = 0.05*.x)) |> 
    set_names(glue('B{0:5}'))
## C-E: mainstream journals
C_alpha = rep.int(list(peak_alpha(k, 1, peak = 0)), 10) |> 
    set_names(LETTERS[3:12])
C_alpha$C

alphas = c(list(A = A_alpha), 
           B_alpha, 
           C_alpha)

## 17 journals/journals-at-time-t
length(alphas)
```

Each paper has a simplex $\theta \sim \Theta(\mathbf{\alpha})$, representing a distribution over topics. We draw `r N` papers from each journal.

```{r}
theta = map(alphas, ~ rdirichlet(N, .x)) |> 
    imap(~ magrittr::set_rownames(.x, glue('{.y}-{str_pad(1:N, 
                                           2, 
                                           pad = "0")}'))) |> 
    reduce(rbind)

head(theta)

theta_df = theta |> 
    as_tibble(rownames = 'doc', 
              .name_repair = tmfast:::make_colnames) |> 
    pivot_longer(cols = -doc, 
                 names_to = 'topic', 
                 values_to = 'theta')

theta_df
```

We next define $\phi$, the distribution over words for each topic.  We use a vocabulary of `r vocab` words, and draw each topic's $\phi$ as a simplex from a symmetric Dirichlet distribution with $\alpha = 0.1$. Note that words are simply natural numbers.

```{r}
## Word distributions ----
## 20 topics, 1000 terms, 
## uniform Dirichlet distribution w/ param 0.1
phi = rdirichlet(k, 0.1, k = vocab)
str(phi)
```

For simplicity, we use a uniform length of `r paper_size` words for each paper. `draw_corpus` assumes that this is a vector with one element for each document.

```{r}
## Document length ----
## 500 words each
doc_len = rep.int(paper_size, N * length(alphas))
```

With these elements constructed, `draw_corpus()` generates a synthetic document-term matrix. Because this process is relatively slow (approximately 75 seconds on the laptop used to prepare this simulation), we use `memoise` to create a cache. As a pre-processing step for the topic model, we apply the `log(x+1)` transform.

```{r}
## Draw corpus ----
draw_corpus_cached = memoise(draw_corpus, 
                              cache = cachem::cache_disk('draw_corpus_cache'))
set.seed(2024-03-25)  ## NB draw_corpus() is non-deterministic! 
tic()
corpus = draw_corpus_cached(doc_len, theta, phi) |> 
    mutate(doc = rownames(theta)[doc], 
           n = log1p(n))
toc()
corpus
```

We can now fit the topic model.

```{r}
## Fit topic model ----
fitted = tmfast(corpus, c(10, 15, 20, 25, 30))
```

The tile plot indicates that (1) documents in journal A generally have a high probability for topic 1; journals C-E generally have a low probability; and journal B gradually has more articles in topic 1.

```{r}
## Topics by journal ----
add_journal = function(x) {
    mutate(x, journal = str_extract(document, '^[A-Z]*[0-9]?'))
}

tidy(fitted, k = 20, matrix = 'gamma') |> 
    add_journal() |> 
    ggplot(aes(document, topic, fill = gamma)) +
    geom_tile() +
    facet_wrap(vars(journal), scales = 'free', 
               ncol = 7) +
    scale_x_discrete(guide = 'none')
```

Focusing on topic 1, we first calculate the average probability across each journal.  A has the highest value; C-E are consistently low; and B shows an increase over time.

```{r}
## Topic model clearly detects the focal topic in A, increase in B over time, minimal presence in other journals
tidy(fitted, k = 20, matrix = 'gamma') |> 
    filter(topic == 'V01') |> 
    add_journal() |> 
    ggplot(aes(journal, gamma)) +
    ggforce::geom_sina() +
    stat_summary(color = 'red')
```

While the journals are largely distinguishable on the other topics, except for the unusually low value for A. 
```{r}
## With no pattern for the other topics
tidy(fitted, k = 20, matrix = 'gamma') |> 
    filter(topic == 'V02') |> 
    add_journal() |> 
    ggplot(aes(journal, gamma)) +
    ggforce::geom_sina() +
    stat_summary(color = 'red')
```

Next we calculate the rate of documents with probability greater than 0.1 in topic 1. As expected, this is almost all papers in journal A, no papers in the other mainstream journals, and in B the rate increases over time.

```{r}
tidy(fitted, k = 20, matrix = 'gamma') |> 
    filter(topic == 'V01', 
           gamma > .1) |> 
    add_journal() |> 
    count(journal) |> 
    mutate(rate = n / N)
```

We therefore conclude that this topic model method can detect the "mainstreaming" from a marginal topic over time.

As a limitation, we note that, for $\theta > 0.05$ or so, the estimated topic probabilities are below the true values. 
```{r}
tidy(fitted, k = 20, matrix = 'gamma') |> 
    filter(topic == 'V01') |> 
    add_journal() |> 
    inner_join(theta_df, by = c('document' = 'doc', 'topic')) |> 
    ggplot(aes(theta, gamma, color = journal)) +
    geom_point() +
    stat_function(fun = identity, 
                  inherit.aes = FALSE, color = 'black') +
    scale_color_viridis_d()
```

*[tmfast paper]* notes this problem, and implements a "renormalization" option when extracting the probability distributions from the fitted model.  This "renormalization" increases the "peakedness" of the distribution (lower entropy), adjusting the highest probabilities up and making the small probabilities several orders of magnitude smaller.  Renormalization would just amplify the differences found here, and so is not implemented for this proof-of-concept.  Renormalization is used in the primary, empirical analysis.

### 2024-03-29 scripts/08_topics_over_time.R

@@ -236,11 +236,11 @@ big_grid = function(model_file, exp_file, name, plot = TRUE, verbose = TRUE) {
     exp = read_exp(exp_file)
     
     ## Identify topics of interest: 
-    ## race and/or intelligence keywords in top 10 terms
+    ## race and/or intelligence keywords in top 15 terms
     topics = tidy_all(model) |> 
         filter(k > 5) |> 
         group_by(k, topic) |> 
-        top_n(10, beta) |> 
+        top_n(15, beta) |> 
         arrange(k, topic, beta) |> 
         summarize(race = any(str_detect(token, 'race|whites|blacks')), 
                   intelligence = any(str_detect(token, 'intelligence|iq') & 

### 2024-03-29 scripts/12_stm.R

@@ -1,0 +1,311 @@
+library(tidyverse)
+theme_set(theme_bw())
+library(stm)
+library(tidytext)
+library(ggh4x)
+
+library(arrow)
+library(here)
+library(tictoc)
+library(glue)
+
+data_dir = here('data')
+k = 40
+vocab = 'md'
+
+## Load data ----
+meta_ar = open_dataset(here(data_dir, '01_metadata'))
+
+## Medium vocab
+vocabulary = here(data_dir, glue('03-vocab-{vocab}.Rds')) |> 
+    read_rds()
+source(here('R', 'phrases.R'))
+phrases_mx = phrases() |> 
+    filter(phrase %in% vocabulary$phrase) |> 
+    collect() |> 
+    arrange(article_id, phrase) |> 
+    cast_sparse(article_id, phrase, n)
+
+article_ids = rownames(phrases_mx)
+
+## Fit topic model ----
+model_path = glue('12_stm_{vocab}_{k}.Rds') %>%
+    here(data_dir, .)
+if (!file.exists(model_path)) {
+    ## w/ emtol = 1e-4, fitting model takes 1190 sec ~ 20 minutes
+    ## (versus 80+ with 1e-5)
+    tic()
+    stm_fit = stm(phrases_mx, K = k, seed = 2024-03-29, emtol = 1e-4)
+    toc()
+    write_rds(stm_fit, model_path)
+} else {
+    stm_fit = read_rds(model_path)
+}
+
+
+## Match stm and tmfast topics ----
+beta_stm = tidy(stm_fit, matrix = 'beta') |>
+    mutate(topic = str_pad(topic, 2)) |>
+    arrange(term, topic)
+
+beta_tmfast = glue('04_{vocab}_tmfast.Rds') %>%
+    here(data_dir, '04_tm', .) |>
+    read_rds() |>
+    tmfast::tidy(k = 40, matrix = 'beta') |>
+    arrange(token, topic)
+
+# magrittr::equals(
+#     {tmfast:::build_matrix(beta_stm, topic, term, beta, sparse = FALSE) |>
+#             colnames()},
+#     {tmfast:::build_matrix(beta_tmfast, topic, token, beta, sparse = FALSE) |>
+#             colnames()}
+# ) |>
+#     all()
+
+hellingers = tmfast:::hellinger.data.frame(
+    beta_stm,    id1 = topic, cat1 = term,  prob1 = beta,
+    beta_tmfast, id2 = topic, cat2 = token, prob2 = beta,
+    df = TRUE)
+
[…]

### 2024-03-29 scripts/13_pf_authors.R
first pass at author analysis

@@ -1,0 +1,111 @@
+library(tidyverse)
+theme_set(theme_bw())
+library(tmfast)
+
+library(arrow)
+library(here)
+
+data_dir = here('data')
+tm_dir = here(data_dir, '04_tm')
+
+## Load data ----
+source(here('R', 'pf_authors.R'))
+
+meta_df = open_dataset(here(data_dir, '01_metadata')) |> 
+    mutate(article_id = tolower(article_id)) |> 
+    inner_join(pf_authors, by = 'article_id') |>
+    collect()
+
+## Available documents by author
+count(meta_df, pf_author)
+
+## Load topic models ----
+model_files = list.files(tm_dir, '*_tmfast.Rds') %>%
+    here(tm_dir, .) |> 
+    set_names('lg', 'md', 'sm')
+exp_files = list.files(data_dir, '*_exponents.Rds') %>%
+    here(data_dir, .) |> 
+    set_names('lg', 'md', 'sm')
+## Helper to read exponents as a named num
+read_exp = function(file) {
+    file |> 
+        read_rds() |> 
+        pull(exponent, name = k)
+}
+
+
+this_vocab = 'md'
+
+this_model_file = model_files[this_vocab]
+this_model = read_rds(this_model_file)
+this_exp = exp_files[this_vocab] |> 
+    read_exp()
+
+gamma_df = tidy_all(this_model, 
+                    matrix = 'gamma', 
+                    exponent = this_exp) |> 
+    inner_join(meta_df, 
+               by = c('document' = 'article_id'), 
+               multiple = 'all')
+
+## Tile plot ----
+## Difficult to read bc so many authors
+ggplot(gamma_df, aes(topic, document, fill = gamma)) +
+    geom_raster() +
+    geom_vline(xintercept = c(10, 20, 30, 40) + .5,
+               color = 'black') +
+    facet_grid(rows = vars(pf_author), 
+               cols = vars(k),
+               scale = 'free', 
+               switch = 'y',
+               labeller = label_wrap_gen(width = 15), 
+               # space = 'free_y'
+               ) +
+    scale_x_discrete(breaks = c('V10', 'V20', 'V30', 'V40')) +
+    scale_y_discrete(breaks = NULL,
+                     name = '') +
+    scale_fill_viridis_c(option = 'viridis', direction = 1) +
+    theme_minimal() +
[…]

### 2024-04-15 paper/paper.qmd
first pass integrating proof-of-concept

@@
     link-citations: true
+    mathspec: true
 
@@
 #  - \linenumbers
+
+execute:
+  cache: true
 ---
-
-<!-- NHB submission guidelines: <https://www.nature.com/nathumbehav/submission-guidelines> -->
 
@@
 
+### Proof of concept
+
+{{< include ../scripts/11-proof-of-concept.qmd >}}
+
 ### Topic model interpretation and quality assessment

### 2024-04-15 scripts/13_pf_authors.R

@@ -42,8 +42,9 @@ this_exp = exp_files[this_vocab] |>
     read_exp()
 
 gamma_df = tidy_all(this_model, 
-                    matrix = 'gamma', 
-                    exponent = this_exp) |> 
+                    matrix = 'gamma') |> 
+    group_split(k) |> 
+    map2_dfr(this_exp, ~ renorm(.x, document, gamma, .y)) |> 
     inner_join(meta_df, 
                by = c('document' = 'article_id'), 
                multiple = 'all')
@@ -72,10 +73,31 @@ ggplot(gamma_df, aes(topic, document, fill = gamma)) +
 
 gamma_df |> 
     filter(k == 40) |> 
+    ggplot(aes(topic, document, fill = gamma)) +
+    geom_raster() +
+    geom_vline(xintercept = c(10, 20, 30, 40) + .5,
+               color = 'black') +
+    facet_wrap(vars(pf_author), 
+               scale = 'free', 
+               labeller = label_wrap_gen(width = 15), 
+               # space = 'free_y'
+    ) +
+    scale_x_discrete(breaks = c('V10', 'V20', 'V30', 'V40')) +
+    scale_y_discrete(breaks = NULL,
+                     name = '') +
+    scale_fill_viridis_c(option = 'viridis', direction = 1) +
+    theme_minimal() +
+    theme(axis.ticks.x = element_line(),
+          legend.position = 'bottom',
+          plot.margin = margin(r = 10))
+
+gamma_df |> 
+    filter(k == 40) |> 
+    mutate(author_count = n(), .by = pf_author) |> 
     ggplot(aes(topic, gamma, group = document)) +
+    geom_line(aes(alpha = 1/author_count), show.legend = FALSE) +
     geom_vline(xintercept = c('V07', 'V22', 'V24'), 
-               color = 'blue', alpha = .5) +
-    geom_line(alpha = .2) +
+               color = 'blue', alpha = 1) +
     facet_wrap(vars(pf_author)) +
     scale_x_discrete(breaks = c('V10', 'V20', 'V30'))
 
@@ -96,6 +118,25 @@ gamma_df |>
     filter(k == 40) |> 
     group_by(pf_author, document) |> 
     filter(gamma == max(gamma)) |> 
+    group_by(pf_author) |> 
+    count(topic) |> 
+    complete(topic = {1:40 |> 
+            str_pad(2, pad = '0') %>% 
[…]

### 2024-05-03 paper/paper.qmd
prep to merge back into master

@@
 title: "Race Science in Mainstream Psychology, 1960-2010"
-# authors:
-#   - name: "Daniel J. Hicks"
-#     email: "dhicks4@ucmerced.edu"
-#     affiliations: 
-#       - id: ucm
-#         name: "University of California, Merced"
-#         city: "Merced"
-#         country: "CA, USA"
-#     corresponding: true
-#   - name: "Emilio J. C. Lobato"
-#     affiliations:
-#       - ref: ucm
+authors:
+  - name: "Daniel J. Hicks"
+    email: "dhicks4@ucmerced.edu"
+    affiliations: 
+      - id: a
+        name: "University of California, Merced"
+        city: "Merced"
+        country: "CA, USA"
+    corresponding: true
+  - name: "Emilio J. C. Lobato"
+    affiliations:
+      - ref: a
+
 abstract: |
@@
 bibliography: [references.bib, refs-manual.bib]
-cite-method: "citeproc"
-csl: "nature-no-superscript.csl"
+# cite-method: "citeproc"
+# csl: "nature-no-superscript.csl"
 
 format: 
-  elsevier-pdf:
-    journal: 
-      name: "journal"
+  pnas-pdf:
     keep_tex: true
     latex_engine: "xelatex"
-    number-sections: false
     link-citations: true
     mathspec: true
+    classoption:
+      - twocolumn
+      - twoside
+    fig.env: "figure*"
 
@@
 ---
+
+<!-- 
+PNAS guidelines
+Research Report/Direct Submission: The preferred length of these articles is 6 pages, but PNAS allows articles up to a maximum of 12 pages.
+A standard 6-page article is approximately 4,000 words, 50 references, and 4 medium-size graphical elements (i.e., figures and tables).
+
+- a manuscript file (in any format) including the following:
+    title page (title, author list, classification, keywords)
+    abstract
+    significance statement
+    main text
+    references
+    figures or tables with appropriate legends (may be uploaded separately)
+    SI files (may be uploaded separately)
+- contact and competing interest information for all authors
+- data sharing plans (for all data, documentation, and code used in analysis)
+- funding information and whether an open access license has been selected
+- list of appropriate Editorial Board members, NAS members, and qualified reviewers (minimum of three each) who are experts in the paper's scientific area.
+(A brief justification for suggested reviewers is welcome, particularly for interdisciplinary papers)
+
+- revised submission will require track changes version
+
+figure sizes: 
+Small: approximately 9 cm x 6 cm
+Medium: approximately 11 cm x 11 cm
+Large: approximately 18 cm x 22 cm
+
+ -->
+
+*[TODO: 
+- NHB reviews todo
+- sort git tree
+- reorganize into PNAS ms + SI
+- check cite formatting
+- copy updated plots? 
+- abstract formatting
+- figure sizes
+]*
 
@@
 
+\onecolumn
+
 | | | |
@@
 
-![Journals publishing 6 or more Pioneer-funded researchers, Web of Science author search results.](img/wos_results.png){#fig-wos width="4.76in" height="2.6in"}
+\twocolumn
+
+![Journals publishing 6 or more Pioneer-funded researchers, Web of Science author search results.](img/wos_results.png){#fig-wos width="4.76in" height="2.6in" fig-env="figure*"}
 
@@
 Thin lines give annual values, thick lines give 5-year running averages.
-Because *Behavior Genetics* and *Behavioral and Brain Sciences* are not prominent in any panel, neither is given a direct label.](img/focal_topics.png){#fig-focal width="4.76in" height="3.57in"}
+Because *Behavior Genetics* and *Behavioral and Brain Sciences* are not prominent in any panel, neither is given a direct label.](img/focal_topics.png){#fig-focal width="4.76in" height="3.57in" fig-env="figure*"}
 
@@
 
-![Corpus assembly.](img/dataset.png){#fig-corpus width="120%"}
+![Corpus assembly.](img/dataset.png){#fig-corpus width="120%" fig-env="figure*"}
 
@@
 
-![Count of documents in the corpus, by journal and year.](img/02_count.png){#fig-counts width="4.76in" height="3.6in"}
+![Count of documents in the corpus, by journal and year.](img/02_count.png){#fig-counts width="11cm" height="11cm" fig-env="figure*"}
+
+\onecolumn
 
@@
 |Behavioral and Brain Sciences          |   898|  1978| 2010|
+
+\twocolumn
 
@@
 
-{{< include ../scripts/11-proof-of-concept.qmd >}}
+<!-- {{< include ../scripts/11-proof-of-concept.qmd >}} -->
 
@@
 However, "false negatives" are still possible, as some documents that human readers would classify as race science discourse might have been associated with other topics.
+
+### STM robustness check
+
+As a robustness check, we compared the `tmfast` results with a topic model fit using `STM`, a widely used package for topic modeling in R *[cite]*.
+Using the medium vocabulary, $k=40$, and a relaxed convergence threshold (`1e-4` compared to the default `1e-5`), fitting a single `STM` model took approximately 20 minutes.
+This was approximately two orders of magnitude slower than `tmfast`.
+And, unlike `tmfast`, `STM` can only fit a single value of $k$ at a time.
+We therefore only fit this one `STM` model.
+
+*[xref]* shows four topics extracted from the `STM` model, automatically identified by the use of race keywords, intelligence keywords, or both, among the top 15 terms.
+As in the `tmfast` analysis, the "race only" topic is almost exclusive to MQ; the "race and intelligence" topic emerges in mainstream journals before appearing in both the mainstream and MQ; and there are distinctive "intelligence only" topics.
+Compare *[xref]*.
+One notable difference between these results is the role of PR in publishing race and intelligence.
+
+![](img/12-stm.png){fig-env="figure*"}
+
+### *[author analysis]*
 

### 2024-05-03 paper/test.qmd
prep to merge back into master

--- 2024-03-29 scripts/11-proof-of-concept.qmd
@@
 paper_size = 500  ## length of each paper
+
+paper_size
 ```
-
-We construct a simulated corpus as follows.
-We work with five journals, labelled A through E.
-Journal A primarily publishes the marginalized topic (correspond to race science).
-Ten journals, C-L, are mainstream journals, and publish an even mix from `r k-1` other topics, but avoid the marginalized topic (weighted 0 in the Dirichlet parameterization below).
-These journals represent the context or comparison group for identifying the race science topic; their exact number is not important here.
-
-Journal B is the venue by which the marginalized topic is "mainstreamed."  This journal is represented in a time-indexed way, $B_t$, with $t$ ranging from 0 to 5.
-For simplicity, in the text each time-slice of B is treated as a distinct journal, giving us 17 journals total (A; the ten C-L; plus 6 time-slices of B).
-
-This gives us three sets of journals: A, B (0-5), and C-L.
-In the LDA model, these journal sets correspond to $\mathbf{\alpha}$, the parameter for the "prior" Dirichlet distribution $\Theta(\mathbf{\alpha})$.
-The `tmfast` package provides a helper function `peak_alpha()` for constructing $\mathbf{\alpha}$.
-
-```{r}
-## Construct journal theta ----
-## A: marginal topic journal
-A_alpha = peak_alpha(k, 1, peak = .9)
-A_alpha
-## B: mainstream journal that picks up marginal topic over time
-B_alpha = map(0:5, ~peak_alpha(k, 1, peak = 0.05*.x)) |> 
-    set_names(glue('B{0:5}'))
-## C-E: mainstream journals
-C_alpha = rep.int(list(peak_alpha(k, 1, peak = 0)), 10) |> 
-    set_names(LETTERS[3:12])
-C_alpha$C
-
-alphas = c(list(A = A_alpha), 
-           B_alpha, 
-           C_alpha)
-
-## 17 journals/journals-at-time-t
-length(alphas)
-```
-
-Each paper has a simplex $\theta \sim \Theta(\mathbf{\alpha})$, representing a distribution over topics. We draw `r N` papers from each journal.
-
-```{r}
-theta = map(alphas, ~ rdirichlet(N, .x)) |> 
-    imap(~ magrittr::set_rownames(.x, glue('{.y}-{str_pad(1:N, 
-                                           2, 
-                                           pad = "0")}'))) |> 
-    reduce(rbind)
-
-head(theta)
-
-theta_df = theta |> 
-    as_tibble(rownames = 'doc', 
-              .name_repair = tmfast:::make_colnames) |> 
-    pivot_longer(cols = -doc, 
-                 names_to = 'topic', 
-                 values_to = 'theta')
-
-theta_df
-```
-
-We next define $\phi$, the distribution over words for each topic.
-We use a vocabulary of `r vocab` words, and draw each topic's $\phi$ as a simplex from a symmetric Dirichlet distribution with $\alpha = 0.1$.
-Note that words are simply natural numbers.
-
-```{r}
-## Word distributions ----
-## 20 topics, 1000 terms, 
-## uniform Dirichlet distribution w/ param 0.1
-phi = rdirichlet(k, 0.1, k = vocab)
-str(phi)
-```
-
-For simplicity, we use a uniform length of `r paper_size` words for each paper. `draw_corpus` assumes that this is a vector with one element for each document.
-
-```{r}
-## Document length ----
-## 500 words each
-doc_len = rep.int(paper_size, N * length(alphas))
-```
-
-With these elements constructed, `draw_corpus()` generates a synthetic document-term matrix.
-Because this process is relatively slow (approximately 75 seconds on the laptop used to prepare this simulation), we use `memoise` to create a cache.
-As a pre-processing step for the topic model, we apply the `log(x+1)` transform.
-
-```{r}
-## Draw corpus ----
-draw_corpus_cached = memoise(draw_corpus, 
-                              cache = cachem::cache_disk('draw_corpus_cache'))
-set.seed(2024-03-25)  ## NB draw_corpus() is non-deterministic! 
-tic()
-corpus = draw_corpus_cached(doc_len, theta, phi) |> 
-    mutate(doc = rownames(theta)[doc], 
-           n = log1p(n))
-toc()
-corpus
-```
-
-We can now fit the topic model.
-
-```{r}
-## Fit topic model ----
-fitted = tmfast(corpus, c(10, 15, 20, 25, 30))
-```
-
-The tile plot indicates that (1) documents in journal A generally have a high probability for topic 1; journals C-E generally have a low probability; and journal B gradually has more articles in topic 1.
-
-```{r}
-## Topics by journal ----
-add_journal = function(x) {
-    mutate(x, journal = str_extract(document, '^[A-Z]*[0-9]?'))
-}
-
-tidy(fitted, k = 20, matrix = 'gamma') |> 
-    add_journal() |> 
-    ggplot(aes(document, topic, fill = gamma)) +
-    geom_tile() +
-    facet_wrap(vars(journal), scales = 'free', 
-               ncol = 7) +
-    scale_x_discrete(guide = 'none')
-```
-
-Focusing on topic 1, we first calculate the average probability across each journal.  A has the highest value; C-E are consistently low; and B shows an increase over time.
-
-```{r}
-## Topic model clearly detects the focal topic in A, increase in B over time, minimal presence in other journals
-tidy(fitted, k = 20, matrix = 'gamma') |> 
-    filter(topic == 'V01') |> 
-    add_journal() |> 
-    ggplot(aes(journal, gamma)) +
-    ggforce::geom_sina() +
-    stat_summary(color = 'red')
-```
-
-While the journals are largely distinguishable on the other topics, except for the unusually low value for A. 
-```{r}
-## With no pattern for the other topics
-tidy(fitted, k = 20, matrix = 'gamma') |> 
-    filter(topic == 'V02') |> 
-    add_journal() |> 
-    ggplot(aes(journal, gamma)) +
-    ggforce::geom_sina() +
-    stat_summary(color = 'red')
-```
-
-Next we calculate the rate of documents with probability greater than 0.1 in topic 1.
-As expected, this is almost all papers in journal A, no papers in the other mainstream journals, and in B the rate increases over time.
-
-```{r}
-tidy(fitted, k = 20, matrix = 'gamma') |> 
-    filter(topic == 'V01', 
-           gamma > .1) |> 
-    add_journal() |> 
-    count(journal) |> 
-    mutate(rate = n / N)
-```
-
-We therefore conclude that this topic model method can detect the "mainstreaming" from a marginal topic over time.
-
-As a limitation, we note that, for $\theta > 0.05$ or so, the estimated topic probabilities are below the true values. 
-```{r}
-tidy(fitted, k = 20, matrix = 'gamma') |> 
-    filter(topic == 'V01') |> 
-    add_journal() |> 
-    inner_join(theta_df, by = c('document' = 'doc', 'topic')) |> 
-    ggplot(aes(theta, gamma, color = journal)) +
-    geom_point() +
-    stat_function(fun = identity, 
-                  inherit.aes = FALSE, color = 'black') +
-    scale_color_viridis_d()
-```
-
-*[tmfast paper]* notes this problem, and implements a "renormalization" option when extracting the probability distributions from the fitted model.
-This "renormalization" increases the "peakedness" of the distribution (lower entropy), adjusting the highest probabilities up and making the small probabilities several orders of magnitude smaller.
-Renormalization would just amplify the differences found here, and so is not implemented for this proof-of-concept.
-Renormalization is used in the primary, empirical analysis.

### 2024-05-03 scripts/12_stm.R

@@ -1,8 +1,10 @@
+## This scripts fits a STM topic model, as a robustness check
 library(tidyverse)
 theme_set(theme_bw())
 library(stm)
 library(tidytext)
 library(ggh4x)
+library(cowplot)
 
 library(arrow)
 library(here)
@@ -29,6 +31,8 @@ phrases_mx = phrases() |>
 article_ids = rownames(phrases_mx)
 
 ## Fit topic model ----
+## per note below, fitting model is very slow, 
+## so we use a cache
 model_path = glue('12_stm_{vocab}_{k}.Rds') %>%
     here(data_dir, .)
 if (!file.exists(model_path)) {
@@ -306,6 +310,32 @@ big_grid = function(model, name, plot = TRUE, verbose = TRUE) {
         labs(caption = glue('{name} vocabulary'))
 }
 # debugonce(big_grid)
-big_grid(stm_fit, 'stm', plot = TRUE)
-ggsave(here('out', glue('12-grid-{Sys.time()}.png')), 
+grid_gg = big_grid(stm_fit, glue('stm-{vocab}-{k}'), plot = TRUE)
+grid_gg
+ggsave(here('out', glue('12-grid-stm.png')), 
        height = 4, width = 8, bg = 'white')
+
+## Combine w/ Silge plot
+focal_topics = c(16, 40, 9, 33)
+silge_gg = tidy(stm_fit, matrix = 'beta') |> 
+    filter(topic %in% focal_topics) |> 
+    mutate(topic = factor(topic, levels = focal_topics)) |> 
+    group_by(topic) |> 
+    top_n(15, beta) |> 
+    arrange(topic, desc(beta)) |> 
+    mutate(term = reorder_within(term, beta, topic)) |> 
+    ggplot(aes(term, beta)) +
+    geom_point() +
+    geom_linerange(aes(ymax = beta), ymin = 0) +
+    coord_flip() +
+    scale_x_reordered() + 
+    scale_y_continuous(name = 'β', 
+                       breaks = scales::pretty_breaks(3), 
+                       limits = c(0, NA)) +
+    facet_wrap(vars(topic), scales = 'free_y', nrow = 1)
+silge_gg
+
+plot_grid(silge_gg, grid_gg, ncol = 1, align = 'v', 
+          rel_heights = c(.7, 1))
+ggsave(here('out', glue('12-stm.png')), 
+       height = 6, width = 8, bg = 'white', scale = 1.5)

### 2024-05-03 scripts/13_pf_authors.R

@@ -1,152 +1,0 @@
-library(tidyverse)
-theme_set(theme_bw())
-library(tmfast)
-
-library(arrow)
-library(here)
-
-data_dir = here('data')
-tm_dir = here(data_dir, '04_tm')
-
-## Load data ----
-source(here('R', 'pf_authors.R'))
-
-meta_df = open_dataset(here(data_dir, '01_metadata')) |> 
-    mutate(article_id = tolower(article_id)) |> 
-    inner_join(pf_authors, by = 'article_id') |>
-    collect()
-
-## Available documents by author
-count(meta_df, pf_author)
-
-## Load topic models ----
-model_files = list.files(tm_dir, '*_tmfast.Rds') %>%
-    here(tm_dir, .) |> 
-    set_names('lg', 'md', 'sm')
-exp_files = list.files(data_dir, '*_exponents.Rds') %>%
-    here(data_dir, .) |> 
-    set_names('lg', 'md', 'sm')
-## Helper to read exponents as a named num
-read_exp = function(file) {
-    file |> 
-        read_rds() |> 
-        pull(exponent, name = k)
-}
-
-
-this_vocab = 'md'
-
-this_model_file = model_files[this_vocab]
-this_model = read_rds(this_model_file)
-this_exp = exp_files[this_vocab] |> 
-    read_exp()
-
-gamma_df = tidy_all(this_model, 
-                    matrix = 'gamma') |> 
-    group_split(k) |> 
-    map2_dfr(this_exp, ~ renorm(.x, document, gamma, .y)) |> 
-    inner_join(meta_df, 
-               by = c('document' = 'article_id'), 
-               multiple = 'all')
-
-## Tile plot ----
-## Difficult to read bc so many authors
-ggplot(gamma_df, aes(topic, document, fill = gamma)) +
-    geom_raster() +
-    geom_vline(xintercept = c(10, 20, 30, 40) + .5,
-               color = 'black') +
-    facet_grid(rows = vars(pf_author), 
-               cols = vars(k),
-               scale = 'free', 
-               switch = 'y',
-               labeller = label_wrap_gen(width = 15), 
-               # space = 'free_y'
-               ) +
-    scale_x_discrete(breaks = c('V10', 'V20', 'V30', 'V40')) +
-    scale_y_discrete(breaks = NULL,
-                     name = '') +
-    scale_fill_viridis_c(option = 'viridis', direction = 1) +
[…]

### 2024-05-03 scripts/14_term_lists.R

@@ -1,0 +1,65 @@
+library(tidyverse)
+theme_set(theme_minimal())
+library(tmfast)
+
+library(gt)
+
+library(arrow)
+library(here)
+library(glue)
+
+data_dir = here('data')
+tm_dir = here(data_dir, '04_tm')
+
+model = read_rds(here(tm_dir, '04_md_tmfast.Rds'))
+
+k = 40
+topics_of_interest = c('V07', 'V22', 'V24')
+
+phrases_ar = open_dataset(here(data_dir, '01_phrases.csv'), 
+                          format = 'csv')
+
+## Extract word-topic distributions ----
+## Top 250 words for each topic, ranked
+top_words = tidy(model, k, matrix = 'beta') |> 
+    filter(topic %in% topics_of_interest) |> 
+    group_by(topic) |> 
+    top_n(250, wt = beta) |> 
+    arrange(topic, desc(beta)) |> 
+    mutate(rank = row_number()) |> 
+    ungroup() |> 
+    group_split(topic)
+
+make_edges = function(x, y) {
+    inner_join(x, y, by = 'token')
+}
+
+edges = bind_rows(
+    make_edges(top_words[[1]], top_words[[2]]), 
+    make_edges(top_words[[2]], top_words[[3]]), 
+    {make_edges(top_words[[3]], top_words[[1]]) |> 
+            mutate(topic.y = str_c(topic.y, ' '))}) |> 
+    mutate(delta = rank.x - rank.y)
+
+ggplot(edges, aes(x = topic.x, xend = topic.y, 
+                      y = rank.x, yend = rank.y, 
+                      text = token)) +
+    geom_segment(aes(color = token, group = token)) +
+    scale_x_discrete(limits = c('V07', 'V22', 'V24', 'V07 '), 
+                     name = 'topic') +
+    scale_y_reverse(name = 'rank') +
+    scale_color_viridis_d(guide = 'none')
+
+## Beta rather than rank scale; harder to parse
+last_plot() + 
+    aes(y = beta.x, yend = beta.y) +
+    scale_y_continuous()
+
+## Table ----
+bind_rows(
+    make_edges(top_words[[1]], top_words[[2]]), 
+    make_edges(top_words[[2]], top_words[[3]]), 
+    make_edges(top_words[[3]], top_words[[1]])) |> 
+    group_by(topic.x, topic.y) |> 
+    summarize(n()) |> 
+    ungroup()

### 2024-05-03 scripts/scratch/all-umap.R

@@ -1,0 +1,48 @@
+library(tidyverse)
+
+library(arrow)
+library(here)
+
+data_dir = here('data')
+tm_dir = here(data_dir, '04_tm')
+
+meta_df = open_dataset(here(data_dir, '01_metadata')) |> 
+    collect()
+
+umap_files = list.files(tm_dir, 'umap')
+umap_df = umap_files %>%
+    here(tm_dir, .) |> 
+    set_names(umap_files) |> 
+    map_dfr(read_rds, .id = 'file') |> 
+    separate(file, 
+             into = c('script', 'vocabulary', 'k', 'umap'), 
+             sep = '_') |> 
+    select(-script, -umap) |> 
+    mutate(k = as.integer(k))
+
+bounds = umap_df |> 
+    group_by(vocabulary, k) |> 
+    summarize(across(c(x, y), 
+                     c(bot = ~ quantile(.x, .05), 
+                       top = ~ quantile(.x, .95)))) |> 
+    ungroup()
+
+
+umap_df |> 
+    left_join(bounds, by = c('vocabulary', 'k')) |> 
+    filter(x_bot <= x, x <= x_top,
+           y_bot <= y, y <= y_top) |> 
+    left_join(meta_df, by = c('document' = 'article_id')) |> 
+    ggplot(aes(x, y, color = container.title)) +
+    geom_point(alpha = .1) +
+    facet_wrap(vars(k, vocabulary),
+               ncol = 3,
+               scales = 'free', 
+               labeller = label_wrap_gen(multi_line = FALSE)) +
+    guides(colour = guide_legend(override.aes = list(alpha = 1))) +
+    theme_bw() +
+    theme(legend.position = 'bottom')
+ggsave('test.png', 
+       width = 3*4, 
+       height = 6*4 + 2)
+

### 2024-05-03 scripts/scratch/args.R

@@ -1,0 +1,1 @@
+source('../../R/args.R')

### 2024-05-03 scripts/scratch/hellinger_graph.R

@@ -1,0 +1,30 @@
+library(tidygraph)
+library(ggraph)
+
+
+graph = comp_df |> 
+    mutate(start_id = str_c(model_a, '-', topic_a), 
+           end_id = str_c(model_b, '-', topic_b)) |> 
+    select(start_id, end_id, everything()) |> 
+    as_tbl_graph() |> 
+    mutate(model = str_split(name, '-', simplify = TRUE)[,1], 
+           topic = str_split(name, '-', simplify = TRUE)[,2], 
+           across(c(model, topic), 
+                  as.integer))
+
+graph |> 
+    activate(edges) |> 
+    filter(similarity > .5) |> 
+    activate(nodes) |> 
+    filter(!node_is_isolated()) |> 
+    ggraph(layout = 'dendrogram') +
+    geom_edge_link(aes(color = similarity), width = 1) +
+    geom_node_label(aes(label = name), 
+                    alpha = .5, 
+                    nudge_y = -.15) +
+    scale_x_reverse() +
+    scale_y_reverse() +
+    scale_edge_color_viridis(limits = c(0.5, 1),
+                          direction = -1) +
+    coord_flip() +
+    theme(legend.position = c(.2, .2))

### 2024-05-03 scripts/scratch/pf_authors.R

@@ -1,0 +1,49 @@
+library(tidyverse)
+library(tmfast)
+
+library(here)
+library(arrow)
+
+source(here('R', 'author_data.R'))
+
+author_data() |> 
+    inner_join(mq, by = 'author') |> 
+    count(author)
+
+meta_df = open_dataset(here('data', '01_metadata')) |> 
+    inner_join(author_data(), by = 'article_id') |> 
+    inner_join(mq, by = 'author') |> 
+    collect()
+
+## Publications by journal ----
+ggplot(meta_df, aes(fct_rev(author), fill = container.title)) +
+    geom_bar(color = 'black', size = .25) +
+    # scale_y_sqrt(breaks = seq(100, 900, by = 200)) +
+    scale_fill_brewer(palette = 'Set1', direction = -1L) +
+    coord_flip() +
+    labs(x = 'author', 
+         fill = '') +
+    theme_minimal()
+meta_df |> 
+    count(author, container.title) |> 
+    complete(author, container.title, fill = list(n = 0)) |> 
+    group_by(author) |> 
+    mutate(share = n/sum(n)) |> 
+    ggplot(aes(author, container.title, fill = share)) +
+    geom_raster() +
+    geom_text(aes(label = n), color = 'white') +
+    scale_fill_viridis_c(option = 'A')
+
+## med vocab, k = 30 topics ----
+model = read_rds(here('data', '04_tm', '04_md_tmfast.Rds'))
+
+tidy(model, k = 30, matrix = 'gamma') |> 
+    filter(gamma > .05) |> 
+    inner_join(meta_df, by = c('document' = 'article_id')) |> 
+    count(author, topic) |> 
+    complete(author, topic, fill = list(n = 0)) |> 
+    inner_join(count(meta_df, author), by = 'author') |> 
+    mutate(share = n.x / n.y) |> 
+    ggplot(aes(topic, author, fill = share)) +
+    geom_raster() +
+    scale_fill_viridis_c(option = 'A')

### 2024-05-03 scripts/scratch/stm comparison.R

@@ -1,0 +1,28 @@
+library(stm)
+
+tic()
+test_stm = phrases_ar |> 
+    filter(phrase %in% vocabs$sm$phrase) |> 
+    mutate(n = as.integer(exp(n) - 1)) |> 
+    collect() |> 
+    build_matrix(article_id, phrase, n) |> 
+    stm(K = 30)
+toc()
+
+
+
+tmf = read_rds(here(data_dir, '04_tm', '04_sm_tmfast.Rds'))
+
+beta_tmf = tidy(tmf, k = 30, matrix = 'beta')
+beta_stm = tidy(test_stm, matrix = 'beta')
+
+source(here('R', 'compare_betas.R'))
+compare_betas(beta_tmf, rename(beta_stm, token = term), 
+              vocab = unique(vocabs$sm$phrase)) |> 
+    as.matrix() |> 
+    as_tibble(rownames = 'tmf_topic') |> 
+    pivot_longer(cols = -tmf_topic, 
+                 names_to = 'stm_topic', 
+                 values_to = 'distance') |> 
+    ggplot(aes(tmf_topic, stm_topic, fill = distance)) +
+    geom_raster()

### 2024-05-03 scripts/scratch/tmfast.R

@@ -1,0 +1,98 @@
+## This is super promising.  Relatively fast to fit, and the one really expensive operation (SVD) only needs to happen once to fit models across different values of k.  Looks like we'll need to go further than k = 10.  Try up through 50? 
+library(tidyverse)
+theme_set(theme_minimal())
+library(arrow)
+library(irlba)
+library(tidytext)
+
+library(tmfast)
+
+library(tictoc)
+library(here)
+
+data_dir = here('data')
+
+## Load metadata ----
+meta_ar = open_dataset(here(data_dir, '01_metadata'))
+
+set.seed(2022-02-14)
+subset_meta = meta_ar |> 
+    group_by(container.title) |> 
+    collect() |> 
+    slice_sample(n = 1000)
+
+subset_phr = open_dataset(here(data_dir, '01_phrases.csv'),
+                          format = 'csv') |>
+    filter(article_id %in% subset_meta$article_id)
+
+## Vocabulary selection ----
+## R(doc) draws docs inversely proportional to their length
+r_df = subset_phr |> 
+    group_by(article_id) |> 
+    summarize(len = sum(n)) |> 
+    collect() |> 
+    mutate(alpha = sum(len), 
+           r = len/alpha)
+
+## Information gain from R to tokens
+d_df = subset_phr |> 
+    left_join(r_df, by = 'article_id') |>
+    collect() |>
+    group_by(phrase) |> 
+    mutate(p = n/sum(n), 
+           H_term = - p * log2(p),
+           d_term = p * log2(p / r)) |> 
+    summarize(n = sum(n), 
+              H = sum(H_term),
+              d = sum(d_term)) |> 
+    mutate(nd = log2(n)*d) |> 
+    arrange(desc(nd)) # |> view()
+
+terms = d_df |> 
+    filter(H > 0) |> 
+    top_n(10*nrow(subset_meta), nd) |> 
+    pull(phrase)
+
+subset_phr |> 
+    filter(phrase %in% terms) |> 
+    count(article_id) |> 
+    inner_join(subset_meta, by = 'article_id') |> 
+    count(container.title) |> 
+    collect()
+
+
+## PCA ----
+dtm = subset_phr |> 
+    filter(phrase %in% terms) |> 
+    mutate(n = log1p(n)) |> 
+    collect() |> 
+    cast_sparse(row = article_id, col = phrase, value = n)
+
+## ~200 sec
+tic()
[…]

### 2024-05-03 scripts/scratch/topics over time.R

@@ -1,0 +1,94 @@
+library(tidyverse)
+theme_set(theme_minimal())
+library(tmfast)
+
+library(arrow)
+library(here)
+library(assertthat)
+library(glue)
+
+data_dir = here('data')
+tm_dir = here(data_dir, '04_tm')
+out_dir = here('out')
+
+model = here(tm_dir, '04_md_tmfast.Rds') |>
+    read_rds()
+meta_ar = open_dataset(here(data_dir, '01_metadata'))
+
+gamma_df = tidy_all(model, matrix = 'gamma')
+
+gamma_df |> 
+    filter(k == 30, 
+           topic %in% c('V05', 'V07', 'V19')) |> 
+    left_join(collect(meta_ar), 
+              by = c('document' = 'article_id')) |> 
+    ggplot(aes(container.title, gamma, 
+               color = container.title)) +
+    # geom_boxplot() +
+    geom_violin(draw_quantiles = .5, 
+                scale = 'width') +
+    scale_color_discrete(guide = 'none') +
+    facet_wrap(vars(topic), 
+               ncol = 1)
+
+gamma_df |> 
+    filter(k == 30, 
+           topic %in% c('V05', 'V07', 'V19')) |> 
+    left_join(collect(meta_ar), 
+              by = c('document' = 'article_id')) |> 
+    ggplot(aes(year, gamma, 
+               color = container.title)) +
+    # geom_point(alpha = .2) +
+    stat_summary(fun = median, geom = 'line') +
+    facet_wrap(vars(topic), 
+               ncol = 1)
+
+threshold = .1
+
+## TODO: properly handle missing values
+gamma_df |> 
+    filter(k == 30, 
+           topic %in% c('V05', 'V07', 'V19')) |> 
+    left_join(collect(meta_ar), 
+              by = c('document' = 'article_id')) |> 
+    filter(gamma > threshold) |> 
+    count(k, topic, year, container.title) |> 
+    group_by(k, topic, container.title) |>
+    mutate(smooth = slider::slide_index_dbl(n, year, 
+                                            mean, 
+                                            .before = 2, 
+                                            .after = 2, 
+                                            .complete = TRUE)) |> 
+    ungroup() |> 
+    ggplot(aes(year, 
+               color = container.title, group = container.title)) +
+    geom_line(aes(y = n), alpha = .2) +
+    geom_line(aes(y = smooth), size = 1) +
+    facet_wrap(vars(topic), 
+               scales = 'free_y',
+               ncol = 1) +
[…]

### 2024-05-03 scripts/scratch/topics_05_07_19.R

@@ -1,0 +1,272 @@
+library(tidyverse)
+theme_set(theme_minimal())
+library(tmfast)
+
+library(gt)
+
+library(arrow)
+library(here)
+library(glue)
+
+data_dir = here('data')
+tm_dir = here(data_dir, '04_tm')
+
+model = read_rds(here(tm_dir, '04_md_tmfast.Rds'))
+
+topics_of_interest = c('V05', 'V07', 'V19')
+
+phrases_ar = open_dataset(here(data_dir, '01_phrases.csv'), 
+                          format = 'csv')
+
+
+## Extra word-topic distributions ----
+beta_19 = tidy(model, 30, matrix = 'beta') |> 
+    filter(topic == 'V19') |> 
+    arrange(desc(beta)) |> 
+    top_n(250, beta) |> 
+    mutate(rank_19 = row_number())
+
+beta_05 = tidy(model, 30, matrix = 'beta') |> 
+    filter(topic == 'V05') |> 
+    arrange(desc(beta)) |> 
+    top_n(250, beta) |> 
+    mutate(rank_05 = row_number())
+
+beta_07 = tidy(model, 30, matrix = 'beta') |> 
+    filter(topic == 'V07') |> 
+    arrange(desc(beta)) |> 
+    top_n(250, beta) |> 
+    mutate(rank_07 = row_number())
+
+## 19 appears to be "race & intelligence"
+## 05 appears to be "intelligence testing"
+# full_join(beta_19, beta_05, by = 'token') |> 
+#     view()
+
+## 07 is "anthropological race science"; 19 is "psychological race science"? 
+# full_join(beta_07, beta_19, by = 'token') |> 
+#     view()
+
+# full_join(beta_07, beta_05, by = 'token') |> 
+#     view()
+
+
+## Compare term lists ----
+make_edges = function(df1, df2, by = 'token') {
+    df1 = df1 |> 
+        rename(rank = starts_with('rank'))
+    df2 = df2 |> 
+        rename(rank = starts_with('rank'))
+    
+    inner_join(df1, df2, by = by)
+}
+
+term_diff = bind_rows(
+    make_edges(beta_05, beta_07),
+    make_edges(beta_07, beta_19),
+    {make_edges(beta_19, beta_05) |> 
+            mutate(topic.y = str_c(topic.y, ' '))}) |> 
+    mutate(delta = rank.x - rank.y)
+
+ggplot(term_diff, aes(x = topic.x, xend = topic.y, 
[…]

### 2024-05-03 scripts/scratch/vocab.R

@@ -1,0 +1,128 @@
+library(tidyverse)
+theme_set(theme_minimal())
+library(arrow)
+
+library(here)
+
+data_dir = here('data')
+
+## Load metadata ----
+meta_ar = open_dataset(here(data_dir, '01_metadata'))
+
+set.seed(2022-02-14)
+subset_meta = meta_ar |> 
+    group_by(container.title) |> 
+    collect() |> 
+    slice_sample(n = 250)
+
+subset_phr = open_dataset(here(data_dir, '01_phrases.csv'), 
+                          format = 'csv') |> 
+    filter(article_id %in% subset_meta$article_id) |> 
+    collect()
+
+## Entropy ----
+## Pr(document | phrase)
+# subset_phr |> 
+#     group_by(phrase) |> 
+#     mutate(p = n/sum(n))
+# 
+# ## Conditional entropy
+# subset_phr |> 
+#     group_by(phrase) |> 
+#     mutate(p = n/sum(n),
+#            H_term = -p*log2(p)) |> 
+#     summarize(H = sum(H_term)) |> 
+#     arrange(H)
+# 
+# ## Information gain
+# max_H = log2(nrow(subset_meta))
+# 
+# subset_phr |> 
+#     group_by(phrase) |> 
+#     mutate(p = n/sum(n),
+#            H_term = -p*log2(p)) |> 
+#     summarize(H = sum(H_term)) |> 
+#     mutate(delta_H = max_H - H) |> 
+#     arrange(desc(delta_H))
+# 
+# ## ndH
+# H_df = subset_phr |> 
+#     group_by(article_id) |> 
+#     mutate(doc_len = sum(n)) |> 
+#     group_by(phrase) |> 
+#     mutate(p = n/sum(n), 
+#            H_term = -p*log2(p)) |> 
+#     summarize(n = sum(n), 
+#               H = sum(H_term)) |> 
+#     mutate(delta_H = max_H - H, 
+#            ndH = log10(n)*delta_H) |> 
+#     arrange(desc(ndH))
+# 
+# # terms = c('sleep', 'dreams', 'rem_sleep', 'creativity')
+# terms = H_df |> 
+#     head(24) |> 
+#     pull(phrase)
+# 
+# subset_phr |> 
+#     filter(phrase %in% terms) |> 
+#     #view() |> 
+#     left_join(subset_meta, by = 'article_id') |> 
+#     ggplot(aes(article_id, n, fill = container.title)) +
[…]

### 2024-05-03 scripts/scratch/vsp.R

@@ -1,0 +1,128 @@
+library(tidyverse)
+theme_set(theme_minimal())
+library(vsp)
+
+library(arrow)
+
+library(here)
+
+data_dir = here('data')
+
+## Two helper functions ----
+trim_negative = function(x) {
+    x[x < 0] = 0
+    return(x)
+}
+
+normalize = function(x) {
+    if (sum(x) != 0) {
+        x = x / sum(x)
+    }
+    return(x)
+}
+
+## Load metadata ----
+meta_ar = open_dataset(here(data_dir, '01_metadata'))
+
+set.seed(2022-02-14)
+subset_meta = meta_ar |> 
+    group_by(container.title) |> 
+    collect() |> 
+    slice_sample(n = 1000)
+
+subset_phr = open_dataset(here(data_dir, '01_phrases.csv'), 
+                          format = 'csv') |> 
+    filter(article_id %in% subset_meta$article_id) |> 
+    collect()
+
+
+## Vocabulary selection ----
+## R(doc) draws docs inversely proportional to their length
+r_df = subset_phr |> 
+    group_by(article_id) |> 
+    summarize(len = sum(n)) |> 
+    collect() |> 
+    mutate(alpha = sum(len), 
+           r = len/alpha)
+
+## Information gain from R to tokens
+d_df = subset_phr |> 
+    left_join(r_df, by = 'article_id') |> 
+    collect() |> 
+    group_by(phrase) |> 
+    mutate(p = n/sum(n), 
+           H_term = - p * log2(p),
+           d_term = p * log2(p / r)) |> 
+    summarize(n = sum(n), 
+              H = sum(H_term),
+              d = sum(d_term)) |> 
+    mutate(nd = log2(n)*d) |> 
+    arrange(desc(nd)) # |> view()
+
+terms = d_df |> 
+    filter(H > 0) |> 
+    top_n(10*nrow(subset_meta), nd) |> 
+    pull(phrase)
+
+subset_phr |> 
+    filter(phrase %in% terms) |> 
+    count(article_id) |> 
+    inner_join(subset_meta, by = 'article_id') |> 
+    count(container.title) |> 
+    collect()
+
+## Using vsp ----
+## <https://rss.org.uk/RSS/media/File-library/Events/Discussion%20meetings/Preprint_Rohe_Zeng_11-May-2022.pdf>
+tic()
+vsp_fit = subset_phr |> 
+    filter(phrase %in% terms) |> 
+    collect() |> 
[…]

### 2024-05-03 scripts/11-proof-of-concept.qmd

@@
+---
+format:
+  - html
+  - pdf: 
+      mathspec: true
+---
+
 This section uses simulated data to provide a proof-of-concept for our use of topic models to detect the "mainstreaming" of a marginalized topic.
@@
 library(tidyverse)
+library(ggforce)
 theme_set(theme_bw())
@@
 This gives us three sets of journals: A, B (0-5), and C-L.
-In the LDA model, these journal sets correspond to $\mathbf{\alpha}$, the parameter for the "prior" Dirichlet distribution $\Theta(\mathbf{\alpha})$.
-The `tmfast` package provides a helper function `peak_alpha()` for constructing $\mathbf{\alpha}$.
+In the LDA model, these journal sets correspond to $\boldsymbol{\alpha}$, the parameter for the "prior" Dirichlet distribution $\Theta(\boldsymbol{\alpha})$.
+The `tmfast` package provides a helper function `peak_alpha()` for constructing $\boldsymbol{\alpha}$.
 
@@
 
-Each paper has a simplex $\theta \sim \Theta(\mathbf{\alpha})$, representing a distribution over topics. We draw `r N` papers from each journal.
+Each paper has a simplex $\theta \sim \Theta(\boldsymbol{\alpha})$, representing a distribution over topics. We draw `r N` papers from each journal.
 
@@
 
-head(theta)
+theta[1:5, 1:5]
 
@@
 ## Fit topic model ----
+tic()
 fitted = tmfast(corpus, c(10, 15, 20, 25, 30))
+toc()
 ```
@@
 ```{r}
+#| fig-width: 7
+#| fig-height: 4
+#| fig-cap: "Fitted topic-document distributions for the proof-of-concept simulation."
 ## Topics by journal ----
@@
     ggplot(aes(document, topic, fill = gamma)) +
-    geom_tile() +
-    facet_wrap(vars(journal), scales = 'free', 
+    geom_raster() +
+    facet_wrap(vars(journal), scales = 'free_x', 
                ncol = 7) +
-    scale_x_discrete(guide = 'none')
+    scale_y_discrete(breaks = c('V01', 'V10', 'V20')) +
+    scale_x_discrete(breaks = 'none')
 ```
@@
 ```{r}
-## Topic model clearly detects the focal topic in A, increase in B over time, minimal presence in other journals
+#| fig-cap: "In the proof-of-concept simulation, the topic model clearly detects the marginal topic in journal A, increase in journal B over time, and minimal presence in other journals"
 tidy(fitted, k = 20, matrix = 'gamma') |> 
@@
 
-While the journals are largely distinguishable on the other topics, except for the unusually low value for A. 
+While the journals are largely not distinguishable from each other on the other topics, except for the unusually low value for A.
+
 ```{r}
-## With no pattern for the other topics
+#| fig-cap: "In the proof-of-concept simulation, the topic model does not produce a false negative pattern with the mainstream topics"
 tidy(fitted, k = 20, matrix = 'gamma') |> 
@@
     ggplot(aes(journal, gamma)) +
-    ggforce::geom_sina() +
+    geom_sina() +
     stat_summary(color = 'red')
@@
 ```{r}
+#| tbl-cap: "In the proof-of-concept simulation, almost all articles in journal A contain the marginal topic; the rate in B increases over time; and the marginal topic does not appear in the other journals"
 tidy(fitted, k = 20, matrix = 'gamma') |> 
@@
     add_journal() |> 
-    count(journal) |> 
-    mutate(rate = n / N)
+    count(journal, .drop = FALSE) |> 
+    mutate(rate = n / N) |> 
+    knitr::kable()
 ```
@@
 As a limitation, we note that, for $\theta > 0.05$ or so, the estimated topic probabilities are below the true values.
+
 ```{r}
+#| fig-cap: "In the proof-of-concept simulation, without normalization the estimated topic-document probabilities (gamma) are well below the true probabilities (theta) for approx. $\\theta > 0.05$"
 tidy(fitted, k = 20, matrix = 'gamma') |> 
@@
                   inherit.aes = FALSE, color = 'black') +
-    scale_color_viridis_d()
+    scale_color_viridis_d(guide = 'none')
 ```

### 2024-05-03 scripts/13_pf_authors.qmd

This section examines topic-document distributions at the author level, focusing on Pioneer-funded authors.  First, a separate helper script uses regular expressions to identify Pioneer-funded authors in the mainstream journal and MQ metadata.

```{r}
library(tidyverse)
theme_set(theme_bw())
library(tmfast)

library(arrow)
library(here)

data_dir = here('data')
tm_dir = here(data_dir, '04_tm')

## Load data ----
source(here('R', 'pf_authors.R'))

meta_df = open_dataset(here(data_dir, '01_metadata')) |> 
    mutate(article_id = tolower(article_id)) |> 
    inner_join(pf_authors, by = 'article_id') |>
    collect()

## Available documents by author
count(meta_df, pf_author)
```

We then load the topic models.  For this supplemental analysis, we're going to focus on the medium vocabulary models.  We use pre-calculated normalization exponents for the topic-document distributions.

```{r}
## Load topic models ----
model_files = list.files(tm_dir, '*_tmfast.Rds') %>%
    here(tm_dir, .) |> 
    set_names('lg', 'md', 'sm')
exp_files = list.files(data_dir, '*_exponents.Rds') %>%
    here(data_dir, .) |> 
    set_names('lg', 'md', 'sm')
## Helper to read exponents as a named num
read_exp = function(file) {
    file |> 
        read_rds() |> 
        pull(exponent, name = k)
}

this_vocab = 'md'

this_model_file = model_files[this_vocab]
this_model = read_rds(this_model_file)
this_exp = exp_files[this_vocab] |> 
    read_exp()

gamma_df = tidy_all(this_model, 
                    matrix = 'gamma') |> 
    group_split(k) |> 
    map2_dfr(this_exp, ~ renorm(.x, document, gamma, .y)) |> 
    inner_join(meta_df, 
               by = c('document' = 'article_id'), 
               multiple = 'all')
```

A tile plot of across all values of $k$ is difficult to read, but suggests that individual researchers often work in just a few topics.

```{r}
#| fig.height: 8
#| fig.width: 8
#| fig.cap: "Topic-document distributions, medium vocabulary, all values of k, by Pioneer-funded author"
## Tile plot ----
ggplot(gamma_df, aes(topic, document, fill = gamma)) +
    geom_raster() +
    geom_vline(xintercept = c(10, 20, 30, 40) + .5,
               color = 'black') +
    facet_grid(rows = vars(pf_author), 
               cols = vars(k),
               scale = 'free', 
               switch = 'y',
               labeller = label_wrap_gen(width = 15), 
               # space = 'free_y'
               ) +
    scale_x_discrete(breaks = c('V10', 'V20', 'V30', 'V40', 
                                'V50', 'V60')) +
    scale_y_discrete(breaks = NULL,
                     name = '') +
    scale_fill_viridis_c(option = 'viridis', direction = 1) +
    theme_minimal() +
    theme(axis.ticks.x = element_line(),
          legend.position = 'bottom',
          plot.margin = margin(r = 10), 
          strip.text.y.left = element_text(angle = 0))
```

Focusing on $k=40$ suggests both overlaps and differences across authors.

```{r}
#| fig.height: 4
#| fig.width: 7
#| fig.cap: "Topic-document distributions, medium vocabulary, $k=40$, by Pioneer-funded author"
gamma_df |> 
    filter(k == 40) |> 
    ggplot(aes(topic, document, fill = gamma)) +
    geom_raster() +
    geom_vline(xintercept = c(10, 20, 30, 40) + .5,
               color = 'black') +
    facet_wrap(vars(pf_author), 
               scale = 'free', 
               labeller = label_wrap_gen(width = 15), 
               # space = 'free_y'
    ) +
    scale_x_discrete(breaks = c('V10', 'V20', 'V30', 'V40')) +
    scale_y_discrete(breaks = NULL,
                     name = '') +
    scale_fill_viridis_c(option = 'viridis', direction = 1) +
    theme_minimal() +
    theme(axis.ticks.x = element_line(),
          legend.position = 'bottom',
          plot.margin = margin(r = 10))
```

In particular, for $k=40$, all Pioneer-funded authors publish a significant number of papers in at least one of the focal topics for this model:  V07 (MQ race), V22 (mainstream intelligence), and V24 (race-and-intelligence).

```{r}
#| fig.height: 4
#| fig.width: 6
#| fig.cap: "Topic-document distributions, for $k=40$ and medium vocabulary, by Pioneer-funded author. Each line represents the topic distribution for a single document. The use of transparency means that the darkness of a line corresponds to the frequency of that topic in the author's documents. Blue vertical lines indicate the focal topics for this model."
gamma_df |> 
    filter(k == 40) |> 
    mutate(author_count = n(), .by = pf_author) |> 
    ggplot(aes(topic, gamma, group = document)) +
    geom_line(aes(alpha = 1/author_count), show.legend = FALSE) +
    geom_vline(xintercept = c('V07', 'V22', 'V24'), 
               color = 'blue', alpha = 1) +
    facet_wrap(vars(pf_author)) +
    scale_x_discrete(breaks = c('V10', 'V20', 'V30'))
```

```{r}
## Document max count ----
## Due to filtering, topic breaks look weird
gamma_df |> 
    group_by(k, pf_author, document) |> 
    filter(gamma == max(gamma)) |> 
    group_by(k, pf_author) |> 
    count(topic) |> 
    ungroup() |> 
    # complete(pf_author, nesting(k, topic), 
    #          fill = list(n = 0L)) |> 
    ggplot(aes(topic, n)) +
    # geom_vline(xintercept = c('V10', 'V20', 'V30', 'V40', 
    #                           'V50', 'V60'), 
    #            alpha = .1) +
    geom_col() +
    facet_grid(rows = vars(pf_author), 
               cols = vars(k), 
               scales = 'free') +
    scale_x_discrete(breaks = c('V10', 'V20', 'V30', 'V40', 
                                'V50', 'V60')) +
    scale_y_continuous(breaks = scales::pretty_breaks(3)) +
    theme(strip.text.y = element_text(angle = 0))

ggsave(here('out', '13_doc_max.png'), 
       height = 8, width = 8, bg = 'white', scale = 2)

gamma_df |> 
    filter(k == 40) |> 
    group_by(pf_author, document) |> 
    filter(gamma == max(gamma)) |> 
    group_by(pf_author) |> 
    count(topic) |> 
    complete(topic = {1:40 |> 
            str_pad(2, pad = '0') %>% 
            str_c('V', .)}, 
            fill = list(n = 0L)) |> 
    ggplot(aes(topic, n)) +
    geom_vline(xintercept = c('V07', 'V22', 'V24'), 
               color = 'blue', alpha = .5) +
    geom_col() +
    facet_wrap(vars(pf_author), 
               scales = 'free_y') +
    scale_x_discrete(breaks = c('V10', 'V20', 'V30')) +
    scale_y_continuous(breaks = scales::pretty_breaks())

## Document threshold count ----
gamma_df |> 
    filter(k == 40) |> 
    group_by(pf_author, document) |> 
    filter(gamma > .5) |> 
    group_by(pf_author) |> 
    count(topic) |> 
    complete(topic = {1:40 |> 
            str_pad(2, pad = '0') %>% 
            str_c('V', .)}, 
            fill = list(n = 0L)) |> 
    ggplot(aes(topic, n)) +
    geom_vline(xintercept = c('V07', 'V22', 'V24'), 
               color = 'blue', alpha = .5) +
    geom_col() +
    facet_wrap(vars(pf_author), 
               scales = 'free_y') +
    scale_x_discrete(breaks = c('V10', 'V20', 'V30', 'V40'))
```

### 2024-05-03 scripts/scratch/vocab.Rmd

---
title: "How to choose words to do things with"
author: Dan Hicks
output:
    html_document:
        code_folding: show
        toc: true
        toc_float: true
---

After extracting text from the documents in our corpus, we typically want to work with a small subset of tokens (text units).  Extremely common terms, like "it" and "she," don't carry much information; and similarly with misspelled words or OCR errors.  Identifying the subset of tokens to use in further analysis is called *vocabulary selection*.  This document will discuss the problems with the widely-used TF-IDF method for vocabulary selection, then present my information-theoretic alternative.

# Setup #

We'll use a small part of the combined race science corpus, with 50 articles from each of the 5 journals.

```{r, setup, echo = 'hide', results = 'hide', message = FALSE, warning = FALSE}
library(tidyverse)
theme_set(theme_bw())
library(tidytext)

library(arrow)
library(here)

data_dir = here('data')
```
```{r, load-data}
meta_ar = open_dataset(here(data_dir, '01_metadata'))

set.seed(2022-02-14)
subset_meta = meta_ar |> 
    group_by(container.title) |> 
    collect() |> 
    slice_sample(n = 50)

count(subset_meta, container.title)

subset_phr = open_dataset(here(data_dir, '00_phrases')) |> 
    filter(article_id %in% subset_meta$article_id) |> 
    collect()
```
```{r}
subset_meta
subset_phr
```

# TF-IDF #

TF-IDF stands for *token frequency*-*inverse document frequency*.  Token frequency is the relative occurrence of the given token in the given document; inverse document frequency is the log ratio of documents in the corpus to documents that contain the token.  TF-IDF is the product of these two.  TF-IDF is widely used for vocabulary selection in text mining, and [has a one-liner in the `tidytext` package](https://www.tidytextmining.com/tfidf.html).

```{r, tf-idf}
bind_tf_idf(subset_phr, phrase, article_id, n) |> 
    arrange(desc(tf_idf))
```

This example demonstrates two problems with TF-IDF.  First, the TF-IDF value is associated with a token *in a document*, not a token across the entire corpus.  After calculating TF-IDF scores, a common next step is to keep the top $n$ tokens for each document.  Suppose a token $t$ appears in the top $n$ tokens for document 1, and as the $n+1$th token in documents $2, 3, \ldots, k$.  Token $t$ would be eliminated in the other documents, as though it never occurred in them at all.  So $t$ would be in the analysis vocabulary — but only for document 1.  This creates a weird discretization, and it's not hard to imagine that it might lead to artifacts.

The second problem is that, when taken across the entire corpus, TF-IDF favors rare terms from short documents.  Here's the entire text of [`10.1023/a:1017357211509`](https://link.springer.com/content/pdf/10.1023/A:1017357211509.pdf), which has the top-scoring token according to TF-IDF, `john_k_hewitt`:

> The editors of Behavior Genetics are pleased to
> announce that John K. Hewitt, University of Colorado,
> Boulder, assumed the position of Executive Editor of
> the journal in July 2001. He is now receiving new
> manuscripts for the journal. Contributors should send
> manuscripts to [mailing address]

This document only has a few noun phrases, so `john_k_hewitt` has a high term frequency; and the idf score indicates that it appears in only one documents in the (small) corpus: 
```{r}
nrow(subset_meta)/exp(5.5214609)
```

# The document guessing game #

To develop a better approach, suppose we played a game where I pick a document from the corpus, and you had to guess which one I picked.  I give you a clue: a single token from the document.  Highly informative tokens would be ones that significantly narrowed down the possible documents.  We can express this quantitatively using [Kullback-Leibler divergence](https://en.wikipedia.org/wiki/Kl-divergence) or *KL divergence*, a measure of "distance" between two probability distributions.

Let $P$ and $Q$ be two (discrete) probability distributions on the same space $X$.  The KL divergence from $Q$ to $P$ is defined as
$$ D_{KL}(P || Q) = \sum_{x \in X} P(x) \log \left( \frac{P(x)}{Q(x)} \right) $$
KL divergence is also called *information gain*, and in this sense is interpreted as the increase in information (decrease in entropy) when we move from $Q$ to $P$.

(Note that KL divergence is asymmetric, with $D(P||Q) \neq D(Q||P)$.  KL divergence isn't a [metric](https://en.wikipedia.org/wiki/Metric_(mathematics)), and doesn't have other properties we usually associate with distance.)

```{r}
subset_phr |> 
    filter(phrase == 'evolution') |> 
    ggplot(aes(article_id, n)) +
    geom_col() +
    scale_x_discrete(breaks = NULL)
```

To use KL divergence with our document guessing game, let $Q$ be the uniform distribution across the $N$ documents in the corpus: $Q(d) = \frac{1}{N}$ for all documents $d$.  Then, given a token $t$, let $P_t(d) = Pr(d | t)$ be the share of all occurrences of token $t$ (across all documents) that appear in document $d$.  For example, if $t$ occurs 5 times in $d$ and 25 times across the entire corpus, then $P(d) = 5/25 = 0.2$.  In our game, suppose I pick the document uniformly at random from the corpus, then pick the token uniformly at random from all token occurrences in that document (so that common tokens are more likely to be picked than uncomon ones).  Then KL divergence tells us how informative $t$ is.  
\begin{align*}
D_{KL}(P_t || Q) &= \sum_{x \in X} P_t(x) \log \left( \frac{P_t(x)}{Q(x)} \right) \\
      &= \sum P_t(x) \left(\log P_t(x) - \log \frac{1}{N}\right) \\
      &= \left[ \sum P_t(x) \log P_t(x) \right] + \left[\log N \sum P_t(x)\right] \\
      &= [-H(P)] + [\log N] \\
      &= \log N - H(P)
\end{align*}
where $H$ is the (Shannon) entropy of $P$.

To translate this into R, first let's calculate $P_t(d)$. 
```{r}
subset_phr |> 
    group_by(phrase) |> 
    mutate(p = n/sum(n)) |> 
    select(article_id, phrase, n, p) |> 
    filter(phrase == 'evolution')
```

Next we extend this to calculate $H(P_t)$ and information gain, which we represent as $\Delta H$. 
```{r}
max_H = log2(nrow(subset_meta))

H_df = subset_phr |> 
    group_by(phrase) |> 
    mutate(p = n/sum(n),
           H_term = -p*log2(p)) |> 
    summarize(H = sum(H_term)) |> 
    mutate(delta_H = max_H - H) |> 
    arrange(desc(delta_H))

head(H_df, 20)

H_df |> 
    filter(phrase == 'evolution')
```

Whoops, lots of garbage.  The highest information gain comes from tokens that only appear in a single document.  They're maximally informative in our game, but are semantically meaningless and don't correspond to topics that occur across the corpus. Filtering to $H > 0$ eliminates unique tokens, and initially seems to help.

```{r}
H_df |> 
    filter(H > 0) |> 
    head(20)
```

But these terms still only appear in *approximately* one document.

```{r}
subset_phr |> 
    filter(phrase %in% c('solms', 'dreaming', 'grodzinsky')) |> 
    arrange(phrase, desc(n))
```

# `ndH` #

KL divergence plays a role similar to IDF:  tokens that only occur in a few documents generally have high KL divergence as well as high IDF.  We need to balance this "selectivity" with a measure of how frequently tokens appear across the corpus.  This will be analogous to TF; but instead of counting within each document, we'll use the total number of occurrences across the corpus.  To get this frequency count on the same scale as entropy, we'll take its log.  I refer to the resulting measure as $\log n \Delta H$ or `ndH`.

```{r}
ndH_df = subset_phr |> 
    group_by(phrase) |> 
    mutate(p = n/sum(n),
           H_term = -p*log2(p)) |> 
    summarize(n = sum(n), 
              H = sum(H_term)) |> 
    filter(H > 0) |> 
    mutate(delta_H = max_H - H, 
           ndH = log2(n)*delta_H) |> 
    arrange(desc(ndH))

head(ndH_df, 20)
```

This looks good at first, and in previous text mining projects it's done well.  But notice how many of the top terms have to do with sleep?  Let's plot frequency counts for the top 25 terms.  We'll put the frequencies on a square root scale, so it's easier to see both large and small-but-nonzero values.

```{r, fig.width = 10, fig.height = 8, out.width = "100%"}
terms_ndH = ndH_df |> 
    pull(phrase) |> 
    magrittr::extract(1:25)

subset_phr |> 
    filter(phrase %in% terms_ndH) |> 
    left_join(subset_meta, by = 'article_id') |> 
    ggplot(aes(article_id, n, fill = container.title)) +
    geom_col() +
    facet_wrap(vars(phrase), scales = 'free_x') +
    scale_x_discrete(breaks = NULL) +
    scale_y_sqrt() +
    coord_flip() +
    theme(axis.text.y=element_blank())
```

Because the *Behavioral and Brain Science* articles (actually, collections of articles + commentaries) are so long, tokens that occur in just a few of these articles can still have large frequencies.  The high-`ndH` tokens come from this weird little corner of the distribution of tokens.

```{r}
ggplot(ndH_df, aes(log2(n), delta_H)) +
    geom_bin_2d(bins = 40) +
    stat_function(fun = \(x) 55/x, xlim = c(6, 13)) +
    scale_fill_viridis_c(trans = 'log10')
```

# Changing the game #

Consider what the two terms in the construction of `ndH` are supposed to be doing.  Frequency is supposed to pick out tokens that occur widely across the corpus, while information gain is supposed to pick out tokens that are distinctive to different documents.  But, because the *Behavior and Brain Sciences* documents are so long, their distinctive tokens also end up with a high frequency.  To balance this out, we can adjust information gain, penalizing tokens that come from larger documents.

We do this by changing the background conditions of the game.  Suppose I pick the document not uniformly at random, but instead in proportion to length.  Then you'll gain more information from tokens that come from shorter (less likely) documents.

More formally, let $R(d) = n_d / \alpha$, where $n_d$ is the total number of tokens in document $d$ and $\alpha = \sum_d n_d$ is the total number of tokens across the entire corpus.  We then calculate information gain for token $t$ using $P_t$ and KL divergence, but from $R$ instead of $Q$.  
$$ D_{KL}(P_t || R) = \sum_{d} P_t(d) \log \left( \frac{P_t(d)}{R(d)} \right) $$
I don't think this one simplifies nicely, so we'll just use the right-hand side directly in the calculation.

First we need to calculate $R(d)$. 
```{r}
r_df = subset_phr |> 
    group_by(article_id) |> 
    summarize(len = sum(n)) |> 
    mutate(alpha = sum(len), 
           r = len/alpha)

r_df

r_df |> 
    left_join(subset_meta, by = 'article_id') |> 
    ggplot(aes(len, r, color = container.title)) +
    geom_point() +
    scale_x_log10() +
    scale_y_log10()
```

Then we calculate the divergence and, as with `ndH`, multiply by log frequency.  
```{r}
d_df = subset_phr |> 
    left_join(r_df, by = 'article_id') |> 
    group_by(phrase) |> 
    mutate(p = n/sum(n), 
           H_term = - p * log2(p),
           d_term = p * log2(p / r)) |> 
    summarize(n = sum(n), 
              H = sum(H_term),
              delta_H = max_H - H,
              d = sum(d_term)) |> 
    filter(H > 0) |> 
    mutate(nd = log2(n)*d) |> 
    arrange(desc(nd))

head(d_df, 20)
```

The plot of these two components doesn't have the weird peninsula in the upper-right.  Instead the highest values come from combining moderate information gain with moderate frequency, which is probably what we want.

```{r}
ggplot(d_df, aes(log2(n), d)) +
    geom_bin_2d(bins = 40) +
    stat_function(fun = \(x) 40/x, xlim = c(3, 13)) +
    scale_fill_viridis_c(trans = 'log10')
```

We can also compare $\Delta H$ to information gain from $R$.  They're correlated; but high $\Delta H$ corresponds to a wide range of values with this new information gain calculation.  Distinctive terms from large documents are nudged down, while distinctive terms from small documents are nudged up.  
```{r}
ggplot(d_df, aes(delta_H, d)) +
    geom_bin_2d(bins = 40) +
    stat_function(fun = identity) +
    scale_fill_viridis_c(trans = 'log10')
```

Plotting frequencies of the top 25 terms indicates that this approach is picking out terms from across the different journals in the corpus.  They also appear to correspond to different topics, rather than the excess emphasis on sleep and dreaming that we got using `ndH`.

```{r, fig.width = 10, fig.height = 8, out.width = "100%"}
terms_nd = d_df |> 
    pull(phrase) |> 
    magrittr::extract(1:25)

subset_phr |> 
    filter(phrase %in% terms_nd) |> 
    left_join(subset_meta, by = 'article_id') |> 
    ggplot(aes(article_id, n, fill = container.title)) +
    geom_col() +
    facet_wrap(vars(phrase), scales = 'free_x') +
    scale_x_discrete(breaks = NULL) +
    scale_y_sqrt() +
    coord_flip() +
    theme(axis.text.y=element_blank())
```

### 2024-05-03 paper/paper.qmd
various todos

@@
 
-acknowledgements: |
-  | Thanks to Emily Merchant and John Jackson for some initial discussions that helped clarify the scope of the project and identify some key resources related to the Pioneer Fund and *Mankind Quarterly*.
-Thanks to Anthony Sainez for help retrieving the *Mankind Quarterly* articles.
-Thanks to Derek Devnich and James Dooley for their work attempting to secure the APA-published articles.
-Thanks to Kevin Bird, John Jackson, and Manuela Fernández Pinto for comments on a draft of this paper.
-  | **Funding:** EJCL's work on this project was partially supported by UC Merced.  DJH's work on this project was not supported by any specific funding. 
-  | **Author contributions:** DJH (conceptualization, data curation, formal analysis, investigation, methodology, project administration, software, validation, visualization, writing-original draft, writing-review & editing); EJCL (conceptualization, methodology, software, validation, visualization, writing-original draft, writing-review & editing)
-  | **Competing interests:** The authors have no competing interests to declare. 
-  | **Data and materials availability:** Web of Science search results and article fulltext are not publicly available due to intellectual property restrictions.
-Preprocessed analysis data (document-term counts and article metadata) and analysis scripts are available at *[doi]*.
-
 bibliography: [references.bib, refs-manual.bib]
-# cite-method: "citeproc"
-# csl: "nature-no-superscript.csl"
+csl: "pnas.csl"
 
@@
     fig.env: "figure*"
+    number-sections: false
 
@@
 <!-- 
-PNAS guidelines
+PNAS guidelines: <https://www.pnas.org/author-center/submitting-your-manuscript#article-types>
+
 Research Report/Direct Submission: The preferred length of these articles is 6 pages, but PNAS allows articles up to a maximum of 12 pages.
@@
 - NHB reviews todo
-- sort git tree
-- reorganize into PNAS ms + SI
-- check cite formatting
+- de-anonymize
 - copy updated plots? 
-- abstract formatting
 - figure sizes
+- SI xrefs
 ]*
 
-<!-- # Introduction {.unnumbered} -->
+# Significance
+
+*[120 words, "undergrad-educated scientist outside their field of specialty"]*
+
+# Introduction
 
@@
 *Race science discourse* refers to a broader category, any treatment of race science as a legitimate area of scientific research.
-This includes methodological and empirical critiques of race science.
-For example, the Flynn effect (the finding that IQ scores have increased over time) poses a serious empirical challenge to hereditarian claims of racial differences in intelligence.
-But technical debates over the extent and causes of the Flynn effect likely create the impression among the general public of ordinary scientific disagreement.
-In this way technical critiques of race science based on the Flynn effect could potentially legitimize race science.
+
+We stress that race science discourse includes methodological and empirical critiques of race science.
+We argue that such technical critiques — even when they are intended to undermine race science — can instead legitimize it, by creating an impression among the broader public that race science is a matter of ordinary, mainstream scientific debate.
+For example, the Flynn effect (the finding that IQ scores have increased over time) poses a serious empirical challenge to hereditarian claims of racial differences in intelligence, and has been extensively debated in the mainstream scientific literature.
+Most members of the general public lack the time and skills needed to assess the evidence and arguments concerning the Flynn effect.
+Instead, we suggest, they are likely to understand the Flynn effect — and, by extension, race science — as an area of active, mainstream scientific research and ordinary scientific disagreement.
+In this way technical critiques of race science based on the Flynn effect can legitimize race science.
 So empirical research supporting the Flynn effect is usually not itself race science, but can still be race science discourse.
 
-These distinction allow us to bracket the intentions and attitudes of the researchers involved in any particular piece of research or debate.
+The distinction between scientific racism, race science, and race science discourse allows us to bracket the intentions and attitudes of the researchers involved in any particular piece of research or debate.
 For the purposes of the current study, it is not important whether, for instance, psychologist Arthur Jensen is or is not correctly labelled a scientific racist [@JacksonJr.ArthurJensenEvolutionary2022].
@@
 
-<!-- # Scientific racism, eugenics, and *Brown* {.unnumbered} -->
+## Scientific racism, eugenics, and *Brown* {.unnumbered}
 
@@
 
-In light of the scientific racist origins of Pioneer and MQ, and the significant attention MQ has received in recent historiography, we hypothesized that bibliometric and text mining analyses would show race science ideas originating in MQ, and from there being disseminated to mainstream publications.
+In light of the scientific racist origins of Pioneer and MQ, and the significant attention MQ has received in recent historiography, we hypothesized that bibliometric and text mining analyses would show race science discourse originating in MQ, and from there being disseminated to mainstream publications.
 This hypothesis was not supported.
-Instead, mainstream psychology journals provided a venue for a distinct form of race science, based on intelligence research and originating outside of MQ.
+Instead, mainstream psychology journals provided a venue for a distinct form of race science discourse, based on intelligence research and originating outside of MQ.
 
@@
 
-Figure @fig-focal and this discussion focuses on three topics identified in one of the 24 fitted topic models; see figures [@fig-gamma-lg; @fig-gamma-md; @fig-gamma-sm; @fig-silge-lg; @fig-silge-md; @fig-silge-sm] for all topics from all models, and figures [@fig-grid-lg; @fig-grid-md; @fig-grid-sm] for selected topics from all models.
+Figure @fig-focal focuses on three topics identified in one of the 24 fitted topic models; see figures [@fig-gamma-lg; @fig-gamma-md; @fig-gamma-sm; @fig-silge-lg; @fig-silge-md; @fig-silge-sm] for all topics from all models, and figures [@fig-grid-lg; @fig-grid-md; @fig-grid-sm] for selected topics from all models.
 In the focal model, topic 07 is strongly associated with MQ:  MQ published dozens of articles in this topic each year, and no other journal ever published more than a handful.
@@
 However, the other authors named in this topic — John Horn and Peter Bentler — do not appear to have contributed to scientific racism.
-Instead this topic appears to identify "mainstream" (non-race science) intelligence research, especially factor analysis and the debate over whether intelligence is unidimensional or multidimensional.
+Instead this topic appears to identify "mainstream" (non-race science discourse) intelligence research, especially factor analysis and the debate over whether intelligence is unidimensional or multidimensional.
 (Three other topics only associated with mainstream intelligence journals and terms were also identified by this topic model.)  This topic indicates that the model is able to distinguish "mainstream" intelligence research from race science discourse.
@@
 The Silge plot for topic 24, by contrast, suggests a distinct race science discourse topic, with multiple racial terms (`whites`, `blacks`, `racial_differences`, `races`, `race_differences`) and the names of three prominent Pioneer-funded researchers, `jensen`, `lynn`, and `rushton`.
-Independent qualitative coding of 121 papers in this topic (those with $\gamma > 0.97$ and published in journals other than PR) confirmed this interpretation, with 108 (89%) coded as race science discourse by both authors (Cohen's $\kappa = 0.86$).
+Independent qualitative coding of 121 papers in this topic (those with $\gamma > 0.97$ and published in journals other than PR) confirmed this interpretation, with 108 (89%) coded as race science discourse by both authors (Cohen's $\kappa = 0.86$); see @sec-topic-quality.
 
@@
 
-## Race science, behavior genetics, and psychology {.unnumbered}
+*[across all models]*
+
+- As $k$ increases, multiple different "intelligence only" topics
+- lg, md: usually 1 "race only" topic and/or 1 race-and-intelligence topic
+    - "race only" always almost exclusively in MQ
+    - md: race-and-intelligence predominantly in mainstream
+    - lg: for k=30, race-and-intelligence predominantly MQ; no race-int for k=40, 50; then similar to md for k=60, 70
+- sm: 
+    - seems to have a distinct "race only" topic in PR (V04)
+        - top terms include race_differences, rushton, crime: race science, but poss. either not about intelligence or int-related terms didn't make it into top terms
+    - other one is almost exclusive to MQ
+    - except for k=40, race-int topic, predominantly mainstream
+
+## Race science discourse, behavior genetics, and psychology {.unnumbered}
 
@@
 However, the topic model analysis suggests that, by the 1990s, behavior genetics may have distanced itself somewhat from race science, albeit without directly repudiating it.
-This interpretation was supported by a supplemental analysis that focused on BG specifically (see methods and materials)<!-- [@sec-bg] -->.
+This interpretation was supported by a supplemental analysis that focused on BG specifically; see @sec-bg.
 In addition, the 1995 BGA presidential address by Glayde Whitney --- in which he criticized the "Marxist-Lysenkoist denial of genetics" and proposed that differences in murder rates between countries and cities were caused by racial genetic differences in intelligence, empathy, aggression, and impulsivity --- was published in MQ rather than BG.
@@
 
-# Discussion {.unnumbered}
+# Discussion
 
 Using bibliometric and text-mining methods, this study finds that Pioneer-funded researchers published heavily in certain mainstream psychology journals, much more than in *Mankind Quarterly*; and that a distinct topic of race science discourse, centered on race and intelligence, can be identified in these same mainstream journals across the period 1960-2010.
-These findings indicate that *Mankind Quarterly* (MQ) was less important as venue for late 20th century race science than mainstream psychology journals, especially *Intelligence* (Int) and *Personality and Individual Differences* (PID).
-Indeed, until the 1990s, MQ published a very different kind of race science from the race-and-intelligence research published in Int and PID.
+These findings indicate that *Mankind Quarterly* (MQ) was less important as venue for late 20th century race science discourse than mainstream psychology journals, especially *Intelligence* (Int) and *Personality and Individual Differences* (PID).
+Indeed, until the 1990s, MQ published a very different kind of race science discourse from the race-and-intelligence research published in Int and PID.
 During this same period of time, race scientists served in prominent and influential positions in behavior genetics and psychology, including as society presidents and members of journal editorial boards.
@@
 
-<!-- 
-\clearpage
-\pagenumbering{arabic}
-\renewcommand*{\thepage}{\thesection.\arabic{page}}
-\appendix
-\renewcommand{\thesection}{S\arabic{section}}
-\renewcommand\thefigure{\thesection.\arabic{figure}}    
-\setcounter{figure}{0} 
-\renewcommand\thetable{\thesection.\arabic{table}}    
-\setcounter{table}{0}
-
-\clearpage
-
-# Supplemental Materials {.unnumbered}
-
-| 
-| S1. Materials and methods
-| Figures S2-S7
-| Data S8, S9
-
-\clearpage
- -->
-
-# Materials and methods
+# Methods and materials
 
 ## Corpus assembly
-
-![Corpus assembly.](img/dataset.png){#fig-corpus width="120%" fig-env="figure*"}
 
@@
 We then used the author search tool in Clarivate's Web of Science platform (<https://www.webofscience.com/wos/author/search>), retrieving publication lists for 14 researchers.
-These searches were conducted between 2021-09-24 and 2021-10-05 by *[AUTHOR 1]*.
+These searches were conducted between 2021-09-24 and 2021-10-05 by DH.
 After parsing these results, we counted how many of the 14 researchers had published in each journal.
@@
 
-![Count of documents in the corpus, by journal and year.](img/02_count.png){#fig-counts width="11cm" height="11cm" fig-env="figure*"}
-
-\onecolumn
-
-Table: Document counts, by journal, and years included in the corpus. {#tbl-counts}
-
-|journal                                |     n| start|  end|
-|:--------------------------------------|-----:|-----:|----:|
-|Behavior Genetics                      |  2268|  1970| 2010|
-|Intelligence                           |  1237|  1977| 2010|
-|Mankind Quarterly                      |  1821|  1960| 2004|
-|Personality and Individual Differences |  7274|  1980| 2010|
-|Psychological Reports                  | 21398|  1960| 2010|
-|Behavioral and Brain Sciences          |   898|  1978| 2010|
-
-\twocolumn
-
 ## Text preparation
@@
 
-The public analysis repository, <!-- <https://github.com/dhicks/race-science> --> *[URL and DOI will go here]*, includes document metadata and documentwise counts of NLP-extracted noun phrases ("document-term matrices"), along with code to reproduce the analysis of the following sections.
+The public analysis repository, <https://github.com/dhicks/race-science> *DOI will go here]*, includes document metadata and documentwise counts of NLP-extracted noun phrases ("document-term matrices"), along with code to reproduce the analysis of the following sections.
 
@@
 To fit topic models, we followed the approach proposed by Rohe and Zeng [@RoheVintageFactorAnalysis2020], which uses varimax-rotated partial principal components instead of the variational inference methods used by standard topic model packages such as `stm` [@RobertsStmPackageStructural2019].
-This novel approach was implemented in the R package *[redacted for anonymous review]* <!-- `tmfast` [@HicksTmfastFitsTopic2023] -->.
-<!-- We conducted a simulation study of `tmfast`, which found that it was significantly faster and only slightly less accurate at reconstructing known word-topic and topic-document distributions, compared to `stm` [@HicksTmfastFitsTopic2023].
--->
+This novel approach was implemented in the R package `tmfast` [@HicksTmfastFitsTopic2023].
+We conducted a simulation study of `tmfast`, which found that it was significantly faster and only slightly less accurate at reconstructing known word-topic and topic-document distributions, compared to `stm` [@HicksTmfastFitsTopic2023].
+We report a proof-of-concept simulation for the current study in @sec-poc.
 
@@
 Following the approach of [@HicksProductivityInterdisciplinaryImpacts2021], we did not attempt to identify a unique best fitted model for further analysis.
-While the manuscript text focuses on the medium vocabulary, $k=40$ model as the "median" among the 24 fitted models, we compare and contrast findings from this model with those from the other models.
+While the main text of this paper focuses on the medium vocabulary, $k=40$ model as the "median" among the 24 fitted models, qualitatively similar results appear across all models with $k \geq 10$ for all vocabularies; see @sec-topic-quality and @fig-grid-lg, @fig-grid-md, @fig-grid-sm.
 
-### Proof of concept
-
-<!-- {{< include ../scripts/11-proof-of-concept.qmd >}} -->
-
-### Topic model interpretation and quality assessment
-
-After fitting topic models, our first research question was whether we could identify distinctive "race science discourse" topics.
-Figures [@fig-gamma-lg; @fig-gamma-md; and @fig-gamma-sm] show the gamma (topic-document) distributions for each value of $k$ for the three vocabularies.
-In these figures, each panel corresponds to a single journal-$k$ combination, each row of cells is a single document in that journal, and each column of cells is a single topic in that particular model.
-Topics correspond within columns of panels, but not within rows; for example, topic 04 for $k=10$ does not necessarily correspond to topic 04 for $k=20$.
-Color intensity indicates the value of gamma for that particular topic-document combination.
-
-For $k=5, 10$ the distributions are quite noisy, difficult to interpret, and generally don't form very strong visual clusters.
-As $k$ increases from 20 through 40, coherent bands start to appear for journal-distinctive topics in MQ, BG, Int and, to a lesser degree, PID.
-Note that the distinctive topics for MQ and Int do not obviously coincide.
-Above about $k=50$, the models appear to identify more fine-grained topics and the coherent bands fade.
-
-We next constructed Silge plots [@SilgeTopicModeling2017], showing the top 15 (highest-probability) terms (noun phrases) from each topic in the model; see figures [@fig-silge-lg; @fig-silge-md; @fig-silge-sm].
-Perusing these term lists, we focused on topics in each model that contained racial terms.
-Every fitted topic model included at least one racial topic.
-Importantly, in some cases these topics contained terms related to intelligence research, but in other cases they did not.
-
-The gamma distributions and Silge plots suggested that, once $k$ was sufficiently large, the topic models were distinguishing between two types of race science articles.
-We therefore examined the prevalence of topics that used racial terms, intelligence-research terms, or both, by journal and across all 24 models; see figures [@fig-grid-lg; @fig-grid-md; @fig-grid-sm].
-In these visualizations, each row of facets corresponds to a value of $k$ (excluding 5 as too noisy), and each column of facets corresponds to one topic.
-Topics are clustered based on whether the top 15 terms contain racial terms ("race"), terms related to intelligence research ("intelligence"), or both.
-Line plots show the count of articles with $\gamma > 0.50$ for the given topic, by journal; thick lines are 5-year running averages and thin lines are raw counts.
-
-Across all three vocabularies, for sufficiently large values of $k$ the models identify two or more different racial topics, one that appears all but exclusively in MQ and at least one that appears primarily in mainstream journals.
-
-We next conducted a topic quality check of topic 24 from the medium vocabulary, $k=40$ model, which appeared to identify race science research on intelligence published in mainstream journals.
-A spreadsheet of all articles that had a maximum value of $\gamma$ for this topic was extracted (excluding articles published in *Psychological Reports*), and the top 121 articles ($\gamma > 0.97$) were reviewed manually by both authors.
-We coded each article as *race science discourse*, or not, using our definition of race science discourse as treating race science as a legitimate area of scientific reearch; this includes methodological critiques of race science and empirical tests that falsify race science hypotheses.
-
-For the first round of review, both authors worked independently,  dichotomously coding each article as race science discourse or not.
-We calculated the "false positive rate" — documents with maximum $\gamma$ in this topic that were not race science discourse — and interrater reliability using Cohen's $\kappa$.
-The consensus true positive rate was 89.3% (108/121), while the consensus false positive rate was 8.3% (10/121), with 98% agreement across the two raters ($\kappa = 0.86)$.
-Coding spreadsheets are included in the supplemental materials S8-S9.
-For each of the 10 consensus false positive documents, their inclusion in the topic was readily explained, but in different ways for different documents.
-Some included text from other documents, as when an article started in the middle of a page; others discussed perceptions of race relations or racial animus; one piece in BG discussed differences in the herding behavior of two cattle breeds, referred to as "white" and "black."
-
-The 3 non-consensus documents reflected essential ambiguity in the operationalization of race science discourse.
-One is a table of contents from a 1980 issue of Int; arguably this document should have been excluded as irrelevant front matter, but arguably it legitimizes the research published in the issue.
-The second case is a paper on racial differences in emotional intelligence that was classified as race science discourse by one author, while the other author felt it was more focused on psychometrics [@GignacGroupDifferencesEI2010].
-And a 1990 article in MQ (<https://www.unz.com/print/MankindQuarterly-1990q3-00108/>) proposed to offer an economic explanation for crime.
-While it has a few pages connecting race/ethnicity to both crime and poverty, it ultimately focuses on unemployment and alcohol sales, rather than race, as the key predictors of crime.
-
-Based on this qualitative review, we judged that the topic model approach was sufficiently sensitive (low "false positive" rate) in identifying race science discourse.
-However, "false negatives" are still possible, as some documents that human readers would classify as race science discourse might have been associated with other topics.
-
-### STM robustness check
-
-As a robustness check, we compared the `tmfast` results with a topic model fit using `STM`, a widely used package for topic modeling in R *[cite]*.
-Using the medium vocabulary, $k=40$, and a relaxed convergence threshold (`1e-4` compared to the default `1e-5`), fitting a single `STM` model took approximately 20 minutes.
-This was approximately two orders of magnitude slower than `tmfast`.
-And, unlike `tmfast`, `STM` can only fit a single value of $k$ at a time.
-We therefore only fit this one `STM` model.
-
-*[xref]* shows four topics extracted from the `STM` model, automatically identified by the use of race keywords, intelligence keywords, or both, among the top 15 terms.
-As in the `tmfast` analysis, the "race only" topic is almost exclusive to MQ; the "race and intelligence" topic emerges in mainstream journals before appearing in both the mainstream and MQ; and there are distinctive "intelligence only" topics.
-Compare *[xref]*.
-One notable difference between these results is the role of PR in publishing race and intelligence.
-
-![](img/12-stm.png){fig-env="figure*"}
-
-### *[author analysis]*
-
-## Supplemental analysis of *Behavior Genetics* {#sec-bg}
-
-To check the finding that very few race-and-intelligence documents had been published in BG, we ran an independent search for documents with the keywords "race" and "intelligence" published in BG from 1972 to 2020 using Springer's journal search website.
-This search returned 125 documents.
-Thirty-three were presentation abstracts from meetings of the Behavior Genetics Association; we did not examine these further.
-Fourteen of the remaining documents had been published since 2010 (after the scope of the primary study).
-Among these, one obituary and two papers celebrated the work of John Loehlin, a prominent race-and-intelligence researcher who had been Director of the American Eugenics Society from 1968-1972 [@JohnLoehlin19262020; @WaldmanIntroductionFestschriftJohn2014; @PlominGenotypeEnvironmentCorrelationEra2014].
-Another was a retrospective of the work of Lindon Eaves, a geneticist who made at least one notable contribution to debates on the Scarr-Rowe hypothesis (interactions between heritability, race, and class).
-None of these 14 documents reported any studies of racial differences.
-
-Twenty-eight documents in this sample from BG were published between 1990 and 2010.
-One was an obituary of Jerry Hirsch, a critic of hereditarianism in general and Jensen in particular [@RoubertouxJerryHirsch202008], and another was an obituary of David Rowe [@RodgersObituaryDavidChristian2003].
-Only 2 of these 28 documents reported race differences of any kind: one examining interactions among race, sex, and heritability for adolescent BMI (Body Mass Index, used as a measure of overweight/obesity) [@JacobsonGeneticSharedEnvironmental1998]; and the other interactions among race, family history of alcoholism, and visuospatial performance [@BermanReducedVisuospatialPerformance1995].
-Finally, Philip Vernon — one of the Pioneer-funded scientists we identified earlier — published a critical review of a book that attempted to address hereditarianism, the Flynn effect, race-and-intelligence research, and some related issues.
-
-At the same time, we were unable to find any articles published in BG that acknowledged the involvement of the field with scientific racism with anything like the force of the statements that have been made by biological anthropologists [@FuentesAAPAStatementRace2019] and human geneticists [@JacksonFacingOurHistory2023].
-Notably, the ASHG report includes discussions of behavior genetics race science, and in particular is critical of the field and the organization for failing to publicly reject the race-and-intelligence claims made by Shockley and Jensen [@JacksonFacingOurHistory2023].
-A search for "eugenics" in BG found articles that presented the issue as part of the distant past, and a search for "racism" turned up studies of racist attitudes.
-A statement on the Behavior Genetics Association website, dated 2021, acknowledges that "The history of our field is inextricably linked with racism, including the misuse of behavior genetic research to support violent eugenic policies," but primarily focuses on BGA member demographics (<https://www.bga.org/content.aspx?page_id=22&club_id=971921&module_id=567723>).
+We did conduct two further assessments of topic 24 from the medium vocabulary, $k = 40$ model.
+A manual quality assessment achieved consensus that 89% (108/121) of the top papers in this topic were race science discourse; see @sec-topic-quality.
+We also replicated the medium vocabulary, $k = 40$ model using the popular `STM` package for topic modeling; see @sec-stm.
+This model also gave qualitatively similar results.
+Due to the time and computational resources required for these robustness checks, we determined that it would not be feasible for us to repeat them across all 855 topics ($3 \times (5 + 10 + 20 + \cdots + 70)$) and 24 topic models.
 
@@
 
+# Acknowledgments
+
+Thanks to Emily Merchant and John Jackson for some initial discussions that helped clarify the scope of the project and identify some key resources related to the Pioneer Fund and *Mankind Quarterly*.
+Thanks to Anthony Sainez for help retrieving the *Mankind Quarterly* articles.
+Thanks to Derek Devnich and James Dooley for their work attempting to secure the APA-published articles.
+Thanks to Kevin Bird, John Jackson, and Manuela Fernández Pinto for comments on a draft of this paper.
+
+**Funding:** EJCL's work on this project was partially supported by UC Merced.  DJH's work on this project was not supported by any specific funding.
+
+**Author contributions:** DJH (conceptualization, data curation, formal analysis, investigation, methodology, project administration, software, validation, visualization, writing-original draft, writing-review & editing); EJCL (conceptualization, methodology, software, validation, visualization, writing-original draft, writing-review & editing)
+
+**Competing interests:** The authors have no competing interests to declare.
+
+**Data and materials availability:** Web of Science search results and article fulltext are not publicly available due to intellectual property restrictions.
+Preprocessed analysis data (document-term counts and article metadata) and analysis scripts are available at *[doi]*.
+
 # References
@@
 \clearpage
+\pagenumbering{arabic}
+\renewcommand*{\thepage}{S\arabic{page}}
+\renewcommand{\thesection}{\emph{SI Appendix}, section \arabic{section}}
+\renewcommand\thefigure{S\arabic{figure}}    
+\setcounter{figure}{0} 
+\renewcommand\thetable{S\arabic{table}}    
+\setcounter{table}{0}
 
-# Extended Data
+\onecolumn
 
-\renewcommand{\thefigure}{ED-\arabic{figure}}
-\setcounter{figure}{0}
+# SI Appendix
 
-![Fitted topic models: Gamma (topic-document) distributions, large vocabulary.](img/05_lg_gamma.png){#fig-gamma-lg width="6in" height="4.8in"}
+![Corpus assembly.](img/dataset.png){#fig-corpus width="120%" fig-env="figure*"}
 
-![Fitted topic models: Gamma (topic-document) distributions, medium vocabulary.](img/05_md_gamma.png){#fig-gamma-md width="6in" height="4.8in"}
+![Count of documents in the corpus, by journal and year.](img/02_count.png){#fig-counts width="11cm" height="11cm" fig-env="figure*"}
 
-![Fitted topic models: Gamma (topic-document) distributions, small vocabulary.](img/05_sm_gamma.png){#fig-gamma-sm width="6in" height="4.8in"}
+Table: Document counts, by journal, and years included in the corpus. {#tbl-counts}
 
-![Silge plots (top 15 terms) for each topic and model, large vocabulary.](supplements/S2-lg-silge.pdf){#fig-silge-lg}
+|journal                                |     n| start|  end|
+|:--------------------------------------|-----:|-----:|----:|
+|Behavior Genetics                      |  2268|  1970| 2010|
+|Intelligence                           |  1237|  1977| 2010|
+|Mankind Quarterly                      |  1821|  1960| 2004|
+|Personality and Individual Differences |  7274|  1980| 2010|
+|Psychological Reports                  | 21398|  1960| 2010|
+|Behavioral and Brain Sciences          |   898|  1978| 2010|
 
-![Silge plots (top 15 terms) for each topic and model, medium vocabulary.](supplements/S3-md-silge.pdf){#fig-silge-md}
+## Proof of concept {#sec-poc}
 
-![Silge plots (top 15 terms) for each topic and model, small vocabulary.](supplements/S4-sm-silge.pdf){#fig-silge-sm}
+{{< include ../scripts/11-proof-of-concept.qmd >}}
 
-![Topics with "race," "intelligence," or both occurring in the top 15 terms, large vocabulary.](supplements/S5_grid_lg.pdf){#fig-grid-lg}
+## Topic model interpretation and quality assessment {#sec-topic-quality}
 
-![Topics with "race," "intelligence," or both occurring in the top 15 terms, medium vocabulary.](supplements/S6_grid_md.pdf){#fig-grid-md}
+After fitting topic models, our first research question was whether we could identify distinctive "race science discourse" topics.
+Figures [@fig-gamma-lg; @fig-gamma-md; and @fig-gamma-sm] show the gamma (topic-document) distributions for each value of $k$ for the three vocabularies.
+In these figures, each panel corresponds to a single journal-$k$ combination, each row of cells is a single document in that journal, and each column of cells is a single topic in that particular model.
+Topics correspond within columns of panels, but not within rows; for example, topic 04 for $k=10$ does not necessarily correspond to topic 04 for $k=20$.
+Color intensity indicates the value of gamma for that particular topic-document combination.
 
-![Topics with "race," "intelligence," or both occurring in the top 15 terms, small vocabulary.](supplements/S7_grid_sm.pdf){#fig-grid-sm}
+For $k=5, 10$ the distributions are quite noisy, difficult to interpret, and generally don't form very strong visual clusters.
+As $k$ increases from 20 through 40, coherent bands start to appear for journal-distinctive topics in MQ, BG, Int and, to a lesser degree, PID.
+Note that the distinctive topics for MQ and Int do not obviously coincide.
+Above about $k=50$, the models appear to identify more fine-grained topics and the coherent bands fade.
+
+We next constructed Silge plots [@SilgeTopicModeling2017], showing the top 15 (highest-probability) terms (noun phrases) from each topic in the model; see figures [@fig-silge-lg; @fig-silge-md; @fig-silge-sm].
+Perusing these term lists, we focused on topics in each model that contained racial terms.
+Every fitted topic model included at least one racial topic.
+Importantly, in some cases these topics contained terms related to intelligence research, but in other cases they did not.
+
+The gamma distributions and Silge plots suggested that, once $k$ was sufficiently large, the topic models were distinguishing between two types of race science discourse articles.
+We therefore examined the prevalence of topics that used racial terms, intelligence-research terms, or both, by journal and across all 24 models; see figures [@fig-grid-lg; @fig-grid-md; @fig-grid-sm].
+In these visualizations, each row of facets corresponds to a value of $k$ (excluding 5 as too noisy), and each column of facets corresponds to one topic.
+Topics are clustered based on whether the top 15 terms contain racial terms ("race"), terms related to intelligence research ("intelligence"), or both.
+Line plots show the count of articles with $\gamma > 0.50$ for the given topic, by journal; thick lines are 5-year running averages and thin lines are raw counts.
+
+Across all three vocabularies, for sufficiently large values of $k$ the models identify two or more different racial topics, one that appears all but exclusively in MQ and at least one that appears primarily in mainstream journals.
+
+We next conducted a topic quality check of topic 24 from the medium vocabulary, $k=40$ model, which appeared to identify race science discourse on intelligence published in mainstream journals.
+A spreadsheet of all articles that had a maximum value of $\gamma$ for this topic was extracted (excluding articles published in *Psychological Reports*), and the top 121 articles ($\gamma > 0.97$) were reviewed manually by both authors.
+We coded each article as *race science discourse*, or not, using our definition of race science discourse as treating race science as a legitimate area of scientific research; this includes methodological critiques of race science and empirical tests that falsify race science hypotheses.
+
+For the first round of review, both authors worked independently,  dichotomously coding each article as race science discourse or not.
+We calculated the "false positive rate" — documents with maximum $\gamma$ in this topic that were not race science discourse — and interrater reliability using Cohen's $\kappa$.
+The consensus true positive rate was 89.3% (108/121), while the consensus false positive rate was 8.3% (10/121), with 98% agreement across the two raters ($\kappa = 0.86)$.
+Coding spreadsheets are included in the supplemental materials S8-S9.
+For each of the 10 consensus false positive documents, their inclusion in the topic was readily explained, but in different ways for different documents.
+Some included text from other documents, as when an article started in the middle of a page; others discussed perceptions of race relations or racial animus; one piece in BG discussed differences in the herding behavior of two cattle breeds, referred to as "white" and "black."
+
+The 3 non-consensus documents reflected essential ambiguity in the operationalization of race science discourse.
+One is a table of contents from a 1980 issue of Int; arguably this document should have been excluded as irrelevant front matter, but arguably it legitimizes the research published in the issue.
+The second case is a paper on racial differences in emotional intelligence that was classified as race science discourse by one author, while the other author felt it was more focused on psychometrics [@GignacGroupDifferencesEI2010].
+And a 1990 article in MQ (<https://www.unz.com/print/MankindQuarterly-1990q3-00108/>) proposed to offer an economic explanation for crime.
+While it has a few pages connecting race/ethnicity to both crime and poverty, it ultimately focuses on unemployment and alcohol sales, rather than race, as the key predictors of crime.
+
+Based on this qualitative review, we judged that the topic model approach was sufficiently sensitive (low "false positive" rate) in identifying race science discourse.
+However, "false negatives" are still possible, as some documents that human readers would classify as race science discourse might have been associated with other topics.
+
+## STM robustness check {#sec-stm}
+
+As a robustness check, we compared the `tmfast` results with a topic model fit using `STM`, a widely used package for topic modeling in R *[cite]*.
+Using the medium vocabulary, $k=40$, and a relaxed convergence threshold (`1e-4` compared to the default `1e-5`), fitting a single `STM` model took approximately 20 minutes.
+This was approximately two orders of magnitude slower than `tmfast`.
+And, unlike `tmfast`, `STM` can only fit a single value of $k$ at a time.
+We therefore only fit this one `STM` model.
+
+*[xref]* shows four topics extracted from the `STM` model, automatically identified by the use of race keywords, intelligence keywords, or both, among the top 15 terms.
+As in the `tmfast` analysis, the "race only" topic is almost exclusive to MQ; the "race and intelligence" topic emerges in mainstream journals before appearing in both the mainstream and MQ; and there are distinctive "intelligence only" topics.
+Compare *[xref]*.
+One notable difference between these results is the role of PR in publishing race and intelligence.
+
+![](img/12-stm.png){fig-env="figure*"}
+
+## Supplemental analysis of *Behavior Genetics* {#sec-bg}
+
+To check the finding that very few race-and-intelligence documents had been published in BG, we ran an independent search for documents with the keywords "race" and "intelligence" published in BG from 1972 to 2020 using Springer's journal search website.
+This search returned 125 documents.
+Thirty-three were presentation abstracts from meetings of the Behavior Genetics Association; we did not examine these further.
+Fourteen of the remaining documents had been published since 2010 (after the scope of the primary study).
+Among these, one obituary and two papers celebrated the work of John Loehlin, a prominent race-and-intelligence researcher who had been Director of the American Eugenics Society from 1968-1972 [@JohnLoehlin19262020; @WaldmanIntroductionFestschriftJohn2014; @PlominGenotypeEnvironmentCorrelationEra2014].
+Another was a retrospective of the work of Lindon Eaves, a geneticist who made at least one notable contribution to debates on the Scarr-Rowe hypothesis (interactions between heritability, race, and class).
+None of these 14 documents reported any studies of racial differences.
+
+Twenty-eight documents in this sample from BG were published between 1990 and 2010.
+One was an obituary of Jerry Hirsch, a critic of hereditarianism in general and Jensen in particular [@RoubertouxJerryHirsch202008], and another was an obituary of David Rowe [@RodgersObituaryDavidChristian2003].
+Only 2 of these 28 documents reported race differences of any kind: one examining interactions among race, sex, and heritability for adolescent BMI (Body Mass Index, used as a measure of overweight/obesity) [@JacobsonGeneticSharedEnvironmental1998]; and the other interactions among race, family history of alcoholism, and visuospatial performance [@BermanReducedVisuospatialPerformance1995].
+Finally, Philip Vernon — one of the Pioneer-funded scientists we identified earlier — published a critical review of a book that attempted to address hereditarianism, the Flynn effect, race-and-intelligence research, and some related issues.
+
+At the same time, we were unable to find any articles published in BG that acknowledged the involvement of the field with scientific racism with anything like the force of the statements that have been made by biological anthropologists [@FuentesAAPAStatementRace2019] and human geneticists [@JacksonFacingOurHistory2023].
+Notably, the ASHG report includes discussions of behavior genetics race science, and in particular is critical of the field and the organization for failing to publicly reject the race-and-intelligence claims made by Shockley and Jensen [@JacksonFacingOurHistory2023].
+A search for "eugenics" in BG found articles that presented the issue as part of the distant past, and a search for "racism" turned up studies of racist attitudes.
+A statement on the Behavior Genetics Association website, dated 2021, acknowledges that "The history of our field is inextricably linked with racism, including the misuse of behavior genetic research to support violent eugenic policies," but primarily focuses on BGA member demographics (<https://www.bga.org/content.aspx?page_id=22&club_id=971921&module_id=567723>).
+
+![Fitted topic models: Gamma (topic-document) distributions, large vocabulary.](img/05_lg_gamma.png){#fig-gamma-lg width="6in" height="4.8in" fig-env="figure*"}
+
+![Fitted topic models: Gamma (topic-document) distributions, medium vocabulary.](img/05_md_gamma.png){#fig-gamma-md width="6in" height="4.8in" fig-env="figure*"}
+
+![Fitted topic models: Gamma (topic-document) distributions, small vocabulary.](img/05_sm_gamma.png){#fig-gamma-sm width="6in" height="4.8in" fig-env="figure*"}
+
+![Silge plots (top 15 terms) for each topic and model, large vocabulary.](supplements/S2-lg-silge.pdf){#fig-silge-lg fig-env="figure*"}
+
+![Silge plots (top 15 terms) for each topic and model, medium vocabulary.](supplements/S3-md-silge.pdf){#fig-silge-md fig-env="figure*"}
+
+![Silge plots (top 15 terms) for each topic and model, small vocabulary.](supplements/S4-sm-silge.pdf){#fig-silge-sm fig-env="figure*"}
+
+![Topics with "race," "intelligence," or both occurring in the top 15 terms, large vocabulary.](supplements/S5_grid_lg.pdf){#fig-grid-lg fig-env="figure*"}
+
+![Topics with "race," "intelligence," or both occurring in the top 15 terms, medium vocabulary.](supplements/S6_grid_md.pdf){#fig-grid-md fig-env="figure*"}
+
+![Topics with "race," "intelligence," or both occurring in the top 15 terms, small vocabulary.](supplements/S7_grid_sm.pdf){#fig-grid-sm fig-env="figure*"}

### 2024-05-03 scripts/11-proof-of-concept.qmd
various todos

@@
 format:
-  - html
-  - pdf: 
+  html: default
+  pdf: 
     mathspec: true
@@
     scale_y_discrete(breaks = c('V01', 'V10', 'V20')) +
-    scale_x_discrete(breaks = 'none')
+    scale_x_discrete(breaks = 'none') +
+    theme_bw()
 ```
@@
     ggforce::geom_sina() +
-    stat_summary(color = 'red')
+    stat_summary(color = 'red') +
+    theme_bw()
 ```
@@
     geom_sina() +
-    stat_summary(color = 'red')
+    stat_summary(color = 'red') +
+    theme_bw()
 ```
@@
                   inherit.aes = FALSE, color = 'black') +
-    scale_color_viridis_d(guide = 'none')
+    scale_color_viridis_d(guide = 'none') +
+    theme_bw()
 ```

### 2024-05-06 paper/paper.qmd
revisioning

@@
 *[TODO: 
-- NHB reviews todo
+- NHB reviews todo: <https://docs.google.com/document/d/1xzG95pZOZgZQFfSk6yYMFhrnUQ8AWk9U1O5ngm3n_uo/edit>
 - de-anonymize
@@
 - SI xrefs
+- bookmarks on wrong pages
+- align abstract w/ intro
+- proofread proof-of-concept
 ]*
@@
 We stress that race science discourse includes methodological and empirical critiques of race science.
-We argue that such technical critiques — even when they are intended to undermine race science — can instead legitimize it, by creating an impression among the broader public that race science is a matter of ordinary, mainstream scientific debate.
+We argue that such technical critiques — even when they are intended to undermine race science — can instead legitimize it, by creating an impression among the broader public that race science is a matter of ordinary, reasonable, mainstream scientific debate.
 For example, the Flynn effect (the finding that IQ scores have increased over time) poses a serious empirical challenge to hereditarian claims of racial differences in intelligence, and has been extensively debated in the mainstream scientific literature.
-Most members of the general public lack the time and skills needed to assess the evidence and arguments concerning the Flynn effect.
-Instead, we suggest, they are likely to understand the Flynn effect — and, by extension, race science — as an area of active, mainstream scientific research and ordinary scientific disagreement.
-In this way technical critiques of race science based on the Flynn effect can legitimize race science.
-So empirical research supporting the Flynn effect is usually not itself race science, but can still be race science discourse.
+Most members of the general public lack the time and skills needed to assess these debates for themselves.
+So, we suggest, they are quite likely to use a "middle ground" heuristic to understand these debates as ordinary, reasonable disagreement among mainstream scientists.
+Some experts say one thing, other experts say something else, and so "the truth is somewhere in the middle."  Tobacco and other industries have exploited this heuristic to "manufacture doubt" about the hazards of their products *[cites]*.
+In this way technical critiques of race science based on the Flynn effect could legitimize race science.
+So, while empirical research supporting the Flynn effect is usually not itself race science, it can still be race science discourse.
 
@@
 
-## Scientific racism, eugenics, and *Brown* {.unnumbered}
+<!-- ## Scientific racism, eugenics, and *Brown* {.unnumbered} -->
 
@@
 
-In light of the scientific racist origins of Pioneer and MQ, and the significant attention MQ has received in recent historiography, we hypothesized that bibliometric and text mining analyses would show race science discourse originating in MQ, and from there being disseminated to mainstream publications.
+In light of the scientific racist origins of Pioneer and MQ, and the significant attention MQ has received in recent historiography, we hypothesized that MQ might have served as an "incubator" for race science discourse, with ideas originating in MQ before appearing within the scientific mainstream.
+Establishing venues for favorable research is another key element of the "tobacco strategy" for manufacturing doubt *[FP]*.
+Specifically, we hypothesized that bibliometric and text mining analyses would show race science discourse appearing first in MQ, and subsequently appearing in mainstream publications.
 This hypothesis was not supported.
@@
 See @tbl-researchers and figure @fig-wos.
+As the affiliations in @tbl-researchers indicate, almost all of these researchers were faculty at major research universities.
 
@@
 
-| | | |
-|:-----|-----:|----:|
-| Thomas J. Bouchard, Jr. (1937-) | psychology | 184 |
-| Brunetto Chiarelli (?-?) | anthropology | 91 |
-| Hans Eysenck (1916-1997) | psychology | 661 |
-| Robert A. Gordon (1932-) | sociology | 0|
-| Linda Gottfredson (1947-) | psychology | 71 |
-| Garrett Hardin (1915-2003) | ecology | 75 |
-| Joseph M. Horn (1940-) | psychology | 68 |
-| Lloyd Humphreys (1913-2003) | psychology | 0 |
-| Arthur Jensen (1923-2003) | psychology | 235 |
-| Michael Levin (1943-) | philosophy | 96 |
-| Richard Lynn (1930-2023) | psychology | 288 |
-| R. Travis Osborne (1913-2013) | psychology | 59 |
-| J. Phillippe Rushton (1943-2012) | psychology | 277 |
-| Audrey M. Shuey (1900-1977) | psychology | 10 |
-| Philip A. Vernon (1950-) | psychology | 227 |
-| Daniel Vining, Jr. (1944-) | demography | 33 |
+| | | | |
+|:--------|:----|:--------|-:|
+| Thomas J. Bouchard, Jr. (1937-) | psychology | Univ. Minnesota | 184 |
+| Brunetto Chiarelli (?-?) | anthropology | | 91 |
+| Hans Eysenck (1916-1997) | psychology | King's College London | 661 |
+| Robert A. Gordon (1932-) | sociology | Johns Hopkins | 0|
+| Linda Gottfredson (1947-) | psychology | Univ. Delaware | 71 |
+| Garrett Hardin (1915-2003) | ecology | UC Santa Barbara | 75 |
+| Joseph M. Horn (1940-) | psychology | UT Austin | 68 |
+| Lloyd Humphreys (1913-2003) | psychology | Univ. Illinois | 0 |
+| Arthur Jensen (1923-2003) | psychology | UC Berkeley | 235 |
+| Michael Levin (1943-) | philosophy | CCNY | 96 |
+| Richard Lynn (1930-2023) | psychology | Ulster University\* | 288 |
+| R. Travis Osborne (1913-2013) | psychology | Univ. Georgia | 59 |
+| J. Phillippe Rushton (1943-2012) | psychology | Univ. Western Ontario | 277 |
+| Audrey M. Shuey (1900-1977) | psychology | Randolph-Macon Women's College | 10 |
+| Philip A. Vernon (1950-) | psychology | Univ. Western Ontario | 227 |
+| Daniel Vining, Jr. (1944-) | demography | Univ. Pennsylvania | 33 |
 
 Table: Pioneer-funded researchers.
-Either identified in [@MillerPioneerFundBankrolling1994] or named on an archive copy of Pioneer's website, along with birth and death dates from Wikipedia, attributed discipline, and WoS author search result counts.
-Brunetto Chiarelli does not have a Wikipedia page.
+Either identified in [@MillerPioneerFundBankrolling1994] or named on an archive copy of Pioneer's website; along with birth and death dates from Wikipedia, attributed discipline, academic affiliation, and WoS author search result counts.
+Brunetto Chiarelli does not have a Wikipedia page; Richard Lynn's title of Professor Emeritus at Ulster University was withdrawn in 2018 *[<https://www.bbc.co.uk/news/uk-northern-ireland-43768132>]*.
 {#tbl-researchers}
@@
 The fact that Pioneer-funded researchers published heavily in two mainstream psychology journals does not tell us anything about the content of their publications or the claims of their research.
-We assembled a full-text corpus of articles published in MQ, PID, Int, BG, PR, and Behavior and Brain Sciences (BBS) between 1960-2010 and used topic modeling to identify topics discussing race, intelligence, and both.
+We assembled a full-text corpus of articles published in MQ, PID, Int, BG, PR, and Behavior and Brain Sciences (BBS) between 1960-2010 and used topic modeling across three different sizes of vocabulary ("large," "medium," and "small") to identify topics discussing race, intelligence, and both.
 (See the materials and methods for a discussion of some important limitations with this corpus: journals published by the American Psychological Association could not be included, BBS bundles focus articles and commentaries together in a way that confounds simple document counts, and an unknown number of relevant documents in PID could not be included.)
 
-![Silge plots [@SilgeTopicModeling2017] and smoothed time series for three focal topics.
+![Silge plots [@SilgeTopicModeling2017] and smoothed time series for three focal topics from the medium vocabulary, $k = 40$ model.
 Top row: top 15 terms by $\beta$ (term-topic distribution) for each topic.
@@
 
-*[across all models]*
+These qualitative results are similar across models using the large and medium vocabularies; see @fig-grid-lg, @fig-grid-md, @fig-grid-sm.
+Across all vocabularies, as $k$ increases, multiple different "intelligence only" topics emerge, apparently picking out different methods or aspects of intelligence research.
+For example, there is often an "intelligence only" topic that is highly associated with BG.
+MQ never has many publications in an "intelligence only" topic, except for $k=10, 20$ in the large vocabulary; in these two models, there is neither a "race only" nor "race-and-intelligence" topic.
 
-- As $k$ increases, multiple different "intelligence only" topics
-- lg, md: usually 1 "race only" topic and/or 1 race-and-intelligence topic
-    - "race only" always almost exclusively in MQ
-    - md: race-and-intelligence predominantly in mainstream
-    - lg: for k=30, race-and-intelligence predominantly MQ; no race-int for k=40, 50; then similar to md for k=60, 70
-- sm: 
-    - seems to have a distinct "race only" topic in PR (V04)
-        - top terms include race_differences, rushton, crime: race science, but poss. either not about intelligence or int-related terms didn't make it into top terms
-    - other one is almost exclusive to MQ
-    - except for k=40, race-int topic, predominantly mainstream
+In the large and medium vocabularies, there is usually one "race only" topic and/or one "race-and-intelligence" topic, and the "race only" topic is always virtually exclusive to MQ.
+In the medium vocabulary, "race-and-intelligence" is more prevalent in mainstream journals than in MQ; except for $k=20$, where there is no "race only" topic.
+In the large vocabulary, a "race only" topic does not emerge until $k=40$, and there are not simultaneous "race only" and "race-and-intelligence" topics until $k=60$.
+However, in $k=60, 70$ there are both topics, and they follow the same pattern seen in the medium vocabulary: "race only" is almost exclusive to MQ, and "race-and-intelligence" is more prevalent in mainstream journals.
+
+Unlike the other two vocabularies, in the small vocabulary there are two "race only" topics for $k \geq 50$.
+Across values of $k$, topic V04 is more prevalent in PS through about 1990; the other "race only" topic is typically more prevalent in MQ, but also includes a few articles from PS.
+For $k \geq 40$, the small vocabulary models have distinct "race only," "race-and-intelligence," and "intelligence only" topics; "race-and-intelligence" is always predominantly mainstream, and increases in prevalence over the course of the 1980s.
+The "extra" "race only" topic appars to fade out quickly over the course of the 1980s, and its top terms include `race differences`, `rushton`, and `crime`.
+So one possibility is that that the small vocabulary is able to pick out a distinct line of race science discourse that is focused on crime and/or other behaviors, rather than intelligence, and was replaced with a more intelligence-focused line, perhaps in a way that anticipated publication of *The Bell Curve* in 1994.
+However, we think it's more plausible that this separation is an artifact due to the very small size of this vocabulary, with many fewer terms than documents.
 
@@
 BG published very few articles in any race-and-intelligence topic identified by any of the 24 topic models.
-Among both academics and the general public, the field of behavior genetics is strongly associated with race science, and specifically race-and-intelligence research.
+But, among both (other) academics and the general public, the field of behavior genetics is strongly associated with race science, and specifically race-and-intelligence research.
 Panofsky [@PanofskyMisbehavingScienceControversy2014] argues that, prior to Jensen's 1969 paper, behavior genetics emphasized the study of non-human animals and intentionally avoided associations with eugenics and public controversy more generally.
 In response to Jensen, critics such as Lewontin offered broad critiques of behavior genetics as such, and behavior geneticists in turn adopted a radical conception of academic freedom (without any sense of responsibility for the social implications of academic research) and a siege or wartime mentality, as illustrated by Sandra Scarr's 1986 presidential address to the Behavior Genetics Association (BGA) [@ScarrThreeCheersBehavior1987].
-Scarr's address coincided with the period between 1970-1990 when BG published articles in topic 24 (including Scarr's address itself).
+Scarr's address coincided with the period between 1970-1990 when BG published articles (including Scarr's address itself) in the race science discourse topic (24) of the medium vocabulary, $k = 40$ model.
 
@@
 This interpretation was supported by a supplemental analysis that focused on BG specifically; see @sec-bg.
-In addition, the 1995 BGA presidential address by Glayde Whitney --- in which he criticized the "Marxist-Lysenkoist denial of genetics" and proposed that differences in murder rates between countries and cities were caused by racial genetic differences in intelligence, empathy, aggression, and impulsivity --- was published in MQ rather than BG.
+For example, the 1995 BGA presidential address by Glayde Whitney --- in which he criticized the "Marxist-Lysenkoist denial of genetics" and proposed that differences in murder rates between countries and cities were caused by racial genetic differences in intelligence, empathy, aggression, and impulsivity --- was published in MQ rather than BG.
 Behavior genetics may have been prominent in promoting race science in the past, but appears to have been less receptive to such ideas in more recent history.
-Instead, our results suggest that other subdisciplines of psychology have provided the primary venue for mainstreaming race science.
+Instead, our results suggest that other subdisciplines of psychology — perhaps especially cognitive psychology — have provided the primary venue for mainstreaming race science.
 

### 2024-05-08 paper/paper.qmd
revisions, mostly proof-of-concept

@@
         country: "CA, USA"
+    orcid: "0000-0001-7945-4416"
     corresponding: true
@@
       - ref: a
+    orcid: "*[todo]*"
 
 abstract: |
-  | While anthropologists and human geneticists have long rejected the biological race concept, both social movements and prominent researchers have purported to justify racial inequality and colonialism by appealing to the epistemic authority of science.
+  | Anthropologists and human geneticists have long rejected race science, research that can be utilized to purportedly justify racial inequality and colonialism.
+But race science has persisted, at the very least on the margins of the mainstream scientific community.
 In this study, bibliometric and text mining methods were used to trace the appearance of race science discourse in both mainstream behavioral science journals and the parascholarly journal *Mankind Quarterly* from 1960 to 2010.
 Our analysis found two distinguishable lines of race science discourse.
-One is strongly associated with *Mankind Quarterly*, while the other — emphasizing race and intelligence — is strongly associated with mainstream journals, especially *Intelligence* and *Personality and Individual Differences*.
-  Psychology, and especially intelligence research, has played a significant role in mainstreaming race science.
+One is strongly associated with *Mankind Quarterly*, while the other — emphasizing race and intelligence — is strongly associated with mainstream psychology journals, especially *Intelligence* and *Personality and Individual Differences*.
+We compare psychology's response to race science to those of other fields, and discuss the social responsibility of scientists with respect to race science.
+
+significance: |
+  | Anthropologists and human geneticists have long rejected efforts to purportedly scientifically justify racial inequality and colonialism.
+This study examines the persistence of scientific racism throughout the second half of the twentieth century, examining both mainstream behavioral science journals and a notorious fringe journal, *Mankind Quarterly* (MQ).
+While we expected to find evidence that race science emerged first in MQ and later spread to mainstream journals, we instead found two distinct lines of race science discourse.
+Race-and-intelligence discourse appears first and primarily in two mainstream journals, *Intelligence* and *Personality and Individual Differences*.
+Unlike anthropology and human genetics, psychology has not strongly condemned race science in the mainstream of the field.
+
+keywords: 
+  - race science
+  - race and intelligence
+  - psychology
+  - behavior genetics
+  - text mining
+  - topic modeling
 
@@
   pnas-pdf:
-    keep_tex: true
-    latex_engine: "xelatex"
+    keep-tex: true
     link-citations: true
@@
     fig.env: "figure*"
-    number-sections: false
+    number-sections: true
 
@@
   fig-prefix: ""
-
-# header-includes:
-#  - \usepackage{lineno}
-#  - \linenumbers
+  sec-prefix: "*SI Appendix*, section"
 
@@
 - NHB reviews todo: <https://docs.google.com/document/d/1xzG95pZOZgZQFfSk6yYMFhrnUQ8AWk9U1O5ngm3n_uo/edit>
-- de-anonymize
-- copy updated plots? 
+- copy updated plots
 - figure sizes
-- SI xrefs
-- bookmarks on wrong pages
 - align abstract w/ intro
-- proofread proof-of-concept
+- figures and tables to end
 ]*
 
-# Significance
-
-*[120 words, "undergrad-educated scientist outside their field of specialty"]*
-
-# Introduction
+# Introduction {.unnumbered}
 
@@
 
-# Results
+# Results {.unnumbered}
 
@@
 Either identified in [@MillerPioneerFundBankrolling1994] or named on an archive copy of Pioneer's website; along with birth and death dates from Wikipedia, attributed discipline, academic affiliation, and WoS author search result counts.
-Brunetto Chiarelli does not have a Wikipedia page; Richard Lynn's title of Professor Emeritus at Ulster University was withdrawn in 2018 *[<https://www.bbc.co.uk/news/uk-northern-ireland-43768132>]*.
+Brunetto Chiarelli does not have a Wikipedia page; Richard Lynn's title of Professor Emeritus at Ulster University was withdrawn in 2018.
 {#tbl-researchers}
@@
 
-# Methods and materials
+# Methods and materials {.unnumbered}
 
-## Corpus assembly
+## Corpus assembly {.unnumbered}
 
@@
 
-### Mankind Quarterly
+### Mankind Quarterly {.unnumbered}
 
@@
 
-### Mainstream journals
+### Mainstream journals {.unnumbered}
 
@@
 
-## Text preparation
+## Text preparation {.unnumbered}
 
@@
 
-### Data and code availability
+### Data and code availability {.unnumbered}
 
@@
 
-## Vocabulary selection
+## Vocabulary selection {.unnumbered}
 
@@
 
-## Topic modeling
+## Topic modeling {.unnumbered}
 
@@
 
-# Acknowledgments
+# Acknowledgments {.unnumbered}
 
@@
 
-# References
+# References {.unnumbered}
 
@@
 \renewcommand*{\thepage}{S\arabic{page}}
-\renewcommand{\thesection}{\emph{SI Appendix}, section \arabic{section}}
 \renewcommand\thefigure{S\arabic{figure}}    
@@
 
-# SI Appendix
+# SI Appendix {.unnumbered}
+
+Supporting Information for "Race Science in Mainstream Psychology, 1960-2010"
+
+Daniel J. Hicks, Emilio J.C. Lobato
+
+\clearpage
 
@@
 
-## Proof of concept {#sec-poc}
+# Proof of concept {#sec-poc}
 
@@
 
-## Topic model interpretation and quality assessment {#sec-topic-quality}
+# Topic model interpretation and quality assessment {#sec-topic-quality}
 
@@
 
-## STM robustness check {#sec-stm}
+# STM robustness check {#sec-stm}
 
-As a robustness check, we compared the `tmfast` results with a topic model fit using `STM`, a widely used package for topic modeling in R *[cite]*.
+As a robustness check, we compared the `tmfast` results with a topic model fit using `STM`, a widely used package for topic modeling in R [@RobertsStmPackageStructural2019].
 Using the medium vocabulary, $k=40$, and a relaxed convergence threshold (`1e-4` compared to the default `1e-5`), fitting a single `STM` model took approximately 20 minutes.
@@
 
-*[xref]* shows four topics extracted from the `STM` model, automatically identified by the use of race keywords, intelligence keywords, or both, among the top 15 terms.
+Figure @fig-stm shows four topics extracted from the `STM` model, automatically identified by the use of race keywords, intelligence keywords, or both, among the top 15 terms.
 As in the `tmfast` analysis, the "race only" topic is almost exclusive to MQ; the "race and intelligence" topic emerges in mainstream journals before appearing in both the mainstream and MQ; and there are distinctive "intelligence only" topics.
-Compare *[xref]*.
+Compare figures @fig-focal, @fig-grid-md.
 One notable difference between these results is the role of PR in publishing race and intelligence.
 
-![](img/12-stm.png){fig-env="figure*"}
+![Silge plots and thresholded document counts for "race only," "race-and-intelligence," and "intelligence only" topics in the STM model.](img/12-stm.png){#fig-stm fig-env="figure*"}
 
-## Supplemental analysis of *Behavior Genetics* {#sec-bg}
+# Supplemental analysis of *Behavior Genetics* {#sec-bg}
 

### 2024-05-08 scripts/11-proof-of-concept.qmd
revisions, mostly proof-of-concept

@@
 
-This section uses simulated data to provide a proof-of-concept for our use of topic models to detect the "mainstreaming" of a marginalized topic.
+This section uses simulated data to provide a proof-of-concept for our use of topic models to detect the "mainstreaming" of a marginalized topic (corresponding to race science).
+The `tmfast` package [@HicksTmfastFitsTopic2023] provides samplers that can be used to generate a simulated document-term matrix.
 
@@
 ## Simulation parameters
+j = 10            ## number of journals
 k = 20            ## total number of topics
@@
 We construct a simulated corpus as follows.
-We work with five journals, labelled A through E.
-Journal A primarily publishes the marginalized topic (correspond to race science).
-Ten journals, C-L, are mainstream journals, and publish an even mix from `r k-1` other topics, but avoid the marginalized topic (weighted 0 in the Dirichlet parameterization below).
-These journals represent the context or comparison group for identifying the race science topic; their exact number is not important here.
+We work with `r j` journals, labelled A through `r LETTERS[j]`.
+Journal A primarily publishes the marginalized topic.
+Journals C-`r LETTERS[j]` are mainstream journals, and publish an even mix from `r k-1` other topics, but avoid the marginalized topic (weighted 0 in the Dirichlet parameterization below).
+These journals represent the context or comparison group for identifying the race science topic.
 
-Journal B is the venue by which the marginalized topic is "mainstreamed."  This journal is represented in a time-indexed way, $B_t$, with $t$ ranging from 0 to 5.
-For simplicity, in the text each time-slice of B is treated as a distinct journal, giving us 17 journals total (A; the ten C-L; plus 6 time-slices of B).
+Journal B is the venue by which the marginalized topic is "mainstreamed."  This journal is represented as a series of time slices, $B_t$, with $t$ ranging from 0 to 5.
+For simplicity, in the simulation each time-slice of B is treated as a distinct journal, giving us `r j + 5` "journals" total.
 
-This gives us three sets of journals: A, B (0-5), and C-L.
+This gives us three sets of journals: A, B (0-5), and C-`r LETTERS[j]`.
 In the LDA model, these journal sets correspond to $\boldsymbol{\alpha}$, the parameter for the "prior" Dirichlet distribution $\Theta(\boldsymbol{\alpha})$.
@@
 
-```{r}
+```{r journal_priors}
 ## Construct journal theta ----
@@
 ## C-E: mainstream journals
-C_alpha = rep.int(list(peak_alpha(k, 1, peak = 0)), 10) |> 
-    set_names(LETTERS[3:12])
+C_alpha = rep.int(list(peak_alpha(k, 1, peak = 0)), j - 2) |> 
+    set_names(LETTERS[3:j])
 C_alpha$C
@@
 
-## 17 journals/journals-at-time-t
-length(alphas)
+## j + 5 journals/journals-at-time-t
+identical(length(alphas), j + 5)
 ```
@@
 
-```{r}
+```{r journal_theta}
 theta = map(alphas, ~ rdirichlet(N, .x)) |> 
@@
 
-```{r}
+```{r words}
 ## Word distributions ----
@@
 
-```{r}
+```{r doc_length}
 ## Document length ----
@@
 With these elements constructed, `draw_corpus()` generates a synthetic document-term matrix.
-Because this process is relatively slow (approximately 75 seconds on the laptop used to prepare this simulation), we use `memoise` to create a cache.
-As a pre-processing step for the topic model, we apply the `log(x+1)` transform.
+Because this process is relatively slow (~100 seconds on the laptop used to prepare this simulation), we use `memoise` to create a cache.
+(Due to memoization, the wall time reported in the chunk output may be much faster than the actual time to generate the document-term matrix.) As a pre-processing step for the topic model, we apply the `log(x+1)` transform.
 
-```{r}
+```{r draw_corpus}
 ## Draw corpus ----
@@
 
-We can now fit the topic model.
+We can now fit the topic model.  (Note that this is not memoized, so the wall time reported in the chunk output corresponds to the actual time to fit.)
 
-```{r}
+```{r fit_model}
 ## Fit topic model ----
@@
 
-The tile plot indicates that (1) documents in journal A generally have a high probability for topic 1; journals C-E generally have a low probability; and journal B gradually has more articles in topic 1.
+The tile plot indicates that (1) documents in journal A generally have a high probability for topic 1; journals C-`r LETTERS[j]` generally have a low probability; and journal B gradually has more articles in topic 1.
 
-```{r}
+```{r tile_plot}
 #| fig-width: 7
@@
     facet_wrap(vars(journal), scales = 'free_x', 
-               ncol = 7) +
+               ncol = 5) +
     scale_y_discrete(breaks = c('V01', 'V10', 'V20')) +
@@
 
-Focusing on topic 1, we first calculate the average probability across each journal.  A has the highest value; C-E are consistently low; and B shows an increase over time.
+Focusing on topic 1, we first calculate the average probability across each journal.  A has the highest value; C-`r LETTERS[j]` are consistently low; and B shows an increase over time.
 
-```{r}
+```{r marginal_topic_plot}
 #| fig-cap: "In the proof-of-concept simulation, the topic model clearly detects the marginal topic in journal A, increase in journal B over time, and minimal presence in other journals"
@@
     ggplot(aes(journal, gamma)) +
-    ggforce::geom_sina() +
+    geom_sina() +
     stat_summary(color = 'red') +
@@
 
-```{r}
+```{r ms_topic_plot}
 #| fig-cap: "In the proof-of-concept simulation, the topic model does not produce a false negative pattern with the mainstream topics"
@@
 
-```{r}
+```{r marginal_topic_rate}
 #| tbl-cap: "In the proof-of-concept simulation, almost all articles in journal A contain the marginal topic; the rate in B increases over time; and the marginal topic does not appear in the other journals"
@@
 
-As a limitation, we note that, for $\theta > 0.05$ or so, the estimated topic probabilities are below the true values.
+As a limitation, we note that, for $\theta > 0.05$ or so, the estimated topic probabilities (`gamma`) are below the true values (`theta`).
 
-```{r}
+```{r gamma_theta_scatter}
 #| fig-cap: "In the proof-of-concept simulation, without normalization the estimated topic-document probabilities (gamma) are well below the true probabilities (theta) for approx. $\\theta > 0.05$"
@@
 
-*[tmfast paper]* notes this problem, and implements a "renormalization" option when extracting the probability distributions from the fitted model.
+@HicksTmfastFitsTopic2023 notes this problem, and implements a "renormalization" option when extracting the probability distributions from the fitted model.
 This "renormalization" increases the "peakedness" of the distribution (lower entropy), adjusting the highest probabilities up and making the small probabilities several orders of magnitude smaller.

### 2024-05-08 paper/paper.qmd
scales on Silge plots

@@
 - NHB reviews todo: <https://docs.google.com/document/d/1xzG95pZOZgZQFfSk6yYMFhrnUQ8AWk9U1O5ngm3n_uo/edit>
+- figure out renv::restore()
+    - re-make
     - copy updated plots

### 2024-05-08 scripts/06_hellinger_betas.R

@@ -66,7 +66,7 @@ hellinger_betas = function(model_file, name) {
                     breaks = scales::pretty_breaks(n = 3)) +
                 coord_flip(ylim = c(0, NA)) +
                 facet_wrap(vars(topic),
-                           scales = 'free',
+                           scales = 'free_y',
                            ncol = 5) +
                 theme(plot.background = 
                           element_rect(colour = "black", 

### 2024-05-08 scripts/08_topics_over_time.R

@@ -214,7 +214,7 @@ silge_gg = tidy(md_tmf, matrix = 'beta', k = 40) |>
     coord_flip() +
     scale_x_reordered() + 
     labs(y = 'β') +
-    facet_wrap(vars(topic), scales = 'free')
+    facet_wrap(vars(topic), scales = 'free_y')
 if (interactive()) silge_gg
 
 # silge_gg / time_gg +

### 2024-06-20 paper/paper.qmd
re-snapshot renv bc compiler errors

--- 2024-05-08 paper/paper.qmd
@@
       - ref: a
-    orcid: "*[todo]*"
+    orcid: "0000-0002-3066-2932"
 
@@
 figure sizes: 
-Small: approximately 9 cm x 6 cm
-Medium: approximately 11 cm x 11 cm
-Large: approximately 18 cm x 22 cm
+Small: approximately 9 cm x 6 cm (3.5 in x 2.4 in)
+Medium: approximately 11 cm x 11 cm (4.3 in x 4.3 in)
+Large: approximately 18 cm x 22 cm (7.1 in x 8.7 in)
 
@@
 *[TODO: 
-- NHB reviews todo: <https://docs.google.com/document/d/1xzG95pZOZgZQFfSk6yYMFhrnUQ8AWk9U1O5ngm3n_uo/edit>
+- figure out renv::restore()
+    - re-make
     - copy updated plots
-- figure sizes
-- align abstract w/ intro
-- figures and tables to end
 ]*
@@
 So, we suggest, they are quite likely to use a "middle ground" heuristic to understand these debates as ordinary, reasonable disagreement among mainstream scientists.
-Some experts say one thing, other experts say something else, and so "the truth is somewhere in the middle."  Tobacco and other industries have exploited this heuristic to "manufacture doubt" about the hazards of their products *[cites]*.
+Some experts say one thing, other experts say something else, and so "the truth is somewhere in the middle."  Tobacco and other industries have exploited this heuristic to "manufacture doubt" about the hazards of their products [@FernandezPintoKnowBetterNot2017].
 In this way technical critiques of race science based on the Flynn effect could legitimize race science.
@@
 In light of the scientific racist origins of Pioneer and MQ, and the significant attention MQ has received in recent historiography, we hypothesized that MQ might have served as an "incubator" for race science discourse, with ideas originating in MQ before appearing within the scientific mainstream.
-Establishing venues for favorable research is another key element of the "tobacco strategy" for manufacturing doubt *[FP]*.
+Establishing venues for favorable research is another key element of the "tobacco strategy" for manufacturing doubt [@FernandezPintoKnowBetterNot2017].
 Specifically, we hypothesized that bibliometric and text mining analyses would show race science discourse appearing first in MQ, and subsequently appearing in mainstream publications.
@@
 
-![Journals publishing 6 or more Pioneer-funded researchers, Web of Science author search results.](img/wos_results.png){#fig-wos width="4.76in" height="2.6in" fig-env="figure*"}
+![Journals publishing 6 or more Pioneer-funded researchers, Web of Science author search results.](img/pf_wos_results.png){#fig-wos width="7.1in" height="3.5in" fig-env="figure*"}
 
@@
 Thin lines give annual values, thick lines give 5-year running averages.
-Because *Behavior Genetics* and *Behavioral and Brain Sciences* are not prominent in any panel, neither is given a direct label.](img/focal_topics.png){#fig-focal width="4.76in" height="3.57in" fig-env="figure*"}
+Because *Behavior Genetics* and *Behavioral and Brain Sciences* are not prominent in any panel, neither is given a direct label.](img/08_focal_topics.png){#fig-focal width="7.1in" height="5.2in" fig-env="figure*"}
 
@@
 
-![Count of documents in the corpus, by journal and year.](img/02_count.png){#fig-counts width="11cm" height="11cm" fig-env="figure*"}
+![Count of documents in the corpus, by journal and year.](img/02_count.png){#fig-counts width="3.5in" height="2.4in" fig-env="figure*"}
 
@@
 
-![Silge plots (top 15 terms) for each topic and model, large vocabulary.](supplements/S2-lg-silge.pdf){#fig-silge-lg fig-env="figure*"}
+![Silge plots (top 15 terms) for each topic and model, large vocabulary.](img/06-lg-silge.pdf){#fig-silge-lg fig-env="figure*"}
 
-![Silge plots (top 15 terms) for each topic and model, medium vocabulary.](supplements/S3-md-silge.pdf){#fig-silge-md fig-env="figure*"}
+![Silge plots (top 15 terms) for each topic and model, medium vocabulary.](img/06-md-silge.pdf){#fig-silge-md fig-env="figure*"}
 
-![Silge plots (top 15 terms) for each topic and model, small vocabulary.](supplements/S4-sm-silge.pdf){#fig-silge-sm fig-env="figure*"}
+![Silge plots (top 15 terms) for each topic and model, small vocabulary.](img/06-sm-silge.pdf){#fig-silge-sm fig-env="figure*"}
 
-![Topics with "race," "intelligence," or both occurring in the top 15 terms, large vocabulary.](supplements/S5_grid_lg.pdf){#fig-grid-lg fig-env="figure*"}
+![Topics with "race," "intelligence," or both occurring in the top 15 terms, large vocabulary.](img/08_grid_lg.pdf){#fig-grid-lg fig-env="figure*"}
 
-![Topics with "race," "intelligence," or both occurring in the top 15 terms, medium vocabulary.](supplements/S6_grid_md.pdf){#fig-grid-md fig-env="figure*"}
+![Topics with "race," "intelligence," or both occurring in the top 15 terms, medium vocabulary.](img/08_grid_md.pdf){#fig-grid-md fig-env="figure*"}
 
-![Topics with "race," "intelligence," or both occurring in the top 15 terms, small vocabulary.](supplements/S7_grid_sm.pdf){#fig-grid-sm fig-env="figure*"}
+![Topics with "race," "intelligence," or both occurring in the top 15 terms, small vocabulary.](img/08_grid_sm.pdf){#fig-grid-sm fig-env="figure*"}

### 2024-06-20 scripts/02_describe_corpus.R

@@ -37,7 +37,7 @@ count_df = meta_ar |>
 label_df = count_df |> 
     group_by(container.title) |> 
     slice_max(n, n = 1) |> 
-    mutate(container.title = str_wrap(container.title, width = 25))
+    mutate(container.title_w = str_wrap(container.title, width = 25))
 
 ggplot(count_df, 
        aes(year, n, color = container.title, group = container.title)) +
@@ -46,7 +46,7 @@ ggplot(count_df,
     geom_label_repel(
         data = label_df, 
         mapping = aes(fill = container.title, 
-                      label = container.title), 
+                      label = container.title_w), 
         hjust = 0, nudge_y = 20,
         xlim = c(1960, 2100), direction = 'x',
         color = 'black', alpha = .9) +
@@ -57,7 +57,7 @@ ggplot(count_df,
     theme(plot.margin = margin(t = 10, r = 50, b = 10, l = 10))
 
 ggsave(here(out_dir, '02_count.png'), 
-       width = 4.76, height = 3/4 * 4.76, scale = 1.75,
+       width = 3.5, height = 2.4, scale = 2,
        bg = 'white')
 
 meta_ar |> 

### 2024-06-20 scripts/08_topics_over_time.R

@@ -10,6 +10,7 @@ library(arrow)
 library(cowplot)
 library(glue)
 library(ggh4x)
+library(slider)
 
 ## Load data ----
 data_dir = here('data')
@@ -136,7 +137,7 @@ count_plot = function(tmf, vocab, exp, this_k, topics, threshold,
         group_by(k, topic, container.title) |> 
         complete(year = seq.int(min(year), max(year)), 
                  fill = list('gamma' = 0)) |> 
-        mutate(gamma_sm = slider::slide_index_dbl(gamma, year, 
+        mutate(gamma_sm = slide_index_dbl(gamma, year, 
                                                   mean, 
                                                   .complete = TRUE,
                                                   .before = 2, 
@@ -188,7 +189,7 @@ jr_labels = tribble(
     'V22', 'Personality and Individual Differences', 'Person. & Indiv. Diff.', 1985, 13,
     'V22', 'Intelligence', 'Intelligence', 1985, 10,
     'V07', 'Mankind Quarterly', 'Mankind Quarterly', 1995, 30, 
-    'V24', 'Psychological Reports', 'Psychological Reports', 1997, 8
+    'V24', 'Psychological Reports', 'Psychological Reports', 1987, 7
 )
 
 time_gg = plot_50 +
@@ -226,7 +227,8 @@ if (interactive()) focal_topics_gg
 
 ggsave(here(out_dir, '08_focal_topics.png'), 
        plot = focal_topics_gg,
-       height = 3*3, width = 4*3, bg = 'white')
+       width = 7.1, height = .6*8.7, scale = 1.4,
+       bg = 'white')
 
 ## Big grid ----
 big_grid = function(model_file, exp_file, name, plot = TRUE, verbose = TRUE) {

### 2024-06-20 scripts/scratch/beta_table.R

@@ -1,265 +1,0 @@
-library(tidyverse)
-theme_set(theme_minimal())
-library(tmfast)
-library(gt)
-
-# devtools::install_github("ddsjoberg/bstfun")
-library(bstfun)
-library(patchwork)
-
-library(arrow)
-library(here)
-library(glue)
-
-vocab = 'md'
-k = 40
-focal_topics = c('MQ race science' = 'V07', 
-                 'intelligence' = 'V22', 
-                 'intelligence & race' = 'V24')
-topic_labs = names(focal_topics)
-
-out_prefix = '??'
-
-## Load data ----
-out_dir = here('out')
-data_dir = here('data')
-tm_dir = here('data', '04_tm')
-
-model = here(tm_dir, glue('04_{vocab}_tmfast.Rds')) |> 
-    read_rds()
-
-phrases_ar = open_dataset(here(data_dir, '01_phrases.csv'), 
-                          format = 'csv')
-
-source(here('R', 'author_data.R'))
-## Canonicalized names
-canonicalized_df = read_csv(here('data', 
-                                 '00_authors_canonicalized.csv')) |> 
-    filter(!is.na(canonical)) |> 
-    select(-...3)
-
-meta_df = open_dataset(here('data', '01_metadata')) |> 
-    left_join(author_data(), by = 'article_id') |> 
-    collect() |>
-    nest(authors = author)
-
-## Extract matrices
-beta  = tidy(model, k, matrix = 'beta')
-
-
-## Table of highest-beta phrases ----
-beta_gt = beta |> 
-    filter(topic %in% focal_topics) |> 
-    group_by(topic) |> 
-    top_n(15, beta) |> 
-    ungroup() |> 
-    arrange(topic, desc(beta)) |> 
-    select(topic, phrase = token, beta) |> 
-    group_split(topic) |> 
-    bind_cols() |> 
-    select(-starts_with('topic')) |> 
-    gt() |> 
-    tab_spanner(label = topic_labs[1], id = 't1',
-                1:2) |> 
-    tab_spanner(label = topic_labs[2], id = 't2',
-                3:4) |> 
-    tab_spanner(label = topic_labs[3], id = 't3',
-                5:6) |> 
-    fmt_number(starts_with('beta'), 
-               decimals = 4) |> 
-    cols_label(phrase...2 = 'phrase', 
-               beta...3   = 'beta', 
-               phrase...5 = 'phrase', 
-               beta...6   = 'beta', 
[…]

### 2024-06-20 scripts/12_stm.R
ready to submit

@@ -1,3 +1,4 @@
+renv::load(here::here())
 ## This scripts fits a STM topic model, as a robustness check
 library(tidyverse)
 theme_set(theme_bw())

### 2024-07-16 paper/paper.qmd
ready to submit
headers on table 1
commit before reformat

@@
 <!-- 
-PNAS guidelines: <https://www.pnas.org/author-center/submitting-your-manuscript#article-types>
+QSS submission guidelines: 
+<https://direct.mit.edu/qss/pages/submission-guidelines>
 
-Research Report/Direct Submission: The preferred length of these articles is 6 pages, but PNAS allows articles up to a maximum of 12 pages.
-A standard 6-page article is approximately 4,000 words, 50 references, and 4 medium-size graphical elements (i.e., figures and tables).
+- 300-word cover letter on significance and fit
+- non-anonymous
+- 200-word abstract
+- 6 keywords
+- number sections, subsections, and lines
+- figures and tables near first ref
+- author-date APA citations
+- supplemental materials okay
+- methods in "body," not end
+- Acknowledgments, author contribution statements, competing interests statements, funding information, data availability statements, and appendices all at end
 
-- a manuscript file (in any format) including the following:
-    title page (title, author list, classification, keywords)
-    abstract
-    significance statement
-    main text
-    references
-    figures or tables with appropriate legends (may be uploaded separately)
-    SI files (may be uploaded separately)
-- contact and competing interest information for all authors
-- data sharing plans (for all data, documentation, and code used in analysis)
-- funding information and whether an open access license has been selected
-- list of appropriate Editorial Board members, NAS members, and qualified reviewers (minimum of three each) who are experts in the paper's scientific area.
-(A brief justification for suggested reviewers is welcome, particularly for interdisciplinary papers)
-
-- revised submission will require track changes version
-
-figure sizes: 
-Small: approximately 9 cm x 6 cm (3.5 in x 2.4 in)
-Medium: approximately 11 cm x 11 cm (4.3 in x 4.3 in)
-Large: approximately 18 cm x 22 cm (7.1 in x 8.7 in)
+- Articles: original theoretical, empirical, or methodological research (typically between 5,000 and 8,000 words)
 
  -->
-
-*[TODO: 
-- figure out renv::restore()
-    - re-make
-    - copy updated plots
-]*
 
@@
 
-| | | | |
+| name (birth-death) | field | affiliation | WoS doc. count |
 |:--------|:----|:--------|-:|
@@
 
-The public analysis repository, <https://github.com/dhicks/race-science> *DOI will go here]*, includes document metadata and documentwise counts of NLP-extracted noun phrases ("document-term matrices"), along with code to reproduce the analysis of the following sections.
+The public analysis repository, <https://github.com/dhicks/race-science> *[DOI will go here]*, includes document metadata and documentwise counts of NLP-extracted noun phrases ("document-term matrices"), along with code to reproduce the analysis of the following sections.
 

### 2024-07-16 paper/paper.qmd
working conversion to new format

@@
 bibliography: [references.bib, refs-manual.bib]
-csl: "pnas.csl"
 
 format: 
-  pnas-pdf:
+  elsevier-pdf:  # https://github.com/quarto-journals/elsevier
     keep-tex: true
@@
     mathspec: true
-    classoption:
-      - twocolumn
-      - twoside
     fig.env: "figure*"
     number-sections: true
+    journal:
+      name: "NA"
+      cite-style: "authoryear"
+      layout: "onecolumn"
 
@@
   fig-prefix: ""
-  sec-prefix: "*SI Appendix*, section"
 
@@
 
-# Introduction {.unnumbered}
+# Introduction
 
@@
 
-<!-- ## Scientific racism, eugenics, and *Brown* {.unnumbered} -->
+<!-- ## Scientific racism, eugenics, and *Brown*  -->
 
@@
 
-# Results {.unnumbered}
+# Methods and materials
 
-## *Mankind Quarterly* and Pioneer-funded researchers {.unnumbered}
-
-We identified 16 researchers who had received funding from Pioneer; 14 of these researchers had profiles in the Web of Science (WoS) author search, allowing us to identify 13 WoS-indexed journals that had published 6 or more of these authors.
-See @tbl-researchers and figure @fig-wos.
-As the affiliations in @tbl-researchers indicate, almost all of these researchers were faculty at major research universities.
-
-\onecolumn
-
-| name (birth-death) | field | affiliation | WoS doc. count |
-|:--------|:----|:--------|-:|
-| Thomas J. Bouchard, Jr. (1937-) | psychology | Univ. Minnesota | 184 |
-| Brunetto Chiarelli (?-?) | anthropology | | 91 |
-| Hans Eysenck (1916-1997) | psychology | King's College London | 661 |
-| Robert A. Gordon (1932-) | sociology | Johns Hopkins | 0|
-| Linda Gottfredson (1947-) | psychology | Univ. Delaware | 71 |
-| Garrett Hardin (1915-2003) | ecology | UC Santa Barbara | 75 |
-| Joseph M. Horn (1940-) | psychology | UT Austin | 68 |
-| Lloyd Humphreys (1913-2003) | psychology | Univ. Illinois | 0 |
-| Arthur Jensen (1923-2003) | psychology | UC Berkeley | 235 |
-| Michael Levin (1943-) | philosophy | CCNY | 96 |
-| Richard Lynn (1930-2023) | psychology | Ulster University\* | 288 |
-| R. Travis Osborne (1913-2013) | psychology | Univ. Georgia | 59 |
-| J. Phillippe Rushton (1943-2012) | psychology | Univ. Western Ontario | 277 |
-| Audrey M. Shuey (1900-1977) | psychology | Randolph-Macon Women's College | 10 |
-| Philip A. Vernon (1950-) | psychology | Univ. Western Ontario | 227 |
-| Daniel Vining, Jr. (1944-) | demography | Univ. Pennsylvania | 33 |
-
-Table: Pioneer-funded researchers.
-Either identified in [@MillerPioneerFundBankrolling1994] or named on an archive copy of Pioneer's website; along with birth and death dates from Wikipedia, attributed discipline, academic affiliation, and WoS author search result counts.
-Brunetto Chiarelli does not have a Wikipedia page; Richard Lynn's title of Professor Emeritus at Ulster University was withdrawn in 2018.
-{#tbl-researchers}
-
-\twocolumn
-
-![Journals publishing 6 or more Pioneer-funded researchers, Web of Science author search results.](img/pf_wos_results.png){#fig-wos width="7.1in" height="3.5in" fig-env="figure*"}
-
-Figure @fig-wos shows that, while MQ is among the "Pioneer-publishing" journals, a number of mainstream journals are more prominent: *Personality and Individual Differences* (PID), *Intelligence* (Int), *Behavior Genetics* (BG), and *Psychological Reports* (PR).
-In addition, only psychologist Richard Lynn appears to have published heavily in MQ.
-Lynn became an assistant editor of MQ in 1979 (vol.
-XX, Nos.
-1 & 2) and is listed as editor-in-chief on MQ's current website as of 2023-07-21.
-He was also president of Pioneer from the death of psychologist J.
-Phillippe Rushton in 2012 [@BeirichPioneerFundAssets2013] until his own death in 2023.
-By contrast, a number of Pioneer-funded researchers have published a significant number of papers in PID, Int, and to a lesser degree BG: Bouchard, Eysenck, Jensen, Rushton, Vernon, and also Lynn.
-
-*Personality and Individual Differences* (PID) was founded in 1980, with Eysenck as editor-in-chief and an editorial board including Jensen and Lynn.
-In the inaugural editorial, Eysenck identified "studies of the genetic determinants of individual differences in the areas of personality and intelligence" as one of the journal's eight major areas of interest.
-Eysenck remained editor-in-chief until his death in 1997.
-In 2005 the editorial board still included Jensen and Lynn.
-PID was first published by Pergamon Press, a mainstream academic press, and today is published by Elsevier.
-
-*Intelligence* (Int) was founded in 1977, with psychologist Douglas Detterman as editor-in-chief from the founding until 2016.
-Lloyd Humphreys was on the editorial board starting from 1977; by 1990 he had been joined by Jensen and Philip Vernon.
-Richard Lynn joined the editorial board sometime between 1998 and 2002.
-(Archive copies of the Int editorial board page are not available from the journal's website from 1999 through 2001.)  Int has been criticized for including Lynn and Gerhard Meisenberg — who was editor-in-chief of MQ in 2015-18 — on its editorial board until 2018 [@SainiSuperiorReturnRace2019].
-Int is published by Elsevier.
-
-## Topic model analysis identifies race science discourse {.unnumbered}
-
-The fact that Pioneer-funded researchers published heavily in two mainstream psychology journals does not tell us anything about the content of their publications or the claims of their research.
-We assembled a full-text corpus of articles published in MQ, PID, Int, BG, PR, and Behavior and Brain Sciences (BBS) between 1960-2010 and used topic modeling across three different sizes of vocabulary ("large," "medium," and "small") to identify topics discussing race, intelligence, and both.
-(See the materials and methods for a discussion of some important limitations with this corpus: journals published by the American Psychological Association could not be included, BBS bundles focus articles and commentaries together in a way that confounds simple document counts, and an unknown number of relevant documents in PID could not be included.)
-
-![Silge plots [@SilgeTopicModeling2017] and smoothed time series for three focal topics from the medium vocabulary, $k = 40$ model.
-Top row: top 15 terms by $\beta$ (term-topic distribution) for each topic.
-Bottom row: count of articles associated with each topic, by journal and year.
-Article counts use a threshold approach, with $\gamma$ (topic-document distribution) greater than 0.5.
-Thin lines give annual values, thick lines give 5-year running averages.
-Because *Behavior Genetics* and *Behavioral and Brain Sciences* are not prominent in any panel, neither is given a direct label.](img/08_focal_topics.png){#fig-focal width="7.1in" height="5.2in" fig-env="figure*"}
-
-Figure @fig-focal focuses on three topics identified in one of the 24 fitted topic models; see figures [@fig-gamma-lg; @fig-gamma-md; @fig-gamma-sm; @fig-silge-lg; @fig-silge-md; @fig-silge-sm] for all topics from all models, and figures [@fig-grid-lg; @fig-grid-md; @fig-grid-sm] for selected topics from all models.
-In the focal model, topic 07 is strongly associated with MQ:  MQ published dozens of articles in this topic each year, and no other journal ever published more than a handful.
-The "Silge plot," showing the top 15 terms in the topic [@SilgeTopicModeling2017], contains a number of racial terms (`races`, `whites`, `negroes` and potentially `europe`, `africa`, `india`, and `japan`); as well as `book`, likely reflecting the fact that MQ published a number of book reviews, while the mainstream journals either did not publish book reviews or these were not available for the corpus.
-
-Topic 22 is strongly associated with Int and PID in the same way that 07 is associated with MQ.
-The Silge plot does contain `jensen`, as well as a reference to Raymond Cattell, who played a major role in the development of factor analysis and intelligence testing but also advocated for eugenics, fascism, and Nazi race science [@MehlerBeyondismRaymondCattell1997].
-However, the other authors named in this topic — John Horn and Peter Bentler — do not appear to have contributed to scientific racism.
-Instead this topic appears to identify "mainstream" (non-race science discourse) intelligence research, especially factor analysis and the debate over whether intelligence is unidimensional or multidimensional.
-(Three other topics only associated with mainstream intelligence journals and terms were also identified by this topic model.)  This topic indicates that the model is able to distinguish "mainstream" intelligence research from race science discourse.
-
-The Silge plot for topic 24, by contrast, suggests a distinct race science discourse topic, with multiple racial terms (`whites`, `blacks`, `racial_differences`, `races`, `race_differences`) and the names of three prominent Pioneer-funded researchers, `jensen`, `lynn`, and `rushton`.
-Independent qualitative coding of 121 papers in this topic (those with $\gamma > 0.97$ and published in journals other than PR) confirmed this interpretation, with 108 (89%) coded as race science discourse by both authors (Cohen's $\kappa = 0.86$); see @sec-topic-quality.
-
-In almost all years, most papers in topic 24 were published in mainstream journals rather than MQ.
-Jensen's "How Much Can We Boost IQ and Scholastic Achievement?" [@JensenHowMuchCan1969] was published in 1969 in *Harvard Educational Review* (not included in this study), and the time series indicates that, in the early 1970s, there was an increase in articles in topic 24 in both PR and MQ (the only two journals in our corpus that were active at the time).
-MQ shows another sharp increase in the late 1980s; using content analysis, Adams and Pilloud found that psychology was the dominant discipline in MQ in the period 1992-2018 [@AdamsMisAppropriationBiological2021].
-PID published multiple papers in topic 24 almost immediately after it was founded, with Int showing a more gradual increase between the mid-1970s and mid-1990s.
-
-These qualitative results are similar across models using the large and medium vocabularies; see @fig-grid-lg, @fig-grid-md, @fig-grid-sm.
-Across all vocabularies, as $k$ increases, multiple different "intelligence only" topics emerge, apparently picking out different methods or aspects of intelligence research.
-For example, there is often an "intelligence only" topic that is highly associated with BG.
-MQ never has many publications in an "intelligence only" topic, except for $k=10, 20$ in the large vocabulary; in these two models, there is neither a "race only" nor "race-and-intelligence" topic.
-
-In the large and medium vocabularies, there is usually one "race only" topic and/or one "race-and-intelligence" topic, and the "race only" topic is always virtually exclusive to MQ.
-In the medium vocabulary, "race-and-intelligence" is more prevalent in mainstream journals than in MQ; except for $k=20$, where there is no "race only" topic.
-In the large vocabulary, a "race only" topic does not emerge until $k=40$, and there are not simultaneous "race only" and "race-and-intelligence" topics until $k=60$.
-However, in $k=60, 70$ there are both topics, and they follow the same pattern seen in the medium vocabulary: "race only" is almost exclusive to MQ, and "race-and-intelligence" is more prevalent in mainstream journals.
-
-Unlike the other two vocabularies, in the small vocabulary there are two "race only" topics for $k \geq 50$.
-Across values of $k$, topic V04 is more prevalent in PS through about 1990; the other "race only" topic is typically more prevalent in MQ, but also includes a few articles from PS.
-For $k \geq 40$, the small vocabulary models have distinct "race only," "race-and-intelligence," and "intelligence only" topics; "race-and-intelligence" is always predominantly mainstream, and increases in prevalence over the course of the 1980s.
-The "extra" "race only" topic appars to fade out quickly over the course of the 1980s, and its top terms include `race differences`, `rushton`, and `crime`.
-So one possibility is that that the small vocabulary is able to pick out a distinct line of race science discourse that is focused on crime and/or other behaviors, rather than intelligence, and was replaced with a more intelligence-focused line, perhaps in a way that anticipated publication of *The Bell Curve* in 1994.
-However, we think it's more plausible that this separation is an artifact due to the very small size of this vocabulary, with many fewer terms than documents.
-
-## Race science discourse, behavior genetics, and psychology {.unnumbered}
-
-BG published very few articles in any race-and-intelligence topic identified by any of the 24 topic models.
-But, among both (other) academics and the general public, the field of behavior genetics is strongly associated with race science, and specifically race-and-intelligence research.
-Panofsky [@PanofskyMisbehavingScienceControversy2014] argues that, prior to Jensen's 1969 paper, behavior genetics emphasized the study of non-human animals and intentionally avoided associations with eugenics and public controversy more generally.
-In response to Jensen, critics such as Lewontin offered broad critiques of behavior genetics as such, and behavior geneticists in turn adopted a radical conception of academic freedom (without any sense of responsibility for the social implications of academic research) and a siege or wartime mentality, as illustrated by Sandra Scarr's 1986 presidential address to the Behavior Genetics Association (BGA) [@ScarrThreeCheersBehavior1987].
-Scarr's address coincided with the period between 1970-1990 when BG published articles (including Scarr's address itself) in the race science discourse topic (24) of the medium vocabulary, $k = 40$ model.
-
-However, the topic model analysis suggests that, by the 1990s, behavior genetics may have distanced itself somewhat from race science, albeit without directly repudiating it.
-This interpretation was supported by a supplemental analysis that focused on BG specifically; see @sec-bg.
-For example, the 1995 BGA presidential address by Glayde Whitney --- in which he criticized the "Marxist-Lysenkoist denial of genetics" and proposed that differences in murder rates between countries and cities were caused by racial genetic differences in intelligence, empathy, aggression, and impulsivity --- was published in MQ rather than BG.
-Behavior genetics may have been prominent in promoting race science in the past, but appears to have been less receptive to such ideas in more recent history.
-Instead, our results suggest that other subdisciplines of psychology — perhaps especially cognitive psychology — have provided the primary venue for mainstreaming race science.
-
-# Discussion {.unnumbered}
-
-Using bibliometric and text-mining methods, this study finds that Pioneer-funded researchers published heavily in certain mainstream psychology journals, much more than in *Mankind Quarterly*; and that a distinct topic of race science discourse, centered on race and intelligence, can be identified in these same mainstream journals across the period 1960-2010.
-These findings indicate that *Mankind Quarterly* (MQ) was less important as venue for late 20th century race science discourse than mainstream psychology journals, especially *Intelligence* (Int) and *Personality and Individual Differences* (PID).
-Indeed, until the 1990s, MQ published a very different kind of race science discourse from the race-and-intelligence research published in Int and PID.
-During this same period of time, race scientists served in prominent and influential positions in behavior genetics and psychology, including as society presidents and members of journal editorial boards.
-In some cases, some race scientists simultaneously maintained active connections to both the mainstream scientific community and white supremacist organizations [@WinstonScientificRacismNorth2020; @JacksonJr.ArthurJensenEvolutionary2022; @SlobodianUnequalMindHow2023].
-
-We are not identifying any particular scientist as a racist, scientific racist, or race scientist based on the topic model results alone.
-Such claims require an analysis of documentary evidence that goes beyond the scope of the current study [@JacksonJr.ArthurJensenEvolutionary2022].
-We do assume that our readers, like us, regard white supremacy and scientific racism as morally odious and beyond the scope of reasonable debate [@SchroederLimitsDemocratizingScience2022].
-However, scientific research can be appropriated to promote scientific racism — and thus count as race science — even when this is contrary to the intentions of the original researchers themselves [@TaberyWhyStudyingGenetics2015; @GillbornSoftlySoftlyGenetics2016; @CarlsonQuantifyingContextualizingImpact2020; @HennWhyDNANo2021].
-Panofsky et al.
-show how the term "human biodiversity," originally developed for anti-racist purposes by biological anthropologist Jonathan Marks, has been inverted by white supremacists and is now a dogwhistle for biological racial hierarchy [@PanofskyHowWhiteNationalists2021].
-
-Over the past few decades, many professional organizations in genetics and anthropology have made formal statements rejecting race as a biologically meaningful concept [@anthropology1996aapa; @american2018ashg; @board1998aaa; @committee2020asa; @FuentesAAPAStatementRace2019; @nhgri2023; @rotimi2022ashg; @wynshawboris2020ashg], undermining the core assumption of race science.
-The American Psychological Association (APA) --- whose members are predominantly clinical psychologists --- the Federation of Associations in Behavioral & Brain Sciences (FABBS) --- an organization whose purpose is to provide policy recommendations --- and the editors of *Nature Human Behavior* have openly rejected race as biologically meaningful [@american2021apology; @baron2022fabbs; @ScienceMustRespect2022].
-But research-focused psychological organizations like the Association for Psychological Science (APS) and Psychonomic Society (PS) stop short of rejecting race as biologically meaningful when they denounce racism [@aps2021statement; @board2020psychonomic].
-Psychology as a discipline maintains space for several mythological race science narratives, such as allegations of a taboo against race and intelligence research [@jackson2021mythical], claims that scientists engaging in or calling for any kind of anti-racism in the field is an ideological corruption of dispassionate and value-neutral science [@roberts_2022], or arguments that holding race science to the same evidentiary standards as other psychological research is a violation of academic freedom [@herbert2023academic].
-And historically several race scientists have secured important gatekeeping positions within the scientific community, such as on the editorial boards of *Intelligence* and *Personality and Individual Differences*.
-
-Research on the genetics of intelligence has been socially harmful, not just in the somewhat abstract sense of promoting racial stigma, but in the concrete sense of being used to rationalize mass shootings and other acts of racial violence [@PronczukRacistResearcherExposed2022; @MeyerWrestlingSocialBehavioral2023].
-While scientists do not fully control the downstream social effects of their research, like other citizens scientists are responsible for mitigating reasonably foreseeable harms that result from their actions [@BlockIQHeritabilityInequality1974a; @DouglasSciencePolicyValuefree2009; @KitcherArgumentFreeInquiry1997; @KouranyShouldKnowledgeBe2016].
-These actions include not only individual decisions to research this topic or that, but also collective decisions about the allocation of scarce resources such as journal space, research funding, and positions of power and influence.
-Given the way research on intelligence and behavior genetics has been used historically, these fields may be especially susceptible to appropriation by scientific racists.
-This only strengthens the obligations of researchers in these areas to prevent their research from causing harm [@CarlsonQuantifyingContextualizingImpact2020; @OgbunugaforDNABasketballBirthday2022].
-
-How should psychologists exercise these responsibilities?
-We believe restorative or reparative justice provides a useful model [@WenzelRetributiveRestorativeJustice2008; @WormerRestorativeJustice2013].
-Unlike retributive justice (used by almost all criminal justice systems today), the aim of restorative justice is to repair the damaged relationships between victims and perpetrators of injustice.
-Typically, restorative justice requires perpetrators to not only acknowledge their actions and the harms that these actions had on victims, but also to work with victims to identify concrete actions that perpetrators can take to redress or mitigate these harms.
-From this perspective, explicit apologies for contributions to scientific racism are essential as a first step [@FuentesAAPAStatementRace2019; @american2021apology; @JacksonFacingOurHistory2023], but not fully sufficient on their own.
-
-# Methods and materials {.unnumbered}
-
-## Corpus assembly {.unnumbered}
+## Corpus assembly
 
@@
 
-### Mankind Quarterly {.unnumbered}
+### Mankind Quarterly
 
@@
 
-### Mainstream journals {.unnumbered}
+### Mainstream journals
 
@@
 
-## Text preparation {.unnumbered}
+## Text preparation
 
@@
 
-### Data and code availability {.unnumbered}
+### Data and code availability
 
@@
 
-## Vocabulary selection {.unnumbered}
+## Vocabulary selection
 
-We took an information-theoretic approach to vocabulary selection[@HicksProductivityInterdisciplinaryImpacts2021].
+We took an information-theoretic approach to vocabulary selection [@HicksProductivityInterdisciplinaryImpacts2021].
 Consider a game in which I draw a document from the corpus, then a single token from that document.
@@
 
-## Topic modeling {.unnumbered}
+## Topic modeling
 
@@
 Following the approach of [@HicksProductivityInterdisciplinaryImpacts2021], we did not attempt to identify a unique best fitted model for further analysis.
-While the main text of this paper focuses on the medium vocabulary, $k=40$ model as the "median" among the 24 fitted models, qualitatively similar results appear across all models with $k \geq 10$ for all vocabularies; see @sec-topic-quality and @fig-grid-lg, @fig-grid-md, @fig-grid-sm.
+While the results section of this paper focuses on the medium vocabulary, $k=40$ model as the "median" among the 24 fitted models, qualitatively similar results appear across all models with $k \geq 10$ for all vocabularies; see @sec-topic-quality and @fig-grid-lg, @fig-grid-md, @fig-grid-sm.
 
@@
 
-## Limitations {.unnumbered}
+# Results
+
+## *Mankind Quarterly* and Pioneer-funded researchers
+
+We identified 16 researchers who had received funding from Pioneer; 14 of these researchers had profiles in the Web of Science (WoS) author search, allowing us to identify 13 WoS-indexed journals that had published 6 or more of these authors.
+See @tbl-researchers and figure @fig-wos.
+As the affiliations in @tbl-researchers indicate, almost all of these researchers were faculty at major research universities.
+
+| name (birth-death) | field | affiliation | WoS doc. count |
+|:--------|:----|:--------|-:|
+| Thomas J. Bouchard, Jr. (1937-) | psychology | Univ. Minnesota | 184 |
+| Brunetto Chiarelli (?-?) | anthropology | | 91 |
+| Hans Eysenck (1916-1997) | psychology | King's College London | 661 |
+| Robert A. Gordon (1932-) | sociology | Johns Hopkins | 0|
+| Linda Gottfredson (1947-) | psychology | Univ. Delaware | 71 |
+| Garrett Hardin (1915-2003) | ecology | UC Santa Barbara | 75 |
+| Joseph M. Horn (1940-) | psychology | UT Austin | 68 |
+| Lloyd Humphreys (1913-2003) | psychology | Univ. Illinois | 0 |
+| Arthur Jensen (1923-2003) | psychology | UC Berkeley | 235 |
+| Michael Levin (1943-) | philosophy | CCNY | 96 |
+| Richard Lynn (1930-2023) | psychology | Ulster University\* | 288 |
+| R. Travis Osborne (1913-2013) | psychology | Univ. Georgia | 59 |
+| J. Phillippe Rushton (1943-2012) | psychology | Univ. Western Ontario | 277 |
+| Audrey M. Shuey (1900-1977) | psychology | Randolph-Macon Women's College | 10 |
+| Philip A. Vernon (1950-) | psychology | Univ. Western Ontario | 227 |
+| Daniel Vining, Jr. (1944-) | demography | Univ. Pennsylvania | 33 |
+
+Table: Pioneer-funded researchers.
+Either identified in [@MillerPioneerFundBankrolling1994] or named on an archive copy of Pioneer's website; along with birth and death dates from Wikipedia, attributed discipline, academic affiliation, and WoS author search result counts.
+Brunetto Chiarelli does not have a Wikipedia page; Richard Lynn's title of Professor Emeritus at Ulster University was withdrawn in 2018.
+{#tbl-researchers}
+
+![Journals publishing 6 or more Pioneer-funded researchers, Web of Science author search results.](img/pf_wos_results.png){#fig-wos fig-env="figure*"}
+
+Figure @fig-wos shows that, while MQ is among the "Pioneer-publishing" journals, a number of mainstream journals are more prominent: *Personality and Individual Differences* (PID), *Intelligence* (Int), *Behavior Genetics* (BG), and *Psychological Reports* (PR).
+In addition, only psychologist Richard Lynn appears to have published heavily in MQ.
+Lynn became an assistant editor of MQ in 1979 (vol.
+XX, Nos.
+1 & 2) and is listed as editor-in-chief on MQ's current website as of 2023-07-21.
+He was also president of Pioneer from the death of psychologist J.
+Phillippe Rushton in 2012 [@BeirichPioneerFundAssets2013] until his own death in 2023.
+By contrast, a number of Pioneer-funded researchers have published a significant number of papers in PID, Int, and to a lesser degree BG: Bouchard, Eysenck, Jensen, Rushton, Vernon, and also Lynn.
+
+*Personality and Individual Differences* (PID) was founded in 1980, with Eysenck as editor-in-chief and an editorial board including Jensen and Lynn.
+In the inaugural editorial, Eysenck identified "studies of the genetic determinants of individual differences in the areas of personality and intelligence" as one of the journal's eight major areas of interest.
+Eysenck remained editor-in-chief until his death in 1997.
+In 2005 the editorial board still included Jensen and Lynn.
+PID was first published by Pergamon Press, a mainstream academic press, and today is published by Elsevier.
+
+*Intelligence* (Int) was founded in 1977, with psychologist Douglas Detterman as editor-in-chief from the founding until 2016.
+Lloyd Humphreys was on the editorial board starting from 1977; by 1990 he had been joined by Jensen and Philip Vernon.
+Richard Lynn joined the editorial board sometime between 1998 and 2002.
+(Archive copies of the Int editorial board page are not available from the journal's website from 1999 through 2001.)  Int has been criticized for including Lynn and Gerhard Meisenberg — who was editor-in-chief of MQ in 2015-18 — on its editorial board until 2018 [@SainiSuperiorReturnRace2019].
+Int is published by Elsevier.
+
+## Topic model analysis identifies race science discourse
+
+The fact that Pioneer-funded researchers published heavily in two mainstream psychology journals does not tell us anything about the content of their publications or the claims of their research.
+We assembled a full-text corpus of articles published in MQ, PID, Int, BG, PR, and Behavior and Brain Sciences (BBS) between 1960-2010 and used topic modeling across three different sizes of vocabulary ("large," "medium," and "small") to identify topics discussing race, intelligence, and both.
+(See the materials and methods for a discussion of some important limitations with this corpus: journals published by the American Psychological Association could not be included, BBS bundles focus articles and commentaries together in a way that confounds simple document counts, and an unknown number of relevant documents in PID could not be included.)
+
+![Silge plots [@SilgeTopicModeling2017] and smoothed time series for three focal topics from the medium vocabulary, $k = 40$ model.
+Top row: top 15 terms by $\beta$ (term-topic distribution) for each topic.
+Bottom row: count of articles associated with each topic, by journal and year.
+Article counts use a threshold approach, with $\gamma$ (topic-document distribution) greater than 0.5.
+Thin lines give annual values, thick lines give 5-year running averages.
+Because *Behavior Genetics* and *Behavioral and Brain Sciences* are not prominent in any panel, neither is given a direct label.](img/08_focal_topics.png){#fig-focal}
+
+Figure @fig-focal focuses on three topics identified in one of the 24 fitted topic models; see figures [@fig-gamma-lg; @fig-gamma-md; @fig-gamma-sm; @fig-silge-lg; @fig-silge-md; @fig-silge-sm] for all topics from all models, and figures [@fig-grid-lg; @fig-grid-md; @fig-grid-sm] for selected topics from all models.
+In the focal model, topic 07 is strongly associated with MQ:  MQ published dozens of articles in this topic each year, and no other journal ever published more than a handful.
+The "Silge plot," showing the top 15 terms in the topic [@SilgeTopicModeling2017], contains a number of racial terms (`races`, `whites`, `negroes` and potentially `europe`, `africa`, `india`, and `japan`); as well as `book`, likely reflecting the fact that MQ published a number of book reviews, while the mainstream journals either did not publish book reviews or these were not available for the corpus.
+
+Topic 22 is strongly associated with Int and PID in the same way that 07 is associated with MQ.
+The Silge plot does contain `jensen`, as well as a reference to Raymond Cattell, who played a major role in the development of factor analysis and intelligence testing but also advocated for eugenics, fascism, and Nazi race science [@MehlerBeyondismRaymondCattell1997].
+However, the other authors named in this topic — John Horn and Peter Bentler — do not appear to have contributed to scientific racism.
+Instead this topic appears to identify "mainstream" (non-race science discourse) intelligence research, especially factor analysis and the debate over whether intelligence is unidimensional or multidimensional.
+(Three other topics only associated with mainstream intelligence journals and terms were also identified by this topic model.)  This topic indicates that the model is able to distinguish "mainstream" intelligence research from race science discourse.
+
+The Silge plot for topic 24, by contrast, suggests a distinct race science discourse topic, with multiple racial terms (`whites`, `blacks`, `racial_differences`, `races`, `race_differences`) and the names of three prominent Pioneer-funded researchers, `jensen`, `lynn`, and `rushton`.
+Independent qualitative coding of 121 papers in this topic (those with $\gamma > 0.97$ and published in journals other than PR) confirmed this interpretation, with 108 (89%) coded as race science discourse by both authors (Cohen's $\kappa = 0.86$); see @sec-topic-quality.
+
+In almost all years, most papers in topic 24 were published in mainstream journals rather than MQ.
+Jensen's "How Much Can We Boost IQ and Scholastic Achievement?" [@JensenHowMuchCan1969] was published in 1969 in *Harvard Educational Review* (not included in this study), and the time series indicates that, in the early 1970s, there was an increase in articles in topic 24 in both PR and MQ (the only two journals in our corpus that were active at the time).
+MQ shows another sharp increase in the late 1980s; using content analysis, Adams and Pilloud found that psychology was the dominant discipline in MQ in the period 1992-2018 [@AdamsMisAppropriationBiological2021].
+PID published multiple papers in topic 24 almost immediately after it was founded, with Int showing a more gradual increase between the mid-1970s and mid-1990s.
+
+These qualitative results are similar across models using the large and medium vocabularies; see @fig-grid-lg, @fig-grid-md, @fig-grid-sm.
+Across all vocabularies, as $k$ increases, multiple different "intelligence only" topics emerge, apparently picking out different methods or aspects of intelligence research.
+For example, there is often an "intelligence only" topic that is highly associated with BG.
+MQ never has many publications in an "intelligence only" topic, except for $k=10, 20$ in the large vocabulary; in these two models, there is neither a "race only" nor "race-and-intelligence" topic.
+
+In the large and medium vocabularies, there is usually one "race only" topic and/or one "race-and-intelligence" topic, and the "race only" topic is always virtually exclusive to MQ.
+In the medium vocabulary, "race-and-intelligence" is more prevalent in mainstream journals than in MQ; except for $k=20$, where there is no "race only" topic.
+In the large vocabulary, a "race only" topic does not emerge until $k=40$, and there are not simultaneous "race only" and "race-and-intelligence" topics until $k=60$.
+However, in $k=60, 70$ there are both topics, and they follow the same pattern seen in the medium vocabulary: "race only" is almost exclusive to MQ, and "race-and-intelligence" is more prevalent in mainstream journals.
+
+Unlike the other two vocabularies, in the small vocabulary there are two "race only" topics for $k \geq 50$.
+Across values of $k$, topic V04 is more prevalent in PS through about 1990; the other "race only" topic is typically more prevalent in MQ, but also includes a few articles from PS.
+For $k \geq 40$, the small vocabulary models have distinct "race only," "race-and-intelligence," and "intelligence only" topics; "race-and-intelligence" is always predominantly mainstream, and increases in prevalence over the course of the 1980s.
+The "extra" "race only" topic appars to fade out quickly over the course of the 1980s, and its top terms include `race differences`, `rushton`, and `crime`.
+So one possibility is that that the small vocabulary is able to pick out a distinct line of race science discourse that is focused on crime and/or other behaviors, rather than intelligence, and was replaced with a more intelligence-focused line, perhaps in a way that anticipated publication of *The Bell Curve* in 1994.
+However, we think it's more plausible that this separation is an artifact due to the very small size of this vocabulary, with many fewer terms than documents.
+
+## Race science discourse, behavior genetics, and psychology
+
+BG published very few articles in any race-and-intelligence topic identified by any of the 24 topic models.
+But, among both (other) academics and the general public, the field of behavior genetics is strongly associated with race science, and specifically race-and-intelligence research.
+Panofsky [@PanofskyMisbehavingScienceControversy2014] argues that, prior to Jensen's 1969 paper, behavior genetics emphasized the study of non-human animals and intentionally avoided associations with eugenics and public controversy more generally.
+In response to Jensen, critics such as Lewontin offered broad critiques of behavior genetics as such, and behavior geneticists in turn adopted a radical conception of academic freedom (without any sense of responsibility for the social implications of academic research) and a siege or wartime mentality, as illustrated by Sandra Scarr's 1986 presidential address to the Behavior Genetics Association (BGA) [@ScarrThreeCheersBehavior1987].
+Scarr's address coincided with the period between 1970-1990 when BG published articles (including Scarr's address itself) in the race science discourse topic (24) of the medium vocabulary, $k = 40$ model.
+
+However, the topic model analysis suggests that, by the 1990s, behavior genetics may have distanced itself somewhat from race science, albeit without directly repudiating it.
+This interpretation was supported by a supplemental analysis that focused on BG specifically; see @sec-bg.
+For example, the 1995 BGA presidential address by Glayde Whitney --- in which he criticized the "Marxist-Lysenkoist denial of genetics" and proposed that differences in murder rates between countries and cities were caused by racial genetic differences in intelligence, empathy, aggression, and impulsivity --- was published in MQ rather than BG.
+Behavior genetics may have been prominent in promoting race science in the past, but appears to have been less receptive to such ideas in more recent history.
+Instead, our results suggest that other subdisciplines of psychology — perhaps especially cognitive psychology — have provided the primary venue for mainstreaming race science.
+
+# Discussion
+
+Using bibliometric and text-mining methods, this study finds that Pioneer-funded researchers published heavily in certain mainstream psychology journals, much more than in *Mankind Quarterly*; and that a distinct topic of race science discourse, centered on race and intelligence, can be identified in these same mainstream journals across the period 1960-2010.
+These findings indicate that *Mankind Quarterly* (MQ) was less important as venue for late 20th century race science discourse than mainstream psychology journals, especially *Intelligence* (Int) and *Personality and Individual Differences* (PID).
+Indeed, until the 1990s, MQ published a very different kind of race science discourse from the race-and-intelligence research published in Int and PID.
+During this same period of time, race scientists served in prominent and influential positions in behavior genetics and psychology, including as society presidents and members of journal editorial boards.
+In some cases, some race scientists simultaneously maintained active connections to both the mainstream scientific community and white supremacist organizations [@WinstonScientificRacismNorth2020; @JacksonJr.ArthurJensenEvolutionary2022; @SlobodianUnequalMindHow2023].
+
+We are not identifying any particular scientist as a racist, scientific racist, or race scientist based on the topic model results alone.
+Such claims require an analysis of documentary evidence that goes beyond the scope of the current study [@JacksonJr.ArthurJensenEvolutionary2022].
+We do assume that our readers, like us, regard white supremacy and scientific racism as morally odious and beyond the scope of reasonable debate [@SchroederLimitsDemocratizingScience2022].
+However, scientific research can be appropriated to promote scientific racism — and thus count as race science — even when this is contrary to the intentions of the original researchers themselves [@TaberyWhyStudyingGenetics2015; @GillbornSoftlySoftlyGenetics2016; @CarlsonQuantifyingContextualizingImpact2020; @HennWhyDNANo2021].
+Panofsky et al.
+show how the term "human biodiversity," originally developed for anti-racist purposes by biological anthropologist Jonathan Marks, has been inverted by white supremacists and is now a dogwhistle for biological racial hierarchy [@PanofskyHowWhiteNationalists2021].
+
+Over the past few decades, many professional organizations in genetics and anthropology have made formal statements rejecting race as a biologically meaningful concept [@anthropology1996aapa; @american2018ashg; @board1998aaa; @committee2020asa; @FuentesAAPAStatementRace2019; @nhgri2023; @rotimi2022ashg; @wynshawboris2020ashg], undermining the core assumption of race science.
+The American Psychological Association (APA) --- whose members are predominantly clinical psychologists --- the Federation of Associations in Behavioral & Brain Sciences (FABBS) --- an organization whose purpose is to provide policy recommendations --- and the editors of *Nature Human Behavior* have openly rejected race as biologically meaningful [@american2021apology; @baron2022fabbs; @ScienceMustRespect2022].
+But research-focused psychological organizations like the Association for Psychological Science (APS) and Psychonomic Society (PS) stop short of rejecting race as biologically meaningful when they denounce racism [@aps2021statement; @board2020psychonomic].
+Psychology as a discipline maintains space for several mythological race science narratives, such as allegations of a taboo against race and intelligence research [@jackson2021mythical], claims that scientists engaging in or calling for any kind of anti-racism in the field is an ideological corruption of dispassionate and value-neutral science [@roberts_2022], or arguments that holding race science to the same evidentiary standards as other psychological research is a violation of academic freedom [@herbert2023academic].
+And historically several race scientists have secured important gatekeeping positions within the scientific community, such as on the editorial boards of *Intelligence* and *Personality and Individual Differences*.
+
+Research on the genetics of intelligence has been socially harmful, not just in the somewhat abstract sense of promoting racial stigma, but in the concrete sense of being used to rationalize mass shootings and other acts of racial violence [@PronczukRacistResearcherExposed2022; @MeyerWrestlingSocialBehavioral2023].
+While scientists do not fully control the downstream social effects of their research, like other citizens scientists are responsible for mitigating reasonably foreseeable harms that result from their actions [@BlockIQHeritabilityInequality1974a; @DouglasSciencePolicyValuefree2009; @KitcherArgumentFreeInquiry1997; @KouranyShouldKnowledgeBe2016].
+These actions include not only individual decisions to research this topic or that, but also collective decisions about the allocation of scarce resources such as journal space, research funding, and positions of power and influence.
+Given the way research on intelligence and behavior genetics has been used historically, these fields may be especially susceptible to appropriation by scientific racists.
+This only strengthens the obligations of researchers in these areas to prevent their research from causing harm [@CarlsonQuantifyingContextualizingImpact2020; @OgbunugaforDNABasketballBirthday2022].
+
+How should psychologists exercise these responsibilities?
+We believe restorative or reparative justice provides a useful model [@WenzelRetributiveRestorativeJustice2008; @WormerRestorativeJustice2013].
+Unlike retributive justice (used by almost all criminal justice systems today), the aim of restorative justice is to repair the damaged relationships between victims and perpetrators of injustice.
+Typically, restorative justice requires perpetrators to not only acknowledge their actions and the harms that these actions had on victims, but also to work with victims to identify concrete actions that perpetrators can take to redress or mitigate these harms.
+From this perspective, explicit apologies for contributions to scientific racism are essential as a first step [@FuentesAAPAStatementRace2019; @american2021apology; @JacksonFacingOurHistory2023], but not fully sufficient on their own.
+
+## Limitations
 
@@
 
-# Acknowledgments {.unnumbered}
+# Acknowledgments
 
@@
 
-# References {.unnumbered}
+# References
 
@@
 \setcounter{table}{0}
-
-\onecolumn
-
-# SI Appendix {.unnumbered}
+\renewcommand*{\thesection}{S\arabic{section}}
+\setcounter{section}{0}
 

### 2024-07-16 paper/test.qmd
working conversion to new format

@@
-This section uses simulated data to provide a proof-of-concept for our use of topic models to detect the "mainstreaming" of a marginalized topic.
+---
+title: "Race Science in Mainstream Psychology, 1960-2010"
+authors:
+  - name: "Daniel J. Hicks"
+    email: "dhicks4@ucmerced.edu"
+    affiliations: 
+      - id: a
+        name: "University of California, Merced"
+        city: "Merced"
+        country: "CA, USA"
+    orcid: "0000-0001-7945-4416"
+    corresponding: true
+  - name: "Emilio J. C. Lobato"
+    affiliations:
+      - ref: a
+    orcid: "0000-0002-3066-2932"
 
-```{r setup}
-#| message: false
-#| warning: false
-library(tidyverse)
-theme_set(theme_bw())
-library(tmfast)
+abstract: |
+  | Anthropologists and human geneticists have long rejected race science, research that can be utilized to purportedly justify racial inequality and colonialism.
+But race science has persisted, at the very least on the margins of the mainstream scientific community.
+In this study, bibliometric and text mining methods were used to trace the appearance of race science discourse in both mainstream behavioral science journals and the parascholarly journal *Mankind Quarterly* from 1960 to 2010.
+Our analysis found two distinguishable lines of race science discourse.
+One is strongly associated with *Mankind Quarterly*, while the other — emphasizing race and intelligence — is strongly associated with mainstream psychology journals, especially *Intelligence* and *Personality and Individual Differences*.
+We compare psychology's response to race science to those of other fields, and discuss the social responsibility of scientists with respect to race science.
 
-library(glue)
-library(tictoc)
-library(memoise)
+significance: |
+  | Anthropologists and human geneticists have long rejected efforts to purportedly scientifically justify racial inequality and colonialism.
+This study examines the persistence of scientific racism throughout the second half of the twentieth century, examining both mainstream behavioral science journals and a notorious fringe journal, *Mankind Quarterly* (MQ).
+While we expected to find evidence that race science emerged first in MQ and later spread to mainstream journals, we instead found two distinct lines of race science discourse.
+Race-and-intelligence discourse appears first and primarily in two mainstream journals, *Intelligence* and *Personality and Individual Differences*.
+Unlike anthropology and human genetics, psychology has not strongly condemned race science in the mainstream of the field.
 
-set.seed(2024-03-18)
+keywords: 
+  - race science
+  - race and intelligence
+  - psychology
+  - behavior genetics
+  - text mining
+  - topic modeling
 
-## Simulation parameters
-k = 20            ## total number of topics
-N = 100           ## papers per journal
-vocab = 1000      ## vocabulary size
-paper_size = 500  ## length of each paper
+bibliography: [references.bib, refs-manual.bib]
 
-paper_size
-```
+format: 
+  elsevier-pdf:
+    keep-tex: true
+    link-citations: true
+    mathspec: true
+    fig.env: "figure*"
+    number-sections: true
+    journal:
+      name: "NA"
+      layout: "twocolumn"
+      cite-style: "authoryear"
+
+crossref:
+  fig-prefix: ""
+  sec-prefix: "*SI Appendix*, section"
+
+execute:
+  cache: true
+---
+
+# Introduction
+
+By the end of the Second World War, a combination of technical developments in anthropology and genetics, along with revelations of Nazi atrocities, had led most members of the scientific community to reject attempts to scientifically justify racial inequalities and colonialism [@BarkanRetreatScientificRacism1992].
+But eugenic thinking and scientific racism persisted throughout the second half of the twentieth century, and recent work has found that white supremacists continue to attempt to use scientific research to justify racial violence [@DusterBackdoorEugenics2003; @PanofskyHowWhiteNationalists2021; @PronczukRacistResearcherExposed2022].
+
+{{< include ../scripts/11-proof-of-concept.qmd >}}

### 2024-07-19 paper/paper.qmd
ready to submit, pending a query about length

@@
 But race science has persisted, at the very least on the margins of the mainstream scientific community.
-In this study, bibliometric and text mining methods were used to trace the appearance of race science discourse in both mainstream behavioral science journals and the parascholarly journal *Mankind Quarterly* from 1960 to 2010.
+In this study, scientometric and text mining methods were used to trace the appearance of race science discourse in both mainstream behavioral science journals and the parascholarly journal *Mankind Quarterly* from 1960 to 2010.
 Our analysis found two distinguishable lines of race science discourse.
-One is strongly associated with *Mankind Quarterly*, while the other — emphasizing race and intelligence — is strongly associated with mainstream psychology journals, especially *Intelligence* and *Personality and Individual Differences*.
+One is strongly associated with *Mankind Quarterly*, while the other — emphasizing race and intelligence — is more prevalent with mainstream psychology journals, especially *Intelligence* and *Personality and Individual Differences*.
 We compare psychology's response to race science to those of other fields, and discuss the social responsibility of scientists with respect to race science.
@@
       layout: "onecolumn"
+    include-in-header: 
+      text: |
+        \usepackage{lineno}
+        % \linenumbers
 
@@
 - 300-word cover letter on significance and fit
-- non-anonymous
-- 200-word abstract
-- 6 keywords
-- number sections, subsections, and lines
-- figures and tables near first ref
-- author-date APA citations
-- supplemental materials okay
-- methods in "body," not end
-- Acknowledgments, author contribution statements, competing interests statements, funding information, data availability statements, and appendices all at end
-
 - Articles: original theoretical, empirical, or methodological research (typically between 5,000 and 8,000 words)
@@
 @SainiDraperMillionsPhilanthropic2022].
-Pioneer was formed in 1937, during the waning years of the eugenics movement in North America [@BarkanRetreatScientificRacism1992], with the aim of "support[ing] academic research and the \'[sic]dissemination of information, into the 'problem of heredity and eugenics' and 'the problems of race betterment'" [@MehlerFoundationFascismNew1989 p.21, quoting Laughlin].
+Pioneer was formed in 1937, during the waning years of the eugenics movement in North America [@BarkanRetreatScientificRacism1992; @TuckerFundingScientificRacism2002], with the aim of "support[ing] academic research and the \'[sic]dissemination of information, into the 'problem of heredity and eugenics' and 'the problems of race betterment'" [@MehlerFoundationFascismNew1989 p.21, quoting Laughlin].
 MQ was founded in 1960 by biologist R.
@@
 In the landmark case *Brown v Board of Education of Topeka* (1954), the US Supreme Court banned *de jure* educational segregation.
-The Court's decision relied on expert testimony from psychologists and education researchers; but the segregationists also put forward their own experts, including Henry Garrett [@WinstonScienceServiceFar1998; @JacksonScienceSegregationRace2005; @SchafferScientificRacismAgain2007].
-As part of this segregationist reaction, MQ was created to provide a favorable venue for race scientists to publish their views, on the grounds that an "equalitarian dogma" created a censorious "taboo" against their research in mainstream publications [@TuckerFundingScientificRacism2002; @JacksonMythicalTabooRace2020].
+The Court's decision relied on expert testimony from psychologists and education researchers; but segregationists also put forward their own experts, including Henry Garrett [@WinstonScienceServiceFar1998; @JacksonScienceSegregationRace2005; @SchafferScientificRacismAgain2007].
+After *Brown*, Pioneer funded various segregationist efforts across the United States, including lectures on eugenics by Stanford physicist William Shockley [@JacksonScienceSegregationRace2005].
+MQ was created to provide a favorable venue for race scientists to publish their views, on the grounds that an "equalitarian dogma" created a censorious "taboo" against their research in mainstream publications [@TuckerFundingScientificRacism2002; @JacksonMythicalTabooRace2020].
 
@@
 Establishing venues for favorable research is another key element of the "tobacco strategy" for manufacturing doubt [@FernandezPintoKnowBetterNot2017].
-Specifically, we hypothesized that bibliometric and text mining analyses would show race science discourse appearing first in MQ, and subsequently appearing in mainstream publications.
+Specifically, we hypothesized that scientometric and text mining analyses would show race science discourse appearing first and more prominently in MQ, and appearing in mainstream publications only subsequently.
 This hypothesis was not supported.
@@
 # Methods and materials
+
+The primary data for this study was the full text of all available articles published in journals that had published multiple Pioneer-funded researchers, between 1960 (the year MQ was founded) and 2010.
+The end year was chosen for a combination of reasons: easy access to MQ ended at 2004, and we did not want to extend the data too far beyond that point; 2010 gave us a "tidy" 50-year window of coverage; and an end year that wasn't too recent would hopefully allow us to avoid distracting controversies about researchers who are still active today.
 
@@
 
-An electronic archive of *Mankind Quarterly*, from the first issue in 1969 through 2004, is available for free on the open web at the white nationalist website *The Unz Review*.
+An electronic archive of *Mankind Quarterly*, from the first issue in 1960 through 2004, is available for free on the open web at the white nationalist website *The Unz Review*.
 The Python package Beautiful Soup (<https://www.crummy.com/software/BeautifulSoup/>, version not recorded) was used to retrieve every PDF available in this archive.
@@
 To identify suitable mainstream journals for inclusion in the corpus, we first identified academic researchers who had been funded by the Pioneer Fund.
-Pioneer is an American non-profit organization founded in 1937 to fund eugenics research and propaganda [@TuckerFundingScientificRacism2002].
-After the *Brown v.
-Board of Education* ruling in 1954, Pioneer funded various segregationist efforts across the United States, including lectures on eugenics by Stanford physicist William Shockley [@JacksonScienceSegregationRace2005].
 We reviewed a critical profile of Pioneer [@MillerPioneerFundBankrolling1994] as well as an archived page from the organization's own web site (<https://web.archive.org/web/20130103005545/http://www.pioneerfund.org/Grantees.html>), which together listed 16 researchers who had received Pioneer funds.
@@
 Thirteen journals had published 6 or more of the 14 Pioneer-funded researchers.
-We excluded *Mankind Quarterly* (as already included) as well as *Science* and *Nature* (as too general) from further consideration in this side of the corpus.
-Three journals published by the American Psychological Association (APA; *American Psychologist*, *Contemporary Psychology*, and *Journal of Educational Psychology*) had to be excluded due to confusion over who could give us permission to use the archives for a text mining project, with both APA and ProQuest asserting that we needed to get permission from the other entity.
-*European Journal of Personality*, published by SAGE, also had to be excluded because our institutional access only went back to 1999.
-
-The remaining 5 journals are all published by major academic publishers — Elsevier, Springer, or Cambridge University Press — and each item in the entire run of each journal has been assigned a DOI (digital object identifier) for archival purposes.
-We used the Crossref API and `rcrossref` R interface to this API [@ChamberlainRcrossrefClientVarious2019 version 1.1.0.99] to retrieve metadata for each item published from 1960-2010 in each of these journals.
-These metadata included item-level license information — confirming that the text of each item could be used for text mining projects — and a URL to an electronic version of the item.
-These URLs were used to retrieve a XML or PDF version of each item, except for *Personality and Individual Differences*.
-This journal has published a relatively large number of non-article documents, such as book reviews and commentaries, that are not available at the URL included in the Crossref metadata.
-(This is unfortunate, as it was not difficult to find highly relevant documents that we could not automatically retrieve and therefore had to be excluded from our corpus.
-One set of examples is an exchange between Rushton and Flynn on the Flynn effect [@FlynnEvidenceRushtonGenetic1998; @RushtonSecularGainsIQ1998; @FlynnReplyRushtonGang1998].)  Instead we used Elsevier's ScienceDirect API (<https://dev.elsevier.com/>) to independently search and retrieve all available items from *Personality and Individual Differences*.
-This stage of corpus assembly was conducted between 2021-11 and 2022-05.
-
-*Behavioral and Brain Sciences* typically uses a target article + commentary format; in some cases there are dozens of commentaries for a single article.
-In the Crossref DOI metadata, each target article and individual commentary is given its own DOI, with no distinction between contribution types or metadata links between a commentary and the corresponding target article.
-But text is only available in aggregate PDFs that bundle together the target article and all of the commentaries.
-In somes cases this results in PDFs that are hundreds of pages long and might be linked from the Crossref metadata 30+ times.
-In addition, the retrieved PDFs are not perfectly identical, because Cambridge UP's servers add a timestamped watermark to each page when the PDF is requested.
-We attempted to contact Cambridge for assistance but did not receive a response.
-We ultimately used a series of ad hoc measures to mitigate text duplication.
-Approximately 100,000 documents that were 1-2 pages long (per Crossref metadata) were excluded before PDF retrieval.
-After PDF retrieval and text extraction, the timestamped watermarks were removed using a regular expression and the text was hashed using SHA 256.
-Hashes were used to efficiently construct groups of identical documents, and a single document (whichever one happened to be first in the dataframe) was chosen from each hash group for inclusion in the corpus.
-It is plausible that some duplicates made it through this process: where the watermark overlapped with the text, the regular expression likely would have been unable to identify and remove the watermark (a false negative result), and this small difference in the watermarks (not the text) would produce different hashes.
-
-After XML/PDF versions were retrieved, text was extracted using either the `xml2` package [@WickhamXml2ParseXML2023] or `pdftools`.
-Figure @fig-counts and @tbl-counts show the number of fulltext documents included in the corpus.
-All together, the corpus includes 34,896 documents from 6 journals.
-
-## Text preparation
-
-After document retrieval and text extraction, we pre-processed the text using the `spaCy` NLP (natural language processing) Python library [@SpaCyIndustrialstrengthNatural2018 version 1.9.0] and the R API `spacyr` [@BenoitSpacyrWrapperSpaCy2020 version 1.2.1].
-Specifically, we applied regular expressions to remove header/footer copyright notices and hyphenation, used spaCy to annotate and extract noun phrases (eg, "the intelligence test items"), and then cleaned and standardized these phrases (eg, removing the/an/a, coverting all text to lowercase, and replacing all whitespace with underscores: "intelligence_test_items").
-We then counted the occurrence of each noun phrase in the document.
-The aggregated "document-term matrix" was stored in Parquet format [@VohraApacheParquet2016] for performance reasons, and written and read using the `arrow` package for R [@RichardsonArrowIntegrationApache2023 version 7.0.0].
-
-NLP-extracted noun phrases offer a number of advantages over the more traditional unigram ("single word") terms.
-First, noun phrase extraction removes many standard stopwords (articles, common verbs, numbers) without relying on a fixed, a priori stopword list.
-Phrases can be more informative than single terms, for example, distinguishing "intelligence test" from "hypothesis test."  While simple n-gram extraction will include numerous phrases that are not especially meaningful.
-Consider the sentence "Since our first analyses of feeding patterns in rats, we had been using a criterion of 40 minutes (Le Magnen & Tallon 1963; 1966)"[@SclafaniCorrelationCausationStudy1981].
-Bigrams such as "since our" and "criterion 40" will likely be discarded in vocabulary selection, but significantly increase the computational cost of vocabulary selection.
-Noun phrase extraction is therefore more efficient.
-
-After noun phrase extraction, the corpus comprised 43,162,055 total tokens of 6,320,783 distinct phrases.
-
-### Data and code availability
-
-Document metadata retrieved from Crossref does not appear to be covered by any copyright or other intellectual property restrictions.
-However, due to copyright restrictions, we are unable to make document fulltext or Web of Science search results publicly available.
-Code for the corpus assembly and text preparation steps described above is available by request.
-
-The public analysis repository, <https://github.com/dhicks/race-science> *[DOI will go here]*, includes document metadata and documentwise counts of NLP-extracted noun phrases ("document-term matrices"), along with code to reproduce the analysis of the following sections.
-
-## Vocabulary selection
-
-We took an information-theoretic approach to vocabulary selection [@HicksProductivityInterdisciplinaryImpacts2021].
-Consider a game in which I draw a document from the corpus, then a single token from that document.
-I tell you the term, and you have to guess which document I picked.
-Intuitively, highly informative terms (in this project, noun phrase types) are distinctive, allowing you to dramatically narrow down the list of potential documents.
-This "informativeness" of a term can be quantified as the KL (Kullback-Leibler) divergence from a "baseline" uniform distribution across documents to the distribution conditional on the term, which we abbreviate as $\Delta H$.
-Because the most informative terms tend to be typos and OCR errors — these are unique to a single document — we multiply the KL divergence by the logarithm of the total number of occurrences of the term across the entire corpus.
-We refer to the resulting measure as $log(n) \Delta H$ or `ndH`.
-
-In the current project, we found that this `ndH` approach heavily favored recurrent noun phrases in the longest documents.
-Many of the documents published in *Psychological Reports* are extremely short, 1-2 page brief notices of a single study; while many of the documents published in *Brain and Behavioral Sciences* are book-length collections that include a long review article and sometimes dozens of commentaries.
-Very generic noun phrases that happen to appear in BBS can occur orders of magnitude more often than highly distinctive phrases in PS, and so the $log(n)$ factor overwhelms the $\Delta H$ factor.
-
-To address this, we used a different baseline distribution of documents, namely, one in which document probability is proportional to length.
-This makes phrases from short documents much more "surprising" (much less likely to occur according to the baseline), and hence substantially increases their informativeness.
-This was more effective at identifying useful phrases from across the corpus.
-
-A common rule of thumb in topic modeling is that the vocabulary should have about 10 times as many distinct terms as the number of documents in the corpus.
-However, we had some concerns with computational demands here:  the resulting document-term matrix would have roughly $10 \times 33,000^2$ or 10.9 billion entries; with 10% density this would require on the order of 4 GB of memory just for a single copy of the matrix; and most of the analysis was to be conducted on the authors' laptops.
-We therefore chose to work with three smaller vocabularies, $5 \times$, $1 \times$ and $\frac{1}{5} \times$ the number of documents.
-We refer to these as the "large" (174,480 distinct phrases), "medium" (34,896) and "small" (6,979) vocabulary, respectively.
-We compare findings across vocabularies as a robustness check.
-
-## Topic modeling
-
-To fit topic models, we followed the approach proposed by Rohe and Zeng [@RoheVintageFactorAnalysis2020], which uses varimax-rotated partial principal components instead of the variational inference methods used by standard topic model packages such as `stm` [@RobertsStmPackageStructural2019].
-This novel approach was implemented in the R package `tmfast` [@HicksTmfastFitsTopic2023].
-We conducted a simulation study of `tmfast`, which found that it was significantly faster and only slightly less accurate at reconstructing known word-topic and topic-document distributions, compared to `stm` [@HicksTmfastFitsTopic2023].
-We report a proof-of-concept simulation for the current study in @sec-poc.
-
-Topic models were fit for all three vocabularies (large, medium, and small) with $k = 5, 10, 20, ..., 70$ (number of topics), resulting in a total of $24 = 3 \times 8$ models.
-
-Following the approach of [@HicksProductivityInterdisciplinaryImpacts2021], we did not attempt to identify a unique best fitted model for further analysis.
-While the results section of this paper focuses on the medium vocabulary, $k=40$ model as the "median" among the 24 fitted models, qualitatively similar results appear across all models with $k \geq 10$ for all vocabularies; see @sec-topic-quality and @fig-grid-lg, @fig-grid-md, @fig-grid-sm.
-
-We did conduct two further assessments of topic 24 from the medium vocabulary, $k = 40$ model.
-A manual quality assessment achieved consensus that 89% (108/121) of the top papers in this topic were race science discourse; see @sec-topic-quality.
-We also replicated the medium vocabulary, $k = 40$ model using the popular `STM` package for topic modeling; see @sec-stm.
-This model also gave qualitatively similar results.
-Due to the time and computational resources required for these robustness checks, we determined that it would not be feasible for us to repeat them across all 855 topics ($3 \times (5 + 10 + 20 + \cdots + 70)$) and 24 topic models.
-
-# Results
-
-## *Mankind Quarterly* and Pioneer-funded researchers
-
-We identified 16 researchers who had received funding from Pioneer; 14 of these researchers had profiles in the Web of Science (WoS) author search, allowing us to identify 13 WoS-indexed journals that had published 6 or more of these authors.
 See @tbl-researchers and figure @fig-wos.
-As the affiliations in @tbl-researchers indicate, almost all of these researchers were faculty at major research universities.
 
@@
 
+For further consideration in this side of the corpus, we excluded *Mankind Quarterly* (as already included on the other side) as well as *Science* and *Nature* (as too general).
+Three journals published by the American Psychological Association (APA; *American Psychologist*, *Contemporary Psychology*, and *Journal of Educational Psychology*) had to be excluded due to confusion over who could give us permission to use the archives for a text mining project, with both APA and ProQuest asserting that permission could only be granted by the other entity.
+*European Journal of Personality*, published by SAGE, also had to be excluded because our institutional access only went back to 1999.
+
+The remaining 5 journals are all published by major academic publishers — Elsevier, Springer, or Cambridge University Press — and each item in the entire run of each journal has been assigned a DOI (digital object identifier) for archival purposes.
+We used the Crossref API and `rcrossref` R interface to this API [@ChamberlainRcrossrefClientVarious2019 version 1.1.0.99] to retrieve metadata for each item published from 1960-2010 in each of these journals.
+These metadata included item-level license information — confirming that the text of each item could be used for text mining projects — and a URL to an electronic version of the item.
+These URLs were used to retrieve a XML or PDF version of each item, except for *Personality and Individual Differences*.
+This journal has published a relatively large number of non-article documents, such as book reviews and commentaries, that are not available at the URL included in the Crossref metadata.
+(This is unfortunate, as it was not difficult to find highly relevant documents that we could not automatically retrieve and therefore had to be excluded from our corpus.
+One set of examples is an exchange between Rushton and Flynn on the Flynn effect: @FlynnEvidenceRushtonGenetic1998; @RushtonSecularGainsIQ1998; @FlynnReplyRushtonGang1998.)  Instead we used Elsevier's ScienceDirect API (<https://dev.elsevier.com/>) to independently search and retrieve all available items from *Personality and Individual Differences*.
+This stage of corpus assembly was conducted between 2021-11 and 2022-05.
+
+*Behavioral and Brain Sciences* typically uses a target article + commentary format; in some cases there are dozens of commentaries for a single article.
+In the Crossref DOI metadata, each target article and individual commentary is given its own DOI, with no distinction between contribution types and no metadata link between a commentary and the corresponding target article.
+But text is only available in aggregate PDFs that bundle together the target article and all of the commentaries.
+In somes cases this results in PDFs that are hundreds of pages long and might be linked from the Crossref metadata 30+ times.
+In addition, the retrieved PDFs are not perfectly identical, because Cambridge UP's servers add a timestamped watermark to each page when the PDF is requested.
+We attempted to contact Cambridge for assistance but did not receive a response.
+We ultimately used a series of ad hoc measures to mitigate text duplication.
+Approximately 100,000 documents that were 1-2 pages long (per Crossref metadata) were excluded before PDF retrieval.
+After PDF retrieval and text extraction, the timestamped watermarks were removed using a regular expression and the text was hashed using SHA 256.
+Hashes were used to efficiently construct groups of identical documents, and a single document (whichever one happened to be first in the dataframe) was chosen from each hash group for inclusion in the corpus.
+It is plausible that some duplicates made it through this process: where the watermark overlapped with the text, the regular expression likely would have been unable to identify and remove the watermark (a false negative result), and this small difference in the watermarks (not the text) would produce different hashes.
+
+After XML/PDF versions were retrieved, text was extracted using either the `xml2` package [@WickhamXml2ParseXML2023] or `pdftools`.
+Figure @fig-counts and @tbl-counts show the number of fulltext documents included in the corpus.
+All together, the corpus includes 34,896 documents from 6 journals.
+
+## Text preparation
+
+After document retrieval and text extraction, we pre-processed the text using the `spaCy` NLP (natural language processing) Python library [@SpaCyIndustrialstrengthNatural2018 version 1.9.0] and the R API `spacyr` [@BenoitSpacyrWrapperSpaCy2020 version 1.2.1].
+Specifically, we applied regular expressions to remove header/footer copyright notices and hyphenation, used spaCy to annotate and extract noun phrases (eg, "the intelligence test items"), and then cleaned and standardized these phrases (eg, removing the/an/a, coverting all text to lowercase, and replacing all whitespace with underscores: "intelligence_test_items").
+We then counted the occurrence of each noun phrase in the document.
+The aggregated "document-term matrix" was stored in Parquet format [@VohraApacheParquet2016] for performance reasons, and written and read using the `arrow` package for R [@RichardsonArrowIntegrationApache2023 version 7.0.0].
+
+NLP-extracted noun phrases offer a number of advantages over the more traditional unigram ("single word") terms.
+First, noun phrase extraction removes many standard stopwords (articles, common verbs, numbers) without relying on a fixed, a priori stopword list.
+Phrases can be more informative than single terms, for example, distinguishing "intelligence test" from "hypothesis test."  While simple n-gram extraction will include numerous phrases that are not especially meaningful.
+Consider the sentence "Since our first analyses of feeding patterns in rats, we had been using a criterion of 40 minutes (Le Magnen & Tallon 1963; 1966)" [@SclafaniCorrelationCausationStudy1981].
+Bigrams such as "since our" and "criterion 40" will likely be discarded in vocabulary selection, but significantly increase the computational cost of vocabulary selection.
+Noun phrase extraction is therefore more efficient.
+
+After noun phrase extraction, the corpus comprised 43,162,055 total tokens of 6,320,783 distinct phrases.
+
+### Data and code availability
+
+Document metadata retrieved from Crossref does not appear to be covered by any copyright or other intellectual property restrictions.
+However, due to copyright restrictions, we are unable to make document fulltext or Web of Science search results publicly available.
+Code for the corpus assembly and text preparation steps described above is available by request.
+
+The public analysis repository, <https://github.com/dhicks/race-science> *[DOI will go here]*, includes document metadata and documentwise counts of NLP-extracted noun phrases ("document-term matrices"), along with code to reproduce the analysis of the following sections.
+
+## Vocabulary selection
+
+We used a variation on the information-theoretic "`ndH`" approach to vocabulary selection from @HicksProductivityInterdisciplinaryImpacts2021.
+In the current project, we found that that the uniform baseline distribution over documents in this approach heavily favored recurrent noun phrases in the longest documents.
+Many of the documents published in *Psychological Reports* are extremely short, 1-2 page brief notices of a single study; while many of the documents published in *Brain and Behavioral Sciences* are book-length collections that include a long review article and sometimes dozens of commentaries.
+Very generic noun phrases that happen to appear in BBS can occur orders of magnitude more often than highly distinctive phrases in PS, and so the term frequency $log(n)$ factor overwhelms the information gain $\Delta H$ factor.
+
+To address this, we used a different baseline distribution of documents, namely, one in which document probability is proportional to length.
+This makes phrases from short documents much more "surprising" (much less likely to occur according to the baseline), and hence substantially increases their informativeness.
+This was more effective at identifying useful phrases from across the corpus.
+
+A common rule of thumb in topic modeling is that the vocabulary should have about 10 times as many distinct terms as the number of documents in the corpus.
+However, we had some concerns with computational demands here:  the resulting document-term matrix would have roughly $10 \times 33,000^2$ or 10.9 billion entries; with 10% density this would require on the order of 4 GB of memory just for a single copy of the matrix; and most of the analysis was to be conducted on the authors' laptops.
+We therefore chose to work with three smaller vocabularies, $5 \times$, $1 \times$ and $\frac{1}{5} \times$ the number of documents.
+We refer to these as the "large" (174,480 distinct phrases), "medium" (34,896) and "small" (6,979) vocabulary, respectively.
+We compare findings across vocabularies as a robustness check.
+
+## Topic modeling
+
+We used topic models as a clustering (unsupervised machine learning) technique, to identify race science discourse within the corpus.
+
+To fit these topic models, we followed the approach proposed by @RoheVintageFactorAnalysis2020, which uses varimax-rotated partial principal components instead of the variational inference methods used by standard topic model packages such as `stm` [@RobertsStmPackageStructural2019].
+This novel approach was implemented in the R package `tmfast` [@HicksTmfastFitsTopic2023].
+We conducted a simulation study of `tmfast`, which found that it was significantly faster and only slightly less accurate at reconstructing known word-topic and topic-document distributions, compared to `stm` [@HicksTmfastFitsTopic2023].
+We report a specific proof-of-concept simulation for the current study in @sec-poc.
+
+Topic models were fit for all three vocabularies (large, medium, and small) with $k = 5, 10, 20, ..., 70$ (number of topics), resulting in a total of $24 = 3 \times 8$ models.
+
+Following the approach of @HicksProductivityInterdisciplinaryImpacts2021, we did not attempt to identify a unique best fitted model for further analysis.
+While the results section of this paper focuses on the medium vocabulary, $k=40$ model as the "median" among the 24 fitted models, qualitatively similar results appear across all models with $k \geq 10$ for all vocabularies; see @sec-topic-quality and figures @fig-grid-lg, @fig-grid-md, @fig-grid-sm.
+
+To analyze the fitted topic models, we first constructed and manually inspected visualizations of the topic-document distributions (see figures @fig-gamma-lg, @fig-gamma-md, @fig-gamma-sm) and the top (highest-probability) terms for each topic (see figures @fig-silge-lg, @fig-silge-md, @fig-silge-sm).
+This suggested that the models often identified two distinct topics with racial terms, that some of these topics used terms from intelligence research, and that these topics had different occurrence patterns across journals.
+To examine these patterns more systematically, we programmatically identified every topic across all 24 models that used some combination of racial and intelligence terms, and assembled visualization grids of the occurrence of these topics by journal over time (see figures @fig-grid-lg, @fig-grid-md, @fig-grid-sm).
+
+This analysis relied on an interpretive assumption that topics with racial top terms were reliable indicators of race science discourse.
+To confirm this assumption, we conducted two quality assessments of topic 24 from the medium vocabulary, $k = 40$ model.
+A manual assessment by both authors achieved consensus that 89% (108/121) of the top (highest likelihood) papers in this topic were race science discourse; see @sec-topic-quality.
+We also replicated the medium vocabulary, $k = 40$ model using the popular `STM` package for topic modeling; see @sec-stm.
+This model also gave qualitatively similar results.
+Due to the time and computational resources required for these interpretive checks, we determined that it would not be feasible for us to repeat them across all 855 topics ($3 \times (5 + 10 + 20 + \cdots + 70)$) and 24 topic models.
+
+## Limitations
+
+A key limitation of the current study stems from the inability of text mining techniques to draw on a broader cultural context than what is represented in the corpus.
+Scholars of racism and US race relations have noted a shift in the way racist attitudes are understood and expressed, from overt and direct to a subtle "color-blind" racism [@Bonilla-SilvaRacismRacistsColorBlind2006].
+Color-blind racism represents a change in the primary mechanisms by which white privilege is maintained, using non-racialized language in furtherance of the racial status quo both institutionally and individually, and increasingly reliant on the invisibility of socio-cultural mechanisms to preserve inequalities.
+Color-blind racism involves the use of racialized "codes" or "dogwhistles," language that is superficially non-racial but carries racial implications, such as (in the US context) "violent inner city criminals" or "welfare queen" [@SaulDogwhistlesPoliticalManipulation2018].
+Alderfer notes that the racist implications of *The Bell Curve* only emerge gradually, with the opening of the book using the color-blind language of "group differences" [@AlderferScienceNonsciencePsychologists2003].
+These cases suggest further that color-blindness can facilitate white supremacist appropriation of non-racial research — even contrary to the researchers' own intentions — as scientific racist readers re-interpret racial-neutral language as dogwhistles [@GillbornSoftlySoftlyGenetics2016; @WillsAreClustersRaces2017].
+
+The phenomenon of color-blind racism is closely related to ambiguities that cannot be resolved even by human coders familiar with the broader cultural context.
+Two relevant examples are the "Flynn effect," a secular increase in intelligence test scores noted by philosopher James Flynn; and studies of "national IQ," exemplified by the work of Richard Lynn.
+The Flynn effect implies that environmental factors can create group differences in IQ that are comparable to Black-White differences, and thus directly challenges claims that the racial differences are biological.
+Using methods widely regarded as ad hoc, cherry-picking, and generally unreliable, Lynn and collaborators have claimed that national IQ averages are correlated with national GDP [@SamorodnitskyJournalsThatPublished2024].
+Both the Flynn effect and "national IQ" are strongly associated with race-and-intelligence discourse in the US cultural context.
+But scientific journal articles on these topics may not include any racial language at all.
+For example, a meta-analysis of the Flynn effect does not discuss racial differences or use terms such as "White" or "Black" outside the reference list [@TrahanFlynnEffectMetaanalysis2014].
+Should such articles be coded as race science discourse?
+We expect that even readers who accept our definitions of race science and race science discourse will reasonably disagree on how to answer this question.
+Topic models, of course, are not even capable of representing this essential ambiguity.
+
+# Results
+
+## *Mankind Quarterly* and Pioneer-funded researchers
+
+We identified 16 researchers who had received funding from Pioneer; 14 of these researchers had profiles in the Web of Science (WoS) author search, allowing us to identify 13 WoS-indexed journals that had published 6 or more of these authors.
+See @tbl-researchers and figure @fig-wos.
+As the affiliations in @tbl-researchers indicate, almost all of these researchers were faculty at major research universities.
+
 Figure @fig-wos shows that, while MQ is among the "Pioneer-publishing" journals, a number of mainstream journals are more prominent: *Personality and Individual Differences* (PID), *Intelligence* (Int), *Behavior Genetics* (BG), and *Psychological Reports* (PR).
@@
 
-The fact that Pioneer-funded researchers published heavily in two mainstream psychology journals does not tell us anything about the content of their publications or the claims of their research.
-We assembled a full-text corpus of articles published in MQ, PID, Int, BG, PR, and Behavior and Brain Sciences (BBS) between 1960-2010 and used topic modeling across three different sizes of vocabulary ("large," "medium," and "small") to identify topics discussing race, intelligence, and both.
-(See the materials and methods for a discussion of some important limitations with this corpus: journals published by the American Psychological Association could not be included, BBS bundles focus articles and commentaries together in a way that confounds simple document counts, and an unknown number of relevant documents in PID could not be included.)
+Per our hypothesis that MQ served as an "incubator" or "safe space" for race science, we had expected the topic models to identify a distinct race science discourse topic, that this topic would appear first and most prominently in MQ, and only later "spread" to mainstream journals.
+Instead, the models robustly identified two distinct kinds of race science discourse.
+One of these was almost exclusive to MQ; the other originated outside of MQ and was almost always more prevalent in mainstream journals, especially Int and PID.
 
@@
 
-Figure @fig-focal focuses on three topics identified in one of the 24 fitted topic models; see figures [@fig-gamma-lg; @fig-gamma-md; @fig-gamma-sm; @fig-silge-lg; @fig-silge-md; @fig-silge-sm] for all topics from all models, and figures [@fig-grid-lg; @fig-grid-md; @fig-grid-sm] for selected topics from all models.
+Figure @fig-focal focuses on three topics identified in one of the 24 fitted topic models; see figures [@fig-gamma-lg; @fig-gamma-md; @fig-gamma-sm; @fig-silge-lg; @fig-silge-md; @fig-silge-sm] for all topics from all models, and figures [@fig-grid-lg; @fig-grid-md; @fig-grid-sm] for selected topics from all models; we discuss the robustness of our findings across all models below.
 In the focal model, topic 07 is strongly associated with MQ:  MQ published dozens of articles in this topic each year, and no other journal ever published more than a handful.
@@
 
-These qualitative results are similar across models using the large and medium vocabularies; see @fig-grid-lg, @fig-grid-md, @fig-grid-sm.
+These qualitative results are robust across models using the large and medium vocabularies; see @fig-grid-lg, @fig-grid-md, @fig-grid-sm.
 Across all vocabularies, as $k$ increases, multiple different "intelligence only" topics emerge, apparently picking out different methods or aspects of intelligence research.
@@
 
-Using bibliometric and text-mining methods, this study finds that Pioneer-funded researchers published heavily in certain mainstream psychology journals, much more than in *Mankind Quarterly*; and that a distinct topic of race science discourse, centered on race and intelligence, can be identified in these same mainstream journals across the period 1960-2010.
+Using scientometric and text-mining methods, this study finds that Pioneer-funded researchers published heavily in certain mainstream psychology journals, much more than in *Mankind Quarterly*; and that a distinct topic of race science discourse, centered on race and intelligence, can be identified in these same mainstream journals across the period 1960-2010.
 These findings indicate that *Mankind Quarterly* (MQ) was less important as venue for late 20th century race science discourse than mainstream psychology journals, especially *Intelligence* (Int) and *Personality and Individual Differences* (PID).
@@
 
-## Limitations
-
-A key limitation of the current study stems from the inability of text mining techniques to draw on a broader cultural context than what is represented in the corpus.
-Scholars of racism and US race relations have noted a shift in the way racist attitudes are understood and expressed, from overt and direct to a subtle "color-blind" racism [@Bonilla-SilvaRacismRacistsColorBlind2006].
-Color-blind racism represents a change in the primary mechanisms by which white privilege is maintained, using non-racialized language in furtherance of the racial status quo both institutionally and individually, and increasingly reliant on the invisibility of socio-cultural mechanisms to preserve inequalities.
-Color-blind racism involves the use of racialized "codes" or "dogwhistles," language that is superficially non-racial but carries racial implications, such as (in the US context) "violent inner city criminals" or "welfare queen" [@SaulDogwhistlesPoliticalManipulation2018].
-Alderfer notes that the racist implications of *The Bell Curve* only emerge gradually, with the opening of the book using the color-blind language of "group differences" [@AlderferScienceNonsciencePsychologists2003].
-These cases suggest further that color-blindness can facilitate white supremacist appropriation of non-racial research — even contrary to the researchers' own intentions — as scientific racist readers re-interpret racial-neutral language as dogwhistles [@GillbornSoftlySoftlyGenetics2016; @WillsAreClustersRaces2017].
-
-Techniques such as topic modeling may be able to account for discursive shifts from explicit to coded racial language, but only if the corpus contains a critical mass of documents using both sets of terms together.
-This would seem to require either a relatively gradual transition, the inclusion of reflective commentary --- such as scholarship identifying and analyzing the transition --- or both.
-It is unlikely that these requirements are met by the corpus used in this study.
-So it is plausible that the topic models used here have some rate of "false negatives," documents in the corpus that engage in race science discourse but in a color-blind way that avoids the use of the explicit racial terms in the race-and-intelligence topic.
-
-The phenomenon of color-blind racism is closely related to ambiguities that cannot be resolved even by human coders familiar with the broader cultural context.
-Two relevant examples are the "Flynn effect," a secular increase in intelligence test scores noted by philosopher James Flynn; and studies of "national IQ," exemplified by the work of Richard Lynn.
-The Flynn effect implies that environmental factors can create group differences in IQ that are comparable to Black-White differences, and thus directly challenges claims that the racial differences are biological.
-Using methods widely regarded as ad hoc, cherry-picking, and generally unreliable, Lynn and collaborators have claimed that national IQ averages are correlated with national GDP.
-Both the Flynn effect and "national IQ" are strongly associated with race-and-intelligence discourse in the US cultural context.
-But scientific journal articles on these topics may not include any racial language at all.
-For example, a meta-analysis of the Flynn effect does not discuss racial differences or use terms such as "White" or "Black" outside the reference list [@TrahanFlynnEffectMetaanalysis2014].
-Should such articles be coded as race science discourse?
-We expect that even readers who accept our definitions of race science and race science discourse will reasonably disagree on how to answer this question.
-Topic models, of course, are not even capable of representing this essential ambiguity.
-
 # Acknowledgments
@@
 
+\appendix
 \clearpage
@@
 \renewcommand*{\thesection}{S\arabic{section}}
-\setcounter{section}{0}
 
@@
 
-![Count of documents in the corpus, by journal and year.](img/02_count.png){#fig-counts width="3.5in" height="2.4in" fig-env="figure*"}
+![Count of documents in the corpus, by journal and year.](img/02_count.png){#fig-counts  fig-env="figure*"}
 
@@
 
-![Fitted topic models: Gamma (topic-document) distributions, large vocabulary.](img/05_lg_gamma.png){#fig-gamma-lg width="6in" height="4.8in" fig-env="figure*"}
+![Fitted topic models: Gamma (topic-document) distributions, large vocabulary.](img/05_lg_gamma.png){#fig-gamma-lg  fig-env="figure*"}
 
-![Fitted topic models: Gamma (topic-document) distributions, medium vocabulary.](img/05_md_gamma.png){#fig-gamma-md width="6in" height="4.8in" fig-env="figure*"}
+![Fitted topic models: Gamma (topic-document) distributions, medium vocabulary.](img/05_md_gamma.png){#fig-gamma-md  fig-env="figure*"}
 
-![Fitted topic models: Gamma (topic-document) distributions, small vocabulary.](img/05_sm_gamma.png){#fig-gamma-sm width="6in" height="4.8in" fig-env="figure*"}
+![Fitted topic models: Gamma (topic-document) distributions, small vocabulary.](img/05_sm_gamma.png){#fig-gamma-sm  fig-env="figure*"}
 

### 2025-01-03 paper/paper.qmd
table S2

@@
 authors:
-  - name: "Daniel J. Hicks"
+  - name: "D.J. Hicks"
     email: "dhicks4@ucmerced.edu"
@@
         \usepackage{lineno}
-        % \linenumbers
+        \linenumbers
 
@@
   fig-prefix: ""
+  tbl-prefix: ""
 
@@
 @SainiDraperMillionsPhilanthropic2022].
-Pioneer was formed in 1937, during the waning years of the eugenics movement in North America [@BarkanRetreatScientificRacism1992; @TuckerFundingScientificRacism2002], with the aim of "support[ing] academic research and the \'[sic]dissemination of information, into the 'problem of heredity and eugenics' and 'the problems of race betterment'" [@MehlerFoundationFascismNew1989 p.21, quoting Laughlin].
+Pioneer was formed in 1937, during the waning years of the eugenics movement in North America [@BarkanRetreatScientificRacism1992; @TuckerFundingScientificRacism2002], with the aim of "support[ing] academic research and the \'dissemination of information, into the 'problem of heredity and eugenics' and 'the problems of race betterment'" [@MehlerFoundationFascismNew1989 p.21, quoting Laughlin, unbalanced quotation marks in Mehler].
 MQ was founded in 1960 by biologist R.
@@
 
-In light of the scientific racist origins of Pioneer and MQ, and the significant attention MQ has received in recent historiography, we hypothesized that MQ might have served as an "incubator" for race science discourse, with ideas originating in MQ before appearing within the scientific mainstream.
-Establishing venues for favorable research is another key element of the "tobacco strategy" for manufacturing doubt [@FernandezPintoKnowBetterNot2017].
-Specifically, we hypothesized that scientometric and text mining analyses would show race science discourse appearing first and more prominently in MQ, and appearing in mainstream publications only subsequently.
-This hypothesis was not supported.
+Scholars in the agnotology ("theory of ignorance") literature have examined the ways that powerful economic interests, especially the fossil fuel and tobacco industries, have "manufactured doubt," creating a inaccurate public perception of scientific uncertainty to delay public protective regulation [@OreskesMerchantsDoubtHow2011; @FernandezPintoKnowBetterNot2017; @KouranyScienceProductionIgnorance2020].
+Following @OreskesMerchantsDoubtHow2011, @FernandezPintoKnowBetterNot2017 identified five "core strategies or mechanisms" of the "tobacco strategy," showing how this strategy was used to manufacture doubt surrounding climate change, pharmaceuticals, and the causes of the 2008 financial crisis.
+
+Race science has not been heavily studied using an agnotology framework [though see @MillsRacialContract2014 and @KouranyMightScientificIgnorance2020].
+But the qualitative historiography suggests that actors such as the Pioneer Fund have deployed a version of the tobacco strategy to promote misleading public perceptions of scientific uncertainty about (or even scientific support for) racial hierarchy.
+(Indeed, the origins of the "tobacco strategy" are often dated to a 1954 advertising campaign by the tobacco industry, 17 years after Pioneer was founded.
+Scientific racists and the eugenics movement might have been deploying the "tobacco strategy" for decades before the tobacco industry.)
+
+The Pioneer Fund itself clearly exemplifies the strategies of supporting friendly research; and figures such as Shockley and Jensen exemplify the strategy of recruiting high-profile, distinguished scientists *[Jackson?
+on funding Shockley]*.
+Using agent-based models, *[cailin etc.]* have shown that *[st]*.
+
+For the current project, we hypothesized that MQ and certain mainstream journals played a key role in another strategy identified by @FernandezPintoKnowBetterNot2017, the "echo chamber effect." We theorize this effect as involving three components: the "margins," the "mainstream," and the "boundary."  @FernandezPintoKnowBetterNot2017 argues that, using sponsored journals and conferences, "the tobacco industry circumvented peer review standards in publication, without compromising its façade of research supporter" (59).
+These marginal venues provide a "safe space" in which favorable conceptual frameworks and empirical research (or pseudo-research) could be "incubated," protected from the destructive critical scrutiny they would receive in the mainstream.
+Boundary venues are accepted as part of the mainstream, but cite and draw on work from the margins, thereby mainstreaming these marginal ideas, that is, making them appear to be part of the legitimate body of mainstream scientific research.
+
+MQ does not appear to have ever been considered a mainstream journal; it has always existed firmly on the margins.
+If MQ provided the "safe space" for "incubating" race science, away from mainstream criticism, then boundary venues would be necessary to disseminate race science into the scientific mainstream.
+We therefore hypothesized (H1) that scientometric and text mining analyses would show race science discourse appearing first and most prominently in MQ, and appearing in mainstream publications only subsequently.
+
+Specifically, among both (other) academics and the general public, the field of behavior genetics is strongly associated with race science [@PanofskyMisbehavingScienceControversy2014].
+We therefore hypothesized (H2) that the journal *Behavior Genetics* (BG) would be a major boundary venue for the dissemination of race science discourse.
+
+Neither of these hypotheses were supported.
 Instead, mainstream psychology journals provided a venue for a distinct form of race science discourse, based on intelligence research and originating outside of MQ.
+And BG has not been a major venue for either form of race science discourse.
 
@@
 Noun phrase extraction is therefore more efficient.
+For examples of previous research using noun phrases, see @HicksImpactingCapabilitiesConceptual2018; @HicksProductivityInterdisciplinaryImpacts2021.
 
@@
 ![Silge plots [@SilgeTopicModeling2017] and smoothed time series for three focal topics from the medium vocabulary, $k = 40$ model.
-Top row: top 15 terms by $\beta$ (term-topic distribution) for each topic.
+Top row: top 15 terms by $\beta$ (term-topic distribution) for each topic; text version in table @tbl-silge.
 Bottom row: count of articles associated with each topic, by journal and year.
@@
 
-Figure @fig-focal focuses on three topics identified in one of the 24 fitted topic models; see figures [@fig-gamma-lg; @fig-gamma-md; @fig-gamma-sm; @fig-silge-lg; @fig-silge-md; @fig-silge-sm] for all topics from all models, and figures [@fig-grid-lg; @fig-grid-md; @fig-grid-sm] for selected topics from all models; we discuss the robustness of our findings across all models below.
+Figure @fig-focal focuses on three topics identified in one of the 24 fitted topic models; see table @tbl-silge for a table version of the Silge plot in @fig-focal; figures [@fig-gamma-lg; @fig-gamma-md; @fig-gamma-sm; @fig-silge-lg; @fig-silge-md; @fig-silge-sm] for all topics from all models, and figures [@fig-grid-lg; @fig-grid-md; @fig-grid-sm] for selected topics from all models; we discuss the robustness of our findings across all models below.
 In the focal model, topic 07 is strongly associated with MQ:  MQ published dozens of articles in this topic each year, and no other journal ever published more than a handful.
@@
 Panofsky [@PanofskyMisbehavingScienceControversy2014] argues that, prior to Jensen's 1969 paper, behavior genetics emphasized the study of non-human animals and intentionally avoided associations with eugenics and public controversy more generally.
-In response to Jensen, critics such as Lewontin offered broad critiques of behavior genetics as such, and behavior geneticists in turn adopted a radical conception of academic freedom (without any sense of responsibility for the social implications of academic research) and a siege or wartime mentality, as illustrated by Sandra Scarr's 1986 presidential address to the Behavior Genetics Association (BGA) [@ScarrThreeCheersBehavior1987].
+In response to Jensen, critics such as Lewontin offered broad critiques of behavior genetics as such, and behavior geneticists in turn adopted a radical conception of academic freedom (without any sense of responsibility for the social implications of academic research) along with a siege or wartime mentality, as illustrated by Sandra Scarr's 1986 presidential address to the Behavior Genetics Association (BGA) [@ScarrThreeCheersBehavior1987].
 Scarr's address coincided with the period between 1970-1990 when BG published articles (including Scarr's address itself) in the race science discourse topic (24) of the medium vocabulary, $k = 40$ model.
@@
 For example, the 1995 BGA presidential address by Glayde Whitney --- in which he criticized the "Marxist-Lysenkoist denial of genetics" and proposed that differences in murder rates between countries and cities were caused by racial genetic differences in intelligence, empathy, aggression, and impulsivity --- was published in MQ rather than BG.
-Behavior genetics may have been prominent in promoting race science in the past, but appears to have been less receptive to such ideas in more recent history.
+Behavior genetics may have been prominent in promoting race science in the past, but appear to have been less receptive to such ideas in more recent history.
 Instead, our results suggest that other subdisciplines of psychology — perhaps especially cognitive psychology — have provided the primary venue for mainstreaming race science.
@@
 
+```{r}
+#| tbl-cap: "Top 15 terms, by beta value, from selected topics of the medium vocabulary, $k = 40$ model"
+#| label: tbl-silge
+#| echo: false
+library(gt)
+
+readRDS(file.path('..', 'out', '08_silge_gt.Rds')) |>
+    # cols_width(everything() ~ pct(100/6)) |>
+    tab_options(table.font.size = 10)
+```
+
 # Proof of concept {#sec-poc}
@@
 For $k=5, 10$ the distributions are quite noisy, difficult to interpret, and generally don't form very strong visual clusters.
-As $k$ increases from 20 through 40, coherent bands start to appear for journal-distinctive topics in MQ, BG, Int and, to a lesser degree, PID.
+For moderate values of $k$, 20 through 40, coherent bands start to appear for journal-distinctive topics in MQ, BG, Int and, to a lesser degree, PID.
 Note that the distinctive topics for MQ and Int do not obviously coincide.
-Above about $k=50$, the models appear to identify more fine-grained topics and the coherent bands fade.
+At a low level, these journal-distinctive topics reflect different distributions of noun phrases, and prima facie are likely explained by the relatively narrow scope of these four journals; PR is a generalist journal covering all of psychology and BBS is an interdisciplinary cognitive science journal.
+Above about $k=50$, the models appear to identify more fine-grained topics, the coherent bands fade, and the topics do not obviously distinguish different journals.
 

### 2025-01-03 scripts/08_topics_over_time.R

@@ -1,7 +1,8 @@
-renv::load(here::here())
+# renv::load(here::here())
 library(tidyverse)
 library(tidytext)
 theme_set(theme_bw())
+library(gt)
 
 library(tmfast)
 library(here)
@@ -217,6 +218,24 @@ silge_gg = tidy(md_tmf, matrix = 'beta', k = 40) |>
     labs(y = 'β') +
     facet_wrap(vars(topic), scales = 'free_y')
 if (interactive()) silge_gg
+
+## Table version of Silge plot
+silge_gt = tidy(md_tmf, matrix = 'beta', k = 40) |> 
+    filter(topic %in% c('V07', 'V22', 'V24')) |> 
+    group_by(topic) |> 
+    top_n(15, beta) |> 
+    arrange(topic, desc(beta)) |> 
+    ungroup() |> 
+    pivot_wider(names_from = topic,
+                values_from = c(token, beta), 
+                names_vary = 'slowest') |> 
+    unnest(everything()) |> 
+    gt() |> 
+    fmt_number(decimals = 4) |> 
+    tab_spanner_delim(delim = '_', reverse = TRUE)
+silge_gt
+write_rds(silge_gt, here(out_dir, '08_silge_gt.Rds'))
+
 
 # silge_gg / time_gg +
 #     plot_layout()

### 2025-01-18 paper/paper.qmd
switch quarto format and fix references

@@
 
-bibliography: [references.bib, refs-manual.bib]
+bibliography: [references.yaml, refs-manual.bib]
 
 format: 
-  elsevier-pdf:  # https://github.com/quarto-journals/elsevier
+  aft-pdf:
     keep-tex: true
@@
     number-sections: true
-    journal:
-      name: "NA"
-      cite-style: "authoryear"
-      layout: "onecolumn"
+    cite-method: citeproc
     include-in-header: 
@@
             \linenumbers
+    geometry:
+        - left = 1.5in
+        - right = 1.5in
+
+#  elsevier-pdf:  # https://github.com/quarto-journals/elsevier
+#    keep-tex: true
+#    link-citations: true
+#    mathspec: true
+#    fig.env: "figure*"
+#    number-sections: true
+#    journal:
+#      name: "NA"
+#      cite-style: "authoryear"
+#      layout: "onecolumn"
+#    include-in-header: 
+#      text: |
+#        \usepackage{lineno}
+#        \linenumbers
+#    cite_method: "citeproc"
 
@@
 
-By the end of the Second World War, a combination of technical developments in anthropology and genetics, along with revelations of Nazi atrocities, had led most members of the scientific community to reject attempts to scientifically justify racial inequalities and colonialism [@BarkanRetreatScientificRacism1992].
-But eugenic thinking and scientific racism persisted throughout the second half of the twentieth century, and recent work has found that white supremacists continue to attempt to use scientific research to justify racial violence [@DusterBackdoorEugenics2003; @PanofskyHowWhiteNationalists2021; @PronczukRacistResearcherExposed2022].
+By the end of the Second World War, a combination of technical developments in anthropology and genetics, intellectual campaigns by the likes of Franz Boas and W.E.B.
+Du Bois, and revelations of Nazi atrocities meant that the dominant position within the scientific community was one that rejected attempts to scientifically justify racial inequalities and colonialism [@BarkanRetreatScientificRacism1992; @YudellRaceUnmaskedBiology2014; @JacksonDarwinismDemocracyRace2017].
+But eugenic thinking and scientific racism persisted throughout the second half of the twentieth century, and recent work has found that white supremacists continue to attempt to use scientific research to justify racial violence [@DusterBackdoorEugenics2003; @JacksonScienceSegregationRace2005; @PanofskyHowWhiteNationalists2021; @BirdTypologicalThinkingHuman2024; @PronczukRacistResearcherExposed2022].
 
@@
 *Race science discourse* refers to a broader category, any treatment of race science as a legitimate area of scientific research.
+(@sec-operationalizing discusses how this conception of "race science" might be operationalized in practice.
+@BirdTypologicalThinkingHuman2024 give a concrete example of how a particular figure was transformed into a scientific racist meme.)
 
@@
 @WinstonScientificRacismNorth2020;
-@AdamsMisAppropriationBiological2021;
+@AdamsMisappropriationBiologicalAnthropology2021;
 @SainiDraperMillionsPhilanthropic2022].
@@
 The Court's decision relied on expert testimony from psychologists and education researchers; but segregationists also put forward their own experts, including Henry Garrett [@WinstonScienceServiceFar1998; @JacksonScienceSegregationRace2005; @SchafferScientificRacismAgain2007].
-After *Brown*, Pioneer funded various segregationist efforts across the United States, including lectures on eugenics by Stanford physicist William Shockley [@JacksonScienceSegregationRace2005].
+After *Brown*, Pioneer funded various segregationist efforts across the United States, including lectures on eugenics by Stanford physicist William Shockley [@TuckerFundingScientificRacism2002; @JacksonScienceSegregationRace2005].
 MQ was created to provide a favorable venue for race scientists to publish their views, on the grounds that an "equalitarian dogma" created a censorious "taboo" against their research in mainstream publications [@TuckerFundingScientificRacism2002; @JacksonMythicalTabooRace2020].
@@
 Race science has not been heavily studied using an agnotology framework [though see @MillsRacialContract2014 and @KouranyMightScientificIgnorance2020].
-But the qualitative historiography suggests that actors such as the Pioneer Fund have deployed a version of the tobacco strategy to promote misleading public perceptions of scientific uncertainty about (or even scientific support for) racial hierarchy.
+The qualitative historiography suggests that actors such as the Pioneer Fund have deployed a version of the tobacco strategy to promote misleading public perceptions of scientific uncertainty about (or even scientific support for) racial hierarchy.
 (Indeed, the origins of the "tobacco strategy" are often dated to a 1954 advertising campaign by the tobacco industry, 17 years after Pioneer was founded.
@@
 
-The Pioneer Fund itself clearly exemplifies the strategies of supporting friendly research; and figures such as Shockley and Jensen exemplify the strategy of recruiting high-profile, distinguished scientists *[Jackson?
-on funding Shockley]*.
-Using agent-based models, *[cailin etc.]* have shown that *[st]*.
+The Pioneer Fund itself clearly exemplifies the strategies of supporting friendly research; and figures such as Shockley and Jensen exemplify the strategy of recruiting high-profile, distinguished scientists [@TuckerFundingScientificRacism2002].
 
-For the current project, we hypothesized that MQ and certain mainstream journals played a key role in another strategy identified by @FernandezPintoKnowBetterNot2017, the "echo chamber effect." We theorize this effect as involving three components: the "margins," the "mainstream," and the "boundary."  @FernandezPintoKnowBetterNot2017 argues that, using sponsored journals and conferences, "the tobacco industry circumvented peer review standards in publication, without compromising its façade of research supporter" (59).
+For the current project, we hypothesized that MQ and certain mainstream journals played a key role in another strategy identified by @FernandezPintoKnowBetterNot2017, the "echo chamber effect." We theorize this effect as involving three components: the "margins," the "mainstream," and the "bridge."  @FernandezPintoKnowBetterNot2017 argues that, using sponsored journals and conferences, "the tobacco industry circumvented peer review standards in publication, without compromising its façade of research supporter" (59).
 These marginal venues provide a "safe space" in which favorable conceptual frameworks and empirical research (or pseudo-research) could be "incubated," protected from the destructive critical scrutiny they would receive in the mainstream.
-Boundary venues are accepted as part of the mainstream, but cite and draw on work from the margins, thereby mainstreaming these marginal ideas, that is, making them appear to be part of the legitimate body of mainstream scientific research.
+Bridge venues are accepted as part of the mainstream, but cite and draw on work from the margins, thereby mainstreaming these marginal ideas, that is, making them appear to be part of the legitimate body of mainstream scientific research.
 
 MQ does not appear to have ever been considered a mainstream journal; it has always existed firmly on the margins.
-If MQ provided the "safe space" for "incubating" race science, away from mainstream criticism, then boundary venues would be necessary to disseminate race science into the scientific mainstream.
-We therefore hypothesized (H1) that scientometric and text mining analyses would show race science discourse appearing first and most prominently in MQ, and appearing in mainstream publications only subsequently.
+If MQ provided the "safe space" for "incubating" race science, away from mainstream criticism, then bridge venues would be necessary to disseminate race science into the scientific mainstream.
+We therefore hypothesized (H1) that scientometric and text mining analyses would show race science discourse appearing first and most prominently in MQ, and appearing in mainstream publications only subsequently, with certain venues serving as the bridge.
 
 Specifically, among both (other) academics and the general public, the field of behavior genetics is strongly associated with race science [@PanofskyMisbehavingScienceControversy2014].
-We therefore hypothesized (H2) that the journal *Behavior Genetics* (BG) would be a major boundary venue for the dissemination of race science discourse.
+We therefore hypothesized (H2) that the journal *Behavior Genetics* (BG) would be a major bridge venue for the dissemination of race science discourse.
 
@@
 
-After document retrieval and text extraction, we pre-processed the text using the `spaCy` NLP (natural language processing) Python library [@SpaCyIndustrialstrengthNatural2018 version 1.9.0] and the R API `spacyr` [@BenoitSpacyrWrapperSpaCy2020 version 1.2.1].
+After document retrieval and text extraction, we pre-processed the text using the `spaCy` NLP (natural language processing) Python library [@HonnibalSpaCyIndustrialstrengthNatural2018 version 1.9.0] and the R API `spacyr` [@BenoitSpacyrWrapperSpaCy2020 version 1.2.1].
 Specifically, we applied regular expressions to remove header/footer copyright notices and hyphenation, used spaCy to annotate and extract noun phrases (eg, "the intelligence test items"), and then cleaned and standardized these phrases (eg, removing the/an/a, coverting all text to lowercase, and replacing all whitespace with underscores: "intelligence_test_items").
@@
 
-A key limitation of the current study stems from the inability of text mining techniques to draw on a broader cultural context than what is represented in the corpus.
+With respect to the aim of detecting and studying the prevalence of race science, a critical limitation of our methods is that they cannot distinguish race science from the broader category of race science discourse.
+Topic modeling represents a document as an unstructured "bag of words"; while it can detect that the document discusses both intelligence and racial groups, for example, it cannot detect whether the author endorses or critiques race-and-intelligence research.
+For the same reason, topic modeling cannot determine whether an article purports — implicitly or explicitly — to justify racial inequality and colonialism by appealing to scientific (or pseudo-scientific) findings, and thus it cannot distinguish race science discourse from scientific racism.
+We therefore can only make claims about finding race science discourse, except when manual inspection or prior research indicates otherwise [e.g., @JacksonJr.ArthurJensenEvolutionary2022].
+
+To generalize this limitation, the text mining techniques we use here are unable to draw on a broader cultural context than what is represented in the corpus.
 Scholars of racism and US race relations have noted a shift in the way racist attitudes are understood and expressed, from overt and direct to a subtle "color-blind" racism [@Bonilla-SilvaRacismRacistsColorBlind2006].
@@
 Topic models, of course, are not even capable of representing this essential ambiguity.
+
+Recently, some computational scholars have argued that foundational large language models (LLMs) are able to encode broad cultural contexts [@UnderwoodWhyAINeeds2024] and produce sophisticated analyses of arguments [@CaulfieldTwoMeaningsReasoning2024].
+If they are right, future LLM-based methods might possibly be able to distinguish supporters from critics of race-and-intelligence research, detect the use of racial dogwhistles, and perhaps even reliably identify the fuzzy boundaries of essentially contestable categories such as race science.
+However, such methods are beyond the scope of the project we present here.
 
@@
 ![Silge plots [@SilgeTopicModeling2017] and smoothed time series for three focal topics from the medium vocabulary, $k = 40$ model.
+Topics 07 and 24 are examples of the two distinct types of race science discourse, while topic 22 is "mainstream" or non-race science intelligence research.
 Top row: top 15 terms by $\beta$ (term-topic distribution) for each topic; text version in table @tbl-silge.
@@
 Jensen's "How Much Can We Boost IQ and Scholastic Achievement?" [@JensenHowMuchCan1969] was published in 1969 in *Harvard Educational Review* (not included in this study), and the time series indicates that, in the early 1970s, there was an increase in articles in topic 24 in both PR and MQ (the only two journals in our corpus that were active at the time).
-MQ shows another sharp increase in the late 1980s; using content analysis, Adams and Pilloud found that psychology was the dominant discipline in MQ in the period 1992-2018 [@AdamsMisAppropriationBiological2021].
+MQ shows another sharp increase in the late 1980s; using content analysis, Adams and Pilloud found that psychology was the dominant discipline in MQ in the period 1992-2018 [@AdamsMisappropriationBiologicalAnthropology2021].
 PID published multiple papers in topic 24 almost immediately after it was founded, with Int showing a more gradual increase between the mid-1970s and mid-1990s.
@@
 
-Using scientometric and text-mining methods, this study finds that Pioneer-funded researchers published heavily in certain mainstream psychology journals, much more than in *Mankind Quarterly*; and that a distinct topic of race science discourse, centered on race and intelligence, can be identified in these same mainstream journals across the period 1960-2010.
+Using scientometric and text-mining methods, this study finds that Pioneer-funded researchers published heavily in certain mainstream psychology journals, much more than in *Mankind Quarterly*; and that a distinct version of race science discourse, centered on race and intelligence, can be identified in these same mainstream journals across the period 1960-2010.
 These findings indicate that *Mankind Quarterly* (MQ) was less important as venue for late 20th century race science discourse than mainstream psychology journals, especially *Intelligence* (Int) and *Personality and Individual Differences* (PID).
@@
 In some cases, some race scientists simultaneously maintained active connections to both the mainstream scientific community and white supremacist organizations [@WinstonScientificRacismNorth2020; @JacksonJr.ArthurJensenEvolutionary2022; @SlobodianUnequalMindHow2023].
+
+## *[implications for agnotology]*
+
+In terms of the margins-mainstream-bridge model of the "echo chamber effect," there was no particular bridge that carried ideas originating in MQ on the margins to mainstream journals.
+Within the scope of our data, the race-and-intelligence version of race science discourse was always a mainstream phenomenon.
+
+The circulation of ideas is a prominent theme in the empirical agnotology literature.
+Researchers have often examined the public circulation of ideas — how an industry attempts to manage public concerns about its products in marketing, press releases, and interviews — as well as the circulation of ideas within a company or industry itself — as in memos among intramural scientists and management [@ProctorGoldenHolocaustOrigins2012].
+In a series of articles, Supran and Oreskes have shown that Exxon/ExxonMobil lied — knowingly made false statements — about climate change, contrasting skeptical and denialist statements in public-facing documents with skillful climate projections by the company's own scientists [@SupranAssessingExxonMobilsClimate2017; @SupranAddendumAssessingExxonMobils2020; @SupranRhetoricFrameAnalysis2021; @SupranAssessingExxonMobilsGlobal2023].
+
+Less empirical research has examined the circulation of ideas within the scientific community.
+@SupranAssessingExxonMobilsGlobal2023 include several peer-reviewed journal articles in their content analysis, but compare them to intramural documents and public statements, rather than the broader scientific literature.
+@FernandezPintoKnowBetterNot2017 gives three examples to illustrate the "tobacco strategy": the joint public relations-research funding campaign by the tobacco industry itself; medical ghostwriting by the pharmaceutical industry; and efforts by conservative think tanks to promote a simplistic account of the causes of the Great Recession that shifted blame away from deregulation of the finance industry.
+Both the first and last cases emphasize the circulation of ideas to the public, and the third does not involve academic researchers in any significant way.
+Medical ghostwriting is concerning as a strategy for circulating false or misleading claims in the scientific literature.
+But it has not been a major case in the agnotology literature in recent years.
+This might reflect less concern about ghosting in biomedical research over time: on 2025-01-16, a Web of Science search for "medic* ghostwrit*" returned 85 total papers published between 2003-2024, with the number of publications per year peaking at 11 in both 2010 and 2011 and declining to 1-2 per year in 2023 and 2024.
+
+Some researchers have used agent-based models (computer simulation methods) to study the circulation of misinformation within scientific communities under agnogenic ("ignorance-creating") influences such as "industry selection" and increased dissemination of favorable views [@HolmanExperimentationIndustrialSelection2017; @WeatherallHowBeatScience2018; @OConnorMisinformationAgeHow2019].
+While these models can give insight into how certain mechanisms might work, it is difficult to extrapolate from them to actual cases [@FernandezPintoEpistemicLandscapesReloaded2018; @HicksOpenScienceReplication2023].
+
+Understanding the public circulation of race science — its translation into scientific racism — is essential.
+But our findings suggest the circulation of race science within the scientific community warrants further investigation.
+Using the set of race science discourse articles identified in this project as a starting point, future research might examine authorship and/or citation patterns.
+For example, have race science discourse articles primarily produced by a small and stable set of authors, or a larger set that has changed over time?
+Or, to what extent have race science discourse articles been cited within non-race science articles?
+
+## *[implications for psychology]*
 
@@
 Over the past few decades, many professional organizations in genetics and anthropology have made formal statements rejecting race as a biologically meaningful concept [@anthropology1996aapa; @american2018ashg; @board1998aaa; @committee2020asa; @FuentesAAPAStatementRace2019; @nhgri2023; @rotimi2022ashg; @wynshawboris2020ashg], undermining the core assumption of race science.
-The American Psychological Association (APA) --- whose members are predominantly clinical psychologists --- the Federation of Associations in Behavioral & Brain Sciences (FABBS) --- an organization whose purpose is to provide policy recommendations --- and the editors of *Nature Human Behavior* have openly rejected race as biologically meaningful [@american2021apology; @baron2022fabbs; @ScienceMustRespect2022].
+The American Psychological Association (APA) --- whose members are predominantly clinical psychologists --- the Federation of Associations in Behavioral & Brain Sciences (FABBS) --- an organization whose purpose is to provide policy recommendations --- and the editors of *Nature Human Behavior* have openly rejected race as biologically meaningful [@american2021apology; @baron2022fabbs; @EditorsScienceMustRespect2022].
 But research-focused psychological organizations like the Association for Psychological Science (APS) and Psychonomic Society (PS) stop short of rejecting race as biologically meaningful when they denounce racism [@aps2021statement; @board2020psychonomic].
@@
 
+# Operationalizing "race science" in practice {#sec-operationalizing}
+
+A reviewer raised a concern that our conception of race science involved a counterfactual that "could be difficult to confirm or refute." A full discussion of how, for example, a research funder might operationalize race science in order to subject it to heightened critical scrutiny before funding, is beyond the scope of the current project.
+Here we sketch three frameworks that might be used to operationalize race science, either independently or in combination.
+
+First, some research has actually been utilized in actual scientific racist propaganda, for example, in memes that circulate on fringe discussion boards or as citations in the manifestos of mass shooters.
+This research is ipso facto race science.
+Taken on its own, this framework is only useful after the fact.
+But it might still play a substantial role in, for example, a decision whether to retract a publication or adjudicate whether a researcher has engaged in misconduct.
+
+This first framework can also provide important evidence for the second framework, which draws on the *reasonable person* standard from the law of torts.
+In other words, the second framework asks whether a reasonable scientific peer — well-informed about what kinds of research have been utilized by scientific racists and how, and concerned to avoid this happening in the future — would consider the research likely to be utilized in a similar way in the future.
+@DouglasSciencePolicyValuefree2009 argues that, because we all have a moral responsibility to exercise reasonable foresight over our activities, scientists have a specific responsibility to exercise such foresight over the social consequences of their research.
+Reasonable foresight does require projecting counterfactual scenarios; but these can be based in casuistical reasoning (comparisons with actual cases, as per the first framework) and are not different in kind from the projections made by research funders and hiring committees.
+
+@PenceChallengesCommunityScience2023 argue that a reasonable foresight framework encounters difficulties as scientific communities become more nebulous and vaguely defined, and propose that exercising reasonable foresight should include community engagement and incorporating outsider perspectives.
+So, specifically, cognitive psychologists might make a point of engaging with groups that have been marginalized by race-and-intelligence research and working with scholars who have studied how and why certain kinds of research are taken up by scientific racists.
+
+The third framework starts with exactly this question of how and why research is taken up by scientific racists. In other words, what features of a project make it race science?
+
+Schematically, a research project can be thought of as using *methods* to produce *findings* that (aim to) answer *research questions*.
+Each of these three elements requires various assumptions to be logically relevant to the other two  [@LonginoScienceSocialKnowledge1990; @LloydAdaptationismLogicResearch2015].
+For example, a particular project might use data from a national genome bank and a genome-wide association study (GWAS; method) to develop a polygenic score estimator of educational attainment (finding) with the aim of answering the question "To what extent is cognitive ability genetic?" This project must assume that a sample from a single nation is representative of humanity in general (to be relevant to the generalized framing of the question), that "cognitive ability" is a well-defined construct and worth investigating, that educational attainment is an accurate measure of cognitive ability, that the SNPs included in the genome data capture genetically-caused variation, that analysis of the polygenic scores can recover/measure this variation, that variation can be partitioned into genetic and environmental causes [@KellerMirageSpaceNature2010], and so on.
+
+Some of these assumptions — or the findings themselves — can support scientific racist claims, either through strict deductive entailment (the kind studied in formal logic courses) or through *implicature* [@DavisImplicature2024].
+The concept of "implicature" refers to the phenomenon that language use goes beyond the strict deductive entailments — "we mean more than we say." Consider the following exchange between a teenager and their parent.
+
+| 
+| Teen: Sasha's going to Cold Stone. 
+| Parent: The dishes are still in the sink. 
+| Teen: I don't have practice in the morning. 
+| Parent: Be home by 10.
+
+Taken strictly, the exchange is nonsensical: Sasha going to Cold Stone has nothing to do with the dishes, which in turn has nothing to do with whether the teen has practice or what time they should be home.
+
+Instead the exchange runs on implicature.
+The teen's first statement implies but does not strictly entail a request, that they be granted permission to go to Cold Stone as well.
+The parent's first response is an implied denial of this request, on the grounds that the teen hasn't done their chores.
+The teen responds with an implied promise to finish their chores in the morning, and the parent implies permission with a proviso.
+
+Implicature is context-dependent; the exchange above assumes that the teen wants to do the same thing as Sasha, that they're responsible for doing the dishes, usually have practice in the mornings, and so on.
+In a different context, without this assumptions in play, the implicature doesn't work.
+
+We can think of a social context as providing unstated assumptions that serve as additional premises, and thereby allow language users to draw additional inferences from a statement that go beyond its strict entailments.
+Returning to the hypothetical GWAS of educational attainment study, suppose the study found that "educational attainment is 21% heritable." This supports the simplification "educational attainment is substantially heritable." Then, in a context in which educational attainment is assumed to reflect intelligence, this has the implicature "intelligence is substantially heritable." Next, with the genetic/environmental partition assumption and in a context where genetic determinism is widely held, this has the further implicature that "intelligence is innate." Combined with studies that claim differences in mean IQ across racial groups (or even just a contextual assumption to this effect), an assumption that IQ also measures intelligence, and statistical essentialism (the idea that statistical averages reflect group essences), we get to the scientific racist implicature that "some races are innately more intelligent than others."
+
+In this way, a study that did not mention race — or, indeed, might have mentioned race only to highlight that the sample was racially homogeneous and therefore the results could not be generalized — could still be utilized to support scientific racist implicatures and therefore be race science [@KitcherArgumentFreeInquiry1997; @WillsAreClustersRaces2017; @BirdTypologicalThinkingHuman2024].
+(Note that denying the generalizability of results entails that the study is not deductively relevant to the generalized research question.)
+
+Our example in the main text, of the Flynn effect as legitimizing race science even though it poses an empirical challenge to certain aspects of race science (hereditarianism), can be analyzed in a similar way.
+A project investigating the Flynn effect might pose a research question that assumes — implicitly or explicitly — that it is scientifically possible but not certain than hereditarianism is false.
+This entails that it is scientifically possible but not certain that hereditarianism is true.
+Then, in a context where it is widely assumed that any scientifically possible but uncertain claim is a legitimate subject for scientific research, this supports the implicature that hereditarianism is a legitimate subject for scientific research.
+In this way, Flynn effect research can be race science discourse.
+
+Because key assumptions that enable scientific racist implicatures are often implicit and context-specific, it will often be difficult to empirically prove to the satisfaction of everyone concerned that a particular project is (or is not) race science — outside of cases where the research has actually been utilized in actual scientific racist propaganda.
+Authorities within scientific communities — funders, journal editors, hiring and tenure committees — will need to decide what kinds of evidence are appropriate for adjudicating claims of race science, and whether there is sufficient evidence to reach a conclusion (one way or the other) in any given case.
+The frameworks we have offered here can help these authorities systematically and empirically — that is to say, scientifically — adjudicate such cases.
+But race science is still an unavoidably political concept.
+
 # Proof of concept {#sec-poc}
@@
 Fourteen of the remaining documents had been published since 2010 (after the scope of the primary study).
-Among these, one obituary and two papers celebrated the work of John Loehlin, a prominent race-and-intelligence researcher who had been Director of the American Eugenics Society from 1968-1972 [@JohnLoehlin19262020; @WaldmanIntroductionFestschriftJohn2014; @PlominGenotypeEnvironmentCorrelationEra2014].
+Among these, one obituary and two papers celebrated the work of John Loehlin, a prominent race-and-intelligence researcher who had been Director of the American Eugenics Society from 1968-1972 [@TurkheimerJohnLoehlin192620202020; @WaldmanIntroductionFestschriftJohn2014; @PlominGenotypeEnvironmentCorrelationEra2014].
 Another was a retrospective of the work of Lindon Eaves, a geneticist who made at least one notable contribution to debates on the Scarr-Rowe hypothesis (interactions between heritability, race, and class).

### 2025-01-18 scripts/07_umap_interactive.R

@@ -1,5 +1,4 @@
 ## Formerly, this script generated two-dimensional UMAP embeddings, for use in interactive and static visualizations.  These visualizations were dropped from the analysis pipeline.  Because calculating these embeddings is computationally intensive, the relevant lines (last 3 lines at the end of the script) are commented out.  This script is still used in the workflow to generate a quality check spreadsheet for the focal topic.  
-renv::load(here::here())
 library(tidyverse)
 theme_set(theme_minimal())
 library(tmfast)
@@ -217,3 +216,5 @@ make_umap('md', 40, spreadsheet = 'V24', visualization = FALSE)
 # make_umap('md', c(20, 30, 40, 50, 60, 70), seed = seed, force = force)
 # make_umap('sm', c(20, 30, 40, 50, 60, 70), seed = seed, force = force)
 # make_umap('lg', c(20, 30, 40, 50, 60, 70), seed = seed, force = force)
+make_umap('md', 40, spreadsheet = c('V07', 'V22', 'V24'), 
+          visualization = FALSE)

### 2025-01-21 paper/paper.qmd
finished w/ revisions?

@@
 abstract: |
-  | Anthropologists and human geneticists have long rejected race science, research that can be utilized to purportedly justify racial inequality and colonialism.
-But race science has persisted, at the very least on the margins of the mainstream scientific community.
+  | Anthropologists and human geneticists have long rejected efforts to purportedly scientifically justify racial inequality and colonialism.
+But race science has persisted, at least on the margins of the mainstream scientific community.
 In this study, scientometric and text mining methods were used to trace the appearance of race science discourse in both mainstream behavioral science journals and the parascholarly journal *Mankind Quarterly* from 1960 to 2010.
@@
 By the end of the Second World War, a combination of technical developments in anthropology and genetics, intellectual campaigns by the likes of Franz Boas and W.E.B.
-Du Bois, and revelations of Nazi atrocities meant that the dominant position within the scientific community was one that rejected attempts to scientifically justify racial inequalities and colonialism [@BarkanRetreatScientificRacism1992; @YudellRaceUnmaskedBiology2014; @JacksonDarwinismDemocracyRace2017].
+Du Bois, and revelations of Nazi atrocities meant that the dominant position within the scientific community rejected attempts to scientifically justify racial inequalities and colonialism [@BarkanRetreatScientificRacism1992; @YudellRaceUnmaskedBiology2014; @JacksonDarwinismDemocracyRace2017].
 But eugenic thinking and scientific racism persisted throughout the second half of the twentieth century, and recent work has found that white supremacists continue to attempt to use scientific research to justify racial violence [@DusterBackdoorEugenics2003; @JacksonScienceSegregationRace2005; @PanofskyHowWhiteNationalists2021; @BirdTypologicalThinkingHuman2024; @PronczukRacistResearcherExposed2022].
@@
 *Race science discourse* refers to a broader category, any treatment of race science as a legitimate area of scientific research.
-(@sec-operationalizing discusses how this conception of "race science" might be operationalized in practice.
-@BirdTypologicalThinkingHuman2024 give a concrete example of how a particular figure was transformed into a scientific racist meme.)
+(@sec-operationalizing discusses how this conception of race science might be operationalized in practice.
+@BirdTypologicalThinkingHuman2024 give a concrete example of how a particular data visualization was transformed into a scientific racist meme.)
 
@@
 So, we suggest, they are quite likely to use a "middle ground" heuristic to understand these debates as ordinary, reasonable disagreement among mainstream scientists.
-Some experts say one thing, other experts say something else, and so "the truth is somewhere in the middle."  Tobacco and other industries have exploited this heuristic to "manufacture doubt" about the hazards of their products [@FernandezPintoKnowBetterNot2017].
-In this way technical critiques of race science based on the Flynn effect could legitimize race science.
+Some experts say one thing, other experts say something else, and so "the truth is somewhere in the middle."  In this way technical critiques of race science based on the Flynn effect could legitimize race science.
 So, while empirical research supporting the Flynn effect is usually not itself race science, it can still be race science discourse.
@@
 
-Scholars in the agnotology ("theory of ignorance") literature have examined the ways that powerful economic interests, especially the fossil fuel and tobacco industries, have "manufactured doubt," creating a inaccurate public perception of scientific uncertainty to delay public protective regulation [@OreskesMerchantsDoubtHow2011; @FernandezPintoKnowBetterNot2017; @KouranyScienceProductionIgnorance2020].
+Scholars in the agnotology ("theory of ignorance") literature have examined the ways that powerful economic interests — especially the fossil fuel and tobacco industries — have "manufactured doubt," creating an inaccurate public perception of scientific uncertainty to delay public protective regulation [@OreskesMerchantsDoubtHow2011; @FernandezPintoKnowBetterNot2017; @KouranyScienceProductionIgnorance2020].
 Following @OreskesMerchantsDoubtHow2011, @FernandezPintoKnowBetterNot2017 identified five "core strategies or mechanisms" of the "tobacco strategy," showing how this strategy was used to manufacture doubt surrounding climate change, pharmaceuticals, and the causes of the 2008 financial crisis.
 
-Race science has not been heavily studied using an agnotology framework [though see @MillsRacialContract2014 and @KouranyMightScientificIgnorance2020].
+Race science has not been heavily studied using an agnotology framework (though see @MillsRacialContract2014 and @KouranyMightScientificIgnorance2020).
 The qualitative historiography suggests that actors such as the Pioneer Fund have deployed a version of the tobacco strategy to promote misleading public perceptions of scientific uncertainty about (or even scientific support for) racial hierarchy.
@@
 
-The Pioneer Fund itself clearly exemplifies the strategies of supporting friendly research; and figures such as Shockley and Jensen exemplify the strategy of recruiting high-profile, distinguished scientists [@TuckerFundingScientificRacism2002].
+The Pioneer Fund itself clearly exemplifies the strategy of supporting friendly research; and figures such as Shockley and Jensen exemplify the strategy of recruiting high-profile, distinguished scientists [@TuckerFundingScientificRacism2002].
 
 For the current project, we hypothesized that MQ and certain mainstream journals played a key role in another strategy identified by @FernandezPintoKnowBetterNot2017, the "echo chamber effect." We theorize this effect as involving three components: the "margins," the "mainstream," and the "bridge."  @FernandezPintoKnowBetterNot2017 argues that, using sponsored journals and conferences, "the tobacco industry circumvented peer review standards in publication, without compromising its façade of research supporter" (59).
-These marginal venues provide a "safe space" in which favorable conceptual frameworks and empirical research (or pseudo-research) could be "incubated," protected from the destructive critical scrutiny they would receive in the mainstream.
+These kinds of marginal venues provide a "safe space" in which favorable conceptual frameworks and empirical research (or pseudo-research) could be developed, protected from the destructive critical scrutiny they would receive in the mainstream.
 Bridge venues are accepted as part of the mainstream, but cite and draw on work from the margins, thereby mainstreaming these marginal ideas, that is, making them appear to be part of the legitimate body of mainstream scientific research.
@@
 MQ does not appear to have ever been considered a mainstream journal; it has always existed firmly on the margins.
-If MQ provided the "safe space" for "incubating" race science, away from mainstream criticism, then bridge venues would be necessary to disseminate race science into the scientific mainstream.
+If MQ provided the "safe space" for developing race science, protected away from mainstream criticism, then bridge venues would be necessary to disseminate these ideas into the scientific mainstream.
 We therefore hypothesized (H1) that scientometric and text mining analyses would show race science discourse appearing first and most prominently in MQ, and appearing in mainstream publications only subsequently, with certain venues serving as the bridge.
@@
 Specifically, among both (other) academics and the general public, the field of behavior genetics is strongly associated with race science [@PanofskyMisbehavingScienceControversy2014].
-We therefore hypothesized (H2) that the journal *Behavior Genetics* (BG) would be a major bridge venue for the dissemination of race science discourse.
+We therefore hypothesized (H2) that the journal *Behavior Genetics* (BG) would be a significant bridge venue for the dissemination of race science discourse.
 
 Neither of these hypotheses were supported.
-Instead, mainstream psychology journals provided a venue for a distinct form of race science discourse, based on intelligence research and originating outside of MQ.
-And BG has not been a major venue for either form of race science discourse.
+Instead, mainstream psychology journals provided a venue for a distinct form of race science discourse, based on intelligence research and originating in two mainstream journals — *Intelligence* and *Personality and Individual Differences* — rather than MQ.
+And BG has not been a significant venue for either form of race science discourse.
 
@@
 An electronic archive of *Mankind Quarterly*, from the first issue in 1960 through 2004, is available for free on the open web at the white nationalist website *The Unz Review*.
-The Python package Beautiful Soup (<https://www.crummy.com/software/BeautifulSoup/>, version not recorded) was used to retrieve every PDF available in this archive.
+The Python package Beautiful Soup [@BeautifulSoupScreenscrapingLibrary, version not recorded] was used to retrieve every PDF available in this archive.
 After identifying some gaps (missing issues) in the archive from *The Unz Review*, we manually retrieved PDFs for further articles using ProQuest.
@@
 Thirteen journals had published 6 or more of the 14 Pioneer-funded researchers.
-See @tbl-researchers and figure @fig-wos.
+See table @tbl-researchers and figure @fig-wos.
 
@@
 In addition, the retrieved PDFs are not perfectly identical, because Cambridge UP's servers add a timestamped watermark to each page when the PDF is requested.
-We attempted to contact Cambridge for assistance but did not receive a response.
+We contacted Cambridge UP for assistance but did not receive a response.
 We ultimately used a series of ad hoc measures to mitigate text duplication.
@@
 After XML/PDF versions were retrieved, text was extracted using either the `xml2` package [@WickhamXml2ParseXML2023] or `pdftools`.
-Figure @fig-counts and @tbl-counts show the number of fulltext documents included in the corpus.
+Figure @fig-counts and table @tbl-counts show the number of fulltext documents included in the corpus.
 All together, the corpus includes 34,896 documents from 6 journals.
@@
 We identified 16 researchers who had received funding from Pioneer; 14 of these researchers had profiles in the Web of Science (WoS) author search, allowing us to identify 13 WoS-indexed journals that had published 6 or more of these authors.
-See @tbl-researchers and figure @fig-wos.
-As the affiliations in @tbl-researchers indicate, almost all of these researchers were faculty at major research universities.
+See table @tbl-researchers and figure @fig-wos.
+As the affiliations in table @tbl-researchers indicate, almost all of these researchers were faculty at major research universities.
 
@@
 In the focal model, topic 07 is strongly associated with MQ:  MQ published dozens of articles in this topic each year, and no other journal ever published more than a handful.
-The "Silge plot," showing the top 15 terms in the topic [@SilgeTopicModeling2017], contains a number of racial terms (`races`, `whites`, `negroes` and potentially `europe`, `africa`, `india`, and `japan`); as well as `book`, likely reflecting the fact that MQ published a number of book reviews, while the mainstream journals either did not publish book reviews or these were not available for the corpus.
+The "Silge plot," showing the top 15 terms in the topic [@SilgeTopicModeling2017], contains a number of racial terms (`races`, `whites`, `negroes` and potentially `europe`, `africa`, `india`, and `japan`); as well as `book`, likely reflecting the fact that MQ published a number of book reviews, while the mainstream journals either did not publish book reviews or (in the case of PID) these were not available for the corpus.
 
@@
 The Silge plot does contain `jensen`, as well as a reference to Raymond Cattell, who played a major role in the development of factor analysis and intelligence testing but also advocated for eugenics, fascism, and Nazi race science [@MehlerBeyondismRaymondCattell1997].
-However, the other authors named in this topic — John Horn and Peter Bentler — do not appear to have contributed to scientific racism.
+However, the other authors named in this topic — John Horn and Peter Bentler — do not appear to have contributed to race science.
 Instead this topic appears to identify "mainstream" (non-race science discourse) intelligence research, especially factor analysis and the debate over whether intelligence is unidimensional or multidimensional.
@@
 
-## Race science discourse, behavior genetics, and psychology
+## Race science discourse and behavior genetics
 
@@
 But, among both (other) academics and the general public, the field of behavior genetics is strongly associated with race science, and specifically race-and-intelligence research.
-Panofsky [@PanofskyMisbehavingScienceControversy2014] argues that, prior to Jensen's 1969 paper, behavior genetics emphasized the study of non-human animals and intentionally avoided associations with eugenics and public controversy more generally.
-In response to Jensen, critics such as Lewontin offered broad critiques of behavior genetics as such, and behavior geneticists in turn adopted a radical conception of academic freedom (without any sense of responsibility for the social implications of academic research) along with a siege or wartime mentality, as illustrated by Sandra Scarr's 1986 presidential address to the Behavior Genetics Association (BGA) [@ScarrThreeCheersBehavior1987].
+@PanofskyMisbehavingScienceControversy2014 argues that, prior to Jensen's 1969 paper, behavior genetics emphasized the study of non-human animals and intentionally avoided associations with eugenics and public controversy more generally.
+In response to Jensen, critics such as biologist Richard Lewontin offered broad critiques of behavior genetics as such, and behavior geneticists in turn adopted a radical conception of academic freedom (namely, one that denies any sense of responsibility for the social implications of academic research) along with a siege or wartime mentality, as illustrated by Sandra Scarr's 1986 presidential address to the Behavior Genetics Association (BGA) [@ScarrThreeCheersBehavior1987].
 Scarr's address coincided with the period between 1970-1990 when BG published articles (including Scarr's address itself) in the race science discourse topic (24) of the medium vocabulary, $k = 40$ model.
@@
 For example, the 1995 BGA presidential address by Glayde Whitney --- in which he criticized the "Marxist-Lysenkoist denial of genetics" and proposed that differences in murder rates between countries and cities were caused by racial genetic differences in intelligence, empathy, aggression, and impulsivity --- was published in MQ rather than BG.
-Behavior genetics may have been prominent in promoting race science in the past, but appear to have been less receptive to such ideas in more recent history.
+Behavior genetics may have been prominent in promoting race science in the past, but appears to have been less receptive to such ideas in more recent history.
 Instead, our results suggest that other subdisciplines of psychology — perhaps especially cognitive psychology — have provided the primary venue for mainstreaming race science.
@@
 Using scientometric and text-mining methods, this study finds that Pioneer-funded researchers published heavily in certain mainstream psychology journals, much more than in *Mankind Quarterly*; and that a distinct version of race science discourse, centered on race and intelligence, can be identified in these same mainstream journals across the period 1960-2010.
-These findings indicate that *Mankind Quarterly* (MQ) was less important as venue for late 20th century race science discourse than mainstream psychology journals, especially *Intelligence* (Int) and *Personality and Individual Differences* (PID).
+These findings indicate that *Mankind Quarterly* (MQ) was a less important venue for late 20th century race science discourse than mainstream psychology journals — especially *Intelligence* (Int) and *Personality and Individual Differences* (PID).
 Indeed, until the 1990s, MQ published a very different kind of race science discourse from the race-and-intelligence research published in Int and PID.
@@
 In some cases, some race scientists simultaneously maintained active connections to both the mainstream scientific community and white supremacist organizations [@WinstonScientificRacismNorth2020; @JacksonJr.ArthurJensenEvolutionary2022; @SlobodianUnequalMindHow2023].
+Thus the emphasis on MQ in race science historiography may have been misplaced.
 
-## *[implications for agnotology]*
+## Agnogenesis within the scientific community
 
-In terms of the margins-mainstream-bridge model of the "echo chamber effect," there was no particular bridge that carried ideas originating in MQ on the margins to mainstream journals.
-Within the scope of our data, the race-and-intelligence version of race science discourse was always a mainstream phenomenon.
+In terms of the margins-mainstream-bridge model of the "echo chamber effect," there was no bridge that carried ideas originating in MQ on the margins to mainstream journals, because the kind of race science discourse that originated in MQ did not spread to the mainstream.
+At the same time, within the scope of our data, the race-and-intelligence version of race science discourse was always a mainstream phenomenon.
 
-The circulation of ideas is a prominent theme in the empirical agnotology literature.
+The circulation of ideas is a prominent theme in the agnotology literature.
 Researchers have often examined the public circulation of ideas — how an industry attempts to manage public concerns about its products in marketing, press releases, and interviews — as well as the circulation of ideas within a company or industry itself — as in memos among intramural scientists and management [@ProctorGoldenHolocaustOrigins2012].
-In a series of articles, Supran and Oreskes have shown that Exxon/ExxonMobil lied — knowingly made false statements — about climate change, contrasting skeptical and denialist statements in public-facing documents with skillful climate projections by the company's own scientists [@SupranAssessingExxonMobilsClimate2017; @SupranAddendumAssessingExxonMobils2020; @SupranRhetoricFrameAnalysis2021; @SupranAssessingExxonMobilsGlobal2023].
+In a series of articles, Supran and Oreskes have shown that Exxon/ExxonMobil lied about climate change — knowingly made false statements — by contrasting skeptical and denialist statements in public-facing documents with skillful climate projections by the company's own scientists [@SupranAssessingExxonMobilsClimate2017; @SupranAddendumAssessingExxonMobils2020; @SupranRhetoricFrameAnalysis2021; @SupranAssessingExxonMobilsGlobal2023].
 
-Less empirical research has examined the circulation of ideas within the scientific community.
-@SupranAssessingExxonMobilsGlobal2023 include several peer-reviewed journal articles in their content analysis, but compare them to intramural documents and public statements, rather than the broader scientific literature.
+However, empirical research in agntology has paid less attention to the circulation of ideas within the scientific community.
+@SupranAssessingExxonMobilsGlobal2023 include several peer-reviewed journal articles in their content analysis; but compare them to intramural documents and public statements, rather than the broader scientific literature.
 @FernandezPintoKnowBetterNot2017 gives three examples to illustrate the "tobacco strategy": the joint public relations-research funding campaign by the tobacco industry itself; medical ghostwriting by the pharmaceutical industry; and efforts by conservative think tanks to promote a simplistic account of the causes of the Great Recession that shifted blame away from deregulation of the finance industry.
 Both the first and last cases emphasize the circulation of ideas to the public, and the third does not involve academic researchers in any significant way.
-Medical ghostwriting is concerning as a strategy for circulating false or misleading claims in the scientific literature.
+Medical ghostwriting is indeed concerning as a strategy for circulating false or misleading claims in the scientific literature.
 But it has not been a major case in the agnotology literature in recent years.
-This might reflect less concern about ghosting in biomedical research over time: on 2025-01-16, a Web of Science search for "medic* ghostwrit*" returned 85 total papers published between 2003-2024, with the number of publications per year peaking at 11 in both 2010 and 2011 and declining to 1-2 per year in 2023 and 2024.
+This might reflect less concern about ghosting in biomedical research over time: on 2025-01-16, a Web of Science search for `medic* ghostwrit*` returned 85 total papers published during 2003-2024, with the number of publications per year peaking at 11 in both 2010 and 2011 and declining to 1-2 per year in 2023 and 2024.
 
-Some researchers have used agent-based models (computer simulation methods) to study the circulation of misinformation within scientific communities under agnogenic ("ignorance-creating") influences such as "industry selection" and increased dissemination of favorable views [@HolmanExperimentationIndustrialSelection2017; @WeatherallHowBeatScience2018; @OConnorMisinformationAgeHow2019].
+Some researchers have used agent-based models (computer simulation methods) to study the circulation of misinformation within scientific communities under agnogenic ("ignorance-creating") influences such as "industrial selection" and increased dissemination of favorable views [@HolmanExperimentationIndustrialSelection2017; @WeatherallHowBeatScience2018; @OConnorMisinformationAgeHow2019].
 While these models can give insight into how certain mechanisms might work, it is difficult to extrapolate from them to actual cases [@FernandezPintoEpistemicLandscapesReloaded2018; @HicksOpenScienceReplication2023].
@@
 Using the set of race science discourse articles identified in this project as a starting point, future research might examine authorship and/or citation patterns.
-For example, have race science discourse articles primarily produced by a small and stable set of authors, or a larger set that has changed over time?
+For example, have race science discourse articles been produced primarily by a small and stable set of authors, or a larger set that has changed over time?
 Or, to what extent have race science discourse articles been cited within non-race science articles?
 
-## *[implications for psychology]*
+## When the race science is coming from inside the house
 
@@
 Over the past few decades, many professional organizations in genetics and anthropology have made formal statements rejecting race as a biologically meaningful concept [@anthropology1996aapa; @american2018ashg; @board1998aaa; @committee2020asa; @FuentesAAPAStatementRace2019; @nhgri2023; @rotimi2022ashg; @wynshawboris2020ashg], undermining the core assumption of race science.
-The American Psychological Association (APA) --- whose members are predominantly clinical psychologists --- the Federation of Associations in Behavioral & Brain Sciences (FABBS) --- an organization whose purpose is to provide policy recommendations --- and the editors of *Nature Human Behavior* have openly rejected race as biologically meaningful [@american2021apology; @baron2022fabbs; @EditorsScienceMustRespect2022].
+The American Psychological Association (APA) --- whose members are predominantly clinical psychologists --- the Federation of Associations in Behavioral & Brain Sciences (FABBS) --- an organization whose purpose is to provide policy recommendations --- and the editors of *Nature Human Behavior* have openly rejected race as biologically meaningful [@american2021apology; @baron2022fabbs; @ScienceMustRespect2022].
 But research-focused psychological organizations like the Association for Psychological Science (APS) and Psychonomic Society (PS) stop short of rejecting race as biologically meaningful when they denounce racism [@aps2021statement; @board2020psychonomic].
@@
 
-Daniel J. Hicks, Emilio J.C. Lobato
+D.J. Hicks, Emilio J.C. Lobato
 

### 2025-02-03 paper/paper.qmd
ready to resubmit?

@@
 In contrast, Garrett had been president of the American Psychological Association in 1946 and chair of Psychology at Columbia from 1941 to 1955 [@WinstonScienceServiceFar1998].
+[@ShukmanRaceScienceInc2024 report that the Pioneer Fund has been effectively renamed the Human Diversity Foundation (HDF).]
 
@@
 
+(Late in the publication cycle, it occurred to us that evolutionary psychology is often associated with race science in the same way as behavior genetics.
+Because no evolutionary psychology journals were included in the corpus, we are unable to examine this association here.
+However, it would be a worthwhile topic for future research.)
+
 # Discussion

### 2025-02-04 paper/paper.qmd
submitted on 2025-02-04

@@
 Our analysis found two distinguishable lines of race science discourse.
-One is strongly associated with *Mankind Quarterly*, while the other — emphasizing race and intelligence — is more prevalent with mainstream psychology journals, especially *Intelligence* and *Personality and Individual Differences*.
+One is strongly associated with *Mankind Quarterly*, while the other — emphasizing race and intelligence — is more prevalent in mainstream psychology journals, especially *Intelligence* and *Personality and Individual Differences*.
 We compare psychology's response to race science to those of other fields, and discuss the social responsibility of scientists with respect to race science.
@@
 
-The Pioneer Fund itself clearly exemplifies the strategy of supporting friendly research; and figures such as Shockley and Jensen exemplify the strategy of recruiting high-profile, distinguished scientists [@TuckerFundingScientificRacism2002].
+Considering the five strategies identified by @FernandezPintoKnowBetterNot2017, the Pioneer Fund itself clearly exemplifies the strategy of supporting friendly research; and figures such as Shockley and Jensen exemplify the strategy of recruiting high-profile, distinguished scientists [@TuckerFundingScientificRacism2002].
 
@@
 If MQ provided the "safe space" for developing race science, protected away from mainstream criticism, then bridge venues would be necessary to disseminate these ideas into the scientific mainstream.
-We therefore hypothesized (H1) that scientometric and text mining analyses would show race science discourse appearing first and most prominently in MQ, and appearing in mainstream publications only subsequently, with certain venues serving as the bridge.
+We therefore hypothesized that (H1) scientometric and text mining analyses would show race science discourse appearing first and most prominently in MQ, and appearing in mainstream publications only subsequently, with certain venues serving as the bridge.
 
 Specifically, among both (other) academics and the general public, the field of behavior genetics is strongly associated with race science [@PanofskyMisbehavingScienceControversy2014].
-We therefore hypothesized (H2) that the journal *Behavior Genetics* (BG) would be a significant bridge venue for the dissemination of race science discourse.
+We therefore hypothesized that (H2) the journal *Behavior Genetics* (BG) would be a significant bridge venue for the dissemination of race science discourse.
 
@@
 Text was extracted using the R package `pdftools` [@OomsPdftoolsTextExtraction2023 version 3.0.1].
-This stage of corpus assembly was conducted in Fall 2021 with the assistance of Anthony Sainez.
+This stage of corpus assembly was conducted in fall 2021 with the assistance of Anthony Sainez.
 
@@
 We then used the author search tool in Clarivate's Web of Science platform (<https://www.webofscience.com/wos/author/search>), retrieving publication lists for 14 researchers.
-These searches were conducted between 2021-09-24 and 2021-10-05 by DH.
+These searches were conducted between 2021-09-24 and 2021-10-05 by DJH.
 After parsing these results, we counted how many of the 14 researchers had published in each journal.
@@
 Color-blind racism represents a change in the primary mechanisms by which white privilege is maintained, using non-racialized language in furtherance of the racial status quo both institutionally and individually, and increasingly reliant on the invisibility of socio-cultural mechanisms to preserve inequalities.
-Color-blind racism involves the use of racialized "codes" or "dogwhistles," language that is superficially non-racial but carries racial implications, such as (in the US context) "violent inner city criminals" or "welfare queen" [@SaulDogwhistlesPoliticalManipulation2018].
+Color-blind racism involves the use of racialized "codes" or "dogwhistles," language that is superficially non-racial but carries racial implicatures, such as (in the US context) "violent inner-city criminals" or "welfare queen" [@SaulDogwhistlesPoliticalManipulation2018].
 Alderfer notes that the racist implications of *The Bell Curve* only emerge gradually, with the opening of the book using the color-blind language of "group differences" [@AlderferScienceNonsciencePsychologists2003].
@@
 @FernandezPintoKnowBetterNot2017 gives three examples to illustrate the "tobacco strategy": the joint public relations-research funding campaign by the tobacco industry itself; medical ghostwriting by the pharmaceutical industry; and efforts by conservative think tanks to promote a simplistic account of the causes of the Great Recession that shifted blame away from deregulation of the finance industry.
-Both the first and last cases emphasize the circulation of ideas to the public, and the third does not involve academic researchers in any significant way.
+Both the first and third cases emphasize the circulation of ideas to the public, and the third does not involve academic researchers in any significant way.
 Medical ghostwriting is indeed concerning as a strategy for circulating false or misleading claims in the scientific literature.
@@
 
-## When the race science is coming from inside the house
+## Moral responsibilities of cognitive psychology
 
@@
 show how the term "human biodiversity," originally developed for anti-racist purposes by biological anthropologist Jonathan Marks, has been inverted by white supremacists and is now a dogwhistle for biological racial hierarchy [@PanofskyHowWhiteNationalists2021].
+
+A natural question, then, is what responsibilities does cognitive psychology have to mitigate the field’s contributions to race science and scientific racism?
 
@@
 Implicature is context-dependent; the exchange above assumes that the teen wants to do the same thing as Sasha, that they're responsible for doing the dishes, usually have practice in the mornings, and so on.
-In a different context, without this assumptions in play, the implicature doesn't work.
+In a different context, without these assumptions in play, the implicature doesn't work.
 

### 2025-04-07 paper/paper.qmd
updating racist agnotology discussion
adding paragraph connecting racist agnotology to norms and practices in psychology

--- 2025-02-03 paper/paper.qmd
@@
 
+Furthermore, the maintenance of doubt or ignorance about race within psychological science can be achieved not just via the explicit race science research of a small minority of psychologists who seek to create that doubt or ignorance in the first place [cf.
+@WinstonWhyMainstreamResearch2020], but also through the norms and traditions of the discipline that are engaged in by the average psychological scientist.
+For example, Roberts and Mortensen analyzed over 1700 articles published between 1995 and 2018 in seven psychology journals, finding that research using White samples was more likely than research using samples of color to be presented in race-neutral language [@roberts2023challenging].
+They reported this finding in the context of a larger argument about how psychology as a discipline treats White as a default and nonracial description of humanity and offering suggestions for how psychology can move past this longstanding implicit yet discipline-wide framework.
+The treatment of Whiteness as racially neutral by psychology can, like the reluctance of psychological science organizations to reject racial categories as biologically meaningful, further the goals of those who wish to create racialized ignorance, and it is the responsibility of psychological scientists to rebuke efforts that create ignorance.
+
 How should psychologists exercise these responsibilities?

### 2025-04-09 paper/paper.qmd
Pioneer 'presence'

--- 2025-02-04 paper/paper.qmd
@@
 
-Race science has not been heavily studied using an agnotology framework (though see @MillsRacialContract2014 and @KouranyMightScientificIgnorance2020).
+Race science has not been heavily studied using an agnotology framework [though see @MillsRacialContract2014; @KouranyMightScientificIgnorance2020; @DasguptaBiologicalEssentialismWhite2024; @DasguptaRacistAgnotologyHow2025].
 The qualitative historiography suggests that actors such as the Pioneer Fund have deployed a version of the tobacco strategy to promote misleading public perceptions of scientific uncertainty about (or even scientific support for) racial hierarchy.
@@
 Considering the five strategies identified by @FernandezPintoKnowBetterNot2017, the Pioneer Fund itself clearly exemplifies the strategy of supporting friendly research; and figures such as Shockley and Jensen exemplify the strategy of recruiting high-profile, distinguished scientists [@TuckerFundingScientificRacism2002].
+
+In @FernandezPintoKnowBetterNot2017, the Tobacco Industry Research Committee/Council for Tobacco Research provides the paradigm for supporting friendly research.
+@FernandezPintoKnowBetterNot2017 points out that the TIRC/CTR funded "scientific experts in major universities and research institutions" rather than marginal, fringe, or industry-employed scientists.
+The purpose was "to use 'good' research as a distraction" from the harmful effects of smoking by "fund[ing] research on heredity, infection, nutrition, hormones, nervous tension, and environmental factors" as causes of lung cancer and heart disease (57).
+This process of "industrial selection" for favorable research [@HolmanExperimentationIndustrialSelection2017] created the misleading impression that these non-tobacco causes were just as important, perhaps even more important, than smoking [compare @WeatherallHowBeatScience2018; @OConnorMisinformationAgeHow2019].
+Similarly, per Laughlin, the Pioneer Fund was explicitly created to promote research on eugenics and race science as just as important, if not more important, than a supposed "equalitarian dogma" and "environmentalist" explanations of racial inequality *[cite]*.
+
+A key aim of the TIRC/CTR was to recruit friendly scientists, "distinguished scientists, whose authority and academic connections would help the industry’s goal of fostering uncertainty about scientific claims" [@FernandezPintoKnowBetterNot2017 57].
+One prominent example not discussed by Fernández is statistician Ronald Fisher [@StolleyWhenGeniusErrs1991; @GundleProveThisIndustrys2010], who proposed a series of genetic explanations for the correlation between smoking and lung cancer in the last few years of his life.
+*[Shockley and Jensen]*
 
@@
 
+## Disproportionate presence of Pioneer-funded authors in race science discourse
+
+![*[caption]*](img/15_presence.png){#fig-presence}
+
+Post hoc, we were interested in trying to understand the impact that Pioneer might have had as an institution for supporting friendly scientists.
+(We thank a reviewer for suggesting this further analysis.) Because we did not have citation data, we were unable to examine "impact" in its usual sense of citations.
+Instead we chose to examine "presence," which we operationalize as the difference between the share of papers (published in a given topic in a given year) with at least one Pioneer-funded author ("paper share") against the share of all authors (again, for papers published in a given topic in a given year; individuated by family name) who were on our list of Pioneer-funded authors ("author share").
+Positive values of "presence" would indicate that Pioneer-funded authors publish more (in a given topic) than might be expected.
+
+Because this was a post hoc analysis, we elected to examine "presence" only for the three focal topics for the medium vocabulary, $k=40$ model (@fig-focal).
+@fig-presence shows the results of this analysis.
+
+@fig-presence A shows that topics 7 (MQ-style race science) and 24 (race-and-intelligence) had stable numbers of authors over the scope of the study, while topic 22 (mainstream intelligence) saw an exponential-like increase.
+@fig-presence B indicates that Pioneer-funded authors were a relatively high share of authors in topic 24, often comprising more than 10% of all authors in this topic after 1980.
+In contrast, these authors were almost never more than 10% of all authors in the other two topics.
+
+@fig-presence C shows that a greater share of papers in topic 24 had Pioneer-funded authors, compared to the other two topics, throughout the scope of the study but especially after about 1980.
+
+@fig-presence D is a scatterplot of paper share (subfigure C) against author share (subfigure B).
+Points above the dark diagonal line (length of the thin colored lines) indicate that, in that year-topic combination, Pioneer-funded authors' paper share was greater than their author share, and in this sense had a disproportionate "presence." For topic 7 (MQ-style race science), Pioneer-funded authors frequently had a small, disproportionately *negative* presence.
+While for topic 22 (mainstream intelligence) Pioneer-funded authors had a small or modest disproportionately positive presence.
+In contrast, for topic 24 (race-and-intelligence), Pioneer-funded authors had a substantial, disproportionately positive presence, with in some cases differences between paper share and author share of 20 points or more.
+
+To contextualize this measure of "presence," for each topic we used a permutation-style approach, drawing random sets of 15 authors (who ever published in the topic) and calculating paper share and author share in the same way as for Pioneer-funded authors.
+@fig-presence E shows the resulting distribution of mean presence (difference between paper share and author share) across 1,000 such permutations, with vertical lines for the observed values for the Pioneer-funded authors.
+Compared to the permutations, the Pioneer-funded authors do not have a distinctive presence for topics 7 and 22, but are at the rightmost extreme for topic 24.
+
+All together, this post hoc analysis suggests that the Pioneer Fund had a disproportionate presence in race-and-intelligence research, but not in MQ-style race science or mainstream intelligence research.
+
 ## Race science discourse and behavior genetics
@@
 Because no evolutionary psychology journals were included in the corpus, we are unable to examine this association here.
-However, it would be a worthwhile topic for future research.)
+However, this would be a worthwhile topic for future research.)
 
@@
 
-Using scientometric and text-mining methods, this study finds that Pioneer-funded researchers published heavily in certain mainstream psychology journals, much more than in *Mankind Quarterly*; and that a distinct version of race science discourse, centered on race and intelligence, can be identified in these same mainstream journals across the period 1960-2010.
+Using scientometric and text-mining methods, this study finds that Pioneer-funded researchers published heavily in certain mainstream psychology journals, much more than in *Mankind Quarterly*; that a distinct version of race science discourse, centered on race and intelligence, can be identified in these same mainstream journals across the period 1960-2010; and that authors funded by the Pioneer Fund had a disproportionate presence in this latter race science discourse.
 These findings indicate that *Mankind Quarterly* (MQ) was a less important venue for late 20th century race science discourse than mainstream psychology journals — especially *Intelligence* (Int) and *Personality and Individual Differences* (PID).
@@
 Thus the emphasis on MQ in race science historiography may have been misplaced.
+At the same time, the Pioneer Fund may have played a role in promoting race-and-intelligence research in mainstream psychology.
 
@@
 In a series of articles, Supran and Oreskes have shown that Exxon/ExxonMobil lied about climate change — knowingly made false statements — by contrasting skeptical and denialist statements in public-facing documents with skillful climate projections by the company's own scientists [@SupranAssessingExxonMobilsClimate2017; @SupranAddendumAssessingExxonMobils2020; @SupranRhetoricFrameAnalysis2021; @SupranAssessingExxonMobilsGlobal2023].
+@DasguptaBiologicalEssentialismWhite2024 and @DasguptaRacistAgnotologyHow2025 examine how white nationalists deploy agnotology-like analytical frames to critique and appropriate mainstream racial disparities research — research often explicitly motivated by an anti-racist agenda.
 
@@
 But it has not been a major case in the agnotology literature in recent years.
-This might reflect less concern about ghosting in biomedical research over time: on 2025-01-16, a Web of Science search for `medic* ghostwrit*` returned 85 total papers published during 2003-2024, with the number of publications per year peaking at 11 in both 2010 and 2011 and declining to 1-2 per year in 2023 and 2024.
+This might reflect less concern about ghostwriting in biomedical research over time: on 2025-01-16, a Web of Science search for `medic* ghostwrit*` returned 85 total papers published during 2003-2024, with the number of publications per year peaking at 11 in both 2010 and 2011 and declining to 1-2 per year in 2023 and 2024.
 
 Some researchers have used agent-based models (computer simulation methods) to study the circulation of misinformation within scientific communities under agnogenic ("ignorance-creating") influences such as "industrial selection" and increased dissemination of favorable views [@HolmanExperimentationIndustrialSelection2017; @WeatherallHowBeatScience2018; @OConnorMisinformationAgeHow2019].
-While these models can give insight into how certain mechanisms might work, it is difficult to extrapolate from them to actual cases [@FernandezPintoEpistemicLandscapesReloaded2018; @HicksOpenScienceReplication2023].
+Combined with our findings that Pioneer-funded authors had a disproportionate presence in race-and-intelligence research, these models suggest that the Pioneer Fund might have "maintained the controversy" within the scientific community by supporting friendly scientists.
+However, while these models can give insight into how certain mechanisms might work, it is difficult to link them to empirical data [@FernandezPintoEpistemicLandscapesReloaded2018; @HicksOpenScienceReplication2023].
 
@@
 But our findings suggest the circulation of race science within the scientific community warrants further investigation.
-Using the set of race science discourse articles identified in this project as a starting point, future research might examine authorship and/or citation patterns.
-For example, have race science discourse articles been produced primarily by a small and stable set of authors, or a larger set that has changed over time?
-Or, to what extent have race science discourse articles been cited within non-race science articles?
+Using the set of race science discourse articles identified in this project as a starting point, future research might examine patterns of citations or — perhaps by applying large language models (LLMs) — the appearance of certain modes of argument or types of evidence.
+For example, to what extent have race science discourse articles been cited within non-race science articles?
+Do race science discourse articles appeal to distinctive kinds of evidence, or deploy distinctive modes of interpretation?
 

### 2025-04-09 scripts/15_pf_presence.R

@@ -1,0 +1,304 @@
+library(tidyverse)
+library(tmfast)
+theme_set(theme_bw())
+library(patchwork)
+library(tinytable)
+
+library(here)
+library(arrow)
+
+## Used for simulation draws
+library(furrr)
+plan(multisession, workers = availableCores() - 2)
+
+
+## Load data ----
+data_dir = here('data')
+tm_dir = here(data_dir, '04_tm')
+out_dir = here('out')
+
+meta_ar = open_dataset(here(data_dir, '01_metadata'))
+
+source(here('R', 'author_data.R'))
+
+model_files = list.files(tm_dir, '*_tmfast.Rds') %>%
+    here(tm_dir, .) |> 
+    set_names('lg', 'md', 'sm')
+
+## med vocab, k = 40, topics 7, 22, 24 ----
+doc_gamma = read_rds(model_files['md']) |> 
+    tidy(k = 40, matrix = 'gamma') |> 
+    filter(gamma == max(gamma) & topic %in% c('V07', 'V22', 'V24'), 
+           .by = document)
+
+## Link metadata to documents of interest
+docs_df = inner_join(meta_ar, doc_gamma, by = c('article_id' = 'document')) |> 
+    ## And then to authors
+    left_join(author_data(), by = 'article_id') |> 
+    collect() |> 
+    mutate(family = {str_split(author, ' ') |> 
+            map_chr(last)})
+
+
+## Diversity as number of authors, irrespective of pub count
+n_authors_df = docs_df |> 
+    count(topic, year, family, name = 'nn') |> 
+    count(topic, year) |> 
+    complete(topic, year, fill = list(n = 0L))
+
+n_authors_gg = ggplot(n_authors_df, 
+                      aes(year, n, color = topic, group = topic)) +
+    geom_line(linewidth = 1.5, color = 'black') +
+    geom_line(linewidth = 1) +
+    scale_color_brewer(palette = 'Set2', guide = 'none') +
+    labs(y = 'authors')
+n_authors_gg
+
+## Diversity as perplexity (exp(entropy))
+## Basically the same as author count
+# docs_df |> 
+#     count(topic, year, family) |> 
+#     group_by(topic, year) |> 
+#     mutate(share = n / sum(n)) |> 
+#     summarize(H = sum(-share * log2(share))) |> 
+#     mutate(perplexity = 2^H) |> 
+#     ungroup() |> 
+#     complete(topic, year, fill = list(H = 0)) |> 
[…]

### 2025-04-09 paper/paper.qmd
giant caption

@@
 
-![*[caption]*](img/15_presence.png){#fig-presence}
+![Disproportionate presence of Pioneer-funded authors in three focal topics.\
+(A) Number of authors (distinguished by family name) for three focal topics from the medium vocabulary, k= 40 model, by year.
+Mainstream intelligence (topic 22) shows exponential-like increase while MQ race science (topic 7) and race-and-intelligence research (topic 24) remain flat.
+\
+(B) Pioneer-funded authors as a share (%) of all authors publishing in a given topic, by year (thin lines) and 5-year running mean (thick lines).
+Usually fewer than 15% of authors in a topic are Pioneer-funded.
+\
+(C) Share of papers (%) with at least one Pioneer-funded author, by year (thin lines) and 5-year running mean (thick lines).
+After about 1985, more than 15% of papers in topic 24 have at least one Pioneer-funded author.
+\
+(D) Comparison of author share (panel B) to paper share (panel C).
+Each point represents one year.
+If Pioneer-funded authors had publishing profiles (y-axis) proportionate to their authorship profiles (x-axis), points would be close to the dark line $y = x$.
+Points above this line indicate that Pioneer-funded authors had a relatively greater publishing presence than would be expected given their authorship profile.
+Length of thin vertical lines indicates difference between paper share and author share.
+\
+(E) Empirical density plots of mean difference between paper share and author share for 1,000 randomly drawn sets of 15 authors.
+Bold vertical lines are the mean values for the Pioneer-funded authors (signed length of lines in panel D).
+Pioneer-funded authors are similar to randomly drawn sets for topics 7 and 22, but very different for topic 24.
+](img/15_presence.png){#fig-presence}
+
+*[more details in supplement]*
 
@@
 
-Because this was a post hoc analysis, we elected to examine "presence" only for the three focal topics for the medium vocabulary, $k=40$ model (@fig-focal).
+Because this was a post hoc analysis, we elected to examine "presence" only for the three focal topics for the medium vocabulary, $k=40$ model (fig.
+@fig-focal).
+Fig.
 @fig-presence shows the results of this analysis.
 
+Fig.
 @fig-presence A shows that topics 7 (MQ-style race science) and 24 (race-and-intelligence) had stable numbers of authors over the scope of the study, while topic 22 (mainstream intelligence) saw an exponential-like increase.
+Fig.
 @fig-presence B indicates that Pioneer-funded authors were a relatively high share of authors in topic 24, often comprising more than 10% of all authors in this topic after 1980.
@@
 
-@fig-presence C shows that a greater share of papers in topic 24 had Pioneer-funded authors, compared to the other two topics, throughout the scope of the study but especially after about 1980.
+Fig. @fig-presence C shows that a greater share of papers in topic 24 had Pioneer-funded authors, compared to the other two topics, throughout the scope of the study but especially after about 1980.
 
+Fig.
 @fig-presence D is a scatterplot of paper share (subfigure C) against author share (subfigure B).
@@
 To contextualize this measure of "presence," for each topic we used a permutation-style approach, drawing random sets of 15 authors (who ever published in the topic) and calculating paper share and author share in the same way as for Pioneer-funded authors.
+Fig.
 @fig-presence E shows the resulting distribution of mean presence (difference between paper share and author share) across 1,000 such permutations, with vertical lines for the observed values for the Pioneer-funded authors.

### 2025-04-12 paper/paper.qmd
ready?

@@
 
-In @FernandezPintoKnowBetterNot2017, the Tobacco Industry Research Committee/Council for Tobacco Research provides the paradigm for supporting friendly research.
+In @FernandezPintoKnowBetterNot2017, the paradigm for supporting friendly research is the Tobacco Industry Research Committee, later known as the Council for Tobacco Research.
 @FernandezPintoKnowBetterNot2017 points out that the TIRC/CTR funded "scientific experts in major universities and research institutions" rather than marginal, fringe, or industry-employed scientists.
@@
 This process of "industrial selection" for favorable research [@HolmanExperimentationIndustrialSelection2017] created the misleading impression that these non-tobacco causes were just as important, perhaps even more important, than smoking [compare @WeatherallHowBeatScience2018; @OConnorMisinformationAgeHow2019].
-Similarly, per Laughlin, the Pioneer Fund was explicitly created to promote research on eugenics and race science as just as important, if not more important, than a supposed "equalitarian dogma" and "environmentalist" explanations of racial inequality *[cite]*.
+Similarly, per Laughlin, the Pioneer Fund was explicitly created to promote research on eugenics and race science as just as important, if not more important, than non-biological ("environmentalist") explanations of racial social inequality.
 
-A key aim of the TIRC/CTR was to recruit friendly scientists, "distinguished scientists, whose authority and academic connections would help the industry’s goal of fostering uncertainty about scientific claims" [@FernandezPintoKnowBetterNot2017 57].
-One prominent example not discussed by Fernández is statistician Ronald Fisher [@StolleyWhenGeniusErrs1991; @GundleProveThisIndustrys2010], who proposed a series of genetic explanations for the correlation between smoking and lung cancer in the last few years of his life.
-*[Shockley and Jensen]*
-
-For the current project, we hypothesized that MQ and certain mainstream journals played a key role in another strategy identified by @FernandezPintoKnowBetterNot2017, the "echo chamber effect." We theorize this effect as involving three components: the "margins," the "mainstream," and the "bridge."  @FernandezPintoKnowBetterNot2017 argues that, using sponsored journals and conferences, "the tobacco industry circumvented peer review standards in publication, without compromising its façade of research supporter" (59).
+For the current project, we hypothesized that MQ and certain mainstream journals played a key role in another strategy identified by @FernandezPintoKnowBetterNot2017, the "echo chamber effect." We theorize this effect as involving three components: the "margins," the "mainstream," and the "bridge" connecting them.
+@FernandezPintoKnowBetterNot2017 argues that, using sponsored journals and conferences, "the tobacco industry circumvented peer review standards in publication, without compromising its façade of research supporter" (59).
 These kinds of marginal venues provide a "safe space" in which favorable conceptual frameworks and empirical research (or pseudo-research) could be developed, protected from the destructive critical scrutiny they would receive in the mainstream.
 Bridge venues are accepted as part of the mainstream, but cite and draw on work from the margins, thereby mainstreaming these marginal ideas, that is, making them appear to be part of the legitimate body of mainstream scientific research.
+Where supporting friendly research is a strategy more-or-less entirely within the mainstream of the scientific community, the "echo chamber" strategy is located on the margins or fringes.
 
@@
 
-## Disproportionate presence of Pioneer-funded authors in race science discourse
-
-![Disproportionate presence of Pioneer-funded authors in three focal topics.\
-(A) Number of authors (distinguished by family name) for three focal topics from the medium vocabulary, k= 40 model, by year.
-Mainstream intelligence (topic 22) shows exponential-like increase while MQ race science (topic 7) and race-and-intelligence research (topic 24) remain flat.
-\
-(B) Pioneer-funded authors as a share (%) of all authors publishing in a given topic, by year (thin lines) and 5-year running mean (thick lines).
-Usually fewer than 15% of authors in a topic are Pioneer-funded.
-\
-(C) Share of papers (%) with at least one Pioneer-funded author, by year (thin lines) and 5-year running mean (thick lines).
-After about 1985, more than 15% of papers in topic 24 have at least one Pioneer-funded author.
-\
-(D) Comparison of author share (panel B) to paper share (panel C).
-Each point represents one year.
-If Pioneer-funded authors had publishing profiles (y-axis) proportionate to their authorship profiles (x-axis), points would be close to the dark line $y = x$.
-Points above this line indicate that Pioneer-funded authors had a relatively greater publishing presence than would be expected given their authorship profile.
-Length of thin vertical lines indicates difference between paper share and author share.
-\
-(E) Empirical density plots of mean difference between paper share and author share for 1,000 randomly drawn sets of 15 authors.
-Bold vertical lines are the mean values for the Pioneer-funded authors (signed length of lines in panel D).
-Pioneer-funded authors are similar to randomly drawn sets for topics 7 and 22, but very different for topic 24.
-](img/15_presence.png){#fig-presence}
-
-*[more details in supplement]*
-
-Post hoc, we were interested in trying to understand the impact that Pioneer might have had as an institution for supporting friendly scientists.
-(We thank a reviewer for suggesting this further analysis.) Because we did not have citation data, we were unable to examine "impact" in its usual sense of citations.
-Instead we chose to examine "presence," which we operationalize as the difference between the share of papers (published in a given topic in a given year) with at least one Pioneer-funded author ("paper share") against the share of all authors (again, for papers published in a given topic in a given year; individuated by family name) who were on our list of Pioneer-funded authors ("author share").
-Positive values of "presence" would indicate that Pioneer-funded authors publish more (in a given topic) than might be expected.
-
-Because this was a post hoc analysis, we elected to examine "presence" only for the three focal topics for the medium vocabulary, $k=40$ model (fig.
-@fig-focal).
-Fig.
-@fig-presence shows the results of this analysis.
-
-Fig.
-@fig-presence A shows that topics 7 (MQ-style race science) and 24 (race-and-intelligence) had stable numbers of authors over the scope of the study, while topic 22 (mainstream intelligence) saw an exponential-like increase.
-Fig.
-@fig-presence B indicates that Pioneer-funded authors were a relatively high share of authors in topic 24, often comprising more than 10% of all authors in this topic after 1980.
-In contrast, these authors were almost never more than 10% of all authors in the other two topics.
-
-Fig. @fig-presence C shows that a greater share of papers in topic 24 had Pioneer-funded authors, compared to the other two topics, throughout the scope of the study but especially after about 1980.
-
-Fig.
-@fig-presence D is a scatterplot of paper share (subfigure C) against author share (subfigure B).
-Points above the dark diagonal line (length of the thin colored lines) indicate that, in that year-topic combination, Pioneer-funded authors' paper share was greater than their author share, and in this sense had a disproportionate "presence." For topic 7 (MQ-style race science), Pioneer-funded authors frequently had a small, disproportionately *negative* presence.
-While for topic 22 (mainstream intelligence) Pioneer-funded authors had a small or modest disproportionately positive presence.
-In contrast, for topic 24 (race-and-intelligence), Pioneer-funded authors had a substantial, disproportionately positive presence, with in some cases differences between paper share and author share of 20 points or more.
-
-To contextualize this measure of "presence," for each topic we used a permutation-style approach, drawing random sets of 15 authors (who ever published in the topic) and calculating paper share and author share in the same way as for Pioneer-funded authors.
-Fig.
-@fig-presence E shows the resulting distribution of mean presence (difference between paper share and author share) across 1,000 such permutations, with vertical lines for the observed values for the Pioneer-funded authors.
-Compared to the permutations, the Pioneer-funded authors do not have a distinctive presence for topics 7 and 22, but are at the rightmost extreme for topic 24.
-
-All together, this post hoc analysis suggests that the Pioneer Fund had a disproportionate presence in race-and-intelligence research, but not in MQ-style race science or mainstream intelligence research.
-
 ## Race science discourse and behavior genetics
@@
 
+## Disproportionate presence of Pioneer-funded authors in race science discourse
+
+![Disproportionate presence of Pioneer-funded authors in three focal topics.\
+(A) Number of authors for three focal topics from the medium vocabulary, k= 40 model, by year.
+Mainstream intelligence (topic 22) shows exponential-like increase while MQ race science (topic 7) and race-and-intelligence research (topic 24) remain flat.
+\
+(B) Pioneer-funded authors as a share (%) of all authors publishing in a given topic, by year (thin lines) and 5-year running mean (thick lines).
+Usually fewer than 15% of authors in a topic are Pioneer-funded.
+\
+(C) Share of papers (%) with at least one Pioneer-funded author, by year (thin lines) and 5-year running mean (thick lines).
+After about 1985, more than 15% of papers in topic 24 have at least one Pioneer-funded author.
+\
+(D) Comparison of author share (panel B) to paper share (panel C).
+Each point is one year-topic combination.
+Points above the dark line $y = x$ indicate Pioneer-funded authors had a greater paper share than would be expected based on author share.
+Length of thin vertical lines measures "presence." \
+(E) Empirical density plots of "presence," mean difference between paper share and author share for 1,000 randomly drawn sets of 15 authors.
+Bold vertical lines are the mean values for the Pioneer-funded authors (signed length of lines in panel D).
+Pioneer-funded authors are similar to randomly drawn sets for topics 7 and 22, but very different for topic 24.
+](img/15_presence.png){#fig-presence}
+
+Post hoc, we were interested in trying to understand the impact that Pioneer might have had as an institution for supporting friendly scientists.
+(We thank a reviewer for comments that suggested this further analysis.) Because we did not have citation data, we were unable to examine "impact" in its usual sense of citations.
+Instead we chose to examine "presence," which we operationalized as the difference between the share of papers with at least one Pioneer-funded author ("paper share") against the share of all authors who were on our list of Pioneer-funded authors ("author share").
+Positive values of "presence" would indicate that Pioneer-funded authors publish more (in a given topic) than might be expected.
+
+Because this was a post hoc analysis, we elected to examine "presence" only for the three focal topics for the medium vocabulary, $k=40$ model (figure @fig-focal).
+Figure @fig-presence shows the results of this analysis.
+Details of the methods used here are reported in @sec-presence.
+
+Subfigure @fig-presence A shows that topics 7 (MQ-style race science) and 24 (race-and-intelligence) had stable numbers of authors over the scope of the study, while topic 22 (mainstream intelligence) saw an exponential-like increase.
+Fig.
+@fig-presence B indicates that Pioneer-funded authors were a relatively high share of authors in topic 24, often comprising more than 10% of all authors in this topic after 1980.
+In contrast, these authors were almost never more than 10% of all authors in the other two topics.
+
+Subfigure @fig-presence C shows that a greater share of papers in topic 24 had Pioneer-funded authors, compared to the other two topics, throughout the scope of the study but especially after about 1980.
+
+Subfigure @fig-presence D is a scatterplot of paper share (subfigure C) against author share (subfigure B), by year and topic.
+The dark diagonal line $y = x$ corresponds to equality between paper share and author share.
+Distance above (below) this line indicates that Pioneer-funded authors had disproportionately greater (less) "presence" than would be expected based on their authorship profile.
+For topics 7 (MQ-style race science) and 22 (mainstream intelligence), Pioneer-funded authors consistently had a small negative to modestly positive presence.
+In contrast, for topic 24 (race-and-intelligence), Pioneer-funded authors consistently had a substantial, disproportionately positive presence, with in some cases differences between paper share and author share of 20 points or more.
+
+Subfigure @fig-presence E uses a resampling method to compare the observed value of presence for the Pioneer-funded authors (dark vertical lines) to a "random" set of other authors publishing in the topic (density curves and rugs).
+Compared to other authors, the Pioneer-funded authors do not have a distinctive presence for topics 7 and 22, but are at the rightmost extreme for topic 24.
+
+All together, this post hoc analysis suggests that the Pioneer Fund had a disproportionate presence in race-and-intelligence research, but not in MQ-style race science or mainstream intelligence research.
+
 # Discussion
 
-Using scientometric and text-mining methods, this study finds that Pioneer-funded researchers published heavily in certain mainstream psychology journals, much more than in *Mankind Quarterly*; that a distinct version of race science discourse, centered on race and intelligence, can be identified in these same mainstream journals across the period 1960-2010; and that authors funded by the Pioneer Fund had a disproportionate presence in this latter race science discourse.
-These findings indicate that *Mankind Quarterly* (MQ) was a less important venue for late 20th century race science discourse than mainstream psychology journals — especially *Intelligence* (Int) and *Personality and Individual Differences* (PID).
+Using scientometric and text-mining methods, this study finds that Pioneer-funded researchers published heavily in certain mainstream psychology journals, much more than in *Mankind Quarterly* (MQ); that a distinct version of race science discourse, centered on race and intelligence, can be identified in these same mainstream journals across the period 1960-2010; and that authors funded by the Pioneer Fund had a disproportionate presence in this latter race science discourse, and not in the variety of race science discourse published by MQ.
+These findings indicate that MQ was a less important venue for late 20th century race science discourse than mainstream psychology journals — especially *Intelligence* (Int) and *Personality and Individual Differences* (PID).
 Indeed, until the 1990s, MQ published a very different kind of race science discourse from the race-and-intelligence research published in Int and PID.
+These results contrast with much of the qualitative historiography of scientific racism, which has often emphasized MQ [@MehlerFoundationFascismNew1989; @WinstonScienceServiceFar1998; @TuckerFundingScientificRacism2002; @SchafferScientificRacismAgain2007; @WinstonScientificRacismNorth2020; @SainiDraperMillionsPhilanthropic2022; @AdamsMisappropriationBiologicalAnthropology2021 also emphasize MQ, but in relation to anthropology rather than psychology].
+
 During this same period of time, race scientists served in prominent and influential positions in behavior genetics and psychology, including as society presidents and members of journal editorial boards.
@@
 Thus the emphasis on MQ in race science historiography may have been misplaced.
-At the same time, the Pioneer Fund may have played a role in promoting race-and-intelligence research in mainstream psychology.
+At the same time, our findings indicate that the Pioneer Fund may have played an important role in promoting race-and-intelligence research in mainstream psychology.
 
@@
 In terms of the margins-mainstream-bridge model of the "echo chamber effect," there was no bridge that carried ideas originating in MQ on the margins to mainstream journals, because the kind of race science discourse that originated in MQ did not spread to the mainstream.
-At the same time, within the scope of our data, the race-and-intelligence version of race science discourse was always a mainstream phenomenon.
+At the same time, within the scope of our data, the race-and-intelligence version of race science discourse was always a mainstream phenomenon, and as a mainstream phenomenon was potentially supported by the Pioneer Fund.
+All together, the Pioneer Fund, *Intelligence*, and *Personality and Individual Differences* appear to fit with the strategy of promoting friendly research, rather than the echo chamber strategy [@FernandezPintoKnowBetterNot2017].
 
@@
 However, while these models can give insight into how certain mechanisms might work, it is difficult to link them to empirical data [@FernandezPintoEpistemicLandscapesReloaded2018; @HicksOpenScienceReplication2023].
+
+Furthermore, doubt or ignorance about race within the community of psychological science is maintained not only by the Pioneer Fund and its small group of "friendly scientists" [@WinstonWhyMainstreamResearch2020] but also by the ordinary norms and standards of the broader scientific community.
+@roberts2023challenging analyzed over 1700 articles published between 1995 and 2018 in seven psychology journals, finding that research using White samples was more likely to be presented in race-neutral language than research using samples of people of color.
+@roberts2023challenging further argued that psychology as a discipline treats White as a default and "nonracial" description of humanity.
 
@@
 
-Furthermore, the maintenance of doubt or ignorance about race within psychological science can be achieved not just via the explicit race science research of a small minority of psychologists who seek to create that doubt or ignorance in the first place [cf.
+<!-- Furthermore, the maintenance of doubt or ignorance about race within psychological science can be achieved not just via the explicit race science research of a small minority of psychologists who seek to create that doubt or ignorance in the first place [cf.
 @WinstonWhyMainstreamResearch2020], but also through the norms and traditions of the discipline that are engaged in by the average psychological scientist.
@@
 The treatment of Whiteness as racially neutral by psychology can, like the reluctance of psychological science organizations to reject racial categories as biologically meaningful, further the goals of those who wish to create racialized ignorance, and it is the responsibility of psychological scientists to rebuke efforts that create ignorance.
+-->
 
@@
 
+# Supplemental analysis of "presence" of Pioneer-funded authors {#sec-presence}
+
+For this supplemental analysis, we focused on the medium vocabulary, $k = 40$ and the same focal topics as in figure @fig-focal.
+We characterized a paper as "in" one of these topics based on its maximum value of $\gamma$, the topic-document distribution.
+
+## Author name handling and ambiguity error
+
+The Crossref API does not standardize or disambiguate author names.
+We manually disambiguated the names of the 15 Pioneer-funded authors, e.g., searching the author field for every occurrence of "Rushton" and then identifying all the variations for J.
+Philippe Rushton.
+A simple crosswalk table allowed us to standardize all of these variations to a "canonical" version of the name of each Pioneer-funded author, e.g., J.P.
+Rushton would be replaced with the "canonical" version J.
+Philippe Rushton before proceeding with further analysis.
+Fig.
+@fig-pf-count shows the number of Pioneer-funded authors who published in the three focal topics, by year.
+Across the entire scope of the data, 12 of the 15 authors ever published in any of these three topics, with Rushton, Lynn, Jensen, Humphreys, and Vernon each publishing 10 or more papers (aggregated across topics).
+
+![Count of Pioneer-funded authors publishing in three focal topics, by year](img/15_pf_count.png){#fig-pf-count}
+
+Names of other authors were not disambiguated, and so, e.g., D.J.
+Hicks and Daniel J.
+Hicks would be treated as two different authors for the purposes of this analysis.
+As a consequence, the total number of authors is probably too high in the estimate of `author_share`, and so the estimate of `author_share` is probably too low.
+Assuming no issues with the manual disambiguation of Pioneer-funded authors, `paper_share` does not have this problem.
+Therefore the estimated difference `paper_share - author_share` is probably larger than the true difference.
+In terms of subfigure #fig-presence E, the vertical bar is probably too far to the right.
+
+This error is unlikely to have significant implications for topics 7 (MQ-style race science) and 22 (mainstream intelligence research).
+For topic 24 (race-and-intelligence), we judge that a significant error would be 5 percentage points or more, i.e., rather than the estimated mean difference of about 11%, the true mean difference would be about 6% or less.
+Algebraically,
+
+$$ \frac{p}{t} - \frac{p}{r} = \frac{p}{r - \varepsilon} - \frac{p}{r} > \alpha $$
+
+where $p$ is the count of Pioneer-funded authors; $t$ is the true author count, i.e., after perfect disambiguation; $r$ is the "raw" author count estimated using no disambiguation (except for the Pioneer-funded authors); $\varepsilon$ is the "ambiguity error" $r - t$; and $\alpha$ is the threshold .05 for a significant error in the conclusion.
+Solving for $\varepsilon$ gives
+
+$$\varepsilon > r - \frac{p}{\alpha + p/r}$$
+
+To estimate what size of ambiguity error $\varepsilon$ would be required, we first note from @fig-pf-count that at most 5 Pioneer-funded authors are active in topic 24 in any given year.
+And subfigure @fig-presence A indicates about 20-25 authors are active in this topic each year from the mid 1970s on.
+Setting $p = 5$ and $r = 20$ gives us $\varepsilon > 3.33$; $p = 5$ and $r = 25$ gives $\varepsilon > 5$; and $p = 3, r = 20$ also gives $\varepsilon > 5$.
+That is, a disambiguation error of 4-5 authors could be enough to undermine our finding for topic 24.
+
+We therefore manually inspected the list of authors for topic 24 in 1989, 1990, and 1991, to gauge the potential magnitude of ambiguity error.
+In 1989, topic 24 had 21 distinct authors, including one paper without an author name, and no author name variations that would require disambiguation.
+In 1990, topic 24 had 24 authors, 1 missing, and no variations.
+And in 1991 there were 30 authors, none missing, no variations.
+
+For the purposes of this ad hoc analysis, we are satisfied that ambiguity error is probably sufficiently small.
+
+## Paper share and presence
+
+Paper share was calculated simply by asking whether any (at least one) of a given paper's authors appeared on our list of Pioneer-funded authors.
+This resulted in a dichotomous (yes/no) variable at the paper level, and paper share was calculated as the fraction of papers with a "yes" value.
+
+In subfigure @fig-presence D, years when Pioneer-funded authors had 0 author share, i.e., none of these authors published in the topic, were excluded from calculating the mean value of presence.
+
+## Resampling to contextualize presence
+
+To contextualize this measure of "presence" within topics, we first identified all distinct authors (Pioneer-funded or otherwise, again without disambiguation for the others) across all papers in the given topic.
+We then used a permutation-style resampling approach: drawing a set of 15 authors and calculating paper share, author share, their difference, and the mean of this difference, exactly as with the Pioneer-funded authors.
+(In particular, years where author share was 0, i.e., the authors had no publications in the topic, were excluded from calculating the mean difference.)
+
+This approach is equivalent to randomly assigning group labels in a permutation test, and can be interpreted as comparing the presence of Pioneer-funded authors to a "random" set of authors.
+We conducted 1,000 resamples for each of the three focal topics.
+
 ![Fitted topic models: Gamma (topic-document) distributions, large vocabulary.](img/05_lg_gamma.png){#fig-gamma-lg  fig-env="figure*"}

### 2025-04-12 scripts/15_pf_presence.R

@@ -7,7 +7,7 @@ library(tinytable)
 library(here)
 library(arrow)
 
-## Used for simulation draws
+## Used for resampling draws
 library(furrr)
 plan(multisession, workers = availableCores() - 2)
 
@@ -35,14 +35,12 @@ doc_gamma = read_rds(model_files['md']) |>
 docs_df = inner_join(meta_ar, doc_gamma, by = c('article_id' = 'document')) |> 
     ## And then to authors
     left_join(author_data(), by = 'article_id') |> 
-    collect() |> 
-    mutate(family = {str_split(author, ' ') |> 
-            map_chr(last)})
+    collect()
 
 
-## Diversity as number of authors, irrespective of pub count
+## Number of authors ----
 n_authors_df = docs_df |> 
-    count(topic, year, family, name = 'nn') |> 
+    count(topic, year, author, name = 'nn') |> 
     count(topic, year) |> 
     complete(topic, year, fill = list(n = 0L))
 
@@ -67,7 +65,7 @@ n_authors_gg
 #     ggplot(aes(year, perplexity, color = topic)) +
 #     geom_line()
 
-## What share of papers have a PF-funded author? 
+## What share of papers have a PF-funded author? ----
 share_papers_df = docs_df |> 
     mutate(pf = author %in% mq$author) |> 
     summarize(pf = any(pf), authors = list(author),
@@ -95,7 +93,51 @@ share_papers_gg = ggplot(share_papers_df, aes(year,
 share_papers_gg
 
 
-## % of authors PF-funded
+## % of authors PF-funded ----
+## Count of PF-funded authors
+docs_df |> 
+    count(topic, year, author) |> 
+    mutate(pf = author %in% mq$author) |> 
+    filter(pf) |> 
+    count(topic, year) |> 
+    complete(topic, year = 1960:2010, 
+             fill = list(n = 0L)) |> 
+    ggplot(aes(year, n, color = topic, group = topic)) +
+    geom_line() +
+    # geom_point() +
+    ylim(0, 5) +
+    scale_color_brewer(palette = 'Set2')
+ggsave(here(out_dir, '15_pf_count.png'), 
+       height = 3, width = 4)
+
+## Which? 
+## 12/15 show up at some point
+## Rushton, Lynn, Jensen, Humphreys, and Vernon all have 10+
+docs_df |> 
+    count(topic, year, author) |> 
[…]

### 2025-04-14 paper/paper.qmd
fix issue w/ PF author list and downstream changes in 15

@@
 In @FernandezPintoKnowBetterNot2017, the paradigm for supporting friendly research is the Tobacco Industry Research Committee, later known as the Council for Tobacco Research.
-@FernandezPintoKnowBetterNot2017 points out that the TIRC/CTR funded "scientific experts in major universities and research institutions" rather than marginal, fringe, or industry-employed scientists.
+@FernandezPintoKnowBetterNot2017 points out that the TIRC/CTR funded "scientific experts in major universities and research institutions" rather than marginal, fringe, or industry-employed scientists (57).
 The purpose was "to use 'good' research as a distraction" from the harmful effects of smoking by "fund[ing] research on heredity, infection, nutrition, hormones, nervous tension, and environmental factors" as causes of lung cancer and heart disease (57).
@@
 (B) Pioneer-funded authors as a share (%) of all authors publishing in a given topic, by year (thin lines) and 5-year running mean (thick lines).
-Usually fewer than 15% of authors in a topic are Pioneer-funded.
+Usually less than 15% of authors in a topic are Pioneer-funded.
 \
 (C) Share of papers (%) with at least one Pioneer-funded author, by year (thin lines) and 5-year running mean (thick lines).
+Not shown is 1963 for topic 22, where Robert A.
+Gordon was one of two coauthors of the single publication in that topic that year.
 After about 1985, more than 15% of papers in topic 24 have at least one Pioneer-funded author.
@@
 Each point is one year-topic combination.
+Not shown here is 1963 for topic 22.
 Points above the dark line $y = x$ indicate Pioneer-funded authors had a greater paper share than would be expected based on author share.
 Length of thin vertical lines measures "presence." \
-(E) Empirical density plots of "presence," mean difference between paper share and author share for 1,000 randomly drawn sets of 15 authors.
+(E) Empirical density plots of "presence," mean difference between paper share and author share for 1,000 randomly drawn sets of 13 authors.
 Bold vertical lines are the mean values for the Pioneer-funded authors (signed length of lines in panel D).
@@
 Distance above (below) this line indicates that Pioneer-funded authors had disproportionately greater (less) "presence" than would be expected based on their authorship profile.
-For topics 7 (MQ-style race science) and 22 (mainstream intelligence), Pioneer-funded authors consistently had a small negative to modestly positive presence.
+For topics 7 (MQ-style race science) and 22 (mainstream intelligence), Pioneer-funded authors consistently had a small negative to modestly positive presence; with the exception of 1963 (not shown on this plot), where Robert A.
+Gordon was one of two coauthors (50% author share) of the sole paper in topic 22 that year (100% author share).
 In contrast, for topic 24 (race-and-intelligence), Pioneer-funded authors consistently had a substantial, disproportionately positive presence, with in some cases differences between paper share and author share of 20 points or more.
 
-Subfigure @fig-presence E uses a resampling method to compare the observed value of presence for the Pioneer-funded authors (dark vertical lines) to a "random" set of other authors publishing in the topic (density curves and rugs).
-Compared to other authors, the Pioneer-funded authors do not have a distinctive presence for topics 7 and 22, but are at the rightmost extreme for topic 24.
+Subfigure @fig-presence E uses a resampling method to compare the observed value of presence for the Pioneer-funded authors (dark vertical lines) to a "random" set of 13 other authors publishing in the topic (density curves and rugs).
+Compared to other authors, the Pioneer-funded authors do not have a distinctive presence for topic 7, are somewhat greater than the mode for topic 22, and are at the rightmost extreme for topic 24.
 
@@
 The Crossref API does not standardize or disambiguate author names.
-We manually disambiguated the names of the 15 Pioneer-funded authors, e.g., searching the author field for every occurrence of "Rushton" and then identifying all the variations for J.
+We manually disambiguated the names of the Pioneer-funded authors, e.g., searching the author field for every occurrence of "Rushton" and then identifying all the variations for J.
 Philippe Rushton.
@@
 @fig-pf-count shows the number of Pioneer-funded authors who published in the three focal topics, by year.
-Across the entire scope of the data, 12 of the 15 authors ever published in any of these three topics, with Rushton, Lynn, Jensen, Humphreys, and Vernon each publishing 10 or more papers (aggregated across topics).
+Across the entire scope of the data, 13 of the Pioneer-funded authors ever published in any of these three topics, with Rushton, Lynn, Jensen, Humphreys, and Vernon each publishing 10 or more papers (aggregated across topics).
+(Note that Lloyd Humphreys and Robert A.
+Gordon did not have any results in the Web of Science searches but did have papers in the corpus.
+Hardin, Horn, and Shuey do not appear in the papers for these focal topics.)
 
@@
 Therefore the estimated difference `paper_share - author_share` is probably larger than the true difference.
-In terms of subfigure #fig-presence E, the vertical bar is probably too far to the right.
+In terms of subfigure @fig-presence E, the vertical bar is probably too far to the right.
 
@@
 To contextualize this measure of "presence" within topics, we first identified all distinct authors (Pioneer-funded or otherwise, again without disambiguation for the others) across all papers in the given topic.
-We then used a permutation-style resampling approach: drawing a set of 15 authors and calculating paper share, author share, their difference, and the mean of this difference, exactly as with the Pioneer-funded authors.
+We then used a permutation-style resampling approach: drawing a set of 13 authors and calculating paper share, author share, their difference, and the mean of this difference, exactly as with the Pioneer-funded authors.
 (In particular, years where author share was 0, i.e., the authors had no publications in the topic, were excluded from calculating the mean difference.)

### 2025-04-14 R/author_data.R

@@ -42,18 +42,19 @@ author_data = function(data_dir = here::here('data'),
 }
 
 
-mq = tibble(author = c('Arthur R. Jensen', 
-                       'J. Philippe Rushton', 
-                       'Michael Levin', 
-                       'Robert M. Gordon', 
+pf = tibble(author = c('Thomas J. Bouchard, Jr.', 
+                       'Brunetto Chiarelli',
+                       'Hans J. Eysenck', 
+                       'Robert A. Gordon',
                        'Linda S. Gottfredson', 
-                       'Daniel R. Vining, Jr.', 
+                       'Joseph M. Horn', 
+                       'Lloyd G. Humphreys', 
+                       'Arthur R. Jensen', 
+                       'Michael Levin', 
                        'Richard Lynn', 
                        'R. Travis Osborne', 
-                       'Hans J. Eysenck', 
-                       'Thomas J. Bouchard, Jr.', 
-                       'Joseph M. Horn', 
+                       'J. Philippe Rushton', 
+                       'Audrey M. Shuey',
                        'Philip A. Vernon', 
-                       'Brunetto Chiarelli',
-                       'Lloyd G. Humphreys', 
-                       'Audrey M. Shuey'))
+                       'Daniel R. Vining, Jr.'
+                       ))

### 2025-04-14 scripts/15_pf_presence.R

@@ -1,6 +1,7 @@
 library(tidyverse)
-library(tmfast)
 theme_set(theme_bw())
+library(tmfast)
+
 library(patchwork)
 library(tinytable)
 
@@ -67,7 +68,7 @@ n_authors_gg
 
 ## What share of papers have a PF-funded author? ----
 share_papers_df = docs_df |> 
-    mutate(pf = author %in% mq$author) |> 
+    mutate(pf = author %in% pf$author) |> 
     summarize(pf = any(pf), authors = list(author),
               .by = c(article_id, year, topic)) |> 
     count(topic, year, pf) |> 
@@ -89,7 +90,8 @@ share_papers_gg = ggplot(share_papers_df, aes(year,
     geom_line(aes(y = share_sm), linewidth = 1) +
     scale_color_brewer(palette = 'Set2', guide = 'none') +
     scale_y_continuous(labels = scales::percent_format(), 
-                       name = 'paper share')
+                       name = 'paper share') +
+    coord_cartesian(ylim = c(0, .5))
 share_papers_gg
 
 
@@ -97,7 +99,7 @@ share_papers_gg
 ## Count of PF-funded authors
 docs_df |> 
     count(topic, year, author) |> 
-    mutate(pf = author %in% mq$author) |> 
+    mutate(pf = author %in% pf$author) |> 
     filter(pf) |> 
     count(topic, year) |> 
     complete(topic, year = 1960:2010, 
@@ -111,25 +113,26 @@ ggsave(here(out_dir, '15_pf_count.png'),
        height = 3, width = 4)
 
 ## Which? 
-## 12/15 show up at some point
+## 13/16 show up at some point
 ## Rushton, Lynn, Jensen, Humphreys, and Vernon all have 10+
+## Hardin, Horn, and Shuey do not appear
 docs_df |> 
     count(topic, year, author) |> 
-    mutate(pf = author %in% mq$author) |> 
+    mutate(pf = author %in% pf$author) |> 
     filter(pf) |> 
     group_by(author) |> 
     summarize(n = sum(n)) |> 
     arrange(desc(n))
 
 ## Authors in topic 24, 1989
-docs_df |> 
-    filter(year == 1991, topic == 'V24') |> 
-    count(author) |> 
-    view()
+# docs_df |> 
+#     filter(year == 1991, topic == 'V24') |> 
+#     count(author) |> 
+#     view()
 
 # docs_df |> 
[…]

### 2025-04-15 paper/paper.qmd
proofread

@@
 
-Considering the five strategies identified by @FernandezPintoKnowBetterNot2017, the Pioneer Fund itself clearly exemplifies the strategy of supporting friendly research; and figures such as Shockley and Jensen exemplify the strategy of recruiting high-profile, distinguished scientists [@TuckerFundingScientificRacism2002].
+Considering the five strategies identified by @FernandezPintoKnowBetterNot2017, the Pioneer Fund itself likely exemplifies the strategy of supporting friendly research; and figures such as Shockley and Jensen exemplify the strategy of recruiting high-profile, distinguished scientists [@TuckerFundingScientificRacism2002].
 
@@
 This process of "industrial selection" for favorable research [@HolmanExperimentationIndustrialSelection2017] created the misleading impression that these non-tobacco causes were just as important, perhaps even more important, than smoking [compare @WeatherallHowBeatScience2018; @OConnorMisinformationAgeHow2019].
-Similarly, per Laughlin, the Pioneer Fund was explicitly created to promote research on eugenics and race science as just as important, if not more important, than non-biological ("environmentalist") explanations of racial social inequality.
+Similarly, per the quote from Laughlin above, the Pioneer Fund was all-but-explicitly created to promote research on eugenics and race science as just as important, if not more important, than non-biological ("environmentalist") explanations of racial social inequality.
 
@@
 @FernandezPintoKnowBetterNot2017 argues that, using sponsored journals and conferences, "the tobacco industry circumvented peer review standards in publication, without compromising its façade of research supporter" (59).
-These kinds of marginal venues provide a "safe space" in which favorable conceptual frameworks and empirical research (or pseudo-research) could be developed, protected from the destructive critical scrutiny they would receive in the mainstream.
+These kinds of marginal venues provide a "safe space" in which favorable conceptual frameworks and empirical research (or pseudo-research) could be developed, protected from the destructive critical scrutiny they would receive in the mainstream [@HicksYoungsPValuePlot2022].
 Bridge venues are accepted as part of the mainstream, but cite and draw on work from the margins, thereby mainstreaming these marginal ideas, that is, making them appear to be part of the legitimate body of mainstream scientific research.
@@
 
-The primary data for this study was the full text of all available articles published in journals that had published multiple Pioneer-funded researchers, between 1960 (the year MQ was founded) and 2010.
+The primary data for this study were the full text of all available articles published in journals that had published multiple Pioneer-funded researchers, between 1960 (the year MQ was founded) and 2010.
 The end year was chosen for a combination of reasons: easy access to MQ ended at 2004, and we did not want to extend the data too far beyond that point; 2010 gave us a "tidy" 50-year window of coverage; and an end year that wasn't too recent would hopefully allow us to avoid distracting controversies about researchers who are still active today.
@@
 Pioneer-funded authors are similar to randomly drawn sets for topics 7 and 22, but very different for topic 24.
-](img/15_presence.png){#fig-presence}
+](img/15_presence.png){c}
 
@@
 Details of the methods used here are reported in @sec-presence.
+Note that papers were assigned to topics in a different, less conservative way than in @fig-focal.
 
 Subfigure @fig-presence A shows that topics 7 (MQ-style race science) and 24 (race-and-intelligence) had stable numbers of authors over the scope of the study, while topic 22 (mainstream intelligence) saw an exponential-like increase.
-Fig.
-@fig-presence B indicates that Pioneer-funded authors were a relatively high share of authors in topic 24, often comprising more than 10% of all authors in this topic after 1980.
+Subfigure @fig-presence B indicates that Pioneer-funded authors were a relatively high share of authors in topic 24, often comprising more than 10% of all authors in this topic after 1980.
 In contrast, these authors were almost never more than 10% of all authors in the other two topics.
@@
 
-However, empirical research in agntology has paid less attention to the circulation of ideas within the scientific community.
+However, empirical research in agnotology has paid less attention to the circulation of ideas within the scientific community.
 @SupranAssessingExxonMobilsGlobal2023 include several peer-reviewed journal articles in their content analysis; but compare them to intramural documents and public statements, rather than the broader scientific literature.
@@
 Each of these three elements requires various assumptions to be logically relevant to the other two  [@LonginoScienceSocialKnowledge1990; @LloydAdaptationismLogicResearch2015].
-For example, a particular project might use data from a national genome bank and a genome-wide association study (GWAS; method) to develop a polygenic score estimator of educational attainment (finding) with the aim of answering the question "To what extent is cognitive ability genetic?" This project must assume that a sample from a single nation is representative of humanity in general (to be relevant to the generalized framing of the question), that "cognitive ability" is a well-defined construct and worth investigating, that educational attainment is an accurate measure of cognitive ability, that the SNPs included in the genome data capture genetically-caused variation, that analysis of the polygenic scores can recover/measure this variation, that variation can be partitioned into genetic and environmental causes [@KellerMirageSpaceNature2010], and so on.
+For example, a particular project might use data from a national genome bank and a genome-wide association study (GWAS; method) to develop a polygenic score estimator of educational attainment (finding) with the aim of answering the question "To what extent is cognitive ability genetic?" This project must assume that a sample from a single nation is representative of humanity in general (to be relevant to the generalized framing of the question), that "cognitive ability" is a well-defined construct and worth investigating, that educational attainment is an accurate measure of cognitive ability, that the SNPs included in the genome data capture genetically-caused variation, that analysis of the polygenic scores can recover/measure this variation, that variation can be partitioned into genetic and environmental causes [@KellerMirageSpaceNature2010; @TaberyStruggleUnderstandInteraction2014], and so on.
 
 Some of these assumptions — or the findings themselves — can support scientific racist claims, either through strict deductive entailment (the kind studied in formal logic courses) or through *implicature* [@DavisImplicature2024].
-The concept of "implicature" refers to the phenomenon that language use goes beyond the strict deductive entailments — "we mean more than we say." Consider the following exchange between a teenager and their parent.
+The concept of "implicature" refers to the phenomenon that language use goes beyond strict deductive entailments — "we mean more than we say." Consider the following exchange between a teenager and their parent.
 

### 2025-04-15 scripts/15_pf_presence.R

@@ -27,16 +27,31 @@ model_files = list.files(tm_dir, '*_tmfast.Rds') %>%
     set_names('lg', 'md', 'sm')
 
 ## med vocab, k = 40, topics 7, 22, 24 ----
+## NB In script 08, docs are assigned to topics based on a threshold of normalized gamma > 0.5
+## For simplicity, here we use the max value of gamma
+## This results in much higher document counts than in figure 2, 
 doc_gamma = read_rds(model_files['md']) |> 
     tidy(k = 40, matrix = 'gamma') |> 
     filter(gamma == max(gamma) & topic %in% c('V07', 'V22', 'V24'), 
            .by = document)
+
+# # A tibble: 3 × 2
+# topic     n
+# <chr> <int>
+# 1 V07    1436
+# 2 V22     963
+# 3 V24     611
+count(doc_gamma, topic)
 
 ## Link metadata to documents of interest
 docs_df = inner_join(meta_ar, doc_gamma, by = c('article_id' = 'document')) |> 
     ## And then to authors
     left_join(author_data(), by = 'article_id') |> 
     collect()
+
+ggplot(docs_df, 
+       aes(year, color = topic, group = topic)) +
+    geom_line(stat = 'count', aes(y = after_stat(count)))
 
 
 ## Number of authors ----

### 2025-04-23 paper/paper.qmd
accidentally deleted fig label
working on final formatting

@@
 bibliography: [references.yaml, refs-manual.bib]
+csl: apa.csl
 
@@
     cite-method: citeproc
-    include-in-header: 
-        text: |
-            \usepackage{lineno}
-            \linenumbers
-    geometry:
-        - left = 1.5in
-        - right = 1.5in
 
@@
 
-<!-- 
-QSS submission guidelines: 
-<https://direct.mit.edu/qss/pages/submission-guidelines>
+keywords:
 
-- 300-word cover letter on significance and fit
-- Articles: original theoretical, empirical, or methodological research (typically between 5,000 and 8,000 words)
-
- -->
+- race science
+- race and intelligence
+- psychology
+- behavior genetics
+- text mining
+- topic modeling
 
@@
 
-![Journals publishing 6 or more Pioneer-funded researchers, Web of Science author search results.](img/pf_wos_results.png){#fig-wos fig-env="figure*"}
+![Journals publishing 6 or more Pioneer-funded researchers, Web of Science author search results.](img/fig-01.png){#fig-wos fig-env="figure*"}
 
@@
 Following the approach of @HicksProductivityInterdisciplinaryImpacts2021, we did not attempt to identify a unique best fitted model for further analysis.
-While the results section of this paper focuses on the medium vocabulary, $k=40$ model as the "median" among the 24 fitted models, qualitatively similar results appear across all models with $k \geq 10$ for all vocabularies; see @sec-topic-quality and figures @fig-grid-lg, @fig-grid-md, @fig-grid-sm.
+While the results section of this paper focuses on the medium vocabulary, $k=40$ model as the "median" among the 24 fitted models, qualitatively similar results appear across all models with $k \geq 10$ for all vocabularies; see @sec-topic-quality and supplements SM7, SM8, SM9<!-- figures @fig-grid-lg, @fig-grid-md, @fig-grid-sm -->.
 
-To analyze the fitted topic models, we first constructed and manually inspected visualizations of the topic-document distributions (see figures @fig-gamma-lg, @fig-gamma-md, @fig-gamma-sm) and the top (highest-probability) terms for each topic (see figures @fig-silge-lg, @fig-silge-md, @fig-silge-sm).
+To analyze the fitted topic models, we first constructed and manually inspected visualizations of the topic-document distributions (see figures @fig-gamma-lg, @fig-gamma-md, @fig-gamma-sm) and the top (highest-probability) terms for each topic (see supplements SM4-6<!-- figures @fig-silge-lg, @fig-silge-md, @fig-silge-sm -->).
 This suggested that the models often identified two distinct topics with racial terms, that some of these topics used terms from intelligence research, and that these topics had different occurrence patterns across journals.
-To examine these patterns more systematically, we programmatically identified every topic across all 24 models that used some combination of racial and intelligence terms, and assembled visualization grids of the occurrence of these topics by journal over time (see figures @fig-grid-lg, @fig-grid-md, @fig-grid-sm).
+To examine these patterns more systematically, we programmatically identified every topic across all 24 models that used some combination of racial and intelligence terms, and assembled visualization grids of the occurrence of these topics by journal over time (see supplements SM7, SM8, SM9<!-- figures @fig-grid-lg, @fig-grid-md, @fig-grid-sm -->).
 
@@
 Thin lines give annual values, thick lines give 5-year running averages.
-Because *Behavior Genetics* and *Behavioral and Brain Sciences* are not prominent in any panel, neither is given a direct label.](img/08_focal_topics.png){#fig-focal}
+Because *Behavior Genetics* and *Behavioral and Brain Sciences* are not prominent in any panel, neither is given a direct label.](img/fig-02.png){#fig-focal}
 
-Figure @fig-focal focuses on three topics identified in one of the 24 fitted topic models; see table @tbl-silge for a table version of the Silge plot in @fig-focal; figures [@fig-gamma-lg; @fig-gamma-md; @fig-gamma-sm; @fig-silge-lg; @fig-silge-md; @fig-silge-sm] for all topics from all models, and figures [@fig-grid-lg; @fig-grid-md; @fig-grid-sm] for selected topics from all models; we discuss the robustness of our findings across all models below.
+Figure @fig-focal focuses on three topics identified in one of the 24 fitted topic models; see table @tbl-silge for a table version of the Silge plot in @fig-focal; figures [@fig-gamma-lg; @fig-gamma-md; @fig-gamma-sm; and supplements SM4, SM5, SM6<!--  @fig-silge-lg; @fig-silge-md; @fig-silge-sm -->] for all topics from all models, and supplements SM7, SM8, SM9<!-- figures [@fig-grid-lg; @fig-grid-md; @fig-grid-sm] --> for selected topics from all models; we discuss the robustness of our findings across all models below.
 In the focal model, topic 07 is strongly associated with MQ:  MQ published dozens of articles in this topic each year, and no other journal ever published more than a handful.
@@
 
-These qualitative results are robust across models using the large and medium vocabularies; see @fig-grid-lg, @fig-grid-md, @fig-grid-sm.
+These qualitative results are robust across models using the large and medium vocabularies; see supplements SM7, SM8, SM9<!-- @fig-grid-lg, @fig-grid-md, @fig-grid-sm -->.
 Across all vocabularies, as $k$ increases, multiple different "intelligence only" topics emerge, apparently picking out different methods or aspects of intelligence research.
@@
 Pioneer-funded authors are similar to randomly drawn sets for topics 7 and 22, but very different for topic 24.
-](img/15_presence.png){c}
+](img/fig-03.png){#fig-presence}
 
@@
 
-![Corpus assembly.](img/dataset.png){#fig-corpus width="120%" fig-env="figure*"}
+![Corpus assembly.](img/fig-S01.png){#fig-corpus width="120%" fig-env="figure*"}
 
-![Count of documents in the corpus, by journal and year.](img/02_count.png){#fig-counts  fig-env="figure*"}
+![Count of documents in the corpus, by journal and year.](img/fig-S02.png){#fig-counts  fig-env="figure*"}
 
@@
 
-We next constructed Silge plots [@SilgeTopicModeling2017], showing the top 15 (highest-probability) terms (noun phrases) from each topic in the model; see figures [@fig-silge-lg; @fig-silge-md; @fig-silge-sm].
+We next constructed Silge plots [@SilgeTopicModeling2017], showing the top 15 (highest-probability) terms (noun phrases) from each topic in the model; see supplements SM4, SM5, SM6<!-- figures [@fig-silge-lg; @fig-silge-md; @fig-silge-sm] -->.
 Perusing these term lists, we focused on topics in each model that contained racial terms.
@@
 The gamma distributions and Silge plots suggested that, once $k$ was sufficiently large, the topic models were distinguishing between two types of race science discourse articles.
-We therefore examined the prevalence of topics that used racial terms, intelligence-research terms, or both, by journal and across all 24 models; see figures [@fig-grid-lg; @fig-grid-md; @fig-grid-sm].
+We therefore examined the prevalence of topics that used racial terms, intelligence-research terms, or both, by journal and across all 24 models; see supplements SM7, SM8, SM9<!-- figures [@fig-grid-lg; @fig-grid-md; @fig-grid-sm] -->.
 In these visualizations, each row of facets corresponds to a value of $k$ (excluding 5 as too noisy), and each column of facets corresponds to one topic.
@@
 The consensus true positive rate was 89.3% (108/121), while the consensus false positive rate was 8.3% (10/121), with 98% agreement across the two raters ($\kappa = 0.86)$.
-Coding spreadsheets are included in the supplemental materials S8-S9.
+Coding spreadsheets are included in the supplemental materials SM2 and SM3.
 For each of the 10 consensus false positive documents, their inclusion in the topic was readily explained, but in different ways for different documents.
@@
 
+At the request of a reviewer, we also extracted a list of the top documents for each of the three focal topics in the $k=40$, medium vocabulary model. This list is supplement SM10.
+
 # STM robustness check {#sec-stm}
@@
 As in the `tmfast` analysis, the "race only" topic is almost exclusive to MQ; the "race and intelligence" topic emerges in mainstream journals before appearing in both the mainstream and MQ; and there are distinctive "intelligence only" topics.
-Compare figures @fig-focal, @fig-grid-md.
+Compare figure @fig-focal with supplement SM8<!-- @fig-grid-md -->.
 One notable difference between these results is the role of PR in publishing race and intelligence.
 
-![Silge plots and thresholded document counts for "race only," "race-and-intelligence," and "intelligence only" topics in the STM model.](img/12-stm.png){#fig-stm fig-env="figure*"}
+![Silge plots and thresholded document counts for "race only," "race-and-intelligence," and "intelligence only" topics in the STM model.](img/fig-S07.png){#fig-stm fig-env="figure*"}
 
@@
 
-![Count of Pioneer-funded authors publishing in three focal topics, by year](img/15_pf_count.png){#fig-pf-count}
+![Count of Pioneer-funded authors publishing in three focal topics, by year](img/fig-S08.png){#fig-pf-count}
 
@@
 
-![Fitted topic models: Gamma (topic-document) distributions, large vocabulary.](img/05_lg_gamma.png){#fig-gamma-lg  fig-env="figure*"}
+![Fitted topic models: Gamma (topic-document) distributions, large vocabulary.](img/fig-S09.png){#fig-gamma-lg  fig-env="figure*"}
 
-![Fitted topic models: Gamma (topic-document) distributions, medium vocabulary.](img/05_md_gamma.png){#fig-gamma-md  fig-env="figure*"}
+![Fitted topic models: Gamma (topic-document) distributions, medium vocabulary.](img/fig-S10.png){#fig-gamma-md  fig-env="figure*"}
 
-![Fitted topic models: Gamma (topic-document) distributions, small vocabulary.](img/05_sm_gamma.png){#fig-gamma-sm  fig-env="figure*"}
+![Fitted topic models: Gamma (topic-document) distributions, small vocabulary.](img/fig-S11.png){#fig-gamma-sm  fig-env="figure*"}
 
-![Silge plots (top 15 terms) for each topic and model, large vocabulary.](img/06-lg-silge.pdf){#fig-silge-lg fig-env="figure*"}
+<!-- 
+![Silge plots (top 15 terms) for each topic and model, large vocabulary.](img/fig-S12.pdf){#fig-silge-lg fig-env="figure*"}
 
-![Silge plots (top 15 terms) for each topic and model, medium vocabulary.](img/06-md-silge.pdf){#fig-silge-md fig-env="figure*"}
+![Silge plots (top 15 terms) for each topic and model, medium vocabulary.](img/fig-S13.pdf){#fig-silge-md fig-env="figure*"}
 
-![Silge plots (top 15 terms) for each topic and model, small vocabulary.](img/06-sm-silge.pdf){#fig-silge-sm fig-env="figure*"}
+![Silge plots (top 15 terms) for each topic and model, small vocabulary.](img/fig-S14.pdf){#fig-silge-sm fig-env="figure*"}
 
-![Topics with "race," "intelligence," or both occurring in the top 15 terms, large vocabulary.](img/08_grid_lg.pdf){#fig-grid-lg fig-env="figure*"}
+![Topics with "race," "intelligence," or both occurring in the top 15 terms, large vocabulary.](img/fig-S15.pdf){#fig-grid-lg fig-env="figure*"}
 
-![Topics with "race," "intelligence," or both occurring in the top 15 terms, medium vocabulary.](img/08_grid_md.pdf){#fig-grid-md fig-env="figure*"}
+![Topics with "race," "intelligence," or both occurring in the top 15 terms, medium vocabulary.](img/fig-S16.pdf){#fig-grid-md fig-env="figure*"}
 
-![Topics with "race," "intelligence," or both occurring in the top 15 terms, small vocabulary.](img/08_grid_sm.pdf){#fig-grid-sm fig-env="figure*"}
+![Topics with "race," "intelligence," or both occurring in the top 15 terms, small vocabulary.](img/fig-S17.pdf){#fig-grid-sm fig-env="figure*"}
+ -->

### 2025-04-23 paper/paper.qmd
data and code DOI

@@
 
-The public analysis repository, <https://github.com/dhicks/race-science> *[DOI will go here]*, includes document metadata and documentwise counts of NLP-extracted noun phrases ("document-term matrices"), along with code to reproduce the analysis of the following sections.
+The public analysis repository (<https://github.com/dhicks/race-science> and <https://doi.org/10.5281/zenodo.15272241>) includes document metadata and documentwise counts of NLP-extracted noun phrases ("document-term matrices"), along with code to reproduce the analysis of the following sections.
 
@@
 **Data and materials availability:** Web of Science search results and article fulltext are not publicly available due to intellectual property restrictions.
-Preprocessed analysis data (document-term counts and article metadata) and analysis scripts are available at *[doi]*.
+Preprocessed analysis data (document-term counts and article metadata) and analysis scripts are available at <https://doi.org/10.5281/zenodo.15272241>.
 
